### 2020-06-10 explanatory text.Rmd
explanatory text

---
title: "Application Outline"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Application aims

The aim of this application is to provide a statistical foundation for Bateson challenges on the basis of race and gender. Building on the work of Kadane^[Kadane, 2018. _Statistics for Bateson Challenges_, **17**, 1-13.] and Barrett^[Barrett, 2007. _The American Statistician_, **61**, 4, 296–301.], this application makes use of _Bayesian statistical methods_ to identify potential baises in jury selection, while taking into account historical (from previous trials) jury selection behavior by both prosecutor and defense attorneys.

## Components of the analysis

This application relies on two components. The **first component** is historical jury selection data (stored in a compiled database) that is used to inform our 'prior expectations' about jury selection behavior by a given attorney. The **second component** is the jury selection pattern observed in the trial at hand (and entered into this application). The combination of these two elements allow us to make determinations of bias, based on the current trials data that are _conditioned_, or accounting for, historical jury selection practices. Combining information from the current trial with historical information is the essence of the Bayesian analysis: we have updated our prior beliefs based on the new data presented to us.

## Interpreting the results

Looking at the plots above, we can see this playing out. In these plots $d$ is our measure of bias: **positive** values of $d$ reflect bias _against_ the cognizable class, while **negative** values  of $d$ reflect bias _for_ the cognizable class. If $d$ is **zero**, then jury selection is _neutral_.  Given that we are interested in accounting for uncertainty (or, our degree of confidence) in our estimates, we represent $d$ as a distribution rather than a single value.

The greyed-out distribution in the background represents our 'prior expectation' of what the value of $d$ should be based on historical trial data. Here we have one prior expectation for prosecutors and for the defense. The blue (for prosecution) and red (for the defense) distributions in the foreground represent our _updated expectations_ based on the data entered into the application. This updated information is simply the product of the prior information and the information contained in the table on the left. The dashed lines indicate _credible intervals_^[These are 95% credible intervals, i.e., they contain 95% of the values of $d$.] for each of our distributions of $d$. These intervals indicate the regions of the distributions that we have a high degree of confidence contain the true values of $d$. If these credible intervals do not contain zero, this suggests evidence of bias for or against the cognizable class. Alternatively, if these credible intervals do contain zero, we have less confidence that the true value of $d$ is not, in fact zero, and so have no evidence of bias.

### 2020-06-11 explanatory text.Rmd
revised explanatory statement; added .bib file for references

@@
 title: "Application Outline"
-output: github_document
+output: 
+  github_document:
+       pandoc_args: --webtex
+bibliography: references.bib
 ---
@@
 
-## Application aims
+## Aim
 
-The aim of this application is to provide a statistical foundation for Bateson challenges on the basis of race and gender.
-Building on the work of Kadane^[Kadane, 2018.
-_Statistics for Bateson Challenges_, **17**, 1-13.] and Barrett^[Barrett, 2007.
-_The American Statistician_, **61**, 4, 296–301.], this application makes use of _Bayesian statistical methods_ to identify potential baises in jury selection, while taking into account historical (from previous trials) jury selection behavior by both prosecutor and defense attorneys.
+This application estimates how likely an attorney used peremptory challenges during jury selection to strike potential jurors based on their race or gender.
+The analysis builds on prior work by @Kadane2018 and @Barrett2007 on statistical analysis for _Batson_ challenges.
 
-## Components of the analysis
+## Analysis
 
-This application relies on two components.
-The **first component** is historical jury selection data (stored in a compiled database) that is used to inform our 'prior expectations' about jury selection behavior by a given attorney.
-The **second component** is the jury selection pattern observed in the trial at hand (and entered into this application).
-The combination of these two elements allow us to make determinations of bias, based on the current trials data that are _conditioned_, or accounting for, historical jury selection practices.
-Combining information from the current trial with historical information is the essence of the Bayesian analysis: we have updated our prior beliefs based on the new data presented to us.
+To estimate bias, this application relies on two kinds of data: (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, pre-compiled historical data on that attorney's use of strikes in past cases.
 
-## Interpreting the results
+In the above plots, $d$ is our measure of bias: **positive** values of $d$ reflect bias _against_ the cognizable class, while **negative** values  of $d$ reflect bias _for_ the cognizable class.
+If $d$ is exactly **zero**, we can be completely certain that there is zero bias.
 
-Looking at the plots above, we can see this playing out.
-In these plots $d$ is our measure of bias: **positive** values of $d$ reflect bias _against_ the cognizable class, while **negative** values  of $d$ reflect bias _for_ the cognizable class.
-If $d$ is **zero**, then jury selection is _neutral_.
-Given that we are interested in accounting for uncertainty (or, our degree of confidence) in our estimates, we represent $d$ as a distribution rather than a single value.
+Here, we represent $d$ as a distribution of how likely each possible value of $d$ is, given the data we have. This lets us depict how confident we should be in inferring bias from the data we have.
 
-The greyed-out distribution in the background represents our 'prior expectation' of what the value of $d$ should be based on historical trial data.
-Here we have one prior expectation for prosecutors and for the defense.
-The blue (for prosecution) and red (for the defense) distributions in the foreground represent our _updated expectations_ based on the data entered into the application.
-This updated information is simply the product of the prior information and the information contained in the table on the left.
-The dashed lines indicate _credible intervals_^[These are 95% credible intervals, i.e., they contain 95% of the values of $d$.] for each of our distributions of $d$.
-These intervals indicate the regions of the distributions that we have a high degree of confidence contain the true values of $d$.
-If these credible intervals do not contain zero, this suggests evidence of bias for or against the cognizable class.
-Alternatively, if these credible intervals do contain zero, we have less confidence that the true value of $d$ is not, in fact zero, and so have no evidence of bias.
+The dashed lines indicate a pre-set _credible interval_, i.e., here the values of $d$ between which 80% of the likely values of $d$ fall.
+If zero falls within this interval, then we should be wary about (have low confidence in) inferring bias (for or against the cognizable class) from the strike data.
+If zero falls outside this interval, we should feel confident in inferring bias from the strike data.
+As we get more data, we should feel more (or less) confident about particular values of $d$.
+
+The grey graph is the distribution of $d$ values based _only_ on a pre-set distribution of $d$ to reflect a starting point that bias is unlikely, as adjusted by historical strike data, if available.
+This represents our *prior beliefs* about an attorney's possible bias in using peremptory challenges, before we know anything about how any attorney uses peremptory strikes in the case before us.
+The blue (for prosecution) and red (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data (in the table on the left).
+
+# References

### 2020-06-11 prototype/app.R
add selector for cognizable class (race, sex)

@@ -25,7 +25,7 @@ ui <- fluidPage(
                    "Prior Strike History by Attorney",
                    selectInput("atty_levels_p", "Prosecution:", choices=atty_levels),
                    selectInput("atty_levels_d", "Defense:", choices=atty_levels),
-                   #helpText("helper text"),
+                   selectInput("cog_c_levels", "Cognizable Class:", choices=cog_c_levels),
                    hr(),
                    "Current Strike Tally",
                    rHandsontableOutput("hot")

### 2020-06-11 prototype/dummy_strike_data.R

@@ -22,3 +22,5 @@ shist <- data.frame(round = c(1:10),
 atty_levels <- c("None",levels(shist$atty))
 atty_levels_p <- atty_levels
 atty_levels_d <- atty_levels
+
+cog_c_levels <- c("race","gender")

### 2020-06-12 prototype/app.R
moved key files to prototype folder; modified app file to include if_then for calc_prior function

@@ -10,6 +10,8 @@ Rcpp::sourceCpp(here::here("prototype","mh_sampler.cpp"))
     
 # Load dummy dataframe and strike history
 source(here::here("prototype","dummy_strike_data.R"))
+
+source(here::here("prototype","calc_prior"))
  
 # Define UI for application
     
@@ -23,9 +25,9 @@ ui <- fluidPage(
         fluidRow(
             column(4,
                    "Prior Strike History by Attorney",
-                   selectInput("atty_levels_p", "Prosecution:", choices=atty_levels),
-                   selectInput("atty_levels_d", "Defense:", choices=atty_levels),
-                   selectInput("cog_c_levels", "Cognizable Class:", choices=cog_c_levels),
+                   selectInput("atty_p", "Prosecution:", choices=atty_levels),
+                   selectInput("atty_d", "Defense:", choices=atty_levels),
+                   selectInput("cog_c", "Cognizable Class:", choices=cog_c_levels),
                    hr(),
                    "Current Strike Tally",
                    rHandsontableOutput("hot")
@@ -73,10 +75,26 @@ server <- function(input, output, session) {
                 as.matrix()
             
             # specify prior values
-            pp_prior_mean = 0
-            pp_prior_sd = 2
-            pd_prior_mean = 0
-            pd_prior_sd = 2
+            
+            if (atty_p != "None"){
+                infrmtve_prior_p <- calc_prior(atty_p,TRUE,dat0)
+                pp_prior_mean <- infrmtve_prior_p[[1]] 
+                pp_prior_sd <- infrmtve_prior_p[[2]]  
+            }
+            else{
+                pp_prior_mean = 0 
+                pp_prior_sd = 2    
+            }
+            
+            if (atty_d != "None"){
+                infrmtve_prior_d <- calc_prior(atty_d,FALSE,dat0)
+                pd_prior_mean <- infrmtve_prior_d[[1]] 
+                pd_prior_sd <- infrmtve_prior_d[[2]]  
+            }
+            else{
+                pd_prior_mean = 0
+                pd_prior_sd = 2  
+            }
             
             out_p <- make_posterior(x = df_mp, niter = 110000, 
                                     theta_start_val = 0, theta_proposal_sd =.5, 
@@ -101,7 +119,7 @@ server <- function(input, output, session) {
             dat <- rbind(d_p,d_d)
             
             
-            ## generate priors
+            ## generate prior probability distributions
             
[…]

### 2020-06-12 prototype/calc_prior.R

@@ -1,11 +1,14 @@
 ### assume dat0 is all the historical data we have
 
 library(dplyr)
+library(mcmc)
+library(Rcpp)
+Rcpp::sourceCpp(here::here("prototype","mh_sampler.cpp"))
 
 dat0 <- readRDS("jury_data_cleaned_new.RDS")
 
 
-#### extract_atny() function extract the subset of jury_data_cleaned that corresponding to 
+#### extract_atny() function extracts the subset of jury_data_cleaned that corresponds to 
 #### the attorney name inputed in 'atny_name' parameter. 
 #### if attorney name is from PP, set 'pp' parameter as TRUE
 #### if attorney name is from PD, set 'pp' parameter as FALSE
@@ -56,8 +59,8 @@ organize_input <- function(dat,pp){
 }
 
 ### function calc_prior will return prior mean and sd using the subset data containing the attorney name
-### to get the prior mean and sd corresponding to a prosecuter attorney, we need to input 
-### attorney name of the prosecuter to parameter 'atny_name', and TRUE for 'pp' and jury_data_cleaned_new (which is the one with strike_seq) for 'dat'
+### to get the prior mean and sd corresponding to a prosecutor attorney, we need to input 
+### attorney name of the prosecutor to parameter 'atny_name', and TRUE for 'pp' and jury_data_cleaned_new (which is the one with strike_seq) for 'dat'
 ### to get the prior mean and sd corresponding to a defense attorney, we put the according name and set 'pp' as FALSE.
 
 
@@ -73,4 +76,5 @@ calc_prior <- function(atny_name,pp,dat){
 }
 
 ##### test whether the function works #####
-test_prior = calc_prior('Brian P. Leaming',TRUE,dat0)
+#test_prior = calc_prior('Brian P. Leaming',TRUE,dat0)
+

### 2020-06-12 test.R

@@ -1,13 +1,0 @@
-
-x <- rnorm(100)
-
-
-
-x <- rnorm(100)
-
-min(x)
-
-hist(x)
-
-
-plot(x)

### 2020-06-12 prototype/app.R
moved cognitive class choices to app.R. file; removed dummy strike history data

@@ -8,10 +8,14 @@ library(shinythemes)
 
 Rcpp::sourceCpp(here::here("prototype","mh_sampler.cpp"))
     
-# Load dummy dataframe and strike history
+# Load dummy dataframe
 source(here::here("prototype","dummy_strike_data.R"))
 
 source(here::here("prototype","calc_prior"))
+
+# Define Cognitive Class choices
+
+cog_c_levels <- c("race","gender")
  
 # Define UI for application
     

### 2020-06-12 prototype/dummy_strike_data.R

@@ -1,3 +1,5 @@
+#### dummy strike data ####
+
 df0 <- data.frame(round = c(1:10), 
                   num_cog = c(3, 4, 4, 3, 2, 2, 2, 2, 2,2),
                   total = rep(9, 10), 
@@ -8,19 +10,11 @@ df0 <- data.frame(round = c(1:10),
 party_choices <- c("PP","PD")
 
 
-# dummy strike data
+#### attorney name choices ####
 
-shist <- data.frame(round = c(1:10), 
-                    num_cog = c(3, 4, 4, 3, 2, 2, 2, 2, 2,2),
-                    total = rep(9, 10), 
-                    cog = c(0, 1, 1, 1, 1, 0, 1, 1, 1,0),
-                    party = rep(c("PP","PD"),5),
-                    atty = factor(rep(c("Peter Parker","Matt Murdock"),5)),
-                    firm = rep(c("Parker,Stan,and Lee","Hell's Kitchen Defense"),5)    
-)
+dat0 <- readRDS(here::here("prototype","jury_data_cleaned_new.rds"))
+
+atty_levels_p <- factor(dat0$P_atty_l)
+atty_levels_d <- factor(dat0$D_atty_l)
 
 atty_levels <- c("None",levels(shist$atty))
-atty_levels_p <- atty_levels
-atty_levels_d <- atty_levels
-
-cog_c_levels <- c("race","gender")

### 2020-06-13 prototype/dummy_strike_data.R
extract attorney names for selection boxes

@@ -14,7 +14,19 @@ party_choices <- c("PP","PD")
 
 dat0 <- readRDS(here::here("prototype","jury_data_cleaned_new.rds"))
 
-atty_levels_p <- factor(dat0$P_atty_l)
-atty_levels_d <- factor(dat0$D_atty_l)
+atty_levels_p <- dat0 %>%
+  pull("P_atty_l") %>%
+  unlist() %>%
+  unique() %>%
+  str_trim() %>%
+  factor()
 
-atty_levels <- c("None",levels(shist$atty))
+atty_levels_d <- dat0 %>%
+  pull("D_atty_l") %>%
+  unlist() %>%
+  unique() %>%
+  str_trim() %>%
+  factor()
+
+atty_levels_p <- c("None",levels(atty_levels_p))
+atty_levels_d <- c("None",levels(atty_levels_d))

### 2020-06-13 prototype/calc_prior.R
align attorney selection boxes with calc_prior, plots; plot xlim(6,-6); changed names of cognitive classes

@@ -9,7 +9,7 @@ dat0 <- readRDS("jury_data_cleaned_new.RDS")
 
 
 #### extract_atny() function extracts the subset of jury_data_cleaned that corresponds to 
-#### the attorney name inputed in 'atny_name' parameter. 
+#### the attorney name inputted in 'atny_name' parameter. 
 #### if attorney name is from PP, set 'pp' parameter as TRUE
 #### if attorney name is from PD, set 'pp' parameter as FALSE
 #### parameter 'pp' is the same for the following functions, used to control whether the case is from PP or PD

### 2020-06-13 prototype/dummy_strike_data.R

@@ -3,7 +3,7 @@
 df0 <- data.frame(round = c(1:10), 
                   num_cog = c(3, 4, 4, 3, 2, 2, 2, 2, 2,2),
                   total = rep(9, 10), 
-                  cog = c(0, 1, 1, 1, 1, 0, 1, 1, 1,0),
+                  cog = c(0, 1, 1, 0, 0, 1, 1, 0, 0,0),
                   party = rep(c("PP","PD"),5)
 )
 

### 2020-06-13 prototype/calc_prior.R
add parameter to specify cognizable class

@@ -29,26 +29,38 @@ extract_atny <- function(atny_name,pp,dat){
 #### round, num of female, total number, and whether the strike is from cognizable class
 #### here I assume cognizable class is female 
 #### the return matrix is of the form of the input of make_posterior function
+#### input of cog should be either 'gender' or 'race'
 
-organize_input <- function(dat,pp){
-  dat_strikes <- dat %>% filter(!is.na(strike_seq)) %>% filter(!is.na(sex))
+organize_input <- function(dat,pp,cog){
+  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
+  if(cog=='gender'){
+    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
+  }else{
+    dat_strikes <- dat_strikes %>% filter(!is.na(race))
+  }
+  
   dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
-  dat_strikes$sexb <- ifelse(dat_strikes$sex=="F", 1, 0)
-  num_female <- sum(dat_strikes$sexb)
+  if(cog=='gender'){
+    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
+  }else{
+    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
+  }
+  
+  num_cog <- sum(dat_strikes$cogb)
   num_t <- nrow(dat_strikes)
   subs <- data.frame(round = c(1:num_t), 
-                     num_female = NA, 
+                     num_cog = NA, 
                      total = NA)
-  subs$num_female[1] <- num_female
+  subs$num_cog[1] <- num_cog
   subs$total[1] <- num_t
   for ( j in 2:num_t){
-    num_female = num_female-dat_strikes$sexb[j-1]
+    num_cog = num_cog-dat_strikes$cogb[j-1]
     num_total = num_t - dat_strikes$strike_seq[j-1]
-    subs$num_female[j] <- num_female
+    subs$num_cog[j] <- num_cog
     # number of females struck in this and previous rounds
     subs$total[j] <- num_total
   }
-  subs <- cbind(subs,dat_strikes$sexb)
+  subs <- cbind(subs,dat_strikes$cogb)
   colnames(subs)[4] = 'cognizable'
   if(pp){
     df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
[…]

### 2020-06-17 prototype/explanatory text.Rmd
moved key files to prototype folder; modified app file to include if_then for calc_prior function
navbar structure for UI; explanatory text in seperate tabset

--- 2020-06-11 explanatory text.Rmd
@@
 ---
-title: "Application Outline"
+title: "About this Prototype"
 output: 
@@
 
-## Aim
-
 This application estimates how likely an attorney used peremptory challenges during jury selection to strike potential jurors based on their race or gender.
@@
 
-## Analysis
+To estimate such bias, this application relies on two kinds of data: (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, available jury-selection data on that attorney's use of strikes in some past cases during 2013-2017 in the federal district court for the District of Connecticut.
 
-To estimate bias, this application relies on two kinds of data: (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, pre-compiled historical data on that attorney's use of strikes in past cases.
+### Analysis
 
@@
 
-Here, we represent $d$ as a distribution of how likely each possible value of $d$ is, given the data we have. This lets us depict how confident we should be in inferring bias from the data we have.
+Here, we represent $d$ as a distribution of how likely each possible value of $d$ is, given the data we have.
+This lets us use the curve's width (the distribution's *standard deviation*) to depict how confident we should be in inferring bias from the data we have: The narrower the curve, the more confident we should be.
 
@@
 If zero falls outside this interval, we should feel confident in inferring bias from the strike data.
-As we get more data, we should feel more (or less) confident about particular values of $d$.
 
-The grey graph is the distribution of $d$ values based _only_ on a pre-set distribution of $d$ to reflect a starting point that bias is unlikely, as adjusted by historical strike data, if available.
+The grey graph depicts $d$ based _only_ on a pre-set distribution of $d$ to reflect a starting point that bias is unlikely, as adjusted by any historical strike data we have.
 This represents our *prior beliefs* about an attorney's possible bias in using peremptory challenges, before we know anything about how any attorney uses peremptory strikes in the case before us.
-The blue (for prosecution) and red (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data (in the table on the left).
+The red (for prosecution) and blue (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data (in the table on the left).
 
-# References
+### References

### 2020-06-17 prototype/app.R

@@ -5,6 +5,7 @@ library(mcmc)
 library(Rcpp)
 library(ggplot2)
 library(shinythemes)
+library(markdown)
 
 Rcpp::sourceCpp(here::here("prototype","mh_sampler.cpp"))
     
@@ -22,14 +23,13 @@ cog_c_levels <- c("racial minority","women")
 # Define UI for application
     
 ui <- fluidPage(
+
+    #Navbar structure for UI
     
-        theme = shinytheme("cerulean"),
-        
-        # Application title
-        titlePanel("Batson App"),
-    
-        fluidRow(
-            column(4,
+    navbarPage("Batson App", theme = shinytheme("paper"),
+               
+        tabPanel("Prototype", fluid = TRUE,
+            sidebarPanel(
                    "Prior Strike History by Attorney",
                    selectInput("atty_p", "Prosecution:", 
                                choices=atty_levels_p,
@@ -41,14 +41,14 @@ ui <- fluidPage(
                    hr(),
                    "Current Strike Tally",
                    rHandsontableOutput("hot")
-            ),
-            column(6,
-                   plotOutput("plot"),
-                   "Explanatory text here."
-                   
-            )
-                )
-    )
+                ),
+            mainPanel(plotOutput("plot"))
+                ),
+        
+        tabPanel("About", fluid = TRUE,
+                 includeMarkdown("explanatory-text.md"))
+        ))
+
     
 # Define server logic 
 server <- function(input, output, session) {

### 2020-06-17 prototype/dummy_strike_data.R

@@ -18,14 +18,14 @@ atty_levels_p <- dat0 %>%
   pull("P_atty_l") %>%
   unlist() %>%
   unique() %>%
-  str_trim() %>%
+  stringr::str_trim() %>%
   factor()
 
 atty_levels_d <- dat0 %>%
   pull("D_atty_l") %>%
   unlist() %>%
   unique() %>%
-  str_trim() %>%
+  stringr::str_trim() %>%
   factor()
 
 atty_levels_p <- c("None",levels(atty_levels_p))

### 2020-06-17 prototype/app.R
account for selection of cognizable class

@@ -18,7 +18,7 @@ source(here::here("prototype","calc_prior.R"))
 
 # Define Cognitive Class choices
 
-cog_c_levels <- c("racial minority","women")
+cog_c_levels <- c("race","gender")
  
 # Define UI for application
     
@@ -86,8 +86,10 @@ server <- function(input, output, session) {
             
             # specify prior values
             
+            # Note: same priors regardless of cognizable class if atty = "None"
+            
             if (input$atty_p != "None"){
-                infrmtve_prior_p <- calc_prior(input$atty_p,TRUE,dat0)
+                infrmtve_prior_p <- calc_prior(input$atty_p,TRUE,dat0,input$cog_c)
                 pp_prior_mean <- infrmtve_prior_p[[1]] 
                 pp_prior_sd <- infrmtve_prior_p[[2]]  
             }
@@ -97,9 +99,9 @@ server <- function(input, output, session) {
             }
             
             if (input$atty_d != "None"){
-                infrmtve_prior_d <- calc_prior(input$atty_d,FALSE,dat0)
-                pd_prior_mean <- infrmtve_prior_d[[1]] 
-                pd_prior_sd <- infrmtve_prior_d[[2]]  
+                infrmtve_prior_d <- calc_prior(input$atty_d,FALSE,dat0,input$cog_c)
+                pd_prior_mean <- infrmtve_prior_d[[1]]
+                pd_prior_sd <- infrmtve_prior_d[[2]]
             }
             else{
                 pd_prior_mean = 0

### 2020-06-17 prototype/calc_prior.R
update code comments on cognizable class

@@ -26,8 +26,7 @@ extract_atny <- function(atny_name,pp,dat){
 } 
 
 #### function organize_input will return a matrix with four columns:
-#### round, num of female, total number, and whether the strike is from cognizable class
-#### here I assume cognizable class is female 
+#### round, num of cog_class, total number, and whether the strike is from cognizable class
 #### the return matrix is of the form of the input of make_posterior function
 #### input of cog should be either 'gender' or 'race'
 

### 2020-06-17 prototype/explanatory text.Rmd
revisions, renaming bias parameter as "b"

--- 2020-06-11 explanatory text.Rmd
@@
 ---
-title: "Application Outline"
+title: "About"
 output: 
@@
 
-## Analysis
+To estimate such bias, this application relies on (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, jury-selection data on that attorney's use of strikes in some past cases during 2013-2017 in the federal district court for the District of Connecticut.
 
-To estimate bias, this application relies on two kinds of data: (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, pre-compiled historical data on that attorney's use of strikes in past cases.
+Based on this data, this application graphs $b$ -- our measure of bias -- for the prosecution and the defense: **positive** values reflect bias _against_ the cognizable class, while **negative** values reflect bias _for_ the cognizable class.
+If $b = 0$, that there is zero bias.
 
-In the above plots, $d$ is our measure of bias: **positive** values of $d$ reflect bias _against_ the cognizable class, while **negative** values  of $d$ reflect bias _for_ the cognizable class.
-If $d$ is exactly **zero**, we can be completely certain that there is zero bias.
+The plots show how likely each possible value of $b$ is, given the data we have.
+The curve's spread (the distribution's *variance*) depicts how confident we should be in any particular value of $b$ from the data we have.
 
-Here, we represent $d$ as a distribution of how likely each possible value of $d$ is, given the data we have. This lets us depict how confident we should be in inferring bias from the data we have.
+The dashed lines indicate a pre-set _credible interval_, i.e., here the values of $d$ between which 80% of the likely values of $b$ fall.
+If zero falls within this interval, then we should be wary about (have low confidence in) inferring bias (for or against the cognizable class) from the strike data alone.
+If zero falls outside this interval, we should feel confident in inferring bias from that strike data.
 
-The dashed lines indicate a pre-set _credible interval_, i.e., here the values of $d$ between which 80% of the likely values of $d$ fall.
-If zero falls within this interval, then we should be wary about (have low confidence in) inferring bias (for or against the cognizable class) from the strike data.
-If zero falls outside this interval, we should feel confident in inferring bias from the strike data.
-As we get more data, we should feel more (or less) confident about particular values of $d$.
+The grey graph depicts $b$ based _only_ on a pre-set distribution of $b$ to reflect a starting point that bias is unlikely, as adjusted by any historical strike data we have for a particular selected attorney.
+This represents our *prior beliefs* about an attorney's possible bias in using peremptory challenges, before we know anything about how any attorney uses peremptory strikes in the case before us.
+The red (for prosecution) and blue (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data.
 
-The grey graph is the distribution of $d$ values based _only_ on a pre-set distribution of $d$ to reflect a starting point that bias is unlikely, as adjusted by historical strike data, if available.
-This represents our *prior beliefs* about an attorney's possible bias in using peremptory challenges, before we know anything about how any attorney uses peremptory strikes in the case before us.
-The blue (for prosecution) and red (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data (in the table on the left).
-
-# References
+### References

### 2020-06-17 prototype/app.R

@@ -166,7 +166,7 @@ server <- function(input, output, session) {
             ## labels and theme
             
              pplot <- pplot  + theme_minimal() +
-                 labs (title = "Posterior density of d") +
+                 labs (title = "Posterior density of b") +
                  xlab("") + 
                  ylab("") + 
                  xlim(c(-6,6)) +

### 2020-06-22 prototype/model_text.Rmd
added a seperate "model" explanation -- currently incomplete, needs likelihood function; simplified "about" text.

---
title: "Model"
output: 
  github_document:
       pandoc_args: --webtex
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Like @Kadane2018 and @Barrett2007, we model the probability that the struck person belonged (or not) to the cognitive class as a function of (1) the number of cognizable class members (e.g., racial minorities, women) and non-class members (e.g., White people, men) in the set of people who could have been struck; and (2) an unobserved variable $b$ that indicates how much weight the striking party placed on the struck juror's membership in the cognitive class (e.g., whether that juror was racial minority or White; women or man).

For any given jury selection $i$, let $j$ denote a peremptory strike used (among $t$ total strikes used), and let $X_{ij}$ denote whether a party used that strike on a person who belongs to a "cognizable class", _i.e._, $X_{ij} = 0$ if no, and $X_{ij} = 1$ if yes. If "race" is the bias type of interest, the cognizable class ($X = 1$) is racial minority jurors ($X = 0$ for White jurors).  If "gender" is the bias of interest, the cognizable class is female jurors ($X=0$ for male jurors).

We are interested in estimating the likely values of $b$, given the strike data we have. There is zero bias only if the probability of striking a cognizable class member is the same as the probability of striking a non-cognitive class member.  Accordingly,

$$w = \frac{Pr(X=1)}{Pr(X = 0)}\\
w = e^b$$

This way, if it is equally likely that a member and non-member of the cognizable class will be stricken, then $w = \frac{Pr(X=1)=0.50}{Pr(X = 0)=0.50}=1$, then our measure of bias is zero, because if $w=e^b$, and $e^b=1$ then $b=log(1)=0$. If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognitive class ($X=1$, e.g., the juror is a racial minority). Where $b<1$, the party has bias toward striking a juror falling outside the cognitive class ($X=0$, e.g., the juror is White).

In turn, let $c$ denote the number of cognizable class members; and $m$ denote the number of cognizable class non-members, such that $c+m$ is the total number of jurors potentially subject to strike.

$$Pr(X_{ij}) =
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }X=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }X=0  
\end{cases}
$$

### References

### 2020-06-22 prototype/app.R

@@ -46,7 +46,9 @@ ui <- fluidPage(
                 ),
         
         tabPanel("About", fluid = TRUE,
-                 includeMarkdown("explanatory-text.md"))
+                 includeMarkdown("explanatory-text.md")),
+        tabPanel("Model", fluid = TRUE,
+                 includeMarkdown("model_text.md"))
         ))
 
     

### 2020-06-24 prototype/app.R
factor party

@@ -149,6 +149,10 @@ server <- function(input, output, session) {
             
             ## calculate credible intervals
             
+            ## specify party as factor
+            dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
+                                ordered = TRUE)
+            
             CI <- dat %>%
                 group_by(party) %>%
                 summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%

### 2020-06-24 prototype/model_text.Rmd
add likelihood function

@@
 
-For any given jury selection $i$, let $j$ denote a peremptory strike used (among $t$ total strikes used), and let $X_{ij}$ denote whether a party used that strike on a person who belongs to a "cognizable class", _i.e._, $X_{ij} = 0$ if no, and $X_{ij} = 1$ if yes.
+For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $X_{ij}$ denote whether a party used that strike on a person who belongs to a "cognizable class", _i.e._, $X_{ij} = 0$ if no, and $X_{ij} = 1$ if yes.
 If "race" is the bias type of interest, the cognizable class ($X = 1$) is racial minority jurors ($X = 0$ for White jurors).
@@
 
-$$w = \frac{Pr(X=1)}{Pr(X = 0)}\\
-w = e^b$$
+$$w = \frac{Pr(X=1)}{Pr(X = 0)}$$
+$$w = e^b$$
+$$b \sim \text{Normal}(0, 2)$$
 
-This way, if it is equally likely that a member and non-member of the cognizable class will be stricken, then $w = \frac{Pr(X=1)=0.50}{Pr(X = 0)=0.50}=1$, then our measure of bias is zero, because if $w=e^b$, and $e^b=1$ then $b=log(1)=0$.
+This way, if it is equally likely that a member and non-member of the cognizable class will be stricken, then $w = \frac{Pr(X=1)=0.50}{Pr(X = 0)=0.50}=1$.
+If so, our measure of bias is zero: given $w=e^b$, then $e^b=1$ and $b=log(1)=0$.
 If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognitive class ($X=1$, e.g., the juror is a racial minority).
@@
 
-$$Pr(X_{ij}) =
+$$
+Pr(X_{ij}) =
 \begin{cases}
@@
 
+The likelihood function of b is 
+$$L(b|X) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$
+where $n_{i}$ is the total number of jury selections and $n_j$ is the total number of peremptory strikes; and where $\delta_{ij} = 1$ when $X_{ij} = 1$ and $\delta_{ij} = 0$ when $X_{ij}=0$.
+
+The posterior probability distribution is estimated with R 4.0 [@R] and mcmc 0.9-7 [@mcmc].
+
 ### References

### 2020-06-24 prototype/explanatory text.Rmd
added a seperate "model" explanation -- currently incomplete, needs likelihood function; simplified "about" text.
edits to "model" text (replace "d" with "b"); edits to explanatory text (changed depiction of prior)

@@
 This application estimates how likely an attorney used peremptory challenges during jury selection to strike potential jurors based on their race or gender.
-The analysis builds on prior work by @Kadane2018 and @Barrett2007 on statistical analysis for _Batson_ challenges.
-
+The analysis builds on prior work on statistical analysis for _Batson_ challenges.
 To estimate such bias, this application relies on (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, jury-selection data on that attorney's use of strikes in some past cases during 2013-2017 in the federal district court for the District of Connecticut.
@@
 
-The dashed lines indicate a pre-set _credible interval_, i.e., here the values of $d$ between which 80% of the likely values of $b$ fall.
+The dashed lines indicate a pre-set _credible interval_, i.e., here the values of $b$ between which 80% of the likely values of $b$ fall.
 If zero falls within this interval, then we should be wary about (have low confidence in) inferring bias (for or against the cognizable class) from the strike data alone.
@@
 The red (for prosecution) and blue (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data.
-
-### References

### 2020-06-24 prototype/model_text.Rmd
edits to "model" text (replace "d" with "b"); edits to explanatory text (changed depiction of prior)

@@
 
-Like @Kadane2018 and @Barrett2007, we model the probability that the struck person belonged (or not) to the cognitive class as a function of (1) the number of cognizable class members (e.g., racial minorities, women) and non-class members (e.g., White people, men) in the set of people who could have been struck; and (2) an unobserved variable $b$ that indicates how much weight the striking party placed on the struck juror's membership in the cognitive class (e.g., whether that juror was racial minority or White; women or man).
+Like @Kadane2018 and @Barrett2007, we model the probability that the struck person belonged (or not) to the cognizable class as a function of (1) the number of cognizable class members (e.g., racial minorities, women) and non-class members (e.g., White people, men) in the set of people who could have been struck; and (2) an unobserved variable $b$ that indicates how much weight the striking party placed on the struck juror's membership in the cognizable class (e.g., whether that juror was racial minority or White; women or man).
 
@@
 We are interested in estimating the likely values of $b$, given the strike data we have.
-There is zero bias only if the probability of striking a cognizable class member is the same as the probability of striking a non-cognitive class member.
+There is zero bias only if the probability of striking a cognizable class member is the same as the probability of striking a non-cognizable class member.
 Accordingly,
@@
 $$w = e^b$$
-$$b \sim \text{Normal}(0, 2)$$
+$$b \sim \text{Normal}(\mu_b,\sigma_b)$$
+$$\mu_b = 0$$
+$$\sigma_b = 2$$
 
@@
 If so, our measure of bias is zero: given $w=e^b$, then $e^b=1$ and $b=log(1)=0$.
-If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognitive class ($X=1$, e.g., the juror is a racial minority).
-Where $b<1$, the party has bias toward striking a juror falling outside the cognitive class ($X=0$, e.g., the juror is White).
+If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class ($X=1$, e.g., the juror is a racial minority).
+Where $b<1$, the party has bias toward striking a juror falling outside the cognizable class ($X=0$, e.g., the juror is White).
 
@@
 $$L(b|X) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$
-where $n_{i}$ is the total number of jury selections and $n_j$ is the total number of peremptory strikes; and where $\delta_{ij} = 1$ when $X_{ij} = 1$ and $\delta_{ij} = 0$ when $X_{ij}=0$.
+where $n_{i}$ is the total number of jury selections; $n_j$ is the total number of peremptory strikes; $\delta_{ij} = 1$ when $X_{ij} = 1$; and $\delta_{ij} = 0$ when $X_{ij}=0$.
 

### 2021-02-24 analysis/paper_draft.Rmd
added paper draft and related .bib file

---
title: "Visualizing Bias for Batson Challenges Using Historical Selection Data"
date: "`r Sys.Date()`"
bibliography: references.bib
author: "Pandya, Li, Moore"
output: 
    html_document:
    theme: readable
---

# Introduction

This paper demonstrates the use of a power prior as part of a software application that, during jury selection, attorneys can use in real time to estimate the probability that a prosecutor or defense attorney used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v. Kentucky_ (1986), and its progeny.

# Background

## Power Prior

The power prior is a way to create informative prior distributions based on
historical data, or more precisely, by controlling how much historical data affects the posterior distribution of the unobserved variables of interest. For a review, see @Ibrahim2015. [MORE]

## Jury Selection

In the US, jury selection proceeds in three steps.  First, a court identifies individuals as part of a qualified jury pool, often based on voter registration and driver's license records.  Second, for a given case, a subset of individuals in the jury pool are summoned to appear for possible jury service. Upon their appearance, this group (the "venire") is subject to "voir dire", that is, they are questioned (by the parties or the trial judge) and are (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise (by one of the parties' attorneys) of a peremptory challenge.

The key feature of a peremptory challenge is that, with it, a party declare a prospective juror ineligible ("strikes" that juror) for a seat on the jury. Each side is assigned a number of peremptory challenges and may use them to strike any potential juror without explanation. The number of peremptory challenges varies by jurisdiction, as does how the procedure by which peremptory challenges are exercised. Where State law prescribes no uniform procedure, that procedure can vary by trial judge in that State. In federal court, practices vary by court and judge, because no federal rule specifies the procedure to be used [@Williams2017].

Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p. 297-298]. This particularly matters where each side does not have the same number of strikes.  For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.

Under _Batson v. Kentucky_ (1986)_, a party's exercise of a peremptory challenge violates the Equal Protection Clause of the US Constitution if motivated by a prospective juror's race, gender, or certain other characteristics. To raise a _Batson_ challenge during jury selection, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status. Once established, the party exercising the strike bears the burden of producing a nondiscriminatory reason for that strike.  Thereafter, the challenger bears the burden of persuading the trial judge that the race or gender was the real motivation for the strike in question.^[Some jurisdictions adopt a modified version of _Batson_ that reduces or eliminates the challenger's burden to establish a prima facie case. E.g., State v. Holloway, 209 Conn. 636, 646 & n.4 (1989); State v. Chapman, 317 S.C. 302, 305-06 (1995); see also State v. Antwine, 743 S.W.2d 51, 64 (Mo. 1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]

Evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also historical data on peremptory challenges in past case. Though usually revealed after painstaking investigation of jury selection records [e.g., @Craft2018], such information, if routinely available, would be relevant, in any particular case, to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in any particular case.^[See, e.g., Miller-El v. Dretke, 545 U.S. 231, 253 (2005); Riley v. Taylor, 277 F.3d 261, 280-81, 283 (3d Cir. 2001); Cochran v. Herring, 43 F.3d 1404, 1412 (11th Cir. 1995).]

More recently, a few States (e.g., Washington State, California) have provided an independent basis to challenge peremptory strikes. For example, by court rule, a trial court in Washington State may deny a peremptory challenge if it determines that "an objective observer could view race or ethnicity as a factor in the use of the peremptory challenge."^[R. 37(e); see also Rule 37(f) ("an obiective obseryer is aware that implicit, institutional, and unconscious biases, in addition to purposeful discrimination, have resulted in the unfair exclusion of potential jurors in Washington State"); Rule 37(h) (identifying presumptively invalid reasons for a peremptory challenge).].

Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018]. This pattern persists, despite the fact that both sides' attorneys can often observe the race and gender of those in the venire who were and were not struck by peremptory challenge.

The research literature suggests several explanations. These include a reluctance to raise a _Batson_ challenge on the premise that such a challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021, 148, 151, 158]. Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge. Moreover, drawing inferences from historical data is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018].

In theory, however, were an attorney's strike history in past cases easily available, such strike history can be compared to what we would expect to happen if, following the same procedure, venire members were struck at random [@Kadane2018; @Kadane2018a]. To do so, one would need to, for any peremptory-selection method, a corresponding likelihood function for how a prosecutor or defense attorney strikes prospective jurors of a protected class (e.g., racial-minorities). Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter to jury selection, unless "statisticans accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]

# Method

[insert intro]

# Model 1

Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides. Accordingly, in a non-capital felony trial in federal district court using this method, the trial judge will, after excusing jurors for cause, make sure that there are 28 potential jurors remaining before the parties exercise their peremptory challenges (12 jurors plus 16 peremptories, assuming no alternate jurors). Then, the parties exercise their strikes on those 28 potential jurors.

In a typical variant of the "struck jury" method, the parties strike in an alternating sequence, with the prosecution striking first. Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn. If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.

Under "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box. Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.  Once that is completed, new potential jurors take the seats of the excused or striken jurors, and the process repeats until all the jury seats are filled.  At any point, strikes can be used on anyone in the jury box, including those there previously but not struck.

For these peremptory challenges, following @Kadane2018 and @Barrett2007, we model the probability that the struck person belonged (or not) to the cognizable class as a function of (1) the number of cognizable class members (e.g., racial minorities, women) and non-class members (e.g., White people, men) in the set of people who could have been struck; and (2) a parameter $b$ that indicates how much weight (or "bias") the striking party placed on the struck juror's membership in the cognizable class (e.g., whether that juror was a racial minority or White; women or man).

For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class". If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).  If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors). In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.

Accordingly, for any given value of the bias parameter $b$, the probability of strike, or $Pr(\delta_{ij})$, is:

$$
Pr(\delta_{ij} | b) = 
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
\end{cases}
$$
This is equivalent to

$$Pr(\delta_{ij}| b) = \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$

If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.

Because $b$ is not directly observable, we are interested in estimating the likely values of $b$, given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$.

Accordingly, we model the jury-selection process as follows.

\[L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}\]
\[b \sim \text{Normal}(0,2)\]

where $n_{i}$ is the total number of jury selections; and $n_j$ is the total number of peremptory strikes.

This way, if it is equally likely that a member and non-member of the cognizable class will be stricken, then $w = \frac{Pr(X=1)=0.50}{Pr(X = 0)=0.50}=1$. If so, our measure of bias is zero: given $w=e^b$, then $e^b=1$ and $b=log(1)=0$. If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class ($X=1$, e.g., the juror is a racial minority). Where $b<1$, the party has bias toward striking a juror falling outside the cognizable class ($X=0$, e.g., the juror is White).

# Model 2 - Sequential Strike Method

Under the "sequential strike" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptory challenges. Then, the remaining prospective jurors appear one at a time before the attorneys for prosecutor and defendant, and those attorneys alternate as the order in which they decide whether to strike or not the prospective juror before them. For example, if the prosecutor had first choice whether to strike juror #1, the defendant's attorney would have first choice whether to strike juror # 2. If neither party exercises a strike on the prospective juror before them, that prospective juror is assigned a seat on the jury and the parties' attorneys cannot thereafter strike that juror. The process continues until the parties run out of strikes or the number seats on the jury (including alternates) have been filled.

Accordingly, we model the sequential-strike process as follows.

[insert statistical model]

# The App

[insert]

The posterior probability distribution is estimated with R 4.0 [@R] and mcmc 0.9-7 [@mcmc]. The app was built using the shiny package [@shiny].

# Discussion

# Conclusion

# CRediT Author Statement

Sachin S. Pandya: Conceptualization; Xiaomeng Li: Conceptualization; [insert]; Timothy Moore: Conceptualization; [insert]. Based on categories in https://casrai.org/credit/].

# References

### 2021-05-08 prototype/app.R
changed plot title

@@ -206,7 +206,7 @@ server <- function(input, output, session) {
             ## labels and theme
             
              pplot <- pplot  + theme_minimal() +
-                 labs (title = "Posterior density of b") +
+                 labs (title = "Likely values of b") +
                  xlab("") + 
                  ylab("") + 
                  xlim(c(-6,6)) +

### 2021-05-08 analysis/paper_draft.Rmd

@@
 If so, our measure of bias is zero: given $w=e^b$, then $e^b=1$ and $b=log(1)=0$.
-If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class ($X=1$, e.g., the juror is a racial minority).
-Where $b<1$, the party has bias toward striking a juror falling outside the cognizable class ($X=0$, e.g., the juror is White).
+If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class ($X=1$, e.g., the juror is a racial minority).
+Where $b<0$, the party has bias toward striking a juror falling outside the cognizable class ($X=0$, e.g., the juror is White).
 
@@
 [insert statistical model]
+
+# Simulation 
+We conduct simulation study under different scenarios, including no bias, bias against cognizable class and preference for cognizable class, for both models.
+We have also studied the scenarios with different amount of historical data and whether the bias parameter of historical data is the same as current data.
+We show the model accuracy under different scenarios and how it improves the model accuracy by adding historical data and modifying the weight parameter.
+
+## Model 1
+As introduced in previous section, parameter b is used to measure the bias.
+We simulate data with 7 different values of $b = -3,-2,-1,0,1,2,3$ representing strong bias, bias, moderate bias toward striking a juror falling outside the cognizable class, no bias striking a juror and moderate bias, bias and strong bias favoring a strike against a juror falling within the cognizable.
+We also generate historical data with bias parameter from the 7 values.
+As for historical data, we consider it from one trial, two trials and three trials to represent different size of historical data.
+Consider the bias parameter of a person might change over time, we allow the historical data generated from a different bias parameter than the current data.
+When we use power prior incorporating the historical information, we can control the weight that we put on historical data by modifying the weight parameter.
+In the simulation study, we also try different values of the weight parameter to show the influence of the weight parameter on modeling accuracy.
+We summarize the results in the figure below.
+
+```{r pressure, echo=FALSE, fig.cap="A caption", out.width = '100%'}
+knitr::include_graphics("dino_plots.pdf")
+```
+
+Historical b means the bias parameter for the generated historical data, b means the bias parameter for the current data.
+The values on the right side represent the size of the histrocial data.
+The values on the top of the plots represent the weight parameter of the power prior.
+Different color represents the estimated bias parameter.
+
+From the plot, our model can capture strong bias and bias, when the bias parameter being -3,-2,2,3.
+When the bias is moderate, that is when the bias parameter b being -1 and 1, the model can detect it when we have historical data with the same direction bias parameter.
+If the bias parameter of historical data is in an ooposite direction, we might get a wrong estimate of current bias parameter.
+Thus if the historical data is not reliable, we should set a small weight parameter on them.
 

### 2021-05-08 prototype/dpp.R
declares all packages to be installed during build of Docker image for the Shiny app

@@ -1,0 +1,13 @@
+# dpp.R declares all the packages needed to run the Shiny app. They will be
+# installed during the building process of the Docker image for the Shiny app
+
+# list all the packages as a string list, e.g., c("shiny", "ggplot2", ...)
+packages <- c("shiny","rhandsontable","dplyr","mcmc","Rcpp","ggplot2","shinythemes")
+
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(
+    setdiff(packages, rownames(installed.packages())),
+    dependencies = T,
+    repos='https://cloud.r-project.org/'
+  )
+}

### 2021-05-08 prototype/app.R
removed tab for explanatory text about model and app

@@ -52,13 +52,7 @@ ui <- fluidPage(
                 actionButton("updateButton", "Update",
                              icon = icon("refresh")),
                 plotOutput("plot"))
-                ),
-        
-        tabPanel("About", fluid = TRUE,
-                 includeMarkdown("explanatory-text.md"))
-        #,
-        #tabPanel("Model", fluid = TRUE,
-        #         includeMarkdown("model_text.md"))
+                )
         ))
 
     

### 2021-05-08 sequential.R
updated

@@ -86,33 +86,23 @@ sim_seq <- function(num_seat,num_strikep,num_striked,p0,p_c){
 
 
 #### test  ###
-a = sim_seq(14,7,7,0.2,0.9)
+#a = sim_seq(14,7,7,0.2,0.9)
 ### delete rows with 0 strikes left
-a = as.data.frame(a)
-if(min(which(a$num_strike_p==0))<nrow(a)){
+#a = as.data.frame(a)
+#if(min(which(a$num_strike_p==0))<nrow(a)){
   a1 = a[-which(a$num_strike_p==0 & a$att=='PP'),]
 }else{
   a1 = a[-which(a$num_strike_d==0 & a$att=='PD'),]
 }
 
-pp2 = a1[which(a1$att=='PP'),]
-pd2 <- a1[which(a1$att=='PD'),]
+#pp2 = a1[which(a1$att=='PP'),]
+#pd2 <- a1[which(a1$att=='PD'),]
 
-b = bayesglm(strike~cognizable,data=pp2,family='binomial',prior.scale.for.intercept=2.5)
-summary(b)
+#b = bayesglm(strike~cognizable,data=pp2,family='binomial',prior.scale.for.intercept=2.5)
+#summary(b)
 
 ### get the confidence interval of the estimated coefficient
-simulates <- coef(sim(b))
-hist(simulates[,1])
-quantile(simulates[,1],c(0.1,0.9))
-
-
-
-
-
-
-
-
-
-
+#simulates <- coef(sim(b))
+#hist(simulates[,1])
+#quantile(simulates[,1],c(0.1,0.9))
 

### 2021-05-08 sim_seq.R
simulating sequential selection

@@ -1,0 +1,93 @@
+sim_seq <- function(num_seat,num_strikep,num_striked,p0,p_c){
+  stopifnot(num_seat > 5, 
+            num_strikep < 1,
+            num_striked < 1, 
+            p0 <=1,
+            p0 >= 0, 
+            p_c <=1,
+            p_c >= 0,
+            is.integer(num_seat),
+            is.integer(num_strikep),
+            is.integer(num_striked),
+            is.numeric(p0),
+            is.numeric(p_c)) # all must be TRUE
+  #### generate round, attorney, juror id, cognizable, strike, num_seats left, num_strike left p\d
+  round = 1
+  att <- c('PP','PD')
+  att_idx <- 1
+  jid <- 1
+  ### jid1 used to consider the situation when the first party didn't strike
+  jid1 <- 1
+  cog <- rbinom(1,1,0.5)
+  stk <- ifelse(cog==1,rbinom(1,1,p_c),rbinom(1,1,p0))
+  nseat <- num_seat
+  nstkp <- num_strikep
+  nstkd <- num_striked
+  dat0 <- c(round,att[att_idx],jid,cog,stk,nseat,nstkp,nstkd)
+  nseat <- nseat
+  nstkp <- nstkp - (2-att_idx)*stk
+  nstkd <- nstkd + (1-att_idx)*stk
+  jid1= (stk==0)
+  while(nseat>=0 && (nstkp>0 || nstkd>0)){
+    round = round+1
+    if(stk==1){
+      jid = jid+1
+      att_idx <- 2-jid%%2
+      cog = rbinom(1,1,0.5)
+      if (att_idx==1 && nstkp==0){
+        stk=0
+      }else if(att_idx==2 && nstkd==0){
+        stk=0
+      }else{
+        stk <- ifelse(cog==1,rbinom(1,1,p_c),rbinom(1,1,p0))
+      }
+      
+      dat0 <- rbind(dat0,c(round,att[att_idx],jid,cog,stk,nseat,nstkp,nstkd))
+      nseat <- nseat
+      nstkp <- nstkp - (2-att_idx)*stk
+      nstkd <- nstkd + (1-att_idx)*stk
+      jid1= (stk==0)
+      
+    }else{
+      if(jid1){
+        jid = jid
+        att_idx <- which(att!=att[att_idx])
+        cog = cog
+        if (att_idx==1 && nstkp==0){
+          stk=0
[…]

### 2021-05-08 tests/testthat.R
app tests - initial commit

@@ -1,0 +1,4 @@
+library(testthat)
+library(batson_app)
+
+test_check("batson_app")

### 2021-05-08 tests/testthat/test-sequential.R

@@ -1,0 +1,7 @@
+test_that("multiplication works", {
+  expect_equal(2 * 2, 4)
+})
+
+test_that('no missing values', {
+  expect_identical(a, na.omit(a))
+})

### 2021-05-08 tests/testthat/test-sim_seq.R

@@ -1,0 +1,6 @@
+load_path <- here::here("sim_seq.R")
+source(load_path)
+
+test_that("multiplication works", {
+  expect_equal(2 * 2, 4)
+})

### 2021-05-08 prototype/app.R
commented out the "Update" button

@@ -49,8 +49,8 @@ ui <- fluidPage(
                    rHandsontableOutput("hot")
                 ),
             mainPanel(
-                actionButton("updateButton", "Update",
-                             icon = icon("refresh")),
+                #actionButton("updateButton", "Update",
+                #             icon = icon("refresh")),
                 plotOutput("plot"))
                 )
         ))

### 2021-05-08 prototype/app.R
removed call to library(markdown)

@@ -5,7 +5,6 @@ library(mcmc)
 library(Rcpp)
 library(ggplot2)
 library(shinythemes)
-library(markdown)
 
 Rcpp::sourceCpp(here::here("prototype","mh_sampler_pp.cpp"))
 Rcpp::sourceCpp(here::here("prototype","mh_sampler.cpp"))

### 2021-05-08 app/app.R
removed use of 'here' package to source files

@@ -6,15 +6,15 @@ library(Rcpp)
 library(ggplot2)
 library(shinythemes)
 
-Rcpp::sourceCpp(here::here("prototype","mh_sampler_pp.cpp"))
-Rcpp::sourceCpp(here::here("prototype","mh_sampler.cpp"))
+Rcpp::sourceCpp("mh_sampler_pp.cpp")
+Rcpp::sourceCpp("mh_sampler.cpp")
 
 # Load dummy strike dataframe
-source(here::here("prototype","dummy_strike_data.R"))
+source("dummy_strike_data.R")
 
 # Load calc_prior function
 
-source(here::here("prototype","calc_prior.R"))
+source("calc_prior.R")
 
 # Define Cognitive Class choices
 

### 2021-05-08 app/calc_prior.R

@@ -3,7 +3,7 @@
 library(dplyr)
 library(mcmc)
 library(Rcpp)
-Rcpp::sourceCpp(here::here("prototype","mh_sampler_pp.cpp"))
+Rcpp::sourceCpp("mh_sampler_pp.cpp")
 
 dat0 <- readRDS("jury_data_cleaned_new.RDS")
 

### 2021-05-08 app/dummy_strike_data.R

@@ -12,7 +12,7 @@ party_choices <- c("PP","PD")
 
 #### attorney name choices ####
 
-dat0 <- readRDS(here::here("prototype","jury_data_cleaned_new.rds"))
+dat0 <- readRDS("jury_data_cleaned_new.rds")
 
 atty_levels_p <- dat0 %>%
   pull("P_atty_l") %>%

### 2021-07-16 analysis/paper_draft.Rmd
revised paper draft

--- 2021-02-24 analysis/paper_draft.Rmd
@@
 ---
-title: "Visualizing Bias for Batson Challenges Using Historical Selection Data"
+title: "Batson Challenges Using Historical Selection Data"
 date: "`r Sys.Date()`"
 bibliography: references.bib
-author: "Pandya, Li, Moore"
-output: 
-    html_document:
-    theme: readable
+author:
+  - name: "Sachin S. Pandya"
+    affiliation: "University of Connecticut"
+    orcid_id: 0000-0001-7387-1307
+  - name: "Xiaomeng Li"
+    affiliation: "University of Connecticut"
+  - name: "Timothy E. Moore"
+    affiliation: "University of Connecticut"
+    orcid_id: 0000-0002-9576-0517
+output: distill::distill_article
 ---
@@
 
-This paper demonstrates the use of a power prior as part of a software application that, during jury selection, attorneys can use in real time to estimate the probability that a prosecutor or defense attorney used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v.
+This paper demonstrates a prototype software application that, with both contemporaneous and historical jury-selection data, attorneys can use in real time to estimate the probability that an attorney has used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v.
 Kentucky_ (1986), and its progeny.
 
-# Background
-
-## Power Prior
-
-The power prior is a way to create informative prior distributions based on
-historical data, or more precisely, by controlling how much historical data affects the posterior distribution of the unobserved variables of interest. For a review, see @Ibrahim2015. [MORE]
+# Background and Related Literature
 
@@
 
-In the US, jury selection proceeds in three steps.
-First, a court identifies individuals as part of a qualified jury pool, often based on voter registration and driver's license records.
-Second, for a given case, a subset of individuals in the jury pool are summoned to appear for possible jury service.
-Upon their appearance, this group (the "venire") is subject to "voir dire", that is, they are questioned (by the parties or the trial judge) and are (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise (by one of the parties' attorneys) of a peremptory challenge.
+In the US, once a group of individuals appear in court, this group is questioned by the parties' attorneys and the trial judge.
+In turn, each individual may be (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise by one of the parties' attorneys of a peremptory challenge.
+The key feature of a peremptory challenge is that, with it, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury, but without the burden of explaining why.
+Each side is assigned a number of peremptory challenges and may use them to strike any potential juror without explanation.
+The number of peremptory challenges varies by jurisdiction, as does peremptory-challenge procedure, including the order of strikes and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].
 
-The key feature of a peremptory challenge is that, with it, a party declare a prospective juror ineligible ("strikes" that juror) for a seat on the jury.
-Each side is assigned a number of peremptory challenges and may use them to strike any potential juror without explanation.
-The number of peremptory challenges varies by jurisdiction, as does how the procedure by which peremptory challenges are exercised.
-Where State law prescribes no uniform procedure, that procedure can vary by trial judge in that State.
-In federal court, practices vary by court and judge, because no federal rule specifies the procedure to be used [@Williams2017].
-
+<!--
 Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p.
@@
 For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.
+-->
 
 Under _Batson v.
-Kentucky_ (1986)_, a party's exercise of a peremptory challenge violates the Equal Protection Clause of the US Constitution if motivated by a prospective juror's race, gender, or certain other characteristics.
-To raise a _Batson_ challenge during jury selection, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
+Kentucky_ (1986)_, however, a party's exercise of a peremptory challenge violates the Equal Protection Clause of the US Constitution if motivated by a prospective juror's race, gender, or certain other protected characteristics.
+When raising a _Batson_ challenge during jury selection, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
 Once established, the party exercising the strike bears the burden of producing a nondiscriminatory reason for that strike.
-Thereafter, the challenger bears the burden of persuading the trial judge that the race or gender was the real motivation for the strike in question.^[Some jurisdictions adopt a modified version of _Batson_ that reduces or eliminates the challenger's burden to establish a prima facie case.
-E.g., State v.
+Thereafter, the challenger bears the burden of persuading the trial judge that the race or gender, not the proffered nondiscriminatory reason, was more likely than not the real motivation for the challenged strike(s).
+For discussion, see [@LaFave2020, sec.
+22.3].
+
+By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
+Such information would be relevant, in any particular case, to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in any particular case.^[See, e.g., Miller-El v.
+Dretke, 545 U.S.
+231, 253 (2005); Riley v.
+Taylor, 277 F.3d 261, 280-81, 283 (3d Cir.
+2001); Cochran v.
+Herring, 43 F.3d 1404, 1412 (11th Cir.
+1995); see also Flowers v.
+Mississippi, 139 S.
+Ct.
+2228, 2243 (2019) (kinds of evidence to support Batson challenge includes "relevant history of the State's peremptory strikes in past cases")]
+
+Some States have adopted variants of _Batson_ that reduces or eliminates the challenger's burden to establish a prima facie case^[E.g., State v.
 Holloway, 209 Conn.
@@
 Antwine, 743 S.W.2d 51, 64 (Mo.
-1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]
-
-Evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also historical data on peremptory challenges in past case.
-Though usually revealed after painstaking investigation of jury selection records [e.g., @Craft2018], such information, if routinely available, would be relevant, in any particular case, to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in any particular case.^[See, e.g., Miller-El v.
-Dretke, 545 U.S.
-231, 253 (2005); Riley v.
-Taylor, 277 F.3d 261, 280-81, 283 (3d Cir.
-2001); Cochran v.
-Herring, 43 F.3d 1404, 1412 (11th Cir.
-1995).]
-
-More recently, a few States (e.g., Washington State, California) have provided an independent basis to challenge peremptory strikes.
-For example, by court rule, a trial court in Washington State may deny a peremptory challenge if it determines that "an objective observer could view race or ethnicity as a factor in the use of the peremptory challenge."^[R.
-37(e); see also Rule 37(f) ("an obiective obseryer is aware that implicit, institutional, and unconscious biases, in addition to purposeful discrimination, have resulted in the unfair exclusion of potential jurors in Washington State"); Rule 37(h) (identifying presumptively invalid reasons for a peremptory challenge).].
+1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]; as well as, in a few States, replaced a focus on the motive for strikes with a query into whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in the use of strikes.^[Wash.
+R.
+37(e)].
 
@@
 This pattern persists, despite the fact that both sides' attorneys can often observe the race and gender of those in the venire who were and were not struck by peremptory challenge.
-
 The research literature suggests several explanations.
-These include a reluctance to raise a _Batson_ challenge on the premise that such a challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021, 148, 151, 158].
+These include a reluctance to raise a _Batson_ challenge on the premise that such a challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021].
 Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
-Moreover, drawing inferences from historical data is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018].
+Moreover, drawing inferences from historical data -- the use of perempetory challenges in _past_ cases -- is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking investigation [e.g., @Craft2018].
 
@@
 To do so, one would need to, for any peremptory-selection method, a corresponding likelihood function for how a prosecutor or defense attorney strikes prospective jurors of a protected class (e.g., racial-minorities).
-Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter to jury selection, unless "statisticans accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter, unless "statisticans accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
 
-# Method
+# Discussion
 
-[insert intro]
+Here, we discuss how we developed a software protoype for attorneys to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of perempetory challenges in a particular case, either alone or in combination with historical data on that attorney's pattern of strikes in past cases.
+In doing so, we highlight two distinctive contributions.
 
-# Model 1
+First, we use the power prior to control how much the strike history in past cases affects the posterior distribution of the bias parameter $b$.
+The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of the unobserved variables of interest.
+The key insight is to raise the likelihood function based on the historical data to a power parameter.
+For a review, see @Ibrahim2015.
+
+Second, we implement the analysis as prototype software that, in principle, an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.
+
+# Model
 
@@
 Then, the parties exercise their strikes on those 28 potential jurors.
-
 In a typical variant of the "struck jury" method, the parties strike in an alternating sequence, with the prosecution striking first.
@@
 If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.
-
 Under "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box.
@@
 
-For these peremptory challenges, following @Kadane2018 and @Barrett2007, we model the probability that the struck person belonged (or not) to the cognizable class as a function of (1) the number of cognizable class members (e.g., racial minorities, women) and non-class members (e.g., White people, men) in the set of people who could have been struck; and (2) a parameter $b$ that indicates how much weight (or "bias") the striking party placed on the struck juror's membership in the cognizable class (e.g., whether that juror was a racial minority or White; women or man).
-
+Following @Kadane2018 and @Barrett2007, we model these peremptory-challenge processes as follows.
 For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
@@
 
-Because $b$ is not directly observable, we are interested in estimating the likely values of $b$, given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$.
+Because $b$ is not directly observable, we estimate likely values of $b$, given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$.
 
-Accordingly, we model the jury-selection process as follows.
+Accordingly, we model the strike-process as follows.
 
@@
 If so, our measure of bias is zero: given $w=e^b$, then $e^b=1$ and $b=log(1)=0$.
-If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class ($X=1$, e.g., the juror is a racial minority).
-Where $b<1$, the party has bias toward striking a juror falling outside the cognizable class ($X=0$, e.g., the juror is White).
+If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
+Where $b<0$, the party has a preference for a juror within the cognizable class.
+
+The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes.
+
+Finally, to control how much weight the strike history in past cases has on the posterior distribution of $b$, we revise the likelihood function to include the power prior:
+
+\[L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}\]
+
+\[b \sim \text{Normal}(0,2)\]
+
+<!--
 
@@
 [insert statistical model]
+-->
 
-# The App
+# Simulation
 
-[insert]
+We conduct simulation study under different scenarios, including no bias, bias against cognizable class and preference for cognizable class.
+We have also studied the scenarios with different amount of historical data and whether the bias parameter of historical data is the same as current data.
+We show the model accuracy under different scenarios and how it improves the model accuracy by adding historical data and modifying the weight parameter.
 
-The posterior probability distribution is estimated with R 4.0 [@R] and mcmc 0.9-7 [@mcmc]. The app was built using the shiny package [@shiny].
+In particular, we simulate data with 7 different values of $b = -3,-2,-1,0,1,2,3$ representing strong bias, bias, moderate bias toward striking a juror falling outside the cognizable class, no bias striking a juror and moderate bias, bias and strong bias favoring a strike against a juror falling within the cognizable.
+
+We also generate historical data with bias parameter from the 7 values.
+As for historical data, we consider it from one trial, two trials and three trials to represent different sizes of historical data.
+
+Finally, to let a person's bias parameter may change over time, we allow the historical data generated from a different bias parameter than the current data.
+When we use power prior incorporating the historical information, we can control the weight that we put on historical data by modifying the weight parameter.
+In the simulation study, we also try different values of the weight parameter to show the influence of the weight parameter on modeling accuracy.
+
+Figure [X] depicts the results of these simulations.
+
+```{r pressure, eval = FALSE, echo=FALSE, fig.cap="A caption", out.width = '100%'}
+
+# knitr::include_graphics("dino_plots.pdf")
+
+# INSERT simulation and plot code
+
+```
+
+_Historical b_ means the bias parameter for the generated historical data, $b$ means the bias parameter for the current data.
+The values on the right side represent the size of the historical data.
+The values on the top of the plots represent the weight parameter of the power prior.
+Different color represents the estimated bias parameter.
+
+From the plot, our model can capture strong bias and bias, when the bias parameter being -3,-2,2,3.
+When the bias is moderate, that is when the bias parameter b being -1 and 1, the model can detect it when we have historical data with the same direction bias parameter.
+If the bias parameter of historical data is in an opposite direction, we might get a wrong estimate of current bias parameter.
+Thus if the historical data is not reliable, we should set a small weight parameter on them.
+
+# The App Prototype
+
+The posterior probability distribution is estimated with R 4.0 [@R]. The app was built using the shiny package [@shiny].
 

### 2021-07-20 analysis/paper_draft.Rmd
update method and simulation

@@
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
+If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$.
+If we define the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$ by adding one parameter $w$, we can measure the bias by different values of $w$.
+When $w=1$, it corresponds to the situation when there is no bias.
+When $w>1$, the probability of striking a cognizable class member is greater than non cognizable class, which means a discriminatory against cognizable class.
+On the contrary, when $w<1$, the probability of striking a cognizable class member is smaller than non cognizable class, which means a preference of cognizable class.
+However, under this set up, the weight of non cognizable class is the reciprocal of the weight of cognizable class, which is not a desirable character.
+Using $b = log(w)$ could avoid this issue.
+For any given value of the bias parameter $b$, the probability of strike od a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
 
-Accordingly, for any given value of the bias parameter $b$, the probability of strike, or $Pr(\delta_{ij})$, is:
-
-$$
+\begin{equation}
 Pr(\delta_{ij} | b) = 
@@
   \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
-\end{cases}
-$$
+\end{cases} (\#eq:model1)
+\end{equation}
+
 This is equivalent to
 
-$$Pr(\delta_{ij}| b) = \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$
+$$Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$
+
+Given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$, by estimating the value of $b$, we can measure whether there is bias when striking potential jurors.
 
 If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.
-
-Because $b$ is not directly observable, we estimate likely values of $b$, given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$.
-
-Accordingly, we model the strike-process as follows.
-
-\[L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}\]
-\[b \sim \text{Normal}(0,2)\]
-
-where $n_{i}$ is the total number of jury selections; and $n_j$ is the total number of peremptory strikes.
-
-This way, if it is equally likely that a member and non-member of the cognizable class will be stricken, then $w = \frac{Pr(X=1)=0.50}{Pr(X = 0)=0.50}=1$.
-If so, our measure of bias is zero: given $w=e^b$, then $e^b=1$ and $b=log(1)=0$.
 If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
@@
 
-The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes.
+The likelihood function of b is shown in equation \@ref(eq:model2)
 
-Finally, to control how much weight the strike history in past cases has on the posterior distribution of $b$, we revise the likelihood function to include the power prior:
+\begin{equation}  
+L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}} (\#eq:model2)
+\end{equation}
+<!-- \[b \sim \text{Normal}(0,2)\] -->
 
-\[L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}\]
+where $n_{i}$ is the total number of jury selections; and $n_j$ is the total number of peremptory strikes.
+In @Kadane2018, $b$ is estimated under Bayesian framework with a uniform prior from -6 to 6.
+In this paper, we also would like to incorporate the historical information.
+In addition, we want to allow different weight for historical information, considering people's preference might change over time.
+For instance, we can set a smaller weight if the historical data is collected many years ago than recent information.
+Thus we use power prior to incorporate the historical information [@Ibrahim2000; @Ibrahim2015].
+The formulation of the power prior is shown in equation \@ref(eq:model3)
 
-\[b \sim \text{Normal}(0,2)\]
+\begin{equation}
+\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b) (\#eq:model3)
+\end{equation}
 
+where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information, and $D_0$ is the observed historical data.
+$L(b|D_0)$ is the likelihood function of $b$ given the historical data.
+$\pi_0(b)$ is the initial prior before the historical data is observed.
+In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
+
+<!-- The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes.
+-->
+
+The posterior distribution of $b$ after including the historical information through the power prior is proportional to the product of likelihood function of $b$ and the power prior of $b$ as shown in \@ref(eq:model4)
+
+\begin{equation}
+L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
+\end{equation}
+
+where $L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$ is the likelihood function of $b$ given historical data.
+$n_{0i}$ is the total number of jury selections in the historical data, and $n_{0j}$ is the total number of peremptory strikes in the historical data.
+$\delta_{0ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class" in the historical trials.
+
+The posterior distribution does not have a closed form, thus Metropolis-Hasting algorithm is implemented to sample $b$ from the posterior distribution.
+Traceplots, Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992] were used to measure the convergence.
 <!--
@@
 
-We conduct simulation study under different scenarios, including no bias, bias against cognizable class and preference for cognizable class.
-We have also studied the scenarios with different amount of historical data and whether the bias parameter of historical data is the same as current data.
-We show the model accuracy under different scenarios and how it improves the model accuracy by adding historical data and modifying the weight parameter.
+We conduct simulation study to measure model performances under different scenarios.
+Different scenarios considers different bias and preference for cognizable class member, different amount of historical information and whether the preference changes over time.
+As for different preference for cognizable class, we generate data using seven different values of the bias parameter $b = (-3,-2,-1,0,1,2,3)$.
+As introduced in previous section, $b<0$ represents a preference for a juror within the cognizable class, and $b>0$ represents a bias against the cognizable class.
+Greater absolute value of $b$ represents greater preference or bias for the cognizable class.
+$b=0$ represents no bias.
+As for the amount of historical data, we consider three sizes of historical data, the same size as current data, double size of current data and triple size of current data.
+We also consider seven different values of the bias parameter for generating the historical data and allow the bias parameter of historical data to be different from current data.
+The different value of bias parameter from historical data to current data represents a change of preference for cognizable class.
 
-In particular, we simulate data with 7 different values of $b = -3,-2,-1,0,1,2,3$ representing strong bias, bias, moderate bias toward striking a juror falling outside the cognizable class, no bias striking a juror and moderate bias, bias and strong bias favoring a strike against a juror falling within the cognizable.
+When we use power prior incorporating the historical information, we can control the weight that we put on historical data by modifying the weight parameter $a$.
+In the simulation study, we also try different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show the influence of the weight parameter on modeling accuracy.
+We consider 735 scenarios in total, which is the combination of 7 different bias parameter of current data and of historical data, 3 different amount of historical data and 5 different values of the power prior weight parameter $a$.
+For each scenario, we generate 1000 dataset and fit the model on the 1000 generated dataset.
+We record the mean value of the posterior mean and credible intervals of the 1000 model fits for each scenario.
+As for the credible interval, we tried three significance levels, 80\%, 90\% and 95\%.
+We also record the corresponding coverage rate.
+For instance, the 90\% coverage rate means the proportion of the 1000 model fits that the credible interval shows the true preference as we set.
+That is, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.
+We show the coverage rate under different scenarios in the figures.
 
-We also generate historical data with bias parameter from the 7 values.
-As for historical data, we consider it from one trial, two trials and three trials to represent different sizes of historical data.
+Figure \@ref(fig:b80) depicts the results of significance of estimated $b$ under 80\% credible interval.
+Figure \@ref(fig:b90) shows the results under 90\% credible interval and figure \@ref(fig:b95) shows the results under 95\% credible interval.
 
-Finally, to let a person's bias parameter may change over time, we allow the historical data generated from a different bias parameter than the current data.
-When we use power prior incorporating the historical information, we can control the weight that we put on historical data by modifying the weight parameter.
-In the simulation study, we also try different values of the weight parameter to show the influence of the weight parameter on modeling accuracy.
+```{r b80,  echo=FALSE, fig.cap="80% significance of bias parameter", out.width = '100%'}
+library(ggplot2)
+# knitr::include_graphics("dino_plots.pdf")
 
-Figure [X] depicts the results of these simulations.
+# INSERT simulation and plot code
+sims <- read.csv("sim_resultc.csv")
+sims$a.f <- as.factor(sims$a)
+sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
+sims$sig <- ifelse(sims$L80 >0 & sims$U80 >0 , 
+                                    "pos", sims$sig) 
+sims$sig <- ifelse(sims$L80 <0 & sims$U80 <0 , 
+                                    "neg", sims$sig) 
+sims$sig.n <- 0
+sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
+sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
 
-```{r pressure, eval = FALSE, echo=FALSE, fig.cap="A caption", out.width = '100%'}
+p = ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                        fill = as.factor(sig.n)),
+            color="black")+
+  facet_grid(h~a)+
+  # scale_fill_gradient2(expression(hat(d)), 
+  #                      low = "darkred", mid = "white", 
+  #                      high = "blue", 
+  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
+  #                      limits = c(-3.2, 3.2),
+  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  scale_fill_manual(values = c("white", "blue"))+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+p
+
+```
+
+```{r b90,  echo=FALSE, fig.cap="90% significance of bias parameter", out.width = '100%'}
 
@@
 # INSERT simulation and plot code
+sims$a.f <- as.factor(sims$a)
+sims$sig <- ifelse(sims$L90 <0 & sims$U90 >0 , "ns", 1) 
+sims$sig <- ifelse(sims$L90 >0 & sims$U90 >0 , 
+                                    "pos", sims$sig) 
+sims$sig <- ifelse(sims$L90 <0 & sims$U90 <0 , 
+                                    "neg", sims$sig) 
+sims$sig.n <- 0
+sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
+sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
 
+p = ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                        fill = as.factor(sig.n)),
+            color="black")+
+  facet_grid(h~a)+
+  # scale_fill_gradient2(expression(hat(d)), 
+  #                      low = "darkred", mid = "white", 
+  #                      high = "blue", 
+  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
+  #                      limits = c(-3.2, 3.2),
+  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  scale_fill_manual(values = c("white", "blue"))+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+p
 ```
 
-_Historical b_ means the bias parameter for the generated historical data, $b$ means the bias parameter for the current data.
+```{r b95,  echo=FALSE, fig.cap="95% significance of bias parameter", out.width = '100%'}
+
+# knitr::include_graphics("dino_plots.pdf")
+
+# INSERT simulation and plot code
+sims$a.f <- as.factor(sims$a)
+sims$sig <- ifelse(sims$L95 <0 & sims$U95 >0 , "ns", 1) 
+sims$sig <- ifelse(sims$L95 >0 & sims$U95 >0 , 
+                                    "pos", sims$sig) 
+sims$sig <- ifelse(sims$L95 <0 & sims$U95 <0 , 
+                                    "neg", sims$sig) 
+sims$sig.n <- 0
+sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
+sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
+
+p = ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                        fill = as.factor(sig.n)),
+            color="black")+
+  facet_grid(h~a)+
+  # scale_fill_gradient2(expression(hat(d)), 
+  #                      low = "darkred", mid = "white", 
+  #                      high = "blue", 
+  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
+  #                      limits = c(-3.2, 3.2),
+  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  scale_fill_manual(values = c("white", "blue"))+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+p
+```
+
+_Historical b_ means the bias parameter for the generated historical data, current $b$ means the bias parameter for the current data.
 The values on the right side represent the size of the historical data.
 The values on the top of the plots represent the weight parameter of the power prior.
-Different color represents the estimated bias parameter.
+Blue means the credible interval of $b$ does not include zero, that is $b$ is significant.
 

### 2021-07-20 analysis/paper_draft.Rmd
minor edits, spelling

@@
 Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
-Moreover, drawing inferences from historical data -- the use of perempetory challenges in _past_ cases -- is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking investigation [e.g., @Craft2018].
+Moreover, drawing inferences from historical data -- the use of peremptory challenges in _past_ cases -- is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking investigation [e.g., @Craft2018].
 
@@
 To do so, one would need to, for any peremptory-selection method, a corresponding likelihood function for how a prosecutor or defense attorney strikes prospective jurors of a protected class (e.g., racial-minorities).
-Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter, unless "statisticans accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
 
@@
 
-Here, we discuss how we developed a software protoype for attorneys to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of perempetory challenges in a particular case, either alone or in combination with historical data on that attorney's pattern of strikes in past cases.
+Here, we discuss how we developed a software prototype for attorneys to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of peremptory challenges in a particular case, either alone or in combination with historical data on that attorney's pattern of strikes in past cases.
 In doing so, we highlight two distinctive contributions.
@@
 Then, the parties exercise their strikes on those 28 potential jurors.
+
 In a typical variant of the "struck jury" method, the parties strike in an alternating sequence, with the prosecution striking first.
@@
 Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.
-Once that is completed, new potential jurors take the seats of the excused or striken jurors, and the process repeats until all the jury seats are filled.
+Once that is completed, new potential jurors take the seats of the excused or stricken jurors, and the process repeats until all the jury seats are filled.
 At any point, strikes can be used on anyone in the jury box, including those there previously but not struck.
@@
 If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
+
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
+
 If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$.
@@
 On the contrary, when $w<1$, the probability of striking a cognizable class member is smaller than non cognizable class, which means a preference of cognizable class.
+
 However, under this set up, the weight of non cognizable class is the reciprocal of the weight of cognizable class, which is not a desirable character.
 Using $b = log(w)$ could avoid this issue.
-For any given value of the bias parameter $b$, the probability of strike od a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
+For any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
 
@@
 We consider 735 scenarios in total, which is the combination of 7 different bias parameter of current data and of historical data, 3 different amount of historical data and 5 different values of the power prior weight parameter $a$.
-For each scenario, we generate 1000 dataset and fit the model on the 1000 generated dataset.
+For each scenario, we generate 1000 datasets and fit the model on the 1000 generated dataset.
 We record the mean value of the posterior mean and credible intervals of the 1000 model fits for each scenario.
-As for the credible interval, we tried three significance levels, 80\%, 90\% and 95\%.
-We also record the corresponding coverage rate.
+
+For the credible interval, we set three significance levels (80\%, 90\% and 95\%) and recorded the corresponding coverage rate.
 For instance, the 90\% coverage rate means the proportion of the 1000 model fits that the credible interval shows the true preference as we set.
 That is, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.
-We show the coverage rate under different scenarios in the figures.
 
+The following figures depict the coverage rate under different scenarios.
 Figure \@ref(fig:b80) depicts the results of significance of estimated $b$ under 80\% credible interval.
@@
 ```{r b80,  echo=FALSE, fig.cap="80% significance of bias parameter", out.width = '100%'}
+
 library(ggplot2)
-# knitr::include_graphics("dino_plots.pdf")
 
 # INSERT simulation and plot code
-sims <- read.csv("sim_resultc.csv")
+
+load_path <- here::here("sim_resultc.csv")
+
+sims <- readr::read_csv(load_path)
 sims$a.f <- as.factor(sims$a)
@@
 
-# knitr::include_graphics("dino_plots.pdf")
-
 # INSERT simulation and plot code
@@
 ```{r b95,  echo=FALSE, fig.cap="95% significance of bias parameter", out.width = '100%'}
-
-# knitr::include_graphics("dino_plots.pdf")
 
@@
 From the plot, our model can capture strong bias and bias, when the bias parameter being -3,-2,2,3.
-When the bias is moderate, that is when the bias parameter b being -1 and 1, the model can detect it when we have historical data with the same direction bias parameter.
+When bias is moderate ($b = \{1, -1\}$), the model can detect it when we have historical data with the same direction bias parameter.
 If the bias parameter of historical data is in an opposite direction, we might get a wrong estimate of current bias parameter.
-Thus if the historical data is not reliable, we should set a small weight parameter on them.
+
+Thus if the historical data is not reliable, we should set a small weight parameter on them. For example, [insert]
 

### 2021-07-26 analysis/paper_draft.Rmd
added motivation for power prior; minor edits

@@
 
-This paper demonstrates a prototype software application that, with both contemporaneous and historical jury-selection data, attorneys can use in real time to estimate the probability that an attorney has used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v.
+This paper demonstrates how to use contemporaneous and historical jury-selection data to estimate the probability that an attorney has used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v.
 Kentucky_ (1986), and its progeny.
@@
 
-In the US, once a group of individuals appear in court, this group is questioned by the parties' attorneys and the trial judge.
-In turn, each individual may be (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise by one of the parties' attorneys of a peremptory challenge.
+In the US, once a group of individuals selected for jury-service appear in court as scheduled, this group is questioned by the parties' attorneys and the trial judge.
+During this "voir dire" process, each prospective juror may be (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise by one of the parties' attorneys of a peremptory challenge.
 The key feature of a peremptory challenge is that, with it, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury, but without the burden of explaining why.
@@
 When raising a _Batson_ challenge during jury selection, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
-Once established, the party exercising the strike bears the burden of producing a nondiscriminatory reason for that strike.
+Once established, the challenged party bears the burden of producing one or more nondiscriminatory reasons for that strike.
 Thereafter, the challenger bears the burden of persuading the trial judge that the race or gender, not the proffered nondiscriminatory reason, was more likely than not the real motivation for the challenged strike(s).
@@
 22.3].
+Some States have adopted variants of _Batson_ that reduce or eliminates the challenger's burden to establish a prima facie case^[E.g., State v.
+Holloway, 209 Conn.
+636, 646 & n.4 (1989); State v.
+Chapman, 317 S.C.
+302, 305-06 (1995); see also State v.
+Antwine, 743 S.W.2d 51, 64 (Mo.
+1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]; and, more recently, replaced a focus on the motive for strikes with a query into whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in the use of strikes.^[Wash.
+R.
+37(e)].
 
 By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
-Such information would be relevant, in any particular case, to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in any particular case.^[See, e.g., Miller-El v.
+For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case.^[See, e.g., Miller-El v.
 Dretke, 545 U.S.
@@
 2228, 2243 (2019) (kinds of evidence to support Batson challenge includes "relevant history of the State's peremptory strikes in past cases")]
-
-Some States have adopted variants of _Batson_ that reduces or eliminates the challenger's burden to establish a prima facie case^[E.g., State v.
-Holloway, 209 Conn.
-636, 646 & n.4 (1989); State v.
-Chapman, 317 S.C.
-302, 305-06 (1995); see also State v.
-Antwine, 743 S.W.2d 51, 64 (Mo.
-1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]; as well as, in a few States, replaced a focus on the motive for strikes with a query into whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in the use of strikes.^[Wash.
-R.
-37(e)].
 
@@
 
-Here, we discuss how we developed a software prototype for attorneys to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of peremptory challenges in a particular case, either alone or in combination with historical data on that attorney's pattern of strikes in past cases.
-In doing so, we highlight two distinctive contributions.
+Here, we discuss how to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of peremptory challenges in a particular case either alone or in combination with historical data on that attorney's pattern of strikes in past cases.
+In doing so, we make two distinctive contributions.
 
@@
 
-Second, we implement the analysis as prototype software that, in principle, an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.
+Second, we implement the analysis as prototype software that an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.
 
@@
 
-However, under this set up, the weight of non cognizable class is the reciprocal of the weight of cognizable class, which is not a desirable character.
+However, under this set up, the weight of the non-cognizable class is the reciprocal of the weight of cognizable class, which is not a desirable character.
 Using $b = log(w)$ could avoid this issue.
@@
 
-The posterior distribution does not have a closed form, thus Metropolis-Hasting algorithm is implemented to sample $b$ from the posterior distribution.
+The posterior distribution does not have a closed form.
+Accordingly, we used the Metropolis-Hasting algorithm sample $b$ from the posterior distribution.
 Traceplots, Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992] were used to measure the convergence.
@@
 
-We conduct simulation study to measure model performances under different scenarios.
-Different scenarios considers different bias and preference for cognizable class member, different amount of historical information and whether the preference changes over time.
-As for different preference for cognizable class, we generate data using seven different values of the bias parameter $b = (-3,-2,-1,0,1,2,3)$.
-As introduced in previous section, $b<0$ represents a preference for a juror within the cognizable class, and $b>0$ represents a bias against the cognizable class.
-Greater absolute value of $b$ represents greater preference or bias for the cognizable class.
-$b=0$ represents no bias.
-As for the amount of historical data, we consider three sizes of historical data, the same size as current data, double size of current data and triple size of current data.
-We also consider seven different values of the bias parameter for generating the historical data and allow the bias parameter of historical data to be different from current data.
+We conducted a simulation study to measure model performance under different values of the bias parameter, different amount of historical information and whether the bias/preference changes over time.
+To do this, we generated data using seven different values of the bias parameter $b = {-3,-2,-1,0,1,2,3}$.
+As introduced above, $b<0$ represents a preference for a juror within the cognizable class, and $b>0$ represents a bias against the cognizable class.
+A greater absolute value of $b$ represents greater preference or bias for the cognizable class, while $b=0$ represents no bias.
+We also considered three sizes of historical data, the same size as current data, double size of current data and triple size of current data.
+We also considered seven different values of the bias parameter for generating the historical data and allow the bias parameter of historical data to be different from current data.
 The different value of bias parameter from historical data to current data represents a change of preference for cognizable class.
 
-When we use power prior incorporating the historical information, we can control the weight that we put on historical data by modifying the weight parameter $a$.
-In the simulation study, we also try different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show the influence of the weight parameter on modeling accuracy.
-We consider 735 scenarios in total, which is the combination of 7 different bias parameter of current data and of historical data, 3 different amount of historical data and 5 different values of the power prior weight parameter $a$.
+By using the power prior to incorporate the historical information, we can control the weight that we put on historical data by modifying the weight parameter $a$.
+We may wish to exercise such control and $a < 1$ if some of the historical data is missing not at random.
+For example, for any given attorney, there may be only incomplete information available as to that attorney's exercise of peremptory challenges in previous trials.
+Possible reasons include limited access to the paper records containing such information, either due to rules governing who may access such records or the expense incurred in obtaining copies of such records.
+Standard jury-selection forms may not always require recording relevant information about prospective jurors (e.g., their race or gender), which may force researchers to combine such records with other data sources (e.g., voter registration rolls) by matching by name.
+And even if standard forms or other recordkeeping rules require collecting all relevant information for each past trial, judges, attorneys, and court staff can vary in how much or how often they actually fill in all the required information.
+
+Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that weight parameter affected modeling accuracy.
+We consider 735 scenarios in total, i.e., the combination of 7 different bias parameter of current data and of historical data, 3 different amount of historical data and 5 different values of the power prior weight parameter $a$.
 For each scenario, we generate 1000 datasets and fit the model on the 1000 generated dataset.
-We record the mean value of the posterior mean and credible intervals of the 1000 model fits for each scenario.
+We record the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
 
@@
 
-Thus if the historical data is not reliable, we should set a small weight parameter on them. For example, [insert]
-
 # The App Prototype
 
-The posterior probability distribution is estimated with R 4.0 [@R]. The app was built using the shiny package [@shiny].
+The posterior probability distribution is estimated with `r R.version$version.string'[@R]. The app was built using the shiny package [@shiny].
 

### 2021-08-02 analysis/paper_draft.Rmd
updated section on prior literature; edits to main text

@@
 ---
-title: "Batson Challenges Using Historical Selection Data"
+title: "Batson Challenges Using Historical Data"
 date: "`r Sys.Date()`"
@@
 
-This paper demonstrates how to use contemporaneous and historical jury-selection data to estimate the probability that an attorney has used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v.
+This paper shows how to use historical jury-selection data to estimate how likely an attorney has used peremptory challenges to strike potential jurors because of juror race or gender in violation of _Batson v.
 Kentucky_ (1986), and its progeny.
+
+In the US, once a group of individuals selected for jury-service appear in court as scheduled, this group is questioned by the parties' attorneys and the trial judge.
+During this "voir dire" process, a prospective juror, if not selected for the jury or excused by the trial judge for cause, may be dismissed if a party's attorney uses a peremptory challenge against them.
 
@@
 
-In the US, once a group of individuals selected for jury-service appear in court as scheduled, this group is questioned by the parties' attorneys and the trial judge.
-During this "voir dire" process, each prospective juror may be (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise by one of the parties' attorneys of a peremptory challenge.
-The key feature of a peremptory challenge is that, with it, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury, but without the burden of explaining why.
-Each side is assigned a number of peremptory challenges and may use them to strike any potential juror without explanation.
-The number of peremptory challenges varies by jurisdiction, as does peremptory-challenge procedure, including the order of strikes and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].
+With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
+While countries such as the United Kingdom and Canada have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1); S.C.
+2019, c.
+25, s 269), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases.
+In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].
+
+Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
+
+Then, the parties exercise their strikes on those potential jurors.
+In one procedure, the parties submit their strikes simultaneously.
+In another variant, the parties strike in an alternating sequence, with the prosecution striking first.
+Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn.
+If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.
+
+Under a "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box.
+Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.
+Once that is completed, new potential jurors take the seats of the excused or stricken jurors, and the process repeats until all the jury seats are filled.
+At any point, strikes can be used on anyone in the jury box.
 
 <!--
+
+Note: Constitutionality of Canada's abolition of strikes upheld in https://decisions.scc-csc.ca/scc-csc/scc-csc/en/item/18932/index.do
+
 Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p.
@@
 For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.
+
 -->
 
-Under _Batson v.
-Kentucky_ (1986)_, however, a party's exercise of a peremptory challenge violates the Equal Protection Clause of the US Constitution if motivated by a prospective juror's race, gender, or certain other protected characteristics.
-When raising a _Batson_ challenge during jury selection, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
-Once established, the challenged party bears the burden of producing one or more nondiscriminatory reasons for that strike.
-Thereafter, the challenger bears the burden of persuading the trial judge that the race or gender, not the proffered nondiscriminatory reason, was more likely than not the real motivation for the challenged strike(s).
-For discussion, see [@LaFave2020, sec.
+This paper focuses on _Batson_ challenges to a party's use of peremptory challenge.
+Under the US Supreme Court's decision in _Batson v.
+Kentucky_ (1986)_, a party's exercise of a peremptory challenge against one or more prosepctive jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
+
+For a _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
+If so, the challenged party must produce one or more nondiscriminatory reasons for that strike.
+Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
 22.3].
-Some States have adopted variants of _Batson_ that reduce or eliminates the challenger's burden to establish a prima facie case^[E.g., State v.
+For similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v.
 Holloway, 209 Conn.
@@
 Chapman, 317 S.C.
-302, 305-06 (1995); see also State v.
-Antwine, 743 S.W.2d 51, 64 (Mo.
-1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]; and, more recently, replaced a focus on the motive for strikes with a query into whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in the use of strikes.^[Wash.
+302, 305-06 (1995)], while other States have shifted the inquiry away from the striking party's motives to whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash.
 R.
@@
 
-By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
-For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case.^[See, e.g., Miller-El v.
+Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
+Possible explanations include attorney fear that raising a _Batson_ challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021].
+Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
+
+By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on the pattern of peremptory challenges in past cases.
+For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v.
 Dretke, 545 U.S.
@@
 Herring, 43 F.3d 1404, 1412 (11th Cir.
-1995); see also Flowers v.
+1995)].
+The US Supreme Court recently identified the kinds of evidence to support _Batson_ challenge as including "relevant history of the State's peremptory strikes in past cases" [e.g, Flowers v.
 Mississippi, 139 S.
 Ct.
-2228, 2243 (2019) (kinds of evidence to support Batson challenge includes "relevant history of the State's peremptory strikes in past cases")]
+2228, 2243 (2019)].
 
-Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
-This pattern persists, despite the fact that both sides' attorneys can often observe the race and gender of those in the venire who were and were not struck by peremptory challenge.
-The research literature suggests several explanations.
-These include a reluctance to raise a _Batson_ challenge on the premise that such a challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021].
-Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
-Moreover, drawing inferences from historical data -- the use of peremptory challenges in _past_ cases -- is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking investigation [e.g., @Craft2018].
+Prior work on statistical methods for _Batson_ challenges, however, have largely focused on strikes used in a single case taken alone.
+For example, @Barrett2007 modeled bias detection for _Batson_ challenges, given a "struck-jury" procedure in which the parties took turns striking prospective jurors.
+@Gastwirth2016 applied Fisher's exact test to detect bias in the pattern of strikes in each of ten published cases.
+For each of three cases, @Kadane2018 compared various frequentist approaches (including Fisher's exact test) with a Bayesian approach.
 
-In theory, however, were an attorney's strike history in past cases easily available, such strike history can be compared to what we would expect to happen if, following the same procedure, venire members were struck at random [@Kadane2018; @Kadane2018a].
-To do so, one would need to, for any peremptory-selection method, a corresponding likelihood function for how a prosecutor or defense attorney strikes prospective jurors of a protected class (e.g., racial-minorities).
-Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+When considering strikes in multiple trials in the same State or judicial district, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
+Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
+
+Moreover, drawing inferences from peremptory challenges in past cases is hard, because jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking efforts to collect [e.g., @Craft2018].
+
+This may matter for how much weight to assign such historical data if data on the use of strikes in past trials is missing not at random.
+For example, for any given attorney, there may be only incomplete information on that attorney's use of strikes in previous trials in a particular State or court.
+Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies of them.
+Moreover, because some standard jury-selection forms do not require recording relevant juror information, such as juror race or gender, lawyers and researchers must turn to other data sources (e.g., voter registration rolls) with that information.
+Finally, regardless of what is supposed to be recorded, judges and court staff vary in how often they actually fill in standard forms with the required information.
+
+Finally, even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
 
@@
 
-Here, we discuss how to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of peremptory challenges in a particular case either alone or in combination with historical data on that attorney's pattern of strikes in past cases.
-In doing so, we make two distinctive contributions.
+Here, we extend the Bayesian approach to _Batson_ challenges to assess whether one can credibly infer illegal bias based solely on an attorney's pattern of peremptory challenges using both present strike data and historical data that attorney's strikes in past cases.
 
+In short, our workflow is this: First, specify a model of the peremptory-strike process that best approximates the peremptory-strike process in the jurisdiction of interest.
+Second, assign a prior distribution to the bias parameter $b$ that reflects the legal requirement that the _Batson_ challenger bear the burden of proof.
+Here, we assume a normal prior for $b$ with mean 0 and standard deviation 2.
+Third, using the strike data in past trials generated by the same peremptory-strike process, estimate a posterior distribution for $b$.
+Finally, using data on the target attorney's strikes in the present case, use that posterior distribution as the prior for $b$ and update.
+
+In so doing, we make two main contributions.
 First, we use the power prior to control how much the strike history in past cases affects the posterior distribution of the bias parameter $b$.
 The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of the unobserved variables of interest.
-The key insight is to raise the likelihood function based on the historical data to a power parameter.
-For a review, see @Ibrahim2015.
+The key insight is to raise the likelihood function based on the historical data to a power parameter [@Ibrahim2000; @Ibrahim2015].
 
@@
 
-Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
-Accordingly, in a non-capital felony trial in federal district court using this method, the trial judge will, after excusing jurors for cause, make sure that there are 28 potential jurors remaining before the parties exercise their peremptory challenges (12 jurors plus 16 peremptories, assuming no alternate jurors).
-Then, the parties exercise their strikes on those 28 potential jurors.
+Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process as follows.
 
-In a typical variant of the "struck jury" method, the parties strike in an alternating sequence, with the prosecution striking first.
-Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn.
-If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.
-Under "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box.
-Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.
-Once that is completed, new potential jurors take the seats of the excused or stricken jurors, and the process repeats until all the jury seats are filled.
-At any point, strikes can be used on anyone in the jury box, including those there previously but not struck.
-
-Following @Kadane2018 and @Barrett2007, we model these peremptory-challenge processes as follows.
 For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
@@
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
+If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$.
 
-If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$.
-If we define the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$ by adding one parameter $w$, we can measure the bias by different values of $w$.
-When $w=1$, it corresponds to the situation when there is no bias.
-When $w>1$, the probability of striking a cognizable class member is greater than non cognizable class, which means a discriminatory against cognizable class.
-On the contrary, when $w<1$, the probability of striking a cognizable class member is smaller than non cognizable class, which means a preference of cognizable class.
-
-However, under this set up, the weight of the non-cognizable class is the reciprocal of the weight of cognizable class, which is not a desirable character.
-Using $b = log(w)$ could avoid this issue.
-For any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
+By adding one parameter $w$, we can measure the bias by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
+As a result, however, the weight of the non-cognizable class is the reciprocal of the weight of cognizable class.
+Using $b = log(w)$ avoids this issue.
+Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
 
@@
 
-Given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$, by estimating the value of $b$, we can measure whether there is bias when striking potential jurors.
-
+Given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$, by estimating the value of $b$, we can measure bias when a party is striking potential jurors.
 If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.
@@
 
-The likelihood function of b is shown in equation \@ref(eq:model2)
+The likelihood function of $b$ is shown in equation \@ref(eq:model2)
 
@@
 
-where $n_{i}$ is the total number of jury selections; and $n_j$ is the total number of peremptory strikes.
+where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
 In @Kadane2018, $b$ is estimated under Bayesian framework with a uniform prior from -6 to 6.
-In this paper, we also would like to incorporate the historical information.
-In addition, we want to allow different weight for historical information, considering people's preference might change over time.
-For instance, we can set a smaller weight if the historical data is collected many years ago than recent information.
-Thus we use power prior to incorporate the historical information [@Ibrahim2000; @Ibrahim2015].
-The formulation of the power prior is shown in equation \@ref(eq:model3)
+
+In this paper, we also would like to incorporate the historical data on strikes in past cases and be able to adjust the weight of that historical information on the posterior distribution of the bias parameter.
+Accordingly, we introduce the power prior \@ref(eq:model3)
 
@@
 
-where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information, and $D_0$ is the observed historical data.
-$L(b|D_0)$ is the likelihood function of $b$ given the historical data.
-$\pi_0(b)$ is the initial prior before the historical data is observed.
+where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
@@
 
-The posterior distribution of $b$ after including the historical information through the power prior is proportional to the product of likelihood function of $b$ and the power prior of $b$ as shown in \@ref(eq:model4)
+After including the historical information through the power prior, the posterior distribution of $b$  is proportional to the product of likelihood function of $b$ and the power prior of $b$ as shown in \@ref(eq:model4)
 
@@
 The posterior distribution does not have a closed form.
-Accordingly, we used the Metropolis-Hasting algorithm sample $b$ from the posterior distribution.
+Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
 Traceplots, Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992] were used to measure the convergence.
+
 <!--
@@
 We conducted a simulation study to measure model performance under different values of the bias parameter, different amount of historical information and whether the bias/preference changes over time.
-To do this, we generated data using seven different values of the bias parameter $b = {-3,-2,-1,0,1,2,3}$.
-As introduced above, $b<0$ represents a preference for a juror within the cognizable class, and $b>0$ represents a bias against the cognizable class.
+
+To do this, we generated strike data for current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$.
+As introduced above, $b<0$ represents a preference for a juror within the cognizable class; and $b>0$ represents a bias against the cognizable class.
 A greater absolute value of $b$ represents greater preference or bias for the cognizable class, while $b=0$ represents no bias.
-We also considered three sizes of historical data, the same size as current data, double size of current data and triple size of current data.
-We also considered seven different values of the bias parameter for generating the historical data and allow the bias parameter of historical data to be different from current data.
-The different value of bias parameter from historical data to current data represents a change of preference for cognizable class.
 
-By using the power prior to incorporate the historical information, we can control the weight that we put on historical data by modifying the weight parameter $a$.
-We may wish to exercise such control and $a < 1$ if some of the historical data is missing not at random.
-For example, for any given attorney, there may be only incomplete information available as to that attorney's exercise of peremptory challenges in previous trials.
-Possible reasons include limited access to the paper records containing such information, either due to rules governing who may access such records or the expense incurred in obtaining copies of such records.
-Standard jury-selection forms may not always require recording relevant information about prospective jurors (e.g., their race or gender), which may force researchers to combine such records with other data sources (e.g., voter registration rolls) by matching by name.
-And even if standard forms or other recordkeeping rules require collecting all relevant information for each past trial, judges, attorneys, and court staff can vary in how much or how often they actually fill in all the required information.
+We also generated historical strike data of three sizes (same, double, and triple the size of current data) as well as seven different values of the bias parameter for generating the historical data ("historical" $b = {-3,-2,-1,0,1,2,3}$).
 
-Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that weight parameter affected modeling accuracy.
-We consider 735 scenarios in total, i.e., the combination of 7 different bias parameter of current data and of historical data, 3 different amount of historical data and 5 different values of the power prior weight parameter $a$.
-For each scenario, we generate 1000 datasets and fit the model on the 1000 generated dataset.
-We record the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
+We let historical $b$ differ from current $b$ (our parameter of interest) for two reasons.
+First, it allows us to simulate noise associated with historical strike data that is incomplete not at random.
+Second, it could also simulate a real difference in the bias parameter that depends on a feature that the current trial has but that _all_ the past trials in the historical data do not.
+To illustrate, assume a prosecutor with two true bias values: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
+If our historical data contains only strike data in cases with non-Black defendants, then that prosecutor's historical $b$ will differ from the current case with a Black defendant.
 
-For the credible interval, we set three significance levels (80\%, 90\% and 95\%) and recorded the corresponding coverage rate.
-For instance, the 90\% coverage rate means the proportion of the 1000 model fits that the credible interval shows the true preference as we set.
-That is, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.
+By using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
+If $a = 1$, the historical strike data is equally weighted with the present data on strikes.
+If $a < 1$, the historical data matters proportionally less.
+As discussed above, if some of the historical data is missing not at random, we may wish to assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter to the historical data on strikes.
+Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that parameter affected modeling accuracy.
 
-The following figures depict the coverage rate under different scenarios.
-Figure \@ref(fig:b80) depicts the results of significance of estimated $b$ under 80\% credible interval.
-Figure \@ref(fig:b90) shows the results under 90\% credible interval and figure \@ref(fig:b95) shows the results under 95\% credible interval.
+We considered 735 scenarios in total, i.e., the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data and 5 different values of the power prior weight parameter $a$.
+For each scenario, we generated 1000 datasets and fit the model on those 1000 dataset.
+We recorded the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
 
-```{r b80,  echo=FALSE, fig.cap="80% significance of bias parameter", out.width = '100%'}
+In turn, we set three credible interval level (80\%, 90\% and 95\%) and recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$.
+For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$.
+Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.
+
+The following figures depict the coverage rates of $b$ under the 80\% credible interval (Figure \@ref(fig:b80)), under the 90\% credible interval (Figure \@ref(fig:b90)), and under the 95\% credible interval (Figure \@ref(fig:b95)).
+
+```{r b80,  echo=FALSE, fig.cap="80% credible interval", out.width = '100%'}
 
@@
 
-```{r b90,  echo=FALSE, fig.cap="90% significance of bias parameter", out.width = '100%'}
+```{r b90, echo=FALSE, fig.cap="90% credible interval"}
 
@@
 
-```{r b95,  echo=FALSE, fig.cap="95% significance of bias parameter", out.width = '100%'}
+```{r b95,  echo=FALSE, fig.cap="95% credible interval"}
 
@@
 
-_Historical b_ means the bias parameter for the generated historical data, current $b$ means the bias parameter for the current data.
+In these figures, _historical b_ denotes the bias parameter for the generated historical data, and current $b$ means the bias parameter for the current data.
 The values on the right side represent the size of the historical data.
-The values on the top of the plots represent the weight parameter of the power prior.
-Blue means the credible interval of $b$ does not include zero, that is $b$ is significant.
+The values on the top of the plots represent the power prior weight parameter $a$.
+Blue means the credible interval for $b$ excludes zero.
 
-From the plot, our model can capture strong bias and bias, when the bias parameter being -3,-2,2,3.
-When bias is moderate ($b = \{1, -1\}$), the model can detect it when we have historical data with the same direction bias parameter.
-If the bias parameter of historical data is in an opposite direction, we might get a wrong estimate of current bias parameter.
+From the plot, our model can detect strong bias and bias, when the bias parameter being -3,-2,2,3.
+When bias is moderate ($b = \{1, -1\}$), the model can detect it if the historical data has a bias parameter in the same direction.
+If the bias parameter of historical data is in an opposite direction, the model may provide a wrong estimate of current bias parameter.
 

### 2021-08-02 analysis/paper_draft.Rmd
revise section on historical data missingness

@@
 
-Prior work on statistical methods for _Batson_ challenges, however, have largely focused on strikes used in a single case taken alone.
-For example, @Barrett2007 modeled bias detection for _Batson_ challenges, given a "struck-jury" procedure in which the parties took turns striking prospective jurors.
-@Gastwirth2016 applied Fisher's exact test to detect bias in the pattern of strikes in each of ten published cases.
-For each of three cases, @Kadane2018 compared various frequentist approaches (including Fisher's exact test) with a Bayesian approach.
-
-When considering strikes in multiple trials in the same State or judicial district, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
+Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strikes in a single case [e.g.
+@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a].
+When considering strike data in multiple trials from the same State or judicial district, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
 Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
 
-Moreover, drawing inferences from peremptory challenges in past cases is hard, because jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking efforts to collect [e.g., @Craft2018].
-
-This may matter for how much weight to assign such historical data if data on the use of strikes in past trials is missing not at random.
-For example, for any given attorney, there may be only incomplete information on that attorney's use of strikes in previous trials in a particular State or court.
-Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies of them.
-Moreover, because some standard jury-selection forms do not require recording relevant juror information, such as juror race or gender, lawyers and researchers must turn to other data sources (e.g., voter registration rolls) with that information.
-Finally, regardless of what is supposed to be recorded, judges and court staff vary in how often they actually fill in standard forms with the required information.
+Part of the issue is that strike data from past cases is hard to collect.
+Jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking efforts to collect [e.g., @Craft2018].
+As a result, even where there, for any given attorney, available strike data on past trials, that data may be missing strikes from trials not at random.
+Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
+Some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck, such as their race or gender.
+As a result, lawyers and researchers must expend time and effort to access other data sources (e.g., voter registration rolls) with that information.
+Finally, regardless of record-keeping policy, judges and court staff vary in how often they fully complete or maintain the required information.
 

### 2021-08-03 analysis/paper_draft.Rmd
updated intro text

@@
 ---
-title: "Batson Challenges Using Historical Data"
+title: "Batson Challenges Using Historical Strike Data"
 date: "`r Sys.Date()`"
@@
 
-This paper shows how to use historical jury-selection data to estimate how likely an attorney has used peremptory challenges to strike potential jurors because of juror race or gender in violation of _Batson v.
+This paper shows how to use historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely an attorney has used peremptory challenges because of prospective juror race or gender in violation of _Batson v.
 Kentucky_ (1986), and its progeny.
 
-In the US, once a group of individuals selected for jury-service appear in court as scheduled, this group is questioned by the parties' attorneys and the trial judge.
-During this "voir dire" process, a prospective juror, if not selected for the jury or excused by the trial judge for cause, may be dismissed if a party's attorney uses a peremptory challenge against them.
+In the US, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
+During this "voir dire" process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed (declared ineligible for a seat on the jury) if a party's attorney uses a peremptory challenge against them.
 
@@
 
-By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on the pattern of peremptory challenges in past cases.
+By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
 For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v.
@@
 1995)].
-The US Supreme Court recently identified the kinds of evidence to support _Batson_ challenge as including "relevant history of the State's peremptory strikes in past cases" [e.g, Flowers v.
+The US Supreme Court recently identified evidence for a _Batson_ challenge as including the "relevant history of the State's peremptory strikes in past cases" [Flowers v.
 Mississippi, 139 S.
@@
 
-Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strikes in a single case [e.g.
-@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a].
-When considering strike data in multiple trials from the same State or judicial district, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
-Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
-
-Part of the issue is that strike data from past cases is hard to collect.
+Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
+@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because strike data from past cases is hard to collect.
 Jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking efforts to collect [e.g., @Craft2018].
-As a result, even where there, for any given attorney, available strike data on past trials, that data may be missing strikes from trials not at random.
+As a result, even where there is, for any given individual attorney or prosecutor's office, available strike data on past trials, that data may be missing strike data from trials not at random [e.g., @Baldus2001, 132-134].
 Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
 Some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck, such as their race or gender.
-As a result, lawyers and researchers must expend time and effort to access other data sources (e.g., voter registration rolls) with that information.
+As a result, lawyers and researchers must resort to other data sources (e.g., voter registration rolls) to find that information.
 Finally, regardless of record-keeping policy, judges and court staff vary in how often they fully complete or maintain the required information.
+
+When collecting strike data in multiple trials, prior studies have aggregated strikes, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
+For example, Eisenberg [-@Eisenberg2017, 334-339] used a 2-tail Fisher's exact test on the observed difference in peremptory strikes by race and gender in a non-random sample of thirty-five trials from 1997-2012 that resulted in death sentences (out of 63 such trials).
+@Baldus2001 similarly analyzed 317 venires (out of 461) in Philadelphia capital murder cases from which juries were selected between 1981 and 1997.
+
+In contrast, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
+Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
 
@@
 
-Here, we extend the Bayesian approach to _Batson_ challenges to assess whether one can credibly infer illegal bias based solely on an attorney's pattern of peremptory challenges using both present strike data and historical data that attorney's strikes in past cases.
+Here, we extend a Bayesian approach for deteching illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in present strike data and and that attorney's strikes in past cases.
 
@@
 Second, assign a prior distribution to the bias parameter $b$ that reflects the legal requirement that the _Batson_ challenger bear the burden of proof.
-Here, we assume a normal prior for $b$ with mean 0 and standard deviation 2.
 Third, using the strike data in past trials generated by the same peremptory-strike process, estimate a posterior distribution for $b$.
@@
 In so doing, we make two main contributions.
-First, we use the power prior to control how much the strike history in past cases affects the posterior distribution of the bias parameter $b$.
-The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of the unobserved variables of interest.
+First, we use the power prior to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
+The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of some parameter of interest.
 The key insight is to raise the likelihood function based on the historical data to a power parameter [@Ibrahim2000; @Ibrahim2015].
@@
 
-Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process as follows.
+Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
 
-For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
+In the "blind strike" variation, both parties simultaneously, rather than alternately, exercise their peremptory challenges on all the potential jurors.
+This may result in both sides using a strike each on the same potential juror.
+
+For any given case $i$ in which jury selection occurs, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
 If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).
 If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
-
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
-If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$.
+If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking someone who does not belong to that class.
 
 By adding one parameter $w$, we can measure the bias by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
-As a result, however, the weight of the non-cognizable class is the reciprocal of the weight of cognizable class.
-Using $b = log(w)$ avoids this issue.
+To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let $b = log(w)$.
 Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
@@
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data) as well as seven different values of the bias parameter for generating the historical data ("historical" $b = {-3,-2,-1,0,1,2,3}$).
+We also generated historical strike data of three sizes (same, double, and triple the size of current data) as well as seven different values of the bias parameter for generating the historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).
+We let historical $b$ differ from current $b$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random.
 
-We let historical $b$ differ from current $b$ (our parameter of interest) for two reasons.
-First, it allows us to simulate noise associated with historical strike data that is incomplete not at random.
-Second, it could also simulate a real difference in the bias parameter that depends on a feature that the current trial has but that _all_ the past trials in the historical data do not.
+Alternatively, it lets us simulate a real difference in the bias parameter that depends on a feature that the current trial has but that _all_ the past trials in the historical data do not.
 To illustrate, assume a prosecutor with two true bias values: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical data contains only strike data in cases with non-Black defendants, then that prosecutor's historical $b$ will differ from the current case with a Black defendant.
+If our historical data contains only strike data in cases with non-Black defendants, then that prosecutor's historical $b$ will differ from that prosecutor's bias in the current case with a Black defendant.
 
-By using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
+In any case, by using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
 If $a = 1$, the historical strike data is equally weighted with the present data on strikes.
 If $a < 1$, the historical data matters proportionally less.
-As discussed above, if some of the historical data is missing not at random, we may wish to assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter to the historical data on strikes.
+For reasons discussed above, we may wish to assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
 Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that parameter affected modeling accuracy.
@@
 For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$.
-Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.
+Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
 

### 2021-08-09 analysis/paper_draft.Rmd
updated intro text
added starter para. on simultaneous strikes

@@
 ---
-title: "Batson Challenges Using Historical Strike Data"
+title: "Detecting Bias in Peremptory Challenges Using Historical Strike Data"
 date: "`r Sys.Date()`"
@@
 
-This paper shows how to use historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely an attorney has used peremptory challenges because of prospective juror race or gender in violation of _Batson v.
+This paper presents a way to use historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely an attorney has used peremptory challenges because of prospective juror race or gender in violation of _Batson v.
 Kentucky_ (1986), and its progeny.
@@
 Under the US Supreme Court's decision in _Batson v.
-Kentucky_ (1986)_, a party's exercise of a peremptory challenge against one or more prosepctive jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
+Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prosepctive jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
 
@@
 @Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because strike data from past cases is hard to collect.
-Jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking efforts to collect [e.g., @Craft2018].
-As a result, even where there is, for any given individual attorney or prosecutor's office, available strike data on past trials, that data may be missing strike data from trials not at random [e.g., @Baldus2001, 132-134].
+Jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking efforts to collect [e.g., @Craft2018].
+
+As a result, even where there is, for any given individual attorney or prosecutor's office, available strike data on past trials, that data may be missing strike data not at random [e.g., @Baldus2001, 132-134].
+
 Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
@@
 
-Here, we extend a Bayesian approach for deteching illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in present strike data and and that attorney's strikes in past cases.
+Here, we use a Bayesian approach for detecting illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in both a current trial such attorney's strikes in past cases.
 
-In short, our workflow is this: First, specify a model of the peremptory-strike process that best approximates the peremptory-strike process in the jurisdiction of interest.
-Second, assign a prior distribution to the bias parameter $b$ that reflects the legal requirement that the _Batson_ challenger bear the burden of proof.
-Third, using the strike data in past trials generated by the same peremptory-strike process, estimate a posterior distribution for $b$.
-Finally, using data on the target attorney's strikes in the present case, use that posterior distribution as the prior for $b$ and update.
+First, we specify a model of the peremptory-strike process in the court of interest.
+Second, we assign an initial prior distribution for a bias parameter $b$.
+Third, we use an attorney's strike data from past trials, as  generated by the same peremptory-strike process, to estimate a posterior distribution for $b$.
+Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.
 
@@
 
-# Model
+# Model 1: Alternating Strikes
 
 Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
-
-In the "blind strike" variation, both parties simultaneously, rather than alternately, exercise their peremptory challenges on all the potential jurors.
-This may result in both sides using a strike each on the same potential juror.
 
@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-Traceplots, Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992] were used to measure the convergence.
-
-<!--
-
-# Model 2 - Sequential Strike Method
-
-Under the "sequential strike" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptory challenges.
-Then, the remaining prospective jurors appear one at a time before the attorneys for prosecutor and defendant, and those attorneys alternate as the order in which they decide whether to strike or not the prospective juror before them.
-For example, if the prosecutor had first choice whether to strike juror #1, the defendant's attorney would have first choice whether to strike juror # 2.
-If neither party exercises a strike on the prospective juror before them, that prospective juror is assigned a seat on the jury and the parties' attorneys cannot thereafter strike that juror.
-The process continues until the parties run out of strikes or the number seats on the jury (including alternates) have been filled.
-
-Accordingly, we model the sequential-strike process as follows.
-
-[insert statistical model]
--->
+TO measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992].
 
@@
 
-We conducted a simulation study to measure model performance under different values of the bias parameter, different amount of historical information and whether the bias/preference changes over time.
+We conducted a simulation study to measure model performance under different values of the bias parameter $b$ and different amounts and kinds of historical information.
 
-To do this, we generated strike data for current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$.
-As introduced above, $b<0$ represents a preference for a juror within the cognizable class; and $b>0$ represents a bias against the cognizable class.
-A greater absolute value of $b$ represents greater preference or bias for the cognizable class, while $b=0$ represents no bias.
+To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$.
+As introduced above, $b>0$ ($b<0$) represents a bias against (preference) for a prospective juror within (outside) the cognizable class, while $b=0$ denotes neither bias or preference.
+A greater absolute value of $b$ represents greater bias against or preference for the cognizable class.
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data) as well as seven different values of the bias parameter for generating the historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).
+We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).
+
 We let historical $b$ differ from current $b$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random.
+Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature that the current trial has but that all the past trials in the historical data do not.
+To illustrate, assume a prosecutor with two true bias values: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
+If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias $b$ in the current case with a Black defendant.
 
-Alternatively, it lets us simulate a real difference in the bias parameter that depends on a feature that the current trial has but that _all_ the past trials in the historical data do not.
-To illustrate, assume a prosecutor with two true bias values: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical data contains only strike data in cases with non-Black defendants, then that prosecutor's historical $b$ will differ from that prosecutor's bias in the current case with a Black defendant.
-
-In any case, by using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
+By using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
 If $a = 1$, the historical strike data is equally weighted with the present data on strikes.
 If $a < 1$, the historical data matters proportionally less.
-For reasons discussed above, we may wish to assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
+We can assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
 Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that parameter affected modeling accuracy.
@@
 We considered 735 scenarios in total, i.e., the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data and 5 different values of the power prior weight parameter $a$.
-For each scenario, we generated 1000 datasets and fit the model on those 1000 dataset.
+For each scenario, we generated 1000 datasets and fit the model on those datasets.
 We recorded the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
 
-In turn, we set three credible interval level (80\%, 90\% and 95\%) and recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$.
+In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$.
 For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$.
@@
 
-```{r b80,  echo=FALSE, fig.cap="80% credible interval", out.width = '100%'}
+```{r b80, echo=FALSE, fig.cap="80% credible interval"}
 
@@
 sims$a.f <- as.factor(sims$a)
+
 sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
@@
             color="black")+
-  facet_grid(h~a)+
+  facet_grid(h~a, labeller = label_both)+
   # scale_fill_gradient2(expression(hat(d)), 
@@
   coord_equal()+
-  theme_classic()+
+  theme_minimal()+
+  labs(fill = "coverage") +
   theme(axis.text = element_text(size =10), 
-        axis.title = element_text(size = 16), 
-        legend.title.align = 0.05, 
-        legend.title = element_text(size = 16), 
+        #axis.title = element_text(size = 16), 
+        #legend.title.align = 0.05, 
+        #legend.title = element_text(size = 16), 
         legend.text = element_text(size = 14))
@@
   coord_equal()+
-  theme_classic()+
+  theme_minimal()+
+  labs(fill = "coverage") +
   theme(axis.text = element_text(size =10), 
@@
   theme_classic()+  
+  labs(fill = "coverage") +
   theme(axis.text = element_text(size =10), 
@@
 
-From the plot, our model can detect strong bias and bias, when the bias parameter being -3,-2,2,3.
+These plots indicate that our model can detect strong bias and bias ($b = \{-3, -2, 2,3\}$).
 When bias is moderate ($b = \{1, -1\}$), the model can detect it if the historical data has a bias parameter in the same direction.
 If the bias parameter of historical data is in an opposite direction, the model may provide a wrong estimate of current bias parameter.
+
+# Model 2: Simultaneous Strikes
+
+In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.  This may result in both sides using a strike each on the same potential juror.
+
+For any given case $i$ in which jury selection occurs and a party uses $s$ number of strikes under this process, @Barrett2007 models the probability of a party striking at least $x$ cognizable-class members at random as following a hypergeometric distribution:
+
+$$
+Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}.
+$$
+where $x = 0, 1, 2, 3, .
+.
+.
+s$; $c_{i}$ denotes the number of cognizable class members in the venire subject to strike; and $m_{i}$ denotes the number of cognizable class non-members in the venire subject to strike.
+
+Here, we argue that a simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above in which one party engages in an uninterrupted sequence of strikes against members of the venire.
+The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
+
+[MORE]
 

### 2021-08-24 analysis/paper_draft.Rmd
Add some content of app prototype section

@@
 
-```{r b95,  echo=FALSE, fig.cap="95% credible interval"}
+```{r b95,  echo=FALSE,fig.cap="95% credible interval"}
 
@@
 
+There is currently historical information of some prosecutors and defendants.
+More historical information would be updated when available.
+Users can input the current trial information as shown in table \@ref(tab:inpts), and then select the name of the prosecutor and the defendant from the pull-down window if applicable.
+If the name of the prosecutor or the defendant cannot be found, it means the historical information is not available.
+Then the prior density plot and the posterior density plot would be displayed.
+
+```{r inpts, echo=FALSE}
+df <- data.frame(round = c(1:10),
+                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
+                 total = c(15,14,13,12,11,10,9,8,7,6),
+                 cog = c(0,1,1,0,0,1,1,0,0,0),
+                 party = rep(c("PP","PD"),5))
+knitr::kable(df)
+```
+
+For instance, if we collect the data of a new trial with prosecutor being "Brian P.
+Leaming" and defendant being "Frank J.
+Riccio II", and we would like to see if they are biased or not.
+We can select the name from the pull-down window of prosecutor and defendant to extract the historical information.
+Then we can input the current trial information, and select the weight that we would like to put on the historical information.
+Currently we provide three options for the weight, equal weight of historical information and current information, half weight of historical information and low weight of historical information.
+We can also select the cognizable class that we are interested in.
+There are currently two options of cognizable class, race and gender.
+Suppose the current trial information is as we show in table \@ref(tab:inpts), we put equal weight on historical information and we select cognizable class as race, then the posterior density as shown in figure \@ref(fig:brian) will be displayed.
+
+```{r brian, echo=FALSE,fig.cap="Posterior Density Plot"}
+library(dplyr)
+library(Rcpp)
+load_path2 <- here::here("app/mh_sampler_pp.cpp")
+load_path3 <- here::here("app/jury_data_cleaned_new.rds")
+sourceCpp(load_path2)
+sourceCpp(load_path2)
+dat0 <- readRDS(load_path3)
+df <- data.frame(round = c(1:10),
+                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
+                 total = c(15,14,13,12,11,10,9,8,7,6),
+                 cog = c(0,1,1,0,0,1,1,0,0,0),
+                 party = rep(c("PP","PD"),5))
+extract_atny <- function(atny_name,pp,dat){
+  if(pp){
+    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
+    return(dat[which(idx),])
+  }
+  else{
+    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
+    return(dat[which(idx),])
+  }
+} 
+organize_input <- function(dat,pp,cog){
+  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
+  if(cog=='gender'){
+    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
+  }else{
+    dat_strikes <- dat_strikes %>% filter(!is.na(race))
+  }
+
+  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
+  if(cog=='gender'){
+    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
+  }else{
+    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
+  }
+
+  num_cog <- sum(dat_strikes$cogb)
+  num_t <- nrow(dat_strikes)
+  subs <- data.frame(round = c(1:num_t), 
+                     num_cog = NA, 
+                     total = NA)
+  subs$num_cog[1] <- num_cog
+  subs$total[1] <- num_t
+  for ( j in 2:num_t){
+    num_cog = num_cog-dat_strikes$cogb[j-1]
+    num_total = num_t - dat_strikes$strike_seq[j-1]
+    subs$num_cog[j] <- num_cog
+    # number of females struck in this and previous rounds
+    subs$total[j] <- num_total
+  }
+  subs <- cbind(subs,dat_strikes$cogb)
+  colnames(subs)[4] = 'cognizable'
+  if(pp){
+    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
+  }else{
+    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PD'),])
+  }
+  return(df_m)
+}
+subset <- function(atny_name,pp,dat,cog){
+  dat_sub <- extract_atny(atny_name,pp,dat)
+  sub1 <- dat_sub %>% group_split(ID)
+  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
+  df_m = do.call(rbind.data.frame,sub1_l)
+  df_m <- as.matrix(df_m)
+}
+df_mp <- df %>%
+                filter(party == "PP") %>%
+                select(-c(party)) %>%
+                as.matrix()
+
+df_md <- df %>%
+                filter(party == "PD") %>%
+                select(-c(party)) %>%
+                as.matrix()
+
+sub_p <- subset("Brian P. Leaming",TRUE,dat0,"race")
+sub_d <- subset("Frank J. Riccio II",FALSE,dat0,"race")
+out_p <- make_posterior_p(x = df_mp,x_p =sub_p,a0 = 1, niter = 110000, 
+                                          theta_start_val = 0, theta_proposal_sd =.5, 
+                                          prior_mean = 0, prior_sd = 2)
+pp_prior_theta <- make_posterior_prior(x_p=sub_p,a0 = 1, niter = 110000, 
+                                        theta_start_val = 0, theta_proposal_sd =.5, 
+                                                       prior_mean = 0, prior_sd = 2)
+d_p <- data.frame(
+                theta = out_p$theta[10001:110000],
+                party = "Prosecution", 
+                posterior = "Posterior")
+pp_prior <- data.frame(theta = pp_prior_theta$theta[10001:110000], 
+                                       party = "Prosecution", 
+                                       posterior = "Prior")
+
+out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 1, niter = 110000, 
+                                          theta_start_val = 0, theta_proposal_sd =.5, 
+                                          prior_mean = 0, prior_sd = 2)
+pd_prior_theta <- make_posterior_prior(x_p=sub_d,a0 = 1, niter = 110000, 
+                                        theta_start_val = 0, theta_proposal_sd =.5, 
+                                        prior_mean = 0, prior_sd = 2)
+pd_prior <- data.frame(theta = pd_prior_theta$theta[10001:110000], 
+                                       party = "Defense", 
+                                       posterior = "Prior")
+d_d <- data.frame(
+                theta = out_d$theta[10001:110000],
+                party = "Defense", 
+                posterior = "Posterior")
+
+dat <- rbind(d_p,d_d)
+priors <- rbind(pp_prior, pd_prior)
+dat <- rbind(dat, priors)
+dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
+                                ordered = TRUE)
+CI <- dat %>% filter(posterior == 'Posterior') %>%
+                group_by(party) %>%
+                summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%
+                mutate(bias = ifelse(
+                    q1 <= 0 & q2 >= 0, "No Bias", "Bias"))
+pplot <- ggplot(data=dat) + 
+                geom_density(aes(x = theta, 
+                                 fill = interaction(party, posterior),
+                                 color = interaction(party, posterior),
+                                 alpha = interaction(party, posterior), 
+                                 ..scaled..))+
+                facet_wrap(~party, nrow =2)
+
+pplot <- pplot  + theme_minimal() +
+                 labs (title = "Likely values of b") +
+                 xlab("") + 
+                 ylab("") + 
+                 xlim(c(-6,6)) +
+                 scale_fill_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
+                 scale_color_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
+                 scale_alpha_manual("Group", values = c(0.3, 0.3, 0.1, 0.1)) +
+
+                  # edit text sizes for plot
+                 theme(legend.position="none",
+                       axis.text = element_text(size = 16), 
+                       strip.text = element_text(size = 18), 
+                       plot.title = element_text(size = 24),
+                       plot.subtitle = element_text(size = 16)) +
+                  # add line at zero for reference
+                 geom_vline(xintercept = 0, color = "black", lwd=1.5)
+
+ pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
+                               linetype="dashed", size = 0.9)+
+                    geom_vline(data=CI, aes(xintercept=q2), color = c("blue", "darkred"),
+                           linetype="dashed", size = 0.9)
+
+```
+
+The grey curve represents the probability density of the prior, the blue curve represents the posterior probability density of the defendant and the red curve represents the posterior probability density of the prosecutor.
+The two vertical dotted line represents the 80\% credible interval.
+We can find the majority part of both the prior and posterior density curve are on the right side of zero, and the 80\% credible interval is also on the right side of zero, indicating a bias of "Brian P.
+Leaming" and "Frank J.
+Riccio II" against the potential juror within the cognizable class of race.
+If we are interested in cognizable class of gender, we can modify our selection of cognizable class, and the density plot becomes figure \@ref(fig:brian2).
+
+```{r brian2, echo=FALSE,fig.cap="Posterior Density Plot"}
+load_path2 <- here::here("app/mh_sampler_pp.cpp")
+load_path3 <- here::here("app/jury_data_cleaned_new.rds")
+sourceCpp(load_path2)
+sourceCpp(load_path2)
+dat0 <- readRDS(load_path3)
+df <- data.frame(round = c(1:10),
+                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
+                 total = c(15,14,13,12,11,10,9,8,7,6),
+                 cog = c(0,1,1,0,0,1,1,0,0,0),
+                 party = rep(c("PP","PD"),5))
+extract_atny <- function(atny_name,pp,dat){
+  if(pp){
+    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
+    return(dat[which(idx),])
+  }
+  else{
+    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
+    return(dat[which(idx),])
+  }
+} 
+organize_input <- function(dat,pp,cog){
+  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
+  if(cog=='gender'){
+    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
+  }else{
+    dat_strikes <- dat_strikes %>% filter(!is.na(race))
+  }
+
+  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
+  if(cog=='gender'){
+    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
+  }else{
+    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
+  }
+
+  num_cog <- sum(dat_strikes$cogb)
+  num_t <- nrow(dat_strikes)
+  subs <- data.frame(round = c(1:num_t), 
+                     num_cog = NA, 
+                     total = NA)
+  subs$num_cog[1] <- num_cog
+  subs$total[1] <- num_t
+  for ( j in 2:num_t){
+    num_cog = num_cog-dat_strikes$cogb[j-1]
+    num_total = num_t - dat_strikes$strike_seq[j-1]
+    subs$num_cog[j] <- num_cog
+    # number of females struck in this and previous rounds
+    subs$total[j] <- num_total
+  }
+  subs <- cbind(subs,dat_strikes$cogb)
+  colnames(subs)[4] = 'cognizable'
+  if(pp){
+    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
+  }else{
+    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PD'),])
+  }
+  return(df_m)
+}
+subset <- function(atny_name,pp,dat,cog){
+  dat_sub <- extract_atny(atny_name,pp,dat)
+  sub1 <- dat_sub %>% group_split(ID)
+  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
+  df_m = do.call(rbind.data.frame,sub1_l)
+  df_m <- as.matrix(df_m)
+}
+df_mp <- df %>%
+                filter(party == "PP") %>%
+                select(-c(party)) %>%
+                as.matrix()
+
+df_md <- df %>%
+                filter(party == "PD") %>%
+                select(-c(party)) %>%
+                as.matrix()
+
+sub_p <- subset("Brian P. Leaming",TRUE,dat0,"gender")
+sub_d <- subset("Frank J. Riccio II",FALSE,dat0,"gender")
+out_p <- make_posterior_p(x = df_mp,x_p =sub_p,a0 = 1, niter = 110000, 
+                                          theta_start_val = 0, theta_proposal_sd =.5, 
+                                          prior_mean = 0, prior_sd = 2)
+pp_prior_theta <- make_posterior_prior(x_p=sub_p,a0 = 1, niter = 110000, 
+                                        theta_start_val = 0, theta_proposal_sd =.5, 
+                                                       prior_mean = 0, prior_sd = 2)
+d_p <- data.frame(
+                theta = out_p$theta[10001:110000],
+                party = "Prosecution", 
+                posterior = "Posterior")
+pp_prior <- data.frame(theta = pp_prior_theta$theta[10001:110000], 
+                                       party = "Prosecution", 
+                                       posterior = "Prior")
+
+out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 1, niter = 110000, 
+                                          theta_start_val = 0, theta_proposal_sd =.5, 
+                                          prior_mean = 0, prior_sd = 2)
+pd_prior_theta <- make_posterior_prior(x_p=sub_d,a0 = 1, niter = 110000, 
+                                        theta_start_val = 0, theta_proposal_sd =.5, 
+                                        prior_mean = 0, prior_sd = 2)
+pd_prior <- data.frame(theta = pd_prior_theta$theta[10001:110000], 
+                                       party = "Defense", 
+                                       posterior = "Prior")
+d_d <- data.frame(
+                theta = out_d$theta[10001:110000],
+                party = "Defense", 
+                posterior = "Posterior")
+
+dat <- rbind(d_p,d_d)
+priors <- rbind(pp_prior, pd_prior)
+dat <- rbind(dat, priors)
+dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
+                                ordered = TRUE)
+CI <- dat %>% filter(posterior == 'Posterior') %>%
+                group_by(party) %>%
+                summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%
+                mutate(bias = ifelse(
+                    q1 <= 0 & q2 >= 0, "No Bias", "Bias"))
+pplot <- ggplot(data=dat) + 
+                geom_density(aes(x = theta, 
+                                 fill = interaction(party, posterior),
+                                 color = interaction(party, posterior),
+                                 alpha = interaction(party, posterior), 
+                                 ..scaled..))+
+                facet_wrap(~party, nrow =2)
+
+pplot <- pplot  + theme_minimal() +
+                 labs (title = "Likely values of b") +
+                 xlab("") + 
+                 ylab("") + 
+                 xlim(c(-6,6)) +
+                 scale_fill_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
+                 scale_color_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
+                 scale_alpha_manual("Group", values = c(0.3, 0.3, 0.1, 0.1)) +
+
+                  # edit text sizes for plot
+                 theme(legend.position="none",
+                       axis.text = element_text(size = 16), 
+                       strip.text = element_text(size = 18), 
+                       plot.title = element_text(size = 24),
+                       plot.subtitle = element_text(size = 16)) +
+                  # add line at zero for reference
+                 geom_vline(xintercept = 0, color = "black", lwd=1.5)
+
+ pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
+                               linetype="dashed", size = 0.9)+
+                    geom_vline(data=CI, aes(xintercept=q2), color = c("blue", "darkred"),
+                           linetype="dashed", size = 0.9) 
+```
+
+When we consider cognizable class gender, we can find most of the density curve of both prior and posterior are on the left of zero, indicating the possible preference for the potential jurors within the cognizable class.
+However, the 80\% credible interval includes 0, thus the preference for the cognizable class of "Brian P.
+Leaming" and "Frank J.
+Riccio II" is not statistically significant.
+If we give low weight on historical information, the density plot will be as shown in figure \@ref(fig:brian3).
+
+```{r brian3, echo=FALSE,fig.cap="Posterior Density Plot"}
+load_path2 <- here::here("app/mh_sampler_pp.cpp")
+load_path3 <- here::here("app/jury_data_cleaned_new.rds")
+sourceCpp(load_path2)
+sourceCpp(load_path2)
+dat0 <- readRDS(load_path3)
+df <- data.frame(round = c(1:10),
+                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
+                 total = c(15,14,13,12,11,10,9,8,7,6),
+                 cog = c(0,1,1,0,0,1,1,0,0,0),
+                 party = rep(c("PP","PD"),5))
+extract_atny <- function(atny_name,pp,dat){
+  if(pp){
+    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
+    return(dat[which(idx),])
+  }
+  else{
+    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
+    return(dat[which(idx),])
+  }
+} 
+organize_input <- function(dat,pp,cog){
+  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
+  if(cog=='gender'){
+    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
+  }else{
+    dat_strikes <- dat_strikes %>% filter(!is.na(race))
+  }
+
+  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
+  if(cog=='gender'){
+    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
+  }else{
+    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
+  }
+
+  num_cog <- sum(dat_strikes$cogb)
+  num_t <- nrow(dat_strikes)
+  subs <- data.frame(round = c(1:num_t), 
+                     num_cog = NA, 
+                     total = NA)
+  subs$num_cog[1] <- num_cog
+  subs$total[1] <- num_t
+  for ( j in 2:num_t){
+    num_cog = num_cog-dat_strikes$cogb[j-1]
+    num_total = num_t - dat_strikes$strike_seq[j-1]
+    subs$num_cog[j] <- num_cog
+    # number of females struck in this and previous rounds
+    subs$total[j] <- num_total
+  }
+  subs <- cbind(subs,dat_strikes$cogb)
+  colnames(subs)[4] = 'cognizable'
+  if(pp){
+    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
+  }else{
+    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PD'),])
+  }
+  return(df_m)
+}
+subset <- function(atny_name,pp,dat,cog){
+  dat_sub <- extract_atny(atny_name,pp,dat)
+  sub1 <- dat_sub %>% group_split(ID)
+  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
+  df_m = do.call(rbind.data.frame,sub1_l)
+  df_m <- as.matrix(df_m)
+}
+df_mp <- df %>%
+                filter(party == "PP") %>%
+                select(-c(party)) %>%
+                as.matrix()
+
+df_md <- df %>%
+                filter(party == "PD") %>%
+                select(-c(party)) %>%
+                as.matrix()
+
+sub_p <- subset("Brian P. Leaming",TRUE,dat0,"gender")
+sub_d <- subset("Frank J. Riccio II",FALSE,dat0,"gender")
+out_p <- make_posterior_p(x = df_mp,x_p =sub_p,a0 = 0.2, niter = 110000, 
+                                          theta_start_val = 0, theta_proposal_sd =.5, 
+                                          prior_mean = 0, prior_sd = 2)
+pp_prior_theta <- make_posterior_prior(x_p=sub_p,a0 = 0.2, niter = 110000, 
+                                        theta_start_val = 0, theta_proposal_sd =.5, 
+                                                       prior_mean = 0, prior_sd = 2)
+d_p <- data.frame(
+                theta = out_p$theta[10001:110000],
+                party = "Prosecution", 
+                posterior = "Posterior")
+pp_prior <- data.frame(theta = pp_prior_theta$theta[10001:110000], 
+                                       party = "Prosecution", 
+                                       posterior = "Prior")
+
+out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 0.2, niter = 110000, 
+                                          theta_start_val = 0, theta_proposal_sd =.5, 
+                                          prior_mean = 0, prior_sd = 2)
+pd_prior_theta <- make_posterior_prior(x_p=sub_d,a0 = 0.2, niter = 110000, 
+                                        theta_start_val = 0, theta_proposal_sd =.5, 
+                                        prior_mean = 0, prior_sd = 2)
+pd_prior <- data.frame(theta = pd_prior_theta$theta[10001:110000], 
+                                       party = "Defense", 
+                                       posterior = "Prior")
+d_d <- data.frame(
+                theta = out_d$theta[10001:110000],
+                party = "Defense", 
+                posterior = "Posterior")
+
+dat <- rbind(d_p,d_d)
+priors <- rbind(pp_prior, pd_prior)
+dat <- rbind(dat, priors)
+dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
+                                ordered = TRUE)
+CI <- dat %>% filter(posterior == 'Posterior') %>%
+                group_by(party) %>%
+                summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%
+                mutate(bias = ifelse(
+                    q1 <= 0 & q2 >= 0, "No Bias", "Bias"))
+pplot <- ggplot(data=dat) + 
+                geom_density(aes(x = theta, 
+                                 fill = interaction(party, posterior),
+                                 color = interaction(party, posterior),
+                                 alpha = interaction(party, posterior), 
+                                 ..scaled..))+
+                facet_wrap(~party, nrow =2)
+
+pplot <- pplot  + theme_minimal() +
+                 labs (title = "Likely values of b") +
+                 xlab("") + 
+                 ylab("") + 
+                 xlim(c(-6,6)) +
+                 scale_fill_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
+                 scale_color_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
+                 scale_alpha_manual("Group", values = c(0.3, 0.3, 0.1, 0.1)) +
+
+                  # edit text sizes for plot
+                 theme(legend.position="none",
+                       axis.text = element_text(size = 16), 
+                       strip.text = element_text(size = 18), 
+                       plot.title = element_text(size = 24),
+                       plot.subtitle = element_text(size = 16)) +
+                  # add line at zero for reference
+                 geom_vline(xintercept = 0, color = "black", lwd=1.5)
+
+ pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
+                               linetype="dashed", size = 0.9)+
+                    geom_vline(data=CI, aes(xintercept=q2), color = c("blue", "darkred"),
+                           linetype="dashed", size = 0.9) 
+```
+
+We can find both the prior and posterior density become more flat, and the posterior density looks like symmetric around zero now.
+
 # Discussion

### 2021-08-24 analysis/paper_draft.Rmd
minor edits to paper text

--- 2021-08-09 analysis/paper_draft.Rmd
@@
 In the US, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
-During this "voir dire" process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed (declared ineligible for a seat on the jury) if a party's attorney uses a peremptory challenge against them.
+During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed (declared ineligible for a seat on the jury) if a party's attorney uses a peremptory challenge against them.
 
@@
 25, s 269), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases.
-In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].
+In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
 
@@
 In another variant, the parties strike in an alternating sequence, with the prosecution striking first.
-Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn.
+Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes consecutively.
 If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.
@@
 
+## Batson
+
 This paper focuses on _Batson_ challenges to a party's use of peremptory challenge.
 Under the US Supreme Court's decision in _Batson v.
-Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prosepctive jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
+Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prospective jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
 
@@
 As a result, even where there is, for any given individual attorney or prosecutor's office, available strike data on past trials, that data may be missing strike data not at random [e.g., @Baldus2001, 132-134].
-
-Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
-Some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck, such as their race or gender.
-As a result, lawyers and researchers must resort to other data sources (e.g., voter registration rolls) to find that information.
-Finally, regardless of record-keeping policy, judges and court staff vary in how often they fully complete or maintain the required information.
+Possible reasons include limited public access (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
+Some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
+Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
 
@@
 
-Here, we use a Bayesian approach for detecting illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in both a current trial such attorney's strikes in past cases.
-
+Here, we use a Bayesian approach for detecting illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in both a current trial and that attorney's strikes in past cases.
 First, we specify a model of the peremptory-strike process in the court of interest.
@@
 If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
+
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
@@
 where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
-In @Kadane2018, $b$ is estimated under Bayesian framework with a uniform prior from -6 to 6.
-
+In @Kadane2018, $b$ is estimated under a Bayesian framework with a uniform prior from -6 to 6.
 In this paper, we also would like to incorporate the historical data on strikes in past cases and be able to adjust the weight of that historical information on the posterior distribution of the bias parameter.
@@
 In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
-
-<!-- The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes.
--->
 
@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-TO measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992].
+To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992].
 
@@
 
-Here, we argue that a simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above in which one party engages in an uninterrupted sequence of strikes against members of the venire.
+Here, we argue that a simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above, i.e, one party engages in an uninterrupted sequence of strikes against members of the venire.
 The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
 
-[MORE]
+To show this, we conducted the following simulation study. [MORE]
 
@@
 The posterior probability distribution is estimated with `r R.version$version.string'[@R]. The app was built using the shiny package [@shiny].
-
-# Discussion
 

### 2021-08-24 analysis/paper_draft.Rmd

--- 2021-08-24 analysis/paper_draft.Rmd
@@
 
-# Discussion
-
 # Conclusion

### 2021-08-24 analysis/paper_draft.Rmd
edits; added para. justified prior of N(0,2); updated citations

@@
   - name: "Xiaomeng Li"
-    affiliation: "University of Connecticut"
+    affiliation: "University of Connecticut | Ernest & Young"
   - name: "Timothy E. Moore"
@@
 
-## Jury Selection
+With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
+While some countries have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
+2019, c.
+25, s 269 (Canada)), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases.
 
-With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
-While countries such as the United Kingdom and Canada have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1); S.C.
-2019, c.
-25, s 269), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases.
 In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
 
-Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
+Under the "struck jury" method of peremptory challenges, the trial judge decides whether to accept or reject all challenges for cause - for which a reason must be given -- before the parties exercise any peremptories.
+Thereafter, the number of potential jurors who remain is typically the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
 
@@
 Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes consecutively.
-If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.
+Once the parties have used (or waive) their strikes, the jurors are seated based on the order in which they were selected or at random.
 
@@
 
-## Batson
-
-This paper focuses on _Batson_ challenges to a party's use of peremptory challenge.
+This paper focuses on _Batson_ challenges to a party's use of peremptory challenges.
 Under the US Supreme Court's decision in _Batson v.
-Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prospective jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
+Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prospective jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, ethnicity, gender, or certain other protected characteristics.
 
 For a _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
-If so, the challenged party must produce one or more nondiscriminatory reasons for that strike.
+If so, the burden shifts to the challenged party to produce one or more nondiscriminatory reasons for that strike.
 Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
@@
 Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
-Possible explanations include attorney fear that raising a _Batson_ challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021].
+Possible explanations include attorney fear that even raising a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
 Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
@@
 @Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because strike data from past cases is hard to collect.
-Jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking efforts to collect [e.g., @Craft2018].
+Jury-selection records are often not readily available, because they are kept, if at all, by court staff somewhere other than the case file.
+Such records are also largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking effort to collect [e.g., @Craft2018].
 
-As a result, even where there is, for any given individual attorney or prosecutor's office, available strike data on past trials, that data may be missing strike data not at random [e.g., @Baldus2001, 132-134].
-Possible reasons include limited public access (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
-Some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
+Moreover, strike data on past trials may be missing information not at random [e.g., @Baldus2001, 132-134].
+Possible reasons include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
 Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
 
-When collecting strike data in multiple trials, prior studies have aggregated strikes, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
+When collecting strike data in multiple trials, prior studies have aggregated strikes across those trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
 For example, Eisenberg [-@Eisenberg2017, 334-339] used a 2-tail Fisher's exact test on the observed difference in peremptory strikes by race and gender in a non-random sample of thirty-five trials from 1997-2012 that resulted in death sentences (out of 63 such trials).
-@Baldus2001 similarly analyzed 317 venires (out of 461) in Philadelphia capital murder cases from which juries were selected between 1981 and 1997.
-
-In contrast, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
+@Gastwirth2013 has criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
 Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
 
-Finally, even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
+For exmpale, using jury selection data for 1992-2017 for Mississippi's Fifth Circuit Court District, Craft [-@Craft2018a, 9] applied a logistic regression model to analyze how much a prospective juror's race affected the odds of being struck by the state.
+For studies applying other modeling techniques to this Mississippi data, see @DeCamp2020 (propensity score matching) and @Dunn2021 (optimal feature selection, optimal classification trees).
 
@@
 \end{equation}
-<!-- \[b \sim \text{Normal}(0,2)\] -->
 
 where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
-In @Kadane2018, $b$ is estimated under a Bayesian framework with a uniform prior from -6 to 6.
-In this paper, we also would like to incorporate the historical data on strikes in past cases and be able to adjust the weight of that historical information on the posterior distribution of the bias parameter.
+
+Next, we assign an initial prior:
+
+\[b \sim \text{Normal}(0,2)\]
+
+Unlike @Kadane2018, who assigned $b$ a prior of $Uniform (-6,6)$, because the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior.
+
+First, for our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge as well as the initial burden of producing some relevant evidence of illegal bias (the _Batson_ prima facie case).
+Thus, if that challenging party in fact produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
+This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
+
+Second, we take that law as requiring, absent any data, the assumption that higher degrees of illegal bias (preference) are less likely than lower degrees of bias (preference).
+By using a normal (Gausian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
+
+Finally, we want to incorporate historical data on strikes in past cases and adjust the weight of that historical information on the posterior distribution of the bias parameter.
 Accordingly, we introduce the power prior \@ref(eq:model3)
@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992].
+To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998] and Geweke’s diagnostics [@Gewe1992].
 
@@
 
-The posterior probability distribution is estimated with `r R.version$version.string'[@R]. The app was built using the shiny package [@shiny].
+Even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in real time _during_ jury selection, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+
+Following this suggestion, we describe here a prototype software application ("app") that implements the approach described above using `r R.version$version.string`[@R] and shiny [@shiny].
 
@@
 Users can input the current trial information as shown in table \@ref(tab:inpts), and then select the name of the prosecutor and the defendant from the pull-down window if applicable.
-If the name of the prosecutor or the defendant cannot be found, it means the historical information is not available.
-Then the prior density plot and the posterior density plot would be displayed.
+If the name of the prosecutor or the defense attorney cannot be found, it means the historical information is not available.
+
+The main screen displays two graphs -- one for the prosecution and the defense - that displays the prior density plot and the posterior density plot for the bias parameter.
 
@@
 The two vertical dotted line represents the 80\% credible interval.
+
 We can find the majority part of both the prior and posterior density curve are on the right side of zero, and the 80\% credible interval is also on the right side of zero, indicating a bias of "Brian P.
@@
 ```{r brian2, echo=FALSE,fig.cap="Posterior Density Plot"}
+
 load_path2 <- here::here("app/mh_sampler_pp.cpp")
 load_path3 <- here::here("app/jury_data_cleaned_new.rds")
+
 sourceCpp(load_path2)
-sourceCpp(load_path2)
+
 dat0 <- readRDS(load_path3)
+
 df <- data.frame(round = c(1:10),
@@
                  party = rep(c("PP","PD"),5))
+
 extract_atny <- function(atny_name,pp,dat){
@@
 
-Sachin S. Pandya: Conceptualization; Xiaomeng Li: Conceptualization; [insert]; Timothy Moore: Conceptualization; [insert]. Based on categories in https://casrai.org/credit/].
+Sachin S.
+Pandya: Conceptualization, Investigation, Software, Writing; Xiaomeng Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Timothy Moore: Conceptualization, Supervision, Visualization.
+Based on categories in https://casrai.org/credit/].
 

### 2021-09-01 analysis/paper_draft.Rmd
add simultaneous strikes process simulation results

@@
 
-To show this, we conducted the following simulation study. [MORE]
+To show this, we conducted the following simulation study.
+We first generate a trial and then shuffle the order of the strikes to generate 50 trials with the same proportion of struck cognizable class members but different orders.
+We then use the model to fit the 50 trials to see if the estimated bias parameter are close to each other.
+We conduct the simulation study under three scenarios from moderate bias to strong bias, that is $b=1,2,3$.
+The results are shown in figure \@ref(fig:simltns).
+
+```{r simltns, echo=FALSE,fig.cap="simultaneous case"}
+library(dplyr)
+library(ggplot2)
+#library(conflicted)
+#conflict_prefer("select", "dplyr")
+#conflict_prefer("filter", "dplyr")
+load_path_b1 <- here::here("b1.csv")
+load_path_b2 <- here::here("b2.csv")
+load_path_b3 <- here::here("b3.csv")
+
+b1 <- read.csv(load_path_b1)
+b2 <- read.csv(load_path_b2)
+b3 <- read.csv(load_path_b3)
+
+b1.t <- b1 %>% select(-X) %>%
+  t() %>% data.frame() %>% 
+  dplyr::rename_with(~b1$X) %>%
+      mutate(sim = as.character(1:50))
+#b1.t$sim <- factor(b1.t$sim, levels = b1.t$sim, ordered = TRUE)
+
+b2.t <- b2 %>% select(-X) %>%
+  t() %>% data.frame() %>% 
+  dplyr::rename_with(~b2$X) %>%
+  mutate(sim =  as.character(1:50))
+#b2.t$sim <- factor(b2.t$sim, levels = b2.t$sim, ordered = TRUE)
+
+b3.t <- b3 %>% select(-X) %>%
+  t() %>% data.frame() %>% 
+  dplyr::rename_with(~b2$X) %>%
+  mutate(sim = as.character(1:50))
+#b3.t$sim <- factor(b3.t$sim, levels = b3.t$sim, ordered = TRUE)
+
+colnames(b1.t)[1:3] = c('b_s','2.5%','97.5%')
+colnames(b2.t)[1:3] = c('b_s','2.5%','97.5%')
+colnames(b3.t)[1:3] = c('b_s','2.5%','97.5%')
+
+b1.t$b = "1"
+b2.t$b = "2"
+b3.t$b = "3"
+b.com <- rbind(b1.t, b2.t, b3.t)
+b.com$sig <- ifelse(b.com$`2.5%`>0, "Bias", 
+                                  "No Bias")
+b.com$b.num <- as.numeric(b.com$b)
+
+b1.mean <-data.frame(b_s = mean(b1.t$b_s), 
+                     `2.5%` =  mean(b1.t$`2.5%`),
+                     `97.5%` = mean(b1.t$`97.5%`),
+                     sim = "mean", 
+                     b = "1", 
+                     sig = "No Bias", 
+                     b.num = 1)
+b2.mean <-data.frame(b_s = mean(b2.t$b_s), 
+                     `2.5%` =  mean(b2.t$`2.5%`),
+                     `97.5%` = mean(b2.t$`97.5%`),
+                     sim = "mean", 
+                     b = "2", 
+                     sig = "Bias", 
+                     b.num = 2)
+b3.mean <-data.frame(b_s = mean(b3.t$b_s), 
+                     `2.5%` =  mean(b3.t$`2.5%`),
+                     `97.5%` = mean(b3.t$`97.5%`),
+                     sim = "mean", 
+                     b = "3", 
+                     sig = "Bias", 
+                     b.num = 3)
+
+names(b1.mean)<- colnames(b.com)
+names(b2.mean)<- colnames(b.com)
+names(b3.mean)<- colnames(b.com)
+b.com<- rbind(b.com,b1.mean[1,], b2.mean[1,], b3.mean[1,])
+b.com$sim.n <- as.numeric(b.com$sim)
+b.com$sim.n[151:153] <- 52.5
+
+ggplot(data = b.com, aes(x = sim.n, y = b_s, color = sig))+
+  facet_grid(~b)+
+  geom_point(size = 3)+
+  geom_errorbar(aes(x = sim.n, ymin = `2.5%`, 
+                    ymax = `97.5%`), width = 0.4)+
+  geom_hline(data = b.com, aes(yintercept = b.num), col = "red", lwd = 1.1) +
+  geom_hline(yintercept = 0, col = "black", lwd = 1, lty = 1) +
+  geom_vline(xintercept = 51.25, col = "black", lwd = 1, lty = 2) +
+  xlab("Simulation")+
+  scale_color_manual("Bias", values = c("blue", "grey"))+
+  ylab("Posterior Bias Value")+
+  ylim(c(-0.5, 5))+
+  scale_x_continuous(breaks = c(1:50, 52.5), labels = c(1:50, "Average"))+
+  coord_flip()+
+  theme_classic()+
+  theme(axis.title = element_text(size = 16), 
+        axis.text.x = element_text(size = 12), 
+        axis.text.y = element_text(size = 9))
+```
+The points represent the estimated posterior mean and the lines represent the 95\% credible interval.
+The blue line means the credible interval does not include zero, thus indicating a bias, while the grey line means the credible interval includes zero and indicating no bias.
+We can find the estimated bias parameter of the 50 trials with different orders are close to each other for all of the three scenarios, indicating the order does not influence the estimate of the bias parameter.
+This is consistent with our previous premise and confirms that the simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process.
 

### 2021-09-05 analysis/paper_draft.Rmd
ref to AZ

--- 2021-08-24 analysis/paper_draft.Rmd
@@
 With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
-While some countries have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
+While some jurisdictions have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
 2019, c.
-25, s 269 (Canada)), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases.
-
-In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
+25, s 269 (Canada)), federal and State law in the United States (except in Arizona^[Arizona Rule of Criminal Procedure XX]) authorize their use during jury selection in criminal and civil cases.
+In these cases, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
 

### 2021-09-14 analysis/paper_draft.Rmd
updated main draft

@@
 ```{r simltns, echo=FALSE,fig.cap="simultaneous case"}
+
 library(dplyr)
@@
 #conflict_prefer("filter", "dplyr")
+
+# ADD SCRIPT FOR GENERATING b1.csv, b2.csv, b3.csv
+
 load_path_b1 <- here::here("b1.csv")

### 2021-09-14 dat0.R

@@ -1,0 +1,26 @@
+#' Jury Selection in Connecticut federal court
+#'
+#' A dataset containing the jury selection information for cases completed in the federal district court
+#'  of Connecticut for 2013-2018
+#'
+#' @format A data frame with 979 rows and 30 variables:
+#' \describe{
+#'   \item{Docket.x}{case docket number}
+#'   \item{Case_Name}{case name on caption}
+#'   \item{courthouse}{courthouse location}
+#'   \item{Defendant.x}{first defendant name on case caption}
+#'   \item{Judge}{ID (initials) for district court judge}
+#'   \item{date}{date jury selection began}
+#'   \item{R_No}{venireperson sequence order}
+#'   \item{Pool_Seq}{jury pool ID number}
+#'   \item{race}{numerical code for race from juror questionaire}
+#'   \item{hisp}{1 = "Hispanic", 0 = not}
+#'   \item{sex}{M = "male", F = "female"}
+#'   \item{Disp}{voir dire disposition. "PP" - prosecutor strike; "PD" - defense strike; "J" - seated juror}
+#'   \item{strike_num}{order of strikes - each side}
+#'   \item{P_atty}{prosecutors on docket sheet at time of jury selection}
+#'   \item{D_atty}{defense attorneys on docket sheet at time of jury selection}
+#'   ...
+#' }
+#' @source Jury Administrator, US District Court - Connecticut
+"dat0"

### 2021-09-14 analysis/paper_draft2.Rmd

--- 2021-09-14 analysis/paper_draft.Rmd
@@
 ---
-title: "Detecting Bias in Peremptory Challenges Using Historical Strike Data"
+title: "Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data"
 date: "`r Sys.Date()`"
+author:
+  - Sachin S. Pandya
+#    orcid_id: 0000-0001-7387-1307
+  - Xiaomeng Li
+  - Timothy E. Moore
+#    orcid_id: 0000-0002-9576-0517
+#output: distill::distill_article
+output: pdf_document
 bibliography: references.bib
-author:
-  - name: "Sachin S. Pandya"
-    affiliation: "University of Connecticut"
-    orcid_id: 0000-0001-7387-1307
-  - name: "Xiaomeng Li"
-    affiliation: "University of Connecticut | Ernest & Young"
-  - name: "Timothy E. Moore"
-    affiliation: "University of Connecticut"
-    orcid_id: 0000-0002-9576-0517
-output: distill::distill_article
+abstract: |
+  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, gender, or membership in certain other cognizable classes.
+Here, we propose a Bayesian approach for detecting such illegal bias.
+We show how to use historical data on an attorney's use of peremptory strikes in past cases to estimate, in a current trial, how likely an attorney has struck prospective jurors because of cognizable-class membership.
+We also use of a _power prior_ to ajust the weight of such historical data in the analysis.
+We show the utility of our model on simulated data.
+Finally, we extend our approach by presenting a prototype software application to be used to detect bias in peremptory-challenge in real time during jury-selection.
+We demonstrate the app using actual historical strike data from a convenience sample of cases from the federal district court of Connecticut from YYYY to YYYY.
+
 ---
+
+```{r setup, include=FALSE}
+
+library(ggplot2)
+library(dplyr)
+
+knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)
+
+```
 
@@
 
-This paper presents a way to use historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely an attorney has used peremptory challenges because of prospective juror race or gender in violation of _Batson v.
-Kentucky_ (1986), and its progeny.
+In the US, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
+During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them.
+By asserting a peremptory challenge, a party can declare a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
 
-In the US, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
-During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed (declared ineligible for a seat on the jury) if a party's attorney uses a peremptory challenge against them.
-
-# Background and Related Literature
-
-With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
-While some jurisdictions have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
-2019, c.
-25, s 269 (Canada)), federal and State law in the United States (except in Arizona^[Arizona Rule of Criminal Procedure XX]) authorize their use during jury selection in criminal and civil cases.
-In these cases, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
-
-Under the "struck jury" method of peremptory challenges, the trial judge decides whether to accept or reject all challenges for cause - for which a reason must be given -- before the parties exercise any peremptories.
-Thereafter, the number of potential jurors who remain is typically the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
-
-Then, the parties exercise their strikes on those potential jurors.
-In one procedure, the parties submit their strikes simultaneously.
-In another variant, the parties strike in an alternating sequence, with the prosecution striking first.
-Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes consecutively.
-Once the parties have used (or waive) their strikes, the jurors are seated based on the order in which they were selected or at random.
-
-Under a "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box.
-Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.
-Once that is completed, new potential jurors take the seats of the excused or stricken jurors, and the process repeats until all the jury seats are filled.
-At any point, strikes can be used on anyone in the jury box.
-
-<!--
-
-Note: Constitutionality of Canada's abolition of strikes upheld in https://decisions.scc-csc.ca/scc-csc/scc-csc/en/item/18932/index.do
-
-Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p.
-297-298].
-This particularly matters where each side does not have the same number of strikes.
-For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.
-
--->
-
-This paper focuses on _Batson_ challenges to a party's use of peremptory challenges.
-Under the US Supreme Court's decision in _Batson v.
-Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prospective jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, ethnicity, gender, or certain other protected characteristics.
-
-For a _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
-If so, the burden shifts to the challenged party to produce one or more nondiscriminatory reasons for that strike.
-Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
+Since the US Supreme Court's decision in _Batson v.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class.
+The party bringing a _Batson_ challenge bears the burden of proving such illegal bias [@LaFave2020, sec.
 22.3].
-For similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v.
-Holloway, 209 Conn.
-636, 646 & n.4 (1989); State v.
-Chapman, 317 S.C.
-302, 305-06 (1995)], while other States have shifted the inquiry away from the striking party's motives to whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash.
+For similar challenges under State law, a few States accept proof not of motive but only that an "objective observer" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [e.g., Wash.
 R.
 37(e)].
+Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
 
-Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
-Possible explanations include attorney fear that even raising a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
-Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
+In any case, evidence supporting such challenges may include aggregate data on peremptory challenges in past cases.
+For example, in a State criminal case, evidence for a _Batson_ challenge can include the "relevant history of the State's peremptory strikes in past cases." [@flowers2019, 2243].
+Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
+@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because jury-selection records are often not readily available [@Grosso2017; @Wright2018] and otherwise require painstaking effort to collect [e.g., @APM2018].
 
-By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
-For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v.
-Dretke, 545 U.S.
-231, 253 (2005); Riley v.
-Taylor, 277 F.3d 261, 280-81, 283 (3d Cir.
-2001); Cochran v.
-Herring, 43 F.3d 1404, 1412 (11th Cir.
-1995)].
-The US Supreme Court recently identified evidence for a _Batson_ challenge as including the "relevant history of the State's peremptory strikes in past cases" [Flowers v.
-Mississippi, 139 S.
-Ct.
-2228, 2243 (2019)].
+When collecting strike data in multiple trials, prior studies have aggregated strikes across trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
+[@Eisenberg2017, 334-339; see also @Gastwirth2014; @Gastwirth2013].
+Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
+For different approaches to the same historical strike data from from Mississippi's Fifth Circuit Court District (1992-2017), see @Craft2018a (logistic regression); @DeCamp2020 (propensity score matching); and @Dunn2021 (optimal feature selection, optimal classification trees).
 
-Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
-@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because strike data from past cases is hard to collect.
-Jury-selection records are often not readily available, because they are kept, if at all, by court staff somewhere other than the case file.
-Such records are also largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking effort to collect [e.g., @Craft2018].
-
-Moreover, strike data on past trials may be missing information not at random [e.g., @Baldus2001, 132-134].
-Possible reasons include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
-Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-
-When collecting strike data in multiple trials, prior studies have aggregated strikes across those trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
-For example, Eisenberg [-@Eisenberg2017, 334-339] used a 2-tail Fisher's exact test on the observed difference in peremptory strikes by race and gender in a non-random sample of thirty-five trials from 1997-2012 that resulted in death sentences (out of 63 such trials).
-@Gastwirth2013 has criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
-Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
-
-Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
-For exmpale, using jury selection data for 1992-2017 for Mississippi's Fifth Circuit Court District, Craft [-@Craft2018a, 9] applied a logistic regression model to analyze how much a prospective juror's race affected the odds of being struck by the state.
-For studies applying other modeling techniques to this Mississippi data, see @DeCamp2020 (propensity score matching) and @Dunn2021 (optimal feature selection, optimal classification trees).
-
-# Discussion
-
-Here, we use a Bayesian approach for detecting illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in both a current trial and that attorney's strikes in past cases.
-First, we specify a model of the peremptory-strike process in the court of interest.
-Second, we assign an initial prior distribution for a bias parameter $b$.
-Third, we use an attorney's strike data from past trials, as  generated by the same peremptory-strike process, to estimate a posterior distribution for $b$.
+In this paper, we take a Bayesian approach to _Batson_ challenges [@Kadane2018] and extend it to incorporate historical strike data.
+In this approach, we specify a model of the peremptory-strike process in the court of interest that includes bias parameter $b$, to which we assign an initial prior distribution.
+Then, we use an attorney's strike data from past trials, as generated by the same peremptory-strike process, to estimate a posterior distribution for $b$.
+In so doing, we use the _power prior_ [@Ibrahim2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
 Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.
 
-In so doing, we make two main contributions.
-First, we use the power prior to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
-The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of some parameter of interest.
-The key insight is to raise the likelihood function based on the historical data to a power parameter [@Ibrahim2000; @Ibrahim2015].
+Here, we demonstrate this approach, including the results of simulation studies to test model performance.
+Then, we present a software prototype that would let an attorney use the same approach in using historical strike data to help detect bias in peremptory-challenge use in real time during jury selection.
 
-Second, we implement the analysis as prototype software that an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.
+# Method
 
-# Model 1: Alternating Strikes
+## Statistical Procedure
 
-Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Following @Kadane2018 and @Barrett2007, we model a "struck-jury" peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Under the "struck jury" procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
+Then, the parties exercise their strikes on those potential jurors, either submitting strikes simultaneously or in an alternating sequence.
+Once all strikes are used or waived, the jurors are seated based on the order in which they were selected or at random.
 
-For any given case $i$ in which jury selection occurs, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
+Accordingly, for any given case $i$ in which jury selection occurs, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
 If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).
 If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
-
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
@@
 
-By adding one parameter $w$, we can measure the bias by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
+By adding one parameter $w$, we can measure the _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
 To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let $b = log(w)$.
+
 Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
 
-\begin{equation}
+$$
 Pr(\delta_{ij} | b) = 
@@
   \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
-\end{cases} (\#eq:model1)
-\end{equation}
+\end{cases}
+$$
 
@@
 
-\begin{equation}  
-L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}} (\#eq:model2)
-\end{equation}
-
+$$ 
+L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
+$$
 where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
 
-Next, we assign an initial prior:
+## Incorporating historical strike data
 
-\[b \sim \text{Normal}(0,2)\]
+We incorporate the historical data on strikes in past cases and allow for adjustment of the weight of that historical information on the posterior distribution of the bias parameter.
+To do this, we introduce the power prior:
 
-Unlike @Kadane2018, who assigned $b$ a prior of $Uniform (-6,6)$, because the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior.
+$$\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)$$
+where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
-First, for our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge as well as the initial burden of producing some relevant evidence of illegal bias (the _Batson_ prima facie case).
+In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
+This departs from @Kadane2018, who assigned $b$ a prior of $Unif(-6,6)$.
+We assume that the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior.
+For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge as well as the initial burden of producing some relevant evidence of illegal bias (the _Batson_ prima facie case).
 Thus, if that challenging party in fact produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
@@
 
-Second, we take that law as requiring, absent any data, the assumption that higher degrees of illegal bias (preference) are less likely than lower degrees of bias (preference).
-By using a normal (Gausian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
-
-Finally, we want to incorporate historical data on strikes in past cases and adjust the weight of that historical information on the posterior distribution of the bias parameter.
-Accordingly, we introduce the power prior \@ref(eq:model3)
-
-\begin{equation}
-\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b) (\#eq:model3)
-\end{equation}
-
-where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
-In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
+Moreover, we take that law as requiring, absent any data, the assumption that higher degrees of illegal bias (preference) are less likely than lower degrees of bias (preference).
+By using a normal (Gaussian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
 
@@
 
-\begin{equation}
-L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
-\end{equation}
+$$
+L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
+$$
 
-where $L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$ is the likelihood function of $b$ given historical data.
+where
+
+$$L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$$
+
+is the likelihood function of $b$ given historical data.
 $n_{0i}$ is the total number of jury selections in the historical data, and $n_{0j}$ is the total number of peremptory strikes in the historical data.
@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998] and Geweke’s diagnostics [@Gewe1992].
+To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992; @Gelman1998;@Geweke1992].
 
-# Simulation
+## Model Performance on Simulated Data
 
@@
 Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature that the current trial has but that all the past trials in the historical data do not.
-To illustrate, assume a prosecutor with two true bias values: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
+To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
 If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias $b$ in the current case with a Black defendant.
 
-By using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
+Because we use the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
 If $a = 1$, the historical strike data is equally weighted with the present data on strikes.
@@
 
-```{r b80, echo=FALSE, fig.cap="80% credible interval"}
+```{r b80, fig.cap="80% credible interval"}
 
-library(ggplot2)
+# simulated data via "noseq_simulation.R" script
 
-# INSERT simulation and plot code
-
-load_path <- here::here("sim_resultc.csv")
+load_path <- here::here("data-sim","sim_resultc.csv")
 
@@
 sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
-sims$sig <- ifelse(sims$L80 >0 & sims$U80 >0 , 
-                                    "pos", sims$sig) 
-sims$sig <- ifelse(sims$L80 <0 & sims$U80 <0 , 
-                                    "neg", sims$sig) 
+sims$sig <- ifelse(sims$L80 >0 & sims$U80 >0 , "pos", sims$sig) 
+sims$sig <- ifelse(sims$L80 <0 & sims$U80 <0 , "neg", sims$sig)
+
 sims$sig.n <- 0
@@
 
-p = ggplot()+
+library(ggplot2)
+
+p80 <- ggplot()+
   geom_tile(data = sims, aes(x = d, y = d_h,
@@
   scale_fill_manual(values = c("white", "blue"))+
-  coord_equal()+
-  theme_minimal()+
+  coord_equal() +
+  theme_minimal() +
   labs(fill = "coverage") +
@@
         legend.text = element_text(size = 14))
-p
+p80
 
@@
 
-```{r b90, echo=FALSE, fig.cap="90% credible interval"}
+```{r b90, fig.cap="90% credible interval"}
 
@@
 
-p = ggplot()+
+p90 <- ggplot()+
   geom_tile(data = sims, aes(x = d, y = d_h,
@@
   facet_grid(h~a, labeller = label_both)+
-  # scale_fill_gradient2(expression(hat(d)), 
-  #                      low = "darkred", mid = "white", 
-  #                      high = "blue", 
-  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
-  #                      limits = c(-3.2, 3.2),
-  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
   scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
@@
         legend.text = element_text(size = 14))
-p
+p90
+
 ```
 
-```{r b95,  echo=FALSE,fig.cap="95% credible interval"}
+```{r b95, fig.cap="95% credible interval"}
 
@@
 
-p = ggplot()+
+p<- ggplot()+
   geom_tile(data = sims, aes(x = d, y = d_h,
@@
         legend.text = element_text(size = 14))
-p
+
+p90
+
 ```
@@
 
-# Model 2: Simultaneous Strikes
+## Extension: Simultaneous Strikes
 
@@
 
-$$
-Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}.
-$$
+$$Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}$$
+
 where $x = 0, 1, 2, 3, .
@@
 
-```{r simltns, echo=FALSE,fig.cap="simultaneous case"}
+```{r simltns, fig.cap="simultaneous case", eval = FALSE}
 
-library(dplyr)
-library(ggplot2)
 #library(conflicted)
@@
 
-# ADD SCRIPT FOR GENERATING b1.csv, b2.csv, b3.csv
+# SCRIPT FOR GENERATING b1.csv, b2.csv, b3.csv
+
+# generate data
+
+source("R/simultaneous1.R")
+
+# load already-generated data
 
@@
 
-ggplot(data = b.com, aes(x = sim.n, y = b_s, color = sig))+
+simplot <- ggplot(data = b.com, aes(x = sim.n, y = b_s, color = sig))+
   facet_grid(~b)+
@@
         axis.text.y = element_text(size = 9))
+
+simplot
+
 ```
+
 The points represent the estimated posterior mean and the lines represent the 95\% credible interval.
@@
 We can find the estimated bias parameter of the 50 trials with different orders are close to each other for all of the three scenarios, indicating the order does not influence the estimate of the bias parameter.
-This is consistent with our previous premise and confirms that the simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process.
+This confirms that the simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process.
 
-# The App Prototype
+# The Software Prototype
 
-Even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in real time _during_ jury selection, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+We describe here a prototype software application ("app") that implements the approach described above.
+We built this app using `r R.version$version.string`[@R] and shiny [@shiny].
+To demonstrate how this app might be used by lawyers in real cases, we loaded this app with real strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 
-Following this suggestion, we describe here a prototype software application ("app") that implements the approach described above using `r R.version$version.string`[@R] and shiny [@shiny].
+To collect the data, one of us filed a request with the U.S.
+District Court of Connecticut based on 28 U.S.C.
+$\S$ 1868.
+Thereafter, we received copies (pdf files) of certain jury-selection records associated with X trials in that court during this period.
+These included strike sheets that indicated the ID number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
+Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.
 
-There is currently historical information of some prosecutors and defendants.
-More historical information would be updated when available.
-Users can input the current trial information as shown in table \@ref(tab:inpts), and then select the name of the prosecutor and the defendant from the pull-down window if applicable.
-If the name of the prosecutor or the defense attorney cannot be found, it means the historical information is not available.
+These records, however, often did not indicate the identity of the attorneys exercising the strikes.
+While the standard forms included a signature line for the attorney, many were left blank or filled with illegible signatures.
+Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the US Attorney's office) or the criminal defendant(s) on the date(s) of jury selection.
+Where only a single attorney represented a party during jury selection, it was easy to attribute the pattern of strikes to that attorney.
+Where multiple lawyers appeared for one side, we attributed to each of them that side's pattern of peremptory challenges in that case.
+In such cases, neither the jury-selection documents nor the docket sheets indicated any hierarchy among multiple lawyers or any other basis to attribute strikes to only one attorney among them.
+After generating a dataset based on these documents, we kept only strikes where a criminal defendant was represented by an attorney.
+Finally, to de-identify this dataset, we excluded defendant names and substituted fictitious names for the attorneys using the charlatan package [@charlatan].
 
-The main screen displays two graphs -- one for the prosecution and the defense - that displays the prior density plot and the posterior density plot for the bias parameter.
+With the app, the user uses the pull-down menus to select the cognizable class (top left) and enters by hand the strike information in the case before them in the _strike tally_ table (bottom left), either in real time or immediately after jury selection concludes.
+For demonstration purposes, the prototype app comes pre-loaded with a completed strike tally with values that can be changed and rows that can be added or deleted (Table \@ref(tab:inpts)).
+In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prosepective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
 
@@
 knitr::kable(df)
+
 ```
 
-For instance, if we collect the data of a new trial with prosecutor being "Brian P.
-Leaming" and defendant being "Frank J.
-Riccio II", and we would like to see if they are biased or not.
-We can select the name from the pull-down window of prosecutor and defendant to extract the historical information.
-Then we can input the current trial information, and select the weight that we would like to put on the historical information.
-Currently we provide three options for the weight, equal weight of historical information and current information, half weight of historical information and low weight of historical information.
-We can also select the cognizable class that we are interested in.
-There are currently two options of cognizable class, race and gender.
-Suppose the current trial information is as we show in table \@ref(tab:inpts), we put equal weight on historical information and we select cognizable class as race, then the posterior density as shown in figure \@ref(fig:brian) will be displayed.
+For the prototype app, we set two cognizable-class options: race and gender.
+This was simply because the historical data only contained information on prospective juror race and gender, itself a function of data collected as part of the standard juror questionnaire.
+The data did not expressly indicate the race or gender of the criminal defendant(s).
 
-```{r brian, echo=FALSE,fig.cap="Posterior Density Plot"}
+In the default setting, the pull-down menus for prosecutor and defense are set to "None".
+In that setting, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b = N(0,2)$).
+The main screen accordingly displays two graphs -- one for the prosecution and the defense - that each depict the prior and posterior density plots for the bias parameter.
+
+To use historical strike data, the user selects a name of the prosecutor or defense attorney from the pull-down menus.
+If an attorney's name cannot be found, the app has no historical strike data for that attorney.
+Once selected, the prior density plot (colored grey) and the posterior density plot (blue and red for defense and prosecution, respectively) automatically updates to account for the pre-loaded historical strike data for that attorney.
+For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information ($a = 1$).
+The user has two other options: half weight of historical information ($a = 0.5$) and low weight of historical information ($a = 0.2$).
+
+To illustrate, suppose the strike tally in Table \@ref(tab:inpts) represents the pattern of strikes in the present case with prosecutor "X" and defense attorney "Y".
+To detect racial bias in the use of peremptory strikes using both the strike tally and historical strike data, we set the cognizable-class to "Race"; select the names of the prosecutor and defense attorney from their respective pull-down menus; and set the weight setting to "Equal".
+The app will in turn display the posterior density plots in Figure \@ref(fig:brian).
+
+```{r brian, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
 library(dplyr)
@@
 
-The grey curve represents the probability density of the prior, the blue curve represents the posterior probability density of the defendant and the red curve represents the posterior probability density of the prosecutor.
-The two vertical dotted line represents the 80\% credible interval.
+The grey curve represents the probability density of the prior, the blue curve represents the posterior probability density of the defense attorney and the red curve represents the posterior probability density of the prosecutor.
+The two vertical dotted lines depict the 80\% credible interval.
+Most of both the prior and posterior density curve are on the right side of zero, and the 80\% credible interval is also on the right side of zero.
+Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
 
-We can find the majority part of both the prior and posterior density curve are on the right side of zero, and the 80\% credible interval is also on the right side of zero, indicating a bias of "Brian P.
-Leaming" and "Frank J.
-Riccio II" against the potential juror within the cognizable class of race.
-If we are interested in cognizable class of gender, we can modify our selection of cognizable class, and the density plot becomes figure \@ref(fig:brian2).
+Now suppose that the strike tally indicates the pattern of strikes by these attorneys against women.
+If so, we can select "gender" as the cognizable class, and the density plots becomes Figure \@ref(fig:brian2).
 
-```{r brian2, echo=FALSE,fig.cap="Posterior Density Plot"}
+```{r brian2, fig.cap="Posterior Density Plots, a = 1, cognitive class = gender"}
 
@@
 
-When we consider cognizable class gender, we can find most of the density curve of both prior and posterior are on the left of zero, indicating the possible preference for the potential jurors within the cognizable class.
-However, the 80\% credible interval includes 0, thus the preference for the cognizable class of "Brian P.
-Leaming" and "Frank J.
-Riccio II" is not statistically significant.
-If we give low weight on historical information, the density plot will be as shown in figure \@ref(fig:brian3).
+Here, most of the density curve of both prior and posterior are on the left of zero, indicating the possible bias against male jurors.
+However, since the 80\% credible interval includes 0, inferring that preference from the strike data alone is unjustified.
 
-```{r brian3, echo=FALSE,fig.cap="Posterior Density Plot"}
+Finally, if we give low weight on historical information, the density plot will be as shown in figure \@ref(fig:brian3).
+
+```{r brian3, echo=FALSE,fig.cap="Posterior Density Plot, a = 0.2"}
 load_path2 <- here::here("app/mh_sampler_pp.cpp")
@@
 
-We can find both the prior and posterior density become more flat, and the posterior density looks like symmetric around zero now.
+We can find both the prior and posterior density become more flat, and the posterior density appears symmetric around zero now.
+
+# Discussion
+
+In a typical _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
+If so, the burden shifts to the challenged party to produce one or more nondiscriminatory reasons for that strike.
+Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
+22.3].
+
+Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
+Possible explanations include attorney fear that even raising a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
+Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
+
+One virtue of incorporating the power prior: Strike data on past trials may be missing information not at random [e.g., @Baldus2001, 132-134].
+Possible reasons include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
+Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
+
+## Costs and Benefits of Bayesian Approach
 

### 2021-09-14 analysis/paper_draft_TEM.Rmd

--- 2021-09-05 analysis/paper_draft.Rmd
@@
 ---
-title: "Detecting Bias in Peremptory Challenges Using Historical Strike Data"
+title: "A Bayesian Approach to Detecting Bias in Peremptory Challenges Using Historical Strike Data"
 date: "`r Sys.Date()`"
-bibliography: references.bib
 author:
@@
   - name: "Xiaomeng Li"
-    affiliation: "University of Connecticut | Ernest & Young"
+    affiliation: "University of Connecticut"
   - name: "Timothy E. Moore"
@@
 
+# Abstract
+The _Batson_ decision, and subsequent decisions, forbids the use of peremptory challenges for jury selection in ways that discriminate against cognizable classes, e.g., by race, sex, ethnicity, etc.
+Here, we propose a Bayesian approach for detecting illegal bias for _Batson_ challenges based on an attorney's pattern of peremptory challenges in both a current trial, as well as an attorney's strikes in past cases.
+Specifically, we presents the use of historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely [should this rather be based on probability?] an attorney has used peremptory challenges because of prospective juror race or gender in violation of _Batson v.
+Kentucky_ (1986), and its progeny.
+We also introduce the use of a _power prior_ to provide flexibility in the use of historical trial data.
+We demonstrate the utility of our model on simulated data.
+Finally, we present a user-friendly R Shiny application that can be used to examine jury selection data in real time.
+We demonstrate the app using trail data collected from courts in Connecticut.
+
 # Introduction
 
-This paper presents a way to use historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely an attorney has used peremptory challenges because of prospective juror race or gender in violation of _Batson v.
-Kentucky_ (1986), and its progeny.
+## Jury Selection Procedures
 
 In the US, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
-During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed (declared ineligible for a seat on the jury) if a party's attorney uses a peremptory challenge against them.
+During this "voir dire" process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed (declared ineligible for a seat on the jury) if a party's attorney uses a peremptory challenge against them.
+Under the US Supreme Court's decision in _Batson v.
+Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prosepctive jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.
+In bringing a _Bateson_ challenge the burden is on the attorney bringing the challenge to provide evidence of bias against prospective jurors in a cognizable class.
 
-# Background and Related Literature
+This paper focuses on _Batson_ challenges to a party's use of peremptory challenge.
+With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
+While countries such as the United Kingdom and Canada have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1); S.C.
+2019, c.
+25, s 269), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases.[ repetitive] In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].
 
-With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
-While some jurisdictions have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
-2019, c.
-25, s 269 (Canada)), federal and State law in the United States (except in Arizona^[Arizona Rule of Criminal Procedure XX]) authorize their use during jury selection in criminal and civil cases.
-In these cases, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
-
-Under the "struck jury" method of peremptory challenges, the trial judge decides whether to accept or reject all challenges for cause - for which a reason must be given -- before the parties exercise any peremptories.
-Thereafter, the number of potential jurors who remain is typically the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
-
+Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
 Then, the parties exercise their strikes on those potential jurors.
@@
 In another variant, the parties strike in an alternating sequence, with the prosecution striking first.
-Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes consecutively.
-Once the parties have used (or waive) their strikes, the jurors are seated based on the order in which they were selected or at random.
-
+Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn.
+If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.
 Under a "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box.
@@
 <!--
-
 Note: Constitutionality of Canada's abolition of strikes upheld in https://decisions.scc-csc.ca/scc-csc/scc-csc/en/item/18932/index.do
-
 Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p.
@@
 For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.
-
 -->
 
-This paper focuses on _Batson_ challenges to a party's use of peremptory challenges.
-Under the US Supreme Court's decision in _Batson v.
-Kentucky_ (1986), a party's exercise of a peremptory challenge against one or more prospective jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, ethnicity, gender, or certain other protected characteristics.
-
-For a _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
-If so, the burden shifts to the challenged party to produce one or more nondiscriminatory reasons for that strike.
-Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
-22.3].
-For similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v.
-Holloway, 209 Conn.
-636, 646 & n.4 (1989); State v.
-Chapman, 317 S.C.
-302, 305-06 (1995)], while other States have shifted the inquiry away from the striking party's motives to whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash.
-R.
-37(e)].
-
-Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
-Possible explanations include attorney fear that even raising a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
-Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
-
-By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
-For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v.
-Dretke, 545 U.S.
-231, 253 (2005); Riley v.
-Taylor, 277 F.3d 261, 280-81, 283 (3d Cir.
-2001); Cochran v.
-Herring, 43 F.3d 1404, 1412 (11th Cir.
-1995)].
-The US Supreme Court recently identified evidence for a _Batson_ challenge as including the "relevant history of the State's peremptory strikes in past cases" [Flowers v.
-Mississippi, 139 S.
-Ct.
-2228, 2243 (2019)].
-
-Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
-@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because strike data from past cases is hard to collect.
-Jury-selection records are often not readily available, because they are kept, if at all, by court staff somewhere other than the case file.
-Such records are also largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking effort to collect [e.g., @Craft2018].
-
-Moreover, strike data on past trials may be missing information not at random [e.g., @Baldus2001, 132-134].
-Possible reasons include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
-Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-
-When collecting strike data in multiple trials, prior studies have aggregated strikes across those trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
-For example, Eisenberg [-@Eisenberg2017, 334-339] used a 2-tail Fisher's exact test on the observed difference in peremptory strikes by race and gender in a non-random sample of thirty-five trials from 1997-2012 that resulted in death sentences (out of 63 such trials).
-@Gastwirth2013 has criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
-Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
-
-Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
-For exmpale, using jury selection data for 1992-2017 for Mississippi's Fifth Circuit Court District, Craft [-@Craft2018a, 9] applied a logistic regression model to analyze how much a prospective juror's race affected the odds of being struck by the state.
-For studies applying other modeling techniques to this Mississippi data, see @DeCamp2020 (propensity score matching) and @Dunn2021 (optimal feature selection, optimal classification trees).
-
-# Discussion
-
-Here, we use a Bayesian approach for detecting illegal bias for _Batson_ challenges based solely on an attorney's pattern of peremptory challenges in both a current trial and that attorney's strikes in past cases.
-First, we specify a model of the peremptory-strike process in the court of interest.
-Second, we assign an initial prior distribution for a bias parameter $b$.
-Third, we use an attorney's strike data from past trials, as  generated by the same peremptory-strike process, to estimate a posterior distribution for $b$.
-Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.
-
-In so doing, we make two main contributions.
-First, we use the power prior to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
-The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of some parameter of interest.
-The key insight is to raise the likelihood function based on the historical data to a power parameter [@Ibrahim2000; @Ibrahim2015].
-
-Second, we implement the analysis as prototype software that an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.
-
-# Model 1: Alternating Strikes
+## Statistical Procedure
 
 Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
-
 For any given case $i$ in which jury selection occurs, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
@@
 If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
-
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
 If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking someone who does not belong to that class.
-
-By adding one parameter $w$, we can measure the bias by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
+By adding one parameter $w$, we can measure the _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
 To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let $b = log(w)$.
@@
 \end{equation}
+<!-- \[b \sim \text{Normal}(0,2)\] -->
 
@@
 
-Next, we assign an initial prior:
+## Incorporating prior (historical) information
 
-\[b \sim \text{Normal}(0,2)\]
-
-Unlike @Kadane2018, who assigned $b$ a prior of $Uniform (-6,6)$, because the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior.
-
-First, for our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge as well as the initial burden of producing some relevant evidence of illegal bias (the _Batson_ prima facie case).
-Thus, if that challenging party in fact produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
-This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
-
-Second, we take that law as requiring, absent any data, the assumption that higher degrees of illegal bias (preference) are less likely than lower degrees of bias (preference).
-By using a normal (Gausian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
-
-Finally, we want to incorporate historical data on strikes in past cases and adjust the weight of that historical information on the posterior distribution of the bias parameter.
+In @Kadane2018, $b$ is estimated under Bayesian framework with a uniform prior from -6 to 6.
+In this paper, we also would like to incorporate the historical data on strikes in past cases and be able to adjust the weight of that historical information on the posterior distribution of the bias parameter.
 Accordingly, we introduce the power prior \@ref(eq:model3)
@@
 In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
+
+<!-- The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes.
+-->
 
@@
 
+The use of the power prior allows for flexibility in how much weight historical trail information receives in estimating the bias parameter for a current trial.
+[list reasons why.
+should this be in the introduction/discussion?]
+
 The posterior distribution does not have a closed form.
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998] and Geweke’s diagnostics [@Gewe1992].
+To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992].
 
-# Simulation
+## Model Performance on Simulated Data
+
+We propose the following steps to test for the presence of bias, and thus justification for bringing a _Bateson_ challenge.
+First, specify a model of the peremptory-strike process in the court of interest, as outlined above.
+Second, assign an initial prior distribution for a bias parameter $b$ (here we use a normal distribution with mean 0 and standard deviation of 2).
+Third, use an attorney's strike data from past trials, as  generated by the same peremptory-strike process, to estimate a posterior distribution for $b$.
+Finally, use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.
 
@@
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).
+We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two and three previous trials) using seven different values of the bias parameter for generating that historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).
 
@@
 If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias $b$ in the current case with a Black defendant.
-
 By using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
@@
 Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that parameter affected modeling accuracy.
-
 We considered 735 scenarios in total, i.e., the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data and 5 different values of the power prior weight parameter $a$.
@@
 We recorded the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
-
 In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$.
@@
 ```{r b80, echo=FALSE, fig.cap="80% credible interval"}
-
 library(ggplot2)
-
 # INSERT simulation and plot code
-
 load_path <- here::here("sim_resultc.csv")
-
 sims <- readr::read_csv(load_path)
 sims$a.f <- as.factor(sims$a)
-
 sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
@@
 sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
-
 p = ggplot()+
@@
 p
-
 ```
@@
 ```{r b90, echo=FALSE, fig.cap="90% credible interval"}
-
 # INSERT simulation and plot code
@@
 ```{r b95,  echo=FALSE,fig.cap="95% credible interval"}
-
 # INSERT simulation and plot code
@@
 
-These plots indicate that our model can detect strong bias and bias ($b = \{-3, -2, 2,3\}$).
+These plots indicate that our model can detect strong bias ($b = \{-3, -2, 2,3\}$).
 When bias is moderate ($b = \{1, -1\}$), the model can detect it if the historical data has a bias parameter in the same direction.
-If the bias parameter of historical data is in an opposite direction, the model may provide a wrong estimate of current bias parameter.
+If the bias parameter of historical data is in the opposite direction, the model may provide a wrong estimate of current bias parameter [is this true, even for the 80% CI?].
 
-# Model 2: Simultaneous Strikes
+## R Shiny Application to real data
 
-In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.  This may result in both sides using a strike each on the same potential juror.
+Even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in real time _during_ jury selection, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12] Following this suggestion, we describe here a prototype software application ("app") that implements the approach described above using `r R.version$version.string`[@R] and shiny [@shiny].
+We demonstrate the use of the app on some real data obtained from trials conducted in the state of Connecticut and compiled by the authors.
+For [privacy?] reasons, the attorney names have been changed, using (package name).
 
-For any given case $i$ in which jury selection occurs and a party uses $s$ number of strikes under this process, @Barrett2007 models the probability of a party striking at least $x$ cognizable-class members at random as following a hypergeometric distribution:
-
-$$
-Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}.
-$$
-where $x = 0, 1, 2, 3, .
-.
-.
-s$; $c_{i}$ denotes the number of cognizable class members in the venire subject to strike; and $m_{i}$ denotes the number of cognizable class non-members in the venire subject to strike.
-
-Here, we argue that a simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above, i.e, one party engages in an uninterrupted sequence of strikes against members of the venire.
-The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
-
-To show this, we conducted the following simulation study. [MORE]
-
-# The App Prototype
-
-Even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in real time _during_ jury selection, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
-
-Following this suggestion, we describe here a prototype software application ("app") that implements the approach described above using `r R.version$version.string`[@R] and shiny [@shiny].
-
-There is currently historical information of some prosecutors and defendants.
+There is currently historical information of some prosecutors and defendants contained within the app.
 More historical information would be updated when available.
@@
 ```
+
+[these are the fake names, right?]
 
@@
                 as.matrix()
-
 df_md <- df %>%
@@
                 as.matrix()
-
 sub_p <- subset("Brian P. Leaming",TRUE,dat0,"race")
@@
                                        posterior = "Prior")
-
 out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 1, niter = 110000, 
@@
                 posterior = "Posterior")
-
 dat <- rbind(d_p,d_d)
@@
                 facet_wrap(~party, nrow =2)
-
 pplot <- pplot  + theme_minimal() +
@@
                  geom_vline(xintercept = 0, color = "black", lwd=1.5)
-
  pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
@@
 ```{r brian2, echo=FALSE,fig.cap="Posterior Density Plot"}
-
 load_path2 <- here::here("app/mh_sampler_pp.cpp")
 load_path3 <- here::here("app/jury_data_cleaned_new.rds")
-
 sourceCpp(load_path2)
-
 dat0 <- readRDS(load_path3)
-
 df <- data.frame(round = c(1:10),
@@
                  party = rep(c("PP","PD"),5))
-
 extract_atny <- function(atny_name,pp,dat){
@@
                 as.matrix()
-
 sub_p <- subset("Brian P. Leaming",TRUE,dat0,"gender")
@@
                                        posterior = "Prior")
-
 out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 0.2, niter = 110000, 
@@
 
+# Discussion
+
+This paper makes three main contributions.
+First, we introduce historical trial information into the Bayesian framework proposed by [@Kadane2018].
+Second, we use the power prior to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
+The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of some parameter of interest.
+The key insight is to raise the likelihood function based on the historical data to a power parameter [@Ibrahim2000; @Ibrahim2015].
+Third, we implement the analysis as prototype software that an attorney could use _in real time_ during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.
+
+Prior work on statistical methods for _Batson_ challenges, has largely focused on strike data from a single case [e.g.
+@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because strike data from past cases is hard to collect.
+Jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking efforts to collect [e.g., @Craft2018].
+As a result, even where there is, for any given individual attorney or prosecutor's office, available strike data on past trials, that data may be missing strike data not at random [e.g., @Baldus2001, 132-134].
+Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies.
+Some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck, such as their race or gender.
+As a result, lawyers and researchers must resort to other data sources (e.g., voter registration rolls) to find that information.
+Finally, regardless of record-keeping policy, judges and court staff vary in how often they fully complete or maintain the required information.
+
+When collecting strike data in multiple trials, prior studies have aggregated strikes, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
+For example, Eisenberg [-@Eisenberg2017, 334-339] used a 2-tail Fisher's exact test on the observed difference in peremptory strikes by race and gender in a non-random sample of thirty-five trials from 1997-2012 that resulted in death sentences (out of 63 such trials).
+@Baldus2001 similarly analyzed 317 venires (out of 461) in Philadelphia capital murder cases from which juries were selected between 1981 and 1997.
+
+In contrast, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate.
+Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.
+
+Finally, even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]
+
+For a _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
+If so, the challenged party must produce one or more nondiscriminatory reasons for that strike.
+Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
+22.3].
+For similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v.
+Holloway, 209 Conn.
+636, 646 & n.4 (1989); State v.
+Chapman, 317 S.C.
+302, 305-06 (1995)], while other States have shifted the inquiry away from the striking party's motives to whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash.
+R.
+37(e)].
+
+Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
+Possible explanations include attorney fear that raising a _Batson_ challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021].
+Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
+
+By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
+For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v.
+Dretke, 545 U.S.
+231, 253 (2005); Riley v.
+Taylor, 277 F.3d 261, 280-81, 283 (3d Cir.
+2001); Cochran v.
+Herring, 43 F.3d 1404, 1412 (11th Cir.
+1995)].
+The US Supreme Court recently identified evidence for a _Batson_ challenge as including the "relevant history of the State's peremptory strikes in past cases" [Flowers v.
+Mississippi, 139 S.
+Ct.
+2228, 2243 (2019)].
+
+[paragraph here on how our model and application might help]
+
+### Section on data collection
+
+[Do we need more detail (here or earlier) on the collection of the historical data: how it was collected; where; when; total size.
+This could be supplemented too.
+I bring this up, because I feel like this is possibly one limitation: these data don't exist everywhere.
+It might be nice to put in a sentence stating that courts need to be better at collecting data.
+We could also mention that we can use the app without historical data maybe to link to comment above, we can say something like another reason why may not have been used like this before is that it is hard to get.
+Another thing our app does is provide a template for how this data could be collected, and stored in centralised database to facailitate its use for future Bateson challenges].
+
+### Extending the model: Simultaneous Strikes Example
+
+In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.
+This may result in both sides using a strike each on the same potential juror.
+For any given case $i$ in which jury selection occurs and a party uses $s$ number of strikes under this process, @Barrett2007 models the probability of a party striking at least $x$ cognizable-class members at random as following a hypergeometric distribution:
+
+$$
+Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}.
+$$
+
+where $x = 0, 1, 2, 3, .
+.
+.
+s$; $c_{i}$ denotes the number of cognizable class members in the venire subject to strike; and $m_{i}$ denotes the number of cognizable class non-members in the venire subject to strike.
+
+Here, we argue that a simultaneous-strikes process can be modeled as a special case of the  alternating-strikes process we outline above in which one party engages in an uninterrupted sequence of strikes against members of the venire.
+The premise is that, regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
+Results of simulations based on altering the sequence of strikes on estimates of the bias parameter are presented in the supplement.
+
 # Conclusion
 
-# CRediT Author Statement
+We propose a Bayesian approach to identifying bias in jury selection that incorporates historical trial information that is weighted by the use of a power prior.
+It is important to note that, although the primary purpose of this procedure is so that attorneys can include historical information, this is not necessary for its use.
+In the absence of historical information, we propose the use of a weakly informative prior [Gelman reference], centered on zero.
+This is appropriate because it places the 'burden of proof' on the current data to shift the estimate of bias away from zero.
+Finally, we have packaged our model in an R Shiny application to facilitate its use in real time.
+Given the wide variety of jury selection procedures that exist across the United States, future work aims to extend the model presented here to more complex procedures (e.g., "sequential strikes" in New York State [ref]).
 
-Sachin S.
-Pandya: Conceptualization, Investigation, Software, Writing; Xiaomeng Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Timothy Moore: Conceptualization, Supervision, Visualization.
-Based on categories in https://casrai.org/credit/].
-
+# CRedit Author Statement
+Sachin S. Pandya: Conceptualization; Xiaomeng Li: Conceptualization; [insert]; Timothy E. Moore: Conceptualization; [insert]. Based on categories in https://casrai.org/credit/].
 # References

### 2021-09-17 analysis/paper_draft_TEM.Rmd

@@
 If the bias parameter of historical data is in the opposite direction, the model may provide a wrong estimate of current bias parameter [is this true, even for the 80% CI?].
+
+We also show the coverage rate of using the three credible intervals below.
+The coverage rate is calculated as the proportion of correctly estimated bias among the 1000 simulations.
+The coverage rate of 80\% credible interval is shown in figure \@ref(fig:c80).
+
+```{r c80, echo=FALSE, fig.cap="80% credible interval"}
+ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,fill = C80),
+            color="black")+
+  facet_grid(h~a)+
+   scale_fill_gradient2('C80', 
+                        low = "darkred", mid = "white", 
+                        high = "blue", 
+                        breaks = c(0,0.5,0.7,0.9, 1), 
+                        limits = c(0,1),
+           labels = c("low", "", "moderate", "","high"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current d")+
+  ylab("Historical d")+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+
+```
+
+We can first take a look at the first column, for which the weight parameter of historical information is 0.1.
+It means we mostly rely on current information.
+When the bias parameter is 3, 2 or -2, -3, the coverage rate is high, very close to 1.
+Thus if the bias is obvious, we can always detect it using our model with 80\% credible interval.
+However, when the bias parameter is 1, the coverage rate drop below 0.8.
+If the bias is not strong, our model can detect it with a lower accuracy.
+This lower accuracy comes from the small sample size.
+We can find when we incorporate the historical information, and when the historical information is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
+On the other hand, when the bias parameter is 0, that is when there is no bias, the coverage rate is also around 80\%.
+It indicates that if we use the 80\% credible interval, we have around 20\% chance mistakenly concluding one is biased.
+If we want to be conservative and lower this mistake rate, it would be better to use a wider credible interval.
+The coverage rate of 90\% credible interval is shown in figure \@ref(fig:c90).
+
+```{r c90, echo=FALSE, fig.cap="90% credible interval"}
+ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,fill = C90),
+            color="black")+
+  facet_grid(h~a)+
+   scale_fill_gradient2('C90', 
+                        low = "darkred", mid = "white", 
+                        high = "blue", 
+                        breaks = c(0,0.5,0.7,0.9, 1), 
+                        limits = c(0,1),
+           labels = c("low", "", "moderate", "","high"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current d")+
+  ylab("Historical d")+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+
+```
+
+When the color is dark, it means a higher coverage rate.
+While the color is light, it shows a lower coverage rate.
+We can find when the bias parameter is 2, the color is lighter than using a 80\% credible interval, while the color is darker when the bias parameter is 0.
+Thus when using a 90\% credible interval, we lower the probability that we mistakenly concluding a bias, but the accuracy of detecting a bias is also lower when the bias parameter is 2.
+When the bias parameter is 3, the coverage rate is still high.
+
+```{r c95, echo=FALSE, fig.cap="90% credible interval"}
+ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,fill = C95),
+            color="black")+
+  facet_grid(h~a)+
+   scale_fill_gradient2('C95', 
+                        low = "darkred", mid = "white", 
+                        high = "blue", 
+                        breaks = c(0,0.5,0.7,0.9, 1), 
+                        limits = c(0,1),
+           labels = c("low", "", "moderate", "","high"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current d")+
+  ylab("Historical d")+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+
+```
+
+The coverage rate plot when using a 95\% credible interval is shown in figure \@ref(fig:c95).
+When we use a 95\% credible interval, the coverage rate when there is no bias is very close to 1.
+Thus we can ensure that we will not mistakenly conclude a bias.
+When the bias parameter is 3, the coverage rate is still high, thus we can still accurately detect the bias when it is strong.
+However, when the bias is moderate, like $b=2$ or $b=-2$, the coverage rate is only high when we have consistent historical information.
+Thus by observing the coverage rate plot, we conclude that our model can always accurately detect strong bias.
+As for moderate bias, more information is needed for accurate detection.
+The width of the credible interval depends on how conservative we want to be.
 

### 2021-09-20 analysis/paper_draft2.Rmd
updated

@@
 #output: distill::distill_article
-output: pdf_document
+output: 
+  pdf_document:
+    latex_engine: xelatex
+    number_sections: true
+    fig_width: 5
+    fig_height: 5
+    fig_caption: true
+header-includes: 
+  - \usepackage{amsmath}
+fontsize: 12pt
 bibliography: references.bib
@@
 We show how to use historical data on an attorney's use of peremptory strikes in past cases to estimate, in a current trial, how likely an attorney has struck prospective jurors because of cognizable-class membership.
-We also use of a _power prior_ to ajust the weight of such historical data in the analysis.
+We also use the _power prior_ to adjust the weight of such historical data in the analysis.
 We show the utility of our model on simulated data.
-Finally, we extend our approach by presenting a prototype software application to be used to detect bias in peremptory-challenge in real time during jury-selection.
-We demonstrate the app using actual historical strike data from a convenience sample of cases from the federal district court of Connecticut from YYYY to YYYY.
+Finally, we extend this with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection.
+We demonstrate the app using actual historical strike data from a convenience sample of cases from the federal district court of Connecticut.
 
@@
 
-Since the US Supreme Court's decision in _Batson v.
+Since _Batson v.
 Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class.
@@
 22.3].
-For similar challenges under State law, a few States accept proof not of motive but only that an "objective observer" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [e.g., Wash.
-R.
-37(e)].
+For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [e.g.
+@WashR37, subsec.
+(e); Cal.
+Code Civ.
+Proc.
+$\S$ 231.7(d)(1)].
 Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
 
-In any case, evidence supporting such challenges may include aggregate data on peremptory challenges in past cases.
-For example, in a State criminal case, evidence for a _Batson_ challenge can include the "relevant history of the State's peremptory strikes in past cases." [@flowers2019, 2243].
-Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
-@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because jury-selection records are often not readily available [@Grosso2017; @Wright2018] and otherwise require painstaking effort to collect [e.g., @APM2018].
+In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases.
+[@flowers2019, 2243; @WashR37, subsec.
+(g)(v); Cal.
+Code Civ.
+Proc.
+$\S$ 231.7(d)(3)(G)].
+Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
+Typically, they test for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017, 334-339; @Grosso2012, 1548; for discussion, see @Gastwirth2014; @Gastwirth2013].
+Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
+For different approaches using the same historical strike data from Mississippi, see @Craft2018a (logistic regression); @DeCamp2020 (propensity score matching); and @Dunn2021 (optimal feature selection, optimal classification trees).
 
-When collecting strike data in multiple trials, prior studies have aggregated strikes across trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
-[@Eisenberg2017, 334-339; see also @Gastwirth2014; @Gastwirth2013].
-Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
-For different approaches to the same historical strike data from from Mississippi's Fifth Circuit Court District (1992-2017), see @Craft2018a (logistic regression); @DeCamp2020 (propensity score matching); and @Dunn2021 (optimal feature selection, optimal classification trees).
-
-In this paper, we take a Bayesian approach to _Batson_ challenges [@Kadane2018] and extend it to incorporate historical strike data.
-In this approach, we specify a model of the peremptory-strike process in the court of interest that includes bias parameter $b$, to which we assign an initial prior distribution.
-Then, we use an attorney's strike data from past trials, as generated by the same peremptory-strike process, to estimate a posterior distribution for $b$.
-In so doing, we use the _power prior_ [@Ibrahim2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
+In this paper, we extend a Bayesian approach to _Batson_ challenges [@Kadane2018] to incorporate historical strike data.
+In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter $b$, to which we assign an initial prior distribution.
+Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for $b$.
+In so doing, we use the _power prior_ [@Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
 Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.
@@
 
-Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)
+Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation `\@ref(eq:model1)`
 
-$$
+\begin{equation}
 Pr(\delta_{ij} | b) = 
@@
   \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
-\end{cases}
-$$
+\end{cases} (\#eq:model1)
+\end{equation}
 
@@
 
-$$Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$
+\[
+Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
+\]
 
@@
 
-$$ 
-L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
-$$
+\begin{equation}  
+L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}(\#eq:model2)
+\end{equation}
+
 where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
@@
 
-$$\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)$$
+\begin{equation}  
+\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)
+\end{equation}
+
 where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
@@
 In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
-This departs from @Kadane2018, who assigned $b$ a prior of $Unif(-6,6)$.
+This departs from @Kadane2018, who assigned $b$ a prior of $\text{Uniform}(-6,6)$.
 We assume that the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior.
-For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge as well as the initial burden of producing some relevant evidence of illegal bias (the _Batson_ prima facie case).
-Thus, if that challenging party in fact produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
+For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
+Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
 This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
@@
 
-$$
-L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
-$$
+\begin{equation}
+L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
+\end{equation}
 
@@
 
-$$L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$$
+\begin{equation}
+L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}
+\end{equation}
 
-is the likelihood function of $b$ given historical data.
-$n_{0i}$ is the total number of jury selections in the historical data, and $n_{0j}$ is the total number of peremptory strikes in the historical data.
-$\delta_{0ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class" in the historical trials.
+is the likelihood function of $b$ given historical data; $n_{0i}$ is the total number of jury selections (trials) in the historical data; $n_{0j}$ is the total number of peremptory strikes in the historical data; and $\delta_{0ij}$ denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.
 
@@
 To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$.
-As introduced above, $b>0$ ($b<0$) represents a bias against (preference) for a prospective juror within (outside) the cognizable class, while $b=0$ denotes neither bias or preference.
-A greater absolute value of $b$ represents greater bias against or preference for the cognizable class.
+As introduced above, $b>0$ ($b<0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b=0$ denotes no bias.
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).
+We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical b_$ $= {-3,-2,-1,0,1,2,3}$).
 
@@
 To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias $b$ in the current case with a Black defendant.
+If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias ($b$) in the current case with a Black defendant.
 
@@
 
-The following figures depict the coverage rates of $b$ under the 80\% credible interval (Figure \@ref(fig:b80)), under the 90\% credible interval (Figure \@ref(fig:b90)), and under the 95\% credible interval (Figure \@ref(fig:b95)).
+Figures 1, 2, and 3 depict the coverage rates of $b$ under an 80\%, 90\%, 95\% credible interval, respectively.
 
@@
 
-library(ggplot2)
-
 p80 <- ggplot()+
@@
 
-For any given case $i$ in which jury selection occurs and a party uses $s$ number of strikes under this process, @Barrett2007 models the probability of a party striking at least $x$ cognizable-class members at random as following a hypergeometric distribution:
-
-$$Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}$$
-
-where $x = 0, 1, 2, 3, .
-.
-.
-s$; $c_{i}$ denotes the number of cognizable class members in the venire subject to strike; and $m_{i}$ denotes the number of cognizable class non-members in the venire subject to strike.
-
 Here, we argue that a simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above, i.e, one party engages in an uninterrupted sequence of strikes against members of the venire.
@@
 
-load_path_b1 <- here::here("b1.csv")
-load_path_b2 <- here::here("b2.csv")
-load_path_b3 <- here::here("b3.csv")
+load_path_b1 <- here::here("data-sim", "b1.csv")
+load_path_b2 <- here::here("data-sim", "b2.csv")
+load_path_b3 <- here::here("data-sim", "b3.csv")
 
@@
 
-b1.t <- b1 %>% select(-X) %>%
-  t() %>% data.frame() %>% 
+library(dplyr)
+
+b1.t <- b1 %>% 
+  select(-X) %>%
+  t() %>% 
+  data.frame() %>% 
   dplyr::rename_with(~b1$X) %>%
@@
 
-b2.t <- b2 %>% select(-X) %>%
-  t() %>% data.frame() %>% 
+b2.t <- b2 %>% 
+  select(-X) %>%
+  t() %>% 
+  data.frame() %>% 
   dplyr::rename_with(~b2$X) %>%
@@
 
-b3.t <- b3 %>% select(-X) %>%
-  t() %>% data.frame() %>% 
+b3.t <- b3 %>% 
+  select(-X) %>%
+  t() %>% 
+  data.frame() %>% 
   dplyr::rename_with(~b2$X) %>%
@@
 b.com <- rbind(b1.t, b2.t, b3.t)
-b.com$sig <- ifelse(b.com$`2.5%`>0, "Bias", 
+b.com$sig <- ifelse(b.com$`2.5%`> 0, "Bias", 
                                   "No Bias")
@@
                      b.num = 1)
+
 b2.mean <-data.frame(b_s = mean(b2.t$b_s), 
@@
                      b.num = 2)
+
 b3.mean <-data.frame(b_s = mean(b3.t$b_s), 
@@
 
-simplot <- ggplot(data = b.com, aes(x = sim.n, y = b_s, color = sig))+
+library(ggplot2)
+
+simplot <- ggplot(data = b.com, 
+                  aes(x = sim.n, y = b_s, color = sig))+
   facet_grid(~b)+
-  geom_point(size = 3)+
+  geom_point(size = 2)+
   geom_errorbar(aes(x = sim.n, ymin = `2.5%`, 
@@
 To detect racial bias in the use of peremptory strikes using both the strike tally and historical strike data, we set the cognizable-class to "Race"; select the names of the prosecutor and defense attorney from their respective pull-down menus; and set the weight setting to "Equal".
-The app will in turn display the posterior density plots in Figure \@ref(fig:brian).
+The app will in turn display the posterior density plots in Figure `\@ref(fig:brian)`.
 
-```{r brian, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
+```{r brian, eval = FALSE, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
 library(dplyr)
@@
 Now suppose that the strike tally indicates the pattern of strikes by these attorneys against women.
-If so, we can select "gender" as the cognizable class, and the density plots becomes Figure \@ref(fig:brian2).
+If so, we can select "gender" as the cognizable class, and the density plots becomes Figure `\@ref(fig:brian2)`.
 
-```{r brian2, fig.cap="Posterior Density Plots, a = 1, cognitive class = gender"}
+```{r brian2, eval = FALSE, fig.cap="Posterior Density Plots, a = 1, cognitive class = gender"}
 
@@
 
-Finally, if we give low weight on historical information, the density plot will be as shown in figure \@ref(fig:brian3).
+Finally, if we give low weight on historical information, the density plot will be as shown in figure `\@ref(fig:brian3)`.
 
-```{r brian3, echo=FALSE,fig.cap="Posterior Density Plot, a = 0.2"}
+```{r brian3, eval = FALSE, fig.cap="Posterior Density Plot, a = 0.2"}
 load_path2 <- here::here("app/mh_sampler_pp.cpp")
@@
 
+Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
+@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because jury-selection records are often not readily available [@Grosso2017; @Wright2018] and otherwise require painstaking effort to collect [e.g., @APM2018].
+
 # Conclusion

### 2021-09-20 analysis/paper_draft2.Rmd
created separate file for sim_data function; added new coverage plots; revisions to discussion of app

@@
     number_sections: true
-    fig_width: 5
-    fig_height: 5
-    fig_caption: true
 header-includes: 
@@
 Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class.
-The party bringing a _Batson_ challenge bears the burden of proving such illegal bias [@LaFave2020, sec.
-22.3].
-For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [e.g.
-@WashR37, subsec.
-(e); Cal.
-Code Civ.
-Proc.
-$\S$ 231.7(d)(1)].
+The party bringing a _Batson_ challenge bears the burden of proving such illegal bias [@LaFave2020, $\S$ 22.3].
+For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [Wash.
+Rule 37(e); Calif.
+Code of Civil Procedure $\S$ 231.7(d)(1)].
 Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
@@
 In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases.
-[@flowers2019, 2243; @WashR37, subsec.
-(g)(v); Cal.
-Code Civ.
-Proc.
-$\S$ 231.7(d)(3)(G)].
+( _Flowers v.
+Mississippi_ -@flowers2019, p.
+2243); see also Wash.
+R.
+37(g)(v) -@WashR37; Calif.
+Code of Civil Procedure $\S$ 231.7(d)(3)(G)).
 Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
 Typically, they test for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017, 334-339; @Grosso2012, 1548; for discussion, see @Gastwirth2014; @Gastwirth2013].
-Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
-For different approaches using the same historical strike data from Mississippi, see @Craft2018a (logistic regression); @DeCamp2020 (propensity score matching); and @Dunn2021 (optimal feature selection, optimal classification trees).
-
-In this paper, we extend a Bayesian approach to _Batson_ challenges [@Kadane2018] to incorporate historical strike data.
+Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
+For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2020; and @Dunn2021.
+
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2018] to incorporate historical strike data.
 In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter $b$, to which we assign an initial prior distribution.
@@
 Here, we demonstrate this approach, including the results of simulation studies to test model performance.
-Then, we present a software prototype that would let an attorney use the same approach in using historical strike data to help detect bias in peremptory-challenge use in real time during jury selection.
-
-# Method
+Then, we present a software prototype that encodes the same approach with actual historial strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
+
+# Methods
 
@@
 Then, the parties exercise their strikes on those potential jurors, either submitting strikes simultaneously or in an alternating sequence.
-Once all strikes are used or waived, the jurors are seated based on the order in which they were selected or at random.
+Once all strikes are used or waived, the remaining prospective jurors are assigned to seats on the jury.
 
@@
 
-By adding one parameter $w$, we can measure the _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
+By adding one parameter $w$, we can measure _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
 To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let $b = log(w)$.
 
-Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation `\@ref(eq:model1)`
+Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is such that:
 
@@
   \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
-  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
-\end{cases} (\#eq:model1)
+  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  \label{eq:model1}
+\end{cases} 
 \end{equation}
 
-This is equivalent to
-
-\[
+This Equation \eqref{eq:model1} is equivalent to
+
+\begin{equation}
 Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
-\]
+\label{eq:model2}
+\end{equation}
 
@@
 
-The likelihood function of $b$ is shown in equation \@ref(eq:model2)
+The likelihood function of $b$ is
 
 \begin{equation}  
-L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}(\#eq:model2)
+L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
+\label{eq:model3}
 \end{equation}
@@
 
-We incorporate the historical data on strikes in past cases and allow for adjustment of the weight of that historical information on the posterior distribution of the bias parameter.
+We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter.
 To do this, we introduce the power prior:
@@
 \pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)
+\label{eq:powerprior}
 \end{equation}
@@
 In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
-This departs from @Kadane2018, who assigned $b$ a prior of $\text{Uniform}(-6,6)$.
-We assume that the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior.
+Unlike @Kadane2018, who assigned $b$ a prior of $\text{Uniform}(-6,6)$, we assume that the law for _Batson_ and similar challenges entails a weakly-informative prior.
 For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
@@
 This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
-
-Moreover, we take that law as requiring, absent any data, the assumption that higher degrees of illegal bias (preference) are less likely than lower degrees of bias (preference).
-By using a normal (Gaussian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
-
-After including the historical information through the power prior, the posterior distribution of $b$  is proportional to the product of likelihood function of $b$ and the power prior of $b$ as shown in \@ref(eq:model4)
+Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias.
+For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
+
+After including the historical information through the power prior, the posterior distribution of $b$  is proportional to the product of likelihood function of $b$ and the power prior of $b$ is
 
 \begin{equation}
-L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
+L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
+\label{eq:model4}
 \end{equation}
@@
 L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}
+\label{eq:model5}
 \end{equation}
@@
 
-To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$.
+To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" b $= \{-3,-2,-1,0,1,2,3\}$.
 As introduced above, $b>0$ ($b<0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b=0$ denotes no bias.
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical b_$ $= {-3,-2,-1,0,1,2,3}$).
+We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ("historical b" $= \{-3,-2,-1,0,1,2,3\}$).
 
@@
 
-Because we use the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$.
-If $a = 1$, the historical strike data is equally weighted with the present data on strikes.
+Because we use the power prior, we can control how much we account for the historical strike data by modifying the weight parameter $a$.
+If $a = 1$, the historical strike data is equally weighted with the data on strikes in the current trial.
 If $a < 1$, the historical data matters proportionally less.
 We can assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
-Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that parameter affected modeling accuracy.
+Accordingly, in the simulation study, we tried different values of the weight parameter ($a = \{0.1,0.3,0.5,0.7,1\}$) to show how that parameter affected modeling accuracy.
 
@@
 We recorded the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
-
 In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$.
-For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$.
-Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
-
-Figures 1, 2, and 3 depict the coverage rates of $b$ under an 80\%, 90\%, 95\% credible interval, respectively.
-
-```{r b80, fig.cap="80% credible interval"}
+
+```{r b80}
 
@@
         legend.text = element_text(size = 14))
-p80
-
-```
-
-```{r b90, fig.cap="90% credible interval"}
+#p80
+
+```
+
+```{r b90}
 
@@
         legend.text = element_text(size = 14))
-p90
-
-```
-
-```{r b95, fig.cap="95% credible interval"}
-
-# INSERT simulation and plot code
+#p90
+
+```
+
+```{r b95 }
+
 sims$a.f <- as.factor(sims$a)
@@
 
-p<- ggplot()+
+p95 <- ggplot()+
   geom_tile(data = sims, aes(x = d, y = d_h,
@@
 
-p90
-
-```
-
-In these figures, _historical b_ denotes the bias parameter for the generated historical data, and current $b$ means the bias parameter for the current data.
+#p95
+
+```
+
+For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$.
+Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
+
+The following figures depict the coverage rates of $b$ under an 80\%, 90\%, 95\% credible interval, respectively.
+In these figures, _historical b_ denotes the bias parameter for the generated historical strike data, and current $b$ means the bias parameter for the current data.
 The values on the right side represent the size of the historical data.
 The values on the top of the plots represent the power prior weight parameter $a$.
-Blue means the credible interval for $b$ excludes zero.
-
-These plots indicate that our model can detect strong bias and bias ($b = \{-3, -2, 2,3\}$).
-When bias is moderate ($b = \{1, -1\}$), the model can detect it if the historical data has a bias parameter in the same direction.
-If the bias parameter of historical data is in an opposite direction, the model may provide a wrong estimate of current bias parameter.
+For these figures, a darker color denotes a higher coverage rate, and vice versa.
+
+```{r c80, fig.cap="80% credible interval"}
+
+c80plot <- ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,fill = C80),
+            color="black")+
+  facet_grid(h~a)+
+   scale_fill_gradient2('C80', 
+                        low = "darkred", mid = "white", 
+                        high = "blue", 
+                        breaks = c(0,0.5,0.7,0.9, 1), 
+                        limits = c(0,1),
+           labels = c("low", "", "moderate", "","high"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  coord_equal()+
+  theme_classic()+
+  theme(legend.pos = "none",
+        axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        #legend.title.align = 0.05, 
+        #legend.title = element_text(size = 16), 
+        #legend.text = element_text(size = 14)
+        )
+
+c80plot
+
+```
+
+Figure 1 depicts the coverage rate with an 80\% credible interval.
+As the first column ($a=0.1$) depicts, when we place low weight on the historical strike data, we mostly rely on current information.
+When the bias parameter is 3, 2 or -2, -3, the coverage rate is high, very close to 1.
+
+However, when the bias parameter ("current b") is 1, the coverage rate drops below 0.8.
+If the bias is not strong, our model can detect it with a lower accuracy.
+However, as we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
+When the bias parameter is 0 (no bias), however, the coverage rate is also around 80\%.
+Thus, if we use the 80\% credible interval, we have around 20\% chance of erroneously detecting non-zero bias.
+
+```{r c90fig, fig.cap="90% credible interval"}
+
+c90plot <- ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,fill = C90),
+            color="black")+
+  facet_grid(h~a)+
+   scale_fill_gradient2('C90', 
+                        low = "darkred", mid = "white", 
+                        high = "blue", 
+                        breaks = c(0,0.5,0.7,0.9, 1), 
+                        limits = c(0,1),
+           labels = c("low", "", "moderate", "","high"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  coord_equal()+
+  theme_classic()+
+  theme(legend.pos = "none",
+        axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        #legend.title.align = 0.05, 
+        #legend.title = element_text(size = 16), 
+        #legend.text = element_text(size = 14)
+        )
+
+c90plot
+
+```
+
+Using a wider credible interval lowers this chance of error.
+For example, with a 90\% credible interval (Figure 2), when $b = 2$, the color is lighter (lower coverage rate) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
+
+Thus, with a 90\% credible interval, we are less likely to erroneously infer bias than with an 80\% credible interval.
+The tradeoff is that, with a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
+When the bias parameter is 3, the coverage rate is still high.
+
+```{r c95, fig.cap="95% credible interval"}
+c95plot <- ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,fill = C95),
+            color="black")+
+  facet_grid(h~a)+
+   scale_fill_gradient2('C95', 
+                        low = "darkred", mid = "white", 
+                        high = "blue", 
+                        breaks = c(0,0.5,0.7,0.9, 1), 
+                        limits = c(0,1),
+           labels = c("low", "", "moderate", "","high"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  coord_equal()+
+  theme_classic()+
+  theme(legend.pos = "none",
+        axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
+        #legend.title.align = 0.05, 
+        #legend.title = element_text(size = 16), 
+        #legend.text = element_text(size = 14)
+        )
+
+c95plot
+
+```
+
+Figure 3 depicts coverage rates when using a 95\% credible interval.
+Here, given $b = 0$, the coverage rate is very close to 1.
+This implies that we are very unlikely to conclude a bias when there is none.
+When the bias parameter is 3, the coverage rate is still high.
+This implies that we can still accurately detect the bias when it is strong.
+However, when the bias is moderate ($b = \{2, -2\}$, the coverage rate is only high when we have consistent historical information.
+
+Overall, these simulation results indicate that our model can always accurately detect strong bias.
+As for moderate bias, detection accuracy increases with more historical strike information.
+The choice of credible interval ultimately depends on how conservative we want to be when inferring bias.
 
@@
 
-In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.  This may result in both sides using a strike each on the same potential juror.
-
-Here, we argue that a simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above, i.e, one party engages in an uninterrupted sequence of strikes against members of the venire.
+In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.
+This simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
 The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
 
-To show this, we conducted the following simulation study.
-We first generate a trial and then shuffle the order of the strikes to generate 50 trials with the same proportion of struck cognizable class members but different orders.
-We then use the model to fit the 50 trials to see if the estimated bias parameter are close to each other.
-We conduct the simulation study under three scenarios from moderate bias to strong bias, that is $b=1,2,3$.
-The results are shown in figure \@ref(fig:simltns).
-
-```{r simltns, fig.cap="simultaneous case", eval = FALSE}
-
-#library(conflicted)
-#conflict_prefer("select", "dplyr")
-#conflict_prefer("filter", "dplyr")
-
-# SCRIPT FOR GENERATING b1.csv, b2.csv, b3.csv
-
-# generate data
+To test this premise, we conducted the following simulation study.
+We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 32 prospective jurors, 10 of which were member of a cognizable-class (e.g., Black jurors).
+Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
+We then used the model to fit the 50 trials to see if the estimated bias parameters are close to each other.
+We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
+Figure 4 depicts the results.
+
+```{r simgen, eval = FALSE}
+
+# generate simulated data and output b1.csv, b2.csv, b3.csv
 
 source("R/simultaneous1.R")
+
+```
+
+```{r simltns, fig.cap="Simulation Results - simultaneous case"}
 
@@
                      b = "1", 
-                     sig = "No Bias", 
+                     sig = "Bias",  # FIX: code as function
                      b.num = 1)
@@
                      b = "2",
-                     sig = "Bias", 
+                     sig = "Bias", # FIX: code as function
                      b.num = 2)
@@
                      b = "3",
-                     sig = "Bias", 
+                     sig = "Bias", # FIX: code as function
                      b.num = 3)
@@
   geom_hline(yintercept = 0, col = "black", lwd = 1, lty = 1) +
-  geom_vline(xintercept = 51.25, col = "black", lwd = 1, lty = 2) +
+  geom_vline(xintercept = 51.25, col = "black", lwd = .5, lty = 2) +
   xlab("Simulation")+
-  scale_color_manual("Bias", values = c("blue", "grey"))+
+  scale_color_manual("Bias", values = c("blue", "red"))+
   ylab("Posterior Bias Value")+
@@
   theme_classic()+
-  theme(axis.title = element_text(size = 16), 
+  theme(legend.position = "none",
+        axis.title = element_text(size = 16), 
         axis.text.x = element_text(size = 12), 
@@
 
-The points represent the estimated posterior mean and the lines represent the 95\% credible interval.
-The blue line means the credible interval does not include zero, thus indicating a bias, while the grey line means the credible interval includes zero and indicating no bias.
-We can find the estimated bias parameter of the 50 trials with different orders are close to each other for all of the three scenarios, indicating the order does not influence the estimate of the bias parameter.
+The blue points represent the estimated posterior mean for the  bias parameter and the blue lines represent the 95\% credible interval.
+The red vertical line indicates the true value of the bias parameter.
+
+We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
+None of their credible intervals include zero.
+We therefore infer that the order of strikes does not influence the estimate of the bias parameter.
 This confirms that the simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process.
 
-# The Software Prototype
-
-We describe here a prototype software application ("app") that implements the approach described above.
+## The Software Prototype
+
+We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
 We built this app using `r R.version$version.string`[@R] and shiny [@shiny].
-To demonstrate how this app might be used by lawyers in real cases, we loaded this app with real strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
-
-To collect the data, one of us filed a request with the U.S.
+
+To show how this app might be used by lawyers in real cases, we loaded this app with real strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
+To collect this historical strike data, one of us filed a request with the U.S.
 District Court of Connecticut based on 28 U.S.C.
 $\S$ 1868.
-Thereafter, we received copies (pdf files) of certain jury-selection records associated with X trials in that court during this period.
+Thereafter, we received copies (pdf files) of certain jury-selection records associated with [INSERT] trials in that court during this period.
 These included strike sheets that indicated the ID number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
@@
 
-With the app, the user uses the pull-down menus to select the cognizable class (top left) and enters by hand the strike information in the case before them in the _strike tally_ table (bottom left), either in real time or immediately after jury selection concludes.
-For demonstration purposes, the prototype app comes pre-loaded with a completed strike tally with values that can be changed and rows that can be added or deleted (Table \@ref(tab:inpts)).
-In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prosepective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
+With the app, the user uses the pull-down menus to select the cognizable class (top left) and enters by hand the strike information in the case before them in the _strike tally_ table (bottom left).
+For demonstration purposes, the prototype app comes pre-loaded with a completed strike tally with values that can be changed and rows that can be added or deleted.
+In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
+
+For the prototype app, we set two cognizable-class options: race and gender.
+This was because the standard juror questionnaire did not ask about other possible cognizable classes.
+The data also did not expressly indicate the race or gender of the criminal defendant(s).
+
+In the default setting, the pull-down menus for prosecutor and defense are set to "None".
+As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b = N(0,2)$).
 
@@
 
-For the prototype app, we set two cognizable-class options: race and gender.
-This was simply because the historical data only contained information on prospective juror race and gender, itself a function of data collected as part of the standard juror questionnaire.
-The data did not expressly indicate the race or gender of the criminal defendant(s).
-
-In the default setting, the pull-down menus for prosecutor and defense are set to "None".
-In that setting, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b = N(0,2)$).
-The main screen accordingly displays two graphs -- one for the prosecution and the defense - that each depict the prior and posterior density plots for the bias parameter.
-
-To use historical strike data, the user selects a name of the prosecutor or defense attorney from the pull-down menus.
+To illustrate, suppose the strike tally in Table \@ref(tab:inpts) represents the pattern of strikes in the present case with prosecutor "X" and defense attorney "Y" (both aliases for actual attorneys in the historical strike data).
+After the user enters this strike tally, the app displays two graphs -- one for the prosecution and the defense.
+Each graph depicts the prior density plot (colored grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure 5).
+
+```{r figapp1, fig.cap = "Density Plots without Historical Strike Data"}
+
+load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot1.png")
+
+knitr::include_graphics(load_path)
+
+```
+
+The two vertical dotted lines depict the 80\% credible interval. Here, that credible interval includes zero, indicating no credible to infer bias, given the current strike tally.
+
+To use historical strike data, the user selects a name of prosecutor or defense attorney from the pull-down menus.
 If an attorney's name cannot be found, the app has no historical strike data for that attorney.
-Once selected, the prior density plot (colored grey) and the posterior density plot (blue and red for defense and prosecution, respectively) automatically updates to account for the pre-loaded historical strike data for that attorney.
+Once selected, the prior and posterior density plots automatically update to account for the pre-loaded historical strike data for that attorney.
 For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information ($a = 1$).
-The user has two other options: half weight of historical information ($a = 0.5$) and low weight of historical information ($a = 0.2$).
-
-To illustrate, suppose the strike tally in Table \@ref(tab:inpts) represents the pattern of strikes in the present case with prosecutor "X" and defense attorney "Y".
-To detect racial bias in the use of peremptory strikes using both the strike tally and historical strike data, we set the cognizable-class to "Race"; select the names of the prosecutor and defense attorney from their respective pull-down menus; and set the weight setting to "Equal".
-The app will in turn display the posterior density plots in Figure `\@ref(fig:brian)`.
-
-```{r brian, eval = FALSE, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
+The user has two other options: half weight of historical information ($a = 0.5$) and minimal weight of historical information ($a = 0.2$).
+
+In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to "Equal".
+The density plots update accordingly (Figure \@ref(fig:figapp2)).
+
+```{r figapp2, eval = FALSE, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
 library(dplyr)
@@
 
-The grey curve represents the probability density of the prior, the blue curve represents the posterior probability density of the defense attorney and the red curve represents the posterior probability density of the prosecutor.
-The two vertical dotted lines depict the 80\% credible interval.
-Most of both the prior and posterior density curve are on the right side of zero, and the 80\% credible interval is also on the right side of zero.
 Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
 
-Now suppose that the strike tally indicates the pattern of strikes by these attorneys against women.
-If so, we can select "gender" as the cognizable class, and the density plots becomes Figure `\@ref(fig:brian2)`.
+Next, suppose the same scenario, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
+If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:brian2)).
 
@@
 
-Finally, if we give low weight on historical information, the density plot will be as shown in figure `\@ref(fig:brian3)`.
+Finally, if we assign minimal weight to the historical strike data, the density plot will be as shown in figure \@ref(fig:brian3).
 
@@
 
-In a typical _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status.
-If so, the burden shifts to the challenged party to produce one or more nondiscriminatory reasons for that strike.
-Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec.
-22.3].
-
-Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018].
-Possible explanations include attorney fear that even raising a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
-Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.
-
-One virtue of incorporating the power prior: Strike data on past trials may be missing information not at random [e.g., @Baldus2001, 132-134].
-Possible reasons include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender), thus forcing lawyers and researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
+Statistical methods for _Batson_ and similar changes are intended to measure how much the prospective jurors' race, gender, or other cognizable-class membership matters to whether attorneys use peremptory challenges against them.
+The ideal data, therefore, is, for each attorney, not just who they struck in the current trial, but a complete history of who they struck in all their past cases.
+
+The actual available data, however, is not that ideal, because jury-selection records in past cases are hard to get because of court practices and otherwise require painstaking effort to compile [@Grosso2017; @Wright2018].
+As a result, historical strike data may be missing not at random.
+Possible causes include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender).
+This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information [e.g., @Baldus2001, 132-134].
 Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
 
-## Costs and Benefits of Bayesian Approach
-
-Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g.
-@Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because jury-selection records are often not readily available [@Grosso2017; @Wright2018] and otherwise require painstaking effort to collect [e.g., @APM2018].
-
-# Conclusion
+The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter.
+By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
+And as our prototype app shows, if the historical strike data is collected and loaded into the app in advance, this approach can be easily used by attorneys and others in real time during jury selection.
+Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
+Whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges is a question worth studying further.
+
+At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data, given the choice of how to model a court's particular peremptory-challenge process.
+For example, given the model we used, we required data on the cognitive-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
+Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that applied to other court's procedure (as we showed above).
+If not, that attorney's historical strike data in one court could not used to estimate the posterior of that attorney's bias parameter, given a current trial in the other court.
+
+These data challenges, however, might decrease over time, as attorneys and others realize that more complete historical strike data may make it easier, with the approach described here, to credibly infer bias, or rebut that inference, under _Batson_ and similar law.
 
@@
 
-Sachin S.
-Pandya: Conceptualization, Investigation, Software, Writing; Xiaomeng Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Timothy Moore: Conceptualization, Supervision, Visualization.
+Pandya: Conceptualization, Investigation, Software, Writing; 
+Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Moore: Conceptualization, Supervision, Visualization.
+
 Based on categories in https://casrai.org/credit/].

### 2021-09-20 analysis/R/noseq_simulation.R

@@ -1,25 +1,9 @@
 library(Rcpp)
 library(ggplot2)
 library(dplyr)
-sourceCpp('prototype/mh_sampler.cpp')
-sourceCpp('prototype/mh_sampler_pp.cpp')
-sim_data <- function(d,round,total,num_cog){
-  df0 <- data.frame(round = c(1:round), 
-                    num_cog = NA, 
-                    total = NA,
-                    cog = NA)
-  df0$num_cog[1] = num_cog
-  df0$total[1] = total
-  p_cog <- num_cog*exp(d)/(num_cog*exp(d)+total-num_cog)
-  df0$cog[1] = rbinom(1,1,p_cog)
-  for(i in 2:round){
-    df0$num_cog[i] = df0$num_cog[i-1]-df0$cog[i-1]
-    df0$total[i] = df0$total[i-1]-1
-    p_cog <- df0$num_cog[i]*exp(d)/(df0$num_cog[i]*exp(d)+df0$total[i]-df0$num_cog[i])
-    df0$cog[i] = rbinom(1,1,p_cog)
-  }
-  return(df0)
-}
+sourceCpp('R/mh_sampler.cpp')
+sourceCpp('R/mh_sampler_pp.cpp')
+source('R/sim_data.R')
 
 ### simulation repeated 1000 times ###
 ### round = 30
@@ -140,6 +124,3 @@ ggplot()+
         legend.title.align = 0.05, 
         legend.title = element_text(size = 16), 
         legend.text = element_text(size = 14))
-
-
-

### 2021-09-20 analysis/R/sim_data.R

@@ -1,0 +1,17 @@
+sim_data <- function(d,round,total,num_cog){
+  df0 <- data.frame(round = c(1:round), 
+                    num_cog = NA, 
+                    total = NA,
+                    cog = NA)
+  df0$num_cog[1] = num_cog
+  df0$total[1] = total
+  p_cog <- num_cog*exp(d)/(num_cog*exp(d)+total-num_cog)
+  df0$cog[1] = rbinom(1,1,p_cog)
+  for(i in 2:round){
+    df0$num_cog[i] = df0$num_cog[i-1]-df0$cog[i-1]
+    df0$total[i] = df0$total[i-1]-1
+    p_cog <- df0$num_cog[i]*exp(d)/(df0$num_cog[i]*exp(d)+df0$total[i]-df0$num_cog[i])
+    df0$cog[i] = rbinom(1,1,p_cog)
+  }
+  return(df0)
+}

### 2021-09-20 analysis/R/simultaneous1.R

@@ -1,5 +1,6 @@
 library(Rcpp)
-sourceCpp('mh_sampler.cpp')
+sourceCpp('R/mh_sampler.cpp')
+source('R/sim_data.R')
 ##### b = 1  #####
 a1 <- sim_data(1,15,32,10)
 n_1 = sum(a1$cog)
@@ -27,7 +28,8 @@ idx1 = bs_l[1,]<3
 idx2 = bs_l[2,]>3
 sum(idx1&idx2)/50
 b_s1 <- rbind(b_s,bs_l)
-write.csv(b_s1,'b1.csv')
+save_path_b1 <- here::here("data-sim","b1.csv")
+write.csv(b_s1, save_path_b1)
 
 ##### b=2 #####
 a1 <- sim_data(2,15,32,10)
@@ -56,7 +58,8 @@ idx1 = bs_l[1,]<3
 idx2 = bs_l[2,]>3
 sum(idx1&idx2)/50
 b_s1 <- rbind(b_s,bs_l)
-write.csv(b_s1,'b2.csv')
+save_path_b2 <- here::here("data-sim","b2.csv")
+write.csv(b_s1, save_path_b2)
 
 ###### b=3  ######
 a1 <- sim_data(3,15,32,10)
@@ -85,5 +88,5 @@ idx1 = bs_l[1,]<3
 idx2 = bs_l[2,]>3
 sum(idx1&idx2)/50
 b_s1 <- rbind(b_s,bs_l)
-write.csv(b_s1,'b3.csv')
-
+save_path_b3 <- here::here("data-sim","b3.csv")
+write.csv(b_s1,save_path_b3 )

### 2021-09-21 analysis/paper_draft2.Rmd
updated paper draft

@@
     number_sections: true
+    fig_width: 5
+    fig_height: 4
 header-includes: 
@@
 In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases.
-( _Flowers v.
+(_Flowers v.
 Mississippi_ -@flowers2019, p.
-2243); see also Wash.
-R.
-37(g)(v) -@WashR37; Calif.
+2243; Wash.
+Rule 37(g)(v) -@WashR37; Calif.
 Code of Civil Procedure $\S$ 231.7(d)(3)(G)).
 Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
-Typically, they test for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017, 334-339; @Grosso2012, 1548; for discussion, see @Gastwirth2014; @Gastwirth2013].
+Typically, they test for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014; @Gastwirth2013].
 Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
-For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2020; and @Dunn2021.
+For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2021.
 
@@
 Here, we demonstrate this approach, including the results of simulation studies to test model performance.
-Then, we present a software prototype that encodes the same approach with actual historial strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
+Then, we present a software prototype that encodes the same approach with actual historical strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
 
@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992; @Gelman1998;@Geweke1992].
+To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992; @Gelman1998;@Geweke1992].
 
@@
 
-# INSERT simulation and plot code
 sims$a.f <- as.factor(sims$a)
@@
 For example, with a 90\% credible interval (Figure 2), when $b = 2$, the color is lighter (lower coverage rate) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
-
-Thus, with a 90\% credible interval, we are less likely to erroneously infer bias than with an 80\% credible interval.
-The tradeoff is that, with a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
-When the bias parameter is 3, the coverage rate is still high.
+The tradeoff is lower accuracy.
+With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
+When the bias parameter is 3, however, the coverage rate is still high.
 
@@
 
-To test this premise, we conducted the following simulation study.
-We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 32 prospective jurors, 10 of which were member of a cognizable-class (e.g., Black jurors).
-Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
-We then used the model to fit the 50 trials to see if the estimated bias parameters are close to each other.
-We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
-Figure 4 depicts the results.
-
 ```{r simgen, eval = FALSE}
@@
 
-```{r simltns, fig.cap="Simulation Results - simultaneous case"}
+```{r simltns, fig.align='center', fig.cap="Shuffled Sequence Simulation", out.width='50%'}
 
@@
   theme(legend.position = "none",
-        axis.title = element_text(size = 16), 
+        #axis.title = element_text(size = 16), 
         axis.text.x = element_text(size = 12), 
@@
 ```
+
+To test this premise, we conducted the following simulation study.
+We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 32 prospective jurors, 10 of which were member of a cognizable-class (e.g., Black jurors).
+Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
+We then used the model to fit the 50 trials to see if the estimated bias parameters are close to each other.
+We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
+Figure 4 depicts the results.
 
@@
 We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app using `r R.version$version.string`[@R] and shiny [@shiny].
+We built this app using `r R.version$version.string` and shiny [@R; @shiny].
 
@@
 df <- data.frame(round = c(1:10),
-                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
-                 total = c(15,14,13,12,11,10,9,8,7,6),
+                 num_cog = c(5,5,4,3,2,2,2,2,2,2),
+                 total = c(21,20,19,18,17,16,15,14,13,12),
                  cog = c(0,1,1,0,0,1,1,0,0,0),
                  party = rep(c("PP","PD"),5))
-knitr::kable(df)
+knitr::kable(df, caption ="Hypothetical Strike Tally")
 
@@
 
-To illustrate, suppose the strike tally in Table \@ref(tab:inpts) represents the pattern of strikes in the present case with prosecutor "X" and defense attorney "Y" (both aliases for actual attorneys in the historical strike data).
-After the user enters this strike tally, the app displays two graphs -- one for the prosecution and the defense.
+To illustrate, suppose the strike tally in Table 1 depicts the pattern of strikes in the present case with a prosecuting attorney Aaron Waelchi and defense attorney Lawrance Klocko V (both aliases for actual attorneys in the historical strike data).
+After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.
 Each graph depicts the prior density plot (colored grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure 5).
 
-```{r figapp1, fig.cap = "Density Plots without Historical Strike Data"}
+```{r figapp1, out.width='50%', fig.cap = "Density Plots without Historical Strike Data", fig.align='center'}
 
@@
 
-To use historical strike data, the user selects a name of prosecutor or defense attorney from the pull-down menus.
+To use historical strike data, the user selects a name of the prosecutor or defense attorney from the pull-down menus.
 If an attorney's name cannot be found, the app has no historical strike data for that attorney.
@@
 For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information ($a = 1$).
-The user has two other options: half weight of historical information ($a = 0.5$) and minimal weight of historical information ($a = 0.2$).
+The user has two other options: half weight ($a = 0.5$) and minimal weight ($a = 0.2$).
 
 In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to "Equal".
-The density plots update accordingly (Figure \@ref(fig:figapp2)).
+The density plots update accordingly (Figure 6).
 
-```{r figapp2, eval = FALSE, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
-library(dplyr)
-library(Rcpp)
-load_path2 <- here::here("app/mh_sampler_pp.cpp")
-load_path3 <- here::here("app/jury_data_cleaned_new.rds")
-sourceCpp(load_path2)
-sourceCpp(load_path2)
-dat0 <- readRDS(load_path3)
-df <- data.frame(round = c(1:10),
-                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
-                 total = c(15,14,13,12,11,10,9,8,7,6),
-                 cog = c(0,1,1,0,0,1,1,0,0,0),
-                 party = rep(c("PP","PD"),5))
-extract_atny <- function(atny_name,pp,dat){
-  if(pp){
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-  else{
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-} 
-organize_input <- function(dat,pp,cog){
-  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
-  if(cog=='gender'){
-    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
-  }else{
-    dat_strikes <- dat_strikes %>% filter(!is.na(race))
-  }
+```{r figapp2, out.width='50%', fig.cap = "Density Plots with Historical Strike Data (Race)", fig.align='center'}
 
-  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
-  if(cog=='gender'){
-    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
-  }else{
-    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
-  }
+load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot2.png")
 
-  num_cog <- sum(dat_strikes$cogb)
-  num_t <- nrow(dat_strikes)
-  subs <- data.frame(round = c(1:num_t), 
-                     num_cog = NA, 
-                     total = NA)
-  subs$num_cog[1] <- num_cog
-  subs$total[1] <- num_t
-  for ( j in 2:num_t){
-    num_cog = num_cog-dat_strikes$cogb[j-1]
-    num_total = num_t - dat_strikes$strike_seq[j-1]
-    subs$num_cog[j] <- num_cog
-    # number of females struck in this and previous rounds
-    subs$total[j] <- num_total
-  }
-  subs <- cbind(subs,dat_strikes$cogb)
-  colnames(subs)[4] = 'cognizable'
-  if(pp){
-    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
-  }else{
-    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PD'),])
-  }
-  return(df_m)
-}
-subset <- function(atny_name,pp,dat,cog){
-  dat_sub <- extract_atny(atny_name,pp,dat)
-  sub1 <- dat_sub %>% group_split(ID)
-  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
-  df_m = do.call(rbind.data.frame,sub1_l)
-  df_m <- as.matrix(df_m)
-}
-df_mp <- df %>%
-                filter(party == "PP") %>%
-                select(-c(party)) %>%
-                as.matrix()
-
-df_md <- df %>%
-                filter(party == "PD") %>%
-                select(-c(party)) %>%
-                as.matrix()
-
-sub_p <- subset("Brian P. Leaming",TRUE,dat0,"race")
-sub_d <- subset("Frank J. Riccio II",FALSE,dat0,"race")
-out_p <- make_posterior_p(x = df_mp,x_p =sub_p,a0 = 1, niter = 110000, 
-                                          theta_start_val = 0, theta_proposal_sd =.5, 
-                                          prior_mean = 0, prior_sd = 2)
-pp_prior_theta <- make_posterior_prior(x_p=sub_p,a0 = 1, niter = 110000, 
-                                        theta_start_val = 0, theta_proposal_sd =.5, 
-                                                       prior_mean = 0, prior_sd = 2)
-d_p <- data.frame(
-                theta = out_p$theta[10001:110000],
-                party = "Prosecution", 
-                posterior = "Posterior")
-pp_prior <- data.frame(theta = pp_prior_theta$theta[10001:110000], 
-                                       party = "Prosecution", 
-                                       posterior = "Prior")
-
-out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 1, niter = 110000, 
-                                          theta_start_val = 0, theta_proposal_sd =.5, 
-                                          prior_mean = 0, prior_sd = 2)
-pd_prior_theta <- make_posterior_prior(x_p=sub_d,a0 = 1, niter = 110000, 
-                                        theta_start_val = 0, theta_proposal_sd =.5, 
-                                        prior_mean = 0, prior_sd = 2)
-pd_prior <- data.frame(theta = pd_prior_theta$theta[10001:110000], 
-                                       party = "Defense", 
-                                       posterior = "Prior")
-d_d <- data.frame(
-                theta = out_d$theta[10001:110000],
-                party = "Defense", 
-                posterior = "Posterior")
-
-dat <- rbind(d_p,d_d)
-priors <- rbind(pp_prior, pd_prior)
-dat <- rbind(dat, priors)
-dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
-                                ordered = TRUE)
-CI <- dat %>% filter(posterior == 'Posterior') %>%
-                group_by(party) %>%
-                summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%
-                mutate(bias = ifelse(
-                    q1 <= 0 & q2 >= 0, "No Bias", "Bias"))
-pplot <- ggplot(data=dat) + 
-                geom_density(aes(x = theta, 
-                                 fill = interaction(party, posterior),
-                                 color = interaction(party, posterior),
-                                 alpha = interaction(party, posterior), 
-                                 ..scaled..))+
-                facet_wrap(~party, nrow =2)
-
-pplot <- pplot  + theme_minimal() +
-                 labs (title = "Likely values of b") +
-                 xlab("") + 
-                 ylab("") + 
-                 xlim(c(-6,6)) +
-                 scale_fill_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
-                 scale_color_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
-                 scale_alpha_manual("Group", values = c(0.3, 0.3, 0.1, 0.1)) +
-
-                  # edit text sizes for plot
-                 theme(legend.position="none",
-                       axis.text = element_text(size = 16), 
-                       strip.text = element_text(size = 18), 
-                       plot.title = element_text(size = 24),
-                       plot.subtitle = element_text(size = 16)) +
-                  # add line at zero for reference
-                 geom_vline(xintercept = 0, color = "black", lwd=1.5)
-
- pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
-                               linetype="dashed", size = 0.9)+
-                    geom_vline(data=CI, aes(xintercept=q2), color = c("blue", "darkred"),
-                           linetype="dashed", size = 0.9)
+knitr::include_graphics(load_path)
 
@@
 
+Now, the credible intervals clearly exclude zero.
 Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
@@
 Next, suppose the same scenario, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
-If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:brian2)).
+If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure 7).
 
-```{r brian2, eval = FALSE, fig.cap="Posterior Density Plots, a = 1, cognitive class = gender"}
+```{r figapp3, out.width='50%', fig.cap = "Density Plots with Historical Strike Data (Gender)", fig.align='center'}
 
-load_path2 <- here::here("app/mh_sampler_pp.cpp")
-load_path3 <- here::here("app/jury_data_cleaned_new.rds")
+load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot3.png")
 
-sourceCpp(load_path2)
+knitr::include_graphics(load_path)
 
-dat0 <- readRDS(load_path3)
-
-df <- data.frame(round = c(1:10),
-                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
-                 total = c(15,14,13,12,11,10,9,8,7,6),
-                 cog = c(0,1,1,0,0,1,1,0,0,0),
-                 party = rep(c("PP","PD"),5))
-
-extract_atny <- function(atny_name,pp,dat){
-  if(pp){
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-  else{
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-} 
-organize_input <- function(dat,pp,cog){
-  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
-  if(cog=='gender'){
-    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
-  }else{
-    dat_strikes <- dat_strikes %>% filter(!is.na(race))
-  }
-
-  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
-  if(cog=='gender'){
-    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
-  }else{
-    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
-  }
-
-  num_cog <- sum(dat_strikes$cogb)
-  num_t <- nrow(dat_strikes)
-  subs <- data.frame(round = c(1:num_t), 
-                     num_cog = NA, 
-                     total = NA)
-  subs$num_cog[1] <- num_cog
-  subs$total[1] <- num_t
-  for ( j in 2:num_t){
-    num_cog = num_cog-dat_strikes$cogb[j-1]
-    num_total = num_t - dat_strikes$strike_seq[j-1]
-    subs$num_cog[j] <- num_cog
-    # number of females struck in this and previous rounds
-    subs$total[j] <- num_total
-  }
-  subs <- cbind(subs,dat_strikes$cogb)
-  colnames(subs)[4] = 'cognizable'
-  if(pp){
-    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
-  }else{
-    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PD'),])
-  }
-  return(df_m)
-}
-subset <- function(atny_name,pp,dat,cog){
-  dat_sub <- extract_atny(atny_name,pp,dat)
-  sub1 <- dat_sub %>% group_split(ID)
-  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
-  df_m = do.call(rbind.data.frame,sub1_l)
-  df_m <- as.matrix(df_m)
-}
-df_mp <- df %>%
-                filter(party == "PP") %>%
-                select(-c(party)) %>%
-                as.matrix()
-
-df_md <- df %>%
-                filter(party == "PD") %>%
-                select(-c(party)) %>%
-                as.matrix()
-
-sub_p <- subset("Brian P. Leaming",TRUE,dat0,"gender")
-sub_d <- subset("Frank J. Riccio II",FALSE,dat0,"gender")
-out_p <- make_posterior_p(x = df_mp,x_p =sub_p,a0 = 1, niter = 110000, 
-                                          theta_start_val = 0, theta_proposal_sd =.5, 
-                                          prior_mean = 0, prior_sd = 2)
-pp_prior_theta <- make_posterior_prior(x_p=sub_p,a0 = 1, niter = 110000, 
-                                        theta_start_val = 0, theta_proposal_sd =.5, 
-                                                       prior_mean = 0, prior_sd = 2)
-d_p <- data.frame(
-                theta = out_p$theta[10001:110000],
-                party = "Prosecution", 
-                posterior = "Posterior")
-pp_prior <- data.frame(theta = pp_prior_theta$theta[10001:110000], 
-                                       party = "Prosecution", 
-                                       posterior = "Prior")
-
-out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 1, niter = 110000, 
-                                          theta_start_val = 0, theta_proposal_sd =.5, 
-                                          prior_mean = 0, prior_sd = 2)
-pd_prior_theta <- make_posterior_prior(x_p=sub_d,a0 = 1, niter = 110000, 
-                                        theta_start_val = 0, theta_proposal_sd =.5, 
-                                        prior_mean = 0, prior_sd = 2)
-pd_prior <- data.frame(theta = pd_prior_theta$theta[10001:110000], 
-                                       party = "Defense", 
-                                       posterior = "Prior")
-d_d <- data.frame(
-                theta = out_d$theta[10001:110000],
-                party = "Defense", 
-                posterior = "Posterior")
-
-dat <- rbind(d_p,d_d)
-priors <- rbind(pp_prior, pd_prior)
-dat <- rbind(dat, priors)
-dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
-                                ordered = TRUE)
-CI <- dat %>% filter(posterior == 'Posterior') %>%
-                group_by(party) %>%
-                summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%
-                mutate(bias = ifelse(
-                    q1 <= 0 & q2 >= 0, "No Bias", "Bias"))
-pplot <- ggplot(data=dat) + 
-                geom_density(aes(x = theta, 
-                                 fill = interaction(party, posterior),
-                                 color = interaction(party, posterior),
-                                 alpha = interaction(party, posterior), 
-                                 ..scaled..))+
-                facet_wrap(~party, nrow =2)
-
-pplot <- pplot  + theme_minimal() +
-                 labs (title = "Likely values of b") +
-                 xlab("") + 
-                 ylab("") + 
-                 xlim(c(-6,6)) +
-                 scale_fill_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
-                 scale_color_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
-                 scale_alpha_manual("Group", values = c(0.3, 0.3, 0.1, 0.1)) +
-
-                  # edit text sizes for plot
-                 theme(legend.position="none",
-                       axis.text = element_text(size = 16), 
-                       strip.text = element_text(size = 18), 
-                       plot.title = element_text(size = 24),
-                       plot.subtitle = element_text(size = 16)) +
-                  # add line at zero for reference
-                 geom_vline(xintercept = 0, color = "black", lwd=1.5)
-
- pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
-                               linetype="dashed", size = 0.9)+
-                    geom_vline(data=CI, aes(xintercept=q2), color = c("blue", "darkred"),
-                           linetype="dashed", size = 0.9) 
 ```
 
-Here, most of the density curve of both prior and posterior are on the left of zero, indicating the possible bias against male jurors.
-However, since the 80\% credible interval includes 0, inferring that preference from the strike data alone is unjustified.
+Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors.
+However, because the 80\% credible interval includes 0, inferring such bias from the strike data alone is unjustified.
 
-Finally, if we assign minimal weight to the historical strike data, the density plot will be as shown in figure \@ref(fig:brian3).
+Finally, if we assign minimal weight to the historical strike data, the density plot updates accordingly (Figure 8).
 
-```{r brian3, eval = FALSE, fig.cap="Posterior Density Plot, a = 0.2"}
-load_path2 <- here::here("app/mh_sampler_pp.cpp")
-load_path3 <- here::here("app/jury_data_cleaned_new.rds")
-sourceCpp(load_path2)
-sourceCpp(load_path2)
-dat0 <- readRDS(load_path3)
-df <- data.frame(round = c(1:10),
-                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
-                 total = c(15,14,13,12,11,10,9,8,7,6),
-                 cog = c(0,1,1,0,0,1,1,0,0,0),
-                 party = rep(c("PP","PD"),5))
-extract_atny <- function(atny_name,pp,dat){
-  if(pp){
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-  else{
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-} 
-organize_input <- function(dat,pp,cog){
-  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
-  if(cog=='gender'){
-    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
-  }else{
-    dat_strikes <- dat_strikes %>% filter(!is.na(race))
-  }
+```{r figapp4, out.width='50%', fig.cap = "Density Plots with Historical Strike Data (Gender, Minimal Weight)", fig.align='center'}
 
-  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
-  if(cog=='gender'){
-    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
-  }else{
-    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
-  }
+load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot4.png")
 
-  num_cog <- sum(dat_strikes$cogb)
-  num_t <- nrow(dat_strikes)
-  subs <- data.frame(round = c(1:num_t), 
-                     num_cog = NA, 
-                     total = NA)
-  subs$num_cog[1] <- num_cog
-  subs$total[1] <- num_t
-  for ( j in 2:num_t){
-    num_cog = num_cog-dat_strikes$cogb[j-1]
-    num_total = num_t - dat_strikes$strike_seq[j-1]
-    subs$num_cog[j] <- num_cog
-    # number of females struck in this and previous rounds
-    subs$total[j] <- num_total
-  }
-  subs <- cbind(subs,dat_strikes$cogb)
-  colnames(subs)[4] = 'cognizable'
-  if(pp){
-    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PP'),])
-  }else{
-    df_m <- as.matrix(subs[which(dat_strikes$Disp=='PD'),])
-  }
-  return(df_m)
-}
-subset <- function(atny_name,pp,dat,cog){
-  dat_sub <- extract_atny(atny_name,pp,dat)
-  sub1 <- dat_sub %>% group_split(ID)
-  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
-  df_m = do.call(rbind.data.frame,sub1_l)
-  df_m <- as.matrix(df_m)
-}
-df_mp <- df %>%
-                filter(party == "PP") %>%
-                select(-c(party)) %>%
-                as.matrix()
+knitr::include_graphics(load_path)
 
-df_md <- df %>%
-                filter(party == "PD") %>%
-                select(-c(party)) %>%
-                as.matrix()
-
-sub_p <- subset("Brian P. Leaming",TRUE,dat0,"gender")
-sub_d <- subset("Frank J. Riccio II",FALSE,dat0,"gender")
-out_p <- make_posterior_p(x = df_mp,x_p =sub_p,a0 = 0.2, niter = 110000, 
-                                          theta_start_val = 0, theta_proposal_sd =.5, 
-                                          prior_mean = 0, prior_sd = 2)
-pp_prior_theta <- make_posterior_prior(x_p=sub_p,a0 = 0.2, niter = 110000, 
-                                        theta_start_val = 0, theta_proposal_sd =.5, 
-                                                       prior_mean = 0, prior_sd = 2)
-d_p <- data.frame(
-                theta = out_p$theta[10001:110000],
-                party = "Prosecution", 
-                posterior = "Posterior")
-pp_prior <- data.frame(theta = pp_prior_theta$theta[10001:110000], 
-                                       party = "Prosecution", 
-                                       posterior = "Prior")
-
-out_d <- make_posterior_p(x = df_md,x_p = sub_d,a0= 0.2, niter = 110000, 
-                                          theta_start_val = 0, theta_proposal_sd =.5, 
-                                          prior_mean = 0, prior_sd = 2)
-pd_prior_theta <- make_posterior_prior(x_p=sub_d,a0 = 0.2, niter = 110000, 
-                                        theta_start_val = 0, theta_proposal_sd =.5, 
-                                        prior_mean = 0, prior_sd = 2)
-pd_prior <- data.frame(theta = pd_prior_theta$theta[10001:110000], 
-                                       party = "Defense", 
-                                       posterior = "Prior")
-d_d <- data.frame(
-                theta = out_d$theta[10001:110000],
-                party = "Defense", 
-                posterior = "Posterior")
-
-dat <- rbind(d_p,d_d)
-priors <- rbind(pp_prior, pd_prior)
-dat <- rbind(dat, priors)
-dat$party <- factor(dat$party, levels = c("Defense", "Prosecution"), 
-                                ordered = TRUE)
-CI <- dat %>% filter(posterior == 'Posterior') %>%
-                group_by(party) %>%
-                summarise(q1 = quantile(theta,0.1), q2 = quantile(theta,0.9)) %>%
-                mutate(bias = ifelse(
-                    q1 <= 0 & q2 >= 0, "No Bias", "Bias"))
-pplot <- ggplot(data=dat) + 
-                geom_density(aes(x = theta, 
-                                 fill = interaction(party, posterior),
-                                 color = interaction(party, posterior),
-                                 alpha = interaction(party, posterior), 
-                                 ..scaled..))+
-                facet_wrap(~party, nrow =2)
-
-pplot <- pplot  + theme_minimal() +
-                 labs (title = "Likely values of b") +
-                 xlab("") + 
-                 ylab("") + 
-                 xlim(c(-6,6)) +
-                 scale_fill_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
-                 scale_color_manual("Group", values = c("blue", "darkred", "grey", "grey")) +
-                 scale_alpha_manual("Group", values = c(0.3, 0.3, 0.1, 0.1)) +
-
-                  # edit text sizes for plot
-                 theme(legend.position="none",
-                       axis.text = element_text(size = 16), 
-                       strip.text = element_text(size = 18), 
-                       plot.title = element_text(size = 24),
-                       plot.subtitle = element_text(size = 16)) +
-                  # add line at zero for reference
-                 geom_vline(xintercept = 0, color = "black", lwd=1.5)
-
- pplot + geom_vline(data=CI, aes(xintercept=q1), color = c("blue", "darkred"),
-                               linetype="dashed", size = 0.9)+
-                    geom_vline(data=CI, aes(xintercept=q2), color = c("blue", "darkred"),
-                           linetype="dashed", size = 0.9) 
 ```
 
-We can find both the prior and posterior density become more flat, and the posterior density appears symmetric around zero now.
+We can find both the prior and posterior density become flatter, and the credible intervals plainly includes zero.
 
@@
 
-Statistical methods for _Batson_ and similar changes are intended to measure how much the prospective jurors' race, gender, or other cognizable-class membership matters to whether attorneys use peremptory challenges against them.
-The ideal data, therefore, is, for each attorney, not just who they struck in the current trial, but a complete history of who they struck in all their past cases.
-
-The actual available data, however, is not that ideal, because jury-selection records in past cases are hard to get because of court practices and otherwise require painstaking effort to compile [@Grosso2017; @Wright2018].
-As a result, historical strike data may be missing not at random.
-Possible causes include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender).
-This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information [e.g., @Baldus2001, 132-134].
-Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-
+Statistical methods for _Batson_ and similar changes are intended to measure how much prospective jurors' race, gender, or other cognizable-class membership matters to how likely an attorney will use peremptory strikes against them.
 The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
+
 And as our prototype app shows, if the historical strike data is collected and loaded into the app in advance, this approach can be easily used by attorneys and others in real time during jury selection.
@@
 
-These data challenges, however, might decrease over time, as attorneys and others realize that more complete historical strike data may make it easier, with the approach described here, to credibly infer bias, or rebut that inference, under _Batson_ and similar law.
+Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data, therefore, is, for each attorney, not just who they struck in the current trial, but a complete history of who they struck in all their past cases.
+The problem: Jury-selection records in past cases are hard to get because of court practices and otherwise require painstaking effort to compile [@Grosso2017; @Wright2018].
+Possible causes include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender).
+This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
+Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
+This presents an issue for inferring bias if historical strike data is missing not at random.
+These data challenges, however, might decrease over time, as judges, attorneys, and jury clerks realize that more complete historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
 
-# CRediT Author Statement
+# Acknowledgements
+
+## CRediT Author Statement
 
@@
 
-Based on categories in https://casrai.org/credit/].
+Based on categories in https://casrai.org/credit/.
 

### 2021-09-22 analysis/paper_draft2.Rmd
minor edits

@@
 
-And as our prototype app shows, if the historical strike data is collected and loaded into the app in advance, this approach can be easily used by attorneys and others in real time during jury selection.
+And as our prototype app shows, if the historical strike data loaded into the app in advance, attorneys and others can the approach here in real time during jury selection.
 Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
-Whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges is a question worth studying further.
+It is worth studying whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges.
 
-At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data, given the choice of how to model a court's particular peremptory-challenge process.
+At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data.
 For example, given the model we used, we required data on the cognitive-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
@@
 
-Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data, therefore, is, for each attorney, not just who they struck in the current trial, but a complete history of who they struck in all their past cases.
-The problem: Jury-selection records in past cases are hard to get because of court practices and otherwise require painstaking effort to compile [@Grosso2017; @Wright2018].
-Possible causes include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender).
+Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data is, for each attorney, a complete history of who they struck in all their past cases.
+The problem: Jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018].
+For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
 This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
-Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an issue for inferring bias if historical strike data is missing not at random.
-These data challenges, however, might decrease over time, as judges, attorneys, and jury clerks realize that more complete historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
+And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
+Incomplete historical strike data resents an issue for inferring bias if such data is incomplete or missing not at random.
+These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
 

### 2021-09-22 analysis/draft_archive/paper_draft.Rmd
file cleaning

--- 2021-09-14 analysis/paper_draft.Rmd
@@
 ---
-title: "Detecting Bias in Peremptory Challenges Using Historical Strike Data"
+title: "Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data"
 date: "`r Sys.Date()`"
+author:
+  - Sachin S. Pandya, University of Connecticut
+#    orcid_id: 0000-0001-7387-1307
+  - Xiaomeng Li, University of Connecticut | Ernest & Young
+  - Timothy E. Moore, University of Connecticut
+#    orcid_id: 0000-0002-9576-0517
+#output: distill::distill_article
+output: pdf_document
 bibliography: references.bib
-author:
-  - name: "Sachin S. Pandya"
-    affiliation: "University of Connecticut"
-    orcid_id: 0000-0001-7387-1307
-  - name: "Xiaomeng Li"
-    affiliation: "University of Connecticut | Ernest & Young"
-  - name: "Timothy E. Moore"
-    affiliation: "University of Connecticut"
-    orcid_id: 0000-0002-9576-0517
-output: distill::distill_article
+abstract: |
+  US constitutional law bars using peremptory strikes during jury selection because of race, gender, or another cognizable classes.
+Here, we propose a Bayesian approach for detecting such illegal bias based on an attorney's pattern of peremptory challenges in both a current trial as well that attorney's use of strikes in past cases.
+Specifically, we show how to use historical data on an attorney's use of peremptory strikes in past trials to estimate, in a current trial, how likely an attorney has used peremptory challenges because of the cognizable class of prospective jurors.
+We also introduce the use of a _power prior_ to provide flexibility in the use of such historical data.
+We show the utility of our model on simulated data.
+Finally, we extend our approach by presenting an R Shiny application that can be used to examine jury selection data in real time.
+We demonstrate the app using actual jury-selection data from a convenience sample of criminal trials in the federal district court of Connecticut from YYYY to YYYY.
+
 ---
+
+```{r setup, include=FALSE}
+
+library(ggplot2)
+library(dplyr)
+
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
+
+```
 
@@
 
-With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
-While some jurisdictions have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
+With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.^[While some jurisdictions have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1) (UK); S.C.
 2019, c.
-25, s 269 (Canada)), federal and State law in the United States (except in Arizona^[Arizona Rule of Criminal Procedure XX]) authorize their use during jury selection in criminal and civil cases.
-In these cases, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
+25, s 269 (Canada)), federal and State law in the United States authorize their use during jury selection in criminal and civil cases.
+The current exception is Arizona.
+Arizona Rule of Criminal Procedure XX.] Peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes each party may exercise at a time [@nscs-jury2018; @Williams2017].
+
+This paper focuses on _Batson_ challenges to a party's use of peremptory challenges.
 
@@
 22.3].
-For similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v.
+
+For some similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v.
 Holloway, 209 Conn.
@@
 Chapman, 317 S.C.
-302, 305-06 (1995)], while other States have shifted the inquiry away from the striking party's motives to whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash.
+302, 305-06 (1995)], while other States require proving not motive, only that an "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash.
 R.
@@
 
-By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases.
+By law, evidence supporting such challenges may include aggregate data on peremptory challenges in past cases.
 For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v.
@@
 Jury-selection records are often not readily available, because they are kept, if at all, by court staff somewhere other than the case file.
-Such records are also largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking effort to collect [e.g., @Craft2018].
+Such records are also largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise requires painstaking effort to collect [e.g., @APM2018].
 
@@
 When collecting strike data in multiple trials, prior studies have aggregated strikes across those trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference.
+
 For example, Eisenberg [-@Eisenberg2017, 334-339] used a 2-tail Fisher's exact test on the observed difference in peremptory strikes by race and gender in a non-random sample of thirty-five trials from 1997-2012 that resulted in death sentences (out of 63 such trials).
@@
 Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001].
-For exmpale, using jury selection data for 1992-2017 for Mississippi's Fifth Circuit Court District, Craft [-@Craft2018a, 9] applied a logistic regression model to analyze how much a prospective juror's race affected the odds of being struck by the state.
+For example, using jury selection data for 1992-2017 for Mississippi's Fifth Circuit Court District, Craft [-@Craft2018a, 9] applied a logistic regression model to analyze how much a prospective juror's race affected the odds of being struck by the state.
 For studies applying other modeling techniques to this Mississippi data, see @DeCamp2020 (propensity score matching) and @Dunn2021 (optimal feature selection, optimal classification trees).
@@
 Second, we take that law as requiring, absent any data, the assumption that higher degrees of illegal bias (preference) are less likely than lower degrees of bias (preference).
-By using a normal (Gausian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
+By using a normal (Gaussian) distribution with $\sigma = 2$, we encode this assumption by making values of the bias parameter less likely as they move further from zero and by making very unlikely values any value of $b$ greater than two standard deviations from the mean.
 
@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998] and Geweke’s diagnostics [@Gewe1992].
+To measure convergence, we used traceplots, the Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998] and Geweke’s diagnostics [@Geweke1992].
 
@@
 
-```{r b80, echo=FALSE, fig.cap="80% credible interval"}
-
-library(ggplot2)
-
-# INSERT simulation and plot code
-
-load_path <- here::here("sim_resultc.csv")
+```{r b80, fig.cap="80% credible interval"}
+
+# simulated data via "noseq_simulation.R" script
+
+load_path <- here::here("data-sim","sim_resultc.csv")
 
@@
 sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
+
+library(ggplot2)
 
@@
 
-```{r b95,  echo=FALSE,fig.cap="95% credible interval"}
+```{r b95, fig.cap="95% credible interval"}
 
@@
 
-# Model 2: Simultaneous Strikes
+# Extension: Simultaneous Strikes
 
@@
 
-```{r simltns, echo=FALSE,fig.cap="simultaneous case"}
-
-library(dplyr)
-library(ggplot2)
+```{r simltns, fig.cap="simultaneous case", eval = FALSE}
+
 #library(conflicted)
@@
         axis.text.y = element_text(size = 9))
+
 ```
+
 The points represent the estimated posterior mean and the lines represent the 95\% credible interval.
@@
 
-Following this suggestion, we describe here a prototype software application ("app") that implements the approach described above using `r R.version$version.string`[@R] and shiny [@shiny].
-
-There is currently historical information of some prosecutors and defendants.
-More historical information would be updated when available.
-Users can input the current trial information as shown in table \@ref(tab:inpts), and then select the name of the prosecutor and the defendant from the pull-down window if applicable.
-If the name of the prosecutor or the defense attorney cannot be found, it means the historical information is not available.
-
-The main screen displays two graphs -- one for the prosecution and the defense - that displays the prior density plot and the posterior density plot for the bias parameter.
-
-```{r inpts, echo=FALSE}
+Following this suggestion, we describe here a prototype software application ("app") that implements the approach described above using `r R.version$version.string`[@R] and shiny [@shiny] and comes loaded with real jury-selection data from a convenience sample of criminal cases in the federal district court for Connecticut during YYYY through YYYY.
+To partially de-identify this historical data, we excluded defendants' names and substituted fictitious names for the attorneys using @charlatan.
+
+With the app, users can input the current information on the use of peremptory strikes as shown in table \@ref(tab:inpts), and then select the name of the prosecutor and the defendant from the pull-down window, if available, in the historical data.
+If an attorney's name cannot be found, it means that the app does not have any historical information on that attorney's use of peremptory strikes.
+
+The main screen displays two graphs -- one for the prosecution and the defense - that displays the prior and posterior density plots for the bias parameter.
+
+```{r inpts, echo=FALSE, eval = FALSE}
 df <- data.frame(round = c(1:10),
@@
 knitr::kable(df)
+
 ```
@@
 There are currently two options of cognizable class, race and gender.
+
 Suppose the current trial information is as we show in table \@ref(tab:inpts), we put equal weight on historical information and we select cognizable class as race, then the posterior density as shown in figure \@ref(fig:brian) will be displayed.
@@
 Sachin S.
-Pandya: Conceptualization, Investigation, Software, Writing; Xiaomeng Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Timothy Moore: Conceptualization, Supervision, Visualization.
+Pandya: Conceptualization, Investigation, Software, Writing; Xiaomeng Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Timothy Moore: Conceptualization, Supervision, Visualization, Writing.
 Based on categories in https://casrai.org/credit/].

### 2021-10-21 analysis/paper_draft2.Rmd
updated

@@
 
-And as our prototype app shows, if the historical strike data loaded into the app in advance, attorneys and others can the approach here in real time during jury selection.
-Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
-It is worth studying whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges.
+Moreover, if the historical strike data can loaded into a tool like our prototype app, attorneys and others can the approach here in real time during jury selection.
+Despite _Batson_ and similar law, racial disparity persists in the use of peremptory challenges [@Diamond2018].
+One concern is that attorneys fear that arguing _Batson_  will produce the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
+At the same time, we assume that lawyers and judges want to discourage an attorney from raising a _Batson_ challenge without a credible basis.
+It is worth studying whether an app like the prototype presented here might reduce such barriers to bringing otherwise credible _Batson_ challenges.
 
@@
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-Incomplete historical strike data resents an issue for inferring bias if such data is incomplete or missing not at random.
+Incomplete historical strike data presents an issue for inferring bias if such data is incomplete or missing not at random.
 These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.

### 2021-10-26 analysis/paper_draft2.Rmd
revised text of Discussion

--- 2021-09-22 analysis/paper_draft2.Rmd
@@
 
-Statistical methods for _Batson_ and similar changes are intended to measure how much prospective jurors' race, gender, or other cognizable-class membership matters to how likely an attorney will use peremptory strikes against them.
+Statistical methods for _Batson_ and similar changes aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership matters to how likely an attorney will use peremptory strikes against them.
 The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
-
-And as our prototype app shows, if the historical strike data loaded into the app in advance, attorneys and others can the approach here in real time during jury selection.
-Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021].
-It is worth studying whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges.
+And with a tool like our prototype app, attorneys can use historical strike data to detect bias in real time during jury selection, thus making it easier to raise credible _Batson_ challenges.
 
@@
 
-Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data is, for each attorney, a complete history of who they struck in all their past cases.
-The problem: Jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018].
+Finally, for the approach here, incomplete historical strike data presents an issue for inferring bias if such data is missing not at random.
+In practice, jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018].
 For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
-This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information.
+This forces researchers to turn to other data sources, if any, to find that information.
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-Incomplete historical strike data resents an issue for inferring bias if such data is incomplete or missing not at random.
-These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
+We hope these data challenges decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
 

### 2021-10-27 analysis/paper_draft2.Rmd
revised

@@
 
-knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 
@@
 Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
-Typically, they test for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014; @Gastwirth2013].
+Typically, they test for the probability of observing a non-zero difference in strike rates by the race or gender of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014; @Gastwirth2013].
 Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
@@
 In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2018] to incorporate historical strike data.
-In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter $b$, to which we assign an initial prior distribution.
-Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for $b$.
-In so doing, we use the _power prior_ [@Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.
-Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.
+In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
+Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
+In so doing, we use the _power prior_ [@Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
+Finally, we use that posterior distribution as the prior for the bias parameter in the present case and update the posterior based on data on the target attorney's strikes in that present case.
 
@@
 
-Following @Kadane2018 and @Barrett2007, we model a "struck-jury" peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
-Under the "struck jury" procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.
-Then, the parties exercise their strikes on those potential jurors, either submitting strikes simultaneously or in an alternating sequence.
-Once all strikes are used or waived, the remaining prospective jurors are assigned to seats on the jury.
+Following @Kadane2018 and @Barrett2007, we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories.
+Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that corresponds to the number seats on the jury (plus alternates, if any).
+The parties exercise their strikes on anyone among this subset of potential jurors in an alternating sequence.
+Once all strikes are used or waived, the remaining prospective jurors are each assigned to seats on the jury until those seats are filled.
 
@@
 
-In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.
+In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
 This simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
@@
 We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app using `r R.version$version.string` and shiny [@R; @shiny].
+We built this app with `r R.version$version.string` and shiny [@R; @shiny].
 
@@
 $\S$ 1868.
-Thereafter, we received copies (pdf files) of certain jury-selection records associated with [INSERT] trials in that court during this period.
-These included strike sheets that indicated the ID number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
+Thereafter, we received copies of certain jury-selection records associated with twenty-nine criminal cases in that court during this period.
+These included strike sheets that indicated the identification number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
 Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.
@@
 
-```{r inpts, echo=FALSE}
+```{r inpts}
 df <- data.frame(round = c(1:10),
@@
 
-Finally, for the approach here, incomplete historical strike data presents an issue for inferring bias if such data is missing not at random.
-In practice, jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018].
+Finally, in practice, historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
 For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
@@
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-We hope these data challenges decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
+This presents an issue for the approach presented here for detecting bias if the missing strike data is missing not at random.
 
@@
 
-Based on categories in https://casrai.org/credit/.
-
 # References

### 2021-10-27 analysis/paper_draft2.Rmd
revised

@@
 
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)
 
@@
 
-In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
+In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors in a venire.
 This simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
@@
 
-```{r inpts}
+```{r inpts, echo=FALSE}
 df <- data.frame(round = c(1:10),

### 2021-11-02 analysis/paper_draft2.Rmd
edits - ongoing

--- 2021-10-27 analysis/paper_draft2.Rmd
@@
 We also use the _power prior_ to adjust the weight of such historical data in the analysis.
-We show the utility of our model on simulated data.
-Finally, we extend this with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection.
-We demonstrate the app using actual historical strike data from a convenience sample of cases from the federal district court of Connecticut.
+We show how well our model performs on simulated data.
+Finally, we extend this approach with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection.
+We illustrate this application's use with actual historical strike data from a convenience sample of cases from one federal district court.
 
@@
 2243; Wash.
-Rule 37(g)(v) -@WashR37; Calif.
+Rule 37(g)(v); Calif.
 Code of Civil Procedure $\S$ 231.7(d)(3)(G)).
@@
 Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the _power prior_ [@Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
+In so doing, we use the power prior [@Ibrahim2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
 Finally, we use that posterior distribution as the prior for the bias parameter in the present case and update the posterior based on data on the target attorney's strikes in that present case.
@@
 \begin{equation}
-L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
+L(b|\delta) \propto L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
 \label{eq:model4}
@@
 
-To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" b $= \{-3,-2,-1,0,1,2,3\}$.
+To do this, we generated strike data for a current trial using seven different values of the bias parameter ($b_{curr} = \{-3,-2,-1,0,1,2,3\}$.
 As introduced above, $b>0$ ($b<0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b=0$ denotes no bias.
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ("historical b" $= \{-3,-2,-1,0,1,2,3\}$).
+We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ($b_{hist} = \{-3,-2,-1,0,1,2,3\}$).
 
-We let historical $b$ differ from current $b$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random.
+We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random.
 Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature that the current trial has but that all the past trials in the historical data do not.
 To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias ($b$) in the current case with a Black defendant.
+If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias in the current case ($b_{curr}$) with a Black defendant.
 
@@
 For each scenario, we generated 1000 datasets and fit the model on those datasets.
-We recorded the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.
-In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$.
+We recorded the posterior mean and credible intervals of one thousand model fits for each scenario.
 
-```{r b80}
+In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times we can reach the correct conclusion about the bias parameter based on those credible intervals.
+That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero.
+When $b_{curr}$ equals to zero, the credible interval includes zero, and when $b_{curr}$ is less than zero, the credible interval lies to the left of zero.
+Since we use a weakly-informative prior that pulls the estimate of $b_{current}$ towards zero, we focus not how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
+
+```{r b80, eval = FALSE}
 
@@
 
-p80 <- ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,
-                        fill = as.factor(sig.n)),
-            color="black")+
-  facet_grid(h~a, labeller = label_both)+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current b")+
-  ylab("Historical b")+
-  scale_fill_manual(values = c("white", "blue"))+
-  coord_equal() +
-  theme_minimal() +
-  labs(fill = "coverage") +
-  theme(axis.text = element_text(size =10), 
-        #axis.title = element_text(size = 16), 
-        #legend.title.align = 0.05, 
-        #legend.title = element_text(size = 16), 
-        legend.text = element_text(size = 14))
-#p80
-
 ```
 
-```{r b90}
+```{r b90, eval = FALSE}
 
@@
 
-p90 <- ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,
-                        fill = as.factor(sig.n)),
-            color="black")+
-  facet_grid(h~a, labeller = label_both)+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current b")+
-  ylab("Historical b")+
-  scale_fill_manual(values = c("white", "blue"))+
-  coord_equal()+
-  theme_minimal()+
-  labs(fill = "coverage") +
-  theme(axis.text = element_text(size =10), 
-        axis.title = element_text(size = 16),
-        legend.title.align = 0.05, 
-        legend.title = element_text(size = 16), 
-        legend.text = element_text(size = 14))
-#p90
-
 ```
 
-```{r b95 }
+```{r b95, eval = FALSE}
 
@@
 
-p95 <- ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,
-                        fill = as.factor(sig.n)),
-            color="black")+
-  facet_grid(h~a, labeller = label_both)+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current b")+
-  ylab("Historical b")+
-  scale_fill_manual(values = c("white", "blue"))+
-  coord_equal()+
-  theme_classic()+  
-  labs(fill = "coverage") +
-  theme(axis.text = element_text(size =10), 
-        axis.title = element_text(size = 16), 
-        legend.title.align = 0.05, 
-        legend.title = element_text(size = 16), 
-        legend.text = element_text(size = 14))
+```
 
-#p95
+For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of $b_{curr}$.
+Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
+
+Figure 1 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively.
+Here, $b_{hist}$ denotes the bias parameter for the generated historical strike data and $b_{curr}$ denotes the bias parameter for the current data.
+The values on the right side represent the size of the historical data.
+The values on the top of the plots represent the power prior weight parameter $a$.
+For any given combination, a darker color shade denotes a higher coverage rate, and vice versa.
+
+```{r c80}
+
+# simulated data via "noseq_simulation.R" script
+
+load_path <- here::here("data-sim","sim_resultc.csv")
+
+sims <- readr::read_csv(load_path)
+sims$a.f <- as.factor(sims$a)
+
+library(latex2exp)
+xlabel = TeX("$b_{curr}$")
+ylabel = TeX("$b_{hist}$")
+
+source(here::here("analysis","R","coverageplot_func.R"))
+
+C80plot <- coverageplot_func(sims, sims$C80, xlabel, ylabel) +
+           labs(title = "(a) Coverage Rates - 80% Credible Interval")
+
+C90plot <- coverageplot_func(sims, sims$C90, xlabel, ylabel) +
+           labs(title = "(b) Coverage Rates - 90% Credible Interval")
+
+C95plot <- coverageplot_func(sims, sims$C95, xlabel, ylabel) +
+           labs(title = "(c) Coverage Rates - 95% Credible Interval")
 
@@
 
-For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$.
-Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
-
-The following figures depict the coverage rates of $b$ under an 80\%, 90\%, 95\% credible interval, respectively.
-In these figures, _historical b_ denotes the bias parameter for the generated historical strike data, and current $b$ means the bias parameter for the current data.
-The values on the right side represent the size of the historical data.
-The values on the top of the plots represent the power prior weight parameter $a$.
-For these figures, a darker color denotes a higher coverage rate, and vice versa.
-
-```{r c80, fig.cap="80% credible interval"}
-
-c80plot <- ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,fill = C80),
-            color="black")+
-  facet_grid(h~a)+
-   scale_fill_gradient2('C80', 
-                        low = "darkred", mid = "white", 
-                        high = "blue", 
-                        breaks = c(0,0.5,0.7,0.9, 1), 
-                        limits = c(0,1),
-           labels = c("low", "", "moderate", "","high"))+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current b")+
-  ylab("Historical b")+
-  coord_equal()+
-  theme_classic()+
-  theme(legend.pos = "none",
-        axis.text = element_text(size =10), 
-        axis.title = element_text(size = 16), 
-        #legend.title.align = 0.05, 
-        #legend.title = element_text(size = 16), 
-        #legend.text = element_text(size = 14)
-        )
-
+```{r fig1a}
 c80plot
-
 ```
 
-Figure 1 depicts the coverage rate with an 80\% credible interval.
-As the first column ($a=0.1$) depicts, when we place low weight on the historical strike data, we mostly rely on current information.
-When the bias parameter is 3, 2 or -2, -3, the coverage rate is high, very close to 1.
-
-However, when the bias parameter ("current b") is 1, the coverage rate drops below 0.8.
-If the bias is not strong, our model can detect it with a lower accuracy.
-However, as we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
-When the bias parameter is 0 (no bias), however, the coverage rate is also around 80\%.
-Thus, if we use the 80\% credible interval, we have around 20\% chance of erroneously detecting non-zero bias.
-
-```{r c90fig, fig.cap="90% credible interval"}
-
-c90plot <- ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,fill = C90),
-            color="black")+
-  facet_grid(h~a)+
-   scale_fill_gradient2('C90', 
-                        low = "darkred", mid = "white", 
-                        high = "blue", 
-                        breaks = c(0,0.5,0.7,0.9, 1), 
-                        limits = c(0,1),
-           labels = c("low", "", "moderate", "","high"))+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current b")+
-  ylab("Historical b")+
-  coord_equal()+
-  theme_classic()+
-  theme(legend.pos = "none",
-        axis.text = element_text(size =10), 
-        axis.title = element_text(size = 16), 
-        #legend.title.align = 0.05, 
-        #legend.title = element_text(size = 16), 
-        #legend.text = element_text(size = 14)
-        )
-
+```{r fig1b}
 c90plot
-
 ```
 
-Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (Figure 2), when $b = 2$, the color is lighter (lower coverage rate) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
-The tradeoff is lower accuracy.
-With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
-When the bias parameter is 3, however, the coverage rate is still high.
-
-```{r c95, fig.cap="95% credible interval"}
-c95plot <- ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,fill = C95),
-            color="black")+
-  facet_grid(h~a)+
-   scale_fill_gradient2('C95', 
-                        low = "darkred", mid = "white", 
-                        high = "blue", 
-                        breaks = c(0,0.5,0.7,0.9, 1), 
-                        limits = c(0,1),
-           labels = c("low", "", "moderate", "","high"))+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current b")+
-  ylab("Historical b")+
-  coord_equal()+
-  theme_classic()+
-  theme(legend.pos = "none",
-        axis.text = element_text(size =10), 
-        axis.title = element_text(size = 16), 
-        #legend.title.align = 0.05, 
-        #legend.title = element_text(size = 16), 
-        #legend.text = element_text(size = 14)
-        )
+```{r fig1c}
+##| fig.cap=fig1caption
 
@@
 
-Figure 3 depicts coverage rates when using a 95\% credible interval.
-Here, given $b = 0$, the coverage rate is very close to 1.
-This implies that we are very unlikely to conclude a bias when there is none.
-When the bias parameter is 3, the coverage rate is still high.
+Figure 1(a) depicts the coverage rate with an 80\% credible interval.
+As the first column ($a = 0.1$) depicts, when we place low weight on the historical strike data, we mostly rely on current information.
+When $b_{current}$ is 3, 2 or -2, -3, the coverage rate is high, very close to 1.
+
+However, when $b_{curr}$ is 1, the coverage rate drops below 0.8.
+If the bias is not strong, our model can detect it with a lower accuracy.
+However, as we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
+When the bias parameter is 0 (no bias), however, the coverage rate is also around 80\%.
+Thus, if we use the 80\% credible interval, we have about a 20\% chance of erroneously detecting non-zero bias.
+
+Using a wider credible interval lowers this chance of error.
+For example, with a 90\% credible interval (Figure 1(b)), when $b = 2$, the color is lighter (lower coverage rate) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
+The tradeoff is lower accuracy.
+With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
+When the bias parameter is 3, however, the coverage rate is still high.
+
+Figure 1(c) depicts coverage rates when using a 95\% credible interval.
+Here, given $b_{curr} = 0$, the coverage rate is very close to 1.
+This implies that we are very unlikely to infer bias when there is none.
+When $b_{curr} = 3$, the coverage rate is still high.
 This implies that we can still accurately detect the bias when it is strong.
-However, when the bias is moderate ($b = \{2, -2\}$, the coverage rate is only high when we have consistent historical information.
+However, when the bias is moderate ($b_{curr} = \{2, -2\}$), the coverage rate is only high when we have consistent historical information.
 
@@
 
-```{r simltns, fig.align='center', fig.cap="Shuffled Sequence Simulation", out.width='50%'}
+```{r simltns}
+
+#| fig.cap="Simulation results for simultaneous strikes example.
+Columns are values of bias used to simulate data.
+Blue points denote estimated posterior mean for the bias parameter.
+Blue lines denote the 95% credible interval.
+Red line denotes true value of bias parameters used in simulations."
 
@@
                      b = "3",
-                     sig = "Bias", # FIX: code as function
+                     sig = "Bias",
                      b.num = 3)
@@
 
-The blue points represent the estimated posterior mean for the  bias parameter and the blue lines represent the 95\% credible interval.
-The red vertical line indicates the true value of the bias parameter.
-
 We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
@@
 We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app with `r R.version$version.string` and shiny [@R; @shiny].
+We built this app with `r R.version$version.string` and Shiny [@R; @shiny].
 
@@
 
-Statistical methods for _Batson_ and similar changes aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership matters to how likely an attorney will use peremptory strikes against them.
+Statistical methods for _Batson_ and similar legal challenges have the purpose of helping to measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
 The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter.
@@
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an issue for the approach presented here for detecting bias if the missing strike data is missing not at random.
+This presents an issue for the approach presented here if the missing strike data is missing not at random.
 

### 2021-11-02 analysis/R/coverageplot_func.R
function for generating coverage plots for simulations

@@ -1,0 +1,27 @@
+coverageplot_func <- function(dat,CI,xlabel,ylable){
+  library(ggplot2)
+  ggplot()+
+    geom_tile(data = dat, aes(x = d, y = d_h,fill = CI),
+              color="black")+
+    facet_grid(h~a)+
+    scale_fill_gradient2(
+      'coverage', 
+      low = "darkred", mid = "white", 
+      high = "blue", 
+      breaks = c(0,0.5,0.7,0.9, 1), 
+      limits = c(0,1),
+      labels = c("low", "", "medium", "","high"))+
+    scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+    scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+    xlab(xlabel)+
+    ylab(ylabel)+
+    coord_equal()+
+    theme_classic()+
+    theme(legend.pos = "bottom",
+          axis.text = element_text(size = 10), 
+          axis.title = element_text(size = 16), 
+          #legend.title.align = 0.05, 
+          #legend.title = element_text(size = 16), 
+          #legend.text = element_text(size = 14)
+    )
+}

### 2021-11-02 analysis/paper_draft2.Rmd
revisions

@@
     fig_width: 5
-    fig_height: 4
+    fig_height: 5
 header-includes: 
@@
 
-In the US, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
+In the U.S., individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
 During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them.
@@
 Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class.
-The party bringing a _Batson_ challenge bears the burden of proving such illegal bias [@LaFave2020, $\S$ 22.3].
+The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not [@LaFave2020, $\S$ 22.3(d)].
 For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [Wash.
@@
 
-Here, we demonstrate this approach, including the results of simulation studies to test model performance.
+Here, we demonstrate this approach with simulation studies to test model performance.
 Then, we present a software prototype that encodes the same approach with actual historical strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
@@
 We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random.
-Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature that the current trial has but that all the past trials in the historical data do not.
+Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but not of all the past trials in the historical data.
 To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
@@
 In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times we can reach the correct conclusion about the bias parameter based on those credible intervals.
-That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero.
-When $b_{curr}$ equals to zero, the credible interval includes zero, and when $b_{curr}$ is less than zero, the credible interval lies to the left of zero.
-Since we use a weakly-informative prior that pulls the estimate of $b_{current}$ towards zero, we focus not how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
+That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero; when $b_{curr}$ equals to zero, the credible interval includes zero; and when $b_{curr}$ is less than zero, the credible interval lies to the left of zero.
+Since we use a weakly-informative prior that pulls the estimate of $b_{current}$ towards zero, we focus not on how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
 
@@
 
-Figure 1 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively.
+Figures 1-3 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively.
 Here, $b_{hist}$ denotes the bias parameter for the generated historical strike data and $b_{curr}$ denotes the bias parameter for the current data.
@@
 
-```{r c80}
+```{r plotsfig1, eval = FALSE}
 
@@
 
-sims <- readr::read_csv(load_path)
+sims <- readr::read_csv(load_path, show_col_types = FALSE)
 sims$a.f <- as.factor(sims$a)
@@
 C80plot <- coverageplot_func(sims, sims$C80, xlabel, ylabel) +
-           labs(title = "(a) Coverage Rates - 80% Credible Interval")
+           labs(title = "Coverage Rates - 80% Credible Interval")
 
 C90plot <- coverageplot_func(sims, sims$C90, xlabel, ylabel) +
-           labs(title = "(b) Coverage Rates - 90% Credible Interval")
+           labs(title = "Coverage Rates - 90% Credible Interval")
 
 C95plot <- coverageplot_func(sims, sims$C95, xlabel, ylabel) +
-           labs(title = "(c) Coverage Rates - 95% Credible Interval")
+           labs(title = "Coverage Rates - 95% Credible Interval")
+
+# output to figures folder
+
+load_path <- here::here("analysis","figures","C80plot.png")
+ggsave(load_path, plot = C80plot)
+load_path <- here::here("analysis","figures","C90plot.png")
+ggsave(load_path, plot = C90plot)
+load_path <- here::here("analysis","figures","C95plot.png")
+ggsave(load_path, plot = C95plot)
 
@@
 
-```{r fig1a}
-c80plot
-```
+```{r fig1, fig.cap = "Coverage Rates - 80% Credible Interval", fig.align = "center"}
 
-```{r fig1b}
-c90plot
-```
-
-```{r fig1c}
-##| fig.cap=fig1caption
-
-c95plot
+load_path <- here::here("analysis","figures","C80plot.png")
+knitr::include_graphics(load_path)
 
@@
 
-Figure 1(a) depicts the coverage rate with an 80\% credible interval.
+```{r fig2, fig.cap = "Coverage Rates - 90% Credible Interval", fig.align = "center"}
+
+load_path <- here::here("analysis","figures","C90plot.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r fig3, fig.cap = "Coverage Rates - 95% Credible Interval", fig.align = "center"}
+
+load_path <- here::here("analysis","figures","C95plot.png")
+knitr::include_graphics(load_path)
+
+```
+
+Figure 1 depicts the coverage rate with an 80\% credible interval.
 As the first column ($a = 0.1$) depicts, when we place low weight on the historical strike data, we mostly rely on current information.
 When $b_{current}$ is 3, 2 or -2, -3, the coverage rate is high, very close to 1.
+However, when $b_{curr}$ is 1, the coverage rate drops below 0.8.
+If the bias is not strong, our model detects it with a lower accuracy.
 
-However, when $b_{curr}$ is 1, the coverage rate drops below 0.8.
-If the bias is not strong, our model can detect it with a lower accuracy.
-However, as we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
+As we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
 When the bias parameter is 0 (no bias), however, the coverage rate is also around 80\%.
@@
 Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (Figure 1(b)), when $b = 2$, the color is lighter (lower coverage rate) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
+For example, with a 90\% credible interval (Figure 2), when $b = 2$, the coverage rate is lower (the color is lighter) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
 The tradeoff is lower accuracy.
@@
 
-Figure 1(c) depicts coverage rates when using a 95\% credible interval.
+Figure 3 depicts coverage rates when using a 95\% credible interval.
 Here, given $b_{curr} = 0$, the coverage rate is very close to 1.
@@
 
-```{r simltns}
-
-#| fig.cap="Simulation results for simultaneous strikes example.
-Columns are values of bias used to simulate data.
-Blue points denote estimated posterior mean for the bias parameter.
-Blue lines denote the 95% credible interval.
-Red line denotes true value of bias parameters used in simulations."
+```{r simltns, eval = FALSE}
 
@@
 
-simplot
+load_path <- here::here("analysis","figures","simplot.png")
+ggsave(load_path, plot = simplot)
+
+```
+
+```{r fig4}
+
+#| fig.cap="Simulation results for simultaneous strikes example"
+
+load_path <- here::here("analysis","figures","simplot.png")
+knitr::include_graphics(load_path)
 
@@
 We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
+
 Figure 4 depicts the results.
+The columns are values of bias used to simulate the data.
+Blue points denote estimated posterior mean for the bias parameter.
+Blue lines denote the 95% credible interval.
+Red lines denote true value of bias parameters in the simulations.
 
@@
 
-```{r figapp1, out.width='50%', fig.cap = "Density Plots without Historical Strike Data", fig.align='center'}
+```{r figapp1, out.width='50%', fig.align='center'}
+
+#| fig.cap = "Density Plots without Historical Strike Data"
 

### 2021-11-02 analysis/R/coverageplot_func.R

@@ -3,14 +3,17 @@ coverageplot_func <- function(dat,CI,xlabel,ylable){
   ggplot()+
     geom_tile(data = dat, aes(x = d, y = d_h,fill = CI),
               color="black")+
-    facet_grid(h~a)+
-    scale_fill_gradient2(
+    facet_grid(h~a, labeller = labeller(
+              h = label_value, 
+              a = paste0("a = ",label_value))) +
+    scale_fill_gradient(
       'coverage', 
-      low = "darkred", mid = "white", 
+      low = "white",
       high = "blue", 
-      breaks = c(0,0.5,0.7,0.9, 1), 
-      limits = c(0,1),
-      labels = c("low", "", "medium", "","high"))+
+      #breaks = c(0,0.5,0.7,0.9, 1), 
+      #limits = c(0,1),
+      #labels = c("low", "", "medium", "","high")
+      )+
     scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
     scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
     xlab(xlabel)+

### 2021-11-06 analysis/R/coverageplot_func.R
informative facet labels added

@@ -1,11 +1,15 @@
 coverageplot_func <- function(dat,CI,xlabel,ylable){
   library(ggplot2)
+  cust_labeller_a <- function(x) paste0("a = ", x)
+  cust_labeller_h <- function(x) paste0("h = ", x)
   ggplot()+
     geom_tile(data = dat, aes(x = d, y = d_h,fill = CI),
               color="black")+
-    facet_grid(h~a, labeller = labeller(
-              h = label_value, 
-              a = paste0("a = ",label_value))) +
+    facet_grid(h~a, 
+               labeller = labeller(
+                 h = cust_labeller_h, 
+                 a = cust_labeller_a)
+              ) +
     scale_fill_gradient(
       'coverage', 
       low = "white",
@@ -20,7 +24,7 @@ coverageplot_func <- function(dat,CI,xlabel,ylable){
     ylab(ylabel)+
     coord_equal()+
     theme_classic()+
-    theme(legend.pos = "bottom",
+    theme(legend.pos = "right",
           axis.text = element_text(size = 10), 
           axis.title = element_text(size = 16), 
           #legend.title.align = 0.05, 

### 2021-11-06 analysis/R/coverageplot1.R
code snippets for detect/not variant of coverageplots

@@ -1,0 +1,62 @@
+# coverage plot1
+
+p80 <- ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                             fill = as.factor(sig.n)),
+            color="black")+
+  facet_grid(h~a, labeller = label_both)+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  scale_fill_manual(values = c("white", "blue"))+
+  coord_equal() +
+  theme_minimal() +
+  labs(fill = "coverage") +
+  theme(axis.text = element_text(size =10), 
+        #axis.title = element_text(size = 16), 
+        #legend.title.align = 0.05, 
+        #legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+#p80
+
+p90 <- ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                             fill = as.factor(sig.n)),
+            color="black")+
+  facet_grid(h~a, labeller = label_both)+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  scale_fill_manual(values = c("white", "blue"))+
+  coord_equal()+
+  theme_minimal()+
+  labs(fill = "coverage") +
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16),
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+#p90
+
+p95 <- ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                             fill = as.factor(sig.n)),
+            color="black")+
+  facet_grid(h~a, labeller = label_both)+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current b")+
+  ylab("Historical b")+
+  scale_fill_manual(values = c("white", "blue"))+
+  coord_equal()+
+  theme_classic()+  
+  labs(fill = "coverage") +
+  theme(axis.text = element_text(size =10), 
+        axis.title = element_text(size = 16), 
[…]

### 2021-11-06 analysis/paper_draft2.Rmd
revised to asa format; figure widths; revised references

@@
 date: "`r Sys.Date()`"
-author:
-  - Sachin S. Pandya
-#    orcid_id: 0000-0001-7387-1307
-  - Xiaomeng Li
-  - Timothy E. Moore
-#    orcid_id: 0000-0002-9576-0517
-#output: distill::distill_article
+blinded: 0
+authors:
+- name: "[Sachin S. Pandya](https://orcid.org/0000-0001-7387-1307)"
+  #thanks: The authors gratefully acknowledge ...
+  affiliation: University of Connecticut
+  email: sachin.pandya@uconn.edu
+- name: "Xiaomeng Li"
+  affiliation: University of Connecticut
+- name: "[Timothy E. Moore](https://orcid.org/0000-0002-9576-0517)"
+  affiliation: University of Connecticut
 output: 
-  pdf_document:
-    latex_engine: xelatex
-    number_sections: true
-    fig_width: 5
-    fig_height: 5
+  rticles::asa_article:
+    citation_package: "default"
+    keep_tex: TRUE
+    #fig_width: 5
+    #fig_height: 5
+  #output:
+#  pdf_document:
+#    latex_engine: xelatex
+#    number_sections: true
+    #fig_width: 4
+    #fig_height: 4
 header-includes: 
   - \usepackage{amsmath}
-fontsize: 12pt
+#fontsize: 12pt
 bibliography: references.bib
+biblio-style: agsm.bst
 abstract: |
@@
 Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class.
-The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not [@LaFave2020, $\S$ 22.3(d)].
+The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2020, $\S$ 22.3(d)].
 For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [Wash.
@@
 
-In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases.
-(_Flowers v.
-Mississippi_ -@flowers2019, p.
-2243; Wash.
+In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases [@flowers2019, 2243; Wash.
 Rule 37(g)(v); Calif.
-Code of Civil Procedure $\S$ 231.7(d)(3)(G)).
+Code of Civil Procedure $\S$ 231.7(d)(3)(G)].
 Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
@@
 Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the power prior [@Ibrahim2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
+In so doing, we use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
 Finally, we use that posterior distribution as the prior for the bias parameter in the present case and update the posterior based on data on the target attorney's strikes in that present case.
@@
 That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero; when $b_{curr}$ equals to zero, the credible interval includes zero; and when $b_{curr}$ is less than zero, the credible interval lies to the left of zero.
-Since we use a weakly-informative prior that pulls the estimate of $b_{current}$ towards zero, we focus not on how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
+Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus not on how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
 
@@
 
-Figures 1-3 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively.
-Here, $b_{hist}$ denotes the bias parameter for the generated historical strike data and $b_{curr}$ denotes the bias parameter for the current data.
-The values on the right side represent the size of the historical data.
-The values on the top of the plots represent the power prior weight parameter $a$.
+Figures 1-3 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively, for the combinations of $b_{hist}$, $b_{curr}$, the power prior weight parameter $a$, and different sizes of the historical data ($h = {1,2,3}$).
 For any given combination, a darker color shade denotes a higher coverage rate, and vice versa.
@@
 load_path <- here::here("analysis","figures","C80plot.png")
-ggsave(load_path, plot = C80plot)
+ggsave(load_path, plot = C80plot, height = 4, width = 7)
 load_path <- here::here("analysis","figures","C90plot.png")
-ggsave(load_path, plot = C90plot)
+ggsave(load_path, plot = C90plot, height = 4, width = 7)
 load_path <- here::here("analysis","figures","C95plot.png")
-ggsave(load_path, plot = C95plot)
+ggsave(load_path, plot = C95plot, height = 4, width = 7)
 
@@
 
-```{r fig1, fig.cap = "Coverage Rates - 80% Credible Interval", fig.align = "center"}
+```{r fig1, out.width = "95%", fig.cap = "Coverage Rate of bias parameter based on 80 pct credible interval."}
 
@@
 
-```{r fig2, fig.cap = "Coverage Rates - 90% Credible Interval", fig.align = "center"}
+Figure 1 depicts the coverage rate with an 80\% credible interval.
+When we place low weight on the historical strike data ($a = 0.1$), we mostly rely on current strike information.
+When $b_{curr}$ is 3, 2, -2, or -3, the coverage rate is high, very close to 1.
+However, when $b_{curr} = 1$, the coverage rate drops below 0.8.
+If the bias is not strong, our model detects it with a lower accuracy.
+
+As we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when $b_{curr} = 1$.
+When $b_{curr} = 0$ (no bias), however, the coverage rate is also around 80\%.
+Thus, if we use the 80\% credible interval, we have about a 20\% chance of erroneously detecting non-zero bias.
+
+```{r fig2, out.width = "95%", fig.cap = "Coverage Rates - 90 pct. Credible Interval"}
 
@@
 
-```{r fig3, fig.cap = "Coverage Rates - 95% Credible Interval", fig.align = "center"}
+Using a wider credible interval lowers this chance of error.
+For example, with a 90\% credible interval (Figure 2), when $b_{curr} = 2$, the coverage rate is lower than with an 80\% credible interval; when $b_{curr}=0$, the coverage rate is higher.
+The tradeoff is lower accuracy.
+With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
+When the bias parameter is 3, however, the coverage rate is still high.
+
+```{r fig3, out.width = "95%", fig.cap = "Coverage Rates - 95 pct. Credible Interval"}
 
@@
 
-Figure 1 depicts the coverage rate with an 80\% credible interval.
-As the first column ($a = 0.1$) depicts, when we place low weight on the historical strike data, we mostly rely on current information.
-When $b_{current}$ is 3, 2 or -2, -3, the coverage rate is high, very close to 1.
-However, when $b_{curr}$ is 1, the coverage rate drops below 0.8.
-If the bias is not strong, our model detects it with a lower accuracy.
-
-As we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when the bias parameter is 1.
-When the bias parameter is 0 (no bias), however, the coverage rate is also around 80\%.
-Thus, if we use the 80\% credible interval, we have about a 20\% chance of erroneously detecting non-zero bias.
-
-Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (Figure 2), when $b = 2$, the coverage rate is lower (the color is lighter) than with a 80\% credible interval; when $b=0$, the color is darker (higher coverage rate).
-The tradeoff is lower accuracy.
-With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
-When the bias parameter is 3, however, the coverage rate is still high.
-
 Figure 3 depicts coverage rates when using a 95\% credible interval.
@@
 This implies that we can still accurately detect the bias when it is strong.
-However, when the bias is moderate ($b_{curr} = \{2, -2\}$), the coverage rate is only high when we have consistent historical information.
+However, when the bias is moderate ($b_{curr} = \{2, -2\}$), the coverage rate is only high if we have consistent historical information.
 
@@
 
-```{r fig4}
+To test this premise, we conducted the following simulation study.
+We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 32 prospective jurors, 10 of which were member of a cognizable-class (e.g., Black jurors).
+Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
+We then used the model to fit the 50 trials to see if the estimated bias parameters are close to each other.
+We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
 
-#| fig.cap="Simulation results for simultaneous strikes example"
+```{r fig4, out.width = "75%", fig.cap="Simulation results for simultaneous strikes example.
+Columns correspond to values of bias used to simulate data.
+Points and blue lines denote estimated posterior mean and 95% credible interval for bias parameter.
+Red line: true value of the bias parameters used in the 50 simulation s."}
 
@@
 
-To test this premise, we conducted the following simulation study.
-We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 32 prospective jurors, 10 of which were member of a cognizable-class (e.g., Black jurors).
-Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
-We then used the model to fit the 50 trials to see if the estimated bias parameters are close to each other.
-We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
-
 Figure 4 depicts the results.
-The columns are values of bias used to simulate the data.
-Blue points denote estimated posterior mean for the bias parameter.
-Blue lines denote the 95% credible interval.
-Red lines denote true value of bias parameters in the simulations.
-
 We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
@@
                  party = rep(c("PP","PD"),5))
-knitr::kable(df, caption ="Hypothetical Strike Tally")
+
+knitr::kable(df, caption ="Hypothetical Strike Tally", format = "latex")
 
@@
 
-```{r figapp1, out.width='50%', fig.align='center'}
-
-#| fig.cap = "Density Plots without Historical Strike Data"
+```{r figapp1, out.width='50%', fig.cap = "Density Plots without Historical Strike Data"}
 
@@
 
-Finally, if we assign minimal weight to the historical strike data, the density plot updates accordingly (Figure 8).
+Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure 8).
 
@@
 Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that applied to other court's procedure (as we showed above).
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that applied to another court's procedure (as we showed above).
 If not, that attorney's historical strike data in one court could not used to estimate the posterior of that attorney's bias parameter, given a current trial in the other court.

### 2021-11-06 analysis/paper_draft2.Rmd
added keywords; text revisions; facet labels for simplot; revised captions for fig screenshots of app.

@@
 biblio-style: agsm.bst
+keywords: "batson challenge, peremptory strikes"
 abstract: |
@@
 Here, we propose a Bayesian approach for detecting such illegal bias.
-We show how to use historical data on an attorney's use of peremptory strikes in past cases to estimate, in a current trial, how likely an attorney has struck prospective jurors because of cognizable-class membership.
-We also use the _power prior_ to adjust the weight of such historical data in the analysis.
+We show how to combine data on an attorney's strike in a current trial with historical data on that attorney's use of peremptory strikes in past cases to estimate how likely an attorney has struck prospective jurors in the current trial because of their cognizable-class membership.
+In so doing, we use the power prior to adjust the weight of such historical data in the analysis.
 We show how well our model performs on simulated data.
 Finally, we extend this approach with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection.
-We illustrate this application's use with actual historical strike data from a convenience sample of cases from one federal district court.
+We illustrate this application's use with historical strike data from a convenience sample of cases from one court.
 
@@
 Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus not on how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
+For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of $b_{curr}$.
+Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
 
@@
 
-For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of $b_{curr}$.
-Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
-
 Figures 1-3 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively, for the combinations of $b_{hist}$, $b_{curr}$, the power prior weight parameter $a$, and different sizes of the historical data ($h = {1,2,3}$).
@@
 
-```{r fig1, out.width = "95%", fig.cap = "Coverage Rate of bias parameter based on 80 pct credible interval."}
+```{r fig1, out.width = "95%", fig.cap = "Coverage Rates based on 80 pct credible interval.", fig.align='center'}
 
@@
 
-```{r fig2, out.width = "95%", fig.cap = "Coverage Rates - 90 pct. Credible Interval"}
+```{r fig2, out.width = "95%", fig.cap = "Coverage Rates - 90 pct. Credible Interval", fig.align='center'}
 
@@
 
-```{r fig3, out.width = "95%", fig.cap = "Coverage Rates - 95 pct. Credible Interval"}
+```{r fig3, out.width = "95%", fig.cap = "Coverage Rates - 95 pct. Credible Interval", fig.align='center'}
 
@@
                      b = "1", 
-                     sig = "Bias",  # FIX: code as function
+                     sig = "Bias",
                      b.num = 1)
@@
                      b = "2",
-                     sig = "Bias", # FIX: code as function
+                     sig = "Bias",
                      b.num = 2)
@@
 
+cust_labeller_b <- function(x) paste0("b = ", x)
+
 simplot <- ggplot(data = b.com, 
-                  aes(x = sim.n, y = b_s, color = sig))+
-  facet_grid(~b)+
+                  aes(x = sim.n, 
+                      y = b_s, 
+                      color = sig
+                      )
+                  ) +
+  facet_grid(~b, 
+             labeller = labeller(b = cust_labeller_b)
+             ) +
   geom_point(size = 2)+
@@
 Columns correspond to values of bias used to simulate data.
-Points and blue lines denote estimated posterior mean and 95% credible interval for bias parameter.
-Red line: true value of the bias parameters used in the 50 simulation s."}
+Points and horizontal lines denote estimated posterior mean and 95 pct.
+credible interval for bias parameter.
+Vertical lines denote true values of the bias parameters used in the 50 simulations.", fig.align='center'}
 
@@
 We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app with `r R.version$version.string` and Shiny [@R; @shiny].
+We built this app with `r R.version$version.string` and the shiny package [@R; @shiny].
 
@@
 
-```{r figapp1, out.width='50%', fig.cap = "Density Plots without Historical Strike Data"}
+```{r figapp1, out.width='75%', fig.cap = "Screenshot of R-Shiny application showing density plots for bias of prosecutor and defense attorney based on data in 'Current strike tally' table.
+Vertical dotted lines depict 80 pct.
+credible interval.", fig.align='center'}
 
@@
 
-The two vertical dotted lines depict the 80\% credible interval. Here, that credible interval includes zero, indicating no credible to infer bias, given the current strike tally.
+Here, the 80\% credible interval includes zero, indicating no credible to infer bias, given the current strike tally.
 
@@
 
-```{r figapp2, out.width='50%', fig.cap = "Density Plots with Historical Strike Data (Race)", fig.align='center'}
+```{r figapp2, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Race).
+Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.", fig.align='center'}
 
@@
 
-```{r figapp3, out.width='50%', fig.cap = "Density Plots with Historical Strike Data (Gender)", fig.align='center'}
+```{r figapp3, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender).
+Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.
+Equal weight ($a = 1$) applied to historical and current data.", fig.align='center'}
 
@@
 
-Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure 8).
-
-```{r figapp4, out.width='50%', fig.cap = "Density Plots with Historical Strike Data (Gender, Minimal Weight)", fig.align='center'}
+```{r figapp4, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender).
+Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.
+Minimal weight ($a = 0.1$) applied to historical data.", fig.align='center'}
 
@@
 
-We can find both the prior and posterior density become flatter, and the credible intervals plainly includes zero.
+Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure 8).
+Both the prior and posterior density have become flatter and the credible intervals plainly includes zero.
 

### 2021-11-06 analysis/paper_draft2.Rmd
revised references, text spacing after equations

@@
 biblio-style: agsm.bst
+nocite: |
+  @WashR37, @calccp2021
 keywords: "batson challenge, peremptory strikes"
@@
 For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [Wash.
-Rule 37(e); Calif.
-Code of Civil Procedure $\S$ 231.7(d)(1)].
+General Rule.
+37(e); Calif.
+Code Civil Procedure $\S$ 231.7(d)(1)].
 Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
@@
 In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases [@flowers2019, 2243; Wash.
-Rule 37(g)(v); Calif.
-Code of Civil Procedure $\S$ 231.7(d)(3)(G)].
+General Rule 37(g)(v); Calif.
+Code Civil Procedure $\S$ 231.7(d)(3)(G)].
 Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
@@
 
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2018] to incorporate historical strike data.
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2018a; @Kadane2018b] to incorporate historical strike data.
 In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
@@
 
-Following @Kadane2018 and @Barrett2007, we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Following @Kadane2018a and @Barrett2007, we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
 Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories.
@@
 \end{cases} 
-\end{equation}
-
-This Equation \eqref{eq:model1} is equivalent to
+\end{equation} This Equation \eqref{eq:model1} is equivalent to
 
@@
 \label{eq:model3}
-\end{equation}
-
-where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
+\end{equation} where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
 
@@
 \label{eq:powerprior}
-\end{equation}
+\end{equation} where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
-where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
-
-In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
-Unlike @Kadane2018, who assigned $b$ a prior of $\text{Uniform}(-6,6)$, we assume that the law for _Batson_ and similar challenges entails a weakly-informative prior.
+In this paper, we assign a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
+Unlike @Kadane2018a, who assigned $b$ a prior of $\text{Uniform}(-6,6)$, we assume that the law for _Batson_ and similar challenges entails a weakly-informative prior.
 For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
@@
 \label{eq:model4}
-\end{equation}
-
-where
+\end{equation} where
 
@@
 \label{eq:model5}
-\end{equation}
-
-is the likelihood function of $b$ given historical data; $n_{0i}$ is the total number of jury selections (trials) in the historical data; $n_{0j}$ is the total number of peremptory strikes in the historical data; and $\delta_{0ij}$ denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.
+\end{equation} is the likelihood function of $b$ given historical data; $n_{0i}$ is the total number of jury selections (trials) in the historical data; $n_{0j}$ is the total number of peremptory strikes in the historical data; and $\delta_{0ij}$ denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.
 
@@
 
-## Extension: Simultaneous Strikes
+## Simultaneous Strikes
 
@@
 b.com<- rbind(b.com,b1.mean[1,], b2.mean[1,], b3.mean[1,])
-b.com$sim.n <- as.numeric(b.com$sim)
+
+b.com$sim.n <- as.numeric(b.com$sim) # produces warning: NAs introduced by coercion
 b.com$sim.n[151:153] <- 52.5
@@
 In the default setting, the pull-down menus for prosecutor and defense are set to "None".
-As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b = N(0,2)$).
+As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b \sim N(0,2)$).
 

### 2021-11-08 analysis/paper_draft2.Rmd
revised references, text spacing after equations
revised spacing

@@
 Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992; @Gelman1998;@Geweke1992].
+To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].
 
@@
 After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.
-Each graph depicts the prior density plot (colored grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure 5).
+Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure 5).
 

### 2021-11-08 analysis/R/coverageplot1.R
minor revision and addition to R file re: coverage rate simulation

@@ -60,3 +60,30 @@ p95 <- ggplot()+
         legend.text = element_text(size = 14))
 
 #p95
+
+
+#### dino plot #####
+ggplot()+
+  geom_tile(data = sims, aes(x = d, y = d_h,
+                             fill = C80),
+            color="black")+
+  facet_grid(h~a)+
+  scale_fill_gradient2(expression(C80),
+                       low = "white", 
+                       high = "blue",
+                       breaks = c(0,0.2,0.4,0.6,0.8,0.9,1),
+                       limits = c(-0.5, 1.5),
+                       labels = c("Low", "", "", "","", "", "High"))+
+  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
+  xlab("Current d")+
+  ylab("Historical d")+
+  #scale_fill_manual(values = c("white", "blue"))+
+  coord_equal()+
+  theme_classic()+
+  theme(axis.text = element_text(size =14), 
+        axis.title = element_text(size = 16), 
+        legend.title.align = 0.05, 
+        legend.title = element_text(size = 16), 
+        legend.text = element_text(size = 14))
+

### 2021-11-08 analysis/R/mh_sampler_pp.cpp

@@ -1,0 +1,137 @@
+#include <Rcpp.h>
+#include <iostream>
+using namespace Rcpp;
+
+// This is a simple example of exporting a C++ function to R. You can
+// source this function into an R session using the Rcpp::sourceCpp 
+// function (or via the Source button on the editor toolbar). Learn
+// more about Rcpp at:
+//
+//   http://www.rcpp.org/
+//   http://adv-r.had.co.nz/Rcpp.html
+//   http://gallery.rcpp.org/
+//
+
+// [[Rcpp::export]]
+double probcpp(bool cog, int c, int m, double d){
+  if(cog){
+    return log(c)+d-log(c*exp(d)+m);
+  }else{
+    return log(m)-log(c*exp(d)+m);
+  }
+}
+
+
+// x, col 2 should be number cognizable, 3 be total, 4 be cognizable
+// [[Rcpp::export]]
+double Lk_cpp_p(double d,NumericMatrix x, bool L){
+  double lglkr=0;
+  for(int i=0; i < x.nrow();++i){
+    lglkr = lglkr + probcpp(x(i,3),x(i,1),x(i,2)-x(i,1),d);
+  }
+  if(L){
+    return lglkr;
+  }else{
+    return exp(lglkr);
+  }
+}
+
+// [[Rcpp::export]]
+double Priorcpp_p(double theta, double mu, double std,NumericMatrix x,double a0) {
+  return pow(Lk_cpp_p(theta,x,0),a0)*1/(sqrt(2*M_PI)*std)*exp(-pow((theta-mu),2)/(2*pow(std,2)));
+}
+
+// [[Rcpp::export]]
+List make_posterior_p(NumericMatrix x,NumericMatrix x_p,double a0, int niter, 
+                             double theta_start_val, double theta_proposal_sd, double prior_mean, double prior_sd){
+  NumericVector theta(niter);
+  double current_theta;
+  double new_theta;
+  double r;
+  double thresh;
+  double acpt;
+  int acptn=0;
+ 
+  theta[0] = theta_start_val;
+  
+  for(int i = 1; i < niter; ++i){
+    
+    current_theta = theta[i-1];
+    
+    new_theta = rnorm(1,current_theta, theta_proposal_sd)[0];
+    
+    
+    r = log(Priorcpp_p(new_theta,prior_mean,prior_sd,x_p,a0))+Lk_cpp_p(new_theta,x, 1)-log(Priorcpp_p(current_theta,prior_mean,prior_sd,x_p,a0))-
[…]

### 2021-11-08 analysis/R/noseq_simulation.R

@@ -97,30 +97,3 @@ for(j in 1:nrow(sim_result)){
   sim_result$C90[j] = calc_c(theta1$LCL90,theta1$UCL90,sim_result$d[j])
   sim_result$C95[j] = calc_c(theta1$LCL95,theta1$UCL95,sim_result$d[j])
 }
-
-
-
-#### dino plot #####
-ggplot()+
-  geom_tile(data = sims, aes(x = d, y = d_h,
-                             fill = C80),
-            color="black")+
-  facet_grid(h~a)+
-  scale_fill_gradient2(expression(C80),
-                       low = "white", 
-                       high = "blue",
-                       breaks = c(0,0.2,0.4,0.6,0.8,0.9,1),
-                       limits = c(-0.5, 1.5),
-                       labels = c("Low", "", "", "","", "", "High"))+
-  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
-  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
-  xlab("Current d")+
-  ylab("Historical d")+
-  #scale_fill_manual(values = c("white", "blue"))+
-  coord_equal()+
-  theme_classic()+
-  theme(axis.text = element_text(size =14), 
-        axis.title = element_text(size = 16), 
-        legend.title.align = 0.05, 
-        legend.title = element_text(size = 16), 
-        legend.text = element_text(size = 14))

### 2021-11-11 analysis/paper_draft2.Rmd
revised abstract, keywords

@@
   @WashR37, @calccp2021
-keywords: "batson challenge, peremptory strikes"
+keywords: "Batson challenge, peremptory strikes, power prior, Bayesian"
 abstract: |
@@
 Here, we propose a Bayesian approach for detecting such illegal bias.
-We show how to combine data on an attorney's strike in a current trial with historical data on that attorney's use of peremptory strikes in past cases to estimate how likely an attorney has struck prospective jurors in the current trial because of their cognizable-class membership.
-In so doing, we use the power prior to adjust the weight of such historical data in the analysis.
+We show how to incorporate historical data on an attorney's use of peremptory strikes in past cases to estimate how likely that attorney has struck prospective jurors in a current trial because of their cognizable-class membership.
+In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
 We show how well our model performs on simulated data.

### 2021-11-13 analysis/paper_draft2.Rmd
fixed error re: atty aliases, new app screenshot

@@
 
-To illustrate, suppose the strike tally in Table 1 depicts the pattern of strikes in the present case with a prosecuting attorney Aaron Waelchi and defense attorney Lawrance Klocko V (both aliases for actual attorneys in the historical strike data).
+To illustrate, suppose the strike tally in Table 1 depicts the pattern of strikes in the present case with defense attorney Aaron Waelchi and prosecutor Lawrance Klocko V (both aliases for actual attorneys in the historical strike data).
 After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.

### 2021-12-03 analysis/R/coverageplot_func.R
fix typo on ylabel in function

@@ -1,4 +1,4 @@
-coverageplot_func <- function(dat,CI,xlabel,ylable){
+coverageplot_func <- function(dat,CI,xlabel,ylabel){
   library(ggplot2)
   cust_labeller_a <- function(x) paste0("a = ", x)
   cust_labeller_h <- function(x) paste0("h = ", x)

### 2021-12-03 analysis/paper_draft2.Rmd
edit paragraph describing coverage: change to "bias detection rate"

@@
 
-In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the coverage rates, i.e., the proportion of times we can reach the correct conclusion about the bias parameter based on those credible intervals.
-That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero; when $b_{curr}$ equals to zero, the credible interval includes zero; and when $b_{curr}$ is less than zero, the credible interval lies to the left of zero.
-Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus not on how accurately the model can recover the true value of $b_{curr}$, but on how accurately our model can detect that bias.
-For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of $b_{curr}$.
-Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero among the 1000 model fits.
+In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the bias detection rates, i.e., the proportion of times we can reach the correct conclusion about the bias parameter based on those credible intervals.
+That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero; when $b_{curr}$ equals to zero, the credible interval includes zero; and when $b_{curr}$ is less than zero, the credible interval does not include zero.
+Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate), but on how accurately our model can detect that bias.
+For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the $b_{curr}$ is not zero.
+Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits.
 

### 2021-12-03 analysis/R/coverageplot_func.R
edit coverage plot colors for more detail

@@ -1,23 +1,28 @@
 coverageplot_func <- function(dat,CI,xlabel,ylabel){
   library(ggplot2)
+  library(viridis)
   cust_labeller_a <- function(x) paste0("a = ", x)
   cust_labeller_h <- function(x) paste0("h = ", x)
+  CI.d <- cut(CI, breaks = 5, dig.lab = 2)
+  levels(CI.d)[1] <-   "(0,0.1]"
   ggplot()+
-    geom_tile(data = dat, aes(x = d, y = d_h,fill = CI),
-              color="black")+
+    geom_tile(data = dat, aes(x = d, y = d_h,fill = CI.d),
+              color="black", alpha = 0.5)+
     facet_grid(h~a, 
                labeller = labeller(
                  h = cust_labeller_h, 
                  a = cust_labeller_a)
               ) +
-    scale_fill_gradient(
-      'coverage', 
-      low = "white",
-      high = "blue", 
-      #breaks = c(0,0.5,0.7,0.9, 1), 
-      #limits = c(0,1),
-      #labels = c("low", "", "medium", "","high")
-      )+
+
+    scale_fill_manual("Bias Detection Rate", values = rev(viridis(5)))+
+    # scale_fill_gradient(
+    #   'coverage', 
+    #   low = "white",
+    #   high = "blue", 
+    #   #breaks = c(0,0.5,0.7,0.9, 1), 
+    #   #limits = c(0,1),
+    #   #labels = c("low", "", "medium", "","high")
+    #   )+
     scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
     scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
     xlab(xlabel)+

### 2022-04-05 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
Tim and Eric edits to intro and methods

@@
 In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2018a; @Kadane2018b] to incorporate historical strike data.
-In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
-Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
-Finally, we use that posterior distribution as the prior for the bias parameter in the present case and update the posterior based on data on the target attorney's strikes in that present case.
+In this approach, we use two steps.
+In step one, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
+We use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
+We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
+In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posterior distribution developed with the historical information and the power prior.
 
@@
 
-We conducted a simulation study to measure model performance under different values of the bias parameter $b$ and different amounts and kinds of historical information.
+To evaluate the performance of the proposed method, we conducted a simulation study.
+The primary aim of these simulations is to examine the effect of the strength of discounting of the historical data by the power prior ($\alpha$), and the size of the historical data set, on our ability to detect bias.
 
-To do this, we generated strike data for a current trial using seven different values of the bias parameter ($b_{curr} = \{-3,-2,-1,0,1,2,3\}$.
-As introduced above, $b>0$ ($b<0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b=0$ denotes no bias.
+Because we use the power prior, we can control how much we account for the historical strike
+data by modifying the discounting parameter $\alpha$. If $\alpha=1$, the historical strike data is equally
+weighted with the data on strikes in the current trial.
+If $\alpha<1$, the historical data are discounted and weighted proportionally less than the current trial data.
+We assign $\alpha<1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
+Accordingly, in the simulation study, we evaluated different values of $\alpha$: $\alpha=\{0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1\}$ to show how that parameter affected bias detection .
 
-We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ($b_{hist} = \{-3,-2,-1,0,1,2,3\}$).
+We generated historical strike data of three sizes (same, double, and triple the size of
+current data, i.e., data on one previous trial, two previous trials) using seven different values
+of the bias parameter for generating that historical data $b_{hist}=\{-3,-2,-1,0,1,2,3\}$.
 
-We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random.
-Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but not of all the past trials in the historical data.
-To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias in the current case ($b_{curr}$) with a Black defendant.
+To generate data, we generated strike data for a current trial using seven different values of the bias parameter $b_{curr}=\{0,0.5,1,1.5,2, 2.5,3\}$.
+These values were selected to evaluate the strength of the power prior as a function of the compatibility between current and historical trial data.
+As introduced above, $b > 0$ ($b < 0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b = 0$ denotes no bias.
 
-Because we use the power prior, we can control how much we account for the historical strike data by modifying the weight parameter $a$.
-If $a = 1$, the historical strike data is equally weighted with the data on strikes in the current trial.
-If $a < 1$, the historical data matters proportionally less.
-We can assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
-Accordingly, in the simulation study, we tried different values of the weight parameter ($a = \{0.1,0.3,0.5,0.7,1\}$) to show how that parameter affected modeling accuracy.
+We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with
+historical strike data that is incomplete not at random, by allowing $b_{hist}$ to have the same selected bias values, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$ .
+Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but
+not of all the past trials in the historical data. To illustrate, suppose prosecutor with a bias
+value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
+If our historical strike data contains only cases with non-Black defendants, accounting
+for that historical data may lead us to underestimate prosecutor’s bias in the current case
+$b_{curr}$ with a Black defendant. We note that we assume the defense has no bias.
 
-We considered 735 scenarios in total, i.e., the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data and 5 different values of the power prior weight parameter $a$.
-For each scenario, we generated 1000 datasets and fit the model on those datasets.
-We recorded the posterior mean and credible intervals of one thousand model fits for each scenario.
-
-In turn, for three credible interval levels (80\%, 90\% and 95\%), we recorded the bias detection rates, i.e., the proportion of times we can reach the correct conclusion about the bias parameter based on those credible intervals.
-That is, when the true value of $b_{curr}$ is greater than zero, the credible interval lies to the right of zero; when $b_{curr}$ equals to zero, the credible interval includes zero; and when $b_{curr}$ is less than zero, the credible interval does not include zero.
+For our simulation we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $\b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
 Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate), but on how accurately our model can detect that bias.
@@
 Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits.
+For each scenario, we generated 1000 data sets and fit the model on those data sets.
+For each replicate, we generate a MCMC sample of 10000 iterations with a burn-in period of 1000 iterations.
+For each replicate, we calculate the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density (HPD) intervals for $\b_{curr}$.
+
+Additionally, we present sensitivity analysis for the number of available strikes with 6 strikes, 10 strikes, and 15 strikes.
+Thus, in total, we considered 4851 different scenarios, i.e.
+the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data, 11 different values for the power prior weight parameter, and for 3 combinations of the total number of strikes.
 
@@
 
-Figures 1-3 depicts the coverage rates of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively, for the combinations of $b_{hist}$, $b_{curr}$, the power prior weight parameter $a$, and different sizes of the historical data ($h = {1,2,3}$).
-For any given combination, a darker color shade denotes a higher coverage rate, and vice versa.
+Figures 1-3 depicts the bias detection of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively, for the 7 combinations of $b_{hist}$, and 7 combinations of $b_{curr}$, three different levels the power prior discounting parameter $a$, and different sizes of the historical data ($h = {1,2,3}$).
 

### 2022-05-10 analysis/R/data_generation.R
code for simulations and plots

@@ -1,0 +1,2536 @@
+library(rstan)
+library(dplyr)
+
+#sim_data: Function to generate simulated datasets
+#@d_p: bias in prosecution
+#@d_d: bias in defense
+#@round_p: the number round of strikes for prosecution
+#@round_d: the number round fo strikes for defense
+#@total: total number of jurors  
+#@num_cog: number of cognizable people
+sim_data <- function(d_p,d_d,round_p,round_d,total,num_cog){
+  #df0: empty data frame
+  df0 <- data.frame(round = c(1:(round_p+round_d)),
+                    num_cog = NA,
+                    total = NA,
+                    cog = NA,
+                    party = NA)
+  df0$num_cog[1] = num_cog
+  df0$total[1] = total
+  #p_cog=probablity strike a cognizeable class member
+  p_cog <- num_cog*exp(d_p)/(num_cog*exp(d_p)+total-num_cog)
+  df0$cog[1] = rbinom(1,1,p_cog)
+  df0$party[1] = 'PP'
+  df0$num_cog[2] = df0$num_cog[1]-df0$cog[1]
+  df0$total[2] = df0$total[1]-1
+  p_cog <- df0$num_cog[2]*exp(d_d)/(df0$num_cog[2]*exp(d_d)+df0$total[2]-
+                                      df0$num_cog[2])
+  df0$cog[2] = rbinom(1,1,p_cog)
+  df0$party[2] = 'PD'
+  for(i in 2:min(round_p,round_d)){
+    df0$num_cog[2*i-1] = df0$num_cog[2*i-2]-df0$cog[2*i-2]
+    df0$total[2*i-1] = df0$total[2*i-2]-1
+    p_cog <- df0$num_cog[2*i-1]*exp(d_p)/(df0$num_cog[2*i-1]*exp(d_p)+
+                                            df0$total[2*i-1]-df0$num_cog[2*i-1])
+    df0$cog[2*i-1] = rbinom(1,1,p_cog)
+    df0$party[2*i-1] = 'PP'
+    
+    df0$num_cog[2*i] = df0$num_cog[2*i-1]-df0$cog[2*i-1]
+    df0$total[2*i] = df0$total[2*i-1]-1
+    p_cog <- df0$num_cog[2*i]*exp(d_d)/(df0$num_cog[2*i]*exp(d_d)+
+                                          df0$total[2*i]-df0$num_cog[2*i])
+    df0$cog[2*i] = rbinom(1,1,p_cog)
+    df0$party[2*i] = 'PD'
+  }
[…]

### 2022-05-10 analysis/R/model.R

@@ -1,0 +1,1184 @@
+library(MASS)
+library(rstan) 
+
+#stan_mod multiple alpha discounts (e.g. multiple discounting coefficients for sensitivity analysis)
+stan_mod_mult<-'
+data
+{
+// [Current trial]
+//Number of rounds with a strike  
+int<lower=1> N_c;
+//vector cognizeable =1 or not (strikes)
+int cog_c[N_c];
+
+// vector number of cognizeables
+int c_c[N_c];
+// vector number non-cognizeables 
+int m_c[N_c];
+
+// [Historical Trials(pooled as one data set)]
+//Number of rounds with a strike  (H)
+int<lower=1> N_h;
+//vector cognizeable =1 or not (strikes)
+int cog_h[N_h];
+
+// vector number of cognizeables (H)
+int c_h[N_h];
+
+// vector number non-cognizeables (H)
+int m_h[N_h];
+
+//number of historical trials
+int<lower=1> num_htrials;
+
+
+// discounting coefficient 
+real alpha[num_htrials];  
+
+// trial_ind: which historical trial 2=1st historical trial , 3=2nd historical trial....
+int trial_ind[N_h];
+
+
+//inital prior 
+real mean_init;
+real sd_init;
+
+}
+
+parameters
+{
+//b=bias
+real b; 
+}
+
+model
+{
+// priors 
+//initial prior 
+target += normal_lpdf(b|mean_init,sd_init);
+// power prior 
+
+for(j in 1:N_h)
+{
+//trial_ind=2,3,4
+target+=alpha[trial_ind[j]-1]*bernoulli_lpmf(cog_h[j]| 1.0*exp(b)*c_h[j] / (1.0*exp(b)*c_h[j]+m_h[j]));
+}
+
+
+// likelihood
+for (i in 1:N_c)
+{
+target+=bernoulli_lpmf(cog_c[i]| 1.0*exp(b)*c_c[i] / (1.0*exp(b)*c_c[i]+m_c[i]));
+}
+}
+'
+
+fit_mult<- stan_model(model_code=stan_mod_mult) #Need to run this line to use in function pp_sims
+
+
+## Function: pp_sims_mult
+#@dats: list of data sets current and historical (list of lists)
+#@alpha: discounting coefficient in power prior (must have length of number of historical trials)
+#@mean_init: initial prior in power prior's mean 
[…]

### 2022-05-10 analysis/R/simulation_plots_full.R

@@ -1,0 +1,655 @@
+# New simulations - March 2022
+library(dplyr)
+library(tidyr)
+library(stringr)
+
+# STEPS
+
+##STEP1 -  load and reformat output generated by 'model.R' code
+##STEP2 -  combine all into data frame for plotting
+##STEP3 - plot in ggplot2
+
+
+d0_1 <- load('data/01.RData')
+d0_2 <- load('data/02.RData')
+
+#i=1
+for (i in 1:63){
+
+  assign(paste("d0", paste("p", i, sep = ""), sep = "_"),
+         eval(as.symbol(paste("p", i, sep = ""))))
+
+}
+
+
+
+# create holding DF
+out.df.d0 <-  data.frame(
+  iter = rep(paste("p", 1:63, sep = ""), each = 11),
+  b_c = 0,
+  b_h = rep(c(rep(0, 11), rep(0.5, 11), rep(1, 11), rep(1.5, 11),
+              rep(2, 11), rep(2.5, 11), rep(3, 11)), 9),
+  n_h = rep(c(rep(1, 77),rep(2, 77), rep(3, 77)), 3),
+  alpha = rep(1:11, 63),
+  n_strike = c(rep(6, 231), rep(10, 231), rep(15, 231))
+)
+
+get_p_results<-function(p,num_alphas)
+{
+  res=data.frame("post_bias"=rep(NA,num_alphas),
+                 "post_mean"=rep(NA,num_alphas),
+                 "post_sd"=rep(NA,num_alphas),
+                 "post_bias_det80"=rep(NA,num_alphas),
+                 "cp_80"=rep(NA,num_alphas),
+                 "post_bias_det90"=rep(NA,num_alphas),
+                 "cp_90"=rep(NA,num_alphas),
+                 "post_bias_det95"=rep(NA,num_alphas),
+                 "cp_95"=rep(NA,num_alphas))
+  for(i in 1:11)
+  {
+    res[i,"post_bias"]<-as.numeric(p[[i]][[1]]["post_bias"])
+    res[i,"post_mean"]<-as.numeric(p[[i]][[1]]["post_mean"])
+    res[i,"post_sd"]<-as.numeric(p[[i]][[1]]["post_sd"])
+    res[i,"post_bias_det80"]<-as.numeric(p[[i]][[1]]["post_bias_det80"])
[…]

### 2022-05-10 analysis/R/simulation_plots_reduced.R

@@ -1,0 +1,133 @@
+
+# code for generating plots of simulated data from the Rstan models
+# from the 'model.R' code
+
+#--------------------------------------------------------------#
+#--------------------------------------------------------------#
+
+# Data
+comb.df <- read.csv("data/simulation_data_all.csv", header = TRUE) 
+# this .csv file is generated by 'simulation_plots_full.R'
+
+
+# Plotting 
+
+library(ggplot2)
+library(viridis)
+library(RColorBrewer)
+
+df.eq <- data.frame(b_h = seq(from = 0, to = 3, by = 0.5), 
+                    b_c = seq(from = 0, to = 3, by = 0.5))
+
+
+# New facet label names for dose variable
+hist.labs <- c("N. Historical = 1", "N. Historical = 2", "N. Historical = 3")
+names(hist.labs) <- c("1", "2", "3")
+
+# New facet label names for supp variable
+alpha.labs <- c("alpha = 0.1", 
+                "alpha = 0.5",
+                "alpha = 1")
+names(alpha.labs) <- c("0.1", "0.5", "1")
+
+pp_15 <- 
+  comb.df %>%
+  filter(n_strike==15, alpha.val %in% c(0.1, 0.5, 1)) %>%
+  ggplot(data = ., aes(x = b_h, y = b_c, fill = post_bias_det95), 
+         color="black")+
+  geom_tile(color = "black", alpha = 0.9)+
+  geom_rect(data = df.eq, aes (xmin = -0.25, xmax = 0.25, 
+                               ymin = -0.25, ymax = 0.25),
+            color = "black", lwd = 1.2,
+            fill = "transparent")+
+  coord_equal()+
+  geom_text(aes(label = round(post_bias_det95, 2), color = post_bias_det95), 
+            size = 2.3, fontface = "bold")+
+  scale_color_gradient2(low = "grey35",mid = "grey55", 
+                        high = "white", guide = "none")+
+  scale_fill_viridis("Bias Detection Rate",direction=-1,
+                     option = "inferno", limits = c(0, 1))+
+  scale_x_continuous(breaks = c(0:3), expand = c(0,0))+
+  scale_y_continuous(breaks = c(0:3), expand = c(0,0))+
+  facet_grid(n_h~alpha.val, labeller = labeller(n_h = hist.labs, 
+                                                alpha.val = alpha.labs))+
+  theme_classic()+
[…]

### 2022-05-10 analysis/R/simultaneous_plots.R

@@ -1,0 +1,370 @@
+##Purpose: Code to simulate application of simultaneous strikes 
+
+#Libraries to use
+library(rstan)
+library(dplyr)
+
+# Each row in the resulting data set is a strike
+###Function: sim_data######
+#d_p: bias in prosecution
+#d_d: bias in defense
+#round_p: the number round of strikes for prosecution
+#round_d: the number round fo strikes for defense
+#total: total number of jurors 
+#num_cog: number of cognizeable people
+
+
+sim_data <- function(d_p,d_d,round_p,round_d,total,num_cog){
+  #df0: empty data frame
+  df0 <- data.frame(round = c(1:(round_p+round_d)),
+                    num_cog = NA,
+                    total = NA,
+                    cog = NA,
+                    party = NA)
+  df0$num_cog[1] = num_cog
+  df0$total[1] = total
+  #p_cog=probablity strike a cognizeable class member
+  p_cog <- num_cog*exp(d_p)/(num_cog*exp(d_p)+total-num_cog)
+  df0$cog[1] = rbinom(1,1,p_cog)
+  df0$party[1] = 'PP'
+  df0$num_cog[2] = df0$num_cog[1]-df0$cog[1]
+  df0$total[2] = df0$total[1]-1
+  p_cog <- df0$num_cog[2]*exp(d_d)/(df0$num_cog[2]*exp(d_d)+df0$total[2]-
+                                      df0$num_cog[2])
+  df0$cog[2] = rbinom(1,1,p_cog)
+  df0$party[2] = 'PD'
+  for(i in 2:min(round_p,round_d)){
+    df0$num_cog[2*i-1] = df0$num_cog[2*i-2]-df0$cog[2*i-2]
+    df0$total[2*i-1] = df0$total[2*i-2]-1
+    p_cog <- df0$num_cog[2*i-1]*exp(d_p)/(df0$num_cog[2*i-1]*exp(d_p)+
+                                            df0$total[2*i-1]-df0$num_cog[2*i-1])
+    df0$cog[2*i-1] = rbinom(1,1,p_cog)
+    df0$party[2*i-1] = 'PP'
+    
+    df0$num_cog[2*i] = df0$num_cog[2*i-1]-df0$cog[2*i-1]
+    df0$total[2*i] = df0$total[2*i-1]-1
+    p_cog <- df0$num_cog[2*i]*exp(d_d)/(df0$num_cog[2*i]*exp(d_d)+
[…]

### 2022-05-10 analysis/R/simultaneous_rcode.R

@@ -1,0 +1,53 @@
+
+# Code for generating plots of simultaneous strike data
+# output from the 'simultaneous.R' script. 
+#-----------------------------------------------------------------#
+
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(sjPlot)
+
+load("data/simul_4.19.22.RData")
+
+simul_1[[2]]
+simul_2[[2]]
+simul_3[[2]]
+
+
+sm11 <- simul_1[[2]]
+sm22 <- simul_2[[2]]
+sm33 <- simul_3[[2]]
+
+sm11$b <- "b[cur] %~~% 1"
+sm22$b <- "b[cur] %~~% 2"
+sm33$b <- "b[cur] %~~% 3"
+
+comb_sm <- rbind(sm11, sm22, sm33)
+comb_sm$sim <- rep(factor(paste("",50:1, sep =""),
+                    levels = paste("",50:1, sep =""),ordered = TRUE), 
+                   3)
+
+ggplot(data = comb_sm)+
+  facet_grid(~b, labeller = label_parsed)+
+  geom_errorbarh(aes(y = sim,
+                     xmin = LB_CI_95,
+                     xmax = UB_CI_95), height = 0, 
+                 color = "grey")+
+  geom_point(aes(x = mean_b, y = sim))+
+  xlab("Posterior mean of b (bias parameter)")+
+  geom_vline(xintercept = 0, lty = 1, lwd = 1.1)+
+  ylab("Simulation Number")+
+  scale_x_continuous(breaks = c(0:6))+
+  theme_sjplot2()+
+  theme(panel.grid = element_blank(), 
+        axis.title = element_text(size = 14), 
+        axis.text.x = element_text(size = 12), 
+        axis.text.y = element_text(size = 12))
+# ggsave("simultaneous_sims_plot.png", width = 8, height = 8, 
+#        units = "in", dpi = 300)
+
+
+
+
+  

### 2022-05-10 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
edits to paper draft text

@@
 
-Figures 1-3 depicts the bias detection of $b_{curr}$ with an 80\%, 90\%, and 95\% credible interval, respectively, for the 7 combinations of $b_{hist}$, and 7 combinations of $b_{curr}$, three different levels the power prior discounting parameter $a$, and different sizes of the historical data ($h = {1,2,3}$).
-
 ```{r plotsfig1, eval = FALSE}
@@
 
-Figure 1 depicts the coverage rate with an 80\% credible interval.
-When we place low weight on the historical strike data ($a = 0.1$), we mostly rely on current strike information.
-When $b_{curr}$ is 3, 2, -2, or -3, the coverage rate is high, very close to 1.
-However, when $b_{curr} = 1$, the coverage rate drops below 0.8.
-If the bias is not strong, our model detects it with a lower accuracy.
-
-As we place more weight on the historical strike data (as $a$ increases), when the historical strike data is consistent with current information, the accuracy of detecting the bias is also close to 1 when $b_{curr} = 1$.
-When $b_{curr} = 0$ (no bias), however, the coverage rate is also around 80\%.
-Thus, if we use the 80\% credible interval, we have about a 20\% chance of erroneously detecting non-zero bias.
+Figure 1 depicts the bias detection rate with an 95\% credible interval for the n_strikes = 15 scenario.
+When $b_{curr}$ is high (close to 3) the coverage rate is high, very close to 1.
+However, when $b_{curr}$ is lower (close to zero), the bias detection rate drops below 0.8.
+When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$}), our bias detection rate for 95% CI is 0.05 or less.
+As we place more weight on the historical strike data (as $a$ increases), and the historical strike data is consistent with current information (high compatibility between current and historical trials, the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
+Conversely, as alpha decreases, bias detection is only high when there is high bias in the current trial ($b_{curr} >= 1.5).
+Increasing the number of historical trails leads to an increase in bias detection especially when current and historical trials are compatible.
 

### 2022-05-10 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
update to text

@@
 - name: "Xiaomeng Li"
+  affiliation: University of Connecticut
+- name: "Eric Barón"
   affiliation: University of Connecticut
@@
 Conversely, as alpha decreases, bias detection is only high when there is high bias in the current trial ($b_{curr} >= 1.5).
-Increasing the number of historical trails leads to an increase in bias detection especially when current and historical trials are compatible.
+Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible.
+
+As we decrease the number of strikes, from 15 to 10 to 6 (Figs.
+2 and 3), our ability to detect bias decreases.
+Bias detection is only high when alpha is high (close to 1), current trial bias is high, there is compatibility between current and historical trials, and the number of historical trials is large.
 
@@
 
+KEEP/supplement?
 Using a wider credible interval lowers this chance of error.
@@
 
-Figure 3 depicts coverage rates when using a 95\% credible interval.
-Here, given $b_{curr} = 0$, the coverage rate is very close to 1.
-This implies that we are very unlikely to infer bias when there is none.
-When $b_{curr} = 3$, the coverage rate is still high.
-This implies that we can still accurately detect the bias when it is strong.
-However, when the bias is moderate ($b_{curr} = \{2, -2\}$), the coverage rate is only high if we have consistent historical information.
-
-Overall, these simulation results indicate that our model can always accurately detect strong bias.
-As for moderate bias, detection accuracy increases with more historical strike information.
-The choice of credible interval ultimately depends on how conservative we want to be when inferring bias.
+Overall, these simulation results indicate that our model can accurately detect strong bias when it is present in both the current and historical trials.
+increasing alpha strengthens our ability to detect bias when trials are compatible.
+Increasing the number of historical trials, also facilitates bias detection under agreement between trials.
+Larger numbers of strikes make bias detection more likely.
+Under low number of strikes (n_strikes = 6), buas detection is only possible when there is strong bias in the current trial, and alpha is high, and high compatibility.
 
 ## Simultaneous Strikes
+
+(update)
 
@@
 
+(update)
 Finally, in practice, historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
@@
 This presents an issue for the approach presented here if the missing strike data is missing not at random.
+
+TODO: Add piece about small sample effect and fact that intial prior will dominate in small sample setting.
+
+future work could focus on initial prior choice, and measuring compatibility between cur and hist to develope, data driven choice for alpha.
 

### 2022-05-18 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
updated discussion, removed code and added new figures.

@@
 
-Additionally, we present sensitivity analysis for the number of available strikes with 6 strikes, 10 strikes, and 15 strikes.
+Additionally, we present sensitivity analysis for the number of available strikes with 6 strikes, 10 strikes, and 15 strikes for each attorney.
 Thus, in total, we considered 4851 different scenarios, i.e.
@@
 
-```{r b80, eval = FALSE}
+```{r fig1, out.width = "95%", fig.cap = "Coverage Rates based on 95 pct credible interval for 15 strike scenario", fig.align='center'}
 
-# simulated data via "noseq_simulation.R" script
-
-load_path <- here::here("data-sim","sim_resultc.csv")
-
-sims <- readr::read_csv(load_path)
-sims$a.f <- as.factor(sims$a)
-
-sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
-sims$sig <- ifelse(sims$L80 >0 & sims$U80 >0 , "pos", sims$sig) 
-sims$sig <- ifelse(sims$L80 <0 & sims$U80 <0 , "neg", sims$sig)
-
-sims$sig.n <- 0
-sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
-sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
-
-```
-
-```{r b90, eval = FALSE}
-
-sims$a.f <- as.factor(sims$a)
-sims$sig <- ifelse(sims$L90 <0 & sims$U90 >0 , "ns", 1) 
-sims$sig <- ifelse(sims$L90 >0 & sims$U90 >0 , 
-                                    "pos", sims$sig) 
-sims$sig <- ifelse(sims$L90 <0 & sims$U90 <0 , 
-                                    "neg", sims$sig) 
-sims$sig.n <- 0
-sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
-sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
-
-```
-
-```{r b95, eval = FALSE}
-
-sims$a.f <- as.factor(sims$a)
-sims$sig <- ifelse(sims$L95 <0 & sims$U95 >0 , "ns", 1) 
-sims$sig <- ifelse(sims$L95 >0 & sims$U95 >0 , 
-                                    "pos", sims$sig) 
-sims$sig <- ifelse(sims$L95 <0 & sims$U95 <0 , 
-                                    "neg", sims$sig) 
-sims$sig.n <- 0
-sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
-sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)
-
-```
-
-```{r plotsfig1, eval = FALSE}
-
-# simulated data via "noseq_simulation.R" script
-
-load_path <- here::here("data-sim","sim_resultc.csv")
-
-sims <- readr::read_csv(load_path, show_col_types = FALSE)
-sims$a.f <- as.factor(sims$a)
-
-library(latex2exp)
-xlabel = TeX("$b_{curr}$")
-ylabel = TeX("$b_{hist}$")
-
-source(here::here("analysis","R","coverageplot_func.R"))
-
-C80plot <- coverageplot_func(sims, sims$C80, xlabel, ylabel) +
-           labs(title = "Coverage Rates - 80% Credible Interval")
-
-C90plot <- coverageplot_func(sims, sims$C90, xlabel, ylabel) +
-           labs(title = "Coverage Rates - 90% Credible Interval")
-
-C95plot <- coverageplot_func(sims, sims$C95, xlabel, ylabel) +
-           labs(title = "Coverage Rates - 95% Credible Interval")
-
-# output to figures folder
-
-load_path <- here::here("analysis","figures","C80plot.png")
-ggsave(load_path, plot = C80plot, height = 4, width = 7)
-load_path <- here::here("analysis","figures","C90plot.png")
-ggsave(load_path, plot = C90plot, height = 4, width = 7)
-load_path <- here::here("analysis","figures","C95plot.png")
-ggsave(load_path, plot = C95plot, height = 4, width = 7)
-
-```
-
-```{r fig1, out.width = "95%", fig.cap = "Coverage Rates based on 80 pct credible interval.", fig.align='center'}
-
-load_path <- here::here("analysis","figures","C80plot.png")
+load_path <- here::here("analysis","figures","pp15_95CI.png")
 knitr::include_graphics(load_path)
@@
 
+```{r fig2, out.width = "95%", fig.cap = "Coverage Rates based on 95 pct credible interval for 10 strike scenario", fig.align='center'}
+
+load_path <- here::here("analysis","figures","pp10_95CI.png")
+knitr::include_graphics(load_path)
+
+```
+
 As we decrease the number of strikes, from 15 to 10 to 6 (Figs.
-2 and 3), our ability to detect bias decreases.
+2 and 3 respectively), our ability to detect bias decreases.
 Bias detection is only high when alpha is high (close to 1), current trial bias is high, there is compatibility between current and historical trials, and the number of historical trials is large.
 
-```{r fig2, out.width = "95%", fig.cap = "Coverage Rates - 90 pct. Credible Interval", fig.align='center'}
+```{r fig3, out.width = "95%", fig.cap = "Coverage Rates based on 95 pct credible interval for 6 strike scenario", fig.align='center'}
 
-load_path <- here::here("analysis","figures","C90plot.png")
+load_path <- here::here("analysis","figures","pp6_95CI.png")
 knitr::include_graphics(load_path)
@@
 
-```{r simgen, eval = FALSE}
-
-# generate simulated data and output b1.csv, b2.csv, b3.csv
-
-source("R/simultaneous1.R")
-
-```
-
-```{r simltns, eval = FALSE}
-
-# load already-generated data
-
-load_path_b1 <- here::here("data-sim", "b1.csv")
-load_path_b2 <- here::here("data-sim", "b2.csv")
-load_path_b3 <- here::here("data-sim", "b3.csv")
-
-b1 <- read.csv(load_path_b1)
-b2 <- read.csv(load_path_b2)
-b3 <- read.csv(load_path_b3)
-
-library(dplyr)
-
-b1.t <- b1 %>% 
-  select(-X) %>%
-  t() %>% 
-  data.frame() %>% 
-  dplyr::rename_with(~b1$X) %>%
-  mutate(sim = as.character(1:50))
-#b1.t$sim <- factor(b1.t$sim, levels = b1.t$sim, ordered = TRUE)
-
-b2.t <- b2 %>% 
-  select(-X) %>%
-  t() %>% 
-  data.frame() %>% 
-  dplyr::rename_with(~b2$X) %>%
-  mutate(sim =  as.character(1:50))
-#b2.t$sim <- factor(b2.t$sim, levels = b2.t$sim, ordered = TRUE)
-
-b3.t <- b3 %>% 
-  select(-X) %>%
-  t() %>% 
-  data.frame() %>% 
-  dplyr::rename_with(~b2$X) %>%
-  mutate(sim = as.character(1:50))
-#b3.t$sim <- factor(b3.t$sim, levels = b3.t$sim, ordered = TRUE)
-
-colnames(b1.t)[1:3] = c('b_s','2.5%','97.5%')
-colnames(b2.t)[1:3] = c('b_s','2.5%','97.5%')
-colnames(b3.t)[1:3] = c('b_s','2.5%','97.5%')
-
-b1.t$b = "1"
-b2.t$b = "2"
-b3.t$b = "3"
-b.com <- rbind(b1.t, b2.t, b3.t)
-b.com$sig <- ifelse(b.com$`2.5%`> 0, "Bias", 
-                                  "No Bias")
-b.com$b.num <- as.numeric(b.com$b)
-
-b1.mean <-data.frame(b_s = mean(b1.t$b_s), 
-                     `2.5%` =  mean(b1.t$`2.5%`),
-                     `97.5%` = mean(b1.t$`97.5%`),
-                     sim = "mean", 
-                     b = "1", 
-                     sig = "Bias",
-                     b.num = 1)
-
-b2.mean <-data.frame(b_s = mean(b2.t$b_s), 
-                     `2.5%` =  mean(b2.t$`2.5%`),
-                     `97.5%` = mean(b2.t$`97.5%`),
-                     sim = "mean", 
-                     b = "2",
-                     sig = "Bias",
-                     b.num = 2)
-
-b3.mean <-data.frame(b_s = mean(b3.t$b_s), 
-                     `2.5%` =  mean(b3.t$`2.5%`),
-                     `97.5%` = mean(b3.t$`97.5%`),
-                     sim = "mean", 
-                     b = "3",
-                     sig = "Bias",
-                     b.num = 3)
-
-names(b1.mean)<- colnames(b.com)
-names(b2.mean)<- colnames(b.com)
-names(b3.mean)<- colnames(b.com)
-b.com<- rbind(b.com,b1.mean[1,], b2.mean[1,], b3.mean[1,])
-
-b.com$sim.n <- as.numeric(b.com$sim) # produces warning: NAs introduced by coercion
-b.com$sim.n[151:153] <- 52.5
-
-library(ggplot2)
-
-cust_labeller_b <- function(x) paste0("b = ", x)
-
-simplot <- ggplot(data = b.com, 
-                  aes(x = sim.n, 
-                      y = b_s, 
-                      color = sig
-                      )
-                  ) +
-  facet_grid(~b, 
-             labeller = labeller(b = cust_labeller_b)
-             ) +
-  geom_point(size = 2)+
-  geom_errorbar(aes(x = sim.n, ymin = `2.5%`, 
-                    ymax = `97.5%`), width = 0.4)+
-  geom_hline(data = b.com, aes(yintercept = b.num), col = "red", lwd = 1.1) +
-  geom_hline(yintercept = 0, col = "black", lwd = 1, lty = 1) +
-  geom_vline(xintercept = 51.25, col = "black", lwd = .5, lty = 2) +
-  xlab("Simulation")+
-  scale_color_manual("Bias", values = c("blue", "red"))+
-  ylab("Posterior Bias Value")+
-  ylim(c(-0.5, 5))+
-  scale_x_continuous(breaks = c(1:50, 52.5), labels = c(1:50, "Average"))+
-  coord_flip()+
-  theme_classic()+
-  theme(legend.position = "none",
-        #axis.title = element_text(size = 16), 
-        axis.text.x = element_text(size = 12), 
-        axis.text.y = element_text(size = 9))
-
-load_path <- here::here("analysis","figures","simplot.png")
-ggsave(load_path, plot = simplot)
-
-```
-
 To test this premise, we conducted the following simulation study.
-We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 32 prospective jurors, 10 of which were member of a cognizable-class (e.g., Black jurors).
+We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 30 prospective jurors, 15 of which were member of a cognizable-class (e.g., Black jurors).
 Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
-We then used the model to fit the 50 trials to see if the estimated bias parameters are close to each other.
-We conducted the simulation study under three scenarios from moderate bias to strong bias ($b=\{1,2,3\}$).
+We then fit the model to the 50 trials to examine whether the estimated bias parameters across the 50 simulated trials were equivalent.
 
@@
 
-load_path <- here::here("analysis","figures","simplot.png")
+load_path <- here::here("analysis","figures","simultaneous_sims_plot.png")
 knitr::include_graphics(load_path)
@@
 
-Figure 4 depicts the results.
+Figure 4 depicts the results of the simultaneous strike simulations.
 We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
 None of their credible intervals include zero.
-We therefore infer that the order of strikes does not influence the estimate of the bias parameter.
-This confirms that the simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process.
+This demonstrates that the order of strikes does not influence the estimate of the bias parameter, and confirms that the simultaneous-strikes process can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
 
@@
 Statistical methods for _Batson_ and similar legal challenges have the purpose of helping to measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
-The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter.
+The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
@@
 Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that applied to another court's procedure (as we showed above).
-If not, that attorney's historical strike data in one court could not used to estimate the posterior of that attorney's bias parameter, given a current trial in the other court.
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simlataneous strike example shown above).
+There will, howevere be situations when the same model cannot be applied between courts with different jury selection practices, for example [SACHIN ADD EXAMPLE].
 
-(update)
-Finally, in practice, historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
+1. compatability/incompatability (Selection of discouting coeffficient)
+In our simulations we have generated scenarios where there is both compatability ($b_{curr} = b_{hist}$) and incompatability ($b_{curr} \neq b_{hist}$) between current and historical trial data.
+We define compatability here as the agreement in the bias parameter between current and historical trials.
+What could cause incompatability between current and historical trials?
+One possible reason for incompatability could be missing historical data.
+Historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
 For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
@@
 This presents an issue for the approach presented here if the missing strike data is missing not at random.
+Another possible reason for incompatability could be variation in  trial charachteristics which could also influence an attorney's level of bias against cognizable classes.
+Possible causes of difference between trails, however may exist, for example the nature of crime, cognizable status of defendent, federal vs state...
+[SACHIN ADD MORE/Make more 'lawyer' sounding].
+
+2. initial prior/number strikes - the conservatism of the approach
+Under our simulations, the number of strikes and number of histrical trials play a critical role in our ability to detect bias.
+In particular, our results suggest that bias detection rates will be higher in higher strike trials.
+and when more historical data is present.
+This is driven, in part, by our choice of the initial prior of the bias parameter ($b \sim N(0,2)$), which adheres to the notion that, absent any strike data, any given attorney is not biased.
+Recently, Kadane (ref) has highlighted the importance of selecting an appropriate prior for $b$.
+Kadane suggests that a neutral, symmetrical prior be used.
+Such a prior forces the burden of proof to be on the current and historical trial strike data to determine whether bias is present or not.
+In cases, where there is little historical information, and the number of strike is small, only severve bias will be detected by our model.
+
+END PARAGRAPH
+generated a model that can identify bias and incorporates historical trial information - tested this with simulation
+made an app to make the use of the model more accessible - hope is that lawyers will use it in practice
 

### 2022-05-19 analysis/paper_draft2.Rmd
revisions to text of Discussion section

@@
 
-Because we use the power prior, we can control how much we account for the historical strike
-data by modifying the discounting parameter $\alpha$. If $\alpha=1$, the historical strike data is equally
-weighted with the data on strikes in the current trial.
+Because we use the power prior, we can control how much we account for the historical strike data by modifying the discounting parameter $\alpha$.
+If $\alpha=1$, the historical strike data is equally weighted with the data on strikes in the current trial.
 If $\alpha<1$, the historical data are discounted and weighted proportionally less than the current trial data.
@@
 We generated historical strike data of three sizes (same, double, and triple the size of
-current data, i.e., data on one previous trial, two previous trials) using seven different values
-of the bias parameter for generating that historical data $b_{hist}=\{-3,-2,-1,0,1,2,3\}$.
+current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{-3,-2,-1,0,1,2,3\}$.
 
@@
 
-We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with
-historical strike data that is incomplete not at random, by allowing $b_{hist}$ to have the same selected bias values, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$ .
-Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but
-not of all the past trials in the historical data. To illustrate, suppose prosecutor with a bias
-value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
+We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random, by allowing $b_{hist}$ to have the same selected bias values, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$ .
+Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but not of all the past trials in the historical data.
+To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
 If our historical strike data contains only cases with non-Black defendants, accounting
-for that historical data may lead us to underestimate prosecutor’s bias in the current case
-$b_{curr}$ with a Black defendant. We note that we assume the defense has no bias.
+for that historical data may lead us to underestimate prosecutor’s bias in the current case $b_{curr}$ with a Black defendant. For simplicity, we assume the defense attorney has no bias.
 
@@
 When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$}), our bias detection rate for 95% CI is 0.05 or less.
-As we place more weight on the historical strike data (as $a$ increases), and the historical strike data is consistent with current information (high compatibility between current and historical trials, the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
-Conversely, as alpha decreases, bias detection is only high when there is high bias in the current trial ($b_{curr} >= 1.5).
+As we place more weight on the historical strike data (as $a$ increases), and the historical strike data is consistent with current information (high compatibility between current and historical trials), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
+Conversely, as $\alpha$ decreases, bias detection is only high when there is high bias in the current trial ($b_{curr} \geq 1.5$).
 Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible.
@@
 
-As we decrease the number of strikes, from 15 to 10 to 6 (Figs.
+As we decrease the number of strikes, from fifteen to ten to six (Figs.
 2 and 3 respectively), our ability to detect bias decreases.
-Bias detection is only high when alpha is high (close to 1), current trial bias is high, there is compatibility between current and historical trials, and the number of historical trials is large.
+Bias detection is only high when $\alpha$ is high (close to 1), current trial bias is high, there is compatibility between current and historical trials, and the number of historical trials is large.
 
@@
 
-KEEP/supplement?
 Using a wider credible interval lowers this chance of error.
@@
 Overall, these simulation results indicate that our model can accurately detect strong bias when it is present in both the current and historical trials.
-increasing alpha strengthens our ability to detect bias when trials are compatible.
+Increasing alpha strengthens our ability to detect bias when trials are compatible.
 Increasing the number of historical trials, also facilitates bias detection under agreement between trials.
 Larger numbers of strikes make bias detection more likely.
-Under low number of strikes (n_strikes = 6), buas detection is only possible when there is strong bias in the current trial, and alpha is high, and high compatibility.
+Under the low number of strikes (n = 6), bias detection requires strong bias in the current trial, a high alpha, and high compatibility.
 
 ## Simultaneous Strikes
-
-(update)
 
@@
 None of their credible intervals include zero.
-This demonstrates that the order of strikes does not influence the estimate of the bias parameter, and confirms that the simultaneous-strikes process can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
+This demonstrates that the order of strikes does not influence the estimate of the bias parameter.
+Accordingly, the simultaneous-strikes process can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
 
@@
 At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data.
-For example, given the model we used, we required data on the cognitive-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
+For example, given the model we used, we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
+
 Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simlataneous strike example shown above).
-There will, howevere be situations when the same model cannot be applied between courts with different jury selection practices, for example [SACHIN ADD EXAMPLE].
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example shown above).
+In other cases, however, the same model cannot be applied between courts with different jury selection practices.
+For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
 
-1. compatability/incompatability (Selection of discouting coeffficient)
-In our simulations we have generated scenarios where there is both compatability ($b_{curr} = b_{hist}$) and incompatability ($b_{curr} \neq b_{hist}$) between current and historical trial data.
-We define compatability here as the agreement in the bias parameter between current and historical trials.
-What could cause incompatability between current and historical trials?
-One possible reason for incompatability could be missing historical data.
+Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials ($ = |b_{curr} - b_{hist}|$), as our simulations show.
+These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
+
+First, such incompatibility may occur because of missing historical data.
 Historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
@@
 This presents an issue for the approach presented here if the missing strike data is missing not at random.
-Another possible reason for incompatability could be variation in  trial charachteristics which could also influence an attorney's level of bias against cognizable classes.
-Possible causes of difference between trails, however may exist, for example the nature of crime, cognizable status of defendent, federal vs state...
-[SACHIN ADD MORE/Make more 'lawyer' sounding].
 
-2. initial prior/number strikes - the conservatism of the approach
-Under our simulations, the number of strikes and number of histrical trials play a critical role in our ability to detect bias.
-In particular, our results suggest that bias detection rates will be higher in higher strike trials.
-and when more historical data is present.
-This is driven, in part, by our choice of the initial prior of the bias parameter ($b \sim N(0,2)$), which adheres to the notion that, absent any strike data, any given attorney is not biased.
-Recently, Kadane (ref) has highlighted the importance of selecting an appropriate prior for $b$.
-Kadane suggests that a neutral, symmetrical prior be used.
-Such a prior forces the burden of proof to be on the current and historical trial strike data to determine whether bias is present or not.
-In cases, where there is little historical information, and the number of strike is small, only severve bias will be detected by our model.
+Second, even if that past strike data is complete (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics, such as defendant race or the severity of the criminal offenses charged, that take on one value in the current trial but that vary across the historical trials.
+For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
+If so, then bias detection in the current trial will depend on whether we use all the historical data or only the subset of past trials with Black defendants.
+
+The simulations further show that bias detection increases with the number of an attorney's strikes per trial or the number of historical trials.
+This is in part because of our conservative initial prior for the bias parameter ($b \sim N(0,2)$) to accord with the legal burden of proving bias on those who challenge an attorney for strike bias.
+This accords with @Kadane2021, who recommends a neutral and symmetrical prior for estimating bias in _Batson_ challenges [FIX].
+Thus, given little historical strike data for an attorney and a low number of strikes per trial, the initial prior has more sway and, a a result, the model we use will only detect severe bias.
 
@@
 Pandya: Conceptualization, Investigation, Software, Writing; 
-Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Moore: Conceptualization, Supervision, Visualization.
+Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: [INSERT]; Moore: Conceptualization, Supervision, Visualization.
 

### 2022-05-19 analysis/paper_draft2.Rmd
notes/questions re: revisions in draft text

@@
 
+<!-- IS THE ABOVE PARAGRAPH STILL ACCURATE?-->
+
 ## Model Performance on Simulated Data
 
-To evaluate the performance of the proposed method, we conducted a simulation study.
-The primary aim of these simulations is to examine the effect of the strength of discounting of the historical data by the power prior ($\alpha$), and the size of the historical data set, on our ability to detect bias.
+To evaluate the performance of the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
+These simulations primarily show how much our ability to detect bias was affected by how much the power prior ($\alpha$) discounted the weight of the historical data and the size of the historical data.
 
@@
 
-Figure 1 depicts the bias detection rate with an 95\% credible interval for the n_strikes = 15 scenario.
+Figure 1 depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
 When $b_{curr}$ is high (close to 3) the coverage rate is high, very close to 1.
@@
 Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors.
-However, because the 80\% credible interval includes 0, inferring such bias from the strike data alone is unjustified.
+However, because the 80\% credible interval includes zero, inferring such bias from the strike data alone is unjustified.
 
@@
 
-Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials ($ = |b_{curr} - b_{hist}|$), as our simulations show.
-These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
+Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials ($ = |b_{curr} - b_{hist}|$).
+As we explained above when we let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) in our simulations, these bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
 
@@
 For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
-If so, then bias detection in the current trial will depend on whether we use all the historical data or only the subset of past trials with Black defendants.
+If so, then bias detection in the current trial will depend on whether we use all the historical data or only the subset of past trials with Black defendants
 
@@
 This accords with @Kadane2021, who recommends a neutral and symmetrical prior for estimating bias in _Batson_ challenges [FIX].
-Thus, given little historical strike data for an attorney and a low number of strikes per trial, the initial prior has more sway and, a a result, the model we use will only detect severe bias.
 
-END PARAGRAPH
+<!-- what about Kadane's recommendation to use multiple priors for sensitivity? -->
+
+Thus, given little historical strike data for an attorney and a low number of strikes per trial, the initial prior will dominate, and,  as a result, the model we use will only detect severe bias.
+
+# Conclusion
+
 generated a model that can identify bias and incorporates historical trial information - tested this with simulation
@@
 
-TODO: Add piece about small sample effect and fact that intial prior will dominate in small sample setting.
+TODO: Add piece about small sample effect and fact that intial prior will dominate in small sample setting. <!-- WHAT to ADD to LAST paragraph in DISCUSSION section? -->
 
@@
 
-Pandya: Conceptualization, Investigation, Software, Writing; 
-Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: [INSERT]; Moore: Conceptualization, Supervision, Visualization.
+Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: [INSERT]; Moore: Conceptualization, Supervision, Visualization.
 

### 2022-05-20 analysis/paper_draft2.Rmd
updated discussion, internal notes

@@
 For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
-This forces researchers to turn to other data sources, if any, to find that information.
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an issue for the approach presented here if the missing strike data is missing not at random.
+This presents an issue for any approach to detecting strike bias, including the one presented here, if the missing data is missing not at random.
 
-Second, even if that past strike data is complete (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics, such as defendant race or the severity of the criminal offenses charged, that take on one value in the current trial but that vary across the historical trials.
+Second, even with complete past strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity), that take on one value in the current trial but that vary across the historical trials.
 For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
-If so, then bias detection in the current trial will depend on whether we use all the historical data or only the subset of past trials with Black defendants
+If so, then bias detection in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
+
+These reasons, however, only qualify how lawyers can use the approach here to detect strike bias for _Batson_ challenges.
+For any attorney, the posterior distribution of the bias parameter can interact with _Batson_ procedure in three ways.
+First, it can aid the challenger in showing that it has met the initial burden of producing just enough evidence of strike bias to justify shifting the burden onto the striking attorney to provide permissible grounds for the challenged strikes.
+Second, once the striking attorney proffers those reasons, the trial judge can consider the attorney's estimated bias parameter as relevant to how much to believe those reasons.
+Third, the trial judge can rely on the attorney's estimated bias parameter to decide the ultimate issue: whether the alleged illegal strike bias is more likely than not to be true.
 
 The simulations further show that bias detection increases with the number of an attorney's strikes per trial or the number of historical trials.
-This is in part because of our conservative initial prior for the bias parameter ($b \sim N(0,2)$) to accord with the legal burden of proving bias on those who challenge an attorney for strike bias.
-This accords with @Kadane2021, who recommends a neutral and symmetrical prior for estimating bias in _Batson_ challenges [FIX].
-
-<!-- what about Kadane's recommendation to use multiple priors for sensitivity? -->
-
-Thus, given little historical strike data for an attorney and a low number of strikes per trial, the initial prior will dominate, and,  as a result, the model we use will only detect severe bias.
+This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof.
+This initial prior accords with @Kadane2021 [p.
+51], who suggests a prior for estimating _Batson_ strike bias that is at least unimodal, symmetric, and not dependent on the data.
+Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
 
@@
 
-generated a model that can identify bias and incorporates historical trial information - tested this with simulation
-made an app to make the use of the model more accessible - hope is that lawyers will use it in practice
+In this paper, we discussed a Bayesian approach to estimate attorney strike bias that incorporate historical data on an attorney's use of peremptory strikes in past cases and how to use the power prior to adjust the weight of such historical information.
+Our simulations showed that how well our approach detects attorney bias depending on the number of past trials in which that attorney used strikes and the number of strikes per trial.
+Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.
 
-TODO: Add piece about small sample effect and fact that intial prior will dominate in small sample setting. <!-- WHAT to ADD to LAST paragraph in DISCUSSION section? -->
+<!-- future work could focus on initial prior choice, and measuring compatibility between cur and hist to develope, data driven choice for alpha-->
 
-future work could focus on initial prior choice, and measuring compatibility between cur and hist to develope, data driven choice for alpha.
+<!-- what about Kadane's idea of using multiple priors for sensitivity? -->
 

### 2022-05-24 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
edit methods and discussion

@@
 The posterior distribution does not have a closed form.
-Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
+Accordingly, we used the No U-Turn Sampler (NUTS) implemented in Rstan (v 2.21.2, Stan Dev Team, 2020).
 To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].
+For each replicate, we generate an MCMC sample of 10000 iterations with a burn-in period of 1000 iterations of a single chain.
 
@@
 We built this app with `r R.version$version.string` and the shiny package [@R; @shiny].
+In contrast to the simulations, the application was built using Rcpp (version and citation), and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
 
@@
 
-Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials ($ = |b_{curr} - b_{hist}|$).
+Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials (i.e., $ |b_{curr} - b_{hist}|$).
 As we explained above when we let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) in our simulations, these bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
@@
 
-The simulations further show that bias detection increases with the number of an attorney's strikes per trial or the number of historical trials.
+The degree of incompatability between current and historical trial data may also influcence the selection of the value of $\alpha$.
+For example, when there is perfect agreement between current and historical data ($ |b_{curr} - b_{hist}| = 0 $), $\alpha$ should be at its maximum (equal weight between current and historical data).
+This will maximize the model's ability to detect bias, even when the number of strikes is small.
+This effect of increased bias detection can be seen in the simulation plots above, along the positive diagonals of the individual blocks, and is strengthened by increasing the number of historical trials and number of strikes [new figure here].
+
+The simulations further show that bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatability between current and historical trials.
 This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof.
@@
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
+As @Kadane2021 notes, bias detection will sensitive to the choice of the standard deviation of the initial prior.
+Decreasing the magnitude of the this parameter (e.g., $ < 2$), will make the detection of bias more difficult, make the model more conservative, and require more evidence of bias from the likelihood (the current data).
 

### 2022-05-30 analysis/paper_draft2.Rmd
updates to discussion section; added cite for Rcpp; cleaned up

--- 2022-05-20 analysis/paper_draft2.Rmd
@@
 For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
+Furthermore, this prior accords with @Kadane2021 [p.
+51], who suggests a prior for estimating _Batson_ strike bias that is at least unimodal, symmetric, and not dependent on the data.
 
@@
 
-The posterior distribution does not have a closed form.
-Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].
-
-<!-- IS THE ABOVE PARAGRAPH STILL ACCURATE?-->
-
 ## Model Performance on Simulated Data
 
-To evaluate the performance of the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
+To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
+To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].
 These simulations primarily show how much our ability to detect bias was affected by how much the power prior ($\alpha$) discounted the weight of the historical data and the size of the historical data.
@@
 
-We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random, by allowing $b_{hist}$ to have the same selected bias values, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$ .
+We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random, by allowing $b_{hist}$ to have the same selected bias values, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
 Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but not of all the past trials in the historical data.
 To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical strike data contains only cases with non-Black defendants, accounting
-for that historical data may lead us to underestimate prosecutor’s bias in the current case $b_{curr}$ with a Black defendant. For simplicity, we assume the defense attorney has no bias.
+If our historical strike data contains only cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor’s bias in the current case $b_{curr}$ with a Black defendant.
+For simplicity, we assume the defense attorney has no bias.
 
-For our simulation we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $\b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
+For our simulation we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
 Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate), but on how accurately our model can detect that bias.
@@
 For each scenario, we generated 1000 data sets and fit the model on those data sets.
-For each replicate, we generate a MCMC sample of 10000 iterations with a burn-in period of 1000 iterations.
-For each replicate, we calculate the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density (HPD) intervals for $\b_{curr}$.
+For each replicate, we generate a MCMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
+For each replicate, we calculate the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density (HPD) intervals for $b_{curr}$.
 
@@
 
-```{r fig3, out.width = "95%", fig.cap = "Coverage Rates - 95 pct. Credible Interval", fig.align='center'}
+```{r fig3a, out.width = "95%", fig.cap = "Coverage Rates - 95 pct. Credible Interval", fig.align='center', eval = FALSE}
 
@@
 None of their credible intervals include zero.
-This demonstrates that the order of strikes does not influence the estimate of the bias parameter.
+This shows that the order of strikes does not influence the estimate of the bias parameter.
 Accordingly, the simultaneous-strikes process can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
@@
 We built this app with `r R.version$version.string` and the shiny package [@R; @shiny].
+Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
 
-To show how this app might be used by lawyers in real cases, we loaded this app with real strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
+To show how lawyers might use this app in real cases, we loaded this app with real strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 To collect this historical strike data, one of us filed a request with the U.S.
@@
 Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure 8).
-Both the prior and posterior density have become flatter and the credible intervals plainly includes zero.
+Both the prior and posterior density become flatter and the credible interval plainly includes zero.
 
@@
 
-Statistical methods for _Batson_ and similar legal challenges have the purpose of helping to measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
+Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
 The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
-And with a tool like our prototype app, attorneys can use historical strike data to detect bias in real time during jury selection, thus making it easier to raise credible _Batson_ challenges.
 
-At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data.
+Moreover, with a tool like our prototype app, lawyers can use the approach here to detect strike bias for _Batson_ and related challenges, in three ways.
+First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
+Second, it can help the _Batson_ challenger meet the initial burden of producing just enough evidence of strike bias to persuade the trial judge to require the striking attorney to provide (permissible) reasons for their challenged strikes.
+Third, once the striking attorney proffers those reasons, the trial judge can consider the attorney's estimated bias parameter as relevant to how much to believe those proffered reasons as well as whether the alleged illegal strike bias is ultimately more likely than not to be true.
+
+At the same time, the approach here also requires more from those who would collect the appropriate historical strike data.
 For example, given the model we used, we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
@@
 In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example shown above).
-In other cases, however, the same model cannot be applied between courts with different jury selection practices.
+In other cases, however, one cannot apply the same model to courts with different jury selection practices.
 For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
 
-Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials ($ = |b_{curr} - b_{hist}|$).
+Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials, i.e., $\mid b_{curr} - b_{hist} \mid$.
 As we explained above when we let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) in our simulations, these bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
@@
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an issue for any approach to detecting strike bias, including the one presented here, if the missing data is missing not at random.
+This presents an issue for any approach to detecting strike bias, including but not limited to the one presented here, if the missing data is missing not at random.
 
@@
 
-These reasons, however, only qualify how lawyers can use the approach here to detect strike bias for _Batson_ challenges.
-For any attorney, the posterior distribution of the bias parameter can interact with _Batson_ procedure in three ways.
-First, it can aid the challenger in showing that it has met the initial burden of producing just enough evidence of strike bias to justify shifting the burden onto the striking attorney to provide permissible grounds for the challenged strikes.
-Second, once the striking attorney proffers those reasons, the trial judge can consider the attorney's estimated bias parameter as relevant to how much to believe those reasons.
-Third, the trial judge can rely on the attorney's estimated bias parameter to decide the ultimate issue: whether the alleged illegal strike bias is more likely than not to be true.
+In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
+For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), then we should let $\alpha = 1$ (equal weight between current and historical data).
+This will maximize the model's ability to detect bias, even when the number of strikes is small.
+This effect of increased bias detection can be seen in the simulation plots above, along the positive diagonals of the individual blocks, and is strengthened by increasing the number of historical trials and number of strikes.
 
-The simulations further show that bias detection increases with the number of an attorney's strikes per trial or the number of historical trials.
+<!-- new figure here ? -->
+
+Finally, as the simulations also show, bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatibility between current and historical trials.
 This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof.
-This initial prior accords with @Kadane2021 [p.
-51], who suggests a prior for estimating _Batson_ strike bias that is at least unimodal, symmetric, and not dependent on the data.
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
+Thus, as @Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the initial prior.
+As one lowers it (e.g., $\sigma < 2$), bias will be harder to detect, absent more evidence of bias from the likelihood (the current data).
 
@@
 
-<!-- future work could focus on initial prior choice, and measuring compatibility between cur and hist to develope, data driven choice for alpha-->
-
-<!-- what about Kadane's idea of using multiple priors for sensitivity? -->
-
 # Acknowledgements
@@
 
-Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: [INSERT]; Moore: Conceptualization, Supervision, Visualization.
+Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Visualization, Writing - original draft.
 

### 2022-06-07 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd

@@
   affiliation: University of Connecticut
-- name: "Eric Barón"
+- name: "[Eric Barón](https://orcid.org/0000-0003-1934-1932)"
   affiliation: University of Connecticut

### 2022-06-16 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
added text about compatibility plot

@@
 This effect of increased bias detection can be seen in the simulation plots above, along the positive diagonals of the individual blocks, and is strengthened by increasing the number of historical trials and number of strikes.
+Higher bias detection rates are obtained for higher values of alpha, acorss compatibility scenarios (Fig X).
+This effect is most pronounced for scenarios with a low number of strikes and few historical trails (top left paned of fig X).
+There is less of an effect of $\alpha$ on the bias detection rate when bias is high, the number of strikes is high, and the number of historical trials is high  (lower right panel of fig X).
+Overall, though, increasing $\alpha$ will increase bias detection under current and histrotical trial compatiblity.
 

### 2022-06-17 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
updating discussion

@@
 
-In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
-For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), then we should let $\alpha = 1$ (equal weight between current and historical data).
+In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ that is selected.
+For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
 This will maximize the model's ability to detect bias, even when the number of strikes is small.
@@
 
-<!-- new figure here ? -->
+<!-- compatibility figure here ? -->
 
-Finally, as the simulations also show, bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatibility between current and historical trials.
+Finally, bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatibility between current and historical trials.
 This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof.
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
-Thus, as @Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the initial prior.
-As one lowers it (e.g., $\sigma < 2$), bias will be harder to detect, absent more evidence of bias from the likelihood (the current data).
+@Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the prior.
+The sensitivity to the initial prior’s SD, however, is reduced by the inclusion of historical, especially when $\alpha$ is close to 1.
+Fig X shows the effect of the initial prior’s SD on bias detection rate under a number of compatibility scenarios and alpha values.
+We can see that the effect of setting the initial prior SD to 1 (half the value used in our simulations) is most pronounced when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
+As alpha increases, the impact of lowering the initial prior SD is reduced.
+Fig X also illustrates that increasing the prior SD, making the initial prior less informative, has little impact on bias detection rates, regardless of choice of $\alpha$.
+
+<!-- choice of SD figure here ? -->
 

### 2022-06-18 analysis/paper_draft2.Rmd
changed paths for figures; edits to text

@@
 
-load_path <- here::here("analysis","figures","pp15_95CI.png")
+load_path <- here::here("figures","pp15_95CI.png")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("analysis","figures","pp10_95CI.png")
+load_path <- here::here("figures","pp10_95CI.png")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("analysis","figures","pp6_95CI.png")
+load_path <- here::here("figures","pp6_95CI.png")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("analysis","figures","C95plot.png")
+load_path <- here::here("figures","C95plot.png")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("analysis","figures","simultaneous_sims_plot.png")
+load_path <- here::here("figures","simultaneous_sims_plot.png")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot1.png")
-
+load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot1.png")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot2.png")
+load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")
 
@@
 
-load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot3.png")
+load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")
 
@@
 
-load_path <- here::here("analysis","figures","batson_app_screenshots", "batson_app_screenshot4.png")
+load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot4.png")
 
@@
 There is less of an effect of $\alpha$ on the bias detection rate when bias is high, the number of strikes is high, and the number of historical trials is high  (lower right panel of fig X).
-Overall, though, increasing $\alpha$ will increase bias detection under current and histrotical trial compatiblity.
+Overall, though, increasing $\alpha$ will increase bias detection under current and historica trial compatibility.
 
@@
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
-@Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the prior.
-The sensitivity to the initial prior’s SD, however, is reduced by the inclusion of historical, especially when $\alpha$ is close to 1.
-Fig X shows the effect of the initial prior’s SD on bias detection rate under a number of compatibility scenarios and alpha values.
-We can see that the effect of setting the initial prior SD to 1 (half the value used in our simulations) is most pronounced when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
-As alpha increases, the impact of lowering the initial prior SD is reduced.
-Fig X also illustrates that increasing the prior SD, making the initial prior less informative, has little impact on bias detection rates, regardless of choice of $\alpha$.
+As @Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the prior.
+
+This sensitivity, however, reduces after including the historical , especially when $\alpha$ is close to 1.
+Fig X shows the effect of the initial prior’s standard deviation on bias detection rate under a number of compatibility scenarios and alpha values.
 
 <!-- choice of SD figure here ? -->
+
+As Figure X indicates, if we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
+As alpha increases, there is less of an impact of that lower standard deviation.
+Moreover, increasing the prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
 

### 2022-06-18 analysis/paper_draft2.Rmd
moved discussion of sensitivity to initial prior's standard deviation to Results section.

@@
 
+Bias detection will also be sensitive to the choice of the standard deviation of the prior [@Kadane2021].
+The results above depend in part on our choice a conservative initial prior for the bias parameter ($b \sim N(0,2)$).
+Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
+This sensitivity, however, reduces after including the historical data, especially when $\alpha$ is close to 1.
+Fig X shows the effect of the initial prior’s standard deviation on the bias detection rate under a number of compatibility scenarios and alpha values.
+
+<!-- choice of SD figure here ? -->
+
+As Figure X indicates, if we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
+As alpha increases, the initial prior's standard deviation matters less.
+Moreover, increasing the prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
+
 ## Simultaneous Strikes
@@
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an issue for any approach to detecting strike bias, including but not limited to the one presented here, if the missing data is missing not at random.
+This presents an issue for any approach to detecting strike bias, including but not limited to the one here, if the missing data is missing not at random.
 
@@
 There is less of an effect of $\alpha$ on the bias detection rate when bias is high, the number of strikes is high, and the number of historical trials is high  (lower right panel of fig X).
-Overall, though, increasing $\alpha$ will increase bias detection under current and historica trial compatibility.
-
-<!-- compatibility figure here ? -->
-
-Finally, bias detection increases with the number of an attorney's strikes per trial or the number of historical trials, even when there is incompatibility between current and historical trials.
-This is in part because we chose a conservative initial prior for the bias parameter ($b \sim N(0,2)$) to account for _Batson_ challenger's burden of proof.
-Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
-As @Kadane2021 notes, bias detection will be sensitive to the choice of the standard deviation of the prior.
-
-This sensitivity, however, reduces after including the historical , especially when $\alpha$ is close to 1.
-Fig X shows the effect of the initial prior’s standard deviation on bias detection rate under a number of compatibility scenarios and alpha values.
-
-<!-- choice of SD figure here ? -->
-
-As Figure X indicates, if we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
-As alpha increases, there is less of an impact of that lower standard deviation.
-Moreover, increasing the prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
+Overall, though, increasing $\alpha$ will increase bias detection under current and historical trial compatibility.
 

### 2022-06-20 analysis/paper_draft2.Rmd
updated reference dates

@@
 output:
-  rticles::asa_article:
+  bookdown::pdf_book: 
+      base_format: rticles::asa_article
       citation_package: "default"
@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or certain other cognizable classes.
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2020, $\S$ 22.3(d)].
-For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [Wash.
+For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes [Wash.
 General Rule.
@@
 Code Civil Procedure $\S$ 231.7(d)(1)].
-Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
+Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
 
@@
 
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2018a; @Kadane2018b] to incorporate historical strike data.
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data.
 In this approach, we use two steps.
@@
 
-Here, we demonstrate this approach with simulation studies to test model performance.
+Here, we demonstrate this approach with simulation studies.
 Then, we present a software prototype that encodes the same approach with actual historical strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
@@
 The parties exercise their strikes on anyone among this subset of potential jurors in an alternating sequence.
-Once all strikes are used or waived, the remaining prospective jurors are each assigned to seats on the jury until those seats are filled.
+Once all strikes are used or waived, the remaining prospective jurors are each assigned to seats on the jury plus a few alternate jurors.
 
@@
 \begin{equation}  
-\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)
+\pi(b|D_0,\alpha_0) \propto L(b|D_0)^{\alpha_0}\pi_0(b)
 \label{eq:powerprior}
-\end{equation} where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
+\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
 In this paper, we assign a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
-Unlike @Kadane2018a, who assigned $b$ a prior of $\text{Uniform}(-6,6)$, we assume that the law for _Batson_ and similar challenges entails a weakly-informative prior.
-For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
+We assume that the law for _Batson_ and similar challenges entails a weakly-informative prior.
+For our initial prior of $N(\mu = 0, \sigma = 2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
 Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
@@
 For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
-Furthermore, this prior accords with @Kadane2021 [p.
+This prior accords with @Kadane2021 [p.
 51], who suggests a prior for estimating _Batson_ strike bias that is at least unimodal, symmetric, and not dependent on the data.
@@
 \begin{equation}
-L(b|\delta) \propto L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
+L(b|\delta) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{b^2}{8})
 \label{eq:model4}
@@
 To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
-To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].
-These simulations primarily show how much our ability to detect bias was affected by how much the power prior ($\alpha$) discounted the weight of the historical data and the size of the historical data.
+<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much our ability to detect bias was affected the size of the historical data and by how much the power prior ($\alpha$) discounted the weight of the historical data.
+As introduced above, $b > 0$ represents a prosecutor's bias against a prospective juror within the cognizable class, while $b = 0$ denotes no bias.
+For simplicity, we assume the defense attorney has no bias.
 
@@
 To generate data, we generated strike data for a current trial using seven different values of the bias parameter $b_{curr}=\{0,0.5,1,1.5,2, 2.5,3\}$.
-These values were selected to evaluate the strength of the power prior as a function of the compatibility between current and historical trial data.
-As introduced above, $b > 0$ ($b < 0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b = 0$ denotes no bias.
+We selected these values to evaluate the strength of the power prior as a function of the compatibility between current and historical strike data $\mid b_{curr} - b_{hist} \mid$.
+Given perfect compatibility, $\mid b_{curr} - b_{hist} \mid = 0$.
+To allow for imperfect compatibility, we let $b_{hist}$ to have the same selected bias values as $b_{curr}$, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
+We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.
+charge severity) affects $b_curr$ but not of all the past trials in the historical strike data.
 
-We let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random, by allowing $b_{hist}$ to have the same selected bias values, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
-Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature of the current trial but not of all the past trials in the historical data.
-To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not.
-If our historical strike data contains only cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor’s bias in the current case $b_{curr}$ with a Black defendant.
-For simplicity, we assume the defense attorney has no bias.
-
-For our simulation we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
-Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate), but on how accurately our model can detect that bias.
+For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
+Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus on how accurately our model can detect that bias, _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate).
 For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the $b_{curr}$ is not zero.
@@
 
-```{r fig1, out.width = "95%", fig.cap = "Coverage Rates based on 95 pct credible interval for 15 strike scenario", fig.align='center'}
+```{r figbd9515, out.width = "95%", fig.cap = "Bias Detection Rates, 95 pct credible interval, 15 strike scenario", fig.align='center'}
 
@@
 
-Figure 1 depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
+Figure \@ref(fig:figbd9515) depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
 When $b_{curr}$ is high (close to 3) the coverage rate is high, very close to 1.
 However, when $b_{curr}$ is lower (close to zero), the bias detection rate drops below 0.8.
-When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$}), our bias detection rate for 95% CI is 0.05 or less.
-As we place more weight on the historical strike data (as $a$ increases), and the historical strike data is consistent with current information (high compatibility between current and historical trials), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
+When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$), our bias detection rate for 95% credible interval is 0.05 or less.
+As we place more weight on the historical strike data (as $\alpha$ increases), and the historical strike data is consistent with current information (high compatibility), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
 Conversely, as $\alpha$ decreases, bias detection is only high when there is high bias in the current trial ($b_{curr} \geq 1.5$).
@@
 
-```{r fig2, out.width = "95%", fig.cap = "Coverage Rates based on 95 pct credible interval for 10 strike scenario", fig.align='center'}
+```{r figbd9510, out.width = "95%", fig.cap = "Bias Detection Rates (95 pct credible interval, 10 strikes scenario)", fig.align='center'}
 
@@
 
-As we decrease the number of strikes, from fifteen to ten to six (Figs.
-2 and 3 respectively), our ability to detect bias decreases.
-Bias detection is only high when $\alpha$ is high (close to 1), current trial bias is high, there is compatibility between current and historical trials, and the number of historical trials is large.
-
-```{r fig3, out.width = "95%", fig.cap = "Coverage Rates based on 95 pct credible interval for 6 strike scenario", fig.align='center'}
+```{r figbd956, out.width = "95%", fig.cap = "Bias Detection Rates based on 95 pct credible interval for 6 strike scenario", fig.align='center'}
 
@@
 
+As we decrease the number of strikes each attorney has, from fifteen to ten to six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd9506) respectively), our ability to detect bias decreases.
+Bias detection is only high when $\alpha$ is high (close to 1), current trial bias is high, there is high compatibility between current and historical trials, and the number of historical trials is large.
+
 Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (Figure 2), when $b_{curr} = 2$, the coverage rate is lower than with an 80\% credible interval; when $b_{curr}=0$, the coverage rate is higher.
+For example, with a 90\% credible interval (Figure \@ref(fig:figbd9510)), when $b_{curr} = 2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
 The tradeoff is lower accuracy.
 With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
-When the bias parameter is 3, however, the coverage rate is still high.
+When the bias parameter is 3, however, the detection rate is still high.
 
-```{r fig3a, out.width = "95%", fig.cap = "Coverage Rates - 95 pct. Credible Interval", fig.align='center', eval = FALSE}
+Bias detection will also be sensitive to the choice of the standard deviation of the prior.
+The results above depend in part on our choice a conservative initial prior for the bias parameter ($b \sim N(0,2)$).
+Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate.
+As a result, the model we use will only detect severe bias.
+This sensitivity, however, reduces after including the historical data, especially when $\alpha$ is close to 1.
 
-load_path <- here::here("figures","C95plot.png")
+```{r figsdprior, out.width = "95%", fig.cap = "Sensitivity of Initial Prior's Standard Deviation", fig.align='center'}
+
+load_path <- here::here("figures","SD_sensitivity.jpg")
 knitr::include_graphics(load_path)
@@
 
-Overall, these simulation results indicate that our model can accurately detect strong bias when it is present in both the current and historical trials.
-Increasing alpha strengthens our ability to detect bias when trials are compatible.
-Increasing the number of historical trials, also facilitates bias detection under agreement between trials.
-Larger numbers of strikes make bias detection more likely.
-Under the low number of strikes (n = 6), bias detection requires strong bias in the current trial, a high alpha, and high compatibility.
+Figure \@ref(fig:figsdprior) depicts how different values of the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$).
+If we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
+As $\alpha$ increases, the initial prior's standard deviation matters less.
+Moreover, increasing the initial prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
 
-Bias detection will also be sensitive to the choice of the standard deviation of the prior [@Kadane2021].
-The results above depend in part on our choice a conservative initial prior for the bias parameter ($b \sim N(0,2)$).
-Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate, and, as a result, the model we use will only detect severe bias.
-This sensitivity, however, reduces after including the historical data, especially when $\alpha$ is close to 1.
-Fig X shows the effect of the initial prior’s standard deviation on the bias detection rate under a number of compatibility scenarios and alpha values.
+```{r figcmptbly, out.width = "95%", fig.cap = "Compatibility", fig.align='center'}
 
-<!-- choice of SD figure here ? -->
+load_path <- here::here("figures","compatibility_plot.jpg")
+knitr::include_graphics(load_path)
 
-As Figure X indicates, if we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
-As alpha increases, the initial prior's standard deviation matters less.
-Moreover, increasing the prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
+```
+
+Finally, Figure \@ref(fig:figcmptbly) summarizes our simulation results.
+These results indicate that our model can accurately detect strong bias if present in both the current and historical trials.
+Given high compatibility, increasing $\alpha$ improves bias detection, as does increasing the number of historical trials and more strikes per trial.
+Under the low number of strikes per trial (n = 6), bias detection requires strong bias in the current trial, a high $\alpha$, and high compatibility.
 
@@
 In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
-This simultaneous-strikes process can be modeled as a special case of our model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
+This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
 The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
@@
 
-```{r fig4, out.width = "75%", fig.cap="Simulation results for simultaneous strikes example.
-Columns correspond to values of bias used to simulate data.
+```{r fig-simul, out.width = "75%", fig.cap="Simulation results for simultaneous strikes example.
+Columns correspond to values of bias ($b_{curr}$) used to simulate data.
 Points and horizontal lines denote estimated posterior mean and 95 pct.
 credible interval for bias parameter.
-Vertical lines denote true values of the bias parameters used in the 50 simulations.", fig.align='center'}
+Vertical dashed lines denote true value of $b_{curr}$ in the simulations.", fig.align='center'}
 
-load_path <- here::here("figures","simultaneous_sims_plot.png")
+load_path <- here::here("figures","simultaneous_sims_plot.jpg")
 knitr::include_graphics(load_path)
@@
 
-Figure 4 depicts the results of the simultaneous strike simulations.
+Figure \@ref(fig:fig-simul) depicts the results of the simultaneous strike simulations.
 We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
@@
 After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.
-Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure 5).
+Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \@ref(fig:figapp1)).
+Here, the 80\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally.
 
@@
 
-Here, the 80\% credible interval includes zero, indicating no credible to infer bias, given the current strike tally.
-
 To use historical strike data, the user selects a name of the prosecutor or defense attorney from the pull-down menus.
@@
 Once selected, the prior and posterior density plots automatically update to account for the pre-loaded historical strike data for that attorney.
-For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information ($a = 1$).
-The user has two other options: half weight ($a = 0.5$) and minimal weight ($a = 0.2$).
+For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information ($\alpha = 1$).
+The user has two other options: half weight ($\alpha = 0.5$) and minimal weight ($\alpha = 0.2$).
 
 In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to "Equal".
-The density plots update accordingly (Figure 6).
+The density plots update accordingly (Figure \@ref(fig:figapp2)).
 
@@
 Next, suppose the same scenario, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
-If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure 7).
+If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)).
 
@@
 
-Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure 8).
+Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp4)).
 Both the prior and posterior density become flatter and the credible interval plainly includes zero.
@@
 
-Moreover, with a tool like our prototype app, lawyers can use the approach here to detect strike bias for _Batson_ and related challenges, in three ways.
+Moreover, with a tool like our prototype app, attorneys can use the approach here in to detect strike bias for _Batson_ and related challenges, in three ways.
 First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
 Second, it can help the _Batson_ challenger meet the initial burden of producing just enough evidence of strike bias to persuade the trial judge to require the striking attorney to provide (permissible) reasons for their challenged strikes.
-Third, once the striking attorney proffers those reasons, the trial judge can consider the attorney's estimated bias parameter as relevant to how much to believe those proffered reasons as well as whether the alleged illegal strike bias is ultimately more likely than not to be true.
+Third, once the striking attorney proffers those reasons, the trial judge can consider the attorney's estimated bias parameter as relevant to how much to take those proffered reasons as what the attorney actually believed as well as whether the alleged illegal strike bias is ultimately more likely than not to be true.
 
@@
 Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example shown above).
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example above in section \@ref(Simultaneous-Strikes
+)).
 In other cases, however, one cannot apply the same model to courts with different jury selection practices.
@@
 
-Furthermore, the bias detection rate depends on the distance between the bias parameters for the current trial and for all the past (historical) trials, i.e., $\mid b_{curr} - b_{hist} \mid$.
-As we explained above when we let $b_{hist}$ differ from $b_{curr}$ (our parameter of interest) in our simulations, these bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
+Furthermore, the bias detection rate depends on the degree of incompatibility between the bias parameters for the current trial and for the past (historical) trials ($\mid b_{curr} - b_{hist} \mid$).
+These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
 
@@
 And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an issue for any approach to detecting strike bias, including but not limited to the one here, if the missing data is missing not at random.
+This presents an obstacle for any approach to detecting strike bias, including but not limited to the one here, if the missing data is missing not at random.
 
-Second, even with complete past strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity), that take on one value in the current trial but that vary across the historical trials.
+Second, even with complete past strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
 For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
@@
 
-In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ that is selected.
+In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
 For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
-This will maximize the model's ability to detect bias, even when the number of strikes is small.
-This effect of increased bias detection can be seen in the simulation plots above, along the positive diagonals of the individual blocks, and is strengthened by increasing the number of historical trials and number of strikes.
-Higher bias detection rates are obtained for higher values of alpha, acorss compatibility scenarios (Fig X).
-This effect is most pronounced for scenarios with a low number of strikes and few historical trails (top left paned of fig X).
-There is less of an effect of $\alpha$ on the bias detection rate when bias is high, the number of strikes is high, and the number of historical trials is high  (lower right panel of fig X).
-Overall, though, increasing $\alpha$ will increase bias detection under current and historical trial compatibility.
+This will maximize bias detection even when the number of strikes per trial is small.
+This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd9506) along the positive diagonals of the individual blocks.
+
+Bias detection also increases with the number of historical trials and the number of strikes per trial.
+Higher values of $\alpha$ lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \@ref(fig:figcmptbly)).
+The choice of $\alpha$ matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \@ref(fig:figcmptbly)).
+Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
 
@@
 
-In this paper, we discussed a Bayesian approach to estimate attorney strike bias that incorporate historical data on an attorney's use of peremptory strikes in past cases and how to use the power prior to adjust the weight of such historical information.
-Our simulations showed that how well our approach detects attorney bias depending on the number of past trials in which that attorney used strikes and the number of strikes per trial.
+In this paper, we extended a Bayesian approach to estimate attorney strike bias in the use of peremptory challenges by incorporating historical data on that attorney's use of strikes in past cases.
+In so doing, we used the power prior to adjust the weight of such historical information.
+Our simulations showed that how well our approach detects attorney bias depending on the number of past trials in which that attorney used strikes; the number of strikes per trial; the degree of compatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
 Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.

### 2022-06-20 analysis/paper_draft2.Rmd
revised text; figure and equation cross-references

@@
       keep_tex: TRUE
-    #fig_width: 5
-    #fig_height: 5
-  #output:
-#  pdf_document:
-#    latex_engine: xelatex
-#    number_sections: true
-    #fig_width: 4
-    #fig_height: 4
 header-includes: 
@@
 
-In the U.S., individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
+In the United States, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
 During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them.
@@
 Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or certain other cognizable classes.
-The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2020, $\S$ 22.3(d)].
+The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
 For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes [Wash.
@@
 37(e); Calif.
-Code Civil Procedure $\S$ 231.7(d)(1)].
+Code of Civil Procedure $\S$ 231.7(d)(1)].
 Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
@@
 Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
-Typically, they test for the probability of observing a non-zero difference in strike rates by the race or gender of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014; @Gastwirth2013].
+Typically, they test for the probability of observing a non-zero difference in strike rates by the race or gender of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013].
 Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
@@
 To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
-<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much our ability to detect bias was affected the size of the historical data and by how much the power prior ($\alpha$) discounted the weight of the historical data.
+<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much our ability to detect bias was affected the size of the historical data and by how much the power prior discounted the weight of the historical data.
 As introduced above, $b > 0$ represents a prosecutor's bias against a prospective juror within the cognizable class, while $b = 0$ denotes no bias.
@@
 
-Because we use the power prior, we can control how much we account for the historical strike data by modifying the discounting parameter $\alpha$.
+Because we use the power prior, we can control how much we account for the historical strike data by modifying the discounting parameter $\alpha$ (denoted as $\alpha_0$ in Equations \@ref(eq:powerprior) and \@ref(eq:model4)).
 If $\alpha=1$, the historical strike data is equally weighted with the data on strikes in the current trial.
@@
 We assign $\alpha<1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
-Accordingly, in the simulation study, we evaluated different values of $\alpha$: $\alpha=\{0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1\}$ to show how that parameter affected bias detection .
+Accordingly, in the simulation study, we evaluated different values of $\alpha$: $\alpha=\{0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1\}$ to show how that parameter affected bias detection.
 
@@
 To generate data, we generated strike data for a current trial using seven different values of the bias parameter $b_{curr}=\{0,0.5,1,1.5,2, 2.5,3\}$.
-We selected these values to evaluate the strength of the power prior as a function of the compatibility between current and historical strike data $\mid b_{curr} - b_{hist} \mid$.
-Given perfect compatibility, $\mid b_{curr} - b_{hist} \mid = 0$.
-To allow for imperfect compatibility, we let $b_{hist}$ to have the same selected bias values as $b_{curr}$, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
+To evaluate the strength of the power prior as a function of the degree of incompatibility between current and historical strike data ($\lvert b_{curr} - b_{hist} \rvert > 0$), we assigned $b_{hist}$ the same bias values as $b_{curr}$, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
 We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.
@@
 For each replicate, we generate a MCMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
-For each replicate, we calculate the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density (HPD) intervals for $b_{curr}$.
+For each replicate, we calculate the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.
 
-Additionally, we present sensitivity analysis for the number of available strikes with 6 strikes, 10 strikes, and 15 strikes for each attorney.
+Additionally, we present sensitivity analysis for the number of available strikes per trial (6 strikes, 10 strikes, and 15 strikes for each attorney).
 Thus, in total, we considered 4851 different scenarios, i.e.
@@
 Figure \@ref(fig:figbd9515) depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
-When $b_{curr}$ is high (close to 3) the coverage rate is high, very close to 1.
+When $b_{curr}$ is high (close to 3) the bias detection rate is high (very close to 1).
 However, when $b_{curr}$ is lower (close to zero), the bias detection rate drops below 0.8.
-When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$), our bias detection rate for 95% credible interval is 0.05 or less.
+When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$), the bias detection rate for the 95% credible interval is 0.05 or less.
 As we place more weight on the historical strike data (as $\alpha$ increases), and the historical strike data is consistent with current information (high compatibility), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
@@
 
-As we decrease the number of strikes each attorney has, from fifteen to ten to six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd9506) respectively), our ability to detect bias decreases.
-Bias detection is only high when $\alpha$ is high (close to 1), current trial bias is high, there is high compatibility between current and historical trials, and the number of historical trials is large.
+As we decrease the number of strikes each attorney has, from fifteen to ten to six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases.
+Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
 
@@
 Bias detection will also be sensitive to the choice of the standard deviation of the prior.
-The results above depend in part on our choice a conservative initial prior for the bias parameter ($b \sim N(0,2)$).
+The results above depend in part on our choice of a conservative initial prior for the bias parameter ($b \sim N(0,2)$).
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate.
@@
 Figure \@ref(fig:figsdprior) depicts how different values of the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$).
-If we set initial prior's standard deviation to 1 (half the value used in our simulations), that matters most when $\alpha$ is 0.1, i.e., borrowing less information from the historical data.
+In our simulations, we set $\sigma = 2$ for our initial prior.
+If we instead select $\sigma = 1$, that matters most when $\alpha = 0.1$, i.e., when we borrow less information from the historical data.
 As $\alpha$ increases, the initial prior's standard deviation matters less.
-Moreover, increasing the initial prior's standard deviation -- thus making the initial prior less informative -- has little impact on bias detection rates, regardless of choice of $\alpha$.
+Moreover, when we increased the initial prior's standard deviation to make it much less informative ($\sigma = 100$), that had little impact on bias detection rates, regardless of the choice of $\alpha$.
 
@@
 
-## Simultaneous Strikes
+## Simultaneous Strikes {#simul-strikes}
 
@@
 Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example above in section \@ref(Simultaneous-Strikes
-)).
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \@ref(simul-strikes) ).
 In other cases, however, one cannot apply the same model to courts with different jury selection practices.
@@
 
-Furthermore, the bias detection rate depends on the degree of incompatibility between the bias parameters for the current trial and for the past (historical) trials ($\mid b_{curr} - b_{hist} \mid$).
+Furthermore, the bias detection rate depends on the degree of incompatibility between the bias parameters for the current trial and for the past (historical) trials ($\lvert b_{curr} - b_{hist} \rvert$).
 These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
@@
 In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
-For example, given perfect compatibility ($\mid b_{curr} - b_{hist} \mid = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
+For example, given perfect compatibility ($\lvert b_{curr} - b_{hist} \rvert = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
 This will maximize bias detection even when the number of strikes per trial is small.
-This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd9506) along the positive diagonals of the individual blocks.
+This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd956) along the positive diagonals of the individual blocks.
 
@@
 In so doing, we used the power prior to adjust the weight of such historical information.
-Our simulations showed that how well our approach detects attorney bias depending on the number of past trials in which that attorney used strikes; the number of strikes per trial; the degree of compatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
+Our simulations showed that how well our approach detects attorney bias depends on the number of past trials in which that attorney used strikes; the number of strikes per trial; and the degree of incompatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
 Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.

### 2022-06-20 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
updated wording that figure 4 has b_curr=b_hist

@@
 We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.
-charge severity) affects $b_curr$ but not of all the past trials in the historical strike data.
+charge severity) affects $b_{curr}$ but not of all the past trials in the historical strike data.
 
@@
 
-Figure \@ref(fig:figsdprior) depicts how different values of the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$).
+Figure \@ref(fig:figsdprior) depicts how different values of the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where $b_{curr}=b_{hist}$ .
 In our simulations, we set $\sigma = 2$ for our initial prior.

### 2022-06-21 analysis/paper_draft2.Rmd
revised section on initial prior

@@
 
-In this paper, we assign a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.
-We assume that the law for _Batson_ and similar challenges entails a weakly-informative prior.
-For our initial prior of $N(\mu = 0, \sigma = 2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
+In this paper, we assume that the law for _Batson_ and similar challenges entails a weakly-informative initial prior:
+
+\begin{equation}  
+\pi_0(b) \sim Normal(\mu_0 = 0, \sigma_0 = 2)
+\label{eq:initialprior}
+\end{equation}
+
+For this initial prior, we let $\mu_0 = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
 Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
@@
 Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias.
-For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
+For this reason, we use a normal (Gaussian) distribution with $\sigma_0 = 2$.
 This prior accords with @Kadane2021 [p.
@@
 
-As we decrease the number of strikes each attorney has, from fifteen to ten to six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases.
+As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases.
 Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
@@
 Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (Figure \@ref(fig:figbd9510)), when $b_{curr} = 2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
+For example, with a 90\% credible interval (Figure \@ref(fig:figbd9510)), when $b_{curr}=2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
 The tradeoff is lower accuracy.
@@
 
-Bias detection will also be sensitive to the choice of the standard deviation of the prior.
-The results above depend in part on our choice of a conservative initial prior for the bias parameter ($b \sim N(0,2)$).
+Bias detection will also be sensitive to the choice of the standard deviation of the initial prior.
+The results above depend in part on our choice of a conservative initial prior for the bias parameter (Equation \@ref(eq:initialprior)).
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate.
@@
 
-```{r figsdprior, out.width = "95%", fig.cap = "Sensitivity of Initial Prior's Standard Deviation", fig.align='center'}
+```{r figsdprior, out.width = "95%", fig.cap = "Sensitivity to Initial Prior's Standard Deviation", fig.align='center'}
 
@@
 
-Figure \@ref(fig:figsdprior) depicts how different values of the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where $b_{curr}=b_{hist}$ .
-In our simulations, we set $\sigma = 2$ for our initial prior.
-If we instead select $\sigma = 1$, that matters most when $\alpha = 0.1$, i.e., when we borrow less information from the historical data.
+Figure \@ref(fig:figsdprior) depicts how assigning different values for the initial prior’s standard deviation ($\sigma_0 = \{1,2,100\}$) affects the bias detection rate at the 95\% credible interval for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where each attorney had 15 strikes and where $b_{curr} = b_{hist}$.
+In our simulations, we set $\sigma_0 = 2$ for our initial prior.
+If we instead select $\sigma_0 = 1$, that matters most when $\alpha = 0.1$, i.e., when we borrow less information from the historical data.
 As $\alpha$ increases, the initial prior's standard deviation matters less.
@@
 
-Finally, Figure \@ref(fig:figcmptbly) summarizes our simulation results.
+Finally, Figure \@ref(fig:figcmptbly) summarizes our simulation results for bias detection at the 95% credible interval.
 These results indicate that our model can accurately detect strong bias if present in both the current and historical trials.
@@
 In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
-This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
+This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation
+\@ref(eq:model3)), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
 The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.

### 2022-06-21 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
updated equation 5 so that now it's using general notation
changed notation for initial prior in Eq 6; removed subscripts from mu and sigma; changed text to indicate Figure 5 assumes b_curr = b_hist; adjusted formatting of Table 1

@@
 \begin{equation}  
-\pi_0(b) \sim Normal(\mu_0 = 0, \sigma_0 = 2)
+\pi_0(b) \sim Normal(\mu = 0, \sigma = 2)
 \label{eq:initialprior}
@@
 
-For this initial prior, we let $\mu_0 = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
+For this initial prior, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
 Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
@@
 Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias.
-For this reason, we use a normal (Gaussian) distribution with $\sigma_0 = 2$.
+For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
 This prior accords with @Kadane2021 [p.
@@
 \begin{equation}
-L(b|\delta) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{b^2}{8})
+  L(b|\delta) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{(b-\mu)^2}{2\sigma^2})
   \label{eq:model4}
@@
 
-Bias detection will also be sensitive to the choice of the standard deviation of the initial prior.
+Bias detection will also be sensitive to the choice of the standard deviation of the initial prior ($\sigma$).
 The results above depend in part on our choice of a conservative initial prior for the bias parameter (Equation \@ref(eq:initialprior)).
@@
 
-Figure \@ref(fig:figsdprior) depicts how assigning different values for the initial prior’s standard deviation ($\sigma_0 = \{1,2,100\}$) affects the bias detection rate at the 95\% credible interval for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where each attorney had 15 strikes and where $b_{curr} = b_{hist}$.
-In our simulations, we set $\sigma_0 = 2$ for our initial prior.
-If we instead select $\sigma_0 = 1$, that matters most when $\alpha = 0.1$, i.e., when we borrow less information from the historical data.
+Figure \@ref(fig:figsdprior) depicts how assigning different values for the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate at the 95\% credible interval for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where each attorney had 15 strikes and where $b_{curr} = b_{hist}$.
+In our simulations, we set $\sigma = 2$ for our initial prior.
+If we instead select $\sigma = 1$, that matters most when $\alpha = 0.1$, i.e., when we borrow less information from the historical data.
 As $\alpha$ increases, the initial prior's standard deviation matters less.
@@
 
-Finally, Figure \@ref(fig:figcmptbly) summarizes our simulation results for bias detection at the 95% credible interval.
+Finally, Figure \@ref(fig:figcmptbly) summarizes our simulation results for bias detection at the 95% credible interval, given perfect compatibility ($b_{curr} = b_{hist}$).
 These results indicate that our model can accurately detect strong bias if present in both the current and historical trials.
 Given high compatibility, increasing $\alpha$ improves bias detection, as does increasing the number of historical trials and more strikes per trial.
-Under the low number of strikes per trial (n = 6), bias detection requires strong bias in the current trial, a high $\alpha$, and high compatibility.
+Under the low number of strikes per trial (n = 6), bias detection requires strong bias in the current trial and a high $\alpha$.
 
@@
 
-knitr::kable(df, caption ="Hypothetical Strike Tally", format = "latex")
+knitr::kable(df, caption ="Hypothetical Strike Tally", format = "latex", booktabs = TRUE)
 

### 2022-06-24 analysis/paper_draft2.Rmd
Update paper_draft2.Rmd
Fixed typo on number of strikes.

@@
 
-Figure \@ref(fig:figsdprior) depicts how assigning different values for the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate at the 95\% credible interval for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where each attorney had 15 strikes and where $b_{curr} = b_{hist}$.
+Figure \@ref(fig:figsdprior) depicts how assigning different values for the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate at the 95\% credible interval for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where each attorney had 10 strikes and where $b_{curr} = b_{hist}$.
 In our simulations, we set $\sigma = 2$ for our initial prior.

### 2022-09-20 analysis/paper_draft3.Rmd
draft for R&R
added reference to Gastwirth 2005

--- 2022-06-24 analysis/paper_draft2.Rmd
@@
   US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, gender, or membership in certain other cognizable classes.
-Here, we propose a Bayesian approach for detecting such illegal bias.
+Here, we extend a Bayesian approach for detecting such illegal bias.
 We show how to incorporate historical data on an attorney's use of peremptory strikes in past cases to estimate how likely that attorney has struck prospective jurors in a current trial because of their cognizable-class membership.
@@
 
-Following @Kadane2018a and @Barrett2007, we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Following @Kadane2018a and @Barrett2007 [see also @Gastwirth2005, 183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
 Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories.

### 2022-09-27 analysis/paper_draft3.Rmd
Update paper_draft3.Rmd
revision edits (equations, b_hist)

@@
 abstract: |
-  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, gender, or membership in certain other cognizable classes.
+  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
 Here, we extend a Bayesian approach for detecting such illegal bias.
@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or certain other cognizable classes.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes.
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
@@
 Code Civil Procedure $\S$ 231.7(d)(3)(G)].
-Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender.
-Typically, they test for the probability of observing a non-zero difference in strike rates by the race or gender of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013].
+Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or sex.
+Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013].
 Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
@@
 If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).
-If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
+If "sex" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
 In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
@@
 \begin{equation}  
-\pi(b|D_0,\alpha_0) \propto L(b|D_0)^{\alpha_0}\pi_0(b)
+\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
 \label{eq:powerprior}
-\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
+\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
@@
 \begin{equation}
-  L(b|\delta) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{(b-\mu)^2}{2\sigma^2})
+  P(b|\delta,\delta_0) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{(b-\mu)^2}{2\sigma^2})
   \label{eq:model4}
@@
 We generated historical strike data of three sizes (same, double, and triple the size of
-current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{-3,-2,-1,0,1,2,3\}$.
+current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
 

### 2022-09-27 analysis/paper_draft3.Rmd
Update paper_draft3.Rmd
Updated notation, sex/gender, and some comments. Will show in meeting on Friday

@@
 \label{eq:model3}
-\end{equation} where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.
+\end{equation} where $n_{i}$ is the total number of jury selections (trials), $n_j$ is the total number of peremptory strikes, and $\delta=(\delta_{11},\dots,\delta_{12},\dots,\delta_{{n_i}{n_j}})$
 
@@
 \label{eq:powerprior}
-\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
+\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\dots,\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
@@
 \begin{equation}
-  P(b|\delta,\delta_0) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{(b-\mu)^2}{2\sigma^2})
+  \pi(b|\delta,\delta_0) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{(b-\mu)^2}{2\sigma^2})
   \label{eq:model4}
@@
 As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b \sim N(0,2)$).
+Historical data will be included if the name(s) of a prosection or defense attorney are selected from the pull-down menu.
+Subsequently, the amount of historical data to be borrowed can be selected by the choice of $\alpha$.
 

### 2022-09-30 analysis/paper_draft3.Rmd
added text on Batson procedure

@@
 
-Following @Kadane2018a and @Barrett2007 [see also @Gastwirth2005, 183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Following others (e.g., @Kadane2018a; @Barrett2007; @Gastwirth2005, 183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
 Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories.
@@
 First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
-Second, it can help the _Batson_ challenger meet the initial burden of producing just enough evidence of strike bias to persuade the trial judge to require the striking attorney to provide (permissible) reasons for their challenged strikes.
-Third, once the striking attorney proffers those reasons, the trial judge can consider the attorney's estimated bias parameter as relevant to how much to take those proffered reasons as what the attorney actually believed as well as whether the alleged illegal strike bias is ultimately more likely than not to be true.
+
+```{r batsonworkflowfig}
+
+library(DiagrammeR)
+grViz(diagram = "digraph flowchart {
+  rankdir = LR
+  node [fontname = arial, shape = box]
+  tab1 [label = '@@1']
+  tab2 [label = '@@2']
+  tab3 [label = '@@3']
+
+  tab1 -> tab2 -> tab3;
+}
+
+  [1]: 'prima facie case'
+  [2]: 'proffered reasons'    
+  [3]: 'prove bias'    
+  ", height = '100%', width = '100%')
+
+```
+
+Second, if raised, the tool can help the _Batson_ challenger meet its evidentiary burdens at key steps of a _Batson_ challenge.
+A _Batson_ challenge typically proceeds in a three-step sequence (Figure \@ref(fig:batsonworkflowfig)).
+First, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient inference of illegal bias.
+If so, then at step two, the striking attorney must produce (permissible) reasons for their challenged strikes.
+Third, the challenger must discredit those proffered reasons and ultimately persuade the trial judge to find that it is more likely than not that the striking attorney acted with illegal bias.
+
+To meets its evidentiary burdens at the first and third steps, the attorney can introduce the striking attorney's estimated bias parameter as relevant statistical evidence.
+In this way, the tool helps detect bias.
+But like any statistical evidence submitted for a _Batson_ challenge, it cannot alone be the basis of deciding a _Batson_ challenge.
+That is because it alone cannot account for the veracity of the striking attorney's reasons offered at step two and because the _Batson_ procedure allows for any type of circumstantial evidence of bias at steps one and three.
 

### 2022-09-30 R/render_paper.R

@@ -1,0 +1,5 @@
+# render paper to seperate output folder
+
+output_path <- here::here("analysis","_output")
+input_path <- here::here("analysis","paper_draft3.Rmd")
+rmarkdown::render(input = input_path, output_dir = output_path)

### 2022-09-30 analysis/paper_draft3.Rmd
Update paper_draft3.Rmd
Updated equations and parenthesis styling

@@
 \begin{equation}
-Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
+Pr(\delta_{ij}| b) = \left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
 \label{eq:model2}
@@
 \begin{equation}  
-L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
+L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}\left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
 \label{eq:model3}
@@
 \begin{equation}
-  \pi(b|\delta,\delta_0) \propto L(b|\delta)(L(b|\delta_0))^{\alpha_0}\exp(-\frac{(b-\mu)^2}{2\sigma^2})
+ \pi(b|\delta,\delta_0) \propto L(b|\delta)L(b|\delta_0)^{\alpha_0}\exp\left(-\frac{(b-\mu)^2}{2\sigma^2}\right)
   \label{eq:model4}
@@
 \begin{equation}
-L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}
+L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}\left(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{\delta_{0ij}} \left(\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{1-\delta_{0ij}}
 \label{eq:model5}

### 2022-10-27 analysis/paper_draft3.Rmd
added figure of Batson workflow; discussion of testing for alpha sensitivity; right to attend criminal trials

--- 2022-09-30 analysis/paper_draft3.Rmd
@@
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
+
+```{r batsonworkflowfig, fig.cap="Batson workflow"}
+
+library(DiagrammeR)
+grViz(diagram = "digraph flowchart {
+  rankdir = LR
+  node [fontname = arial, shape = rectangle]
+  tab1 [label = '@@1']
+  tab2 [label = '@@2']
+  tab3 [label = '@@3']
+
+  tab1 -> tab2 -> tab3;
+}
+
+  [1]: 'prima facie case'
+  [2]: 'proffered reasons'    
+  [3]: 'prove bias'    
+  ", height = '100%', width = '100%')
+
+```
+
+A _Batson_ challenge typically proceeds in a three-step sequence (Figure \@ref(fig:batsonworkflowfig)).
+First, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
+If so, then at step two, the striking attorney must produce (permissible) reasons for their challenged strikes.
+Third, the challenger must discredit those proffered reasons and ultimately persuade the trial judge to find that it is more likely than not that the striking attorney acted with illegal bias.
+
 For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes [Wash.
-General Rule.
-37(e); Calif.
-Code of Civil Procedure $\S$ 231.7(d)(1)].
+General Rule 37(e); Calif.
+Code of Civil Procedure $\S$ 231.7(d)(1); Conn.
+Superior Court Rule 5-12(d)].
 Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
@@
 Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
-For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2021.
+For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2022.
 
@@
 
-Following others (e.g., @Kadane2018a; @Barrett2007; @Gastwirth2005, 183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Following others [e.g., @Kadane2018a; @Barrett2007; @Gastwirth2005, 183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
 Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories.
@@
 
-We generated historical strike data of three sizes (same, double, and triple the size of
-current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
+We generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
 
@@
 In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
-This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation
-\@ref(eq:model3)), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
+This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation \@ref(eq:model3)), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
 The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
@@
 
-Moreover, with a tool like our prototype app, attorneys can use the approach here in to detect strike bias for _Batson_ and related challenges, in three ways.
+Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges, in three ways.
 First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
 
-```{r batsonworkflowfig}
-
-library(DiagrammeR)
-grViz(diagram = "digraph flowchart {
-  rankdir = LR
-  node [fontname = arial, shape = box]
-  tab1 [label = '@@1']
-  tab2 [label = '@@2']
-  tab3 [label = '@@3']
-
-  tab1 -> tab2 -> tab3;
-}
-
-  [1]: 'prima facie case'
-  [2]: 'proffered reasons'    
-  [3]: 'prove bias'    
-  ", height = '100%', width = '100%')
-
-```
-
-Second, if raised, the tool can help the _Batson_ challenger meet its evidentiary burdens at key steps of a _Batson_ challenge.
-A _Batson_ challenge typically proceeds in a three-step sequence (Figure \@ref(fig:batsonworkflowfig)).
-First, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient inference of illegal bias.
-If so, then at step two, the striking attorney must produce (permissible) reasons for their challenged strikes.
-Third, the challenger must discredit those proffered reasons and ultimately persuade the trial judge to find that it is more likely than not that the striking attorney acted with illegal bias.
-
-To meets its evidentiary burdens at the first and third steps, the attorney can introduce the striking attorney's estimated bias parameter as relevant statistical evidence.
+Second, if raised, the tool can help the _Batson_ challenger meet its evidentiary burdens at steps one and three of a _Batson_ challenge (Figure \@ref(fig:batsonworkflowfig)).
+To meets its evidentiary burdens at the first and third steps, the attorney can introduce the posterior distribution of the striking attorney's bias parameter as relevant statistical evidence.
 In this way, the tool helps detect bias.
-But like any statistical evidence submitted for a _Batson_ challenge, it cannot alone be the basis of deciding a _Batson_ challenge.
-That is because it alone cannot account for the veracity of the striking attorney's reasons offered at step two and because the _Batson_ procedure allows for any type of circumstantial evidence of bias at steps one and three.
+But it in no way suffices for deciding a _Batson_ challenge.
+Such statistical evidence alone cannot account for the veracity of the striking attorney's reasons offered at step two.
+Besides, _Batson_ procedure allows for any type of circumstantial evidence of bias at steps one and three.
 
@@
 
-Second, even with complete past strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
+At the same time, US law currently affords an avenue for reducing this missing data problem by systematically collecting historical strike data prospectively.
+A criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from the voir dire of prospective jurors.
+Similarly, the public has a presumptive First Amendment right to attend jury voir dire in criminal cases, even if prosecutor and defendant agree to exclude the public [@press1984; @presley2010].
+Upon exercise of these rights, courtroom observors can attend jury voir dire and collect relevant strike data based on what they observe in real time.
+Even if historical strike data is hard to collect retrospectively, courtroom observers can over time build up a dataset for strike data that they would have been less likely to obtain from court records after the fact.
+
+Second, even with complete strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
 For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
@@
 Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
+
+More importantly, in cases where lawyers contest the degree of compatibility, those lawyers should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
+(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about compatibility are warranted and the historical strike data, without more, is less helpful in steps one and three of the _Batson_ procedure than it would otherwise would be.
+One virtue of this approach is that the choice of $\alpha$ explicitly denotes the degree of weight placed on the historical strike data in the analysis.
 

### 2022-10-28 R/batson_workflow_fig.R
generate Batson workflow figure

@@ -1,0 +1,33 @@
+library(DiagrammeR)
+
+# string as graphviz input
+
+flow <- "
+
+digraph flowchart {
+  rankdir = LR
+  node [fontname = arial, shape = rectangle]
+  tab1 [label = '@@1', color = red]
+  tab2 [label = '@@2', shape = ellipse]
+  tab3 [label = '@@3', color = red]
+
+  tab1 -> tab2 -> tab3;
+  
+}
+  
+  [1]: 'prima facie case'
+  [2]: 'proffered reasons'    
+  [3]: 'prove bias'   
+  "
+
+# export graph
+
+save_path <- here::here("figures","batson_workflow_fig.png")
+
+library(DiagrammeRsvg)
+library(rsvg)
+library(htmltools)
+
+grViz(flow) %>% export_svg %>% 
+  charToRaw %>% # convert SVG string into raw data
+  rsvg_png(file = save_path)

### 2022-10-28 analysis/_output/paper_draft_anonymous.Rmd
render paper to _output folder

--- 2022-06-24 analysis/paper_draft2.Rmd
@@
 date: "`r Sys.Date()`"
-blinded: 0
+blinded: 1
 authors:
@@
 
-# Acknowledgements
+<!--# Acknowledgements
 
@@
 Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Visualization, Writing - original draft.
-
+-->
 # References

### 2022-10-28 R/render_paper.R

@@ -1,4 +1,4 @@
-# render paper to seperate output folder
+# render paper to separate output folder
 
 output_path <- here::here("analysis","_output")
 input_path <- here::here("analysis","paper_draft3.Rmd")

### 2022-10-28 analysis/paper_draft3.Rmd
updated batson workflow fig; added appendix heading

@@
 
-```{r batsonworkflowfig, fig.cap="Batson workflow"}
+```{r batsonworkflowfig, fig.cap="Batson workflow", fig.align = 'center', out.width="50%", fig.pos = "h"}
 
-library(DiagrammeR)
-grViz(diagram = "digraph flowchart {
-  rankdir = LR
-  node [fontname = arial, shape = rectangle]
-  tab1 [label = '@@1']
-  tab2 [label = '@@2']
-  tab3 [label = '@@3']
-
-  tab1 -> tab2 -> tab3;
-}
-
-  [1]: 'prima facie case'
-  [2]: 'proffered reasons'    
-  [3]: 'prove bias'    
-  ", height = '100%', width = '100%')
+load_path <- here::here("figures","batson_workflow_fig.png")
+knitr::include_graphics(load_path)
 
@@
 # References
+
+<div id="refs"></div>
+
+# Appendix
+
+Insert appendix here.

### 2022-10-29 analysis/paper_draft3.Rmd
updated figure and text about batson workflow; moved portion of protoytpe discussion to appendix

@@
 
-```{r batsonworkflowfig, fig.cap="Batson workflow", fig.align = 'center', out.width="50%", fig.pos = "h"}
-
-load_path <- here::here("figures","batson_workflow_fig.png")
-knitr::include_graphics(load_path)
-
-```
-
-A _Batson_ challenge typically proceeds in a three-step sequence (Figure \@ref(fig:batsonworkflowfig)).
-First, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
-If so, then at step two, the striking attorney must produce (permissible) reasons for their challenged strikes.
-Third, the challenger must discredit those proffered reasons and ultimately persuade the trial judge to find that it is more likely than not that the striking attorney acted with illegal bias.
+A _Batson_ challenge typically proceeds in a four-step sequence.
+The attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
+Second, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
+If so, then at the next step, the striking attorney must proffer permissible reasons for their challenged strikes.
+Finally, the challenger must discredit those proffered reasons and ultimately persuade the trial judge to find that it is more likely than not that the striking attorney acted with illegal bias.
 
@@
 Code Civil Procedure $\S$ 231.7(d)(3)(G)].
-Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or sex.
+Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex.
 Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013].
@@
 For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2022.
+
+In turn, the inference of bias based in part on historical strike data is relevant to lawyers at three of the four steps of the _Batson_ challenge, i.e., for deciding whether to argue a _Batson_ violation, for establishing a prima facie case, and for ultimately proving that attorney strike bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
+
+```{r batsonworkflowfig, fig.cap="Batson workflow", fig.align = 'center', out.width="100%", fig.pos = "h"}
+
+load_path <- here::here("figures","batson_workflow_fig2.pdf")
+knitr::include_graphics(load_path)
+
+```
 
@@
 
-## The Software Prototype
+## The Software Prototype {#prototype}
 
@@
 Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
+To show how lawyers might use this app in real cases, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 
-To show how lawyers might use this app in real cases, we loaded this app with real strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 To collect this historical strike data, one of us filed a request with the U.S.
@@
 Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
+In the Appendix, we present additional illustrations wherein the app is used to detect gender bias and where we assign less weight to the historical strike data.
 
-Next, suppose the same scenario, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
+# Discussion
+
+Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
+The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
+By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
+
+Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges, in three ways.
+First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
+
+Second, if raised, the tool can help the _Batson_ challenger meet the  evidentiary burdens of presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+To meet those burdens, the attorney can introduce the posterior distribution of the striking attorney's bias parameter as relevant statistical evidence.
+In this way, the tool helps detect bias.
+
+But as with _any_ inference from historical strike data, the approach here in no way suffices for deciding a _Batson_ challenge.
+Such statistical evidence alone cannot account for the veracity of the striking attorney's proferred reasons.
+Besides, _Batson_ procedure allows for any type of circumstantial evidence of bias for purposes of the ultimately proving or disproving illegal bias.
+
+At the same time, the approach here also requires more from those who would collect the appropriate historical strike data.
+For example, given the model we used, we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
+
+Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \@ref(simul-strikes) ).
+In other cases, however, one cannot apply the same model to courts with different jury selection practices.
+For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
+
+Furthermore, the bias detection rate depends on the degree of incompatibility between the bias parameters for the current trial and for the past (historical) trials ($\lvert b_{curr} - b_{hist} \rvert$).
+These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
+
+First, such incompatibility may occur because of missing historical data.
+Historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
+For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
+And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
+This presents an obstacle for any approach to detecting strike bias, including but not limited to the one here, if the missing data is missing not at random.
+
+US law currently affords a way to ameliorate this missing data problem.
+The public has a presumptive First Amendment right to attend jury voir dire in criminal cases, even if prosecutor and defendant agree to exclude the public.
+Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from the voir dire of prospective jurors [@press1984; @presley2010].
+If these rights are asserted, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time.
+Thus, though historical strike data is hard to collect retrospectively, such courtroom observers can prospectively build up a dataset for strike data that they would have been less likely to obtain from court records after the fact.
+
+Second, even with complete strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
+For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
+If so, then bias detection in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
+
+In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
+Given perfect compatibility ($\lvert b_{curr} - b_{hist} \rvert = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
+This will maximize bias detection even when the number of strikes per trial is small.
+This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd956) along the positive diagonals of the individual blocks.
+
+Bias detection also increases with the number of historical trials and the number of strikes per trial.
+Higher values of $\alpha$ lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \@ref(fig:figcmptbly)).
+The choice of $\alpha$ matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \@ref(fig:figcmptbly)).
+Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
+
+In cases where lawyers contest the degree of compatibility, those lawyers can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
+(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about incompatibility are warranted and the historical strike data, without more, is less helpful to proving a _Batson_ violation than it would otherwise would be.
+In any case, one virtue of this approach is that the choice of $\alpha$ explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.
+
+# Conclusion
+
+In this paper, we extended a Bayesian approach to estimate attorney strike bias in the use of peremptory challenges by incorporating historical data on that attorney's use of strikes in past cases.
+In so doing, we used the power prior to adjust the weight of such historical information.
+Our simulations showed that how well our approach detects attorney bias depends on the number of past trials in which that attorney used strikes; the number of strikes per trial; and the degree of incompatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
+Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.
+
+# Acknowledgements
+
+## CRediT Author Statement
+
+Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Visualization, Writing - original draft.
+
+# References
+
+<div id="refs"></div>
+
+# Appendix
+
+Here, we extend the discussion in section \@ref(prototype) of the paper by presenting additional illustrations of how a lawyer might use the prototype app.
+
+Suppose the same scenario discussed in section \@ref(prototype), except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
 If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)).
@@
 Both the prior and posterior density become flatter and the credible interval plainly includes zero.
-
-# Discussion
-
-Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
-The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
-
-Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges, in three ways.
-First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
-
-Second, if raised, the tool can help the _Batson_ challenger meet its evidentiary burdens at steps one and three of a _Batson_ challenge (Figure \@ref(fig:batsonworkflowfig)).
-To meets its evidentiary burdens at the first and third steps, the attorney can introduce the posterior distribution of the striking attorney's bias parameter as relevant statistical evidence.
-In this way, the tool helps detect bias.
-But it in no way suffices for deciding a _Batson_ challenge.
-Such statistical evidence alone cannot account for the veracity of the striking attorney's reasons offered at step two.
-Besides, _Batson_ procedure allows for any type of circumstantial evidence of bias at steps one and three.
-
-At the same time, the approach here also requires more from those who would collect the appropriate historical strike data.
-For example, given the model we used, we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
-
-Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \@ref(simul-strikes) ).
-In other cases, however, one cannot apply the same model to courts with different jury selection practices.
-For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
-
-Furthermore, the bias detection rate depends on the degree of incompatibility between the bias parameters for the current trial and for the past (historical) trials ($\lvert b_{curr} - b_{hist} \rvert$).
-These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
-
-First, such incompatibility may occur because of missing historical data.
-Historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
-For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
-And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an obstacle for any approach to detecting strike bias, including but not limited to the one here, if the missing data is missing not at random.
-
-At the same time, US law currently affords an avenue for reducing this missing data problem by systematically collecting historical strike data prospectively.
-A criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from the voir dire of prospective jurors.
-Similarly, the public has a presumptive First Amendment right to attend jury voir dire in criminal cases, even if prosecutor and defendant agree to exclude the public [@press1984; @presley2010].
-Upon exercise of these rights, courtroom observors can attend jury voir dire and collect relevant strike data based on what they observe in real time.
-Even if historical strike data is hard to collect retrospectively, courtroom observers can over time build up a dataset for strike data that they would have been less likely to obtain from court records after the fact.
-
-Second, even with complete strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
-For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
-If so, then bias detection in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
-
-In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
-For example, given perfect compatibility ($\lvert b_{curr} - b_{hist} \rvert = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
-This will maximize bias detection even when the number of strikes per trial is small.
-This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd956) along the positive diagonals of the individual blocks.
-
-Bias detection also increases with the number of historical trials and the number of strikes per trial.
-Higher values of $\alpha$ lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \@ref(fig:figcmptbly)).
-The choice of $\alpha$ matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \@ref(fig:figcmptbly)).
-Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
-
-More importantly, in cases where lawyers contest the degree of compatibility, those lawyers should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about compatibility are warranted and the historical strike data, without more, is less helpful in steps one and three of the _Batson_ procedure than it would otherwise would be.
-One virtue of this approach is that the choice of $\alpha$ explicitly denotes the degree of weight placed on the historical strike data in the analysis.
-
-# Conclusion
-
-In this paper, we extended a Bayesian approach to estimate attorney strike bias in the use of peremptory challenges by incorporating historical data on that attorney's use of strikes in past cases.
-In so doing, we used the power prior to adjust the weight of such historical information.
-Our simulations showed that how well our approach detects attorney bias depends on the number of past trials in which that attorney used strikes; the number of strikes per trial; and the degree of incompatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
-Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.
-
-# Acknowledgements
-
-## CRediT Author Statement
-
-Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Visualization, Writing - original draft.
-
-# References
-
-<div id="refs"></div>
-
-# Appendix
-
-Insert appendix here.

### 2022-10-29 figures/batson_workflow_fig2.tex
updated figure and text about batson workflow; moved portion of protoytpe discussion to appendix

\documentclass[border=0.01cm]{standalone}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,decorations,arrows,calc,arrows.meta,fit,positioning}
\usetikzlibrary{shadows}
\tikzset{
    -Latex,auto,semithick,
    state/.style ={rectangle, draw, minimum width = 0.7 cm},
    point/.style = {circle, draw, opacity=1, inner sep=0.05cm, fill,node contents={}},
    bidirected/.style={Latex-Latex,dashed},
    el/.style = {inner sep=2pt, align=left, sloped}
}

\begin{document}

\begin{tikzpicture}
  \node (A) at (0,0) [label=below:{argue \emph{Batson}},point,fill={blue}];	
  \node (B1) at (4,0) [label=below:{prima facie case},
                       point,fill={blue}];	
  \node (B2) at (8,0) [label=below:{proffer reasons},
                       point, fill={white}];
  \node (B3) at (12,0) [label=below:{prove bias},point,fill={blue}];
   \node (M) at (6,2) [state, drop shadow, fill={white}] {Model};
   \node (PB) at (6,1) [state] {bias estimate};
    \node (H) at (3,3) [state] {historical data};
    \node (C) at (9,3) [state] {current strikes};

    \path (A) edge (B1);
    \path (B1) edge (B2); 
    \path (B2) edge (B3);
    \path (M) edge (PB);
    \path (PB) edge (A) [dashed, blue];
    \path (PB) edge (B1) [dashed, blue];
    \path (PB) edge (B3) [dashed, blue];
    \path (H) edge (M);
    \path (C) edge (M);

\end{tikzpicture}

\end{document}

### 2022-10-31 analysis/paper_draft3.Rmd
updated batson figure caption and related text

@@
 The attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
-Second, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
+If challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
 If so, then at the next step, the striking attorney must proffer permissible reasons for their challenged strikes.
-Finally, the challenger must discredit those proffered reasons and ultimately persuade the trial judge to find that it is more likely than not that the striking attorney acted with illegal bias.
+Finally, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
 
@@
 
-In turn, the inference of bias based in part on historical strike data is relevant to lawyers at three of the four steps of the _Batson_ challenge, i.e., for deciding whether to argue a _Batson_ violation, for establishing a prima facie case, and for ultimately proving that attorney strike bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
+In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, establishing a prima facie case, and proving that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
 
-```{r batsonworkflowfig, fig.cap="Batson workflow", fig.align = 'center', out.width="100%", fig.pos = "h"}
+```{r batsonworkflowfig, fig.cap="Inference from Historical Strike Data for  Batson Challenge", fig.align = 'center', out.width="100%", fig.pos = "h"}
 
@@
 Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-To show how lawyers might use this app in real cases, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
+To show how lawyers might use this app in real cases in real time during selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 
@@
 
-Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges, in three ways.
-First, the posterior distribution of the bias parameter can help the lawyer decide in real time whether it is worth raising a  _Batson_ challenge at all.
+Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
+The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether it is worth arguing a  _Batson_ challenge at all.
+If argued, the attorney can introduce that posterior distribution as relevant evidence for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+To be sure, such evidence alone cannot account for the veracity of the striking attorney's proferred reasons.
+Besides, _Batson_ procedure allows any type of evidence, not just statistical evidence, for ultimately proving or disproving illegal bias in the particular case.
+Thus, as with any use of historical strike data, the approach here cannot always determine whether a _Batson_ challenge should succeed or fail.
 
-Second, if raised, the tool can help the _Batson_ challenger meet the  evidentiary burdens of presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
-To meet those burdens, the attorney can introduce the posterior distribution of the striking attorney's bias parameter as relevant statistical evidence.
-In this way, the tool helps detect bias.
-
-But as with _any_ inference from historical strike data, the approach here in no way suffices for deciding a _Batson_ challenge.
-Such statistical evidence alone cannot account for the veracity of the striking attorney's proferred reasons.
-Besides, _Batson_ procedure allows for any type of circumstantial evidence of bias for purposes of the ultimately proving or disproving illegal bias.
-
-At the same time, the approach here also requires more from those who would collect the appropriate historical strike data.
+The approach here also requires more from those who would collect the appropriate historical strike data.
 For example, given the model we used, we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.

### 2022-11-16 analysis/paper_draft3.Rmd
Edits to intro and discussion
I think this is looking really good.

@@
 Finally, we extend this approach with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection.
-We illustrate this application's use with historical strike data from a convenience sample of cases from one court.
+We illustrate this application's use with real, historical strike data from a convenience sample of cases from one court.
 
@@
 
-A _Batson_ challenge typically proceeds in a four-step sequence.
-The attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
-If challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
-If so, then at the next step, the striking attorney must proffer permissible reasons for their challenged strikes.
-Finally, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
+A _Batson_ challenge typically proceeds in a four-step sequence (Figure 1): 1) The attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias; 2) If challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias; 3) If so, then at the next step, the striking attorney must proffer permissible reasons for their challenged strikes; 4) Finally, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
+These steps(better word?) ensure that no single piece of evidence is used to determine bias.
 
@@
 Then, we present a software prototype that encodes the same approach with actual historical strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
+[editme] sentence on how this can be used to detect biasbut it is till up to attorney to bring a challenge, and, ultimately,  the judge to decide if bias is present.
 
@@
 To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
-<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much our ability to detect bias was affected the size of the historical data and by how much the power prior discounted the weight of the historical data.
+<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much our ability to detect bias was affected by the size of the historical data and by how much the power prior discounted the weight of the historical data.
 As introduced above, $b > 0$ represents a prosecutor's bias against a prospective juror within the cognizable class, while $b = 0$ denotes no bias.
@@
 Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.
+[editme] maybe we can supplement some of this detail, so we state where the data come from, and that we de-itentified it. More details in the supplement.
 
@@
 Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
+[editme] The goal of these methods is not to _prove_ bias, but rather to lend support to an inferrence of bias and a subsequent _Batson_ challenge.
 The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
+[editme] sentence on why this is appealing
 
 Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
-The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether it is worth arguing a  _Batson_ challenge at all.
-If argued, the attorney can introduce that posterior distribution as relevant evidence for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+**The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether it is worth arguing a  _Batson_ challenge at all.** If argued, the attorney can introduce that posterior distribution as relevant evidence for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
 To be sure, such evidence alone cannot account for the veracity of the striking attorney's proferred reasons.
@@
 
-In cases where lawyers contest the degree of compatibility, those lawyers can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about incompatibility are warranted and the historical strike data, without more, is less helpful to proving a _Batson_ violation than it would otherwise would be.
-In any case, one virtue of this approach is that the choice of $\alpha$ explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.
+**In cases where lawyers contest the degree of compatibility, those lawyers can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
+(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about incompatibility are warranted and the historical strike data, without more information, is less helpful to proving a _Batson_ violation than it would otherwise would be.
+In any case, one virtue of this approach is that the choice of $\alpha$ explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.** [editme] We do not advocate the adjustment of alpha be used to arrive at some desired solution of either bias or no bias.
+In any case, adjustment of alpha, and therefore the weighting of historical evidence can be done iteratively along the 4 steps presented in fig 1, and decisions of true bias will not be based on its value alone.
 

### 2022-11-17 analysis/paper_draft3.Rmd
sentence edit

--- 2022-10-31 analysis/paper_draft3.Rmd
@@
 To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
-<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much our ability to detect bias was affected the size of the historical data and by how much the power prior discounted the weight of the historical data.
+<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.
 As introduced above, $b > 0$ represents a prosecutor's bias against a prospective juror within the cognizable class, while $b = 0$ denotes no bias.

### 2022-11-17 analysis/paper_draft3.Rmd

@@
 
-A _Batson_ challenge typically proceeds in a four-step sequence.
-The attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
-If challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
-If so, then at the next step, the striking attorney must proffer permissible reasons for their challenged strikes.
-Finally, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
+A _Batson_ challenge typically proceeds in a four-step sequence (Figure 1).
+First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
+Second, if challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
+Third, the striking attorney must proffer permissible reasons for their challenged strikes; and (4) the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
 
@@
 Here, we demonstrate this approach with simulation studies.
-Then, we present a software prototype that encodes the same approach with actual historical strike data to help attorneys and others detect bias in peremptory-challenge use in real time during jury selection.
+Then, we present a software prototype that encodes the same approach with actual historical strike data.
+With that data, the prototype shows how attorneys and other can this approach to help them decide to argue, and ultimately prove, illegal bias in peremptory-challenge use in real time during jury selection.
 
@@
 To show how lawyers might use this app in real cases in real time during selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
-
-To collect this historical strike data, one of us filed a request with the U.S.
-District Court of Connecticut based on 28 U.S.C.
-$\S$ 1868.
-Thereafter, we received copies of certain jury-selection records associated with twenty-nine criminal cases in that court during this period.
-These included strike sheets that indicated the identification number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
-Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.
-
-These records, however, often did not indicate the identity of the attorneys exercising the strikes.
-While the standard forms included a signature line for the attorney, many were left blank or filled with illegible signatures.
-Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the US Attorney's office) or the criminal defendant(s) on the date(s) of jury selection.
-Where only a single attorney represented a party during jury selection, it was easy to attribute the pattern of strikes to that attorney.
-Where multiple lawyers appeared for one side, we attributed to each of them that side's pattern of peremptory challenges in that case.
-In such cases, neither the jury-selection documents nor the docket sheets indicated any hierarchy among multiple lawyers or any other basis to attribute strikes to only one attorney among them.
-After generating a dataset based on these documents, we kept only strikes where a criminal defendant was represented by an attorney.
-Finally, to de-identify this dataset, we excluded defendant names and substituted fictitious names for the attorneys using the charlatan package [@charlatan].
+Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later de-identified them.
+For details on this dataset, see the Appendix.
 
@@
 For the prototype app, we set two cognizable-class options: race and gender.
-This was because the standard juror questionnaire did not ask about other possible cognizable classes.
+This was because the standard juror questionnaire in this court did not ask about other possible cognizable classes.
 The data also did not expressly indicate the race or gender of the criminal defendant(s).
@@
 Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
-In the Appendix, we present additional illustrations wherein the app is used to detect gender bias and where we assign less weight to the historical strike data.
+In the Appendix, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
 
@@
 Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
+[editme] The goal of these methods is not to _prove_ bias, but rather to lend support to an inferrence of bias and a subsequent _Batson_ challenge.
 The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
+[editme] sentence on why this is appealing
 
 Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
-The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether it is worth arguing a  _Batson_ challenge at all.
+The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether to make a  _Batson_ challenge at all.
 If argued, the attorney can introduce that posterior distribution as relevant evidence for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
@@
 In cases where lawyers contest the degree of compatibility, those lawyers can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about incompatibility are warranted and the historical strike data, without more, is less helpful to proving a _Batson_ violation than it would otherwise would be.
-In any case, one virtue of this approach is that the choice of $\alpha$ explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.
+(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about incompatibility are warranted.
+As a result, the historical strike data, without more information, helps less to prove a _Batson_ violation than it would otherwise would.
+
+By their choice of $\alpha$, the lawyer explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.
+This transparency thwarts lawyer efforts to strategically select $\alpha$ values to prove or disprove illegal strike bias.
+Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on either side can always directly and repeatedly assess how sensitive any bias inference is to the choice of $\alpha$.
+In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to using historical strike data has that ability.
 
@@
 
-Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Visualization, Writing - original draft.
+Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Data Visualization, Writing - original draft.
 
@@
 
-Here, we extend the discussion in section \@ref(prototype) of the paper by presenting additional illustrations of how a lawyer might use the prototype app.
+Here, we extend the discussion in section \@ref(prototype) of the paper by discussing the data collection and presenting additional illustrations of how a lawyer might use the prototype app.
+
+## Data Collection
+
+To collect this historical strike data, one of us filed a request with the U.S.
+District Court of Connecticut based on 28 U.S.C.
+$\S$ 1868.
+Thereafter, we received copies of certain jury-selection records associated with twenty-nine criminal cases in that court during this period.
+These included strike sheets that indicated the identification number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
+Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.
+
+These records, however, often did not indicate the identity of the attorneys exercising the strikes.
+While the standard forms included a signature line for the attorney, many were left blank or filled with illegible signatures.
+Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the US Attorney's office) or the criminal defendant(s) on the date(s) of jury selection.
+Where only a single attorney represented a party during jury selection, it was easy to attribute the pattern of strikes to that attorney.
+Where multiple lawyers appeared for one side, we attributed to each of them that side's pattern of peremptory challenges in that case.
+In such cases, neither the jury-selection documents nor the docket sheets indicated any hierarchy among multiple lawyers or any other basis to attribute strikes to only one attorney among them.
+After generating a dataset based on these documents, we kept only strikes where a criminal defendant was represented by an attorney.
+Finally, to de-identify this dataset, we excluded defendant names and substituted fictitious names for the attorneys using the charlatan package [@charlatan].
+
+## Additional Illustrations
 

### 2022-11-17 analysis/paper_draft3.Rmd
edits, updated caption to Figure 1; numbering of steps in Fig. 1

@@
 Second, if challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
-Third, the striking attorney must proffer permissible reasons for their challenged strikes; and (4) the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
+Third, the striking attorney must proffer permissible reasons for their challenged strikes.
+Fourth, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
 
@@
 
-```{r batsonworkflowfig, fig.cap="Inference from Historical Strike Data for  Batson Challenge", fig.align = 'center', out.width="100%", fig.pos = "h"}
+```{r batsonworkflowfig, fig.cap="Inference from Historical Strike Data at Key Steps of Batson Challenge", fig.align = 'center', out.width="100%", fig.pos = "h"}
 
@@
 By their choice of $\alpha$, the lawyer explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.
-This transparency thwarts lawyer efforts to strategically select $\alpha$ values to prove or disprove illegal strike bias.
-Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on either side can always directly and repeatedly assess how sensitive any bias inference is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to using historical strike data has that ability.
+This transparency matters, because it thwarts lawyer efforts to strategically select $\alpha$ values to prove or disprove illegal strike bias.
+Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias inference is to the choice of $\alpha$.
+In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to inferring bias based on historical strike data has that ability.
 

### 2022-11-17 figures/batson_workflow_fig2.tex
changed label in figure from "historical data" to "past strikes"
edits, updated caption to Figure 1; numbering of steps in Fig. 1

@@
 \begin{tikzpicture}
-  \node (A) at (0,0) [label=below:{argue \emph{Batson}},point,fill={blue}];	
-  \node (B1) at (4,0) [label=below:{prima facie case},
+  \node (A) at (0,0) [label=below:{1. argue \emph{Batson}},point,fill={blue}];	
+  \node (B1) at (4,0) [label=below:{2. prima facie case},
                        point,fill={blue}];	
-  \node (B2) at (8,0) [label=below:{proffer reasons},
+  \node (B2) at (8,0) [label=below:{3. proffer reasons},
                        point, fill={white}];
-  \node (B3) at (12,0) [label=below:{prove bias},point,fill={blue}];
+  \node (B3) at (12,0) [label=below:{4. prove bias},point,fill={blue}];
    \node (M) at (6,2) [state, drop shadow, fill={white}] {Model};
    \node (PB) at (6,1) [state] {bias estimate};
-    \node (H) at (3,3) [state] {historical data};
+    \node (H) at (3,3) [state] {past strikes};
     \node (C) at (9,3) [state] {current strikes};

### 2022-11-20 analysis/TAS_memo/memo_TAS_reviewers.Rmd
TAS reviewer memo draft: initial commit

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    latex_engine: pdflatex
    template: "svm-latex-memo.tex"
    number_sections: true
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes:
   - \linespread{1.05}

from: Authors
to: Editor, The American Statistican
subject: Reply to Reviewer Comments
date: "`r format(Sys.Date(), '%B %d, %Y')`"
memorandum: true
graphics: false
width: .3
---

This memorandum provides our replies to comments provided by the TAS associated editor and two reviewers.  It indicates how we have revised the paper accordingly, as well as our understanding of the issues raised by those comments.

# Comments of Associate Editor

`[T]he article doesn’t compellingly make the case that the methodology is more useful than simply tracking those who have been reprimanded in the past or that it’s ethical/reasonable to actually employ in the courtroom. Specifically, the way that prior information is encoded, the logic behind a detection appears to be circular in that if someone has been reprimanded in the past they should be reprimanded in the future (in some cases for no reason). This issue hasn’t been discussed to a level that makes the utility of such a model clear.`

We have three replies to this concern.

First, as page X, lines Y of the paper indicate, the law of _Batson_ and parallel State law has consistently permitted judges to infer illegal strike bias based in part on an attorney's use of strikes in past cases. In reviewing the kinds of evidence that can prove a _Batson_ violation by a prosecutor, the U.S. Supreme Court included, as examples of types of evidence, the "relevant history of the State’s peremptory strikes in past cases." Flowers v. Mississippi, 139 S. Ct. 2228, pin (2019).

More recent State law variants on _Batson_ are also explicit about this. For example, Washington State law provides that in deciding whether "an objective observer could view race or ethnicity as a factor in the use of the peremptory challenge," the circumstances the court "should" consider include "whether the party has used peremptory challenges disproportionately against a given race or ethnicity, in the present case _or in past cases_." Wash. General Rule 37(g)(v) (emphasis added). California law similarly identifies, as a circumstance a court may consider, whether the striking "has used peremptory challenges disproportionately against a given race, ethnicity, gender, gender identity, sexual orientation, national origin, or religious affiliation, or perceived membership in any of those groups, in the present case _or in past cases_." Calif. Code of Civil Procedure 231.7(d)(3)(G) (emphasis added).

We hope this removes the worry that it is not "ethical/reasonable" to infer present strike bias from strike behavior in past cases.

Second, the phrase "should be reprimanded in the future (in some cases for no reason)" seems to be premised either on worries that our approach (1) requires trial judges to infer present illegal strike bias (i.e., find _Batson_ violations ) _solely_ based on strikes in past cases; and (2) will cause judges to be more likely to find _Batson_ violations when in fact no strike bias exists. Neither is the case, and we have revised the paper to make this more clear.

As an initial matter, the word "reprimanded" refers to a court finding of a _Batson_ violation. Our approach, however, relies on an attorney's use of strikes of past cases, regardless of whether those strikes were ever the subject of a _Batson_ challenge. This is better than just "tracking those who have been reprimanded in the past" for _Batson_ violations, because the probability of a court that finds a _Batson_ violation depends on an attorney having raised a _Batson_ challenge in the first place. Absent an attorney arguing _Batson_, a court will not inquire into whether illegal strike bias existed. If illegal strike bias exists but the attorney erroneously perceives no credible basis for it, the attorney will tend not to argue _Batson_. In cases where there no illegal strike bias but the attorney erroneously perceives some credible basis for it, the attorney will tend to argue _Batson_ but may not persuade the trial judge that illegal strike bias was more likely than not. By relying on past strike data, not whether anyone argued _Batson_ in those past cases, we avoid any selection bias associated with relying only on past court findings of _Batson_ violations.

More importantly, the law does not permit a trial judge to find illegal strike bias based _only_ on an attorney's strikes in past cases while ignoring evidence tending to prove no illegal bias in the present case. As the paper now explains on page X, line Y, _Batson_ procedure permits any type of admissible evidence that is relevant to strike bias (for or against) -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.

For this reason, our approach alone is not intended as an all-purpose detector of _Batson_ violations, because our approach cannot incorporate the weight of all the other kinds of evidence that could be submitted to prove illegal strike bias in the present case. For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons. If the attorney fails to produce such reasons, the trial judge must find a _Batson_ violation. If the attorney offers such reasons, the trial judge must assess evidence as to the credibility of those reasons. Our approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) relevant to illegal strike bias.

Rather, as Figure 1 now depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on an attorney's strikes in the present case and past cases, i.e., _before_ accounting for other evidence relevant to illegal strike bias in the present case (the _Batson_ violation itself). For that task, our approach assigns at most equal weight to past strikes and their strikes in the present case (i.e., when $\alpha = 1$). And if there is incompatibility betwen the historical strike data and the current strike data ($b_{hist} \neq b_{curr}$) for reasons identified on page X, lines Y, that inference of bias will be sensitive to the choice of $\alpha$, and therefore that inference of bias can and should be taken as weaker for that reason.

In turn, attorneys can use that information, _combined with_ other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find that illegal strike bias is more likely than not, given all the evidence, i.e., find a _Batson_ violation.

### Additional Comments

`1. Page 4, line 5: The authors refer to gender but then provide sex options (male/female).`

[comment]

`2. Page 4, line 36: Use \left(\right) to ensure appropriately sized parentheses where necessary.`

Thank you for this suggestion. The paper has been revised accordingly.

`3. Page 5, equation 6: The left-hand side-of the equation is not correct`

Thank you for spotting this.  The equation has been revised accordingly.

`4. Page 6, line 43: The idea behind $\alpha$ is concerning to me. The idea here is that if there is previous problematic strike data, and we choose large $\alpha$, then we’ll detect problematic strikes at higher rates. This type of modeling is helpful when historical data are informative to current random behavior, but in this case it goes against the usual ‘innocent until proven guilty,’ and, in fact, bias can be detected at rates close to 100% when there is no bias currently! That is, in other judicial contexts (and likely this one), this could be highly problematic. Simultaneously, large choices of $\alpha$ can lead to no detection of bias based on good behavior or lack of detection earlier. Note that detection rates are as low as 40% when current bias is set to 3! To me, noting I’m not a law expert, it seems a non-informative prior would be best here. While prior behavior could be used to demonstrate a pattern (guilt), I’m not convinced it’s appropriate for detection. If that is not the case, it should be thoroughly explained to the reader. Perhaps the citation reference by Reviewer 2 can help the authors – It’s possible that only considering small values of $\alpha$ would be reasonable. The simulation results for alpha = 0.1 are less objectionable.`

As explained above, the law of _Batson_ permits the inference of present strike bias based on past strikes in previous trials. The possible incompatability of $b_hist$ and $b_curr$ is thus already an issue when using past strike data to infer bias. Our contribution is to bring that issue to foreground and to advise assessing how sensitive the inference of bias is to the choice of $\alpha$ values by choosing multiple $\alpha$ values.

`5. Page 6, line 50: The description of bhist is not clear; it is defined twice with different values.`

The paper has been revised to clarify this description.

`6. Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?`

We intend to indicate that the larger the credible interval, the greater the chance that zero falls within it.  Accordingly, the choice of the credible interval is a decision about how conservative one wishes to be about inferring bias (defined here as when $b = 0$ falls outside the chosen credible interval.

Just focus on one in the paper, rest in the appendix. It is a decision about how conservative you want to be. We have removed the discussion of the different CIs and foculs on the 95% CI for the manuscript. We note that this does not necessarily fixed the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), and we are interested in "bias detection", i.e. whether the CI includes 0 or not. We also, present sensitivity analysis for the selection of the nominal HPD level. In practice, this needs to be chosen based on how conservative the lawyer bringing up a _Batson_ challenge behaves.

`7. Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior. One would be curious whether the approach with no prior works for detecting bias (e.g., simulations results without prior data). I think I would find this type of model more compelling.`

[REPLY]

`8. Page 18, line 45: If there is benefit to the user for finding bias, the ability to select $\alpha$ is akin to p-hacking.`

We respectfully disagree with the analogy to "p-hacking", for three reasons.  First, the phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results. In contrast, in our approach, the statistical model necessarily precedes any analysis, because it has to qualify as a valid model of the strike _procedure_ of the court of interest, which can and must be known before any data is collected or analyzed.

Second, in the adversarial setting of the courtroom, our approach actually thwarts the strategic selection of $\alpha$ values, because it makes the choice of $\alpha$ transparent.  This is why we stress in the paper (p. X, lines Y) that, where there is suspected incompatibility between $b_{hist}$ and $b_{curr}$, the proper approach is try multiple values of $\alpha$ and determine how sensitive the inference of bias is to the chosen $\alpha$ value.  Because the choice of how much to weigh the historical strike data is transparent, every lawyer knows that any strategic selection of $\alpha$ will be met with scrutiny from the lawyers on the other side. Absent this transparency, lawyers already still argue for how much past strike behavior ought to weigh when inferring or dispute illegal strike bias, except that they do so implicitly and with the opacity of words.

`9. Something unmentioned that should be addressed is confounders. While technological and statistical tools can be helpful to the judicial system, and even decrease instances of discrimination and curb disparities, there are many unintended consequences of such systems. For example, race and class are intertwined in the United States and there is no discussion (at least) about incorporating covariates (e.g., age, education, class, etc.)`

[INSERT]

# Reviewer 1

>1. In Section 2.3, the paragraph in line 47 on p. 6 . . . I am confused about which set of values was used in the simulation for bhist.

Thank you for catching this.  We have revised the paper (p. X, lines Y) to clarify the values used in the the simulation for $b_{hist}$.

>2. About the results presented in Figure 1 of Section 2.3, if we ignore the assumption of compatibility between historical and current data, in the cases of $\alpha$ = 0:5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0? If so, the Type I error seems to be very high in some cases. Similarly, the power (bias detection rate) seems low at the top left corner in each panel where bias is low in historical data but high in current data.

[INSERT]

>3. Regarding the Shiny App in Section 2.5, it would be helpful to include some details about the Metropolis-Hastings scheme. For instance, is a normal distribution used as the proposal distribution for b? What is the length of the Markov chain? Is any thinning performed on the posterior draws? Is any convergence criterion (e.g., the Gelman-Rubin statistic) checked? What is the computation time?

[INSERT]

>4. In Section 3, the authors state that "the anticipated degree of incompatibility should inuence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error. However, based on results in Figures 1 to 3, if this assumption is wrong, choosing a large $\alpha$ leads to either low power (bias detection rate) or large Type I error, depending on the bias in historical data. In the absence of such knowledge of compatibility, how does one choose the value of $\alpha$? In addition, to get around this issue, is it reasonable to treat $\alpha$ itself random and introduce a hyper prior?

[INSERT]

Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$, because the proper weight to place on historical data depends on whether, for that historical data, there is any missing data is missing not at random and whether an inference of bias based on that historical strike data is complicated by plausible confounding or mediating trial-level variables (e.g., defendant race, criminal charge severity). The plausible reasons for non-random missingness and confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things. [MORE].

>5. Some of the Figure and table captions (e.g. Figures 1-5 and Table 1) can benefit from being more descriptive.

[INSERT]

> 6. A couple of notation issues: . . . .

Thank you for pointing out these issues. We have revised Equations (3) and (6) accordingly.

# Reviewer 2

Reviewer 2 questions whether the data required for such a model are readily available for broad application, in addition to pointing out several important questions about the details of the paper.

Data for past trials is not readily available. But, since the 80s, any person can enter a courtroom and observe the jury selection process, so this data could be collected for present and future cases.

* [add paragraph to address this]

Further, in ‘Other Comments’ 3-4, they describe a rather important realities about the application that are not accounted for in the model. While the problem is certainly interesting, the simulations demonstrate some highly problematic results.
Knowledgeable reviewers and I still have several large unanswered questions about how the methodology can be enacted appropriately. I provide my comments below.

3. Page 5, equation 5: The left-hand side-of the equation is not correct
* [EB: fix in formula] - This is now fixed to reflect that it is a posterior distribution. We note that all formulas have now been updated and checked.

We now specify the vector notation for $\delta_0$ and $\delta_1$.

Mention that in the app, it is easy to check the influence of alpha by re-running the model. Make a clearer case that level of alpha should be selected in cases of high incompatibility. Law allows you to rely on the prosecutor's past behavior. The ability to modify alpha, allows you to decide whether inference based on past behavior is warranted.
[TEM & EB]

6. Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you  saying that larger credible interval percentages lead to less accurate bias detection?

7. Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior. One would be curious whether the approach with no prior works for detecting bias (e.g., simulations results without prior data). I think I would find this type of model more compelling.

The real data analysis, as the default option, does not borrow historical data (i.e. $\alpha=0$). In this case the default prior of $N(0,2)$ is used for the bias parameter (i.e. b). This is reflected when no attorney is selected from the pull-down menu. If an attorney is selected from the pull-down menu, then prior information is used for that attorney. The paper has been updated to reflect this point in the same section.

Review 1
Journal: The American Statistician
Manuscript ID: TAS-22-165
Title: Bayesian detection of bias in peremptory challenges using historical strike data
The authors presented a Bayesian model in detecting potential biases against minority groups
in peremptory challenges. The use of the power prior to incorporate historical data is in-
teresting, and the methodology overall seems reasonable. The proposed R Shiny App may
also be helpful to practitioners. My main concerns are about the simulation results and the
choice of the hyperparameter 
in the power prior.
Major comments:

1. In Section 2.3, the paragraph in line 47 on p. 6 ends with bhist = f3;2; : : : ; 3g,but in the very next paragraph, bhist is set to f0; 0:5; : : : ; 3g. I don't see bhist =f3;2; : : : ; 2; 3g mentioned anywhere else, including in the results shown in Figures1 to 3. I am confused about which set of values was used in the simulation for bhist.

* Fix bhist typo[DONE]

* Repeat metal detector response here.

* Ask Xiaomeng about this in the app [TEM:  reach out]

* Repeat metal detector response here. Could mention this as something that we considered, and mention that we would like find a way to measure compatibility. This is also tool intended to be used by lawyers in real-time.

Main Comments:
1. The method is first illustrated in a simulation study, with a range of values of $\alpha$.
This is fine for an audience of statisticians and attorneys specializing in jury
discrimination but is somewhat overwhelming for a general reader. It would be
easier for the reader if one data set, perhaps from the actual data from Connecticut
was used and the reader could see how the data from each additional previous trial
impacts the posterior. Perhaps only a few values of $\alpha$, rather than 11 and two intervales
of the bias parameter b, could be used.

* Start with CT data as example. Take screenshots from the app part.

> 2. From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones. How often is this data currently available? If it isn’t readily available now, how likely are states and locales to implement a system of systematically reporting and preserving the required data (especially in the areas where one expects minorities to encounter bias in jury membership)?

* Repeat comment about accessibility

`3. In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias. If one has been called for jury service in a city with a large minority population and been in a jury room, this is highly implausible. The reviewer once was sitting next to a minority woman and we saw---A minority leave the jury room, then a white, then another minority, then another white etc. At the end both of us noticed the “pattern”. Because one is analyzing a changing pool of possible challenges each time a peremptory challenge is made, when the defense attorney (assuming a minority defendant—the usual situation where Batson issues arise) removes a white, the minority proportion of the pool of possible individuals to be removed increases. This could make it more difficult to detect a biased prosecutor. On the other hand, if the method detects prosecutorial bias when the defense challenges appear biased too, this should strengthen the evidence of prosecutorial bias. This issue deserves some quantitative investigation, so judges can feel confident when using the results.`

* [TEM&EB: Check this to see if adding bias to the defense alters bias detection for the prosecutors. ][see table below]

`4. The paper and software focus on peremptory challenges made by the same lawyer in previous cases. As noted in the Miller-El cases, the training manual used by prosecutors recommended the order in which groups, e.g. African-Americans, women etc. should be challenged. Therefore, data from prosecutors from the same DA’s office or defense firm in similar cases (drug dealing, robbery etc.), should be quite informative. Should one simply add them to the list of prior cases, or make a down-weighting adjustment in the Bayesian analysis?`

* Could use same idea for individual lawyers for offices, by aggregating by firms/office/etc, and remove duplicate strikes/trials.

[SP: add pararaph/sentence about this]

`5. The paper uses 90% and 95% credible intervals on p. 7, 95% in Tables 2 and 3 and 80% on p. 18 and Fig. 7. From the various contexts, the reviewer got the impression that the author(s) were recommending that 0 not in the 95% credible interval should help the defendant establish their Batson claim, while 0 in an 80% credible interval indicates no bias and would help the prosecutor (in a typical Batson case) rebut a Batson challenge. Is this correct? In any case, the author(s) should indicate which credible intervals should be used at various stages of the proceedings. If possible, the standards for deciding a Batson claim used by judges in a jurisdiction should be the same. This is especially true in the present context where the prior data considered in cases in the same jurisdiction is likely to have a substantial overlap.`

* This is not correct. Clarify by focussing on single interval.

`Gastwirth (2005) in Law, Prob. Risk, p. 179-185, . . . had proposed that
approach when potential jurors were removed one at a time in an alternating
sequence. Checking that reference, the reviewer saw that at the end of the 2005
paper, the use of data from prior cases would be helpful and recommended that
procedures, such as the one proposed here. This connection should be noted. Also,
if one had data from an actual case, one could show the additional insight one
obtains by incorporating data from previous cases.`

Thank you for this reference.  The paper now cites to the model in Gastwirth (2005). We also agree about the "additional insight" from "data from an actual case."  The discussion about the prototype software application expressly relies on actual strike data from a sample of trials from the federal district court of Connecticut. [MORE].

`The choice of $\alpha$ should be discussed and recommendations made. Otherwise, both parties can explore which value is most helpful. Also, how can one ensure that the prior data doesn’t overwhelm the data from the current case, without choosing a ridiculously low value? The authors say something about this on page 6 but more guidance to potential users and courts would be helpful.`

* Repeat section about alpha/metal detector argument.

`The scenario depicted in Table 8 and the related text indicates that both the prosecutor and defense removed minority jurors –both 95% credible intervals do not contain 0. This seems unrealistic. More common would be the prosecutor removed more minorities so their 95% credible interval would not contain 0 while the credible interval for the defense removing non-minority potential jurors would be shifted but still might contain 0 (Of course, in some cases when minorities form a substantial portion, e.g. 40% or so of the venire, the data may indicate both sides were probably biased.`

Table 8 = hypothetical data. This is just one scenario. Nothing about the tool makes any assumptions about what defense and prosecutors do/tend to do when it comes to cognizable/non-cognizable jurors. The burden of proof lies with the person bringing the challenge

> At the bottom of p. 17, you don’t really need names for the attorneys, A and B would suffice.

Fair. Although, these are fake names.

Update online appendix
* Move figures (80% and 90% CI)
   * 80 and 90 CI
   * Fig 4, 6, 8, 9,
* Move app screenshots and majority of text about app
* Add sensitivity for defense bias (new)

Add real data example to introduction - make more explicit

Table of bias detection for the prosecutor, when the defense shows bias. N historical = 3, nstrike =15.

	1.0

### 2022-12-01 analysis/TAS_memo/memo_TAS_reviewers.Rmd
Update memo_TAS_reviewers.Rmd

@@
 Neither is the case, and we have revised the paper to make this more clear.
+Places where these clarifications have been made are highlighted in the paragraphs that follow.
 
@@
 
-Rather, as Figure 1 now depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on an attorney's strikes in the present case and past cases, i.e., _before_ accounting for other evidence relevant to illegal strike bias in the present case (the _Batson_ violation itself).
+Rather, as the new Figure 1 depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on an attorney's strikes in the present case and past cases, i.e., _before_ accounting for other evidence relevant to illegal strike bias in the present case (the _Batson_ violation itself).
 For that task, our approach assigns at most equal weight to past strikes and their strikes in the present case (i.e., when $\alpha = 1$).
@@
 In turn, attorneys can use that information, _combined with_ other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find that illegal strike bias is more likely than not, given all the evidence, i.e., find a _Batson_ violation.
+
+We hope that these modifications clarify the placement of our model in the steps from a sequence of peremptory challenges to a _Batson_ challenge actually being brought, and upheld by the trial judge.
+We we have also more clearly outlined the the precedence for the use of historical strike data in the _Batson_ challenge process.
 
@@
 Our contribution is to bring that issue to foreground and to advise assessing how sensitive the inference of bias is to the choice of $\alpha$ values by choosing multiple $\alpha$ values.
+[?
+We have also included the citation from reviewer 2]
 
@@
 
-Just focus on one in the paper, rest in the appendix.
+For the sake of clarity, we now focus on one in the paper, and have placed the results based on alternative CIs in the appendix.
 It is a decision about how conservative you want to be.
@@
 
-[REPLY]
+This comment is not clear to us.
+The real data analyses are based on real _historical_ trial data from the stae of Connecticut (see lines x to y).
+Users of the app have the option to run the models excluding any historical trial data (as suggested by the comment) by not selecting the name of a prosecutor or defense attorney.
+In this case the normal(0,2) prior is used for the bias parameter.
+Similarly, the app gives the user the option of altering the weight the historical strike data (if selected) by adjusting alpha (power prior parameter).
+Thus the user can: 1) choose not to use historical data, 2) use historical data with equal weight as the current trial (alpha = 1), or 3) down-weight the historical trial data.
 
@@
 
-[INSERT]
+[INSERT] We agree.
+Although, unfortunately this additional information (age, etc) is not routinely collected by the courts, and therefore is difficult to consider in the context of our model.
+We have added a sentence (line x) to the discussion that highlights this point by calling for the need to reassess how, and more importantly what, information is collected on prospective jurors, so that models like ours can be used to my the process more fair and equittable.
 
@@
 
-[INSERT]
+[INSERT] The reviewer is correct on 2 counts: 1 the bias detection rate can be as high as 100% when true bias is zero, and 2 power to detect bias is low when there is low bias in historical trials.
+1.
+these situations occur in the presense of compatibility whih we highlight in the manuscript as arising for various reasons (see lines x), and 2.
+The lack of power is low due the limited number of strikes, which is why the use of compatible historical information can increase this power.
+This is the purpose of power prior: adjust the weight of historical data (and therefore type I error) by adjusting alpha.
+In reality, users of the model will not know the values of bias in the current and historical data, and so will need to rely on ...
+to decide what value of alpha to choose, and as we have said above, ultimately whether or not to bring a Batson challenge where they will need to discolse their use of alpha.
 
@@
 
-[INSERT]
+[INSERT] Will get Eric to check this (still no response from Xiaomeng)
 
 >4.
-In Section 3, the authors state that "the anticipated degree of incompatibility should inuence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error.
+In Section 3, the authors state that "the anticipated degree of incompatibility should influence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error.
 However, based on results in Figures 1 to 3, if this assumption is wrong, choosing a large $\alpha$ leads to either low power (bias detection rate) or large Type I error, depending on the bias in historical data.

### 2022-12-01 analysis/TAS_memo/memo_TAS_reviewers.Rmd
Update memo_TAS_reviewers.Rmd

@@
 This comment is not clear to us.
-The real data analyses are based on real _historical_ trial data from the stae of Connecticut (see lines x to y).
+The real data analyses are based on real _historical_ trial data from the state of Connecticut (see lines x to y).
 Users of the app have the option to run the models excluding any historical trial data (as suggested by the comment) by not selecting the name of a prosecutor or defense attorney.
@@
 
-[INSERT] We agree.
-Although, unfortunately this additional information (age, etc) is not routinely collected by the courts, and therefore is difficult to consider in the context of our model.
-We have added a sentence (line x) to the discussion that highlights this point by calling for the need to reassess how, and more importantly what, information is collected on prospective jurors, so that models like ours can be used to my the process more fair and equittable.
+[INSERT] Batson challenge doesn't care if you strike a cognizable class because you assume they have particular confounders (e.g., low income, etc).
+[editme] We do explicitly highlight that trial-level characteristics and impact compatibility between current and historical trials.
+We do advocate for reassessment of the way (and what) data are collected from individual trials.
 
@@
 
-[INSERT]
+[INSERT] Choose it based on beliefs on causes of incompatibility and evaluate bias detections sensitivity to its value.
 

### 2022-12-12 analysis/paper_draft3.Rmd
edit introduction to add power prior justification

@@
 
-In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, establishing a prima facie case, and proving that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
+In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, establishing a prima facie case, and proving that bias in the particu-lar case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
+One challenge for incorporating historical data is the possible incompatibility between current and historical data.
+This incompatibility could arise due to missing data, or if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).[may need fleshing out]
 
@@
 ```
-
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data.
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows us to explicitly incorporate the degree of incompatibility between current and historical trials.
 In this approach, we use two steps.
@@
 We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
-In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posterior distribution developed with the historical information and the power prior.
+In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posteri-or distribution developed with the historical information and the power prior.
 
-Here, we demonstrate this approach with simulation studies.
+The use of the power prior in this context allows the explicit control of the degree of incompatibility between the current and historical trial data.
+Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected when incompatibility is expected to be high, and higher values (closer to 1) of the power prior parameter selected when incompatibility is low.
+For now, the choice of alpha is placed in the hands of the user, who will need to be able to defend their choice when presenting a _Batson_ challenge before the court.
+
+Here, we demonstrate this approach with simulation studies where we vary the degree of incompatibility between current and historical trial data, and test our ability to detect bias using different values of alpha.
 Then, we present a software prototype that encodes the same approach with actual historical strike data.

### 2022-12-13 analysis/paper_draft3.Rmd
revisions to intro re power prior

@@
 
-In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, establishing a prima facie case, and proving that bias in the particu-lar case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
-One challenge for incorporating historical data is the possible incompatibility between current and historical data.
-This incompatibility could arise due to missing data, or if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).[may need fleshing out]
+In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, proving a prima facie case, and ultimately persuading the trial judge that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
+One challenge for incorporating historical data is possible incompatibility between current and historical strike data for a given attorney that may arise because of missing data or because attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).[may need fleshing out]
 
@@
 
-The use of the power prior in this context allows the explicit control of the degree of incompatibility between the current and historical trial data.
+The use of the power prior in this context allows lawyers to explicitly control over the assumed degree of incompatibility between the current and historical trial data.
 Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected when incompatibility is expected to be high, and higher values (closer to 1) of the power prior parameter selected when incompatibility is low.
-For now, the choice of alpha is placed in the hands of the user, who will need to be able to defend their choice when presenting a _Batson_ challenge before the court.
+Accordingly, the lawyer bringing the _Batson_ challenge chooses the value of $\alpha$ and must be able to explain and defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current strike data and historical strike data available in the particular case.
 

### 2022-12-13 analysis/TAS_memo/memo_TAS_reviewers.Rmd
second round of revisions to TAS memo

--- 2022-11-20 analysis/TAS_memo/memo_TAS_reviewers.Rmd
@@
 
-This memorandum provides our replies to comments provided by the TAS associated editor and two reviewers.
+This memorandum replies to comments provided by the TAS associate editor and two reviewers.
 It indicates how we have revised the paper accordingly, as well as our understanding of the issues raised by those comments.
 
-# Comments of Associate Editor
+# Comments of TAS Associate Editor
 
@@
 2228, pin (2019).
-
 More recent State law variants on _Batson_ are also explicit about this.
@@
 
-Second, the phrase "should be reprimanded in the future (in some cases for no reason)" seems to be premised either on worries that our approach (1) requires trial judges to infer present illegal strike bias (i.e., find _Batson_ violations ) _solely_ based on strikes in past cases; and (2) will cause judges to be more likely to find _Batson_ violations when in fact no strike bias exists.
-Neither is the case, and we have revised the paper to make this more clear.
+Second, as the phrase "should be reprimanded in the future (in some cases for no reason)" implies, the associate editor seems worried that our approach will cause judges to infer that illegal strike bias is more likely than not (i.e., find _Batson_ violations ) when in fact no strike bias exists.
 
-As an initial matter, the word "reprimanded" refers to a court finding of a _Batson_ violation.
+This is not the case, and we have revised the paper to make this more clear.
+
+First, we assume the word "reprimanded" refers to a court finding of a _Batson_ violation.
 Our approach, however, relies on an attorney's use of strikes of past cases, regardless of whether those strikes were ever the subject of a _Batson_ challenge.
 This is better than just "tracking those who have been reprimanded in the past" for _Batson_ violations, because the probability of a court that finds a _Batson_ violation depends on an attorney having raised a _Batson_ challenge in the first place.
-Absent an attorney arguing _Batson_, a court will not inquire into whether illegal strike bias existed.
-If illegal strike bias exists but the attorney erroneously perceives no credible basis for it, the attorney will tend not to argue _Batson_.
+(Absent an attorney arguing _Batson_, a court will not inquire into whether illegal strike bias existed.) If illegal strike bias exists but the attorney erroneously perceives no credible basis for it, the attorney will tend not to argue _Batson_.
 In cases where there no illegal strike bias but the attorney erroneously perceives some credible basis for it, the attorney will tend to argue _Batson_ but may not persuade the trial judge that illegal strike bias was more likely than not.
@@
 
-More importantly, the law does not permit a trial judge to find illegal strike bias based _only_ on an attorney's strikes in past cases while ignoring evidence tending to prove no illegal bias in the present case.
-As the paper now explains on page X, line Y, _Batson_ procedure permits any type of admissible evidence that is relevant to strike bias (for or against) -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
-
-For this reason, our approach alone is not intended as an all-purpose detector of _Batson_ violations, because our approach cannot incorporate the weight of all the other kinds of evidence that could be submitted to prove illegal strike bias in the present case.
+Second, _Batson_ procedure permits any type of admissible evidence of strike bias (for or against) -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+For this reason, our approach alone is not an all-purpose detector of _Batson_ violations, because our approach cannot incorporate the weight of all the other kinds of evidence that could be submitted to prove illegal strike bias in the present case.
 For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons.
@@
 If the attorney offers such reasons, the trial judge must assess evidence as to the credibility of those reasons.
-Our approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) relevant to illegal strike bias.
 
-Rather, as Figure 1 now depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on an attorney's strikes in the present case and past cases, i.e., _before_ accounting for other evidence relevant to illegal strike bias in the present case (the _Batson_ violation itself).
+Our approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) of illegal strike bias.
+Rather, as the new Figure 1 depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias.
 For that task, our approach assigns at most equal weight to past strikes and their strikes in the present case (i.e., when $\alpha = 1$).
-And if there is incompatibility betwen the historical strike data and the current strike data ($b_{hist} \neq b_{curr}$) for reasons identified on page X, lines Y, that inference of bias will be sensitive to the choice of $\alpha$, and therefore that inference of bias can and should be taken as weaker for that reason.
 
+The possible incompatability of $b_hist$ and $b_curr$ is already and always an issue when using past strike data to infer bias.
+This is a reality faced by _any_ statistical technique for inferring strike bias using past strike data.
+Because the law of _Batson_ expressly permits inferring current strike bias based on past strikes in previous trials, the lawyers in _Batson_ challenges involving past strike data are already and always arguing about how much weight to assign to that past strike data.
+
+Our approach simply makes the weighting of historical strike data -- usually expressed with words -- more explicit and transparent by requiring a choice of a value for alpha.
+The more sensitive the inference of bias (zero falls outside the credible interval) is to the choice of $\alpha$, the more one should worry about incompatibility.
+In turn, the appropriateness of small vs.
+large values of $\alpha$ vary with the suspected degree of incompatability.
 In turn, attorneys can use that information, _combined with_ other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find that illegal strike bias is more likely than not, given all the evidence, i.e., find a _Batson_ violation.
+
+We hope that these modifications clarify the placement of our model in the steps from a sequence of peremptory challenges to a _Batson_ challenge actually being brought, and upheld by the trial judge.
 
@@
 
-As explained above, the law of _Batson_ permits the inference of present strike bias based on past strikes in previous trials.
-The possible incompatability of $b_hist$ and $b_curr$ is thus already an issue when using past strike data to infer bias.
-Our contribution is to bring that issue to foreground and to advise assessing how sensitive the inference of bias is to the choice of $\alpha$ values by choosing multiple $\alpha$ values.
+Finally, the reviewer's use of the word "guilty" requires two clarifying notes.
+First, _Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more stringent "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the criminal offense charged.
+Second, as explained above, bias "detection" in the paper means only whether the credible interval for the bias parameter ($b_{curr}) excludes zero.
+Because _Batson_ procedure requires weighing any and all evidence of strike bias, not just past strike data, our approach cannot alone identify _Batson_ violations in cases involving other relevant evidence bias besides past strike data.
 
@@
 
-We intend to indicate that the larger the credible interval, the greater the chance that zero falls within it.
-Accordingly, the choice of the credible interval is a decision about how conservative one wishes to be about inferring bias (defined here as when $b = 0$ falls outside the chosen credible interval.
-
-Just focus on one in the paper, rest in the appendix.
-It is a decision about how conservative you want to be.
-We have removed the discussion of the different CIs and foculs on the 95% CI for the manuscript.
-We note that this does not necessarily fixed the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), and we are interested in "bias detection", i.e.
-whether the CI includes 0 or not.
-We also, present sensitivity analysis for the selection of the nominal HPD level.
-In practice, this needs to be chosen based on how conservative the lawyer bringing up a _Batson_ challenge behaves.
+We have revised this discussion to remove such confusion.
+The choice of the credible interval is a decision about how conservative one wishes to be about inferring bias from past strike data alone.
+We define bias "detection" narrowly to mean when $b = 0$ falls outside the chosen credible interval.
+Thus, the larger the credible interval, the greater the chance that zero falls within it.
+For the sake of clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on other credible intervals in the Appendix.
+This does not necessarily fix the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), and we are interested in "bias detection", i.e.
+whether the credible interval includes 0 or not.
+We also present sensitivity analysis for the selection of the nominal HPD level.
 
@@
 
-[REPLY]
+This comment is not clear to us.
+The real data analyses are based on real _historical_ trial data from the federal district court in Connecticut (see lines x to y).
+Users of the app have the option to run the model excluding any historical trial data (as suggested by the comment) by not selecting the name of a prosecutor or defense attorney.
+In this case, the normal(0,2) prior is used for the bias parameter.
+Similarly, the app gives the user the option of altering the weight the historical strike data (if selected) by adjusting alpha (power prior parameter).
+Thus the user can: 1) choose not to use historical data, 2) use historical data with equal weight as the current trial (alpha = 1), or 3) down-weight the historical trial data.
 
@@
 
-We respectfully disagree with the analogy to "p-hacking", for three reasons.
-First, the phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results.
+We respectfully disagree with the analogy to "p-hacking", for two reasons.
+First, the phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify as "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results.
 In contrast, in our approach, the statistical model necessarily precedes any analysis, because it has to qualify as a valid model of the strike _procedure_ of the court of interest, which can and must be known before any data is collected or analyzed.
@@
 
-[INSERT]
+[INSERT] Batson challenge doesn't care if you strike a cognizable class because you assume they have particular confounders (e.g., low income, etc).
+[editme] We do explicitly highlight that trial-level characteristics and impact compatibility between current and historical trials.
+We do advocate for reassessment of the way (and what) data are collected from individual trials.
 
@@
 
-Thank you for catching this.  We have revised the paper (p. X, lines Y) to clarify the values used in the the simulation for $b_{hist}$.
+Thank you for catching this.  We have revised the paper (p. X, lines Y) to clarify the values used in the simulation for $b_{hist}$.
 
@@
 
-[INSERT]
+We agree that the bias "detection" rate (whether the credible interval excludes zero) can be as high as 100% when true bias is zero.
+We also agree that the power to detect bias is low when there is low bias in historical trials.
+These situations can occur because of high incompatibility.
+Similarly, the low power situations occurs due the limited number of strikes, which is why _compatible_ historical strike data can increase this power.
+
+This is an advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting alpha.
+In reality, because $b_{curr}$ and $b_{hist}$ are not directly observable, one will need to rely on background knowledge or assumptions about the degree of incompatibility to decide which value of alpha to choose.
+If they bring a _Batson_ challenge, they must also disclose and justify their choice of alpha and the assumptions on which that choice is based.
 
@@
 
-[INSERT]
-
-Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$, because the proper weight to place on historical data depends on whether, for that historical data, there is any missing data is missing not at random and whether an inference of bias based on that historical strike data is complicated by plausible confounding or mediating trial-level variables (e.g., defendant race, criminal charge severity).
+As indicated in the response to comment 2 above, one chooses a value of $alpha$ based on background beliefs on plausible degree of incompatibility, and then evaluate how sensitive bias detection (whether zero falls outside the credible interval) is to the choice of alpha value.
+Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$, because the proper weight to place on historical data depends on whether, for that historical data, there is any missing data that is missing not at random and whether an inference of bias based on that historical strike data is complicated by plausible confounding or mediating trial-level variables (e.g., defendant race, criminal charge severity).
 The plausible reasons for non-random missingness and confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
-[MORE].
 
@@
 
-Reviewer 2 questions whether the data required for such a model are readily available for broad application, in addition to pointing out several important questions about the details of the paper.
+>1.
+The method is first illustrated in a simulation study, with a range of values of $\alpha$.
+This is fine for an audience of statisticians and attorneys specializing in jury discrimination but is somewhat overwhelming for a general reader.
+It would be easier for the reader if one data set, perhaps from the actual data from Connecticut was used and the reader could see how the data from each additional previous trial impacts the posterior.
+Perhaps only a few values of $\alpha$, rather than 11 and two val[u]es of the bias parameter b, could be used.
+
+* Start with CT data as example. Take screenshots from the app part.
+
+> 2.
+From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones.
+How often is this data currently available?
+If it isn’t readily available now, how likely are states and locales to implement a system of systematically reporting and preserving the required data (especially in the areas where one expects minorities to encounter bias in jury membership)?
+
+In the Discussion, we explained that data on consecutive strike patterns is hard to collect, simply because of the challenges of collecting any historical strike data, regardless of strike procedure.
+We have revised that Discussion to add that, since the 1980s, any person has a First Amendment right to attend jury selection in criminal court proceedings, and criminal defendants have a Sixth Amendment right to public court proceedings during jury selection.
+These legal rights, in turn, enable data collection by court observers, regardless of State and local recordkeeping practices.
+
+Finally, the current availability of the data required by the model seems orthogonal to the merits of our statistical approach of using such data to infer bias. [MORE].
+
+* Repeat comment about accessibility
+
+`3.
+In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias.
+If one has been called for jury service in a city with a large minority population and been in a jury room, this is highly implausible.
+The reviewer once was sitting next to a minority woman and we saw---A minority leave the jury room, then a white, then another minority, then another white etc.
+At the end both of us noticed the “pattern”.
+Because one is analyzing a changing pool of possible challenges each time a peremptory challenge is made, when the defense attorney (assuming a minority defendant—the usual situation where Batson issues arise) removes a white, the minority proportion of the pool of possible individuals to be removed increases.
+This could make it more difficult to detect a biased prosecutor.
+On the other hand, if the method detects prosecutorial bias when the defense challenges appear biased too, this should strengthen the evidence of prosecutorial bias.
+This issue deserves some quantitative investigation, so judges can feel confident when using the results.`
+
+* [TEM&EB: Check this to see if adding bias to the defense alters bias detection for the prosecutors. ][see table below]
 
@@
 
-Review 1
-Journal: The American Statistician
-Manuscript ID: TAS-22-165
-Title: Bayesian detection of bias in peremptory challenges using historical strike data
-The authors presented a Bayesian model in detecting potential biases against minority groups
-in peremptory challenges. The use of the power prior to incorporate historical data is in-
-teresting, and the methodology overall seems reasonable. The proposed R Shiny App may
-also be helpful to practitioners. My main concerns are about the simulation results and the
-choice of the hyperparameter 
-in the power prior.
-Major comments:
-
-1.
-In Section 2.3, the paragraph in line 47 on p.
-6 ends with bhist = f3;2; : : : ; 3g,but in the very next paragraph, bhist is set to f0; 0:5; : : : ; 3g.
-I don't see bhist =f3;2; : : : ; 2; 3g mentioned anywhere else, including in the results shown in Figures1 to 3.
-I am confused about which set of values was used in the simulation for bhist.
-
-* Fix bhist typo[DONE]
-
-* Repeat metal detector response here.
-
-* Ask Xiaomeng about this in the app [TEM:  reach out]
-
-* Repeat metal detector response here.
-Could mention this as something that we considered, and mention that we would like find a way to measure compatibility.
-This is also tool intended to be used by lawyers in real-time.
-
 Main Comments:
-1. The method is first illustrated in a simulation study, with a range of values of $\alpha$.
-This is fine for an audience of statisticians and attorneys specializing in jury
-discrimination but is somewhat overwhelming for a general reader. It would be
-easier for the reader if one data set, perhaps from the actual data from Connecticut
-was used and the reader could see how the data from each additional previous trial
-impacts the posterior. Perhaps only a few values of $\alpha$, rather than 11 and two intervales
-of the bias parameter b, could be used.
-
-* Start with CT data as example. Take screenshots from the app part.
-
-> 2.
-From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones.
-How often is this data currently available?
-If it isn’t readily available now, how likely are states and locales to implement a system of systematically reporting and preserving the required data (especially in the areas where one expects minorities to encounter bias in jury membership)?
-
-* Repeat comment about accessibility
-
-`3.
-In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias.
-If one has been called for jury service in a city with a large minority population and been in a jury room, this is highly implausible.
-The reviewer once was sitting next to a minority woman and we saw---A minority leave the jury room, then a white, then another minority, then another white etc.
-At the end both of us noticed the “pattern”.
-Because one is analyzing a changing pool of possible challenges each time a peremptory challenge is made, when the defense attorney (assuming a minority defendant—the usual situation where Batson issues arise) removes a white, the minority proportion of the pool of possible individuals to be removed increases.
-This could make it more difficult to detect a biased prosecutor.
-On the other hand, if the method detects prosecutorial bias when the defense challenges appear biased too, this should strengthen the evidence of prosecutorial bias.
-This issue deserves some quantitative investigation, so judges can feel confident when using the results.`
-
-* [TEM&EB: Check this to see if adding bias to the defense alters bias detection for the prosecutors. ][see table below]
 
@@
 
-`Gastwirth (2005) in Law, Prob. Risk, p. 179-185, . . . had proposed that
-approach when potential jurors were removed one at a time in an alternating
-sequence. Checking that reference, the reviewer saw that at the end of the 2005
-paper, the use of data from prior cases would be helpful and recommended that
-procedures, such as the one proposed here. This connection should be noted. Also,
-if one had data from an actual case, one could show the additional insight one
-obtains by incorporating data from previous cases.`
+`Gastwirth (2005) in Law, Prob.
+Risk, p.
+179-185, .
+.
+.
+had proposed that approach when potential jurors were removed one at a time in an alternating sequence.
+Checking that reference, the reviewer saw that at the end of the 2005
+paper, the use of data from prior cases would be helpful and recommended that procedures, such as the one proposed here.
+This connection should be noted.
+Also, if one had data from an actual case, one could show the additional insight one obtains by incorporating data from previous cases.`
 
@@
 * Add sensitivity for defense bias (new)
-
-Add real data example to introduction - make more explicit
-
-Table of bias detection for the prosecutor, when the defense shows bias. N historical = 3, nstrike =15.
-
-	1.0

### 2023-01-05 analysis/paper_draft3.Rmd
updated abstract
front end discussion about including historical data

@@
 Here, we extend a Bayesian approach for detecting such illegal bias.
-We show how to incorporate historical data on an attorney's use of peremptory strikes in past cases to estimate how likely that attorney has struck prospective jurors in a current trial because of their cognizable-class membership.
+Current methods for detecting bias make explicit, but unacknowledged, assumptions about how to incorporate historical strike data into the bias detection proess.
+Here, we show how to incorporate historical data on an attorney's use of peremptory strikes in past cases to estimate how likely that attorney has struck prospective jurors in a current trial because of their cognizable-class membership.
 In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
-We show how well our model performs on simulated data.
-Finally, we extend this approach with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection.
+Using simulations, we show how choice of alpha influences bias detection in the face of varying degrees of incompatibility between current and historical trial data.
+Finally, we extend this approach with a prototype software application (R Shiny application) to detect bias in peremptory-challenge in real time during jury-selection.
 We illustrate this application's use with real, historical strike data from a convenience sample of cases from one court.

### 2023-01-05 analysis/paper_draft3.Rmd
attempt at adding sentences on incompatibility

@@
 In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, proving a prima facie case, and ultimately persuading the trial judge that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
-One challenge for incorporating historical data is possible incompatibility between current and historical strike data for a given attorney that may arise because of missing data or because attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).[may need fleshing out]
+One challenge for incorporating historical data is possible incompatibility between current and historical strike data for a given attorney that may arise because of missing data or because attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).
+Current statistical approaches to the detection of bias have two extreme approaches to inclusion of historical data regardless of incompatibility between current and historical trials.
+Firstly,  historical data can be ignored entirely.
+This approach may lead to underpowered estimates of bias due to the low number of strikes in a single trial (typically 6, 10, or 15 strikes per party), and may lead to inflated estimates of bias.
+Secondly, historical and current trial data can be combined and analysed together.
+While this approach increases the sample size, it ignores the possible effects of incompatiblity between current and historical trial.
 
@@
 We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
-In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posteri-or distribution developed with the historical information and the power prior.
+In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posterior distribution developed with the historical information and the power prior.
 
-The use of the power prior in this context allows lawyers to explicitly control over the assumed degree of incompatibility between the current and historical trial data.
+The use of the power prior in this context allows lawyers explicit control over the assumed degree of incompatibility between the current and historical trial data.
 Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected when incompatibility is expected to be high, and higher values (closer to 1) of the power prior parameter selected when incompatibility is low.

### 2023-01-05 analysis/paper_draft3.Rmd
minor edits

--- 2022-12-13 analysis/paper_draft3.Rmd
@@
 We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
-In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posteri-or distribution developed with the historical information and the power prior.
+In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posterior distribution developed with the historical information and the power prior.
 
-The use of the power prior in this context allows lawyers to explicitly control over the assumed degree of incompatibility between the current and historical trial data.
-Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected when incompatibility is expected to be high, and higher values (closer to 1) of the power prior parameter selected when incompatibility is low.
-Accordingly, the lawyer bringing the _Batson_ challenge chooses the value of $\alpha$ and must be able to explain and defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current strike data and historical strike data available in the particular case.
+The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data.
+Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected if incompatibility is expected to be high.
+Higher values (closer to 1) of the power prior parameter should be selected if incompatibility is expected to be low.
+Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current and historical strike data in the particular case.
 

### 2023-01-05 analysis/paper_draft3.Rmd
added to discussion

--- 2023-01-05 analysis/paper_draft3.Rmd
@@
 While this approach increases the sample size, it ignores the possible effects of incompatiblity between current and historical trial.
+Such incompatibility can lead to inflated type I error and reduced power [@Bennett2020].
 
@@
 [editme] The goal of these methods is not to _prove_ bias, but rather to lend support to an inferrence of bias and a subsequent _Batson_ challenge.
-The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
-[editme] sentence on why this is appealing
+Previously used methods have either ignored historical trial data, or conducted analyses where current and historical data are merged, with no regard for possible incompatibility between trials.
+The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
+By incorporating the power prior, one can assess how sensitive that posterior estimate is to the available historical strike data.
+This is appealing because it makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
 
-Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
+Moreover, with a tool like our prototype app, attorneys can use the approach presented here to better detect strike bias for _Batson_ and related challenges.
 The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether to make a  _Batson_ challenge at all.
@@
 In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to inferring bias based on historical strike data has that ability.
+In future work we hope to examine data-driven approached to the selection of $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
+[editme - refs?] Knowing how to quantify incompatibility and how to specify and appropriate prior $\alpha$ is an area of active research [e.g., https://arxiv.org/abs/2206.04379]
 

### 2023-01-05 analysis/paper_draft3.Rmd
Update paper_draft3.Rmd

--- 2023-01-05 analysis/paper_draft3.Rmd
@@
 In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, proving a prima facie case, and ultimately persuading the trial judge that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
-One challenge for incorporating historical data is possible incompatibility between current and historical strike data for a given attorney that may arise because of missing data or because attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).[may need fleshing out]
+Ignoring historical strike data increases the risk of underpowered or inflated estimates of attorney strike bias, because of the low number of strikes per trial (typically 6, 10, or 15 strikes per party).
+
+Relying on historical strike data, however, requires an approach to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
+As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and based on historical strike data for that attorney.
+Simply pooling the the strike data from current and past trials, and analyzing such data together, ignores such incompatibility.
 
@@
 ```
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows us to explicitly incorporate the degree of incompatibility between current and historical trials.
+
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows us to explicitly and transparently adjust for our assumptions about the degree of incompatibility between current and historical trials.
 In this approach, we use two steps.
@@
 Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
-[editme] The goal of these methods is not to _prove_ bias, but rather to lend support to an inferrence of bias and a subsequent _Batson_ challenge.
+These statistical methods do not aim to  _prove_ bias alone, but rather to lend support to an inferrence of bias and a subsequent _Batson_ challenge.
+
 The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
 By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
-[editme] sentence on why this is appealing
-
-Moreover, with a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
+With a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
 The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether to make a  _Batson_ challenge at all.

### 2023-01-05 analysis/paper_draft3.Rmd
update with Eric

--- 2023-01-05 analysis/paper_draft3.Rmd
@@
 Here, we extend a Bayesian approach for detecting such illegal bias.
-Current methods for detecting bias make explicit, but unacknowledged, assumptions about how to incorporate historical strike data into the bias detection proess.
+Current methods for detecting bias make implicit, but unacknowledged, assumptions about how to incorporate historical strike data into the bias detection process.
 Here, we show how to incorporate historical data on an attorney's use of peremptory strikes in past cases to estimate how likely that attorney has struck prospective jurors in a current trial because of their cognizable-class membership.
 In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
-Using simulations, we show how choice of alpha influences bias detection in the face of varying degrees of incompatibility between current and historical trial data.
+Using simulations, we show how the choice of alpha influences the estimation and detection of bias in the face of varying degrees of incompatibility between current and historical trial data.
 Finally, we extend this approach with a prototype software application (R Shiny application) to detect bias in peremptory-challenge in real time during jury-selection.
@@
 
-In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, proving a prima facie case, and ultimately persuading the trial judge that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
+In turn, the estimation of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, proving a prima facie case, and ultimately persuading the trial judge that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
 One challenge for incorporating historical data is possible incompatibility between current and historical strike data for a given attorney that may arise because of missing data or because attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity).
-Current statistical approaches to the detection of bias have two extreme approaches to inclusion of historical data regardless of incompatibility between current and historical trials.
+Current statistical approaches to the detection of bias have two extreme approaches for inclusion of historical data.
 Firstly,  historical data can be ignored entirely.
@@
 The use of the power prior in this context allows lawyers explicit control over the assumed degree of incompatibility between the current and historical trial data.
+A commonly adopted practice is to fix $\alpha$ and assess the sensitivity of results to different values [Carvalho & Ibrahim - add to bib].
 Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected when incompatibility is expected to be high, and higher values (closer to 1) of the power prior parameter selected when incompatibility is low.
 Accordingly, the lawyer bringing the _Batson_ challenge chooses the value of $\alpha$ and must be able to explain and defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current strike data and historical strike data available in the particular case.
+It should be noted that the purpose of this procedure is to estimate bias to support the bringing of a _Batson_ challenge.
+As such the senstivity of bias detection to the value of $\alpha$ is used to evaluate sensitivity of the estimation of bias and is not indicative of the sucess/failure of a _Batson_ challenge.
 
@@
 By incorporating the power prior, one can assess how sensitive that posterior estimate is to the available historical strike data.
-This is appealing because it makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
+This is appealing because it makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether or not to bring a _Batson_ challenge.
 
@@
 This transparency matters, because it thwarts lawyer efforts to strategically select $\alpha$ values to prove or disprove illegal strike bias.
-Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias inference is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to inferring bias based on historical strike data has that ability.
+Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias estimation is to the choice of $\alpha$.
+In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to estimating bias based on historical strike data has that ability.
+
 In future work we hope to examine data-driven approached to the selection of $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
-[editme - refs?] Knowing how to quantify incompatibility and how to specify and appropriate prior $\alpha$ is an area of active research [e.g., https://arxiv.org/abs/2206.04379]
+[editme - refs?] Knowing how to quantify incompatibility and how to specify and appropriate prior for  $\alpha$ is an area of active research [e.g., https://arxiv.org/abs/2206.04379, cite Eric]
 

### 2023-01-06 analysis/paper_draft3.Rmd
updated draft text

--- 2023-01-05 analysis/paper_draft3.Rmd
@@
 As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and based on historical strike data for that attorney.
-Simply pooling the the strike data from current and past trials, and analyzing such data together, ignores such incompatibility.
+Simply pooling the the strike data from current and past trials, and analyzing such data together, ignores such incompatibility, which can lead to inflated type I error and reduced power [@Bennett2020].
 
@@
 The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data.
-Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected if incompatibility is expected to be high.
+A commonly adopted practice is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho & Ibrahim].
+Specifically, lower values (closer to 0) of $\alpha$ should be selected if incompatibility is expected to be high.
 Higher values (closer to 1) of the power prior parameter should be selected if incompatibility is expected to be low.
-Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current and historical strike data in the particular case.
+Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately corresponding to the likely degree of incompatibility between the current and historical strike data in the particular case.
+Since this overall procedure is to estimate bias to support the bringing of a _Batson_ challenge, the senstivity of bias detection to the value of $\alpha$ is used to evaluate sensitivity of the estimation of bias, not to indicate whether a _Batson_ challenge should succeed.
 
@@
 Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
-These statistical methods do not aim to  _prove_ bias alone, but rather to lend support to an inferrence of bias and a subsequent _Batson_ challenge.
+These statistical methods do not aim to  _prove_ bias alone, but rather to lend support to an inference of bias and a subsequent _Batson_ challenge.
+Previously used methods have either ignored historical trial data, or conducted analyses where current and historical data are merged, with no regard for possible incompatibility between trials.
+The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
+By incorporating the power prior, one can assess how sensitive that posterior estimate is to the available historical strike data.
+This is appealing because it makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
 
-The Bayesian approach here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.
-With a tool like our prototype app, attorneys can use the approach here in to better detect strike bias for _Batson_ and related challenges.
+Moreover, with a tool like our prototype app, attorneys can use the approach presented here to better detect strike bias for _Batson_ and related challenges.
 The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether to make a  _Batson_ challenge at all.
 If argued, the attorney can introduce that posterior distribution as relevant evidence for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+
 To be sure, such evidence alone cannot account for the veracity of the striking attorney's proferred reasons.
@@
 Thus, as with any use of historical strike data, the approach here cannot always determine whether a _Batson_ challenge should succeed or fail.
-
 The approach here also requires more from those who would collect the appropriate historical strike data.
@@
 This transparency matters, because it thwarts lawyer efforts to strategically select $\alpha$ values to prove or disprove illegal strike bias.
-Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias inference is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to inferring bias based on historical strike data has that ability.
+Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias estimation is to the choice of $\alpha$.
+In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to estimating bias based on historical strike data has that ability.
+
+In future work we hope to examine data-driven approached to the selection of $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
+Knowing how to quantify incompatibility and how to specify and appropriate prior for  $\alpha$ is an area of active research [e.g., @Pawel2022]
 

### 2023-01-06 analysis/TAS_memo/memo_TAS_reviewers.Rmd

@@
 
+While it is true, based on our simulations, that bias can be detected even if there is no bias in the current trial's strike pattern under certain conditions: if there are a sufficient number of historical cases (e.g., n=3) and if bias in the historical trials is sufficiently strong (i.e., strong incompatibilty).
+Contrary to this comment, this is exactly why we advocate for the use of $\alpha$!
+In scenarios like the one just described (n=3, strong incompatibility), sensitivity of bias detection can be evaluated by changing alpha and noting its effect on the estimation of bias.
+
 Finally, the reviewer's use of the word "guilty" requires two clarifying notes.
@@
 whether the credible interval includes 0 or not.
-We also present sensitivity analysis for the selection of the nominal HPD level.
+We also present sensitivity analysis for the selection of the nominal credible interval level.
 
@@
 >2.
-About the results presented in Figure 1 of Section 2.3, if we ignore the assumption of compatibility between historical and current data, in the cases of $\alpha$ = 0:5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0?
+About the results presented in Figure 1 of Section 2.3, if we **ignore the assumption of compatibility** between historical and current data, in the cases of $\alpha$ = 0.5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0?
 If so, the Type I error seems to be very high in some cases.
@@
 
+[Maybe mention that these issues - low power, type I error, etc are not unique to our method?]
+
 >3.
@@
 
-[INSERT] Will get Eric to check this (still no response from Xiaomeng)
+[INSERT] Will get Eric to check this (got a response from Xiaomeng - waiting for follow-up)
 
@@
 The plausible reasons for non-random missingness and confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
+
+[Mention that measuring incomaptibility and choice of a $\alpha$ in face of incompatibility is an active area of research (refs) with no clear solution.
+Our providing range of alpha values so that users can evaluate sensitivity is common practice (carvallo & ibrahim 2021)]
 

### 2023-01-11 analysis/paper_draft3.Rmd
add sentences on fisher's test and CMH

@@
 Simply pooling the the strike data from current and past trials, and analyzing such data together, ignores such incompatibility, which can lead to inflated type I error and reduced power [@Bennett2020].
+Methods such as the Fishers exact test and the Cochran-Mantel-Haenszel (CMH) test (Gastwirth and Xu, 2014) can be used to test for bias and could include data from past trials, either by pooling all strike sequence data (in the case fo Fisher's test), or by stritifying the data by trial (CHM test).
+While these approaches may increase statistical power, they do not directly address possible incompatibility between current and historical trial data and may, therefore, only be appropriate when, as suggested by Gastwirth and Hu (2014), historical trail data are selected that are 'similar' to the current trial.
+This notion of simliarity require implicit weighint of historical evidence.
 

### 2023-01-11 analysis/paper_draft3.Rmd
minor edit to last line in conclusion

--- 2023-01-06 analysis/paper_draft3.Rmd
@@
 
-In future work we hope to examine data-driven approached to the selection of $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
-Knowing how to quantify incompatibility and how to specify and appropriate prior for  $\alpha$ is an area of active research [e.g., @Pawel2022]
+In future work we hope to examine data-driven approach to selecting $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
+Knowing how to quantify incompatibility and how to specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022]
 

### 2023-01-11 analysis/paper_draft3.Rmd
revision of paragraph on Cochran-Mantel-Haenszel test

--- 2023-01-11 analysis/paper_draft3.Rmd
@@
 As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and based on historical strike data for that attorney.
+
 Simply pooling the the strike data from current and past trials, and analyzing such data together, ignores such incompatibility, which can lead to inflated type I error and reduced power [@Bennett2020].
+To increase power, one may use other methods, such as stratifying strike data by trial and applying the Cochran-Mantel-Haenszel (CMH) test [@Gastwirth2014, 295-297].
+Such approaches, however, do not directly address incompatibility between current and historical strike data.
+For example, applying the CMH test requires first selecting only historical strike data from past trials that are "similar" to the current trial [@Gastwirth2014, 289, 297].
+Assessing such similarity implies weighting the available historical strike data with respect to some attorney-bias-related set of trial-level characteristics that are present or absent in the current trial.
 

### 2023-01-20 analysis/TAS_memo/memo_TAS_reviewers.Rmd
revisions to memo to TAS reviewers

--- 2022-12-13 analysis/TAS_memo/memo_TAS_reviewers.Rmd
@@
 
-This memorandum replies to comments provided by the TAS associate editor and two reviewers.
-It indicates how we have revised the paper accordingly, as well as our understanding of the issues raised by those comments.
+This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is better as a result.
+
+# Key Issues
+
+Here, we address five key issues raised by multiple reviewer comments.
+
+1.
+Our statistical approach is not an all-purpose detector of _Batson_ violations.
+No statistical approach can do this, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons.
+If the attorney fails to produce such reasons, the trial judge must find a _Batson_ violation.
+If the attorney offers such reasons, the trial judge must assess evidence as to the credibility of those reasons.
+Our statistical approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) of illegal strike bias.
+Rather, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias.
+The phrase "bias detection" in the paper means only that the credible interval for the bias parameter ($b_{curr}$) excludes zero, given the strike data and the model.
+For that "detection" task, our approach lets one assign at most equal weight to past strikes and their strikes in the present case ($\alpha = 1$).
+In turn, attorneys can use the resulting bias estimate, _combined with_ other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find  a _Batson_ violation (i.e., illegal strike bias is more likely than not, given all the evidence).
+We have revised the Introduction to emphasize all this to the reader at the outset.
+
+2.
+Possible incompatibility of $b_{hist}$ and $b_{curr}$ is always an issue when using past strike data to infer strike bias in the current trial.
+This is the case for _any_ statistical approach that uses past strike data to infer strike bias in the current trial.
+And in _Batson_ challenges involving past strike data, the lawyers are already and always arguing about how much weight to assign to that past strike data, albeit with words.
+By using the power prior, our approach simply makes the weighting of historical strike data more explicit and transparent as a choice of a value for $\alpha$.
+The more sensitive bias "detection" (i.e.
+whether the posterior estimate of $b_{curr}$ excludes zero) is to the choice of $\alpha$, the more one should worry about incompatibility.
+Thus, small vs.
+large values of $\alpha$ should vary with the suspected degree of incompatability.
+In reality, because $b_{curr}$ and $b_{hist}$ are not directly observable, one will need to rely on background knowledge or assumptions about the degree of incompatibility to decide which value of $\alpha$ to choose.
+In a _Batson_ challenge, the lawyers must disclose and justify their choice of $\alpha$ accordingly.
+We have revised the paper to emphasize this.
+
+3.
+In the adversarial setting of the courtroom, our approach thwarts the strategic selection of $\alpha$ values, because it makes the choice of $\alpha$ transparent, and thus the choice of how much to weigh the historical strike data.
+As a result, every lawyer will know that any strategic selection of $\alpha$ will be met with scrutiny from the lawyers on the other side.
+Absent this transparency, lawyers will continue to argue strategically about how much to weigh historical strike data, except with the opacity of words.
+We have revised the paper to add this to the Discussion section.
+
+4.
+Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias.
+This is because the proper weight to place on historical data depends on whether, for that historical data, there is any missing data that is missing not at random and whether inferring bias from that historical strike data is complicated by plausible confounding due to trial-level variables (e.g., defendant race, criminal charge severity).
+The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
+As we now note in the Discussion section of the paper, measuring incompatibility and choosing $\alpha$ in the face of incompatibility is an active area of research with no clear solution.
+
+5.
+Choosing a credible interval amounts to deciding how conservative one wishes to be about inferring bias from past strike data, because we define bias "detection" narrowly to mean when the posterior for the bias parameter falls outside the chosen credible interval.
+Thus, the larger the credible interval one chooses, the greater the chance that zero falls within it.
+For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on other credible intervals in the Appendix.
+
+In the following responses to particular comments, we will refer to the above items when appropriate.
 
@@
 
-`[T]he article doesn’t compellingly make the case that the methodology is more useful than simply tracking those who have been reprimanded in the past or that it’s ethical/reasonable to actually employ in the courtroom.
+>[T]he article doesn’t compellingly make the case that the methodology is more useful than simply tracking those who have been reprimanded in the past or that it’s ethical/reasonable to actually employ in the courtroom.
 Specifically, the way that prior information is encoded, the logic behind a detection appears to be circular in that if someone has been reprimanded in the past they should be reprimanded in the future (in some cases for no reason).
-This issue hasn’t been discussed to a level that makes the utility of such a model clear.`
+This issue hasn’t been discussed to a level that makes the utility of such a model clear.
 
@@
 
-First, as page X, lines Y of the paper indicate, the law of _Batson_ and parallel State law has consistently permitted judges to infer illegal strike bias based in part on an attorney's use of strikes in past cases.
+First, the law of _Batson_ and parallel State law has consistently permitted judges to infer illegal strike bias based in part on an attorney's use of strikes in past cases.
 In reviewing the kinds of evidence that can prove a _Batson_ violation by a prosecutor, the U.S.
-Supreme Court included, as examples of types of evidence, the "relevant history of the State’s peremptory strikes in past cases." Flowers v.
-Mississippi, 139 S.
+Supreme Court included, as examples of types of evidence, the "relevant history of the State’s peremptory strikes in past cases." _Flowers v.
+Mississippi_, 139 S.
 Ct.
-2228, pin (2019).
-More recent State law variants on _Batson_ are also explicit about this.
+2228, [PIN] (2019).
+State law variants on _Batson_ are also explicit about this.
 For example, Washington State law provides that in deciding whether "an objective observer could view race or ethnicity as a factor in the use of the peremptory challenge," the circumstances the court "should" consider include "whether the party has used peremptory challenges disproportionately against a given race or ethnicity, in the present case _or in past cases_." Wash.
 General Rule 37(g)(v) (emphasis added).
-California law similarly identifies, as a circumstance a court may consider, whether the striking "has used peremptory challenges disproportionately against a given race, ethnicity, gender, gender identity, sexual orientation, national origin, or religious affiliation, or perceived membership in any of those groups, in the present case _or in past cases_." Calif.
+California law similarly identifies, as a circumstance a court may consider, whether the striking party "has used peremptory challenges disproportionately against a given race .
+.
+.
+in the present case _or in past cases_." Calif.
 Code of Civil Procedure 231.7(d)(3)(G) (emphasis added).
@@
 
-Second, as the phrase "should be reprimanded in the future (in some cases for no reason)" implies, the associate editor seems worried that our approach will cause judges to infer that illegal strike bias is more likely than not (i.e., find _Batson_ violations ) when in fact no strike bias exists.
+Second, as the phrase "should be reprimanded in the future (in some cases for no reason)" implies, the apparent worry is that our approach will cause judges to infer, in any particular case, that illegal strike bias is more likely than not (i.e., find _Batson_ violations) when in fact it was less likely than not.
+This is not the case.
+Our approach does not and cannot automate findings of _Batson_ violations.
+See Item 1, above.
 
-This is not the case, and we have revised the paper to make this more clear.
-
-First, we assume the word "reprimanded" refers to a court finding of a _Batson_ violation.
+Third, we assume the word "reprimanded" refers to a court finding of a _Batson_ violation.
 Our approach, however, relies on an attorney's use of strikes of past cases, regardless of whether those strikes were ever the subject of a _Batson_ challenge.
-This is better than just "tracking those who have been reprimanded in the past" for _Batson_ violations, because the probability of a court that finds a _Batson_ violation depends on an attorney having raised a _Batson_ challenge in the first place.
+This is better than just "tracking those who have been reprimanded in the past" for _Batson_ violations, because the probability that a court finds a _Batson_ violation depends on an attorney having raised a _Batson_ challenge in the first place.
 (Absent an attorney arguing _Batson_, a court will not inquire into whether illegal strike bias existed.) If illegal strike bias exists but the attorney erroneously perceives no credible basis for it, the attorney will tend not to argue _Batson_.
 In cases where there no illegal strike bias but the attorney erroneously perceives some credible basis for it, the attorney will tend to argue _Batson_ but may not persuade the trial judge that illegal strike bias was more likely than not.
-By relying on past strike data, not whether anyone argued _Batson_ in those past cases, we avoid any selection bias associated with relying only on past court findings of _Batson_ violations.
+By relying on past strike data, not whether anyone argued _Batson_ in those past cases, we avoid any bias associated with relying only on past court findings of _Batson_ violations.
 
-Second, _Batson_ procedure permits any type of admissible evidence of strike bias (for or against) -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
-For this reason, our approach alone is not an all-purpose detector of _Batson_ violations, because our approach cannot incorporate the weight of all the other kinds of evidence that could be submitted to prove illegal strike bias in the present case.
-For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons.
-If the attorney fails to produce such reasons, the trial judge must find a _Batson_ violation.
-If the attorney offers such reasons, the trial judge must assess evidence as to the credibility of those reasons.
+>Page 4, line 5: The authors refer to gender but then provide sex options (male/female).`
 
-Our approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) of illegal strike bias.
-Rather, as the new Figure 1 depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias.
-For that task, our approach assigns at most equal weight to past strikes and their strikes in the present case (i.e., when $\alpha = 1$).
+Both the model and the prototype software encode "gender" as binary (male or female) simply because the underlying strike data in the app is based on standard juror questionaires that provide "male" and "female" as the only mutually-exclusive options.
 
-The possible incompatability of $b_hist$ and $b_curr$ is already and always an issue when using past strike data to infer bias.
-This is a reality faced by _any_ statistical technique for inferring strike bias using past strike data.
-Because the law of _Batson_ expressly permits inferring current strike bias based on past strikes in previous trials, the lawyers in _Batson_ challenges involving past strike data are already and always arguing about how much weight to assign to that past strike data.
+>Page 4, line 36: Use `\left(\right)` to ensure appropriately sized parentheses where necessary.`
 
-Our approach simply makes the weighting of historical strike data -- usually expressed with words -- more explicit and transparent by requiring a choice of a value for alpha.
-The more sensitive the inference of bias (zero falls outside the credible interval) is to the choice of $\alpha$, the more one should worry about incompatibility.
-In turn, the appropriateness of small vs.
-large values of $\alpha$ vary with the suspected degree of incompatability.
-In turn, attorneys can use that information, _combined with_ other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find that illegal strike bias is more likely than not, given all the evidence, i.e., find a _Batson_ violation.
+Thank you for spotting this.  The paper has been revised accordingly.
 
-We hope that these modifications clarify the placement of our model in the steps from a sequence of peremptory challenges to a _Batson_ challenge actually being brought, and upheld by the trial judge.
-
-### Additional Comments
-
-`1. Page 4, line 5: The authors refer to gender but then provide sex options (male/female).`
-
-[comment]
-
-`2. Page 4, line 36: Use \left(\right) to ensure appropriately sized parentheses where necessary.`
-
-Thank you for this suggestion. The paper has been revised accordingly.
-
-`3. Page 5, equation 6: The left-hand side-of the equation is not correct`
+>Page 5, equation 6: The left-hand side-of the equation is not correct`
 
@@
 
-`4.
-Page 6, line 43: The idea behind $\alpha$ is concerning to me.
+>Page 6, line 43: The idea behind $\alpha$ is concerning to me.
 The idea here is that if there is previous problematic strike data, and we choose large $\alpha$, then we’ll detect problematic strikes at higher rates.
@@
 
-Finally, the reviewer's use of the word "guilty" requires two clarifying notes.
-First, _Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more stringent "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the criminal offense charged.
-Second, as explained above, bias "detection" in the paper means only whether the credible interval for the bias parameter ($b_{curr}) excludes zero.
-Because _Batson_ procedure requires weighing any and all evidence of strike bias, not just past strike data, our approach cannot alone identify _Batson_ violations in cases involving other relevant evidence bias besides past strike data.
+It is true, based on our simulations, that bias can be erroneously detected (i.e.
+even if $b_curr$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibilty.
+On the other hand, given high compatibility, low values of $\alpha$ increase Type I error.
+This is why we recommend using different values of $\alpha$ to check for sensitivity.
+See Item 2 above.
 
-`5. Page 6, line 50: The description of bhist is not clear; it is defined twice with different values.`
+The use of the word "guilty" is error.
+_Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more stringent "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the criminal offense charged.
 
-The paper has been revised to clarify this description.
+>Page 6, line 50: The description of bhist is not clear; it is defined twice with different values.`
 
-`6. Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?`
+We revised the paper to clarify this description.
 
-We have revised this discussion to remove such confusion.
-The choice of the credible interval is a decision about how conservative one wishes to be about inferring bias from past strike data alone.
-We define bias "detection" narrowly to mean when $b = 0$ falls outside the chosen credible interval.
-Thus, the larger the credible interval, the greater the chance that zero falls within it.
-For the sake of clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on other credible intervals in the Appendix.
-This does not necessarily fix the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), and we are interested in "bias detection", i.e.
+>Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?`
+
+We revised to remove such confusion.
+On the choice of credible intervals, see Item 5 above.
+The revision does not necessarily fix the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), whereas we are interested in "bias detection", i.e.
 whether the credible interval includes 0 or not.
-We also present sensitivity analysis for the selection of the nominal HPD level.
+The paper also has a sensitivity analysis for the selection of the nominal credible interval level.
 
-`7.
-Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior.
+>Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior.
 One would be curious whether the approach with no prior works for detecting bias (e.g., simulations results without prior data).
@@
 This comment is not clear to us.
-The real data analyses are based on real _historical_ trial data from the federal district court in Connecticut (see lines x to y).
-Users of the app have the option to run the model excluding any historical trial data (as suggested by the comment) by not selecting the name of a prosecutor or defense attorney.
-In this case, the normal(0,2) prior is used for the bias parameter.
-Similarly, the app gives the user the option of altering the weight the historical strike data (if selected) by adjusting alpha (power prior parameter).
-Thus the user can: 1) choose not to use historical data, 2) use historical data with equal weight as the current trial (alpha = 1), or 3) down-weight the historical trial data.
+The real data analyses are based on real historical strike data from real criminal cases in the federal district court in Connecticut.
+If they wish, users of the app can run the model excluding any historical data (as this reviewer suggested) by not selecting the name of a prosecutor or defense attorney.
+Like our model, the app uses the $N(0,2)$ prior for the bias parameter.
+Similarly, the app user can alter the weight the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
+Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2).
 
-`8. Page 18, line 45: If there is benefit to the user for finding bias, the ability to select $\alpha$ is akin to p-hacking.`
+>Page 18, line 45: If there is benefit to the user for finding bias, the ability to select $\alpha$ is akin to p-hacking.`
 
-We respectfully disagree with the analogy to "p-hacking", for two reasons.
-First, the phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify as "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results.
-In contrast, in our approach, the statistical model necessarily precedes any analysis, because it has to qualify as a valid model of the strike _procedure_ of the court of interest, which can and must be known before any data is collected or analyzed.
+We respectfully disagree with the analogy to "p-hacking".
+The phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify as "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results.
+In contrast, in our approach, the statistical model has to qualify as a valid model of the strike _procedure_ of the court of interest.
+That strike procedure can and must be known and modeled before any strike data is collected or analyzed.
+More generally, our approach thwarts the strategic selection of $\alpha$ values by making the choice of $\alpha$ transparent.
+See Item 4, above.
 
-Second, in the adversarial setting of the courtroom, our approach actually thwarts the strategic selection of $\alpha$ values, because it makes the choice of $\alpha$ transparent.
-This is why we stress in the paper (p.
-X, lines Y) that, where there is suspected incompatibility between $b_{hist}$ and $b_{curr}$, the proper approach is try multiple values of $\alpha$ and determine how sensitive the inference of bias is to the chosen $\alpha$ value.
-Because the choice of how much to weigh the historical strike data is transparent, every lawyer knows that any strategic selection of $\alpha$ will be met with scrutiny from the lawyers on the other side.
-Absent this transparency, lawyers already still argue for how much past strike behavior ought to weigh when inferring or dispute illegal strike bias, except that they do so implicitly and with the opacity of words.
-
-`9.
-Something unmentioned that should be addressed is confounders.
+>Something unmentioned that should be addressed is confounders.
 While technological and statistical tools can be helpful to the judicial system, and even decrease instances of discrimination and curb disparities, there are many unintended consequences of such systems.
@@
 
-[INSERT] Batson challenge doesn't care if you strike a cognizable class because you assume they have particular confounders (e.g., low income, etc).
-[editme] We do explicitly highlight that trial-level characteristics and impact compatibility between current and historical trials.
-We do advocate for reassessment of the way (and what) data are collected from individual trials.
+On confounders, we expressly highlight that variation in trial-level characteristics (e.g.
+defendant race, charge severity) in the historical data can generate incompatibility, which in turn can affect bias detection.
+However, with respect to juror-level characteristics, the comment does not apply.
+Suppose a prosecutor strikes three Black prospective jurors on the belief that Black people are generally less educated and therefore more likely to favor defendants.
+This is illegal under _Batson_, regardless of the actual education levels of the three jurors struck.
+[FIX].
+
+On unintended consequences, we found this comment unclear.
+The paper aims to show the validity of our Bayesian approach for incorporating historical strike data for estimating strike bias in the context of _Batson_ challenges.
+We do not aim to identify any and all consequences of a full-scale implementation of this approach, intended or unintended.
 
@@
 
->1. In Section 2.3, the paragraph in line 47 on p. 6 . . . I am confused about which set of values was used in the simulation for bhist.
+>In Section 2.3, the paragraph in line 47 on p. 6 . . . I am confused about which set of values was used in the simulation for bhist.
 
-Thank you for catching this.  We have revised the paper (p. X, lines Y) to clarify the values used in the simulation for $b_{hist}$.
+Thank you for catching this.  We have revised the paper to clarify the values used in the simulation for $b_{hist}$.
 
->2.
-About the results presented in Figure 1 of Section 2.3, if we ignore the assumption of compatibility between historical and current data, in the cases of $\alpha$ = 0:5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0?
+>About the results presented in Figure 1 of Section 2.3, if we **ignore the assumption of compatibility** between historical and current data, in the cases of $\alpha$ = 0.5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0?
 If so, the Type I error seems to be very high in some cases.
@@
 
-We agree that the bias "detection" rate (whether the credible interval excludes zero) can be as high as 100% when true bias is zero.
-We also agree that the power to detect bias is low when there is low bias in historical trials.
-These situations can occur because of high incompatibility.
-Similarly, the low power situations occurs due the limited number of strikes, which is why _compatible_ historical strike data can increase this power.
+We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_curr$); and (2) bias detection can be low where current bias is high but bias in historical trials is low.
+The tradeoff is that, given compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
+The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $alpha$.
 
-This is an advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting alpha.
-In reality, because $b_{curr}$ and $b_{hist}$ are not directly observable, one will need to rely on background knowledge or assumptions about the degree of incompatibility to decide which value of alpha to choose.
-If they bring a _Batson_ challenge, they must also disclose and justify their choice of alpha and the assumptions on which that choice is based.
-
->3.
-Regarding the Shiny App in Section 2.5, it would be helpful to include some details about the Metropolis-Hastings scheme.
+>Regarding the Shiny App in Section 2.5, it would be helpful to include some details about the Metropolis-Hastings scheme.
 For instance, is a normal distribution used as the proposal distribution for b?
@@
 
-[INSERT] Will get Eric to check this (still no response from Xiaomeng)
+[INSERT] Will get Eric to check this (got a response from Xiaomeng - waiting for follow-up)
 
->4.
-In Section 3, the authors state that "the anticipated degree of incompatibility should influence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error.
+>In Section 3, the authors state that "the anticipated degree of incompatibility should influence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error.
 However, based on results in Figures 1 to 3, if this assumption is wrong, choosing a large $\alpha$ leads to either low power (bias detection rate) or large Type I error, depending on the bias in historical data.
@@
 
-As indicated in the response to comment 2 above, one chooses a value of $alpha$ based on background beliefs on plausible degree of incompatibility, and then evaluate how sensitive bias detection (whether zero falls outside the credible interval) is to the choice of alpha value.
-Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$, because the proper weight to place on historical data depends on whether, for that historical data, there is any missing data that is missing not at random and whether an inference of bias based on that historical strike data is complicated by plausible confounding or mediating trial-level variables (e.g., defendant race, criminal charge severity).
-The plausible reasons for non-random missingness and confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
+See Items 2 and 4 above.
 
->5. Some of the Figure and table captions (e.g. Figures 1-5 and Table 1) can benefit from being more descriptive.
+>Some of the Figure and table captions (e.g. Figures 1-5 and Table 1) can benefit from being more descriptive.
 
-[INSERT]
+We have revised accordingly.
 
-> 6. A couple of notation issues: . . . .
+>A couple of notation issues: . . . .
 
@@
 
->1.
-The method is first illustrated in a simulation study, with a range of values of $\alpha$.
+>The method is first illustrated in a simulation study, with a range of values of $\alpha$.
 This is fine for an audience of statisticians and attorneys specializing in jury discrimination but is somewhat overwhelming for a general reader.
@@
 
-* Start with CT data as example. Take screenshots from the app part.
+[INSERT]
 
-> 2.
-From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones.
+>From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones.
 How often is this data currently available?
@@
 In the Discussion, we explained that data on consecutive strike patterns is hard to collect, simply because of the challenges of collecting any historical strike data, regardless of strike procedure.
-We have revised that Discussion to add that, since the 1980s, any person has a First Amendment right to attend jury selection in criminal court proceedings, and criminal defendants have a Sixth Amendment right to public court proceedings during jury selection.
-These legal rights, in turn, enable data collection by court observers, regardless of State and local recordkeeping practices.
+We have revised that Discussion to add that, since the 1980s, any person has the right to attend jury selection in criminal court proceedings These legal rights enable data collection by court observers, regardless of State and local recordkeeping practices.
+Finally, our statistical approach's validity does not turn on the _current_ availability of the kind of data it requires.
+Indeed, we hope our statistical approach can inform efforts as to what kinds of data about strikes a court system should make more available.
 
-Finally, the current availability of the data required by the model seems orthogonal to the merits of our statistical approach of using such data to infer bias. [MORE].
-
-* Repeat comment about accessibility
-
-`3.
-In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias.
+>In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias.
 If one has been called for jury service in a city with a large minority population and been in a jury room, this is highly implausible.
@@
 
-* [TEM&EB: Check this to see if adding bias to the defense alters bias detection for the prosecutors. ][see table below]
+[INSERT TABLE]
 
-Data for past trials is not readily available.
-But, since the 80s, any person can enter a courtroom and observe the jury selection process, so this data could be collected for present and future cases.
+>Page 5, equation 5: The left-hand side-of the equation is not correct
 
-* [add paragraph to address this]
+We have revised to indicate that it is a posterior distribution. All formulas have now been updated and checked. We now specify the vector notation for $\delta_0$ and $\delta_1$.
 
-Further, in ‘Other Comments’ 3-4, they describe a rather important realities about the application that are not accounted for in the model.
-While the problem is certainly interesting, the simulations demonstrate some highly problematic results.
-Knowledgeable reviewers and I still have several large unanswered questions about how the methodology can be enacted appropriately. I provide my comments below.
+>Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?
 
-3. Page 5, equation 5: The left-hand side-of the equation is not correct
-* [EB: fix in formula] - This is now fixed to reflect that it is a posterior distribution. We note that all formulas have now been updated and checked.
+See Item 5 above.
 
-We now specify the vector notation for $\delta_0$ and $\delta_1$.
-
-Mention that in the app, it is easy to check the influence of alpha by re-running the model.
-Make a clearer case that level of alpha should be selected in cases of high incompatibility.
-Law allows you to rely on the prosecutor's past behavior.
-The ability to modify alpha, allows you to decide whether inference based on past behavior is warranted.
-[TEM & EB]
-
-6. Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you  saying that larger credible interval percentages lead to less accurate bias detection?
-
-7.
-Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior.
-One would be curious whether the approach with no prior works for detecting bias (e.g., simulations results without prior data).
-I think I would find this type of model more compelling.
-
-The real data analysis, as the default option, does not borrow historical data (i.e.
-$\alpha=0$).
-In this case the default prior of $N(0,2)$ is used for the bias parameter (i.e.
-b).
-This is reflected when no attorney is selected from the pull-down menu.
-If an attorney is selected from the pull-down menu, then prior information is used for that attorney.
-The paper has been updated to reflect this point in the same section.
-
-Main Comments:
-
-`4.
-The paper and software focus on peremptory challenges made by the same lawyer in previous cases.
+>The paper and software focus on peremptory challenges made by the same lawyer in previous cases.
 As noted in the Miller-El cases, the training manual used by prosecutors recommended the order in which groups, e.g.
@@
 Therefore, data from prosecutors from the same DA’s office or defense firm in similar cases (drug dealing, robbery etc.), should be quite informative.
-Should one simply add them to the list of prior cases, or make a down-weighting adjustment in the Bayesian analysis?`
+Should one simply add them to the list of prior cases, or make a down-weighting adjustment in the Bayesian analysis?
 
-* Could use same idea for individual lawyers for offices, by aggregating by firms/office/etc, and remove duplicate strikes/trials.
+Thank you for this comment.  We have revised the paper to address it. [INSERT]
 
-[SP: add pararaph/sentence about this]
+Suppose prosecutors from the same office were all exposed to training on how best to strike Black jurors.
+If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the "DA's office").
+Once aggregated, any down-weighting adjustment (choosing $\alpha < 1$) depends on assumptions about how dissimiliarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibilty, notwithstanding the prosecutors' common exposure to the training.
+If none, one could assign equal weight to the historical data ($\alpha$ = 1).
+If one allows for some incompatibilty nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
+Or one could keep all the historical trials; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is values of alpha.
 
-`5.
-The paper uses 90% and 95% credible intervals on p.
+>The paper uses 90% and 95% credible intervals on p.
 7, 95% in Tables 2 and 3 and 80% on p.
@@
 
-* This is not correct. Clarify by focussing on single interval.
+No, this is not correct.
+We have revised the paper to avoid such confusion.
+See Item 5 above.
+We agree that the choice of credible interval should be consistent within different stages of the _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
 
-`Gastwirth (2005) in Law, Prob.
+>Gastwirth (2005) in Law, Prob.
 Risk, p.
@@
 This connection should be noted.
-Also, if one had data from an actual case, one could show the additional insight one obtains by incorporating data from previous cases.`
+Also, if one had data from an actual case, one could show the additional insight one obtains by incorporating data from previous cases.
 
 Thank you for this reference.
-The paper now cites to the model in Gastwirth (2005).
-We also agree about the "additional insight" from "data from an actual case."  The discussion about the prototype software application expressly relies on actual strike data from a sample of trials from the federal district court of Connecticut.
-[MORE].
+The paper now cites to the model in Gastwirth (2005) as well as a suggested statistical approach in Gastwirth (2014).
+On "data from an actual case," the discussion about the prototype software application expressly relies on actual strike data from a sample of trials.
 
-`The choice of $\alpha$ should be discussed and recommendations made.
+>The choice of $\alpha$ should be discussed and recommendations made.
 Otherwise, both parties can explore which value is most helpful.
@@
 
-* Repeat section about alpha/metal detector argument.
+On the choice of $\alpha$, see Items 2 and 3 above.
+At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
+We have revised to provide further guidance on these points.
 
-`The scenario depicted in Table 8 and the related text indicates that both the prosecutor and defense removed minority jurors –both 95% credible intervals do not contain 0.
+>The scenario depicted in Table 8 and the related text indicates that both the prosecutor and defense removed minority jurors –both 95% credible intervals do not contain 0.
 This seems unrealistic.
 More common would be the prosecutor removed more minorities so their 95% credible interval would not contain 0 while the credible interval for the defense removing non-minority potential jurors would be shifted but still might contain 0 (Of course, in some cases when minorities form a substantial portion, e.g.
-40% or so of the venire, the data may indicate both sides were probably biased.`
+40% or so of the venire, the data may indicate both sides were probably biased.
 
-Table 8 = hypothetical data.
-This is just one scenario.
-Nothing about the tool makes any assumptions about what defense and prosecutors do/tend to do when it comes to cognizable/non-cognizable jurors.
-The burden of proof lies with the person bringing the challenge
+Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical data in Table 8 to underscore that, consistent with _Batson_, nothing in our approach (and in the app) makes different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.
 
@@
 
-Fair. Although, these are fake names.
-
-Update online appendix
-* Move figures (80% and 90% CI)
-   * 80 and 90 CI
-   * Fig 4, 6, 8, 9,
-* Move app screenshots and majority of text about app
-* Add sensitivity for defense bias (new)
+We believe the names make it easier to read and follow along while using the app.
+The names are fake, courtesy of the `charlatan` package.
+Please note that, at the time of initial submission, we offered to make the app prototype available to the reviewers (as an R package), so they too could follow along.
+We received no reply.
+The offer remains.

### 2023-01-20 analysis/paper_draft3.Rmd
minor edits; add ref to Carvalho2021

@@
 The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data.
-A commonly adopted practice is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho & Ibrahim].
+A commonly adopted practice is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho2021, 5252; @Ibrahim2015, 3734].
 Specifically, lower values (closer to 0) of $\alpha$ should be selected if incompatibility is expected to be high.
@@
 We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.
-charge severity) affects $b_{curr}$ but not of all the past trials in the historical strike data.
+defendant race, charge severity) affects $b_{curr}$ but not of all the past trials in the historical strike data.
 
@@
 For each replicate, we generate a MCMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
-For each replicate, we calculate the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.
+For each replicate, we calculated the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.
 
@@
 The data also did not expressly indicate the race or gender of the criminal defendant(s).
+For the levels of race and gender, we were limited by choices in the juror questionnaire (e.g., for gender, "male" and "female").
 
@@
 As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b \sim N(0,2)$).
-Historical data will be included if the name(s) of a prosection or defense attorney are selected from the pull-down menu.
+Historical data will be included if the app user selects the name(s) of a prosection or defense attorney from the pull-down menu.
 Subsequently, the amount of historical data to be borrowed can be selected by the choice of $\alpha$.
@@
 Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
-These statistical methods do not aim to  _prove_ bias alone, but rather to lend support to an inference of bias and a subsequent _Batson_ challenge.
-Previously used methods have either ignored historical trial data, or conducted analyses where current and historical data are merged, with no regard for possible incompatibility between trials.
+These statistical methods do not aim to  prove illegal bias alone, but rather to lend support to an inference of bias and a subsequent _Batson_ challenge.
 The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
@@
 
-Moreover, with a tool like our prototype app, attorneys can use the approach presented here to better detect strike bias for _Batson_ and related challenges.
+Moreover, with a tool like our prototype app, attorneys can use the approach presented here to better detect strike bias for purposes of bringing or satisfying proof burden at different stages of _Batson_ and related challenges.
 The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether to make a  _Batson_ challenge at all.
-If argued, the attorney can introduce that posterior distribution as relevant evidence for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+If argued, the attorney can introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
 
@@
 
-US law currently affords a way to ameliorate this missing data problem.
+US law currently affords a way to reduce this missing data problem.
 The public has a presumptive First Amendment right to attend jury voir dire in criminal cases, even if prosecutor and defendant agree to exclude the public.
@@
 Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias estimation is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds illegal bias, because no approach to estimating bias based on historical strike data has that ability.
+In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds that illegal bias is more likely than not (e.g., a _Batson_ violation), because no approach to estimating bias based on historical strike data has that ability.
 

### 2023-01-20 analysis/TAS_memo/memo_TAS_reviewers.Rmd
minor edits

@@
 
-[INSERT]
+[We see expository burden but . . .. ]
 
@@
 If one allows for some incompatibilty nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
-Or one could keep all the historical trials; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is values of alpha.
+Or one could keep all the historical trials; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to values of alpha.
 

### 2023-01-26 analysis/TAS_memo/memo_TAS_reviewers.Rmd
added table with defense bias

@@
 
-[INSERT TABLE]
+```{r inpts}
+df <- data.frame( `Bias Pros` = c(1.5, 1.5, 1.5, 3.0, 3.0, 3.0, 1.5, 1.5, 1.5),
+                  `Bias Def`= c(1.5, 1.5, 1.5, 3.0, 3.0, 3.0, 0.0, 0.0, 0.0),
+                  `alpha`= c(0.1, 0.5, 1.0, 0.1, 0.5, 1.0, 0.1, 0.5, 1.0),
+                  `Bias Det. Pros (95% CI)`= c(0.80, 0.99, 0.99, 0.99, 1.00, 1.00, 0.87, 1.00, 1.00))
+knitr::kable(df, caption ="Simulation results showing effect of adjusting bias in defense on bias detection for the prosecutor.
+Changing the level of bias in defense does not alter bias detection rate for defense.", format = "latex", booktabs = TRUE, align = "c")
+```
 

### 2023-01-27 analysis/paper_draft3.Rmd
small edits to intro and discussion

@@
 The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data.
-A commonly adopted practice is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho2021, 5252; @Ibrahim2015, 3734].
+One approach is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho2021, 5252; @Ibrahim2015, 3734].
 Specifically, lower values (closer to 0) of $\alpha$ should be selected if incompatibility is expected to be high.
@@
 
-In future work we hope to examine data-driven approach to selecting $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
-Knowing how to quantify incompatibility and how to specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022]
+In future work we hope to examine data-driven approaches to selecting $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
+Knowing how to quantify incompatibility and how to specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022].
+In any case, any formulation for estimating $\alpha$ based on incompabitiliy between current and historical trials would require justification in court (or by practitioners).
 

### 2023-01-27 analysis/paper_draft3.Rmd
minor edits, insert cross-reference to first mention of Fig. 1

@@
 
-A _Batson_ challenge typically proceeds in a four-step sequence (Figure 1).
+A _Batson_ challenge typically proceeds in a four-step sequence (Figure \@ref(fig:batsonworkflowfig)).
 First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
@@
 Relying on historical strike data, however, requires an approach to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
-As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and based on historical strike data for that attorney.
+As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and such values based on historical strike data for that attorney.
 

### 2023-01-31 analysis/TAS_memo/memo_TAS_reviewers.Rmd
add information from Xiaomeng

@@
 The tradeoff is that, given compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
-The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $alpha$.
+The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $\alpha$.
 
@@
 
-[INSERT] Will get Eric to check this (got a response from Xiaomeng - waiting for follow-up)
+This information has been added to the text:
+
+A normal distribution was assumed for the distribution of b.
+The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in.
+No thinning was performed, as the correlation was weak and the convergence occured rapidly.
+Convergence was checked from both traceplots and the Gelman-Rubin statistic.
+Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (current data only _versus_ current and historical data).
 

### 2023-01-31 analysis/paper_draft3.Rmd
restructure incompatibility piece in intro

@@
 
-Relying on historical strike data, however, requires an approach to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
-As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and such values based on historical strike data for that attorney.
+Relying on historical strike data, however, requires an approach to  dealing with incompatibility between current and historical trail data.
+Such incompatibility might arise due to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
+As a result, incompatibility might cause an attorney's bias parameter based on strike data for an attorney in the current trial to differ from such values based on historical strike data for that attorney.
 
@@
 Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
+As in the simulations, a normal distribution was assumed for the distribution of $b$.
+The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in.
+No thinning was performed, as the correlation was weak and convergence occured rapidly.
+Convergence was checked from both traceplots and the Gelman-Rubin statistic [@Gelman1992].
+Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current data only _versus_ current plus historical data).
+
 To show how lawyers might use this app in real cases in real time during selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.

### 2023-01-31 analysis/paper_draft3.Rmd
edits to introduction section

@@
 For example, applying the CMH test requires first selecting only historical strike data from past trials that are "similar" to the current trial [@Gastwirth2014, 289, 297].
-Assessing such similarity implies weighting the available historical strike data with respect to some attorney-bias-related set of trial-level characteristics that are present or absent in the current trial.
+Assessing such similarity implies selecting the available historical strike data with respect to some attorney-bias-related set of trial-level characteristics that are present or absent in the current trial with the goal of reducing the incompatibility between current and historical data.
 

### 2023-02-06 analysis/paper_draft3.Rmd
updated introduction, fixed @Bennett reference

--- 2023-01-31 analysis/paper_draft3.Rmd
@@
 In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
-Using simulations, we show how the choice of alpha influences the estimation and detection of bias in the face of varying degrees of incompatibility between current and historical trial data.
-Finally, we extend this approach with a prototype software application (R Shiny application) to detect bias in peremptory-challenge in real time during jury-selection.
-We illustrate this application's use with real, historical strike data from a convenience sample of cases from one court.
+Using simulations, we show how the choice of the power prior's discounting parameter influences the bias detection (whether the credible interval for the bias parameter excludes zero), depending on the degree of incompatibility between current and historical trial data.
+Finally, we extend this approach with a prototype software application that lawyers could use to detect strike bias in real time during jury-selection.
+We illustrate this application's use with real historical strike data from a convenience sample of cases from one court.
 
@@
 
-A _Batson_ challenge typically proceeds in a four-step sequence (Figure \@ref(fig:batsonworkflowfig)).
+A _Batson_ challenge typically proceeds in a four-step sequence.
 First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
@@
 Superior Court Rule 5-12(d)].
-Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].
+Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@Williams2017].
 
@@
 
-In turn, the inference of bias based in part on historical strike data is relevant to lawyers for deciding whether to argue a _Batson_ violation, proving a prima facie case, and ultimately persuading the trial judge that bias in the particular case is more likely than not (Figure \@ref(fig:batsonworkflowfig)).
-Ignoring historical strike data increases the risk of underpowered or inflated estimates of attorney strike bias, because of the low number of strikes per trial (typically 6, 10, or 15 strikes per party).
+Without historical strike data, there is a higher risk of underpowered or inflated estimates of attorney strike bias, because of the low number of strikes per trial (typically 6, 10, or 15 strikes per party, depending on the jurisdiction) [@nscs-jury2018].
+Yet, completely pooling available current and past strike data ignores possible incompatibility between current and historical strike data.
+Such incompatibility might arise due to missingness in the historical data or because an attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
 
-Relying on historical strike data, however, requires an approach to  dealing with incompatibility between current and historical trail data.
-Such incompatibility might arise due to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
 As a result, incompatibility might cause an attorney's bias parameter based on strike data for an attorney in the current trial to differ from such values based on historical strike data for that attorney.
-
-Simply pooling the the strike data from current and past trials, and analyzing such data together, ignores such incompatibility, which can lead to inflated type I error and reduced power [@Bennett2020].
-To increase power, one may use other methods, such as stratifying strike data by trial and applying the Cochran-Mantel-Haenszel (CMH) test [@Gastwirth2014, 295-297].
-Such approaches, however, do not directly address incompatibility between current and historical strike data.
-For example, applying the CMH test requires first selecting only historical strike data from past trials that are "similar" to the current trial [@Gastwirth2014, 289, 297].
-Assessing such similarity implies weighting the available historical strike data with respect to some attorney-bias-related set of trial-level characteristics that are present or absent in the current trial.
+This can lead to inflated type I error and reduced power [@Bennett2021].
+To increase power, one may use other methods, such as stratifying strike data by trial and applying the Cochran-Mantel-Haenszel (CMH) test [@Gastwirth2014, p.
+295-297].
+Such approaches, however, often adjust for incompatibility implicitly.
+For example, applying the CMH test requires first selecting only historical strike data from the subset of past trials that are "similar" in its bias-salient characteristics to the current trial [@Gastwirth2014, p.
+289, 297], a step to reduce incompatibility due to variation in those salient characteristics.
 
@@
 ```
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows for transparently adjusting for the assumed degree of incompatibility between current and historical trials.
+For this approach, we proceed in three steps.
+First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
+Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
+In so doing, we use the power prior [@Chen2000; @Ibrahim2015], which raises the likelihood of the historical strike data to a fixed power ($\alpha_{0}$) between zero and one, to weight the historical strike data in accord with expected compatibility between the historical trials and the current trial.
+Zero denotes no weight to the historical data (complete incompatibility) and one denotes equal weight (complete compatibility).
+Third, we use this posterior distribution as an informative prior for the bias parameter in the current trial.
 
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows us to explicitly and transparently adjust for our assumptions about the degree of incompatibility between current and historical trials.
-In this approach, we use two steps.
-In step one, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
-We use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter.
-In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posterior distribution developed with the historical information and the power prior.
+Here, we demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can "detect" attorney strike bias (i.e., whether the credible interval for the bias parameter excludes zero), given different values of the power prior's alpha parameter.
+Then, we present a software prototype that encodes the same approach with actual historical strike data.
+With that data, the prototype shows how attorneys and others can use this approach to detect strike bias in real time during jury selection.
 
-The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data.
-One approach is to fix the power prior parameter $\alpha$ and assess the sensitivity of results to different values [@Carvalho2021, 5252; @Ibrahim2015, 3734].
-Specifically, lower values (closer to 0) of $\alpha$ should be selected if incompatibility is expected to be high.
-Higher values (closer to 1) of the power prior parameter should be selected if incompatibility is expected to be low.
-Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately corresponding to the likely degree of incompatibility between the current and historical strike data in the particular case.
-Since this overall procedure is to estimate bias to support the bringing of a _Batson_ challenge, the senstivity of bias detection to the value of $\alpha$ is used to evaluate sensitivity of the estimation of bias, not to indicate whether a _Batson_ challenge should succeed.
-
-Here, we demonstrate this approach with simulation studies where we vary the degree of incompatibility between current and historical trial data, and test our ability to detect bias using different values of alpha.
-Then, we present a software prototype that encodes the same approach with actual historical strike data.
-With that data, the prototype shows how attorneys and other can this approach to help them decide to argue, and ultimately prove, illegal bias in peremptory-challenge use in real time during jury selection.
+To be sure, neither this nor any other statistical approach can automatically detect _Batson_ violations, because _Batson_ law permits any evidence of strike bias, not just the attorney's strikes in past cases.
+And it requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the _Batson_ violation).
+Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys "detect" strike bias, given current and historical strike data, the model, and a fixed value for the power prior's $\alpha$ parameter.
+After testing for sensitivity of bias detection to different fixed $\alpha$ values [@Carvalho2021, p.
+5252; @Ibrahim2015, p.
+3734], attorneys can combine the resulting inference with other evidence to decide whether to argue _Batson_; to prove the _Batson_ prima facie case; and to persuade the trial judge to find  a _Batson_ violation (Figure \@ref(fig:batsonworkflowfig)).
 
@@
 ```{r inpts}
+
 df <- data.frame(round = c(1:10),

### 2023-02-07 analysis/paper_draft3.Rmd
draft with updated Discussion section

@@
   - \usepackage{amsmath}
+  - \usepackage{subfig}
 #fontsize: 12pt
@@
 
-In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases [@flowers2019, 2243; Wash.
+In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, 2243; Wash.
 General Rule 37(g)(v); Calif.
@@
 
-Without historical strike data, there is a higher risk of underpowered or inflated estimates of attorney strike bias, because of the low number of strikes per trial (typically 6, 10, or 15 strikes per party, depending on the jurisdiction) [@nscs-jury2018].
-Yet, completely pooling available current and past strike data ignores possible incompatibility between current and historical strike data.
-Such incompatibility might arise due to missingness in the historical data or because an attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity).
+Without historical strike data, there is a higher risk of underpowered or inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
+Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
+Such incompatibility may occur if missing historical strike is missing not at random due to incomplete or inaccessible jury-selection records [@Grosso2017; @Wright2018].
 
-As a result, incompatibility might cause an attorney's bias parameter based on strike data for an attorney in the current trial to differ from such values based on historical strike data for that attorney.
-This can lead to inflated type I error and reduced power [@Bennett2021].
-To increase power, one may use other methods, such as stratifying strike data by trial and applying the Cochran-Mantel-Haenszel (CMH) test [@Gastwirth2014, p.
-295-297].
-Such approaches, however, often adjust for incompatibility implicitly.
-For example, applying the CMH test requires first selecting only historical strike data from the subset of past trials that are "similar" in its bias-salient characteristics to the current trial [@Gastwirth2014, p.
-289, 297], a step to reduce incompatibility due to variation in those salient characteristics.
+Incompatibility may also occur if attorney strike bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
+For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
+If so, then inferring strike bias in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
+
+As a result, for any given estimate of strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power [@Bennett2021].
+Other methods to increase power may in turn require adjusting for incompatibility implicitly.
+For example, @Gastwirth2014 [p.
+295-297] stratify strike data by trial and apply the Cochran-Mantel-Haenszel test, albeit after first selecting only historical strike data from past trials that are "similar" in their bias-salient characteristics to the current trial [p.
+289, 297].
+This initial step presumably aims to reduce incompatibility due to variation in those salient characteristics.
+
+In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows for transparently adjusting for the assumed degree of incompatibility between current and historical trials.
+We proceed in three steps.
+First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
+Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
+In so doing, we use the power prior [@Chen2000; @Ibrahim2015], which raises the likelihood of the historical strike data to a fixed power ($\alpha$), typically between zero and one, to down-weight the historical strike data in accord with expected incompatibility between the historical trials and the current trial.
+Zero denotes no weight to the historical data (for complete incompatibility) and one denotes equal weight (for complete compatibility).
+Third, we use this posterior distribution as an informative prior for the bias parameter in the current trial.
+
+We demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can "detect" attorney strike bias, i.e., whether the credible interval for the bias parameter excludes zero, given different values of the power prior's alpha parameter.
+Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court.
+With that data, the prototype shows how attorneys and others can use this approach to detect strike bias in real time during jury selection.
 
@@
 ```
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows for transparently adjusting for the assumed degree of incompatibility between current and historical trials.
-For this approach, we proceed in three steps.
-First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
-Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the power prior [@Chen2000; @Ibrahim2015], which raises the likelihood of the historical strike data to a fixed power ($\alpha_{0}$) between zero and one, to weight the historical strike data in accord with expected compatibility between the historical trials and the current trial.
-Zero denotes no weight to the historical data (complete incompatibility) and one denotes equal weight (complete compatibility).
-Third, we use this posterior distribution as an informative prior for the bias parameter in the current trial.
-
-Here, we demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can "detect" attorney strike bias (i.e., whether the credible interval for the bias parameter excludes zero), given different values of the power prior's alpha parameter.
-Then, we present a software prototype that encodes the same approach with actual historical strike data.
-With that data, the prototype shows how attorneys and others can use this approach to detect strike bias in real time during jury selection.
-
-To be sure, neither this nor any other statistical approach can automatically detect _Batson_ violations, because _Batson_ law permits any evidence of strike bias, not just the attorney's strikes in past cases.
-And it requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the _Batson_ violation).
-Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys "detect" strike bias, given current and historical strike data, the model, and a fixed value for the power prior's $\alpha$ parameter.
-After testing for sensitivity of bias detection to different fixed $\alpha$ values [@Carvalho2021, p.
+To be sure, neither this nor any other statistical approach can automatically detect _Batson_ violations, because _Batson_ law permits any evidence of strike bias, not just strikes in past cases.
+And _Batson_ law  requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the _Batson_ violation).
+Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, the model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
+After testing for sensitivity of bias detection to different fixed $\alpha$ values for the power prior [@Carvalho2021, p.
 5252; @Ibrahim2015, p.
-3734], attorneys can combine the resulting inference with other evidence to decide whether to argue _Batson_; to prove the _Batson_ prima facie case; and to persuade the trial judge to find  a _Batson_ violation (Figure \@ref(fig:batsonworkflowfig)).
+3734], attorneys can then combine the resulting inference with other evidence to decide whether to argue _Batson_; to prove the _Batson_ prima facie case; and to persuade the trial judge to find  a _Batson_ violation (Figure \@ref(fig:batsonworkflowfig)).
 
@@
 
-Following others [e.g., @Kadane2018a; @Barrett2007; @Gastwirth2005, 183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
-Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptories.
+Following others [e.g., @Kadane2018a; @Barrett2007; @Gastwirth2005, p.
+183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
+Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptory strikes.
 Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that corresponds to the number seats on the jury (plus alternates, if any).
@@
 To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
-<!--To measure convergence, we used trace plots, the Gelman-Rubin convergence diagnostic and Geweke’s diagnostics [@Gelman1992;@Gelman1998;@Geweke1992].--> These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.
+These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.
 As introduced above, $b > 0$ represents a prosecutor's bias against a prospective juror within the cognizable class, while $b = 0$ denotes no bias.
@@
 Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate.
-As a result, the model we use will only detect severe bias.
+As a result, our model will only detect severe bias.
 This sensitivity, however, reduces after including the historical data, especially when $\alpha$ is close to 1.
@@
 Convergence was checked from both traceplots and the Gelman-Rubin statistic [@Gelman1992].
-Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current data only _versus_ current plus historical data).
+Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current data only _versus_ current plus historical data).
 
-To show how lawyers might use this app in real cases in real time during selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
+To show how lawyers might use this app in real cases in real time during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later de-identified them.
@@
 After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.
-Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \@ref(fig:figapp1)).
+Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \@ref(fig:figapp1)(a)).
 Here, the 80\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally.
 
-```{r figapp1, out.width='75%', fig.cap = "Screenshot of R-Shiny application showing density plots for bias of prosecutor and defense attorney based on data in 'Current strike tally' table.
+```{r figapp1, out.width='50%', fig.cap = "Screenshots of R-Shiny application showing density plots for race bias of prosecutor and defense attorney.
 Vertical dotted lines depict 80 pct.
-credible interval.", fig.align='center'}
+credible interval.", fig.align='center', fig.ncol = 2, fig.show = 'hold', fig.subcap=c("Current Strike Data only","Historical and current strike data")}
 
-load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot1.png")
-knitr::include_graphics(load_path)
+load_path1 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot1.png")
+load_path2 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")
+
+par(mfrow=c(2,1))
+knitr::include_graphics(load_path1)
+knitr::include_graphics(load_path2)
 
@@
 In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to "Equal".
-The density plots update accordingly (Figure \@ref(fig:figapp2)).
+The density plots update accordingly (Figure \@ref(fig:figapp1)(b)).
+Now the credible intervals clearly exclude zero.
+Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
+In the Appendix, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
 
-```{r figapp2, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Race).
-Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.", fig.align='center'}
+```{r figapp2, fig.cap = "Density Plots with Historical Strike Data (Race).
+Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.", fig.align='center', eval = FALSE}
 
-load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")
+load_path2 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")
 
-knitr::include_graphics(load_path)
+knitr::include_graphics(load_path2)
 
@@
 
-Now, the credible intervals clearly exclude zero.
-Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
-In the Appendix, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
-
 # Discussion
 
-Statistical methods for _Batson_ and similar challenges aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership affected an attorney's use of peremptory strikes against them.
-These statistical methods do not aim to  prove illegal bias alone, but rather to lend support to an inference of bias and a subsequent _Batson_ challenge.
+Statistical methods for _Batson_ and similar challenges often use historical strike data to infer illegal strike bias.
 The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, one can assess how sensitive that posterior estimate is to the available historical strike data.
-This is appealing because it makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
+By incorporating the power prior, this approach makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
+Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them  decide whether to make a  _Batson_ challenge at all.
+If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and for ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
 
-Moreover, with a tool like our prototype app, attorneys can use the approach presented here to better detect strike bias for purposes of bringing or satisfying proof burden at different stages of _Batson_ and related challenges.
-The posterior distribution of the striking attorney's bias parameter can help the lawyer decide in real time whether to make a  _Batson_ challenge at all.
-If argued, the attorney can introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and of ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+## Selecting the Power Prior Value
 
-To be sure, such evidence alone cannot account for the veracity of the striking attorney's proferred reasons.
-Besides, _Batson_ procedure allows any type of evidence, not just statistical evidence, for ultimately proving or disproving illegal bias in the particular case.
-Thus, as with any use of historical strike data, the approach here cannot always determine whether a _Batson_ challenge should succeed or fail.
-The approach here also requires more from those who would collect the appropriate historical strike data.
-For example, given the model we used, we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
-
-Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \@ref(simul-strikes) ).
-In other cases, however, one cannot apply the same model to courts with different jury selection practices.
-For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
-
-Furthermore, the bias detection rate depends on the degree of incompatibility between the bias parameters for the current trial and for the past (historical) trials ($\lvert b_{curr} - b_{hist} \rvert$).
-These bias parameters may be incompatible ($b_{curr} \neq b_{hist}$) for at least two reasons.
-
-First, such incompatibility may occur because of missing historical data.
-Historical strike data tends to be incomplete at best, because jury-selection records in past cases are hard to get due to court practices and other barriers [@Grosso2017; @Wright2018].
-For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender).
-And judges and court staff vary in how often they fully complete or maintain the information they are required to record.
-This presents an obstacle for any approach to detecting strike bias, including but not limited to the one here, if the missing data is missing not at random.
-
-US law currently affords a way to reduce this missing data problem.
-The public has a presumptive First Amendment right to attend jury voir dire in criminal cases, even if prosecutor and defendant agree to exclude the public.
-Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from the voir dire of prospective jurors [@press1984; @presley2010].
-If these rights are asserted, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time.
-Thus, though historical strike data is hard to collect retrospectively, such courtroom observers can prospectively build up a dataset for strike data that they would have been less likely to obtain from court records after the fact.
-
-Second, even with complete strike data (no missingness), incompatibility may occur if attorney bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
-For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
-If so, then bias detection in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
-
-In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select.
+In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select for the power prior.
 Given perfect compatibility ($\lvert b_{curr} - b_{hist} \rvert = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
@@
 Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
+But our simulations also show that, given high incompatibility, bias detection can be erroneously high when current bias ($b_{curr}$) is low; and bias detection can be erroneously low where current bias is high but bias in historical trials ($b_{hist}$) is low.
 
-In cases where lawyers contest the degree of compatibility, those lawyers can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-(Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.) If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then worries about incompatibility are warranted.
-As a result, the historical strike data, without more information, helps less to prove a _Batson_ violation than it would otherwise would.
+Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
+Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.
+If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then one should worry more about incompatibility and choose a lower $\alpha$ value accordingly.
 
-By their choice of $\alpha$, the lawyer explicitly and transparently denotes the degree of weight placed on the historical strike data in the analysis.
-This transparency matters, because it thwarts lawyer efforts to strategically select $\alpha$ values to prove or disprove illegal strike bias.
-Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides of a _Batson_ challenge know that the lawyers on the other side can always directly and repeatedly assess how sensitive any bias estimation is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot always determine whether a trial judge finds that illegal bias is more likely than not (e.g., a _Batson_ violation), because no approach to estimating bias based on historical strike data has that ability.
+This approach also applies when the unit of analysis is not the attorney, but the law firm or office.
+To illustrate, suppose prosecutors from the same office were all exposed to training that recommended striking Black jurors.
+If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the "district attorney's office").
+Thereafter, any down-weighting adjustment (choosing $\alpha < 1$) depends on assumptions about how dissimilarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibilty, notwithstanding the prosecutors' common exposure to the training.
+If none, one could assign equal weight to the historical data ($\alpha$ = 1).
+If one allows for some incompatibilty nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
+Or one could keep all the historical data; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to different $\alpha$ values.
 
-In future work we hope to examine data-driven approaches to selecting $\alpha$; where the degree of incompatibility can be estimated and used to inform the choice of the degree of discounting to apply to historical data.
-Knowing how to quantify incompatibility and how to specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022].
-In any case, any formulation for estimating $\alpha$ based on incompabitiliy between current and historical trials would require justification in court (or by practitioners).
+In general, by their choice of $\alpha$, the lawyer explicitly and transparently weights the historical strike data in the analysis.
+This transparency thwarts lawyer efforts to strategically select $\alpha$ values.
+Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of $\alpha$.
+In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation).
+Future work might examine how information-criteria or data-driven approaches for selecting $\alpha$ might apply in this context.
+Knowing how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022].
+
+## Modeling the Strike Procedure
+
+The Bayesian approach here also requires a model to match the strike procedure of the court of interest.
+This choice of model can matter in at least three ways.
+First, it can affect the missingness of historical strike data.
+For example, given the model we used (Equation \@ref(eq:model1)), we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
+Such information may be missing from jury selection records of past cases.
+
+Second, the choice of model can affect prospective data collection not only by courts but also by courtroom observers.
+The public has a presumptive First Amendment right to attend jury voir dire in criminal cases.
+Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from voir dire [@press1984; @presley2010].
+As a result, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time.
+Thus, such courtroom observers can prospectively collect strike data that they would have been less likely to obtain from court records after the fact.
+
+Third, the choice of model matters if historical strike data for a particular attorney covers strikes exercised in multiple courts with different strike procedures.
+In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \@ref(simul-strikes) ).
+In other cases, however, one cannot apply the same model to courts with different jury selection practices.
+For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
 
@@
 
-## CRediT Author Statement
+We thank the Associate Editor and two anonymous reviewers for their comments.
+
+# CRediT Author Statement.
 
@@
 
-Here, we extend the discussion in section \@ref(prototype) of the paper by discussing the data collection and presenting additional illustrations of how a lawyer might use the prototype app.
+Here, we extend the discussion in section \@ref(prototype) by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
 

### 2023-02-08 analysis/paper_draft3.Rmd
update Batson app screenshots, references, Discussion

@@
 Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
-But our simulations also show that, given high incompatibility, bias detection can be erroneously high when current bias ($b_{curr}$) is low; and bias detection can be erroneously low where current bias is high but bias in historical trials ($b_{hist}$) is low.
+But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current bias ($b_{curr}$) is low; and bias detection can be (erroneously) low where current bias is high but bias in historical trials ($b_{hist}$) is low.
 
@@
 Suppose the same scenario discussed in section \@ref(prototype), except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
-If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)).
+If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)(left)).
 
-```{r figapp3, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender).
-Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.
-Equal weight ($a = 1$) applied to historical and current data.", fig.align='center'}
+```{r figapp3, out.width='50%', fig.cap = "Screenshots of R-Shiny application show density plots for gender bias of prosecutor and defense attorney based on historical and current strike data.
+Vertical dotted lines depict 80 pct.
+credible interval.", fig.show = 'hold', fig.ncol = 2, fig.subchap=c("Equal weight","Minimal weight") }
 
-load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")
+load_path3 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")
+load_path4 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot4.png")
 
-knitr::include_graphics(load_path)
+#par(mfrow=c(2,1))
+
+knitr::include_graphics(load_path3)
+knitr::include_graphics(load_path4)
 
@@
 However, because the 80\% credible interval includes zero, inferring such bias from the strike data alone is unjustified.
+Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp3)(right)).
+Both the prior and posterior density become flatter and the credible interval plainly includes zero.
 
@@
 Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.
-Minimal weight ($a = 0.1$) applied to historical data.", fig.align='center'}
+Minimal weight ($a = 0.1$) applied to historical data.", fig.align='center', eval=FALSE}
 
@@
 ```
-
-Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp4)).
-Both the prior and posterior density become flatter and the credible interval plainly includes zero.

### 2023-02-08 analysis/TAS_memo/memo_TAS_reviewers.Rmd
updated TAS memo

--- 2023-01-20 analysis/TAS_memo/memo_TAS_reviewers.Rmd
@@
 from: Authors
-to: Editor, The American Statistican
+to: The American Statistican
 subject: Reply to Reviewer Comments
@@
 
-This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is better as a result.
+This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is much better as a result.
 
@@
 
-Here, we address five key issues raised by multiple reviewer comments.
+Here, we address several key issues raised by multiple reviewers.
 
 1.
-Our statistical approach is not an all-purpose detector of _Batson_ violations.
-No statistical approach can do this, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+As our revised Introduction now makes clear (p.
+4-5 and Fig.
+1), neither our statistical approach nor any other can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
 For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons.
@@
 Rather, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias.
-The phrase "bias detection" in the paper means only that the credible interval for the bias parameter ($b_{curr}$) excludes zero, given the strike data and the model.
+The phrase "bias detection" in the paper means only that the credible interval for the bias parameter excludes zero, given the historical strike data, the model, and choice of $\alpha$ value of the power prior.
 For that "detection" task, our approach lets one assign at most equal weight to past strikes and their strikes in the present case ($\alpha = 1$).
-In turn, attorneys can use the resulting bias estimate, _combined with_ other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find  a _Batson_ violation (i.e., illegal strike bias is more likely than not, given all the evidence).
-We have revised the Introduction to emphasize all this to the reader at the outset.
+In turn, attorneys can use the resulting bias estimate, combined with other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find  a _Batson_ violation (i.e., illegal strike bias is more likely than not, given all the evidence).
 
@@
 Thus, small vs.
-large values of $\alpha$ should vary with the suspected degree of incompatability.
+large values of $\alpha$ should vary with the suspected degree of incompatibility.
 In reality, because $b_{curr}$ and $b_{hist}$ are not directly observable, one will need to rely on background knowledge or assumptions about the degree of incompatibility to decide which value of $\alpha$ to choose.
 In a _Batson_ challenge, the lawyers must disclose and justify their choice of $\alpha$ accordingly.
-We have revised the paper to emphasize this.
+We have revised the paper's Introduction and Discussion (sec 3.1) to emphasize this.
 
@@
 Absent this transparency, lawyers will continue to argue strategically about how much to weigh historical strike data, except with the opacity of words.
-We have revised the paper to add this to the Discussion section.
+We have revised the paper to add this to the Discussion section (sec.
+3.1).
 
 4.
-Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias.
-This is because the proper weight to place on historical data depends on whether, for that historical data, there is any missing data that is missing not at random and whether inferring bias from that historical strike data is complicated by plausible confounding due to trial-level variables (e.g., defendant race, criminal charge severity).
+Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges.
+This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data.
 The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
@@
 
+6.
+When we first submitted the paper, we offered to make the app prototype available to the reviewers (as an R package), so they could follow along with section 2.5 of the paper.
+We received no response to this offer.
+The offer remains.
+
 In the following responses to particular comments, we will refer to the above items when appropriate.
@@
 
->[T]he article doesn’t compellingly make the case that the methodology is more useful than simply tracking those who have been reprimanded in the past or that it’s ethical/reasonable to actually employ in the courtroom.
+_[T]he article doesn’t compellingly make the case that the methodology is more useful than simply tracking those who have been reprimanded in the past or that it’s ethical/reasonable to actually employ in the courtroom.
 Specifically, the way that prior information is encoded, the logic behind a detection appears to be circular in that if someone has been reprimanded in the past they should be reprimanded in the future (in some cases for no reason).
-This issue hasn’t been discussed to a level that makes the utility of such a model clear.
+This issue hasn’t been discussed to a level that makes the utility of such a model clear._
 
@@
 
->Page 4, line 5: The authors refer to gender but then provide sex options (male/female).`
+_Page 4, line 5: The authors refer to gender but then provide sex options (male/female)._
 
-Both the model and the prototype software encode "gender" as binary (male or female) simply because the underlying strike data in the app is based on standard juror questionaires that provide "male" and "female" as the only mutually-exclusive options.
+Both the model and the prototype software encode "gender" as binary (male or female) simply because the underlying strike data in the app is based on standard juror questionnaires that provide "male" and "female" as the only mutually-exclusive options.
 
->Page 4, line 36: Use `\left(\right)` to ensure appropriately sized parentheses where necessary.`
+_Page 4, line 36: Use `\left(\right)` to ensure appropriately sized parentheses where necessary._
 
@@
 
->Page 5, equation 6: The left-hand side-of the equation is not correct`
+_Page 5, equation 6: The left-hand side-of the equation is not correct._
 
@@
 
->Page 6, line 43: The idea behind $\alpha$ is concerning to me.
+_Page 6, line 43: The idea behind $\alpha$ is concerning to me.
 The idea here is that if there is previous problematic strike data, and we choose large $\alpha$, then we’ll detect problematic strikes at higher rates.
@@
 Perhaps the citation reference by Reviewer 2 can help the authors – It’s possible that only considering small values of $\alpha$ would be reasonable.
-The simulation results for alpha = 0.1 are less objectionable.`
+The simulation results for alpha = 0.1 are less objectionable._
 
 It is true, based on our simulations, that bias can be erroneously detected (i.e.
-even if $b_curr$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibilty.
+even if $b_curr$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibility.
 On the other hand, given high compatibility, low values of $\alpha$ increase Type I error.
 This is why we recommend using different values of $\alpha$ to check for sensitivity.
-See Item 2 above.
+See Item 1(2) above.
 
-The use of the word "guilty" is error.
+Part of the concern here may turn on an error about the standard of proof, as evinced by the use of the word "guilty" in this comment.
 _Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more stringent "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the criminal offense charged.
 
->Page 6, line 50: The description of bhist is not clear; it is defined twice with different values.`
+_Page 6, line 50: The description of bhist is not clear; it is defined twice with different values._
 
@@
 
->Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?`
+_Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?_
 
 We revised to remove such confusion.
-On the choice of credible intervals, see Item 5 above.
+On the choice of credible intervals, see Item 1(5) above.
 The revision does not necessarily fix the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), whereas we are interested in "bias detection", i.e.
@@
 
->Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior.
+_Page 17, line 46: The real data analyses side-step the use of prior data and don’t represent the analyses or simulation prior.
 One would be curious whether the approach with no prior works for detecting bias (e.g., simulations results without prior data).
-I think I would find this type of model more compelling.`
+I think I would find this type of model more compelling._
 
@@
 Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2).
+See also Item 1(6).
 
->Page 18, line 45: If there is benefit to the user for finding bias, the ability to select $\alpha$ is akin to p-hacking.`
+_Page 18, line 45: If there is benefit to the user for finding bias, the ability to select $\alpha$ is akin to p-hacking._
 
@@
 The phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify as "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results.
-In contrast, in our approach, the statistical model has to qualify as a valid model of the strike _procedure_ of the court of interest.
+In contrast, in our approach, the statistical model has to qualify as a valid model of the strike procedure of the court of interest.
 That strike procedure can and must be known and modeled before any strike data is collected or analyzed.
 More generally, our approach thwarts the strategic selection of $\alpha$ values by making the choice of $\alpha$ transparent.
-See Item 4, above.
+See Item 1(4) and section 3.2 of the revised Discussion.
 
->Something unmentioned that should be addressed is confounders.
+_Something unmentioned that should be addressed is confounders.
 While technological and statistical tools can be helpful to the judicial system, and even decrease instances of discrimination and curb disparities, there are many unintended consequences of such systems.
-For example, race and class are intertwined in the United States and there is no discussion (at least) about incorporating covariates (e.g., age, education, class, etc.)`
+For example, race and class are intertwined in the United States and there is no discussion (at least) about incorporating covariates (e.g., age, education, class, etc.)_`
 
@@
 This is illegal under _Batson_, regardless of the actual education levels of the three jurors struck.
-[FIX].
 
@@
 
-# Reviewer 1
+# Comments of Reviewer 1
 
->In Section 2.3, the paragraph in line 47 on p. 6 . . . I am confused about which set of values was used in the simulation for bhist.
+_In Section 2.3, the paragraph in line 47 on p. 6 . . . I am confused about which set of values was used in the simulation for bhist._
 
@@
 
->About the results presented in Figure 1 of Section 2.3, if we **ignore the assumption of compatibility** between historical and current data, in the cases of $\alpha$ = 0.5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0?
+_About the results presented in Figure 1 of Section 2.3, if we **ignore the assumption of compatibility** between historical and current data, in the cases of $\alpha$ = 0.5 and $\alpha$ = 1, does the bottom right corner in each panel indicate that the bias detection rate can be as high as 100% even when the true current bias is 0?
 If so, the Type I error seems to be very high in some cases.
-Similarly, the power (bias detection rate) seems low at the top left corner in each panel where bias is low in historical data but high in current data.
+Similarly, the power (bias detection rate) seems low at the top left corner in each panel where bias is low in historical data but high in current data._
 
@@
 The tradeoff is that, given compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
-The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $alpha$.
+The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $\alpha$.
 
->Regarding the Shiny App in Section 2.5, it would be helpful to include some details about the Metropolis-Hastings scheme.
+_Regarding the Shiny App in Section 2.5, it would be helpful to include some details about the Metropolis-Hastings scheme.
 For instance, is a normal distribution used as the proposal distribution for b?
@@
 Is any convergence criterion (e.g., the Gelman-Rubin statistic) checked?
-What is the computation time?
+What is the computation time?_
 
-[INSERT] Will get Eric to check this (got a response from Xiaomeng - waiting for follow-up)
+This information has been added to the text (sec. 2.5, p. 18).
 
->In Section 3, the authors state that "the anticipated degree of incompatibility should influence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error.
+<!-- A normal distribution was assumed for the distribution of b.
+The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in.
+No thinning was performed, as the correlation was weak and the convergence occurred rapidly.
+Convergence was checked from both traceplots and the Gelman-Rubin statistic.
+Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (current data only _versus_ current and historical data).-->
+
+_In Section 3, the authors state that "the anticipated degree of incompatibility should influence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error.
 However, based on results in Figures 1 to 3, if this assumption is wrong, choosing a large $\alpha$ leads to either low power (bias detection rate) or large Type I error, depending on the bias in historical data.
 In the absence of such knowledge of compatibility, how does one choose the value of $\alpha$?
-In addition, to get around this issue, is it reasonable to treat $\alpha$ itself random and introduce a hyper prior?
+In addition, to get around this issue, is it reasonable to treat $\alpha$ itself random and introduce a hyper prior?_
 
-See Items 2 and 4 above.
+See Items 1(2) and 1(4) above and section 3.1. of the revised Discussion.
 
->Some of the Figure and table captions (e.g. Figures 1-5 and Table 1) can benefit from being more descriptive.
+_Some of the Figure and table captions (e.g. Figures 1-5 and Table 1) can benefit from being more descriptive._
 
@@
 
->A couple of notation issues: . . . .
+_A couple of notation issues: . . . ._
 
@@
 
-# Reviewer 2
+# Comments of Reviewer 2
 
->The method is first illustrated in a simulation study, with a range of values of $\alpha$.
+_The method is first illustrated in a simulation study, with a range of values of $\alpha$.
 This is fine for an audience of statisticians and attorneys specializing in jury discrimination but is somewhat overwhelming for a general reader.
 It would be easier for the reader if one data set, perhaps from the actual data from Connecticut was used and the reader could see how the data from each additional previous trial impacts the posterior.
-Perhaps only a few values of $\alpha$, rather than 11 and two val[u]es of the bias parameter b, could be used.
+Perhaps only a few values of $\alpha$, rather than 11 and two val[u]es of the bias parameter b, could be used._
 
-[We see expository burden but . . .. ]
+We appreciate this suggestion but we are worried that this would unduly increase the paper's length. Because of this concern, we have already moved other parts of the paper to the (online) Appendix.
 
->From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones.
+_From the likelihood on page 4, it is apparent that the method requires data on the consecutive strike patterns, not only in the current case, but in previous ones.
 How often is this data currently available?
-If it isn’t readily available now, how likely are states and locales to implement a system of systematically reporting and preserving the required data (especially in the areas where one expects minorities to encounter bias in jury membership)?
+If it isn’t readily available now, how likely are states and locales to implement a system of systematically reporting and preserving the required data (especially in the areas where one expects minorities to encounter bias in jury membership)?_
 
-In the Discussion, we explained that data on consecutive strike patterns is hard to collect, simply because of the challenges of collecting any historical strike data, regardless of strike procedure.
-We have revised that Discussion to add that, since the 1980s, any person has the right to attend jury selection in criminal court proceedings These legal rights enable data collection by court observers, regardless of State and local recordkeeping practices.
-Finally, our statistical approach's validity does not turn on the _current_ availability of the kind of data it requires.
-Indeed, we hope our statistical approach can inform efforts as to what kinds of data about strikes a court system should make more available.
+Our statistical approach's validity turns neither on how many courts make available the kind of data it requires nor on proving how likely courts will start collecting such data.
+Indeed, as section 3.2 (paragraphs 1 and 2) of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right at attend jury selection in criminal court proceedings.
 
->In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias.
+_In Section 2.3, it is noted that in the simulations, it is assumed that the defense attorney has no bias.
 If one has been called for jury service in a city with a large minority population and been in a jury room, this is highly implausible.
@@
 On the other hand, if the method detects prosecutorial bias when the defense challenges appear biased too, this should strengthen the evidence of prosecutorial bias.
-This issue deserves some quantitative investigation, so judges can feel confident when using the results.`
+This issue deserves some quantitative investigation, so judges can feel confident when using the results._
 
-[INSERT TABLE]
+We assumed no strike bias for the defense attorney only to simplify exposition of the simulation results.
+Our model of the strike procedure (Equation 1) assumes that the probability of prosecutor strike bias is independent of the probability of defense attorney strike bias.
+And that model explicitly accounts for the racial composition of the pool of prospective jurors at each strike opportunity.
 
->Page 5, equation 5: The left-hand side-of the equation is not correct
+```{r inpts, echo=FALSE}
 
-We have revised to indicate that it is a posterior distribution. All formulas have now been updated and checked. We now specify the vector notation for $\delta_0$ and $\delta_1$.
+df <- data.frame( `Bias Pros` = c(1.5, 1.5, 1.5, 3.0, 3.0, 3.0, 1.5, 1.5, 1.5),
+                  `Bias Def`= c(1.5, 1.5, 1.5, 3.0, 3.0, 3.0, 0.0, 0.0, 0.0),
+                  `alpha`= c(0.1, 0.5, 1.0, 0.1, 0.5, 1.0, 0.1, 0.5, 1.0),
+                  `Bias Det. Pros (95% CI)`= c(0.80, 0.99, 0.99, 0.99, 1.00, 1.00, 0.87, 1.00, 1.00))
 
->Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?
+knitr::kable(df, caption ="Simulation results showing effect of adjusting bias in defense on bias detection for the prosecutor.
+Changing the level of bias in defense does not alter bias detection rate for prosecutor.", format = "latex", booktabs = TRUE, align = "c")
 
-See Item 5 above.
+```
 
->The paper and software focus on peremptory challenges made by the same lawyer in previous cases.
+In any case, we re-ran our simulation and found that, as expected, changing the level of defense attorney bias does not alter the bias detection rates for prosecutors (Table 1).
+
+_Page 5, equation 5: The left-hand side-of the equation is not correct._
+
+We have revised the paper to indicate that it is a posterior distribution. All formulas have now been updated and checked. We now specify the vector notation for $\delta_0$ and $\delta_1$.
+
+_Page 9, line 15: The discussion about selecting credible interval percentages is confusing. Are you saying that larger credible interval percentages lead to less accurate bias detection?_
+
+See Item 1(5) above.
+
+_The paper and software focus on peremptory challenges made by the same lawyer in previous cases.
 As noted in the Miller-El cases, the training manual used by prosecutors recommended the order in which groups, e.g.
@@
 Therefore, data from prosecutors from the same DA’s office or defense firm in similar cases (drug dealing, robbery etc.), should be quite informative.
-Should one simply add them to the list of prior cases, or make a down-weighting adjustment in the Bayesian analysis?
+Should one simply add them to the list of prior cases, or make a down-weighting adjustment in the Bayesian analysis?_
 
-Thank you for this comment.  We have revised the paper to address it. [INSERT]
+Thank you for this comment!  The paper now addresses this issue directly in the Discussion, section 3.1, paragraph 4.
 
-Suppose prosecutors from the same office were all exposed to training on how best to strike Black jurors.
-If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the "DA's office").
-Once aggregated, any down-weighting adjustment (choosing $\alpha < 1$) depends on assumptions about how dissimiliarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibilty, notwithstanding the prosecutors' common exposure to the training.
-If none, one could assign equal weight to the historical data ($\alpha$ = 1).
-If one allows for some incompatibilty nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
-Or one could keep all the historical trials; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to values of alpha.
-
->The paper uses 90% and 95% credible intervals on p.
+_The paper uses 90% and 95% credible intervals on p.
 7, 95% in Tables 2 and 3 and 80% on p.
@@
 If possible, the standards for deciding a Batson claim used by judges in a jurisdiction should be the same.
-This is especially true in the present context where the prior data considered in cases in the same jurisdiction is likely to have a substantial overlap.`
+This is especially true in the present context where the prior data considered in cases in the same jurisdiction is likely to have a substantial overlap._
 
@@
 We have revised the paper to avoid such confusion.
-See Item 5 above.
+See Item 1(5) above.
 We agree that the choice of credible interval should be consistent within different stages of the _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
+[NOTE: Simulations use 95% credible interval, while app uses 80% credible intervals.]
 
->Gastwirth (2005) in Law, Prob.
-Risk, p.
-179-185, .
-.
-.
-had proposed that approach when potential jurors were removed one at a time in an alternating sequence.
-Checking that reference, the reviewer saw that at the end of the 2005
-paper, the use of data from prior cases would be helpful and recommended that procedures, such as the one proposed here.
-This connection should be noted.
-Also, if one had data from an actual case, one could show the additional insight one obtains by incorporating data from previous cases.
+_Gastwirth (2005) . . . should be noted. Also, if one had data from an actual case, one could show the additional insight one obtains by incorporating data from previous cases._
 
 Thank you for this reference.
-The paper now cites to the model in Gastwirth (2005) as well as a suggested statistical approach in Gastwirth (2014).
-On "data from an actual case," the discussion about the prototype software application expressly relies on actual strike data from a sample of trials.
+The paper now cites to the model in Gastwirth (2005).
+On "data from an actual case," section 2.5 of the paper (our discussion about the prototype software application) expressly relies on actual strike data from a sample of criminal cases.
 
->The choice of $\alpha$ should be discussed and recommendations made.
+_The choice of $\alpha$ should be discussed and recommendations made.
 Otherwise, both parties can explore which value is most helpful.
 Also, how can one ensure that the prior data doesn’t overwhelm the data from the current case, without choosing a ridiculously low value?
-The authors say something about this on page 6 but more guidance to potential users and courts would be helpful.`
+The authors say something about this on page 6 but more guidance to potential users and courts would be helpful._
 
-On the choice of $\alpha$, see Items 2 and 3 above.
-At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
-We have revised to provide further guidance on these points.
+On the choice of $\alpha$, see Items 1(2) and 1(3) above and section 3.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
 
->The scenario depicted in Table 8 and the related text indicates that both the prosecutor and defense removed minority jurors –both 95% credible intervals do not contain 0.
+_The scenario depicted in Table 8 and the related text indicates that both the prosecutor and defense removed minority jurors –both 95% credible intervals do not contain 0.
 This seems unrealistic.
 More common would be the prosecutor removed more minorities so their 95% credible interval would not contain 0 while the credible interval for the defense removing non-minority potential jurors would be shifted but still might contain 0 (Of course, in some cases when minorities form a substantial portion, e.g.
-40% or so of the venire, the data may indicate both sides were probably biased.
+40% or so of the venire, the data may indicate both sides were probably biased._
 
-Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical data in Table 8 to underscore that, consistent with _Batson_, nothing in our approach (and in the app) makes different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.
+Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical data in Table 8, however "unrealistic", to underscore that, as  _Batson_ law requires, our approach makes no different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.
 
-> At the bottom of p. 17, you don’t really need names for the attorneys, A and B would suffice.
+_At the bottom of p. 17, you don’t really need names for the attorneys, A and B would suffice._
 
-We believe the names make it easier to read and follow along while using the app.
-The names are fake, courtesy of the `charlatan` package.
-Please note that, at the time of initial submission, we offered to make the app prototype available to the reviewers (as an R package), so they too could follow along.
-We received no reply.
-The offer remains.
+We believe the names make it easier to read and follow along while using the app. The names are fake, courtesy of the `charlatan` package. See also Item 1(6).

### 2023-02-08 analysis/paper_draft3.Rmd
edits to paper discussion of app

@@
 This shows that the order of strikes does not influence the estimate of the bias parameter.
-Accordingly, the simultaneous-strikes process can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
+Accordingly, the simultaneous-strikes procedure can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
 
@@
 As in the simulations, a normal distribution was assumed for the distribution of $b$.
-The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in.
+The length of the Markov chain was 110000 and the first 10,000 iterations were dropped as burn-in.
 No thinning was performed, as the correlation was weak and convergence occured rapidly.
 Convergence was checked from both traceplots and the Gelman-Rubin statistic [@Gelman1992].
-Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current data only _versus_ current plus historical data).
+Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current strike data only _versus_ current plus historical strike data).
 
@@
 
-With the app, the user uses the pull-down menus to select the cognizable class (top left) and enters by hand the strike information in the case before them in the _strike tally_ table (bottom left).
-For demonstration purposes, the prototype app comes pre-loaded with a completed strike tally with values that can be changed and rows that can be added or deleted.
-In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
-
-For the prototype app, we set two cognizable-class options: race and gender.
+With the app, the user first selects the cognizable class.
+For the prototype, we set two cognizable-class options: race and gender.
 This was because the standard juror questionnaire in this court did not ask about other possible cognizable classes.
@@
 
-In the default setting, the pull-down menus for prosecutor and defense are set to "None".
-As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for prosecutor and defense based only on the strike tally data and the initial prior ($b \sim N(0,2)$).
-Historical data will be included if the app user selects the name(s) of a prosection or defense attorney from the pull-down menu.
-Subsequently, the amount of historical data to be borrowed can be selected by the choice of $\alpha$.
+Next, the user enters by hand the strike information in the case before them in the _strike tally_ table (Figure \@ref(fig:figapp1), top left).
+In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
+the user can add or delete rows to the strike tally as needed.
 
@@
 Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \@ref(fig:figapp1)(a)).
-Here, the 80\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally.
+Here, the 80\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally alone.
+In the default setting, the pull-down menus for prosecutor and defense are set to "None".
+As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for this prosecutor and defense attorney based only on the strike tally data and the initial prior ($b \sim N(0,2)$).
 
@@
 
-To use historical strike data, the user selects a name of the prosecutor or defense attorney from the pull-down menus.
+To use historical strike data, the user selects the name of the prosecutor or defense attorney from the pull-down menus.
 If an attorney's name cannot be found, the app has no historical strike data for that attorney.

### 2023-02-08 analysis/paper_draft3.Rmd
edits to abstract

@@
   US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
-Here, we extend a Bayesian approach for detecting such illegal bias.
-Current methods for detecting bias make implicit, but unacknowledged, assumptions about how to incorporate historical strike data into the bias detection process.
-Here, we show how to incorporate historical data on an attorney's use of peremptory strikes in past cases to estimate how likely that attorney has struck prospective jurors in a current trial because of their cognizable-class membership.
+Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how how to incorporate historical data on an attorney's use of peremptory strikes in past cases.
 In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
-Using simulations, we show how the choice of the power prior's discounting parameter influences the bias detection (whether the credible interval for the bias parameter excludes zero), depending on the degree of incompatibility between current and historical trial data.
+Using simulations, we show how the choice of the power prior's discounting parameter influences bias detection (how likely the credible interval for the bias parameter excludes zero), depending on the degree of incompatibility between current and historical trial data.
 Finally, we extend this approach with a prototype software application that lawyers could use to detect strike bias in real time during jury-selection.

### 2023-02-09 analysis/paper_draft3.Rmd
added line about criteria for deriving guide values for alpha

@@
 In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation).
-Future work might examine how information-criteria or data-driven approaches for selecting $\alpha$ might apply in this context.
+Future work might examine how various criteria for deriving a guide value of $\alpha$ for types of models [e.g.
+@Ibrahim2015, 3724-3738] could appropriately apply in this context.
 Knowing how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022].

### 2023-02-09 analysis/paper_draft3.Rmd
add reference to Han et al. (2023).

@@
 In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation).
-Future work might examine how various criteria for deriving a guide value of $\alpha$ for types of models [e.g.
+Future work might examine how various criteria for deriving a guide value of $\alpha$ [e.g.
 @Ibrahim2015, 3724-3738] could appropriately apply in this context.
-Knowing how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g., @Pawel2022].
+Knowing how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.
+@Han2023; @Pawel2022].
 

### 2023-02-16 analysis/paper_draft3.Rmd
small edit to intro

@@
 
-Without historical strike data, there is a higher risk of underpowered or inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
+Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
 Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
@@
 This initial step presumably aims to reduce incompatibility due to variation in those salient characteristics.
+[explain tthat this is something that would need to be disclosed/justified]
 

### 2023-02-16 analysis/paper_draft3.Rmd
changed refers to app credible interval to 95%

--- 2023-02-09 analysis/paper_draft3.Rmd
@@
 Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \@ref(fig:figapp1)(a)).
-Here, the 80\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally alone.
+Here, the 95\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally alone.
 In the default setting, the pull-down menus for prosecutor and defense are set to "None".
@@
 ```{r figapp1, out.width='50%', fig.cap = "Screenshots of R-Shiny application showing density plots for race bias of prosecutor and defense attorney.
-Vertical dotted lines depict 80 pct.
+Vertical dotted lines depict 95 pct.
 credible interval.", fig.align='center', fig.ncol = 2, fig.show = 'hold', fig.subcap=c("Current Strike Data only","Historical and current strike data")}
@@
 ```{r figapp3, out.width='50%', fig.cap = "Screenshots of R-Shiny application show density plots for gender bias of prosecutor and defense attorney based on historical and current strike data.
-Vertical dotted lines depict 80 pct.
+Vertical dotted lines depict 95 pct.
 credible interval.", fig.show = 'hold', fig.ncol = 2, fig.subchap=c("Equal weight","Minimal weight") }
@@
 Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors.
-However, because the 80\% credible interval includes zero, inferring such bias from the strike data alone is unjustified.
+However, because the 95\% credible interval includes zero, inferring such bias from the strike data alone is unjustified.
 Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp3)(right)).

### 2023-02-16 analysis/TAS_memo/memo_TAS_reviewers.Rmd

@@
 
-This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is much better as a result.
+This memorandum replies to comments provided by the TAS associate editor and two reviewers.
+We thank the editor and reviewers for their comments.
+We believe the revised paper is much better as a result.
 
@@
 
+[editme]In any case, other authors analysing _Batson_ challallenge data have suggest other, more implicit forms of weighting cases based on trial-incompatibility.
+As we now mention in the paper [xx], @Gastwirth2014 suggests that trials with 'similar characteristics' be selected before conducting Cochran-Mantel-Haenszel tests.
+This is tantamount to setting the weight of incompatible cases to zero and compatible cases to 1.
+In such an instance, such weighing should be explicit and justified.
+
 _Page 4, line 5: The authors refer to gender but then provide sex options (male/female)._
@@
 Like our model, the app uses the $N(0,2)$ prior for the bias parameter.
-Similarly, the app user can alter the weight the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
+Similarly, the app user can alter the weight of the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
 Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2).

### 2023-02-17 analysis/memo_TAS_reviewers.Rmd
revised TAS memo - more on hyperprior for alpha etc.; updated references, minor edit to paper

--- 2023-02-08 analysis/TAS_memo/memo_TAS_reviewers.Rmd
@@
   pdf_document:
-    citation_package: natbib
+    citation_package: biblatex
     keep_tex: false
     latex_engine: pdflatex
-    template: "svm-latex-memo.tex"
+    template: svm-latex-memo.tex
     number_sections: true
@@
 width: .3
+bibliography: references.bib
 ---
 
-This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is much better as a result.
+This memorandum replies to comments provided by the TAS associate editor and two reviewers.
+We thank the editor and reviewers for their comments.
+We believe the revised paper is much better as a result.
 
@@
 4-5 and Fig.
-1), neither our statistical approach nor any other can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+1), neither our statistical approach (nor any other) can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+
+    a.
 For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons.
@@
 Rather, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias.
+
+    b.
 The phrase "bias detection" in the paper means only that the credible interval for the bias parameter excludes zero, given the historical strike data, the model, and choice of $\alpha$ value of the power prior.
@@
 The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
-As we now note in the Discussion section of the paper, measuring incompatibility and choosing $\alpha$ in the face of incompatibility is an active area of research with no clear solution.
+As we now note in the Discussion section of the paper, the literature on the power prior discusses information criteria for guide values of $\alpha$ as well as treating $\alpha$ as random and assigning a prior to it.
+However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons.
+First, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model), as distinct from the models of strike procedure discussed here.
+Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter.
+Third, whereas some papers treat incompatibility as itself a parameter [e.g., @Ollier2020], the precise approaches rely on strong assumptions that do not obviously make sense in our context.
+Accordingly, we believe the most conservative choice (for now) is to keep $\alpha$ a fixed value and check for sensitivity of results to the choice of $\alpha$ value.
+Proceeding otherwise would require the work of a completely new paper.
 
@@
 
-We have three replies to this concern.
+We have four replies to this concern.
 
@@
 Ct.
-2228, [PIN] (2019).
+2228, 2243 (2019).
 State law variants on _Batson_ are also explicit about this.
@@
 By relying on past strike data, not whether anyone argued _Batson_ in those past cases, we avoid any bias associated with relying only on past court findings of _Batson_ violations.
+
+Fourth, the comment is overinclusive.
+Most work on the use of historical strike data (a) presupposes that _Batson_ law permits prior strikes as relevant evidence of bias in the current trial; presupposes that any statistical approach therein cannot automatically determine _Batson_ violations alone; and (c) adjusts in some way for incompatibility between historical strike data and strikes from the current trial.
+As an example, we now point in the paper's Introduction to @Gastwirth2014.
+Our approach has a distinctive benefit of making incompatibility assumptions transparent via choosing a fixed value for the power prior's $\alpha$.
 
@@
 
-Part of the concern here may turn on an error about the standard of proof, as evinced by the use of the word "guilty" in this comment.
-_Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more stringent "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the criminal offense charged.
+The use of the word "guilty" in this comment implies that part of the concern here may turn on an error about the applicable standard of proof.
+_Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more demanding  "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the elements of a criminal offense.
 
@@
 We revised to remove such confusion.
+This includes changing the software app to use 95% credible intervals.
 On the choice of credible intervals, see Item 1(5) above.
@@
 The real data analyses are based on real historical strike data from real criminal cases in the federal district court in Connecticut.
+On how we collected this data, see the paper's Appendix.
 If they wish, users of the app can run the model excluding any historical data (as this reviewer suggested) by not selecting the name of a prosecutor or defense attorney.
 Like our model, the app uses the $N(0,2)$ prior for the bias parameter.
-Similarly, the app user can alter the weight the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
+Similarly, the app user can alter the weight of the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
 Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2).
@@
 In contrast, in our approach, the statistical model has to qualify as a valid model of the strike procedure of the court of interest.
-That strike procedure can and must be known and modeled before any strike data is collected or analyzed.
+That strike procedure must be known and modeled before any strike data is analyzed.
 More generally, our approach thwarts the strategic selection of $\alpha$ values by making the choice of $\alpha$ transparent.
@@
 
-On confounders, we expressly highlight that variation in trial-level characteristics (e.g.
+On confounders, we expressly highlight (in the Introduction) that variation in trial-level characteristics (e.g.
 defendant race, charge severity) in the historical data can generate incompatibility, which in turn can affect bias detection.
@@
 We do not aim to identify any and all consequences of a full-scale implementation of this approach, intended or unintended.
+Such implementation entails a series of operational choices that are well beyond the paper's scope.
 
@@
 We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_curr$); and (2) bias detection can be low where current bias is high but bias in historical trials is low.
-The tradeoff is that, given compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
+The tradeoff is that, given high compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
 The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $\alpha$.
@@
 
-We appreciate this suggestion but we are worried that this would unduly increase the paper's length. Because of this concern, we have already moved other parts of the paper to the (online) Appendix.
+We appreciate this suggestion but we are worried that this would unduly increase the paper's length, particular given that absence of a closed form solution for the posterior distribution even in a toy example.
+Because of this concern, we have already moved other parts of the paper to the (online) Appendix.
+We intend the reader to gain understanding by following along using our prototype software application.
 
@@
 And that model explicitly accounts for the racial composition of the pool of prospective jurors at each strike opportunity.
+In any case, we re-ran our simulation and found that, as expected, changing the level of defense attorney bias does not alter the bias detection rates for prosecutors (Table 1).
 
@@
 ```
-
-In any case, we re-ran our simulation and found that, as expected, changing the level of defense attorney bias does not alter the bias detection rates for prosecutors (Table 1).
 
@@
 No, this is not correct.
-We have revised the paper to avoid such confusion.
+We have revised the paper and the software application to avoid such confusion.
 See Item 1(5) above.
 We agree that the choice of credible interval should be consistent within different stages of the _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
-[NOTE: Simulations use 95% credible interval, while app uses 80% credible intervals.]
 
@@
 
-On the choice of $\alpha$, see Items 1(2) and 1(3) above and section 3.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
+On the choice of $\alpha$, see Items 1(2)-1(4) above and section 3.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
 
@@
 We believe the names make it easier to read and follow along while using the app. The names are fake, courtesy of the `charlatan` package. See also Item 1(6).
+
+# References

### 2023-02-23 analysis/paper_draft3.Rmd
revised TAS memo - more on hyperprior for alpha etc.; updated references, minor edit to paper
edit to discussion

@@
 
-In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, 2243; Wash.
+In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, p.
+2243; Wash.
 General Rule 37(g)(v); Calif.
@@
 @Ibrahim2015, 3724-3738] could appropriately apply in this context.
-Knowing how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.
+while also considering incompatibility (commeasureability) in selecting $\alpha$ (@Ollier2020).
+Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.
 @Han2023; @Pawel2022].

### 2023-02-23 analysis/memo_TAS_reviewers.Rmd
edits to choice of alpah paragraph

@@
 First, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model), as distinct from the models of strike procedure discussed here.
+In the clinical trial setting, it might also be easier to capture(measure?) trial-level (and patient-level) characteristics that may be responsible for incompatibility.
+Even in these settings, selecting a prior for $\alpha$ is not straightforward.
+@Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate the influence of the historical data, and so one might have to use highly informative priors on the power parameter.
+These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis".
+More recent versions of the power prior are beginning to address this issue, for example using "commensurate power priors" (@Hobbs2011).
 Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter.

### 2023-02-23 analysis/memo_TAS_reviewers.Rmd
small edit

@@
 Even in these settings, selecting a prior for $\alpha$ is not straightforward.
-@Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate the influence of the historical data, and so one might have to use highly informative priors on the power parameter.
+@Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate (more downweighting) the influence of the historical data, and so one might have to use highly informative priors on the power parameter.
 These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis".

### 2023-02-23 analysis/memo_TAS_reviewers.Rmd
fixed minor errors, referred to paper's citation to Viele2014

--- 2023-02-17 analysis/memo_TAS_reviewers.Rmd
@@
 However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons.
-First, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model), as distinct from the models of strike procedure discussed here.
+First, unlike models of strike procedure discussed here, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model).
+In the clinical trial setting, it is easier to measure trial-level (and patient-level) characteristics that may be responsible for incompatibility.
+Even in these settings, selecting a prior for $\alpha$ is not straightforward.
+@Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate (more down-weighting) the influence of the historical data, and so one might have to use highly informative priors on the power parameter.
+These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis".
+More recent versions of the power prior are beginning to address this issue (e.g., using "commensurate power priors" [@Hobbs2011]).
 Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter.
@@
 Our approach does not and cannot automate findings of _Batson_ violations.
-See Item 1, above.
+See Item 1(1) above.
 
@@
 Fourth, the comment is overinclusive.
-Most work on the use of historical strike data (a) presupposes that _Batson_ law permits prior strikes as relevant evidence of bias in the current trial; presupposes that any statistical approach therein cannot automatically determine _Batson_ violations alone; and (c) adjusts in some way for incompatibility between historical strike data and strikes from the current trial.
+Most work on the use of historical strike data (a) presupposes that _Batson_ law permits prior strikes as relevant evidence of bias in the current trial; (b) presupposes that any statistical approach therein cannot automatically determine _Batson_ violations alone; and (c) adjusts in some way for incompatibility between historical strike data and strikes from the current trial.
 As an example, we now point in the paper's Introduction to @Gastwirth2014.
@@
 
-We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_curr$); and (2) bias detection can be low where current bias is high but bias in historical trials is low.
+We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_{curr}$); and (2) bias detection can be low where current bias is high but bias in historical trials is low.
 The tradeoff is that, given high compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
@@
 
-We appreciate this suggestion but we are worried that this would unduly increase the paper's length, particular given that absence of a closed form solution for the posterior distribution even in a toy example.
+We appreciate this suggestion but we are worried that this would unduly increase the paper's length, particularly given the lack of a closed form solution for the posterior distribution even in a toy example.
 Because of this concern, we have already moved other parts of the paper to the (online) Appendix.
 We intend the reader to gain understanding by following along using our prototype software application.
+We also now refer the reader to a gentle introduction to the power prior in @Viele2014.
 

### 2023-02-23 analysis/paper_draft3.Rmd
added ref to Viele2014. minor edits.

@@
 This initial step presumably aims to reduce incompatibility due to variation in those salient characteristics.
-[explain tthat this is something that would need to be disclosed/justified]
+Any such choice of "similar" historical strike data would need to be disclosed and justified.
 
@@
 Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the power prior [@Chen2000; @Ibrahim2015], which raises the likelihood of the historical strike data to a fixed power ($\alpha$), typically between zero and one, to down-weight the historical strike data in accord with expected incompatibility between the historical trials and the current trial.
+In so doing, we use the power prior [@Chen2000; @Ibrahim2015; for a gentle introduction, see @Viele2014].
+The power prior raises the likelihood of the historical strike data to a fixed power ($\alpha$), typically between zero and one, to down-weight the historical strike data in accord with expected incompatibility between the historical trials and the current trial.
 Zero denotes no weight to the historical data (for complete incompatibility) and one denotes equal weight (for complete compatibility).
@@
 The length of the Markov chain was 110000 and the first 10,000 iterations were dropped as burn-in.
-No thinning was performed, as the correlation was weak and convergence occured rapidly.
+No thinning was performed, as the correlation was weak and convergence occurred rapidly.
 Convergence was checked from both traceplots and the Gelman-Rubin statistic [@Gelman1992].
@@
 Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-Our protoype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.
+Our prototype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.
 If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then one should worry more about incompatibility and choose a lower $\alpha$ value accordingly.
@@
 If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the "district attorney's office").
-Thereafter, any down-weighting adjustment (choosing $\alpha < 1$) depends on assumptions about how dissimilarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibilty, notwithstanding the prosecutors' common exposure to the training.
+Thereafter, any down-weighting adjustment (choosing $\alpha < 1$) depends on assumptions about how dissimilarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibility, notwithstanding the prosecutors' common exposure to the training.
 If none, one could assign equal weight to the historical data ($\alpha$ = 1).
-If one allows for some incompatibilty nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
+If one allows for some incompatibility nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
 Or one could keep all the historical data; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to different $\alpha$ values.
@@
 Future work might examine how various criteria for deriving a guide value of $\alpha$ [e.g.
-@Ibrahim2015, 3724-3738] could appropriately apply in this context.
-while also considering incompatibility (commeasureability) in selecting $\alpha$ (@Ollier2020).
+@Ibrahim2015, p.
+3724-3738] could appropriately apply in this context.
+while also considering incompatibility (incommensurability) in selecting $\alpha$ [@Ollier2020].
 Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.

### 2023-03-11 analysis/memo_TAS_reviewers.Rmd
minor edits; added comment on simulation results for alpha = 0

@@
 
-This memorandum replies to comments provided by the TAS associate editor and two reviewers.
-We thank the editor and reviewers for their comments.
-We believe the revised paper is much better as a result.
+This memorandum replies to comments from a TAS associate editor and two reviewers. We believe the revised paper is much better because of their comments.
 
@@
 As our revised Introduction now makes clear (p.
-4-5 and Fig.
-1), neither our statistical approach (nor any other) can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+4-5 and Figure 1), neither our statistical approach (nor any other) can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
 
@@
 Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges.
-This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data.
-The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
+This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data, which in turn likely varies by.
+among other things, strike procedure and courtroom practices.
 As we now note in the Discussion section of the paper, the literature on the power prior discusses information criteria for guide values of $\alpha$ as well as treating $\alpha$ as random and assigning a prior to it.
 However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons.
-First, unlike models of strike procedure discussed here, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model).
+First, unlike models of strike procedure discussed here, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., generalized linear models).
 In the clinical trial setting, it is easier to measure trial-level (and patient-level) characteristics that may be responsible for incompatibility.
@@
 @Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate (more down-weighting) the influence of the historical data, and so one might have to use highly informative priors on the power parameter.
-These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis".
-More recent versions of the power prior are beginning to address this issue (e.g., using "commensurate power priors" [@Hobbs2011]).
+These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis." More recent versions of the power prior are beginning to address this issue (e.g., using "commensurate power priors" [@Hobbs2011]).
 Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter.
@@
 As an example, we now point in the paper's Introduction to @Gastwirth2014.
-Our approach has a distinctive benefit of making incompatibility assumptions transparent via choosing a fixed value for the power prior's $\alpha$.
+Our approach has a distinctive benefit of making incompatibility assumptions transparent via choosing and justifying a fixed value for the power prior's $\alpha$.
 
@@
 
-Thank you for spotting this.  The paper has been revised accordingly.
+Thank you for spotting this.  We revised the paper accordingly.
 
@@
 
-Thank you for spotting this.  The equation has been revised accordingly.
+Thank you for spotting this.  We revised the equation accordingly.
 
@@
 We revised to remove such confusion.
-This includes changing the software app to use 95% credible intervals.
+This includes changing the prototype software app to use 95% credible intervals.
 On the choice of credible intervals, see Item 1(5) above.
@@
 This comment is not clear to us.
-The real data analyses are based on real historical strike data from real criminal cases in the federal district court in Connecticut.
+We base the real data analyses on real historical strike data from real criminal cases in the federal district court in Connecticut.
 On how we collected this data, see the paper's Appendix.
@@
 See also Item 1(6).
+
+Similarly, although we did run simulations for when when $\alpha = 0$, Figures 2-4 do not depict the bias detection rate for that scenario, simply because that scenario is equivalent to evaluating bias detection with only strike data from the current trial (no historical strike data) -- a topic of prior papers [e.g.
+@Kadane2018a].
 
@@
 
-This information has been added to the text (sec. 2.5, p. 18).
-
-<!-- A normal distribution was assumed for the distribution of b.
-The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in.
-No thinning was performed, as the correlation was weak and the convergence occurred rapidly.
-Convergence was checked from both traceplots and the Gelman-Rubin statistic.
-Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (current data only _versus_ current and historical data).-->
+We added this information to the paper (sec. 2.5, p. 18).
 
@@
 
-Thank you for pointing out these issues. We have revised Equations (3) and (6) accordingly.
+Thank you for pointing out these issues. We revised Equations (3) and (6) accordingly.
 
@@
 Our statistical approach's validity turns neither on how many courts make available the kind of data it requires nor on proving how likely courts will start collecting such data.
-Indeed, as section 3.2 (paragraphs 1 and 2) of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right at attend jury selection in criminal court proceedings.
+Indeed, as section 3.2 (paragraphs 1 and 2) of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right to attend jury selection in criminal court proceedings.
 
@@
 See Item 1(5) above.
-We agree that the choice of credible interval should be consistent within different stages of the _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
+We agree that the choice of credible interval should be consistent across different stages of a _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
 
@@
 On "data from an actual case," section 2.5 of the paper (our discussion about the prototype software application) expressly relies on actual strike data from a sample of criminal cases.
+As mentioned above, we intend the reader to gain understanding by following along using our prototype software application.
 

### 2023-03-11 analysis/paper_draft3.Rmd
edits from EB and XL

@@
   US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
-Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how how to incorporate historical data on an attorney's use of peremptory strikes in past cases.
+Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how to incorporate historical data on an attorney's use of peremptory strikes in past cases.
 In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
@@
 Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court.
-With that data, the prototype shows how attorneys and others can use this approach to detect strike bias in real time during jury selection.
+With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.
 
@@
 ```
+
 To be sure, neither this nor any other statistical approach can automatically detect _Batson_ violations, because _Batson_ law permits any evidence of strike bias, not just strikes in past cases.
 And _Batson_ law  requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the _Batson_ violation).
-Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, the model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
+Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
 After testing for sensitivity of bias detection to different fixed $\alpha$ values for the power prior [@Carvalho2021, p.
@@
 \label{eq:model3}
-\end{equation} where $n_{i}$ is the total number of jury selections (trials), $n_j$ is the total number of peremptory strikes, and $\delta=(\delta_{11},\dots,\delta_{12},\dots,\delta_{{n_i}{n_j}})$
+\end{equation} where $n_{i}$ is the total number of jury selections (trials), $n_j$ is the total number of peremptory strikes, and $\delta=(\delta_{11},\delta_{12},\dots,\delta_{{n_i}{n_j}})$
 
@@
 \label{eq:powerprior}
-\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\dots,\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
+\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
@@
 Figure \@ref(fig:figbd9515) depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
-When $b_{curr}$ is high (close to 3) the bias detection rate is high (very close to 1).
+When $b_{curr}$ is high (close to 3), the bias detection rate is high (very close to 1).
 However, when $b_{curr}$ is lower (close to zero), the bias detection rate drops below 0.8.
@@
 Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible.
+However, when current and historical trials are incompatible, both type I and type II error can be high if $\alpha$ is close to one.
+As we discuss in section \@ref(selecting_alpha) below, the choice of $\alpha$ depends on assumptions about the degree of incompatibility of current and historical trials.
 
@@
 Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (Figure \@ref(fig:figbd9510)), when $b_{curr}=2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
+For example, with a 90\% credible interval (not shown), when $b_{curr}=2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
 The tradeoff is lower accuracy.
@@
 No thinning was performed, as the correlation was weak and convergence occurred rapidly.
-Convergence was checked from both traceplots and the Gelman-Rubin statistic [@Gelman1992].
+We checked convergence with both traceplots and the Gelman-Rubin statistic [@Gelman1992].
 Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current strike data only _versus_ current plus historical strike data).
@@
 
-## Selecting the Power Prior Value
+## Selecting the Power Prior Value {#selecting_alpha}
 
@@
 @Ibrahim2015, p.
-3724-3738] could appropriately apply in this context.
-while also considering incompatibility (incommensurability) in selecting $\alpha$ [@Ollier2020].
+3724-3738] could appropriately apply in this context while also considering incompatibility (incommensurability) in selecting $\alpha$ [@Ollier2020].
 Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.

### 2023-03-14 analysis/memo_TAS_reviewers.Rmd
added figures for 80% and 90% credible intervals to Appendix

@@
 Thus, the larger the credible interval one chooses, the greater the chance that zero falls within it.
-For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on other credible intervals in the Appendix.
+For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on 90% and 80% credible intervals in the Appendix.
 

### 2023-03-14 analysis/paper_draft3.Rmd
added figures for 80% and 90% credible intervals to Appendix

@@
 
-## Model Performance on Simulated Data
+## Model Performance on Simulated Data {#simulation-results}
 
@@
 
-Using a wider credible interval lowers this chance of error.
-For example, with a 90\% credible interval (not shown), when $b_{curr}=2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
+The choice of credible interval also affects bias detection.
+For example, with a 90\% credible interval, when $b_{curr}=2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
 The tradeoff is lower accuracy.
@@
 When the bias parameter is 3, however, the detection rate is still high.
+For details on simulation results for 90% and 80% credible intervals, see the Appendix.
 
@@
 
-Here, we extend the discussion in section \@ref(prototype) by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
+Here, we extend section \@ref(simulation-results) by reporting our simulation results for 90% and 80% credible intervals.
+We also extend the discussion in section \@ref(prototype) by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
 
-## Data Collection
+## Bias Detection with 90% and 80% Credible Intervals
 
-To collect this historical strike data, one of us filed a request with the U.S.
+The choice of credible interval can affect bias detection as we define it, because the larger the credible interval one chooses, the greater the chance that zero falls within it, and vice-versa.
+Here, we report simulation results for bias detection for 90% and 80% credible intervals for trials with fifteen, ten, and six strikes per side.
+
+```{r figbd9015, out.width = "95%", fig.cap = "Bias Detection Rates, 90 pct credible interval, 15 strike scenario", fig.align='center'}
+
+load_path <- here::here("figures","pp15_90CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r figbd9010, out.width = "95%", fig.cap = "Bias Detection Rates (90 pct credible interval, 10 strikes scenario)", fig.align='center'}
+
+load_path <- here::here("figures","pp10_90CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r figbd906, out.width = "95%", fig.cap = "Bias Detection Rates based on 90 pct credible interval for 6 strike scenario", fig.align='center'}
+
+load_path <- here::here("figures","pp6_90CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r figbd8015, out.width = "95%", fig.cap = "Bias Detection Rates, 80 pct credible interval, 15 strike scenario", fig.align='center'}
+
+load_path <- here::here("figures","pp15_80CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r figbd8010, out.width = "95%", fig.cap = "Bias Detection Rates (80 pct credible interval, 10 strikes scenario)", fig.align='center'}
+
+load_path <- here::here("figures","pp10_80CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r figbd806, out.width = "95%", fig.cap = "Bias Detection Rates based on 80 pct credible interval for 6 strike scenario", fig.align='center'}
+
+load_path <- here::here("figures","pp6_80CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+## Prototype Software Application
+
+### Data Collection
+
+To collect the historical strike data for the prototype software application, one of us filed a request with the U.S.
 District Court of Connecticut based on 28 U.S.C.
@@
 
-## Additional Illustrations
+### Additional Illustrations
 
@@
 Both the prior and posterior density become flatter and the credible interval plainly includes zero.
-
-```{r figapp4, out.width='75%', fig.cap = "Density Plots with Historical Strike Data (Gender).
-Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.
-Minimal weight ($a = 0.1$) applied to historical data.", fig.align='center', eval=FALSE}
-
-load_path <- here::here("figures","batson_app_screenshots", "batson_app_screenshot4.png")
-
-knitr::include_graphics(load_path)
-
-```

### 2023-03-14 analysis/paper_draft3.Rmd
added Colophon and Disclosure Statement
added references to figures in Appendix; change "MCMC" to "HMC" in main text

@@
 For each scenario, we generated 1000 data sets and fit the model on those data sets.
-For each replicate, we generate a MCMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
+For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
 For each replicate, we calculated the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.
@@
 
+# Colophon
+
+Figures were produced with ggplot2 [@ggplot2] or Tikz [@Tantau2022].
+
 # CRediT Author Statement.
@@
 Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Data Visualization, Writing - original draft.
+
+# Disclosure Statement
+
+The authors report there are no competing interests to declare.
 
@@
 The choice of credible interval can affect bias detection as we define it, because the larger the credible interval one chooses, the greater the chance that zero falls within it, and vice-versa.
-Here, we report simulation results for bias detection for 90% and 80% credible intervals for trials with fifteen, ten, and six strikes per side.
+Here, we report simulation results for bias detection for 90% intervals for trials with fifteen, ten, and six strikes per side (Figures \@ref(fig:figbd9015), \@ref(fig:figbd9010), and \@ref(fig:figbd906), respectively), and the same for 80% credible intervals (Figures \@ref(fig:figbd8015), \@ref(fig:figbd8010), and \@ref(fig:figbd806)).
 

### 2023-03-15 analysis/memo_TAS_reviewers.Rmd
type correction b_curr to b_{curr}

@@
 It is true, based on our simulations, that bias can be erroneously detected (i.e.
-even if $b_curr$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibility.
+even if $b_{curr}$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibility.
 On the other hand, given high compatibility, low values of $\alpha$ increase Type I error.
@@
 
-Similarly, although we did run simulations for when when $\alpha = 0$, Figures 2-4 do not depict the bias detection rate for that scenario, simply because that scenario is equivalent to evaluating bias detection with only strike data from the current trial (no historical strike data) -- a topic of prior papers [e.g.
+Similarly, although we did run simulations for when $\alpha = 0$, Figures 2-4 do not depict the bias detection rate for that scenario, simply because that scenario is equivalent to evaluating bias detection with only strike data from the current trial (no historical strike data) -- a topic of prior papers [e.g.
 @Kadane2018a].

### 2023-03-15 analysis/paper_draft3.Rmd
Added Results section; more descriptive figure captions; other minor edits; reference to charlatan package

@@
 - name: "[Sachin S. Pandya](https://orcid.org/0000-0001-7387-1307)"
-  #thanks: The authors gratefully acknowledge ...
-  affiliation: University of Connecticut
   email: sachin.pandya@uconn.edu
@@
 
-For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes [Wash.
+For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes (Wash.
 General Rule 37(e); Calif.
 Code of Civil Procedure $\S$ 231.7(d)(1); Conn.
-Superior Court Rule 5-12(d)].
+Superior Court Rule 5-12(d)).
 Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@Williams2017].
 
-In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, p.
+In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases (@flowers2019, p.
 2243; Wash.
 General Rule 37(g)(v); Calif.
-Code Civil Procedure $\S$ 231.7(d)(3)(G)].
+Code Civil Procedure $\S$ 231.7(d)(3)(G)).
 Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex.
@@
 
-```{r batsonworkflowfig, fig.cap="Inference from Historical Strike Data at Key Steps of Batson Challenge", fig.align = 'center', out.width="100%", fig.pos = "h"}
+```{r batsonworkflowfig, fig.cap="Inference from Historical Strike Data at Key Steps of Batson Challenge.
+Filled circles indicate steps for which bias estimates could be used.", fig.align = 'center', out.width="100%", fig.pos = "h"}
 
@@
 
-We generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
+We generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one, two, or three previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
 
@@
 For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the $b_{curr}$ is not zero.
-Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits.
+Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 95\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits.
 For each scenario, we generated 1000 data sets and fit the model on those data sets.
@@
 
-```{r figbd9515, out.width = "95%", fig.cap = "Bias Detection Rates, 95 pct credible interval, 15 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp15_95CI.png")
-knitr::include_graphics(load_path)
-
-```
+# Results
 
@@
 However, when current and historical trials are incompatible, both type I and type II error can be high if $\alpha$ is close to one.
-As we discuss in section \@ref(selecting_alpha) below, the choice of $\alpha$ depends on assumptions about the degree of incompatibility of current and historical trials.
+As we discuss in section \@ref(selectingalpha) below, the choice of $\alpha$ depends on assumptions about the degree of incompatibility of current and historical trials.
 
-```{r figbd9510, out.width = "95%", fig.cap = "Bias Detection Rates (95 pct credible interval, 10 strikes scenario)", fig.align='center'}
+```{r figbd9515, out.width = "95%", fig.align='center', fig.cap = "Bias detection rate for simulations based on 15 strikes.
+Rows of larger squares correspond to combinations of alpha values (0.1, 0.5, and 1) for low, median and equal weight of historical data to current trial data.
+Columns of larger squares correspond to 1, 2, or 3 historical trials.
+Small squares correspond to combinations of bias in current (rows) and historical (columns) trials."}
+
+load_path <- here::here("figures","pp15_95CI.png")
+knitr::include_graphics(load_path)
+
+```
+
+```{r figbd9510, out.width = "95%", fig.cap = "Bias Detection Rates (95 pct credible interval, 10 strikes scenario).", fig.align='center'}
+
+# Bias detection rate for simulations based on 10 strikes.
+Rows of larger squares correspond to combinations of alpha valuesa (0.1, 0.5, and 1) for low, median and equal weight of historical data to current trial data.
+Columns of larger squares correspond to 1, 2, or 3 historical trials.
+Small squares correspond to combinations of bias in current (rows) and historical (columns) trials.
 
@@
 
-```{r figbd956, out.width = "95%", fig.cap = "Bias Detection Rates based on 95 pct credible interval for 6 strike scenario", fig.align='center'}
+```{r figbd956, out.width = "95%", fig.cap = "Bias Detection Rates based on 95 pct credible interval for 6 strike scenario.", fig.align='center'}
 
@@
 
-```{r figsdprior, out.width = "95%", fig.cap = "Sensitivity to Initial Prior's Standard Deviation", fig.align='center'}
+```{r figsdprior, out.width = "80%", fig.cap = "Sensitivity to Standard Deviation for Initial Prior by alpha.", fig.align='center'}
+
+#   Bias was equal in current and historical trials and three levels were selected 0 , 1, and 2."
 
@@
 
-```{r figcmptbly, out.width = "95%", fig.cap = "Compatibility", fig.align='center'}
+```{r figcmptbly, out.width = "85%", fig.cap = "Bias detection under compatibility scenarios (equal bias in current and historical trials) for different numbers of strikes (columns), sizes of historical data (rows) and levels of alpha (shapes).", fig.align='center'}
 
@@
 To show how lawyers might use this app in real cases in real time during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
-Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later de-identified them.
+Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later assigned them aliases using the charlatan package [@charlatan].
 For details on this dataset, see the Appendix.
-
-With the app, the user first selects the cognizable class.
-For the prototype, we set two cognizable-class options: race and gender.
-This was because the standard juror questionnaire in this court did not ask about other possible cognizable classes.
-The data also did not expressly indicate the race or gender of the criminal defendant(s).
-For the levels of race and gender, we were limited by choices in the juror questionnaire (e.g., for gender, "male" and "female").
 
@@
 Vertical dotted lines depict 95 pct.
-credible interval.", fig.align='center', fig.ncol = 2, fig.show = 'hold', fig.subcap=c("Current Strike Data only","Historical and current strike data")}
+credible interval.
+Blue and red density plots represent draws from posterior distributions of defense attorneys and prosecutors respectively.
+Grey background density plots show the initial priors (before using current strike pattern).
+(a) shows the result when no historical data are included, and (b) shows the result when historical data are included with equal weight to current data ('Strike History Weight' set to 'Equal', i.e., alpha = 1).", fig.align='center', fig.ncol = 2, fig.show = 'hold', fig.subcap=c("Current Strike Data only","Historical and current strike data")}
 
@@
 
-## Selecting the Power Prior Value {#selecting_alpha}
+## Selecting the Power Prior Value {#selectingalpha}
 

### 2023-03-15 analysis/paper_draft3.Rmd
revised figure captions

@@
 - name: "[Sachin S. Pandya](https://orcid.org/0000-0001-7387-1307)"
+  affiliation: University of Connecticut
   email: sachin.pandya@uconn.edu
@@
 
-```{r figbd9515, out.width = "95%", fig.align='center', fig.cap = "Bias detection rate for simulations based on 15 strikes.
+```{r figbd9515, out.width = "95%", fig.align='center', fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 15 strikes.
 Rows of larger squares correspond to combinations of alpha values (0.1, 0.5, and 1) for low, median and equal weight of historical data to current trial data.
@@
 
-```{r figbd9510, out.width = "95%", fig.cap = "Bias Detection Rates (95 pct credible interval, 10 strikes scenario).", fig.align='center'}
-
-# Bias detection rate for simulations based on 10 strikes.
-Rows of larger squares correspond to combinations of alpha valuesa (0.1, 0.5, and 1) for low, median and equal weight of historical data to current trial data.
-Columns of larger squares correspond to 1, 2, or 3 historical trials.
-Small squares correspond to combinations of bias in current (rows) and historical (columns) trials.
+```{r figbd9510, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.", fig.align='center'}
 
@@
 
-```{r figbd956, out.width = "95%", fig.cap = "Bias Detection Rates based on 95 pct credible interval for 6 strike scenario.", fig.align='center'}
+```{r figbd956, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 6 strikes.", fig.align='center'}
 
@@
 
-```{r figsdprior, out.width = "80%", fig.cap = "Sensitivity to Standard Deviation for Initial Prior by alpha.", fig.align='center'}
+```{r figsdprior, out.width = "80%", fig.cap = "Sensitivity to Initial Prior's Standard Deviation (SD) by level of discounting (alpha).
+Simulations were run with 3 levels of the initial prior SD: low (SD = 1), medium (SD = 2), and hight (SD = 100).
+Bias detection rate corresponds to the proportion of simulations in which bais was detected based on 0 being excluded from the 95 percent credible interval.
+Bias was equal in current and historical trials and three levels were selected: 0 (no bias scenario), 1, and 2.
+In general, bias detection is relatively unaffected by choice initial SD, especially when alpha is close to 1.", fig.align='center'}
 
@@
 
-```{r figcmptbly, out.width = "85%", fig.cap = "Bias detection under compatibility scenarios (equal bias in current and historical trials) for different numbers of strikes (columns), sizes of historical data (rows) and levels of alpha (shapes).", fig.align='center'}
+```{r figcmptbly, out.width = "85%", fig.cap = "Bias detection under compatibility scenarios (equal bias in current and historical trials) for different numbers of strikes (columns), sizes of historical data (rows) and levels of alpha (shapes).
+As the weight applied to historical data increases (higher alpha), bias detection increases.
+Bias detection rates also increase as the number of strikes and historical datasets increase.", fig.align='center'}
 

### 2023-05-26 analysis/paper_draft4.Rmd
split paper and supplement into separate files (paper_draft4.Rmd and paper_draft4supp.Rmd)

--- 2023-03-15 analysis/paper_draft3.Rmd
@@
 abstract: |
-  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
+  United States law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
 Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how to incorporate historical data on an attorney's use of peremptory strikes in past cases.
@@
 
-In the United States, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
+In the United States, individuals selected for jury service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
 During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them.
@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes.
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
@@
 
-To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
+To evaluate the proposed method, we conducted a simulation study using Stan 2.261, a Hamiltonian Monte Carlo engine for Bayesian inference by way of R version 4.2.2 and RStan and RStan [@rstan].
 These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.
@@
 When the bias parameter is 3, however, the detection rate is still high.
-For details on simulation results for 90% and 80% credible intervals, see the Appendix.
+For details on simulation results for 90% and 80% credible intervals, see the Supplementary Material.
 
@@
 As in the simulations, a normal distribution was assumed for the distribution of $b$.
-The length of the Markov chain was 110000 and the first 10,000 iterations were dropped as burn-in.
+The length of the Markov chain was 110,000 and the first 10,000 iterations were dropped as burn-in.
 No thinning was performed, as the correlation was weak and convergence occurred rapidly.
@@
 Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later assigned them aliases using the charlatan package [@charlatan].
-For details on this dataset, see the Appendix.
+For details on this dataset, see the Supplementary Material.
 
@@
 Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
-In the Appendix, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
+In the Supplementary Material, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
 
@@
 
-# CRediT Author Statement.
+# CRediT Author Statement
 
@@
 <div id="refs"></div>
-
-# Appendix
-
-Here, we extend section \@ref(simulation-results) by reporting our simulation results for 90% and 80% credible intervals.
-We also extend the discussion in section \@ref(prototype) by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
-
-## Bias Detection with 90% and 80% Credible Intervals
-
-The choice of credible interval can affect bias detection as we define it, because the larger the credible interval one chooses, the greater the chance that zero falls within it, and vice-versa.
-Here, we report simulation results for bias detection for 90% intervals for trials with fifteen, ten, and six strikes per side (Figures \@ref(fig:figbd9015), \@ref(fig:figbd9010), and \@ref(fig:figbd906), respectively), and the same for 80% credible intervals (Figures \@ref(fig:figbd8015), \@ref(fig:figbd8010), and \@ref(fig:figbd806)).
-
-```{r figbd9015, out.width = "95%", fig.cap = "Bias Detection Rates, 90 pct credible interval, 15 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp15_90CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-```{r figbd9010, out.width = "95%", fig.cap = "Bias Detection Rates (90 pct credible interval, 10 strikes scenario)", fig.align='center'}
-
-load_path <- here::here("figures","pp10_90CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-```{r figbd906, out.width = "95%", fig.cap = "Bias Detection Rates based on 90 pct credible interval for 6 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp6_90CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-```{r figbd8015, out.width = "95%", fig.cap = "Bias Detection Rates, 80 pct credible interval, 15 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp15_80CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-```{r figbd8010, out.width = "95%", fig.cap = "Bias Detection Rates (80 pct credible interval, 10 strikes scenario)", fig.align='center'}
-
-load_path <- here::here("figures","pp10_80CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-```{r figbd806, out.width = "95%", fig.cap = "Bias Detection Rates based on 80 pct credible interval for 6 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp6_80CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-## Prototype Software Application
-
-### Data Collection
-
-To collect the historical strike data for the prototype software application, one of us filed a request with the U.S.
-District Court of Connecticut based on 28 U.S.C.
-$\S$ 1868.
-Thereafter, we received copies of certain jury-selection records associated with twenty-nine criminal cases in that court during this period.
-These included strike sheets that indicated the identification number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror.
-Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.
-
-These records, however, often did not indicate the identity of the attorneys exercising the strikes.
-While the standard forms included a signature line for the attorney, many were left blank or filled with illegible signatures.
-Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the US Attorney's office) or the criminal defendant(s) on the date(s) of jury selection.
-Where only a single attorney represented a party during jury selection, it was easy to attribute the pattern of strikes to that attorney.
-Where multiple lawyers appeared for one side, we attributed to each of them that side's pattern of peremptory challenges in that case.
-In such cases, neither the jury-selection documents nor the docket sheets indicated any hierarchy among multiple lawyers or any other basis to attribute strikes to only one attorney among them.
-After generating a dataset based on these documents, we kept only strikes where a criminal defendant was represented by an attorney.
-Finally, to de-identify this dataset, we excluded defendant names and substituted fictitious names for the attorneys using the charlatan package [@charlatan].
-
-### Additional Illustrations
-
-Suppose the same scenario discussed in section \@ref(prototype), except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
-If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)(left)).
-
-```{r figapp3, out.width='50%', fig.cap = "Screenshots of R-Shiny application show density plots for gender bias of prosecutor and defense attorney based on historical and current strike data.
-Vertical dotted lines depict 95 pct.
-credible interval.", fig.show = 'hold', fig.ncol = 2, fig.subchap=c("Equal weight","Minimal weight") }
-
-load_path3 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot3.png")
-load_path4 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot4.png")
-
-#par(mfrow=c(2,1))
-
-knitr::include_graphics(load_path3)
-knitr::include_graphics(load_path4)
-
-```
-
-Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors.
-However, because the 95\% credible interval includes zero, inferring such bias from the strike data alone is unjustified.
-Finally, if we assign minimal weight to the historical strike data ($a = 0.1$), the density plot updates accordingly (Figure \@ref(fig:figapp3)(right)).
-Both the prior and posterior density become flatter and the credible interval plainly includes zero.

### 2023-05-26 analysis/paper_draft4.tex
split paper and supplement into separate files (paper_draft4.Rmd and paper_draft4supp.Rmd)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{textcomp}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

%% load any required packages here

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{amsmath}
\usepackage{subfig}

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\hypersetup{
  pdftitle={Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data},
  pdfkeywords={Batson challenge, peremptory strikes, power prior, Bayesian},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data}

  \author{
        \href{https://orcid.org/0000-0001-7387-1307}{Sachin S. Pandya} \\
    University of Connecticut\\
     and \\     Xiaomeng Li \\
    University of Connecticut\\
     and \\     \href{https://orcid.org/0000-0003-1934-1932}{Eric Barón} \\
    University of Connecticut\\
     and \\     \href{https://orcid.org/0000-0002-9576-0517}{Timothy E. Moore} \\
    University of Connecticut\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
United States law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes. Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how to incorporate historical data on an attorney's use of peremptory strikes in past cases. In so doing, we use the power prior to adjust the weight of such historical information in the analysis. Using simulations, we show how the choice of the power prior's discounting parameter influences bias detection (how likely the credible interval for the bias parameter excludes zero), depending on the degree of incompatibility between current and historical trial data. Finally, we extend this approach with a prototype software application that lawyers could use to detect strike bias in real time during jury-selection. We illustrate this application's use with real historical strike data from a convenience sample of cases from one court.
\end{abstract}

\noindent%
{\it Keywords:} Batson challenge, peremptory strikes, power prior, Bayesian

\vfill

\newpage
\spacingset{1.9} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In the United States, individuals selected for jury service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge. During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them. By asserting a peremptory challenge, a party can declare a prospective juror ineligible (``strikes'' that juror) for a seat on the jury without the burden of explaining why.

Since \emph{Batson v. Kentucky} (1986), a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes. The party bringing a \emph{Batson} challenge bears the burden of proving such illegal bias is more likely than not to be true (LaFave et al. 2022, \(\S\) 22.3(d)).

A \emph{Batson} challenge typically proceeds in a four-step sequence. First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias. Second, if challenged, the challenger must meet an initial burden (the \emph{prima facie} case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias. Third, the striking attorney must proffer permissible reasons for their challenged strikes. Fourth, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.

For similar challenges under State law, a few States require only that an ``objective observer'' or an ``objectively reasonable person'' would find that race, ethnicity or another cognizable-class was a ``factor'' in that party's use of strikes (Wash. General Rule 37(e); Calif. Code of Civil Procedure \(\S\) 231.7(d)(1); Conn. Superior Court Rule 5-12(d)). Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes (Williams 2017).

In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases (\emph{Flowers v. Mississippi} (2019), p.~2243; Wash. General Rule 37(g)(v); Calif. Code Civil Procedure \(\S\) 231.7(d)(3)(G)). Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex. Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference (e.g., Eisenberg 2017; Grosso and O'Brien 2012; for discussion, see Gastwirth and Xu 2014, 289--97; Gastwirth and Yu 2013). Prior studies have also modeled how much a prospective juror's race affected the odds of being struck. For different modeling approaches using the same historical strike data from Mississippi, see Craft (2018); DeCamp (2021); and Dunn and Zhuo (2022).

Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction (National Center for State Courts: Center for Jury Studies 2018). Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data. Such incompatibility may occur if missing historical strike is missing not at random due to incomplete or inaccessible jury-selection records (Grosso and O'Brien 2017; Wright, Chavis, and Parks 2018).

Incompatibility may also occur if attorney strike bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials. For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants. If so, then inferring strike bias in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.

As a result, for any given estimate of strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power (Bennett et al. 2021). Other methods to increase power may in turn require adjusting for incompatibility implicitly. For example, Gastwirth and Xu (2014, 295--97) stratify strike data by trial and apply the Cochran-Mantel-Haenszel test, albeit after first selecting only historical strike data from past trials that are ``similar'' in their bias-salient characteristics to the current trial {[}p.~289, 297{]}. This initial step presumably aims to reduce incompatibility due to variation in those salient characteristics. Any such choice of ``similar'' historical strike data would need to be disclosed and justified.

In this paper, we extend a Bayesian approach to \emph{Batson} and similar challenges (Kadane 2021, 2018b, 2018a) to incorporate historical strike data in a manner that allows for transparently adjusting for the assumed degree of incompatibility between current and historical trials. We proceed in three steps. First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution. Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter. In so doing, we use the power prior (Chen and Ibrahim 2000; Ibrahim et al. 2015; for a gentle introduction, see Viele et al. 2014). The power prior raises the likelihood of the historical strike data to a fixed power (\(\alpha\)), typically between zero and one, to down-weight the historical strike data in accord with expected incompatibility between the historical trials and the current trial. Zero denotes no weight to the historical data (for complete incompatibility) and one denotes equal weight (for complete compatibility). Third, we use this posterior distribution as an informative prior for the bias parameter in the current trial.

We demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can ``detect'' attorney strike bias, i.e., whether the credible interval for the bias parameter excludes zero, given different values of the power prior's alpha parameter. Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court. With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.

\begin{figure}[h]

{\centering \includegraphics[width=1\linewidth]{../figures/batson_workflow_fig2}

}

\caption{Inference from Historical Strike Data at Key Steps of Batson Challenge. Filled circles indicate steps for which bias estimates could be used.}\label{fig:batsonworkflowfig}
\end{figure}

To be sure, neither this nor any other statistical approach can automatically detect \emph{Batson} violations, because \emph{Batson} law permits any evidence of strike bias, not just strikes in past cases. And \emph{Batson} law requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the \emph{Batson} violation). Still, \emph{before} accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's \(\alpha\) parameter. After testing for sensitivity of bias detection to different fixed \(\alpha\) values for the power prior (Carvalho and Ibrahim 2021, 5252; Ibrahim et al. 2015, 3734), attorneys can then combine the resulting inference with other evidence to decide whether to argue \emph{Batson}; to prove the \emph{Batson} prima facie case; and to persuade the trial judge to find a \emph{Batson} violation (Figure \ref{fig:batsonworkflowfig}).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{statistical-procedure}{%
\subsection{Statistical Procedure}\label{statistical-procedure}}

Following others (e.g., Kadane 2018b; Barrett 2007; Gastwirth 2005, 183), we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence. Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptory strikes. Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that corresponds to the number seats on the jury (plus alternates, if any). The parties exercise their strikes on anyone among this subset of potential jurors in an alternating sequence. Once all strikes are used or waived, the remaining prospective jurors are each assigned to seats on the jury plus a few alternate jurors.

Accordingly, for any given case \(i\) in which jury selection occurs, let \(j\) denote a peremptory strike used, and let \(\delta_{ij}\) denote whether or not a party used that strike on a person who belongs to a ``cognizable class''. If ``race'' is the bias type of interest, the cognizable class is racial minority jurors (\(\delta_{ij} =1\), 0 for White jurors). If ``sex'' is the bias of interest, the cognizable class is female jurors (\(\delta_{ij} = 1\), 0 for male jurors). In turn, let \(c_{ij}\) denote the number of cognizable class members subject to strike; and \(m_{ij}\) denote the number of cognizable class non-members subject to strike, such that \(c_{ij}+m_{ij}\) is the total number of jurors potentially subject to strike. If there is no bias, the probability is \(\frac{c_{ij}}{c_{ij}+m_{ij}}\) for striking a cognizable class member, and \(1-\frac{c_{ij}}{c_{ij}+m_{ij}}\) for striking someone who does not belong to that class.

By adding one parameter \(w\), we can measure \emph{bias} by different values of \(w\) by defining the probability of a cognizable class member being struck to be \(\frac{wc_{ij}}{wc_{ij}+m_{ij}}\). To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let \(b = log(w)\).

Accordingly, for any given value of the bias parameter \(b\), the probability of strike of a member from either class, or \(Pr(\delta_{ij})\), is such that:

\begin{equation}
Pr(\delta_{ij} | b) = 
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  \label{eq:model1}
\end{cases} 
\end{equation} This Equation \eqref{eq:model1} is equivalent to

\begin{equation}
Pr(\delta_{ij}| b) = \left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
\label{eq:model2}
\end{equation}

Given the strike data we have, i.e., \(\delta_{ij}\), \(c_{ij}\), and \(m_{ij}\), by estimating the value of \(b\), we can measure bias when a party is striking potential jurors. If \(b = 0\), there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck. If \(b>0\), we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority). Where \(b<0\), the party has a preference for a juror within the cognizable class.

The likelihood function of \(b\) is

\begin{equation}  
L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}\left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
\label{eq:model3}
\end{equation} where \(n_{i}\) is the total number of jury selections (trials), \(n_j\) is the total number of peremptory strikes, and \(\delta=(\delta_{11},\delta_{12},\dots,\delta_{{n_i}{n_j}})\)

\hypertarget{incorporating-historical-strike-data}{%
\subsection{Incorporating historical strike data}\label{incorporating-historical-strike-data}}

We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter. To do this, we introduce the power prior:

\begin{equation}  
\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
\label{eq:powerprior}
\end{equation} where \(0 \le \alpha_0 \le 1\) is the parameter controlling the weight of the historical information; \(\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})\) is the observed historical data; \(L(b|\delta_0)\) is the likelihood function of \(b\) given the historical data; and \(\pi_0(b)\) is the initial prior before the historical data is observed.

In this paper, we assume that the law for \emph{Batson} and similar challenges entails a weakly-informative initial prior:

\begin{equation}  
\pi_0(b) \sim Normal(\mu = 0, \sigma = 2)
\label{eq:initialprior}
\end{equation}

For this initial prior, we let \(\mu = 0\), because the law assigns the burden of proof in a \emph{Batson} challenge to the party bringing the challenge. Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven. This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data. Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias. For this reason, we use a normal (Gaussian) distribution with \(\sigma = 2\). This prior accords with Kadane (2021, 51), who suggests a prior for estimating \emph{Batson} strike bias that is at least unimodal, symmetric, and not dependent on the data.

After including the historical information through the power prior, the posterior distribution of \(b\) is proportional to the product of likelihood function of \(b\) and the power prior of \(b\) is

\begin{equation}
 \pi(b|\delta,\delta_0) \propto L(b|\delta)L(b|\delta_0)^{\alpha_0}\exp\left(-\frac{(b-\mu)^2}{2\sigma^2}\right)
  \label{eq:model4}
\end{equation} where

\begin{equation}
L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}\left(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{\delta_{0ij}} \left(\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{1-\delta_{0ij}}
\label{eq:model5}
\end{equation} is the likelihood function of \(b\) given historical data; \(n_{0i}\) is the total number of jury selections (trials) in the historical data; \(n_{0j}\) is the total number of peremptory strikes in the historical data; and \(\delta_{0ij}\) denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.

\hypertarget{simulation-results}{%
\subsection{Model Performance on Simulated Data}\label{simulation-results}}

To evaluate the proposed method, we conducted a simulation study using Stan 2.261, a Hamiltonian Monte Carlo engine for Bayesian inference by way of R version 4.2.2 and RStan and RStan (Stan Development Team 2022). These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data. As introduced above, \(b > 0\) represents a prosecutor's bias against a prospective juror within the cognizable class, while \(b = 0\) denotes no bias. For simplicity, we assume the defense attorney has no bias.

Because we use the power prior, we can control how much we account for the historical strike data by modifying the discounting parameter \(\alpha\) (denoted as \(\alpha_0\) in Equations \eqref{eq:powerprior} and \eqref{eq:model4}). If \(\alpha=1\), the historical strike data is equally weighted with the data on strikes in the current trial. If \(\alpha<1\), the historical data are discounted and weighted proportionally less than the current trial data. We assign \(\alpha<1\) to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes. Accordingly, in the simulation study, we evaluated different values of \(\alpha\): \(\alpha=\{0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1\}\) to show how that parameter affected bias detection.

We generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one, two, or three previous trials) using seven different values of the bias parameter for generating that historical data \(b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}\).

To generate data, we generated strike data for a current trial using seven different values of the bias parameter \(b_{curr}=\{0,0.5,1,1.5,2, 2.5,3\}\). To evaluate the strength of the power prior as a function of the degree of incompatibility between current and historical strike data (\(\lvert b_{curr} - b_{hist} \rvert > 0\)), we assigned \(b_{hist}\) the same bias values as \(b_{curr}\), i.e., \(b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}\). We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.~defendant race, charge severity) affects \(b_{curr}\) but not of all the past trials in the historical strike data.

For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., \(b_{curr} > 0\) , based on whether credible intervals of the bias parameter exclude zero. Since we use a weakly-informative prior that pulls the estimate of \(b_{curr}\) towards zero, we focus on how accurately our model can detect that bias, \emph{not} on how accurately the model can recover the true value of \(b_{curr}\) (i.e., a traditional coverage rate). For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the \(b_{curr}\) is not zero. Put another way, if we set \(b_{curr}>0\), we calculate the proportion of the 95\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits. For each scenario, we generated 1000 data sets and fit the model on those data sets. For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1000 iterations. For each replicate, we calculated the posterior mean and \(80\%\), \(90\%\), and \(95\%\) highest posterior density intervals for \(b_{curr}\).

Additionally, we present sensitivity analysis for the number of available strikes per trial (6 strikes, 10 strikes, and 15 strikes for each attorney). Thus, in total, we considered 4851 different scenarios, i.e.~the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data, 11 different values for the power prior weight parameter, and for 3 combinations of the total number of strikes.

\hypertarget{results}{%
\section{Results}\label{results}}

Figure \ref{fig:figbd9515} depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario. When \(b_{curr}\) is high (close to 3), the bias detection rate is high (very close to 1). However, when \(b_{curr}\) is lower (close to zero), the bias detection rate drops below 0.8. When there is no bias in the current or historical trials (\(b_{curr} = b_{hist} = 0\)), the bias detection rate for the 95\% credible interval is 0.05 or less. As we place more weight on the historical strike data (as \(\alpha\) increases), and the historical strike data is consistent with current information (high compatibility), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots). Conversely, as \(\alpha\) decreases, bias detection is only high when there is high bias in the current trial (\(b_{curr} \geq 1.5\)). Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible. However, when current and historical trials are incompatible, both type I and type II error can be high if \(\alpha\) is close to one. As we discuss in section \ref{selectingalpha} below, the choice of \(\alpha\) depends on assumptions about the degree of incompatibility of current and historical trials.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp15_95CI}

}

\caption{Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 15 strikes. Rows of larger squares correspond to combinations of alpha values (0.1, 0.5, and 1) for low, median and equal weight of historical data to current trial data. Columns of larger squares correspond to 1, 2, or 3 historical trials. Small squares correspond to combinations of bias in current (rows) and historical (columns) trials.}\label{fig:figbd9515}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp10_95CI}

}

\caption{Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.}\label{fig:figbd9510}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp6_95CI}

}

\caption{Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 6 strikes.}\label{fig:figbd956}
\end{figure}

As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \ref{fig:figbd9510} and \ref{fig:figbd956} respectively), bias detection decreases. Bias detection is only high when \(\alpha\) is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.

The choice of credible interval also affects bias detection. For example, with a 90\% credible interval, when \(b_{curr}=2\), the bias-detection rate is lower than with an 80\% credible interval; when \(b_{curr}=0\), that detection rate is higher. The tradeoff is lower accuracy. With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias. When the bias parameter is 3, however, the detection rate is still high. For details on simulation results for 90\% and 80\% credible intervals, see the Supplementary Material.

Bias detection will also be sensitive to the choice of the standard deviation of the initial prior (\(\sigma\)). The results above depend in part on our choice of a conservative initial prior for the bias parameter (Equation \eqref{eq:initialprior}). Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate. As a result, our model will only detect severe bias. This sensitivity, however, reduces after including the historical data, especially when \(\alpha\) is close to 1.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../figures/SD_sensitivity}

}

\caption{Sensitivity to Initial Prior's Standard Deviation (SD) by level of discounting (alpha). Simulations were run with 3 levels of the initial prior SD: low (SD = 1), medium (SD = 2), and hight (SD = 100). Bias detection rate corresponds to the proportion of simulations in which bais was detected based on 0 being excluded from the 95 percent credible interval. Bias was equal in current and historical trials and three levels were selected: 0 (no bias scenario), 1, and 2. In general, bias detection is relatively unaffected by choice initial SD, especially when alpha is close to 1.}\label{fig:figsdprior}
\end{figure}

Figure \ref{fig:figsdprior} depicts how assigning different values for the initial prior's standard deviation (\(\sigma = \{1,2,100\}\)) affects the bias detection rate at the 95\% credible interval for three \(\alpha\) values (\(\alpha = \{0.1,0.5,1\}\)) under scenarios where each attorney had 10 strikes and where \(b_{curr} = b_{hist}\). In our simulations, we set \(\sigma = 2\) for our initial prior. If we instead select \(\sigma = 1\), that matters most when \(\alpha = 0.1\), i.e., when we borrow less information from the historical data. As \(\alpha\) increases, the initial prior's standard deviation matters less. Moreover, when we increased the initial prior's standard deviation to make it much less informative (\(\sigma = 100\)), that had little impact on bias detection rates, regardless of the choice of \(\alpha\).

\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{../figures/compatibility_plot}

}

\caption{Bias detection under compatibility scenarios (equal bias in current and historical trials) for different numbers of strikes (columns), sizes of historical data (rows) and levels of alpha (shapes). As the weight applied to historical data increases (higher alpha), bias detection increases. Bias detection rates also increase as the number of strikes and historical datasets increase.}\label{fig:figcmptbly}
\end{figure}

Finally, Figure \ref{fig:figcmptbly} summarizes our simulation results for bias detection at the 95\% credible interval, given perfect compatibility (\(b_{curr} = b_{hist}\)). These results indicate that our model can accurately detect strong bias if present in both the current and historical trials. Given high compatibility, increasing \(\alpha\) improves bias detection, as does increasing the number of historical trials and more strikes per trial. Under the low number of strikes per trial (n = 6), bias detection requires strong bias in the current trial and a high \(\alpha\).

\hypertarget{simul-strikes}{%
\subsection{Simultaneous Strikes}\label{simul-strikes}}

In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes. This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck. The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.

To test this premise, we conducted the following simulation study. We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 30 prospective jurors, 15 of which were member of a cognizable-class (e.g., Black jurors). Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders. We then fit the model to the 50 trials to examine whether the estimated bias parameters across the 50 simulated trials were equivalent.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{../figures/simultaneous_sims_plot}

}

\caption{Simulation results for simultaneous strikes example. Columns correspond to values of bias ($b_{curr}$) used to simulate data. Points and horizontal lines denote estimated posterior mean and 95 pct. credible interval for bias parameter. Vertical dashed lines denote true value of $b_{curr}$ in the simulations.}\label{fig:fig-simul}
\end{figure}

Figure \ref{fig:fig-simul} depicts the results of the simultaneous strike simulations. We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios. None of their credible intervals include zero. This shows that the order of strikes does not influence the estimate of the bias parameter. Accordingly, the simultaneous-strikes procedure can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.

\hypertarget{prototype}{%
\subsection{The Software Prototype}\label{prototype}}

We describe here a prototype software application (``app'') that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges. We built this app with R version 4.3.0 (2023-04-21 ucrt) and the shiny package (R Core Team 2022; Chang et al. 2021). Unlike the simulations, we built the app using Rcpp (Eddelbuettel and Balamuta 2018) and the Metropolis-Hasting algorithm to sample \(b\) from the posterior distribution. As in the simulations, a normal distribution was assumed for the distribution of \(b\). The length of the Markov chain was 110000 and the first 10,000 iterations were dropped as burn-in. No thinning was performed, as the correlation was weak and convergence occurred rapidly. We checked convergence with both traceplots and the Gelman-Rubin statistic (Gelman and Rubin 1992). Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current strike data only \emph{versus} current plus historical strike data).

To show how lawyers might use this app in real cases in real time during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017. Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later assigned them aliases using the charlatan package (Chamberlain and Voytovich 2020). For details on this dataset, see the Supplementary Material.

Next, the user enters by hand the strike information in the case before them in the \emph{strike tally} table (Figure \ref{fig:figapp1}, top left). In the strike tally, \texttt{round} denotes the order of strikes, \texttt{num\_cog} denotes the number of prospective jurors that could be struck that belong to the cognitive class; \texttt{total} denotes the total number of prospective jurors that could be struck; \texttt{cog} indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and \texttt{party} indicates which side used the strike (PP = prosecutor, PD = defense attorney). the user can add or delete rows to the strike tally as needed.

\begin{table}

\caption{\label{tab:inpts}Hypothetical Strike Tally}
\centering
\begin{tabular}[t]{rrrrl}
\toprule
round & num\_cog & total & cog & party\\
\midrule
1 & 5 & 21 & 0 & PP\\
2 & 5 & 20 & 1 & PD\\
3 & 4 & 19 & 1 & PP\\
4 & 3 & 18 & 0 & PD\\
5 & 2 & 17 & 0 & PP\\
\addlinespace
6 & 2 & 16 & 1 & PD\\
7 & 2 & 15 & 1 & PP\\
8 & 2 & 14 & 0 & PD\\
9 & 2 & 13 & 0 & PP\\
10 & 2 & 12 & 0 & PD\\
\bottomrule
\end{tabular}
\end{table}

To illustrate, suppose the strike tally in Table 1 depicts the pattern of strikes in the present case with defense attorney Aaron Waelchi and prosecutor Lawrance Klocko V (both aliases for actual attorneys in the historical strike data). After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense. Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \ref{fig:figapp1}(a)). Here, the 95\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally alone. In the default setting, the pull-down menus for prosecutor and defense are set to ``None''. As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for this prosecutor and defense attorney based only on the strike tally data and the initial prior (\(b \sim N(0,2)\)).

\begin{figure}

{\centering \subfloat[Current Strike Data only\label{fig:figapp1-1}]{\includegraphics[width=0.5\linewidth]{../figures/batson_app_screenshots/batson_app_screenshot1} }\subfloat[Historical and current strike data\label{fig:figapp1-2}]{\includegraphics[width=0.5\linewidth]{../figures/batson_app_screenshots/batson_app_screenshot2} }

}

\caption{Screenshots of R-Shiny application showing density plots for race bias of prosecutor and defense attorney. Vertical dotted lines depict 95 pct. credible interval. Blue and red density plots represent draws from posterior distributions of defense attorneys and prosecutors respectively. Grey background density plots show the initial priors (before using current strike pattern). (a) shows the result when no historical data are included, and (b) shows the result when historical data are included with equal weight to current data ('Strike History Weight' set to 'Equal', i.e., alpha = 1).}\label{fig:figapp1}
\end{figure}

To use historical strike data, the user selects the name of the prosecutor or defense attorney from the pull-down menus. If an attorney's name cannot be found, the app has no historical strike data for that attorney. Once selected, the prior and posterior density plots automatically update to account for the pre-loaded historical strike data for that attorney. For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information (\(\alpha = 1\)). The user has two other options: half weight (\(\alpha = 0.5\)) and minimal weight (\(\alpha = 0.2\)).

In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to ``Equal''. The density plots update accordingly (Figure \ref{fig:figapp1}(b)). Now the credible intervals clearly exclude zero. Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case. In the Supplementary Material, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Statistical methods for \emph{Batson} and similar challenges often use historical strike data to infer illegal strike bias. The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter. By incorporating the power prior, this approach makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a \emph{Batson} challenge or not. Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them decide whether to make a \emph{Batson} challenge at all. If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and for ultimately proving illegal bias (Figure \ref{fig:batsonworkflowfig}).

\hypertarget{selectingalpha}{%
\subsection{Selecting the Power Prior Value}\label{selectingalpha}}

In practice, the anticipated degree of incompatibility should influence which value of \(\alpha\) to select for the power prior. Given perfect compatibility (\(\lvert b_{curr} - b_{hist} \rvert = 0\)), we should let \(\alpha = 1\) (equal weight between current and historical data). This will maximize bias detection even when the number of strikes per trial is small. This effect of increased bias detection can be seen in the simulation results depicted in Figures \ref{fig:figbd9515}, \ref{fig:figbd9510}, and \ref{fig:figbd956} along the positive diagonals of the individual blocks.

Bias detection also increases with the number of historical trials and the number of strikes per trial. Higher values of \(\alpha\) lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \ref{fig:figcmptbly}). The choice of \(\alpha\) matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \ref{fig:figcmptbly}). Overall, given high compatibility, increasing \(\alpha\) will increase bias detection. But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current bias (\(b_{curr}\)) is low; and bias detection can be (erroneously) low where current bias is high but bias in historical trials (\(b_{hist}\)) is low.

Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of \(\alpha\) by selecting multiple \(\alpha\) values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of \(\alpha\) value. Our prototype app permits this kind of testing of sensitivity in real time by providing multiple options for \(\alpha\) values. If the inference of bias is sensitive to the choice of \(\alpha\) values, all else equal, then one should worry more about incompatibility and choose a lower \(\alpha\) value accordingly.

This approach also applies when the unit of analysis is not the attorney, but the law firm or office. To illustrate, suppose prosecutors from the same office were all exposed to training that recommended striking Black jurors. If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the ``district attorney's office''). Thereafter, any down-weighting adjustment (choosing \(\alpha < 1\)) depends on assumptions about how dissimilarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibility, notwithstanding the prosecutors' common exposure to the training. If none, one could assign equal weight to the historical data (\(\alpha\) = 1). If one allows for some incompatibility nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial. Or one could keep all the historical data; choose an \(\alpha\) value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to different \(\alpha\) values.

In general, by their choice of \(\alpha\), the lawyer explicitly and transparently weights the historical strike data in the analysis. This transparency thwarts lawyer efforts to strategically select \(\alpha\) values. Rather, at the key steps of the \emph{Batson} challenge process (Figure \ref{fig:batsonworkflowfig}), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of \(\alpha\). In any case, the choice of \(\alpha\) cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a \emph{Batson} violation). Future work might examine how various criteria for deriving a guide value of \(\alpha\) (e.g. Ibrahim et al. 2015, 3724--38) could appropriately apply in this context while also considering incompatibility (incommensurability) in selecting \(\alpha\) (Ollier et al. 2020). Knowing exactly how to quantify incompatibility and specify an appropriate prior for \(\alpha\) is an area of active research (e.g. Han, Ye, and Wang 2023; Pawel et al. 2022).

\hypertarget{modeling-the-strike-procedure}{%
\subsection{Modeling the Strike Procedure}\label{modeling-the-strike-procedure}}

The Bayesian approach here also requires a model to match the strike procedure of the court of interest. This choice of model can matter in at least three ways. First, it can affect the missingness of historical strike data. For example, given the model we used (Equation \eqref{eq:model1}), we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike. Such information may be missing from jury selection records of past cases.

Second, the choice of model can affect prospective data collection not only by courts but also by courtroom observers. The public has a presumptive First Amendment right to attend jury voir dire in criminal cases. Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from voir dire (\emph{{Press-Enterprise Co. v. Superior Court}} 1984; \emph{Presley v. Georgia} 2010). As a result, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time. Thus, such courtroom observers can prospectively collect strike data that they would have been less likely to obtain from court records after the fact.

Third, the choice of model matters if historical strike data for a particular attorney covers strikes exercised in multiple courts with different strike procedures. In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \ref{simul-strikes} ). In other cases, however, one cannot apply the same model to courts with different jury selection practices. For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In this paper, we extended a Bayesian approach to estimate attorney strike bias in the use of peremptory challenges by incorporating historical data on that attorney's use of strikes in past cases. In so doing, we used the power prior to adjust the weight of such historical information. Our simulations showed that how well our approach detects attorney bias depends on the number of past trials in which that attorney used strikes; the number of strikes per trial; and the degree of incompatibility, i.e., the distance between the bias parameters for the current trial and for the past trials. Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank the Associate Editor and two anonymous reviewers for their comments.

\hypertarget{colophon}{%
\section{Colophon}\label{colophon}}

Figures were produced with ggplot2 (Wickham 2016) or Tikz (Tantau 2022).

\hypertarget{credit-author-statement}{%
\section{CRediT Author Statement}\label{credit-author-statement}}

Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Data Visualization, Writing - original draft.

\hypertarget{disclosure-statement}{%
\section{Disclosure Statement}\label{disclosure-statement}}

The authors report there are no competing interests to declare.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Barrett2007}{}}%
Barrett, Bruce E. 2007. {``Detecting Bias in Jury Selection.''} \emph{The American Statistician} 61 (4): 296--301. \url{https://doi.org/10.1198/000313007X243629}.

\leavevmode\vadjust pre{\hypertarget{ref-batson1986}{}}%
\emph{Batson v. Kentucky}. 1986. 476 U.S. 79. \url{https://www.loc.gov/item/usrep476079/}.

\leavevmode\vadjust pre{\hypertarget{ref-Bennett2021}{}}%
Bennett, Maxine, Simon White, Nicky Best, and Adrian Mander. 2021. {``A Novel Equivalence Probability Weighted Power Prior for Using Historical Control Data in an Adaptive Clinical Trial Design: A Comparison to Standard Methods.''} \emph{Pharmaceutical Statistics} 20 (3): 462--84. \url{https://doi.org/10.1002/pst.2088}.

\leavevmode\vadjust pre{\hypertarget{ref-calccp2021}{}}%
\emph{California Code of Civil Procedure §231.7}. 2022. \url{https://leginfo.legislature.ca.gov/faces/codes_displaySection.xhtml?sectionNum=231.7.\&nodeTreePath=4.3.1\&lawCode=CCP}.

\leavevmode\vadjust pre{\hypertarget{ref-Carvalho2021}{}}%
Carvalho, Luiz Max, and Joseph G. Ibrahim. 2021. {``On the Normalized Power Prior.''} \emph{Statistics in Medicine} 40 (24): 5251--75. \url{https://doi.org/10.1002/sim.9124}.

\leavevmode\vadjust pre{\hypertarget{ref-charlatan}{}}%
Chamberlain, Scott, and Kyle Voytovich. 2020. \emph{Charlatan: Make Fake Data}. \url{https://CRAN.R-project.org/package=charlatan}.

\leavevmode\vadjust pre{\hypertarget{ref-shiny}{}}%
Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges. 2021. \emph{{shiny: Web Application Framework for R}}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\vadjust pre{\hypertarget{ref-Chen2000}{}}%
Chen, Ming-Hui, and Joseph G. Ibrahim. 2000. {``Power Prior Distributions for Regression Models.''} \emph{Statistical Science} 15 (1): 46--60. \url{https://doi.org/10.1214/ss/1009212673}.

\leavevmode\vadjust pre{\hypertarget{ref-Craft2018a}{}}%
Craft, Will. 2018. {``{Peremptory Strikes in Mississippi's Fifth Circuit Court District}.''} \url{https://features.apmreports.org/files/peremptory_strike_methodology.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-DeCamp2021}{}}%
DeCamp, Whitney. 2021. {``Gender and the Peremptory Challenge: Separating the Effects of Race and Gender in Jury Selection.''} \emph{Women \& Criminal Justice} 31 (2): 159--73. \url{https://doi.org/10.1080/08974454.2020.1785372}.

\leavevmode\vadjust pre{\hypertarget{ref-Dunn2022}{}}%
Dunn, Jack, and Ying Daisy Zhuo. 2022. {``Detecting Racial Bias in Jury Selection.''} \emph{Operations Research Forum} 3 (3): 1--17. \url{https://doi.org/10.1007/s43069-022-00151-x}.

\leavevmode\vadjust pre{\hypertarget{ref-Rcpp}{}}%
Eddelbuettel, Dirk, and James Joseph Balamuta. 2018. {``{Extending \emph{R} with \emph{C++}: A Brief Introduction to \emph{Rcpp}}.''} \emph{The American Statistician} 72 (1): 28--36. \url{https://doi.org/10.1080/00031305.2017.1375990}.

\leavevmode\vadjust pre{\hypertarget{ref-Eisenberg2017}{}}%
Eisenberg, Ann M. 2017. {``{Removal of Women and African Americans in Jury Selection in South Carolina Capital Cases, 1997-2012}.''} \emph{Northeastern University Law Review} 9: 299--345.

\leavevmode\vadjust pre{\hypertarget{ref-flowers2019}{}}%
\emph{Flowers v. Mississippi}. 2019. 139 S. Ct. 2228.

\leavevmode\vadjust pre{\hypertarget{ref-Gastwirth2005}{}}%
Gastwirth, Joseph L. 2005. {``{Case comment: statistical tests for the analysis of data on peremptory challenges: clarifying the standard of proof needed to establish a \emph{prima facie} case of discrimination in \emph{Johnson} v. \emph{California}}.''} \emph{Law, Probability and Risk} 4 (3): 179--85. \url{https://doi.org/10.1093/lpr/mgi014}.

\leavevmode\vadjust pre{\hypertarget{ref-Gastwirth2014}{}}%
Gastwirth, Joseph L., and Wenjing Xu. 2014. {``Statistical Tools for Evaluating the Adequacy of the Size of a Sample on Which Statistical Evidence Is Based.''} \emph{Law, Probability and Risk} 13 (3-4): 277--306. \url{https://doi.org/10.1093/lpr/mgu010}.

\leavevmode\vadjust pre{\hypertarget{ref-Gastwirth2013}{}}%
Gastwirth, Joseph L., and Binbing Yu. 2013. {``Case Comment: Appropriate Statistical Methodology Yields Stronger Evidence of Discriminatory Peremptory Challenges in {North Carolina}: Application to the {Randolph County} Data in \emph{North Carolina} v. \emph{Rouse} and Related Cases.''} \emph{Law, Probability and Risk} 12 (2): 155--64. \url{https://doi.org/10.1093/lpr/mgs030}.

\leavevmode\vadjust pre{\hypertarget{ref-Gelman1992}{}}%
Gelman, Andrew, and Donald B. Rubin. 1992. {``{Inference from Iterative Simulation Using Multiple Sequences}.''} \emph{Statistical Science} 7 (4): 457--72. \url{https://doi.org/10.1214/ss/1177011136}.

\leavevmode\vadjust pre{\hypertarget{ref-Grosso2012}{}}%
Grosso, Catherine M., and Barbara O'Brien. 2012. {``A Stubborn Legacy: The Overwhelming Importance of Race in Jury Selection in 173 Post-Batson North Carolina Capital Trials.''} \emph{Iowa Law Review} 97: 1531--59.

\leavevmode\vadjust pre{\hypertarget{ref-Grosso2017}{}}%
---------. 2017. {``A Call to Criminal Courts: Record Rules for Batson.''} \emph{Kentucky Law Journal} 105: 651.

\leavevmode\vadjust pre{\hypertarget{ref-Han2023}{}}%
Han, Zifei, Keying Ye, and Min Wang. 2023. {``A Study on the Power Parameter in Power Prior Bayesian Analysis.''} \emph{The American Statistician} 77 (1): 12--19. \url{https://doi.org/10.1080/00031305.2022.2046160}.

\leavevmode\vadjust pre{\hypertarget{ref-Ibrahim2015}{}}%
Ibrahim, Joseph G., Ming-Hui Chen, Yeongjin Gwon, and Fang Chen. 2015. {``The Power Prior: Theory and Applications.''} \emph{Statistics in Medicine} 34 (28): 3724--49. \url{https://doi.org/10.1002/sim.6728}.

\leavevmode\vadjust pre{\hypertarget{ref-Kadane2018b}{}}%
Kadane, Joseph B. 2018a. {``\emph{Batson} and Reverse-\emph{Batson} Motions in North Carolina: \emph{State v. Hurd} and \emph{State v. Tucker}.''} \emph{Law, Probability and Risk} 17 (3): 263--73. \url{https://doi.org/10.1093/lpr/mgy015}.

\leavevmode\vadjust pre{\hypertarget{ref-Kadane2018a}{}}%
---------. 2018b. {``Statistics for Batson Challenges.''} \emph{Law, Probability and Risk} 17 (1): 1--13. \url{https://doi.org/10.1093/lpr/mgx016}.

\leavevmode\vadjust pre{\hypertarget{ref-Kadane2021}{}}%
---------. 2021. {``{Priors neutral between the parties: the Batson motion in \emph{Idaho v. Ish}}.''} \emph{Law, Probability and Risk} 20 (1): 45--62. \url{https://doi.org/10.1093/lpr/mgab005}.

\leavevmode\vadjust pre{\hypertarget{ref-LaFave2022}{}}%
LaFave, Wayne R., Jerold H. Israel, Nancy J. King, and Orin S. Kerr. 2022. \emph{Criminal Procedure}. 4th ed. Vol. 6.

\leavevmode\vadjust pre{\hypertarget{ref-nscs-jury2018}{}}%
National Center for State Courts: Center for Jury Studies. 2018. {``Comparative Data: Peremptory Challenges.''} 2018. \url{https://public.tableau.com/shared/BYJM8RF8Y?:showVizHome=no}.

\leavevmode\vadjust pre{\hypertarget{ref-Ollier2020}{}}%
Ollier, Adrien, Satoshi Morita, Moreno Ursino, and Sarah Zohar. 2020. {``An Adaptive Power Prior for Sequential Clinical Trials -- Application to Bridging Studies.''} \emph{Statistical Methods in Medical Research} 29 (8): 2282--94. \url{https://doi.org/10.1177/0962280219886609}.

\leavevmode\vadjust pre{\hypertarget{ref-Pawel2022}{}}%
Pawel, Samuel, Frederik Aust, Leonhard Held, and Eric-Jan Wagenmakers. 2022. {``Normalized Power Priors Always Discount Historical Data.''} arXiv. \url{https://doi.org/10.48550/ARXIV.2206.04379}.

\leavevmode\vadjust pre{\hypertarget{ref-presley2010}{}}%
\emph{Presley v. Georgia}. 2010. 558 U.S. 209. \url{https://www.loc.gov/item/usrep558209/}.

\leavevmode\vadjust pre{\hypertarget{ref-press1984}{}}%
\emph{{Press-Enterprise Co. v. Superior Court}}. 1984. 464 U.S. 501. \url{https://www.loc.gov/item/usrep464501/}.

\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
R Core Team. 2022. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-rstan}{}}%
Stan Development Team. 2022. {``{RStan}: The {R} Interface to {Stan}.''} \url{https://mc-stan.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-Tantau2022}{}}%
Tantau, Till. 2022. \emph{The TikZ and PGF Packages: Manual for Version 3.1.9a}. \url{https://github.com/pgf-tikz/pgf}.

\leavevmode\vadjust pre{\hypertarget{ref-Viele2014}{}}%
Viele, Kert, Scott Berry, Beat Neuenschwander, Billy Amzal, Fang Chen, Nathan Enas, Brian Hobbs, et al. 2014. {``Use of Historical Control Data for Assessing Treatment Effects in Clinical Trials.''} \emph{Pharmaceutical Statistics} 13 (1): 41--54. \url{https://doi.org/doi.org/10.1002/pst.1589}.

\leavevmode\vadjust pre{\hypertarget{ref-WashR37}{}}%
\emph{Washingon State General Rule 37}. 2022. \url{https://www.courts.wa.gov/court_rules/pdf/GR/GA_GR_37_00_00.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-Williams2017}{}}%
Williams, C. J. 2017. {``Proposing a Peremptory Methodology for Exercising Peremptory Strikes.''} \emph{American Criminal Law Review} 54: 277--302.

\leavevmode\vadjust pre{\hypertarget{ref-Wright2018}{}}%
Wright, Ronald F., Kami Chavis, and Gregory S. Parks. 2018. {``The Jury Sunshine Project: Jury Selection Data as a Political Issue.''} \emph{Univeristy of Illinois Law Review} 2018 (4): 1407--42.

\end{CSLReferences}

\hypertarget{supplementary-material-for-bayesian-detection-of-bias-in-peremptory-challenges-using-historical-strike-data}{%
\section{Supplementary Material for ``Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data''}\label{supplementary-material-for-bayesian-detection-of-bias-in-peremptory-challenges-using-historical-strike-data}}

Here, we extend section \ref{simulation-results} of the paper by reporting our simulation results for 90\% and 80\% credible intervals. We also extend the discussion in section \ref{prototype} of the paper by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.

\hypertarget{bias-detection-with-90-and-80-credible-intervals}{%
\subsection{Bias Detection with 90\% and 80\% Credible Intervals}\label{bias-detection-with-90-and-80-credible-intervals}}

The choice of credible interval can affect bias detection as we define it, because the larger the credible interval one chooses, the greater the chance that zero falls within it, and vice-versa. Here, we report simulation results for bias detection for 90\% intervals for trials with fifteen, ten, and six strikes per side (Figures \ref{fig:figbd9015}, \ref{fig:figbd9010}, and \ref{fig:figbd906}, respectively), and the same for 80\% credible intervals (Figures \ref{fig:figbd8015}, \ref{fig:figbd8010}, and \ref{fig:figbd806}).

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp15_90CI}

}

\caption{Bias Detection Rates, 90 pct credible interval, 15 strike scenario}\label{fig:figbd9015}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp10_90CI}

}

\caption{Bias Detection Rates (90 pct credible interval, 10 strikes scenario)}\label{fig:figbd9010}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp6_90CI}

}

\caption{Bias Detection Rates based on 90 pct credible interval for 6 strike scenario}\label{fig:figbd906}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp15_80CI}

}

\caption{Bias Detection Rates, 80 pct credible interval, 15 strike scenario}\label{fig:figbd8015}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp10_80CI}

}

\caption{Bias Detection Rates (80 pct credible interval, 10 strikes scenario)}\label{fig:figbd8010}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../figures/pp6_80CI}

}

\caption{Bias Detection Rates based on 80 pct credible interval for 6 strike scenario}\label{fig:figbd806}
\end{figure}

\hypertarget{prototype-software-application}{%
\subsection{Prototype Software Application}\label{prototype-software-application}}

\hypertarget{data-collection}{%
\subsubsection{Data Collection}\label{data-collection}}

To collect the historical strike data for the prototype software application, one of us filed a request with the U.S. District Court of Connecticut based on 28 U.S.C. \(\S\) 1868. Thereafter, we received copies of certain jury-selection records associated with twenty-nine criminal cases in that court during this period. These included strike sheets that indicated the identification number of prospective jurors who were struck by peremptory challenge, the order in which they were struck, and which side (prosecutor or defense) struck which juror. Such records also included a tally of answers to juror questionnaires that asked each prospective juror to report their race and gender.

These records, however, often did not indicate the identity of the attorneys exercising the strikes. While the standard forms included a signature line for the attorney, many were left blank or filled with illegible signatures. Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the United States Attorney's office) or the criminal defendant(s) on the date(s) of jury selection. Where only a single attorney represented a party during jury selection, it was easy to attribute the pattern of strikes to that attorney. Where multiple lawyers appeared for one side, we attributed to each of them that side's pattern of peremptory challenges in that case. In such cases, neither the jury-selection documents nor the docket sheets indicated any hierarchy among multiple lawyers or any other basis to attribute strikes to only one attorney among them. After generating a dataset based on these documents, we kept only strikes where a criminal defendant was represented by an attorney. Finally, to de-identify this dataset, we excluded defendant names and substituted fictitious names for the attorneys using the charlatan package (Chamberlain and Voytovich 2020).

\hypertarget{additional-illustrations}{%
\subsubsection{Additional Illustrations}\label{additional-illustrations}}

Suppose the same scenario discussed in section \ref{prototype}, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors. If so, we select ``gender'' as the cognizable class, and the density plots update accordingly (Figure \ref{fig:figapp3}(left)).

\begin{figure}
\includegraphics[width=0.5\linewidth]{../figures/batson_app_screenshots/batson_app_screenshot3} \includegraphics[width=0.5\linewidth]{../figures/batson_app_screenshots/batson_app_screenshot4} \caption{Screenshots of R-Shiny application show density plots for gender bias of prosecutor and defense attorney based on historical and current strike data. Vertical dotted lines depict 95 pct. credible interval.}\label{fig:figapp3}
\end{figure}

Here, most of the prosecutor's density curve of both prior and posterior are to the left of zero, indicating the possible bias against male jurors. However, because the 95\% credible interval includes zero, inferring such bias from the strike data alone is unjustified. Finally, if we assign minimal weight to the historical strike data (\(a = 0.1\)), the density plot updates accordingly (Figure \ref{fig:figapp3}(right)). Both the prior and posterior density become flatter and the credible interval plainly includes zero.

\end{document}

### 2023-05-26 analysis/paper_draft4supp.Rmd
split paper and supplement into separate files (paper_draft4.Rmd and paper_draft4supp.Rmd)

--- 2023-03-14 analysis/paper_draft3.Rmd
@@
 ---
-title: "Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data"
+title: Supplementary Material for "Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data"
 date: "`r Sys.Date()`"
@@
 - name: "[Sachin S. Pandya](https://orcid.org/0000-0001-7387-1307)"
-  #thanks: The authors gratefully acknowledge ...
   affiliation: University of Connecticut
@@
 biblio-style: agsm.bst
-nocite: |
-  @WashR37, @calccp2021
-keywords: "Batson challenge, peremptory strikes, power prior, Bayesian"
-abstract: |
-  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
-Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how to incorporate historical data on an attorney's use of peremptory strikes in past cases.
-In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
-Using simulations, we show how the choice of the power prior's discounting parameter influences bias detection (how likely the credible interval for the bias parameter excludes zero), depending on the degree of incompatibility between current and historical trial data.
-Finally, we extend this approach with a prototype software application that lawyers could use to detect strike bias in real time during jury-selection.
-We illustrate this application's use with real historical strike data from a convenience sample of cases from one court.
 
@@
 
-# Introduction
+Here, we extend section 3 of the paper by reporting our simulation results for 90% and 80% credible intervals.
+We also extend the discussion in section 3.2 of the paper by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
 
-In the United States, individuals selected for jury-service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
-During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them.
-By asserting a peremptory challenge, a party can declare a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why.
-
-Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes.
-The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
-
-A _Batson_ challenge typically proceeds in a four-step sequence.
-First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
-Second, if challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
-Third, the striking attorney must proffer permissible reasons for their challenged strikes.
-Fourth, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
-
-For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes [Wash.
-General Rule 37(e); Calif.
-Code of Civil Procedure $\S$ 231.7(d)(1); Conn.
-Superior Court Rule 5-12(d)].
-Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@Williams2017].
-
-In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases [@flowers2019, p.
-2243; Wash.
-General Rule 37(g)(v); Calif.
-Code Civil Procedure $\S$ 231.7(d)(3)(G)].
-Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex.
-Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017; @Grosso2012; for discussion, see @Gastwirth2014, 289-297; @Gastwirth2013].
-Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
-For different modeling approaches using the same historical strike data from Mississippi, see @Craft2018a; @DeCamp2021; and @Dunn2022.
-
-Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
-Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
-Such incompatibility may occur if missing historical strike is missing not at random due to incomplete or inaccessible jury-selection records [@Grosso2017; @Wright2018].
-
-Incompatibility may also occur if attorney strike bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
-For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
-If so, then inferring strike bias in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
-
-As a result, for any given estimate of strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power [@Bennett2021].
-Other methods to increase power may in turn require adjusting for incompatibility implicitly.
-For example, @Gastwirth2014 [p.
-295-297] stratify strike data by trial and apply the Cochran-Mantel-Haenszel test, albeit after first selecting only historical strike data from past trials that are "similar" in their bias-salient characteristics to the current trial [p.
-289, 297].
-This initial step presumably aims to reduce incompatibility due to variation in those salient characteristics.
-Any such choice of "similar" historical strike data would need to be disclosed and justified.
-
-In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows for transparently adjusting for the assumed degree of incompatibility between current and historical trials.
-We proceed in three steps.
-First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
-Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the power prior [@Chen2000; @Ibrahim2015; for a gentle introduction, see @Viele2014].
-The power prior raises the likelihood of the historical strike data to a fixed power ($\alpha$), typically between zero and one, to down-weight the historical strike data in accord with expected incompatibility between the historical trials and the current trial.
-Zero denotes no weight to the historical data (for complete incompatibility) and one denotes equal weight (for complete compatibility).
-Third, we use this posterior distribution as an informative prior for the bias parameter in the current trial.
-
-We demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can "detect" attorney strike bias, i.e., whether the credible interval for the bias parameter excludes zero, given different values of the power prior's alpha parameter.
-Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court.
-With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.
-
-```{r batsonworkflowfig, fig.cap="Inference from Historical Strike Data at Key Steps of Batson Challenge", fig.align = 'center', out.width="100%", fig.pos = "h"}
-
-load_path <- here::here("figures","batson_workflow_fig2.pdf")
-knitr::include_graphics(load_path)
-
-```
-
-To be sure, neither this nor any other statistical approach can automatically detect _Batson_ violations, because _Batson_ law permits any evidence of strike bias, not just strikes in past cases.
-And _Batson_ law  requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the _Batson_ violation).
-Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
-After testing for sensitivity of bias detection to different fixed $\alpha$ values for the power prior [@Carvalho2021, p.
-5252; @Ibrahim2015, p.
-3734], attorneys can then combine the resulting inference with other evidence to decide whether to argue _Batson_; to prove the _Batson_ prima facie case; and to persuade the trial judge to find  a _Batson_ violation (Figure \@ref(fig:batsonworkflowfig)).
-
-# Methods
-
-## Statistical Procedure
-
-Following others [e.g., @Kadane2018a; @Barrett2007; @Gastwirth2005, p.
-183], we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
-Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptory strikes.
-Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that corresponds to the number seats on the jury (plus alternates, if any).
-The parties exercise their strikes on anyone among this subset of potential jurors in an alternating sequence.
-Once all strikes are used or waived, the remaining prospective jurors are each assigned to seats on the jury plus a few alternate jurors.
-
-Accordingly, for any given case $i$ in which jury selection occurs, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class".
-If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).
-If "sex" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).
-In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.
-If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking someone who does not belong to that class.
-
-By adding one parameter $w$, we can measure _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
-To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let $b = log(w)$.
-
-Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is such that:
-
-\begin{equation}
-Pr(\delta_{ij} | b) = 
-\begin{cases}
-  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
-  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  \label{eq:model1}
-\end{cases} 
-\end{equation} This Equation \eqref{eq:model1} is equivalent to
-
-\begin{equation}
-Pr(\delta_{ij}| b) = \left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
-\label{eq:model2}
-\end{equation}
-
-Given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$, by estimating the value of $b$, we can measure bias when a party is striking potential jurors.
-If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.
-If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
-Where $b<0$, the party has a preference for a juror within the cognizable class.
-
-The likelihood function of $b$ is
-
-\begin{equation}  
-L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}\left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
-\label{eq:model3}
-\end{equation} where $n_{i}$ is the total number of jury selections (trials), $n_j$ is the total number of peremptory strikes, and $\delta=(\delta_{11},\delta_{12},\dots,\delta_{{n_i}{n_j}})$
-
-## Incorporating historical strike data
-
-We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter.
-To do this, we introduce the power prior:
-
-\begin{equation}  
-\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
-\label{eq:powerprior}
-\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
-
-In this paper, we assume that the law for _Batson_ and similar challenges entails a weakly-informative initial prior:
-
-\begin{equation}  
-\pi_0(b) \sim Normal(\mu = 0, \sigma = 2)
-\label{eq:initialprior}
-\end{equation}
-
-For this initial prior, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge.
-Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
-This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
-Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias.
-For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
-This prior accords with @Kadane2021 [p.
-51], who suggests a prior for estimating _Batson_ strike bias that is at least unimodal, symmetric, and not dependent on the data.
-
-After including the historical information through the power prior, the posterior distribution of $b$ is proportional to the product of likelihood function of $b$ and the power prior of $b$ is
-
-\begin{equation}
- \pi(b|\delta,\delta_0) \propto L(b|\delta)L(b|\delta_0)^{\alpha_0}\exp\left(-\frac{(b-\mu)^2}{2\sigma^2}\right)
-  \label{eq:model4}
-\end{equation} where
-
-\begin{equation}
-L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}\left(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{\delta_{0ij}} \left(\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{1-\delta_{0ij}}
-\label{eq:model5}
-\end{equation} is the likelihood function of $b$ given historical data; $n_{0i}$ is the total number of jury selections (trials) in the historical data; $n_{0j}$ is the total number of peremptory strikes in the historical data; and $\delta_{0ij}$ denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.
-
-## Model Performance on Simulated Data {#simulation-results}
-
-To evaluate the proposed method, we conducted a simulation study using Stan `r rstan::stan_version()`, a Hamiltonian Monte Carlo engine for Bayesian inference by way of `r R.version$version.string` and RStan [@rstan].
-These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.
-As introduced above, $b > 0$ represents a prosecutor's bias against a prospective juror within the cognizable class, while $b = 0$ denotes no bias.
-For simplicity, we assume the defense attorney has no bias.
-
-Because we use the power prior, we can control how much we account for the historical strike data by modifying the discounting parameter $\alpha$ (denoted as $\alpha_0$ in Equations \@ref(eq:powerprior) and \@ref(eq:model4)).
-If $\alpha=1$, the historical strike data is equally weighted with the data on strikes in the current trial.
-If $\alpha<1$, the historical data are discounted and weighted proportionally less than the current trial data.
-We assign $\alpha<1$ to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
-Accordingly, in the simulation study, we evaluated different values of $\alpha$: $\alpha=\{0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1\}$ to show how that parameter affected bias detection.
-
-We generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
-
-To generate data, we generated strike data for a current trial using seven different values of the bias parameter $b_{curr}=\{0,0.5,1,1.5,2, 2.5,3\}$.
-To evaluate the strength of the power prior as a function of the degree of incompatibility between current and historical strike data ($\lvert b_{curr} - b_{hist} \rvert > 0$), we assigned $b_{hist}$ the same bias values as $b_{curr}$, i.e., $b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}$.
-We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.
-defendant race, charge severity) affects $b_{curr}$ but not of all the past trials in the historical strike data.
-
-For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
-Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus on how accurately our model can detect that bias, _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate).
-For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the $b_{curr}$ is not zero.
-Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits.
-For each scenario, we generated 1000 data sets and fit the model on those data sets.
-For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
-For each replicate, we calculated the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.
-
-Additionally, we present sensitivity analysis for the number of available strikes per trial (6 strikes, 10 strikes, and 15 strikes for each attorney).
-Thus, in total, we considered 4851 different scenarios, i.e.
-the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data, 11 different values for the power prior weight parameter, and for 3 combinations of the total number of strikes.
-
-```{r figbd9515, out.width = "95%", fig.cap = "Bias Detection Rates, 95 pct credible interval, 15 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp15_95CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-Figure \@ref(fig:figbd9515) depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
-When $b_{curr}$ is high (close to 3), the bias detection rate is high (very close to 1).
-However, when $b_{curr}$ is lower (close to zero), the bias detection rate drops below 0.8.
-When there is no bias in the current or historical trials ($b_{curr} = b_{hist} = 0$), the bias detection rate for the 95% credible interval is 0.05 or less.
-As we place more weight on the historical strike data (as $\alpha$ increases), and the historical strike data is consistent with current information (high compatibility), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
-Conversely, as $\alpha$ decreases, bias detection is only high when there is high bias in the current trial ($b_{curr} \geq 1.5$).
-Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible.
-However, when current and historical trials are incompatible, both type I and type II error can be high if $\alpha$ is close to one.
-As we discuss in section \@ref(selecting_alpha) below, the choice of $\alpha$ depends on assumptions about the degree of incompatibility of current and historical trials.
-
-```{r figbd9510, out.width = "95%", fig.cap = "Bias Detection Rates (95 pct credible interval, 10 strikes scenario)", fig.align='center'}
-
-load_path <- here::here("figures","pp10_95CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-```{r figbd956, out.width = "95%", fig.cap = "Bias Detection Rates based on 95 pct credible interval for 6 strike scenario", fig.align='center'}
-
-load_path <- here::here("figures","pp6_95CI.png")
-knitr::include_graphics(load_path)
-
-```
-
-As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases.
-Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
-
-The choice of credible interval also affects bias detection.
-For example, with a 90\% credible interval, when $b_{curr}=2$, the bias-detection rate is lower than with an 80\% credible interval; when $b_{curr}=0$, that detection rate is higher.
-The tradeoff is lower accuracy.
-With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
-When the bias parameter is 3, however, the detection rate is still high.
-For details on simulation results for 90% and 80% credible intervals, see the Appendix.
-
-Bias detection will also be sensitive to the choice of the standard deviation of the initial prior ($\sigma$).
-The results above depend in part on our choice of a conservative initial prior for the bias parameter (Equation \@ref(eq:initialprior)).
-Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate.
-As a result, our model will only detect severe bias.
-This sensitivity, however, reduces after including the historical data, especially when $\alpha$ is close to 1.
-
-```{r figsdprior, out.width = "95%", fig.cap = "Sensitivity to Initial Prior's Standard Deviation", fig.align='center'}
-
-load_path <- here::here("figures","SD_sensitivity.jpg")
-knitr::include_graphics(load_path)
-
-```
-
-Figure \@ref(fig:figsdprior) depicts how assigning different values for the initial prior’s standard deviation ($\sigma = \{1,2,100\}$) affects the bias detection rate at the 95\% credible interval for three $\alpha$ values ($\alpha = \{0.1,0.5,1\}$) under scenarios where each attorney had 10 strikes and where $b_{curr} = b_{hist}$.
-In our simulations, we set $\sigma = 2$ for our initial prior.
-If we instead select $\sigma = 1$, that matters most when $\alpha = 0.1$, i.e., when we borrow less information from the historical data.
-As $\alpha$ increases, the initial prior's standard deviation matters less.
-Moreover, when we increased the initial prior's standard deviation to make it much less informative ($\sigma = 100$), that had little impact on bias detection rates, regardless of the choice of $\alpha$.
-
-```{r figcmptbly, out.width = "95%", fig.cap = "Compatibility", fig.align='center'}
-
-load_path <- here::here("figures","compatibility_plot.jpg")
-knitr::include_graphics(load_path)
-
-```
-
-Finally, Figure \@ref(fig:figcmptbly) summarizes our simulation results for bias detection at the 95% credible interval, given perfect compatibility ($b_{curr} = b_{hist}$).
-These results indicate that our model can accurately detect strong bias if present in both the current and historical trials.
-Given high compatibility, increasing $\alpha$ improves bias detection, as does increasing the number of historical trials and more strikes per trial.
-Under the low number of strikes per trial (n = 6), bias detection requires strong bias in the current trial and a high $\alpha$.
-
-## Simultaneous Strikes {#simul-strikes}
-
-In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
-This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation \@ref(eq:model3)), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
-The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
-
-To test this premise, we conducted the following simulation study.
-We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 30 prospective jurors, 15 of which were member of a cognizable-class (e.g., Black jurors).
-Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
-We then fit the model to the 50 trials to examine whether the estimated bias parameters across the 50 simulated trials were equivalent.
-
-```{r fig-simul, out.width = "75%", fig.cap="Simulation results for simultaneous strikes example.
-Columns correspond to values of bias ($b_{curr}$) used to simulate data.
-Points and horizontal lines denote estimated posterior mean and 95 pct.
-credible interval for bias parameter.
-Vertical dashed lines denote true value of $b_{curr}$ in the simulations.", fig.align='center'}
-
-load_path <- here::here("figures","simultaneous_sims_plot.jpg")
-knitr::include_graphics(load_path)
-
-```
-
-Figure \@ref(fig:fig-simul) depicts the results of the simultaneous strike simulations.
-We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
-None of their credible intervals include zero.
-This shows that the order of strikes does not influence the estimate of the bias parameter.
-Accordingly, the simultaneous-strikes procedure can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
-
-## The Software Prototype {#prototype}
-
-We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app with `r R.version$version.string` and the shiny package [@R; @shiny].
-Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-As in the simulations, a normal distribution was assumed for the distribution of $b$.
-The length of the Markov chain was 110000 and the first 10,000 iterations were dropped as burn-in.
-No thinning was performed, as the correlation was weak and convergence occurred rapidly.
-We checked convergence with both traceplots and the Gelman-Rubin statistic [@Gelman1992].
-Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current strike data only _versus_ current plus historical strike data).
-
-To show how lawyers might use this app in real cases in real time during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
-Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later de-identified them.
-For details on this dataset, see the Appendix.
-
-With the app, the user first selects the cognizable class.
-For the prototype, we set two cognizable-class options: race and gender.
-This was because the standard juror questionnaire in this court did not ask about other possible cognizable classes.
-The data also did not expressly indicate the race or gender of the criminal defendant(s).
-For the levels of race and gender, we were limited by choices in the juror questionnaire (e.g., for gender, "male" and "female").
-
-Next, the user enters by hand the strike information in the case before them in the _strike tally_ table (Figure \@ref(fig:figapp1), top left).
-In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
-the user can add or delete rows to the strike tally as needed.
-
-```{r inpts}
-
-df <- data.frame(round = c(1:10),
-                 num_cog = c(5,5,4,3,2,2,2,2,2,2),
-                 total = c(21,20,19,18,17,16,15,14,13,12),
-                 cog = c(0,1,1,0,0,1,1,0,0,0),
-                 party = rep(c("PP","PD"),5))
-
-knitr::kable(df, caption ="Hypothetical Strike Tally", format = "latex", booktabs = TRUE)
-
-```
-
-To illustrate, suppose the strike tally in Table 1 depicts the pattern of strikes in the present case with defense attorney Aaron Waelchi and prosecutor Lawrance Klocko V (both aliases for actual attorneys in the historical strike data).
-After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.
-Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \@ref(fig:figapp1)(a)).
-Here, the 95\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally alone.
-In the default setting, the pull-down menus for prosecutor and defense are set to "None".
-As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for this prosecutor and defense attorney based only on the strike tally data and the initial prior ($b \sim N(0,2)$).
-
-```{r figapp1, out.width='50%', fig.cap = "Screenshots of R-Shiny application showing density plots for race bias of prosecutor and defense attorney.
-Vertical dotted lines depict 95 pct.
-credible interval.", fig.align='center', fig.ncol = 2, fig.show = 'hold', fig.subcap=c("Current Strike Data only","Historical and current strike data")}
-
-load_path1 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot1.png")
-load_path2 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")
-
-par(mfrow=c(2,1))
-knitr::include_graphics(load_path1)
-knitr::include_graphics(load_path2)
-
-```
-
-To use historical strike data, the user selects the name of the prosecutor or defense attorney from the pull-down menus.
-If an attorney's name cannot be found, the app has no historical strike data for that attorney.
-Once selected, the prior and posterior density plots automatically update to account for the pre-loaded historical strike data for that attorney.
-For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information ($\alpha = 1$).
-The user has two other options: half weight ($\alpha = 0.5$) and minimal weight ($\alpha = 0.2$).
-
-In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to "Equal".
-The density plots update accordingly (Figure \@ref(fig:figapp1)(b)).
-Now the credible intervals clearly exclude zero.
-Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
-In the Appendix, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
-
-```{r figapp2, fig.cap = "Density Plots with Historical Strike Data (Race).
-Screenshot of R-Shiny application show density plots based on historical and current strike data for prosecutor and defense attorney.", fig.align='center', eval = FALSE}
-
-load_path2 <- here::here("figures","batson_app_screenshots", "batson_app_screenshot2.png")
-
-knitr::include_graphics(load_path2)
-
-```
-
-# Discussion
-
-Statistical methods for _Batson_ and similar challenges often use historical strike data to infer illegal strike bias.
-The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, this approach makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
-Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them  decide whether to make a  _Batson_ challenge at all.
-If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and for ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
-
-## Selecting the Power Prior Value {#selecting_alpha}
-
-In practice, the anticipated degree of incompatibility should influence which value of $\alpha$ to select for the power prior.
-Given perfect compatibility ($\lvert b_{curr} - b_{hist} \rvert = 0$), we should let $\alpha = 1$ (equal weight between current and historical data).
-This will maximize bias detection even when the number of strikes per trial is small.
-This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd956) along the positive diagonals of the individual blocks.
-
-Bias detection also increases with the number of historical trials and the number of strikes per trial.
-Higher values of $\alpha$ lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \@ref(fig:figcmptbly)).
-The choice of $\alpha$ matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \@ref(fig:figcmptbly)).
-Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
-But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current bias ($b_{curr}$) is low; and bias detection can be (erroneously) low where current bias is high but bias in historical trials ($b_{hist}$) is low.
-
-Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-Our prototype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.
-If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then one should worry more about incompatibility and choose a lower $\alpha$ value accordingly.
-
-This approach also applies when the unit of analysis is not the attorney, but the law firm or office.
-To illustrate, suppose prosecutors from the same office were all exposed to training that recommended striking Black jurors.
-If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the "district attorney's office").
-Thereafter, any down-weighting adjustment (choosing $\alpha < 1$) depends on assumptions about how dissimilarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibility, notwithstanding the prosecutors' common exposure to the training.
-If none, one could assign equal weight to the historical data ($\alpha$ = 1).
-If one allows for some incompatibility nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
-Or one could keep all the historical data; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to different $\alpha$ values.
-
-In general, by their choice of $\alpha$, the lawyer explicitly and transparently weights the historical strike data in the analysis.
-This transparency thwarts lawyer efforts to strategically select $\alpha$ values.
-Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation).
-Future work might examine how various criteria for deriving a guide value of $\alpha$ [e.g.
-@Ibrahim2015, p.
-3724-3738] could appropriately apply in this context while also considering incompatibility (incommensurability) in selecting $\alpha$ [@Ollier2020].
-Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.
-@Han2023; @Pawel2022].
-
-## Modeling the Strike Procedure
-
-The Bayesian approach here also requires a model to match the strike procedure of the court of interest.
-This choice of model can matter in at least three ways.
-First, it can affect the missingness of historical strike data.
-For example, given the model we used (Equation \@ref(eq:model1)), we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
-Such information may be missing from jury selection records of past cases.
-
-Second, the choice of model can affect prospective data collection not only by courts but also by courtroom observers.
-The public has a presumptive First Amendment right to attend jury voir dire in criminal cases.
-Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from voir dire [@press1984; @presley2010].
-As a result, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time.
-Thus, such courtroom observers can prospectively collect strike data that they would have been less likely to obtain from court records after the fact.
-
-Third, the choice of model matters if historical strike data for a particular attorney covers strikes exercised in multiple courts with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \@ref(simul-strikes) ).
-In other cases, however, one cannot apply the same model to courts with different jury selection practices.
-For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
-
-# Conclusion
-
-In this paper, we extended a Bayesian approach to estimate attorney strike bias in the use of peremptory challenges by incorporating historical data on that attorney's use of strikes in past cases.
-In so doing, we used the power prior to adjust the weight of such historical information.
-Our simulations showed that how well our approach detects attorney bias depends on the number of past trials in which that attorney used strikes; the number of strikes per trial; and the degree of incompatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
-Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.
-
-# Acknowledgements
-
-We thank the Associate Editor and two anonymous reviewers for their comments.
-
-# Colophon
-
-Figures were produced with ggplot2 [@ggplot2] or Tikz [@Tantau2022].
-
-# CRediT Author Statement.
-
-Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Data Visualization, Writing - original draft.
-
-# Disclosure Statement
-
-The authors report there are no competing interests to declare.
-
-# References
-
-<div id="refs"></div>
-
-# Appendix
-
-Here, we extend section \@ref(simulation-results) by reporting our simulation results for 90% and 80% credible intervals.
-We also extend the discussion in section \@ref(prototype) by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
-
-## Bias Detection with 90% and 80% Credible Intervals
+# Bias Detection with 90% and 80% Credible Intervals
 
@@
 
-## Prototype Software Application
+# Prototype Software Application
 
-### Data Collection
+## Data Collection
 
@@
 While the standard forms included a signature line for the attorney, many were left blank or filled with illegible signatures.
-Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the US Attorney's office) or the criminal defendant(s) on the date(s) of jury selection.
+Accordingly, we turned to the publicly-available docket sheets for each case for the names of the lawyers who appeared in the case on behalf of the prosecution (the United States Attorney's office) or the criminal defendant(s) on the date(s) of jury selection.
 Where only a single attorney represented a party during jury selection, it was easy to attribute the pattern of strikes to that attorney.
@@
 
-### Additional Illustrations
+## Additional Illustrations
 
-Suppose the same scenario discussed in section \@ref(prototype), except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
+Suppose the same scenario discussed in section 3.2 of the paper, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
 If so, we select "gender" as the cognizable class, and the density plots update accordingly (Figure \@ref(fig:figapp3)(left)).
@@
 Both the prior and posterior density become flatter and the credible interval plainly includes zero.
+
+# References
+
+<div id="refs"></div>

### 2023-05-26 analysis/paper_draft4supp.tex
split paper and supplement into separate files (paper_draft4.Rmd and paper_draft4supp.Rmd)

--- 2023-05-26 analysis/paper_draft4.tex
@@
 \hypersetup{
-  pdftitle={Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data},
-  pdfkeywords={Batson challenge, peremptory strikes, power prior, Bayesian},
+  pdftitle={Supplementary Material for ``Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data''},
   hidelinks,
@@
 {
-  \title{\bf Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data}
+  \title{\bf Supplementary Material for ``Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data''}
 
@@
   \begin{center}
-    {\LARGE\bf Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data}
+    {\LARGE\bf Supplementary Material for ``Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data''}
   \end{center}
@@
 \bigskip
-\begin{abstract}
-United States law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, sex, or membership in certain other cognizable classes.
-Here, we extend a Bayesian approach for detecting such illegal strike bias by showing how to incorporate historical data on an attorney's use of peremptory strikes in past cases.
-In so doing, we use the power prior to adjust the weight of such historical information in the analysis.
-Using simulations, we show how the choice of the power prior's discounting parameter influences bias detection (how likely the credible interval for the bias parameter excludes zero), depending on the degree of incompatibility between current and historical trial data.
-Finally, we extend this approach with a prototype software application that lawyers could use to detect strike bias in real time during jury-selection.
-We illustrate this application's use with real historical strike data from a convenience sample of cases from one court.
-\end{abstract}
 
 \noindent%
-{\it Keywords:} Batson challenge, peremptory strikes, power prior, Bayesian
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-In the United States, individuals selected for jury service appear in court as scheduled and are questioned by the parties' attorneys and the trial judge.
-During this process, a prospective juror, if not excused by the trial judge for cause, may still be dismissed if a party's attorney uses one of their limited number of peremptory challenges against them.
-By asserting a peremptory challenge, a party can declare a prospective juror ineligible (``strikes'' that juror) for a seat on the jury without the burden of explaining why.
-
-Since \emph{Batson v.
-Kentucky} (1986), a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes.
-The party bringing a \emph{Batson} challenge bears the burden of proving such illegal bias is more likely than not to be true (LaFave et al.
-2022, \(\S\) 22.3(d)).
-
-A \emph{Batson} challenge typically proceeds in a four-step sequence.
-First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
-Second, if challenged, the challenger must meet an initial burden (the \emph{prima facie} case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
-Third, the striking attorney must proffer permissible reasons for their challenged strikes.
-Fourth, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
-
-For similar challenges under State law, a few States require only that an ``objective observer'' or an ``objectively reasonable person'' would find that race, ethnicity or another cognizable-class was a ``factor'' in that party's use of strikes (Wash.
-General Rule 37(e); Calif.
-Code of Civil Procedure \(\S\) 231.7(d)(1); Conn.
-Superior Court Rule 5-12(d)).
-Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes (Williams 2017).
-
-In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases (\emph{Flowers v.
-Mississippi} (2019), p.~2243; Wash.
-General Rule 37(g)(v); Calif.
-Code Civil Procedure \(\S\) 231.7(d)(3)(G)).
-Prior studies have collected such historical strike data in past cases and reported the observed difference in strike rates by race or sex.
-Typically, they test for the probability of observing a non-zero difference in strike rates by the race or sex of the struck prospective jurors, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference (e.g., Eisenberg 2017; Grosso and O'Brien 2012; for discussion, see Gastwirth and Xu 2014, 289--97; Gastwirth and Yu 2013).
-Prior studies have also modeled how much a prospective juror's race affected the odds of being struck.
-For different modeling approaches using the same historical strike data from Mississippi, see Craft (2018); DeCamp (2021); and Dunn and Zhuo (2022).
-
-Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction (National Center for State Courts: Center for Jury Studies 2018).
-Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
-Such incompatibility may occur if missing historical strike is missing not at random due to incomplete or inaccessible jury-selection records (Grosso and O'Brien 2017; Wright, Chavis, and Parks 2018).
-
-Incompatibility may also occur if attorney strike bias depends on certain trial-level characteristics (e.g., defendant race, charge severity) that take on one value in the current trial but that vary across the historical trials.
-For example, suppose an attorney who is more likely to strike Black prospective jurors because the defendant is Black; who has zero strike bias in cases with white defendants; and whose historical strike data comes from past trials with an equal number of Black and white defendants.
-If so, then inferring strike bias in the current trial with a Black defendant may depend in part on whether we use all the historical data or only the subset of past trials with Black defendants.
-
-As a result, for any given estimate of strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power (Bennett et al.
-2021).
-Other methods to increase power may in turn require adjusting for incompatibility implicitly.
-For example, Gastwirth and Xu (2014, 295--97) stratify strike data by trial and apply the Cochran-Mantel-Haenszel test, albeit after first selecting only historical strike data from past trials that are ``similar'' in their bias-salient characteristics to the current trial {[}p.~289, 297{]}.
-This initial step presumably aims to reduce incompatibility due to variation in those salient characteristics.
-Any such choice of ``similar'' historical strike data would need to be disclosed and justified.
-
-In this paper, we extend a Bayesian approach to \emph{Batson} and similar challenges (Kadane 2021, 2018b, 2018a) to incorporate historical strike data in a manner that allows for transparently adjusting for the assumed degree of incompatibility between current and historical trials.
-We proceed in three steps.
-First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
-Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
-In so doing, we use the power prior (Chen and Ibrahim 2000; Ibrahim et al.
-2015; for a gentle introduction, see Viele et al.
-2014).
-The power prior raises the likelihood of the historical strike data to a fixed power (\(\alpha\)), typically between zero and one, to down-weight the historical strike data in accord with expected incompatibility between the historical trials and the current trial.
-Zero denotes no weight to the historical data (for complete incompatibility) and one denotes equal weight (for complete compatibility).
-Third, we use this posterior distribution as an informative prior for the bias parameter in the current trial.
-
-We demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can ``detect'' attorney strike bias, i.e., whether the credible interval for the bias parameter excludes zero, given different values of the power prior's alpha parameter.
-Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court.
-With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\linewidth]{../figures/batson_workflow_fig2}
-
-}
-
-\caption{Inference from Historical Strike Data at Key Steps of Batson Challenge. Filled circles indicate steps for which bias estimates could be used.}\label{fig:batsonworkflowfig}
-\end{figure}
-
-To be sure, neither this nor any other statistical approach can automatically detect \emph{Batson} violations, because \emph{Batson} law permits any evidence of strike bias, not just strikes in past cases.
-And \emph{Batson} law requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the \emph{Batson} violation).
-Still, \emph{before} accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's \(\alpha\) parameter.
-After testing for sensitivity of bias detection to different fixed \(\alpha\) values for the power prior (Carvalho and Ibrahim 2021, 5252; Ibrahim et al.
-2015, 3734), attorneys can then combine the resulting inference with other evidence to decide whether to argue \emph{Batson}; to prove the \emph{Batson} prima facie case; and to persuade the trial judge to find a \emph{Batson} violation (Figure \ref{fig:batsonworkflowfig}).
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{statistical-procedure}{%
-\subsection{Statistical Procedure}\label{statistical-procedure}}
-
-Following others (e.g., Kadane 2018b; Barrett 2007; Gastwirth 2005, 183), we model a peremptory-challenge process in which each party strikes prospective jurors in an alternating sequence.
-Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptory strikes.
-Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that corresponds to the number seats on the jury (plus alternates, if any).
-The parties exercise their strikes on anyone among this subset of potential jurors in an alternating sequence.
-Once all strikes are used or waived, the remaining prospective jurors are each assigned to seats on the jury plus a few alternate jurors.
-
-Accordingly, for any given case \(i\) in which jury selection occurs, let \(j\) denote a peremptory strike used, and let \(\delta_{ij}\) denote whether or not a party used that strike on a person who belongs to a ``cognizable class''.
-If ``race'' is the bias type of interest, the cognizable class is racial minority jurors (\(\delta_{ij} =1\), 0 for White jurors).
-If ``sex'' is the bias of interest, the cognizable class is female jurors (\(\delta_{ij} = 1\), 0 for male jurors).
-In turn, let \(c_{ij}\) denote the number of cognizable class members subject to strike; and \(m_{ij}\) denote the number of cognizable class non-members subject to strike, such that \(c_{ij}+m_{ij}\) is the total number of jurors potentially subject to strike.
-If there is no bias, the probability is \(\frac{c_{ij}}{c_{ij}+m_{ij}}\) for striking a cognizable class member, and \(1-\frac{c_{ij}}{c_{ij}+m_{ij}}\) for striking someone who does not belong to that class.
-
-By adding one parameter \(w\), we can measure \emph{bias} by different values of \(w\) by defining the probability of a cognizable class member being struck to be \(\frac{wc_{ij}}{wc_{ij}+m_{ij}}\).
-To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let \(b = log(w)\).
-
-Accordingly, for any given value of the bias parameter \(b\), the probability of strike of a member from either class, or \(Pr(\delta_{ij})\), is such that:
-
-\begin{equation}
-Pr(\delta_{ij} | b) = 
-\begin{cases}
-  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
-  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  \label{eq:model1}
-\end{cases} 
-\end{equation} This Equation \eqref{eq:model1} is equivalent to
-
-\begin{equation}
-Pr(\delta_{ij}| b) = \left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
-\label{eq:model2}
-\end{equation}
-
-Given the strike data we have, i.e., \(\delta_{ij}\), \(c_{ij}\), and \(m_{ij}\), by estimating the value of \(b\), we can measure bias when a party is striking potential jurors.
-If \(b = 0\), there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.
-If \(b>0\), we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
-Where \(b<0\), the party has a preference for a juror within the cognizable class.
-
-The likelihood function of \(b\) is
-
-\begin{equation}  
-L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}\left(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{\delta_{ij}} \left(\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}}\right)^{1-\delta_{ij}}
-\label{eq:model3}
-\end{equation} where \(n_{i}\) is the total number of jury selections (trials), \(n_j\) is the total number of peremptory strikes, and \(\delta=(\delta_{11},\delta_{12},\dots,\delta_{{n_i}{n_j}})\)
-
-\hypertarget{incorporating-historical-strike-data}{%
-\subsection{Incorporating historical strike data}\label{incorporating-historical-strike-data}}
-
-We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter.
-To do this, we introduce the power prior:
-
-\begin{equation}  
-\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
-\label{eq:powerprior}
-\end{equation} where \(0 \le \alpha_0 \le 1\) is the parameter controlling the weight of the historical information; \(\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})\) is the observed historical data; \(L(b|\delta_0)\) is the likelihood function of \(b\) given the historical data; and \(\pi_0(b)\) is the initial prior before the historical data is observed.
-
-In this paper, we assume that the law for \emph{Batson} and similar challenges entails a weakly-informative initial prior:
-
-\begin{equation}  
-\pi_0(b) \sim Normal(\mu = 0, \sigma = 2)
-\label{eq:initialprior}
-\end{equation}
-
-For this initial prior, we let \(\mu = 0\), because the law assigns the burden of proof in a \emph{Batson} challenge to the party bringing the challenge.
-Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven.
-This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
-Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias.
-For this reason, we use a normal (Gaussian) distribution with \(\sigma = 2\).
-This prior accords with Kadane (2021, 51), who suggests a prior for estimating \emph{Batson} strike bias that is at least unimodal, symmetric, and not dependent on the data.
-
-After including the historical information through the power prior, the posterior distribution of \(b\) is proportional to the product of likelihood function of \(b\) and the power prior of \(b\) is
-
-\begin{equation}
- \pi(b|\delta,\delta_0) \propto L(b|\delta)L(b|\delta_0)^{\alpha_0}\exp\left(-\frac{(b-\mu)^2}{2\sigma^2}\right)
-  \label{eq:model4}
-\end{equation} where
-
-\begin{equation}
-L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}\left(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{\delta_{0ij}} \left(\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}}\right)^{1-\delta_{0ij}}
-\label{eq:model5}
-\end{equation} is the likelihood function of \(b\) given historical data; \(n_{0i}\) is the total number of jury selections (trials) in the historical data; \(n_{0j}\) is the total number of peremptory strikes in the historical data; and \(\delta_{0ij}\) denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.
-
-\hypertarget{simulation-results}{%
-\subsection{Model Performance on Simulated Data}\label{simulation-results}}
-
-To evaluate the proposed method, we conducted a simulation study using Stan 2.261, a Hamiltonian Monte Carlo engine for Bayesian inference by way of R version 4.2.2 and RStan and RStan (Stan Development Team 2022).
-These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.
-As introduced above, \(b > 0\) represents a prosecutor's bias against a prospective juror within the cognizable class, while \(b = 0\) denotes no bias.
-For simplicity, we assume the defense attorney has no bias.
-
-Because we use the power prior, we can control how much we account for the historical strike data by modifying the discounting parameter \(\alpha\) (denoted as \(\alpha_0\) in Equations \eqref{eq:powerprior} and \eqref{eq:model4}).
-If \(\alpha=1\), the historical strike data is equally weighted with the data on strikes in the current trial.
-If \(\alpha<1\), the historical data are discounted and weighted proportionally less than the current trial data.
-We assign \(\alpha<1\) to evaluate how sensitive the posterior of the bias parameter is to the historical data on strikes.
-Accordingly, in the simulation study, we evaluated different values of \(\alpha\): \(\alpha=\{0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1\}\) to show how that parameter affected bias detection.
-
-We generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one, two, or three previous trials) using seven different values of the bias parameter for generating that historical data \(b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}\).
-
-To generate data, we generated strike data for a current trial using seven different values of the bias parameter \(b_{curr}=\{0,0.5,1,1.5,2, 2.5,3\}\).
-To evaluate the strength of the power prior as a function of the degree of incompatibility between current and historical strike data (\(\lvert b_{curr} - b_{hist} \rvert > 0\)), we assigned \(b_{hist}\) the same bias values as \(b_{curr}\), i.e., \(b_{hist}=\{0,0.5,1,1.5,2, 2.5,3\}\).
-We did this to simulate noise in the historical strike data that can arise due to not-at-random missingness or because a confounder (e.g.~defendant race, charge severity) affects \(b_{curr}\) but not of all the past trials in the historical strike data.
-
-For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., \(b_{curr} > 0\) , based on whether credible intervals of the bias parameter exclude zero.
-Since we use a weakly-informative prior that pulls the estimate of \(b_{curr}\) towards zero, we focus on how accurately our model can detect that bias, \emph{not} on how accurately the model can recover the true value of \(b_{curr}\) (i.e., a traditional coverage rate).
-For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the \(b_{curr}\) is not zero.
-Put another way, if we set \(b_{curr}>0\), we calculate the proportion of the 95\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits.
-For each scenario, we generated 1000 data sets and fit the model on those data sets.
-For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
-For each replicate, we calculated the posterior mean and \(80\%\), \(90\%\), and \(95\%\) highest posterior density intervals for \(b_{curr}\).
-
-Additionally, we present sensitivity analysis for the number of available strikes per trial (6 strikes, 10 strikes, and 15 strikes for each attorney).
-Thus, in total, we considered 4851 different scenarios, i.e.~the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data, 11 different values for the power prior weight parameter, and for 3 combinations of the total number of strikes.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-Figure \ref{fig:figbd9515} depicts the bias detection rate with an 95\% credible interval for the 15 strikes scenario.
-When \(b_{curr}\) is high (close to 3), the bias detection rate is high (very close to 1).
-However, when \(b_{curr}\) is lower (close to zero), the bias detection rate drops below 0.8.
-When there is no bias in the current or historical trials (\(b_{curr} = b_{hist} = 0\)), the bias detection rate for the 95\% credible interval is 0.05 or less.
-As we place more weight on the historical strike data (as \(\alpha\) increases), and the historical strike data is consistent with current information (high compatibility), the bias detection rate is close to 1, especially when bias is high (upper right-hand corner of plots).
-Conversely, as \(\alpha\) decreases, bias detection is only high when there is high bias in the current trial (\(b_{curr} \geq 1.5\)).
-Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible.
-However, when current and historical trials are incompatible, both type I and type II error can be high if \(\alpha\) is close to one.
-As we discuss in section \ref{selectingalpha} below, the choice of \(\alpha\) depends on assumptions about the degree of incompatibility of current and historical trials.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.95\linewidth]{../figures/pp15_95CI}
-
-}
-
-\caption{Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 15 strikes.
-Rows of larger squares correspond to combinations of alpha values (0.1, 0.5, and 1) for low, median and equal weight of historical data to current trial data.
-Columns of larger squares correspond to 1, 2, or 3 historical trials.
-Small squares correspond to combinations of bias in current (rows) and historical (columns) trials.}\label{fig:figbd9515}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.95\linewidth]{../figures/pp10_95CI}
-
-}
-
-\caption{Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.}\label{fig:figbd9510}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.95\linewidth]{../figures/pp6_95CI}
-
-}
-
-\caption{Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 6 strikes.}\label{fig:figbd956}
-\end{figure}
-
-As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \ref{fig:figbd9510} and \ref{fig:figbd956} respectively), bias detection decreases.
-Bias detection is only high when \(\alpha\) is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
-
-The choice of credible interval also affects bias detection.
-For example, with a 90\% credible interval, when \(b_{curr}=2\), the bias-detection rate is lower than with an 80\% credible interval; when \(b_{curr}=0\), that detection rate is higher.
-The tradeoff is lower accuracy.
-With a 90\% credible interval, when the bias parameter is 2, the model is less accurate in detecting bias.
-When the bias parameter is 3, however, the detection rate is still high.
-For details on simulation results for 90\% and 80\% credible intervals, see the Supplementary Material.
-
-Bias detection will also be sensitive to the choice of the standard deviation of the initial prior (\(\sigma\)).
-The results above depend in part on our choice of a conservative initial prior for the bias parameter (Equation \eqref{eq:initialprior}).
-Given little historical strike data for an attorney and a low number of strikes per trial, this initial prior will dominate.
-As a result, our model will only detect severe bias.
-This sensitivity, however, reduces after including the historical data, especially when \(\alpha\) is close to 1.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{../figures/SD_sensitivity}
-
-}
-
-\caption{Sensitivity to Initial Prior's Standard Deviation (SD) by level of discounting (alpha).
-Simulations were run with 3 levels of the initial prior SD: low (SD = 1), medium (SD = 2), and hight (SD = 100).
-Bias detection rate corresponds to the proportion of simulations in which bais was detected based on 0 being excluded from the 95 percent credible interval.
-Bias was equal in current and historical trials and three levels were selected: 0 (no bias scenario), 1, and 2.
-In general, bias detection is relatively unaffected by choice initial SD, especially when alpha is close to 1.}\label{fig:figsdprior}
-\end{figure}
-
-Figure \ref{fig:figsdprior} depicts how assigning different values for the initial prior's standard deviation (\(\sigma = \{1,2,100\}\)) affects the bias detection rate at the 95\% credible interval for three \(\alpha\) values (\(\alpha = \{0.1,0.5,1\}\)) under scenarios where each attorney had 10 strikes and where \(b_{curr} = b_{hist}\).
-In our simulations, we set \(\sigma = 2\) for our initial prior.
-If we instead select \(\sigma = 1\), that matters most when \(\alpha = 0.1\), i.e., when we borrow less information from the historical data.
-As \(\alpha\) increases, the initial prior's standard deviation matters less.
-Moreover, when we increased the initial prior's standard deviation to make it much less informative (\(\sigma = 100\)), that had little impact on bias detection rates, regardless of the choice of \(\alpha\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.85\linewidth]{../figures/compatibility_plot}
-
-}
-
-\caption{Bias detection under compatibility scenarios (equal bias in current and historical trials) for different numbers of strikes (columns), sizes of historical data (rows) and levels of alpha (shapes).
-As the weight applied to historical data increases (higher alpha), bias detection increases.
-Bias detection rates also increase as the number of strikes and historical datasets increase.}\label{fig:figcmptbly}
-\end{figure}
-
-Finally, Figure \ref{fig:figcmptbly} summarizes our simulation results for bias detection at the 95\% credible interval, given perfect compatibility (\(b_{curr} = b_{hist}\)).
-These results indicate that our model can accurately detect strong bias if present in both the current and historical trials.
-Given high compatibility, increasing \(\alpha\) improves bias detection, as does increasing the number of historical trials and more strikes per trial.
-Under the low number of strikes per trial (n = 6), bias detection requires strong bias in the current trial and a high \(\alpha\).
-
-\hypertarget{simul-strikes}{%
-\subsection{Simultaneous Strikes}\label{simul-strikes}}
-
-In some courts, both parties simultaneously exercise their peremptory challenges on the prospective jurors subject to strikes.
-This simultaneous-strikes process can be modeled as a special case of the model of an alternating-strikes process above (see Equation \eqref{eq:model3}), i.e, as equivalent to one party engaging in an uninterrupted sequence of strikes against a subset of prospective jurors eligible to be struck.
-The premise: Regardless of the order in which a party announced those strikes, the posterior for the bias parameter would be the same.
-
-To test this premise, we conducted the following simulation study.
-We first generated a single trial in which one attorney used fifteen strikes in an uninterrupted sequence against 30 prospective jurors, 15 of which were member of a cognizable-class (e.g., Black jurors).
-Then, we shuffled the order of those strikes to generate 50 trials with the same proportion of struck cognizable-class members but different orders.
-We then fit the model to the 50 trials to examine whether the estimated bias parameters across the 50 simulated trials were equivalent.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{../figures/simultaneous_sims_plot}
-
-}
-
-\caption{Simulation results for simultaneous strikes example.
-Columns correspond to values of bias ($b_{curr}$) used to simulate data.
-Points and horizontal lines denote estimated posterior mean and 95 pct.
-credible interval for bias parameter.
-Vertical dashed lines denote true value of $b_{curr}$ in the simulations.}\label{fig:fig-simul}
-\end{figure}
-
-Figure \ref{fig:fig-simul} depicts the results of the simultaneous strike simulations.
-We find the estimated bias parameter of the 50 trials with different strike orders are close to each other for all of the three scenarios.
-None of their credible intervals include zero.
-This shows that the order of strikes does not influence the estimate of the bias parameter.
-Accordingly, the simultaneous-strikes procedure can be modeled as a special case of our initial model of an alternating-strikes process with an identical likelihood.
-
-\hypertarget{prototype}{%
-\subsection{The Software Prototype}\label{prototype}}
-
-We describe here a prototype software application (``app'') that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app with R version 4.3.0 (2023-04-21 ucrt) and the shiny package (R Core Team 2022; Chang et al.
-2021).
-Unlike the simulations, we built the app using Rcpp (Eddelbuettel and Balamuta 2018) and the Metropolis-Hasting algorithm to sample \(b\) from the posterior distribution.
-As in the simulations, a normal distribution was assumed for the distribution of \(b\).
-The length of the Markov chain was 110000 and the first 10,000 iterations were dropped as burn-in.
-No thinning was performed, as the correlation was weak and convergence occurred rapidly.
-We checked convergence with both traceplots and the Gelman-Rubin statistic (Gelman and Rubin 1992).
-Computation times vary from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current strike data only \emph{versus} current plus historical strike data).
-
-To show how lawyers might use this app in real cases in real time during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
-Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later assigned them aliases using the charlatan package (Chamberlain and Voytovich 2020).
-For details on this dataset, see the Supplementary Material.
-
-Next, the user enters by hand the strike information in the case before them in the \emph{strike tally} table (Figure \ref{fig:figapp1}, top left).
-In the strike tally, \texttt{round} denotes the order of strikes, \texttt{num\_cog} denotes the number of prospective jurors that could be struck that belong to the cognitive class; \texttt{total} denotes the total number of prospective jurors that could be struck; \texttt{cog} indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and \texttt{party} indicates which side used the strike (PP = prosecutor, PD = defense attorney).
-the user can add or delete rows to the strike tally as needed.
-
-\begin{table}
-
-\caption{\label{tab:inpts}Hypothetical Strike Tally}
-\centering
-\begin{tabular}[t]{rrrrl}
-\toprule
-round & num\_cog & total & cog & party\\
-\midrule
-1 & 5 & 21 & 0 & PP\\
-2 & 5 & 20 & 1 & PD\\
-3 & 4 & 19 & 1 & PP\\
-4 & 3 & 18 & 0 & PD\\
-5 & 2 & 17 & 0 & PP\\
-\addlinespace
-6 & 2 & 16 & 1 & PD\\
-7 & 2 & 15 & 1 & PP\\
-8 & 2 & 14 & 0 & PD\\
-9 & 2 & 13 & 0 & PP\\
-10 & 2 & 12 & 0 & PD\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-To illustrate, suppose the strike tally in Table 1 depicts the pattern of strikes in the present case with defense attorney Aaron Waelchi and prosecutor Lawrance Klocko V (both aliases for actual attorneys in the historical strike data).
-After the user enters this strike tally, the app initially displays two graphs -- one for the prosecution and the defense.
-Each graph depicts the prior density plot (colored light grey) and posterior density plot (blue and red for defense and prosecution, respectively) for the bias parameter (Figure \ref{fig:figapp1}(a)).
-Here, the 95\% credible interval includes zero, indicating no credible basis to infer bias, given the current strike tally alone.
-In the default setting, the pull-down menus for prosecutor and defense are set to ``None''.
-As a result, the app ignores any historical strike data and estimates the posterior distributions of the bias parameter for this prosecutor and defense attorney based only on the strike tally data and the initial prior (\(b \sim N(0,2)\)).
-
-\begin{figure}
-
-{\centering \subfloat[Current Strike Data only\label{fig:figapp1-1}]{\includegraphics[width=0.5\linewidth]{../figures/batson_app_screenshots/batson_app_screenshot1} }\subfloat[Historical and current strike data\label{fig:figapp1-2}]{\includegraphics[width=0.5\linewidth]{../figures/batson_app_screenshots/batson_app_screenshot2} }
-
-}
-
-\caption{Screenshots of R-Shiny application showing density plots for race bias of prosecutor and defense attorney.
-Vertical dotted lines depict 95 pct.
-credible interval.
-Blue and red density plots represent draws from posterior distributions of defense attorneys and prosecutors respectively.
-Grey background density plots show the initial priors (before using current strike pattern).
-(a) shows the result when no historical data are included, and (b) shows the result when historical data are included with equal weight to current data ('Strike History Weight' set to 'Equal', i.e., alpha = 1).}\label{fig:figapp1}
-\end{figure}
-
-To use historical strike data, the user selects the name of the prosecutor or defense attorney from the pull-down menus.
-If an attorney's name cannot be found, the app has no historical strike data for that attorney.
-Once selected, the prior and posterior density plots automatically update to account for the pre-loaded historical strike data for that attorney.
-For the weight to assign that attorney's historical strike data, the default is set to equal weight of historical information and current information (\(\alpha = 1\)).
-The user has two other options: half weight (\(\alpha = 0.5\)) and minimal weight (\(\alpha = 0.2\)).
-
-In our illustration, we select the names of the prosecutor and defense attorney from their respective pull-down menus; and leave the weight setting to ``Equal''.
-The density plots update accordingly (Figure \ref{fig:figapp1}(b)).
-Now the credible intervals clearly exclude zero.
-Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
-In the Supplementary Material, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Statistical methods for \emph{Batson} and similar challenges often use historical strike data to infer illegal strike bias.
-The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, this approach makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a \emph{Batson} challenge or not.
-Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them decide whether to make a \emph{Batson} challenge at all.
-If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and for ultimately proving illegal bias (Figure \ref{fig:batsonworkflowfig}).
-
-\hypertarget{selectingalpha}{%
-\subsection{Selecting the Power Prior Value}\label{selectingalpha}}
-
-In practice, the anticipated degree of incompatibility should influence which value of \(\alpha\) to select for the power prior.
-Given perfect compatibility (\(\lvert b_{curr} - b_{hist} \rvert = 0\)), we should let \(\alpha = 1\) (equal weight between current and historical data).
-This will maximize bias detection even when the number of strikes per trial is small.
-This effect of increased bias detection can be seen in the simulation results depicted in Figures \ref{fig:figbd9515}, \ref{fig:figbd9510}, and \ref{fig:figbd956} along the positive diagonals of the individual blocks.
-
-Bias detection also increases with the number of historical trials and the number of strikes per trial.
-Higher values of \(\alpha\) lead to higher bias detection rates across compatibility scenarios, especially in scenarios with a lower number of strikes per trial and few historical trials (top left panel of Figure \ref{fig:figcmptbly}).
-The choice of \(\alpha\) matters less for bias detection when bias is high; the number of strikes per trial is high; and the number of historical trials is high (lower right panel of Figure \ref{fig:figcmptbly}).
-Overall, given high compatibility, increasing \(\alpha\) will increase bias detection.
-But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current bias (\(b_{curr}\)) is low; and bias detection can be (erroneously) low where current bias is high but bias in historical trials (\(b_{hist}\)) is low.
-
-Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of \(\alpha\) by selecting multiple \(\alpha\) values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of \(\alpha\) value.
-Our prototype app permits this kind of testing of sensitivity in real time by providing multiple options for \(\alpha\) values.
-If the inference of bias is sensitive to the choice of \(\alpha\) values, all else equal, then one should worry more about incompatibility and choose a lower \(\alpha\) value accordingly.
-
-This approach also applies when the unit of analysis is not the attorney, but the law firm or office.
-To illustrate, suppose prosecutors from the same office were all exposed to training that recommended striking Black jurors.
-If so, our approach allows for (but does not require) aggregating all those prosecutors' cases, i.e., acting as if all prosecutors from that office were a single individual prosecutor (e.g., the strike bias of the ``district attorney's office'').
-Thereafter, any down-weighting adjustment (choosing \(\alpha < 1\)) depends on assumptions about how dissimilarities between the trial-level characteristics (e.g., defendant race, offense type) in the current trial and in past trials may still produce incompatibility, notwithstanding the prosecutors' common exposure to the training.
-If none, one could assign equal weight to the historical data (\(\alpha\) = 1).
-If one allows for some incompatibility nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
-Or one could keep all the historical data; choose an \(\alpha\) value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to different \(\alpha\) values.
-
-In general, by their choice of \(\alpha\), the lawyer explicitly and transparently weights the historical strike data in the analysis.
-This transparency thwarts lawyer efforts to strategically select \(\alpha\) values.
-Rather, at the key steps of the \emph{Batson} challenge process (Figure \ref{fig:batsonworkflowfig}), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of \(\alpha\).
-In any case, the choice of \(\alpha\) cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a \emph{Batson} violation).
-Future work might examine how various criteria for deriving a guide value of \(\alpha\) (e.g.
-Ibrahim et al.
-2015, 3724--38) could appropriately apply in this context while also considering incompatibility (incommensurability) in selecting \(\alpha\) (Ollier et al.
-2020).
-Knowing exactly how to quantify incompatibility and specify an appropriate prior for \(\alpha\) is an area of active research (e.g.
-Han, Ye, and Wang 2023; Pawel et al.
-2022).
-
-\hypertarget{modeling-the-strike-procedure}{%
-\subsection{Modeling the Strike Procedure}\label{modeling-the-strike-procedure}}
-
-The Bayesian approach here also requires a model to match the strike procedure of the court of interest.
-This choice of model can matter in at least three ways.
-First, it can affect the missingness of historical strike data.
-For example, given the model we used (Equation \eqref{eq:model1}), we required data on the cognizable-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike.
-Such information may be missing from jury selection records of past cases.
-
-Second, the choice of model can affect prospective data collection not only by courts but also by courtroom observers.
-The public has a presumptive First Amendment right to attend jury voir dire in criminal cases.
-Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from voir dire (\emph{{Press-Enterprise Co.
-v.
-Superior Court}} 1984; \emph{Presley v.
-Georgia} 2010).
-As a result, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time.
-Thus, such courtroom observers can prospectively collect strike data that they would have been less likely to obtain from court records after the fact.
-
-Third, the choice of model matters if historical strike data for a particular attorney covers strikes exercised in multiple courts with different strike procedures.
-In some cases, one court's strike procedure can be modeled as a special case of a more general model that can be applied to another court's procedure (e.g., the simultaneous strike example in section \ref{simul-strikes} ).
-In other cases, however, one cannot apply the same model to courts with different jury selection practices.
-For example, the model we used here does not apply to strike data generated by a court's strike procedure in which prospective jurors appear before the attorneys one at a time to be either struck or seated on the jury.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-In this paper, we extended a Bayesian approach to estimate attorney strike bias in the use of peremptory challenges by incorporating historical data on that attorney's use of strikes in past cases.
-In so doing, we used the power prior to adjust the weight of such historical information.
-Our simulations showed that how well our approach detects attorney bias depends on the number of past trials in which that attorney used strikes; the number of strikes per trial; and the degree of incompatibility, i.e., the distance between the bias parameters for the current trial and for the past trials.
-Finally, we discussed a prototype software application with which attorneys could use our approach in real time during jury selection to detect attorney strike bias.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We thank the Associate Editor and two anonymous reviewers for their comments.
-
-\hypertarget{colophon}{%
-\section{Colophon}\label{colophon}}
-
-Figures were produced with ggplot2 (Wickham 2016) or Tikz (Tantau 2022).
-
-\hypertarget{credit-author-statement}{%
-\section{CRediT Author Statement}\label{credit-author-statement}}
-
-Pandya: Conceptualization, Investigation, Software, Writing; Li: Conceptualization, Formal Analysis, Software, Writing - original draft; Barón: Formal Analysis, Software; Moore: Conceptualization, Supervision, Data Visualization, Writing - original draft.
-
-\hypertarget{disclosure-statement}{%
-\section{Disclosure Statement}\label{disclosure-statement}}
-
-The authors report there are no competing interests to declare.
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Barrett2007}{}}%
-Barrett, Bruce E. 2007. {``Detecting Bias in Jury Selection.''} \emph{The American Statistician} 61 (4): 296--301. \url{https://doi.org/10.1198/000313007X243629}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-batson1986}{}}%
-\emph{Batson v. Kentucky}. 1986. 476 U.S. 79. \url{https://www.loc.gov/item/usrep476079/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bennett2021}{}}%
-Bennett, Maxine, Simon White, Nicky Best, and Adrian Mander.
-2021.
-{``A Novel Equivalence Probability Weighted Power Prior for Using Historical Control Data in an Adaptive Clinical Trial Design: A Comparison to Standard Methods.''} \emph{Pharmaceutical Statistics} 20 (3): 462--84.
-\url{https://doi.org/10.1002/pst.2088}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-calccp2021}{}}%
-\emph{California Code of Civil Procedure §231.7}. 2022. \url{https://leginfo.legislature.ca.gov/faces/codes_displaySection.xhtml?sectionNum=231.7.\&nodeTreePath=4.3.1\&lawCode=CCP}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Carvalho2021}{}}%
-Carvalho, Luiz Max, and Joseph G. Ibrahim. 2021. {``On the Normalized Power Prior.''} \emph{Statistics in Medicine} 40 (24): 5251--75. \url{https://doi.org/10.1002/sim.9124}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-charlatan}{}}%
-Chamberlain, Scott, and Kyle Voytovich. 2020. \emph{Charlatan: Make Fake Data}. \url{https://CRAN.R-project.org/package=charlatan}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-shiny}{}}%
-Chang, Winston, Joe Cheng, JJ Allaire, Carson Sievert, Barret Schloerke, Yihui Xie, Jeff Allen, Jonathan McPherson, Alan Dipert, and Barbara Borges.
-2021.
-\emph{{shiny: Web Application Framework for R}}.
-\url{https://CRAN.R-project.org/package=shiny}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chen2000}{}}%
-Chen, Ming-Hui, and Joseph G. Ibrahim. 2000. {``Power Prior Distributions for Regression Models.''} \emph{Statistical Science} 15 (1): 46--60. \url{https://doi.org/10.1214/ss/1009212673}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Craft2018a}{}}%
-Craft, Will. 2018. {``{Peremptory Strikes in Mississippi's Fifth Circuit Court District}.''} \url{https://features.apmreports.org/files/peremptory_strike_methodology.pdf}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-DeCamp2021}{}}%
-DeCamp, Whitney.
-2021.
-{``Gender and the Peremptory Challenge: Separating the Effects of Race and Gender in Jury Selection.''} \emph{Women \& Criminal Justice} 31 (2): 159--73.
-\url{https://doi.org/10.1080/08974454.2020.1785372}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Dunn2022}{}}%
-Dunn, Jack, and Ying Daisy Zhuo. 2022. {``Detecting Racial Bias in Jury Selection.''} \emph{Operations Research Forum} 3 (3): 1--17. \url{https://doi.org/10.1007/s43069-022-00151-x}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rcpp}{}}%
-Eddelbuettel, Dirk, and James Joseph Balamuta.
-2018.
-{``{Extending \emph{R} with \emph{C++}: A Brief Introduction to \emph{Rcpp}}.''} \emph{The American Statistician} 72 (1): 28--36.
-\url{https://doi.org/10.1080/00031305.2017.1375990}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Eisenberg2017}{}}%
-Eisenberg, Ann M. 2017. {``{Removal of Women and African Americans in Jury Selection in South Carolina Capital Cases, 1997-2012}.''} \emph{Northeastern University Law Review} 9: 299--345.
-
-\leavevmode\vadjust pre{\hypertarget{ref-flowers2019}{}}%
-\emph{Flowers v. Mississippi}. 2019. 139 S. Ct. 2228.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gastwirth2005}{}}%
-Gastwirth, Joseph L.
-2005.
-{``{Case comment: statistical tests for the analysis of data on peremptory challenges: clarifying the standard of proof needed to establish a \emph{prima facie} case of discrimination in \emph{Johnson} v.
-\emph{California}}.''} \emph{Law, Probability and Risk} 4 (3): 179--85.
-\url{https://doi.org/10.1093/lpr/mgi014}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gastwirth2014}{}}%
-Gastwirth, Joseph L., and Wenjing Xu.
-2014.
-{``Statistical Tools for Evaluating the Adequacy of the Size of a Sample on Which Statistical Evidence Is Based.''} \emph{Law, Probability and Risk} 13 (3-4): 277--306.
-\url{https://doi.org/10.1093/lpr/mgu010}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gastwirth2013}{}}%
-Gastwirth, Joseph L., and Binbing Yu.
-2013.
-{``Case Comment: Appropriate Statistical Methodology Yields Stronger Evidence of Discriminatory Peremptory Challenges in {North Carolina}: Application to the {Randolph County} Data in \emph{North Carolina} v.
-\emph{Rouse} and Related Cases.''} \emph{Law, Probability and Risk} 12 (2): 155--64.
-\url{https://doi.org/10.1093/lpr/mgs030}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman1992}{}}%
-Gelman, Andrew, and Donald B. Rubin. 1992. {``{Inference from Iterative Simulation Using Multiple Sequences}.''} \emph{Statistical Science} 7 (4): 457--72. \url{https://doi.org/10.1214/ss/1177011136}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Grosso2012}{}}%
-Grosso, Catherine M., and Barbara O'Brien.
-2012.
-{``A Stubborn Legacy: The Overwhelming Importance of Race in Jury Selection in 173 Post-Batson North Carolina Capital Trials.''} \emph{Iowa Law Review} 97: 1531--59.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Grosso2017}{}}%
----------. 2017. {``A Call to Criminal Courts: Record Rules for Batson.''} \emph{Kentucky Law Journal} 105: 651.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Han2023}{}}%
-Han, Zifei, Keying Ye, and Min Wang.
-2023.
-{``A Study on the Power Parameter in Power Prior Bayesian Analysis.''} \emph{The American Statistician} 77 (1): 12--19.
-\url{https://doi.org/10.1080/00031305.2022.2046160}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ibrahim2015}{}}%
-Ibrahim, Joseph G., Ming-Hui Chen, Yeongjin Gwon, and Fang Chen.
-2015.
-{``The Power Prior: Theory and Applications.''} \emph{Statistics in Medicine} 34 (28): 3724--49.
-\url{https://doi.org/10.1002/sim.6728}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kadane2018b}{}}%
-Kadane, Joseph B.
-2018a.
-{``\emph{Batson} and Reverse-\emph{Batson} Motions in North Carolina: \emph{State v.
-Hurd} and \emph{State v.
-Tucker}.''} \emph{Law, Probability and Risk} 17 (3): 263--73.
-\url{https://doi.org/10.1093/lpr/mgy015}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kadane2018a}{}}%
----------. 2018b. {``Statistics for Batson Challenges.''} \emph{Law, Probability and Risk} 17 (1): 1--13. \url{https://doi.org/10.1093/lpr/mgx016}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kadane2021}{}}%
----------. 2021. {``{Priors neutral between the parties: the Batson motion in \emph{Idaho v. Ish}}.''} \emph{Law, Probability and Risk} 20 (1): 45--62. \url{https://doi.org/10.1093/lpr/mgab005}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-LaFave2022}{}}%
-LaFave, Wayne R., Jerold H. Israel, Nancy J. King, and Orin S. Kerr. 2022. \emph{Criminal Procedure}. 4th ed. Vol. 6.
-
-\leavevmode\vadjust pre{\hypertarget{ref-nscs-jury2018}{}}%
-National Center for State Courts: Center for Jury Studies. 2018. {``Comparative Data: Peremptory Challenges.''} 2018. \url{https://public.tableau.com/shared/BYJM8RF8Y?:showVizHome=no}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ollier2020}{}}%
-Ollier, Adrien, Satoshi Morita, Moreno Ursino, and Sarah Zohar.
-2020.
-{``An Adaptive Power Prior for Sequential Clinical Trials -- Application to Bridging Studies.''} \emph{Statistical Methods in Medical Research} 29 (8): 2282--94.
-\url{https://doi.org/10.1177/0962280219886609}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Pawel2022}{}}%
-Pawel, Samuel, Frederik Aust, Leonhard Held, and Eric-Jan Wagenmakers. 2022. {``Normalized Power Priors Always Discount Historical Data.''} arXiv. \url{https://doi.org/10.48550/ARXIV.2206.04379}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-presley2010}{}}%
-\emph{Presley v. Georgia}. 2010. 558 U.S. 209. \url{https://www.loc.gov/item/usrep558209/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-press1984}{}}%
-\emph{{Press-Enterprise Co. v. Superior Court}}. 1984. 464 U.S. 501. \url{https://www.loc.gov/item/usrep464501/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
-R Core Team. 2022. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rstan}{}}%
-Stan Development Team. 2022. {``{RStan}: The {R} Interface to {Stan}.''} \url{https://mc-stan.org/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Tantau2022}{}}%
-Tantau, Till. 2022. \emph{The TikZ and PGF Packages: Manual for Version 3.1.9a}. \url{https://github.com/pgf-tikz/pgf}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Viele2014}{}}%
-Viele, Kert, Scott Berry, Beat Neuenschwander, Billy Amzal, Fang Chen, Nathan Enas, Brian Hobbs, et al.
-2014.
-{``Use of Historical Control Data for Assessing Treatment Effects in Clinical Trials.''} \emph{Pharmaceutical Statistics} 13 (1): 41--54.
-\url{https://doi.org/doi.org/10.1002/pst.1589}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-WashR37}{}}%
-\emph{Washingon State General Rule 37}. 2022. \url{https://www.courts.wa.gov/court_rules/pdf/GR/GA_GR_37_00_00.pdf}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
-Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Williams2017}{}}%
-Williams, C. J. 2017. {``Proposing a Peremptory Methodology for Exercising Peremptory Strikes.''} \emph{American Criminal Law Review} 54: 277--302.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wright2018}{}}%
-Wright, Ronald F., Kami Chavis, and Gregory S. Parks. 2018. {``The Jury Sunshine Project: Jury Selection Data as a Political Issue.''} \emph{Univeristy of Illinois Law Review} 2018 (4): 1407--42.
-
-\end{CSLReferences}
-
-\hypertarget{supplementary-material-for-bayesian-detection-of-bias-in-peremptory-challenges-using-historical-strike-data}{%
-\section{Supplementary Material for ``Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data''}\label{supplementary-material-for-bayesian-detection-of-bias-in-peremptory-challenges-using-historical-strike-data}}
-
-Here, we extend section \ref{simulation-results} of the paper by reporting our simulation results for 90\% and 80\% credible intervals.
-We also extend the discussion in section \ref{prototype} of the paper by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
+Here, we extend section 3 of the paper by reporting our simulation results for 90\% and 80\% credible intervals.
+We also extend the discussion in section 3.2 of the paper by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
 
@@
 
-Suppose the same scenario discussed in section \ref{prototype}, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
+Suppose the same scenario discussed in section 3.2 of the paper, except now the strike tally indicates the pattern of strikes by these attorneys against female prospective jurors.
 If so, we select ``gender'' as the cognizable class, and the density plots update accordingly (Figure \ref{fig:figapp3}(left)).
@@
 
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-charlatan}{}}%
+Chamberlain, Scott, and Kyle Voytovich. 2020. \emph{Charlatan: Make Fake Data}. \url{https://CRAN.R-project.org/package=charlatan}.
+
+\end{CSLReferences}
+
 \end{document}

### 2023-05-26 analysis/memo2_TAS_reviewers.Rmd
template for reply to second round of reviewer comments

---
output: 
  pdf_document:
    citation_package: biblatex
    keep_tex: false
    latex_engine: pdflatex
    template: svm-latex-memo.tex
    number_sections: true
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes:
   - \linespread{1.05}

from: Authors
to: The American Statistican
subject: Reply to Second Round of Reviewer Comments
date: "`r format(Sys.Date(), '%B %d, %Y')`"
memorandum: true
graphics: false
width: .3
bibliography: references.bib
---

This memorandum replies to comments from the TAS editor, a TAS associate editor, and a reviewer received in May 2023.

# TAS Editor

_Do not hyphenate "jury service" on page 2._

We have edited accordingly.

_Are all Figures 2-4 necessary in the paper or can one/two of them be moved to the supplement?_

TBD

_On all figures, please make sure axes and labels are readable. For example, on your current Fig. 7, $b_cur \approx b$ is difficult to read, (and you use b_curr in the paper, not b_cur)._

TBD

_On your next revision, please submit the paper and supplement as separate files. Refer to figures in the supplement by using Figure S1, Figure S2, and so on. Include author information in the supplement file._

Done.

_The referee has suggested that the analysis of a real data set could be put in the supplement (free of space concerns), noting that this could be helpful for instructors of Bayesian methods in the classroom. I will leave that up to you._

See below.

## Associate Editor

_(page 1, lines 21-23) The authors include the ‘real time’ use of the application in the abstract but this isn’t fully demonstrated in the manuscript as written. I would like the authors to link to the application (perhaps they were waiting because the link is identifying). I also would like to see some step by step use of the application, noting the authors do provide screenshots in Figure 8 that probably give enough information for the manuscript._

We have now provided links to the app as well as the code used to generate the results. Instructions for using the app are now provided not only in section 3.2 of the paper and the Supplemental Materials, but also a vignette in the app itself.

_(page 5, lines 20-25) I think the point here can be made more convincingly for a statistical audience. That is, the problem I had with my initial reading was the bias shown in modeling that could be used against attorneys. However, with the clarifications this methodology is helpful for alerting about potentially biased strikes. These issues, if brought to the judge, are fully explored as explored on page 2 (lines 25-38). That is, errors of Type I are okay and errors of Type II can lead to missed bias. This thought process helped me reorient my initial reading of the paper._

TBD

_(page 6, lines 2-19) I found this paragraph difficult to read through once and understand. Specifically what is a “cognizable” class? Is it the majority or non-marginalized class? I think moving the (e.g.,) on lines 50-51 up to the first mention would help with this. Further, “the number of cognizable class members subject to strike” can probably be replaced with “number of cognizable class jurors”? Or maybe using nij and mij to connote these are number of jurors and not those struck. This is perhaps a bit too nitpicky, but I wrote it in the margins._

TBD

_(page 7, lines 31-54) This is a really good explanation. This may have got me thinking along the lines of comment 2 above._

Thanks!

_(page 9, lines 3-19) I’m still a little unsure exactly how bhist and bcurr are used to generate the data. I’m assuming you’re generating it via the probability model described in Equation (2) given those selected values, is that right?_

Yes, that's correct.

_(page 10, lines 25-29) Does the application allow users to revise the historical data used? That is, should there be guidance about choosing between low α with all the historical data or large α with a subset of the historical data? That is, it seems your approach can be compatible with the stratified approach of Gastwirth and Xu (2014) – would that improve the issue with heterogeneous historical data?_

TBD

## Reviewer 2

_On page 1, where you cite the Washington, Calif. and Conn. Codes stating that that membership in a “cognizable” class was a “factor”, you should also mention that in Miller-El, 545 U.S. the Supreme Court used the expression “significant factor”._

TBD

_On page 3, the initial prior given in (5) should be presented before the power prior in (4). Indeed, you state that the initial prior should be specified before the historical data is observed._

TBD

_Both Fig. 4 (and the surrounding discussion) and the discussion on p. 21, indicate that when there is high incompatibility between the historical and current data, the detections may well be erroneous. This is especially important in the situation where bcurr =0, i.e. no evidence of bias in the challenges in the present case. It is difficult for this reviewer to believe that judges would seriously entertain a Batson challenge, even if historical discrimination had occurred, when the data in the present case indicated the removal rates of protected and unprotected members of the venire were (essentially) equal. Of course, a defendant could raise a Batson challenge if they noticed that minorities with a certain characteristic were being challenged but majority members with same characteristic were not. Comparative evidence, rather than statistical, would be the basis of the ultimate decision on the Batson claim. Perhaps the discussion on p. 21 could discuss this further; especially as the reverse situation (serious bias in the current trial but little historically) is also important._

TBD

### 2023-06-16 analysis/R/simultaneous_plots.R
full code for simultaneous plots, with edits per reviewer comments

@@ -1,0 +1,56 @@
+
+# Code for generating plots of simultaneous strike data
+# output from the 'simultaneous.R' script. 
+#-----------------------------------------------------------------#
+
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(sjPlot)
+
+load("data/simul_4.19.22.RData")
+
+simul_1[[2]]
+simul_2[[2]]
+simul_3[[2]]
+
+
+sm11 <- simul_1[[2]]
+sm22 <- simul_2[[2]]
+sm33 <- simul_3[[2]]
+
+sm11$b <- "b_[curr] %~~% 1"
+sm22$b <- "b_[curr] %~~% 2"
+sm33$b <- "b_[curr] %~~% 3"
+
+comb_sm <- rbind(sm11, sm22, sm33)
+comb_sm$sim <- rep(factor(paste("",50:1, sep =""),
+                    levels = paste("",50:1, sep =""),ordered = TRUE), 
+                   3)
+
+b_set <- data.frame(mean_b = c(1, 2, 3), 
+                    b = c("b_[curr] %~~% 1", "b_[curr] %~~% 2", "b_[curr] %~~% 3"))
+
+ggplot(data = comb_sm)+
+  facet_grid(~b, labeller = label_parsed)+
+  geom_errorbarh(aes(y = sim,
+                     xmin = LB_CI_95,
+                     xmax = UB_CI_95), height = 0, 
+                 color = "grey")+
+  geom_point(aes(x = mean_b, y = sim))+
+  xlab("Posterior mean of b (bias parameter)")+
+  geom_vline(xintercept = 0, lty = 1, lwd = 1.1)+
+  ylab("Simulation Number")+
+  scale_x_continuous(breaks = c(0:6))+
+  theme_sjplot2()+
+  geom_vline(data = b_set, aes(xintercept = mean_b), color = "black", lty=2)+
+  theme(panel.grid = element_blank(), 
+        axis.title = element_text(size = 14), 
+        axis.text.x = element_text(size = 12), 
+        axis.text.y = element_text(size = 12), 
+        strip.background = element_rect(fill="grey95"))
+ggsave("simultaneous_sims_plot_23.jpg", width = 8, height = 8,
+       units = "in", dpi = 300, bg = "white")
+
+
+  

### 2023-06-16 analysis/R/simultaneous_plots.R
fixed b_curr to bcurr

@@ -19,9 +19,9 @@ sm11 <- simul_1[[2]]
 sm22 <- simul_2[[2]]
 sm33 <- simul_3[[2]]
 
-sm11$b <- "b_[curr] %~~% 1"
-sm22$b <- "b_[curr] %~~% 2"
-sm33$b <- "b_[curr] %~~% 3"
+sm11$b <- "b[curr] %~~% 1"
+sm22$b <- "b[curr] %~~% 2"
+sm33$b <- "b[curr] %~~% 3"
 
 comb_sm <- rbind(sm11, sm22, sm33)
 comb_sm$sim <- rep(factor(paste("",50:1, sep =""),
@@ -29,7 +29,7 @@ comb_sm$sim <- rep(factor(paste("",50:1, sep =""),
                    3)
 
 b_set <- data.frame(mean_b = c(1, 2, 3), 
-                    b = c("b_[curr] %~~% 1", "b_[curr] %~~% 2", "b_[curr] %~~% 3"))
+                    b = c("b[curr] %~~% 1", "b[curr] %~~% 2", "b[curr] %~~% 3"))
 
 ggplot(data = comb_sm)+
   facet_grid(~b, labeller = label_parsed)+

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
further responses to review comments

@@
 
-This memorandum replies to comments from the TAS editor, a TAS associate editor, and a reviewer received in May 2023.
+This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and a reviewer.
 
@@
 
-TBD
+We have moved Figure 3 to the supplement.
 
@@
 
-TBD
+We have revised the figures accordingly.  To make the labels more readable, we used a lighter color for them.  We also fixed the reference to $b_curr$.
 
@@
 
-Done.
+We have revised accordingly. [TBD]
 
@@
 
-We have now provided links to the app as well as the code used to generate the results.
-Instructions for using the app are now provided not only in section 3.2 of the paper and the Supplemental Materials, but also a vignette in the app itself.
+We have now provided links to the app as well as the code used to generate the simulation results.
+Instructions for using the app are now provided not only in section 3.2 of the paper and the Supplemental Materials, but also a vignette in the app itself [TBD].
 
@@
 This thought process helped me reorient my initial reading of the paper._
-
-TBD
 
@@
 
-TBD
+As we discuss in the paper (page X, lines YY-YY), how one chooses the subsetting criteria (e.g., "similar" cases) would have to be justified with respect to its _assumed_ effect on the degree of compatability between the historical trials and the current trial ($b_hist - $b_curr$).
+We stress _assumed_, because outside simulations like ours, neither $b_hist$ nor $b_curr$ are directly observable in actual cases.
+Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive to the posterior estimate of strike bias to different values of $\alpha$.
+The current version of the software prototype does not contain a feaure that lets users subset the historical data by some characteristic (e.g., date, defendant race).
+The app, however, in principle, could be modified to add that kind of subsetting feature as a step prior to estimating the posterior distribution for bias.
 
@@
 
-TBD
+Thank you for this comment.  We have revised the paper accordingly.
 
@@
 
-TBD
+Thank you for catching that. We have revised the paper accordingly.
 
@@
 
-TBD
+Thank you for this comment.  We have revised the Discussion to discuss this further. [TBD].
+
+First, we agree that high incompatbility may lead to error in bias detection.
+The challenge is that high incompatibility (in the case where $b_curr = 0$ or where $b_hist = 0$) is not directly observable.
+This is why we recommend using sensitivity to choice of $\alpha$ value to ascertain whether to worry about high incompatability.
+
+Second, we wish to address the idea that a judge may be reluctant to "seriously entertain a Batson challenge" if the strikes in the present case indicate essential equal "removal rates of protected and unprotected members of the venire." As an initial matter, we take "seriously entertain" to denote the premise that judge would be less likely to find a _Batson_ prima facie case and ultimately find a _Batson_ violation, given those removal rates.
+We do not dispute that premise about judges are likely to do.
+We d

### 2023-06-16 analysis/paper_draft4.Rmd
moved text on initial prior to come before introduction of power prior
add sentence to clarify bias
fixed minor error in reference to Rstan

@@
 
-## Incorporating historical strike data
-
-We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter.
-To do this, we introduce the power prior:
-
-\begin{equation}  
-\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
-\label{eq:powerprior}
-\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
-
 In this paper, we assume that the law for _Batson_ and similar challenges entails a weakly-informative initial prior:
@@
 
+## Incorporating historical strike data
+
+We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter.
+To do this, we introduce the power prior:
+
+\begin{equation}  
+\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
+\label{eq:powerprior}
+\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
+
 After including the historical information through the power prior, the posterior distribution of $b$ is proportional to the product of likelihood function of $b$ and the power prior of $b$ is
@@
 
-To evaluate the proposed method, we conducted a simulation study using Stan 2.261, a Hamiltonian Monte Carlo engine for Bayesian inference by way of R version 4.2.2 and RStan and RStan [@rstan].
+To evaluate the proposed method, we conducted a simulation study using Stan 2.261, a Hamiltonian Monte Carlo engine for Bayesian inference by way of R version 4.2.2 and RStan [@rstan].
 These simulations primarily show how much the size of the historical data affects our ability to detect bias and how much the power prior discounted the weight of the historical data.

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
additions to reviewer memo

@@
 
-We have revised accordingly. [TBD]
+We have revised accordingly.
 
@@
 
+We suspect the initial reading arose because "bias" has both a statistical meaning and a conventional meaning, and we are using the latter when referring to $b$ as the bias parameter.
+
 _(page 6, lines 2-19) I found this paragraph difficult to read through once and understand.
@@
 
-TBD
+By "cognizable" class, we mean the category to which the struck juror belongs that is the possible basis of strike bias.
+For example, if the set of prospective jurors consists only of Black and White individuals and the focus is on strike bias _against_ Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors".
+If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors." We used the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", because at the time any strike occurs, (1) no one is actually a juror, because no one has yet been seated onto the jury, and (2) not every prospective juror could have been struck.
 
@@
 
-As we discuss in the paper (page X, lines YY-YY), how one chooses the subsetting criteria (e.g., "similar" cases) would have to be justified with respect to its _assumed_ effect on the degree of compatability between the historical trials and the current trial ($b_hist - $b_curr$).
-We stress _assumed_, because outside simulations like ours, neither $b_hist$ nor $b_curr$ are directly observable in actual cases.
+Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$).
+As we discuss in the paper (page X, lines YY-YY), it depends on what one assumes about how much the subsetting criteria (e.g., only "similar" cases) changes the degree of incompatibility.
+We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases.
 Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive to the posterior estimate of strike bias to different values of $\alpha$.
-The current version of the software prototype does not contain a feaure that lets users subset the historical data by some characteristic (e.g., date, defendant race).
-The app, however, in principle, could be modified to add that kind of subsetting feature as a step prior to estimating the posterior distribution for bias.
+The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race).
+However, in principle, it could be modified to add that kind of feature.
 
@@
 
-Thank you for this comment.  We have revised the paper accordingly.
+Thank you for this comment.  We have revised the paper accordingly. [TBD]
 
@@
 
-Second, we wish to address the idea that a judge may be reluctant to "seriously entertain a Batson challenge" if the strikes in the present case indicate essential equal "removal rates of protected and unprotected members of the venire." As an initial matter, we take "seriously entertain" to denote the premise that judge would be less likely to find a _Batson_ prima facie case and ultimately find a _Batson_ violation, given those removal rates.
-We do not dispute that premise about judges are likely to do.
-We d
+Second, we wish to address the idea that a judge may be reluctant to "seriously entertain a Batson challenge" if the strikes in the present case indicate essential equal "removal rates of protected and unprotected members of the venire." we take "seriously entertain" to denote the premise that judge would be less likely to find a _Batson_ prima facie case and ultimately find a _Batson_ violation, given those observed "removal rates".
+We do not dispute that premise about judge behavior.

### 2023-06-16 analysis/paper_draft4.Rmd
added paragraph on incompatibility to discussion.

--- 2023-05-26 analysis/paper_draft4.Rmd
@@
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
+For clarity, we note that here, and throughout this paper, we are referring to bias as prejudice against selecting members of cognizable classes to serve on a jury and not bias in a statistical sense, or bias in a parameter.
+Rather, bias is the parameter we are trying to estimate.
+[EDITME]
 
@@
 51], who suggests a prior for estimating _Batson_ strike bias that is at least unimodal, symmetric, and not dependent on the data.
+
+## Incorporating historical strike data
+
+We incorporate data on strikes in past cases and allow for adjustment of the weight of that historical strike data on the posterior distribution of the bias parameter.
+To do this, we introduce the power prior:
+
+\begin{equation}  
+\pi(b|\delta_0,\alpha_0) \propto L(b|\delta_0)^{\alpha_0}\pi_0(b)
+\label{eq:powerprior}
+\end{equation} where $0 \le \alpha_0 \le 1$ is the parameter controlling the weight of the historical information; $\delta_0=(\delta_{011},\delta_{012},\dots,\delta_{0{n_i}{n_j}})$ is the observed historical data; $L(b|\delta_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.
 
@@
 
+[EDITME]Our simulations demonstrate that when bias is high in either the historical _or_ current data, bias can be detected, especially when $\alpha$ is close to 1.
+For example, when $b_{curr}$ = 0, $b_{hist}$ = 3, and alpha = 1 our simulation bias detection rate is 1, when there are two or three historical trials (Fig X).
+In a scenario where we are using real data, and the true values of alpha are not known, these represent cases of extreme incompatibility, and should be detectible through sentivity checking of the value of $\alpha$.
+For instance, when the value of $\alpha$ is set to 0.1, the bias detection rate goes down to 0.02 for both the two and three historical data set simulations.
+Such sensitivity to the value of $\alpha$ is indicative of incompatibility between current and historical data.
+
 In general, by their choice of $\alpha$, the lawyer explicitly and transparently weights the historical strike data in the analysis.

### 2023-06-16 analysis/paper_draft4.Rmd
edited bias disambiguation sentence

--- 2023-06-16 analysis/paper_draft4.Rmd
@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or certain other cognizable classes.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membershhip in another category that the law prohibits as a basis for striking that juror.
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
+Such illegal strike bias is the parameter of interest, not to be confused with "bias" in a statistical sense, or bias in a parameter [EDITME].
 

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
bias disambig and cognizable-class clarification

@@
 
-We suspect the initial reading arose because "bias" has both a statistical meaning and a conventional meaning, and we are using the latter when referring to $b$ as the bias parameter.
+We have revised the paper in paragraph 2 to expressly distinguish between "bias" as the illegal strike bias (the parameter to be estimated) from "bias" in the statistical sense.
+We hope this reduces any confusion.
 
@@
 
-By "cognizable" class, we mean the category to which the struck juror belongs that is the possible basis of strike bias.
+By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias.
 For example, if the set of prospective jurors consists only of Black and White individuals and the focus is on strike bias _against_ Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors".

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
response to extreme incompatibility scenario.

@@
 By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias.
-For example, if the set of prospective jurors consists only of Black and White individuals and the focus is on strike bias _against_ Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors".
-If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors." We used the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", because at the time any strike occurs, (1) no one is actually a juror, because no one has yet been seated onto the jury, and (2) not every prospective juror could have been struck.
+For example, if the focus is on strike bias against Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors".
+If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."
+
+We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", for two reasons.
+First, at the time any strike occurs, no one is a "juror", because no one has yet been seated onto the jury.
+Second, at the time any strike occurs, not every prospective juror could have been struck.
+As we assume in the paper (see pages XXX), only a subset of prospective jurors present during jury selection are ever subject to strike at any one time.
+The size of that subset is usually based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.
 
@@
 Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$).
-As we discuss in the paper (page X, lines YY-YY), it depends on what one assumes about how much the subsetting criteria (e.g., only "similar" cases) changes the degree of incompatibility.
+As we discuss in the paper (page X, lines YY-YY), it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility.
 We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases.
-Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive to the posterior estimate of strike bias to different values of $\alpha$.
+Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$.
 The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race).
@@
 
-First, we agree that high incompatbility may lead to error in bias detection.
-The challenge is that high incompatibility (in the case where $b_curr = 0$ or where $b_hist = 0$) is not directly observable.
-This is why we recommend using sensitivity to choice of $\alpha$ value to ascertain whether to worry about high incompatability.
-
-Second, we wish to address the idea that a judge may be reluctant to "seriously entertain a Batson challenge" if the strikes in the present case indicate essential equal "removal rates of protected and unprotected members of the venire." we take "seriously entertain" to denote the premise that judge would be less likely to find a _Batson_ prima facie case and ultimately find a _Batson_ violation, given those observed "removal rates".
-We do not dispute that premise about judge behavior.
+We take this comment to posit that a judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias against the cognizable class (e.g.
+Black prospective jurors), because $b = 0$ falls within the credible interval, even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias, i.e., $b = 0$ falls outside the credible interval.
+This might occur, for example, where an attorney is motivated to strike Black prospective jurors because they are Black _only_ in cases with Black defendants; all that attorney's past trials involved Black defendants; but the current trial has a White defendant.
+In this extreme scenario, we agree that a judge might properly worry about high incompatibility and not find a _Batson_ prima facie case for that reason, even though no one can directly observe or know with complete certainty whether this is a case where $b_curr = 0$ and $b_hist >> 0$.
+Or the judge may still find enough for a prima face case, but downweight the historical data when ultimately deciding whether illegal bias likely occurred.
+In the reverse extreme situation, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but does not detect illegal bias when $\alpha > 0$.
+Here, too, we agree that a judge might properly worry about high incompatibility, except that down-weighting the historical data here increases the support for finding at least enough a _Batson_ prima facie case.

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
minor edits

@@
 We have revised the paper in paragraph 2 to expressly distinguish between "bias" as the illegal strike bias (the parameter to be estimated) from "bias" in the statistical sense.
-We hope this reduces any confusion.
+We hope this reduces any confusion.[EDITME]
 
@@
 _(page 10, lines 25-29) Does the application allow users to revise the historical data used?
-That is, should there be guidance about choosing between low α with all the historical data or large α with a subset of the historical data?
+That is, should there be guidance about choosing between low $\alpha$ with all the historical data or large α with a subset of the historical data?
 That is, it seems your approach can be compatible with the stratified approach of Gastwirth and Xu (2014) – would that improve the issue with heterogeneous historical data?_

### 2023-06-16 analysis/paper_draft4.Rmd
minor edits

@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membershhip in another category that the law prohibits as a basis for striking that juror.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membership in another category that the law prohibits as a basis for striking that juror.
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].

### 2023-06-16 analysis/paper_draft4supp.Rmd
added Figure for 95% CI (ten strikes); changed Fig numbering.

@@
   - \usepackage{subfig}
+  - \newcommand{\beginsupplement}{
+    \setcounter{table}{0}  
+    \renewcommand{\thetable}{S\arabic{table}} 
+    \setcounter{figure}{0} 
+    \renewcommand{\thefigure}{S\arabic{figure}}
+    }
 #fontsize: 12pt
@@
 
-Here, we extend section 3 of the paper by reporting our simulation results for 90% and 80% credible intervals.
+\beginsupplement
+
+Here, we extend section 3 of the paper by reporting our simulation results for strike bias detection for the 95% credible interval (ten strikes per side) as well as such results for the 90% and 80% credible intervals for fifteen, ten, and six strikes per side.
 We also extend the discussion in section 3.2 of the paper by discussing data collection and presenting additional illustrations of how a lawyer might use the prototype app.
 
-# Bias Detection with 90% and 80% Credible Intervals
+# Additional Simulation Results for Strike Bias Detection
+
+The number of strikes can affect bias detection as we define it.
+The paper reports simulation results for bias detection for the 95% credible interval for trials with fifteen and six strikes per side.
+Here, Figure \@ref(fig:figbd9510) depicts the simulation results for trials with ten strikes per side.
+
+```{r figbd9510, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.", fig.align='center'}
+
+load_path <- here::here("figures","pp10_95CI.png")
+knitr::include_graphics(load_path)
+
+```
 

### 2023-06-16 analysis/paper_draft4.Rmd
moved Fig for 95%CI (ten strikes) to Supplement

--- 2023-06-16 analysis/paper_draft4.Rmd
@@
 
-Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
+Without historical strike data, there is a higher risk of underpowered and inflated estimates of an attorney's strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
 Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
@@
 Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court.
-With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.
+With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias during jury selection.
 
@@
 
-```{r figbd9510, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.", fig.align='center'}
-
-load_path <- here::here("figures","pp10_95CI.png")
-knitr::include_graphics(load_path)
-
-```
-
 ```{r figbd956, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 6 strikes.", fig.align='center'}
@@
 
-As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases.
-Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
+As we decrease the number of strikes each attorney has from fifteen to six (Figure \@ref(fig:figbd956)), bias detection decreases.
+(For the simulation results for ten strikes, see the Supplementary Material.) Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
 

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
revised "cog class" discussion

--- 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
@@
 
-This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and a reviewer.
+This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and an anonymous reviewer.
 
@@
 
+We now introduce the term "cognizable class" later, in section 2.
 By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias.
 For example, if the focus is on strike bias against Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors".
-If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."
+If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."  Because _Batson_ applies without regard to the "majority" or "marginalized" status of struck jurors (e.g., Black prospective jurors _and_ White prospective jurors), we have avoided those terms.
 
@@
 Second, at the time any strike occurs, not every prospective juror could have been struck.
-As we assume in the paper (see pages XXX), only a subset of prospective jurors present during jury selection are ever subject to strike at any one time.
+As we assume in the paper, only a subset of prospective jurors present during jury selection are ever subject to strike at any one time.
 The size of that subset is usually based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.
@@
 Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$).
-As we discuss in the paper (page X, lines YY-YY), it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility.
+As we discuss in the paper, it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility.
 We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases.

### 2023-06-17 analysis/memo2_TAS_reviewers.Rmd
added response to extreme incompatibility scenario

@@
 
-Thank you for this comment.  We have revised the Discussion to discuss this further. [TBD].
+We took this comment to have two possible intended meanings.
 
-We take this comment to posit that a judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias against the cognizable class (e.g.
-Black prospective jurors), because $b = 0$ falls within the credible interval, even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias, i.e., $b = 0$ falls outside the credible interval.
+First, we read this comment to posit that judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case where "data in the present case indicated the removal rates of protected and unprotected members of the venire were (essentially) equal." However, in some cases, bias can be detected even if equal strikes _rates_.
+To illustrate, consider the following hypothetical patterns of strikes by a prosecutor:
+
+```{r}
+
+round <- c(1:10)
+num_cog <- c(7.6,5,4,3,2,2,2,2,2)
+total <- c(30,28,26,24,22,20,18,16,14,12)
+cog <- c(1,1,1,1,1,0,0,0,0,0)
+
+df <- data.frame(round, num_cog, total, cog)
+
+knitr::kable(df)
+
+```
+
+Here, the strike rates for the cognizable class and the non-cognizable class are equal (5 out of 10).
+However, the 95% credible interval for the posterior estimate of $b$ is excludes zero, thus indicating illegal bias.
+
+Second, we took this comment to posit that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias against the cognizable class (e.g.
+Black prospective jurors), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias.
 This might occur, for example, where an attorney is motivated to strike Black prospective jurors because they are Black _only_ in cases with Black defendants; all that attorney's past trials involved Black defendants; but the current trial has a White defendant.
@@
 In the reverse extreme situation, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but does not detect illegal bias when $\alpha > 0$.
-Here, too, we agree that a judge might properly worry about high incompatibility, except that down-weighting the historical data here increases the support for finding at least enough a _Batson_ prima facie case.
+Here, too, we agree that a judge might properly worry about high incompatibility, except that down-weighting the historical data here increases the support for at least enough a _Batson_ prima facie case.

### 2023-06-17 analysis/paper_draft4.Rmd
added references to NJ, CT, Batson ultimate-inquiry.

@@
 nocite: |
-  @WashR37, @calccp2021
+  @WashR37, @calccp2021, @NJR183A, @CT512
 keywords: "Batson challenge, peremptory strikes, power prior, Bayesian"
@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membership in another category that the law prohibits as a basis for striking that juror.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membership in a cognizable class, i.e., a category that the law prohibits as a basis for striking that juror.
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
@@
 
-For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes (Wash.
+For _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." [@flowers2019, p.
+2241] In contrast, for similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes (Wash.
 General Rule 37(e); Calif.
 Code of Civil Procedure $\S$ 231.7(d)(1); Conn.
-Superior Court Rule 5-12(d)).
+Superior Court Rule 5-12(d); for a similar rule, see N.J.
+Court Rule 1:8-3A(d)).
 Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@Williams2017].
@@
 To be sure, neither this nor any other statistical approach can automatically detect _Batson_ violations, because _Batson_ law permits any evidence of strike bias, not just strikes in past cases.
-And _Batson_ law  requires the trial judge to weigh all such evidence to decide whether strike bias likely exists (the _Batson_ violation).
+And _Batson_ law  requires the trial judge to weigh all such evidence to decide whether illegal strike bias likely exists (the _Batson_ violation).
 Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
@@
 This will maximize bias detection even when the number of strikes per trial is small.
-This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515), \@ref(fig:figbd9510), and \@ref(fig:figbd956) along the positive diagonals of the individual blocks.
+This effect of increased bias detection can be seen in the simulation results depicted in Figures \@ref(fig:figbd9515) and \@ref(fig:figbd956) along the positive diagonals of the individual blocks.
 
@@
 For example, when $b_{curr}$ = 0, $b_{hist}$ = 3, and alpha = 1 our simulation bias detection rate is 1, when there are two or three historical trials (Fig X).
-In a scenario where we are using real data, and the true values of alpha are not known, these represent cases of extreme incompatibility, and should be detectible through sentivity checking of the value of $\alpha$.
+In a scenario where we are using real data, and the true values of $\alpha$ are not known, these represent cases of extreme incompatibility, and should be detectible through sensitivity checking of the value of $\alpha$.
 For instance, when the value of $\alpha$ is set to 0.1, the bias detection rate goes down to 0.02 for both the two and three historical data set simulations.
-Such sensitivity to the value of $\alpha$ is indicative of incompatibility between current and historical data.
+Such sensitivity to the choice of $\alpha$ value indicates incompatibility between current and historical data.
 

### 2023-06-18 analysis/paper_draft4.Rmd
edit to disambig sentence about "bias"; minor edits

@@
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
-Such illegal strike bias is the parameter of interest, not to be confused with "bias" in a statistical sense, or bias in a parameter [EDITME].
+Such illegal strike bias is the parameter of interest, not to be confused with "bias" in the estimator of a parameter.
 
@@
 Fourth, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
+For _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." [@flowers2019, p.
+2241]
 
-For _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." [@flowers2019, p.
-2241] In contrast, for similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes (Wash.
+In contrast, for similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class was a "factor" in that party's use of strikes (Wash.
 General Rule 37(e); Calif.
@@
 If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.
-If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
+If $b>0$, we infer that the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
 Where $b<0$, the party has a preference for a juror within the cognizable class.
@@
 We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges.
-We built this app with `r R.version$version.string` and the shiny package [@R; @shiny].
+We built this app with R version 4.2.2 and the shiny package [@R; @shiny].
 Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
@@
 
-To show how lawyers might use this app in real cases in real time during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
+To show how lawyers might use this app in real cases during jury selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017.
 Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later assigned them aliases using the charlatan package [@charlatan].
@@
 
-Next, the user enters by hand the strike information in the case before them in the _strike tally_ table (Figure \@ref(fig:figapp1), top left).
+To use the app, the user enters by hand the strike information in the case before them in the _strike tally_ table (Figure \@ref(fig:figapp1), top left).
 In the strike tally, `round` denotes the order of strikes, `num_cog` denotes the number of prospective jurors that could be struck that belong to the cognitive class; `total` denotes the total number of prospective jurors that could be struck; `cog` indicates whether the prospective juror actually struck in that round was a member of the cognizable class (1 = yes, 0 = no); and `party` indicates which side used the strike (PP = prosecutor, PD = defense attorney).
-the user can add or delete rows to the strike tally as needed.
+The user can add or delete rows to the strike tally as needed.
 
@@
 Now the credible intervals clearly exclude zero.
-Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how these attorney use their peremptory challenges in the present case.
+Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how this attorney uses their peremptory challenges in the present case.
 In the Supplementary Material, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
@@
 The Bayesian approach presented here accounts for available historical strike data when estimating the posterior for the bias parameter.
-By incorporating the power prior, this approach makes assumptions about the incompatibility between current and historical trial data an explicit part of the process of identifying bias and the decision of whether to bring a _Batson_ challenge or not.
-Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them  decide whether to make a  _Batson_ challenge at all.
-If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case and for ultimately proving illegal bias (Figure \@ref(fig:batsonworkflowfig)).
+By incorporating the power prior, this approach makes assumptions about the degree of incompatibility between current and historical trial data an explicit part of the process of identifying likely strike bias.
+Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them decide whether to bring a  _Batson_ challenge at all.
+If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case or for ultimately proving illegal strike bias (Figure \@ref(fig:batsonworkflowfig)).
 
@@
 Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-Our prototype app permits this kind of testing of sensitivity in real time by providing multiple options for $\alpha$ values.
+Our prototype app permits this kind of testing of sensitivity by providing multiple options for $\alpha$ values.
 If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then one should worry more about incompatibility and choose a lower $\alpha$ value accordingly.
@@
 In a scenario where we are using real data, and the true values of $\alpha$ are not known, these represent cases of extreme incompatibility, and should be detectible through sensitivity checking of the value of $\alpha$.
-For instance, when the value of $\alpha$ is set to 0.1, the bias detection rate goes down to 0.02 for both the two and three historical data set simulations.
+For instance, when the value of $\alpha$ is set to 0.1, the bias detection rate goes down to 0.02 for both the two and three historical data set simulations [INSERT REF TO FIGURE].
 Such sensitivity to the choice of $\alpha$ value indicates incompatibility between current and historical data.
@@
 Rather, at the key steps of the _Batson_ challenge process (Figure \@ref(fig:batsonworkflowfig)), the lawyers on both sides know that at any time the lawyers on the other side can always assess how sensitive any strike bias estimation is to the choice of $\alpha$.
-In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal bias is more likely than not (i.e., a _Batson_ violation).
+In any case, the choice of $\alpha$ cannot alone determine whether a trial judge finds that illegal strike bias is more likely than not (i.e., a _Batson_ violation).
 Future work might examine how various criteria for deriving a guide value of $\alpha$ [e.g.
@@
 Second, the choice of model can affect prospective data collection not only by courts but also by courtroom observers.
-The public has a presumptive First Amendment right to attend jury voir dire in criminal cases.
-Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from voir dire [@press1984; @presley2010].
-As a result, any person can often attend jury voir dire and collect relevant strike data based on what they observe in real time.
+The public has a presumptive First Amendment right to attend jury selection in criminal cases.
+Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from jury selection [@press1984; @presley2010].
+As a result, any person can often attend jury selection and collect relevant strike data based on what they observe in real time.
 Thus, such courtroom observers can prospectively collect strike data that they would have been less likely to obtain from court records after the fact.

### 2023-06-18 analysis/memo2_TAS_reviewers.Rmd
revised response to reviewer 2's comments 1, 3.

@@
 The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race).
-However, in principle, it could be modified to add that kind of feature.
+However, in principle, it could be modified to add that feature.
 
@@
 
-Thank you for this comment.  We have revised the paper accordingly. [TBD]
+We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." [@flowers2019, p.
+2241]
+
+The phrase "significant factor" does not appear in _Miller-El v.
+Dretke_, 545 U.S.
+231 (2005).
+At best, the Court wrote: "Comparing his strike with the treatment of panel members who expressed similar views supports a conclusion that race was significant in determining who was challenged and who was not.
+_Id._ at 252 (footnote omitted).
 
@@
 
-First, we read this comment to posit that judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case where "data in the present case indicated the removal rates of protected and unprotected members of the venire were (essentially) equal." However, in some cases, bias can be detected even if equal strikes _rates_.
-To illustrate, consider the following hypothetical patterns of strikes by a prosecutor:
+First, we read this comment to posit that judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case where "data in the present case indicated the removal rates of protected and unprotected members of the venire were (essentially) equal." However, in some cases, bias can be detected even if there are equal strike _rates_.
+To illustrate, consider the following hypothetical pattern of strikes by a prosecutor in a trial with a strike procedure like the one we model in the paper.
 
@@
 However, the 95% credible interval for the posterior estimate of $b$ is excludes zero, thus indicating illegal bias.
+Thus, in this case, the judge would err by relying on the equal strike rates to infer no strike bias.
 
-Second, we took this comment to posit that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias against the cognizable class (e.g.
-Black prospective jurors), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias.
+Second, we took this comment to say that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias (the credible interval includes zero) against some cognizable class (e.g.
+Black prospective jurors), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias (the credible interval excludes zero).
 This might occur, for example, where an attorney is motivated to strike Black prospective jurors because they are Black _only_ in cases with Black defendants; all that attorney's past trials involved Black defendants; but the current trial has a White defendant.

### 2023-06-19 analysis/paper_draft4.Rmd
add def of statistical bias

--- 2023-06-16 analysis/paper_draft4.Rmd
@@
 The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
-Such illegal strike bias is the parameter of interest, not to be confused with "bias" in a statistical sense, or bias in a parameter [EDITME].
+Such illegal strike bias is the parameter of interest, not to be confused with "bias" in a statistical sense (expected value deviating from the true value of a parameter), or bias in a parameter [EDITME].
 
@@
 
-Without historical strike data, there is a higher risk of underpowered and inflated estimates of an attorney's strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
+Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
 Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
@@
 Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court.
-With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias during jury selection.
+With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.
 
@@
 
+```{r figbd9510, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.", fig.align='center'}
+
+load_path <- here::here("figures","pp10_95CI.png")
+knitr::include_graphics(load_path)
+
+```
+
 ```{r figbd956, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 6 strikes.", fig.align='center'}
@@
 
-As we decrease the number of strikes each attorney has from fifteen to six (Figure \@ref(fig:figbd956)), bias detection decreases.
-(For the simulation results for ten strikes, see the Supplementary Material.) Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
+As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases.
+Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
 

### 2023-06-23 analysis/TAS_memo/memo_TAS_reviewers.Rmd
updated TAS reviewer memo, paper, references

--- 2023-02-08 analysis/TAS_memo/memo_TAS_reviewers.Rmd
@@
 width: .3
+bibliography: references.bib
 ---
 
-This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is much better as a result.
+This memorandum replies to comments from a TAS associate editor and two reviewers. We believe the revised paper is much better because of their comments.
 
@@
 As our revised Introduction now makes clear (p.
-4-5 and Fig.
-1), neither our statistical approach nor any other can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+4-5 and Figure 1), neither our statistical approach (nor any other) can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.
+
+    a.
 For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons.
@@
 In a _Batson_ challenge, the lawyers must disclose and justify their choice of $\alpha$ accordingly.
-We have revised the paper's Introduction and Discussion (sec 3.1) to emphasize this.
+We have revised the paper's Introduction and Discussion (section 4.1) to emphasize this.
 
@@
 Absent this transparency, lawyers will continue to argue strategically about how much to weigh historical strike data, except with the opacity of words.
-We have revised the paper to add this to the Discussion section (sec.
-3.1).
+We have revised the paper to add this to the Discussion section.
 
@@
 Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges.
-This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data.
-The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things.
-As we now note in the Discussion section of the paper, measuring incompatibility and choosing $\alpha$ in the face of incompatibility is an active area of research with no clear solution.
+This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data, which in turn likely varies by.
+among other things, strike procedure and courtroom practices.
+As we now note in the Discussion section of the paper, the literature on the power prior discusses information criteria for guide values of $\alpha$ as well as treating $\alpha$ as random and assigning a prior to it.
+However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons.
+First, unlike models of strike procedure discussed here, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., generalized linear models).
+In the clinical trial setting, it is easier to measure trial-level (and patient-level) characteristics that may be responsible for incompatibility.
+Even in these settings, selecting a prior for $\alpha$ is not straightforward.
+@Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate (more down-weighting) the influence of the historical data, and so one might have to use highly informative priors on the power parameter.
+These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis." More recent versions of the power prior are beginning to address this issue (e.g., using "commensurate power priors" [@Hobbs2011]).
+Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter.
+Third, whereas some papers treat incompatibility as itself a parameter [e.g., @Ollier2020], the precise approaches rely on strong assumptions that do not obviously make sense in our context.
+Accordingly, we believe the most conservative choice (for now) is to keep $\alpha$ a fixed value and check for sensitivity of results to the choice of $\alpha$ value.
+Proceeding otherwise would require the work of a completely new paper.
 
@@
 Thus, the larger the credible interval one chooses, the greater the chance that zero falls within it.
-For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on other credible intervals in the Appendix.
+For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on 90% and 80% credible intervals in the Supplementary Material.
 
 6.
-When we first submitted the paper, we offered to make the app prototype available to the reviewers (as an R package), so they could follow along with section 2.5 of the paper.
+When we first submitted the paper, we offered to make the app prototype available to the reviewers (as an R package), so they could follow along with section 3.2 of the paper.
+That section's primary purpose is to show with real data how our approach can be used in practice, including by lawyers in real time during jury selection.
 We received no response to this offer.
@@
 
+Fourth, the comment is overinclusive.
+Most work on the use of historical strike data (a) presupposes that _Batson_ law permits prior strikes as relevant evidence of bias in the current trial; (b) presupposes that any statistical approach therein cannot automatically determine _Batson_ violations alone; and (c) adjusts in some way for incompatibility between historical strike data and strikes from the current trial.
+As an example, we now point in the paper's Introduction to @Gastwirth2014.
+Our approach has a distinctive benefit of making incompatibility assumptions transparent via choosing and justifying a fixed value for the power prior's $\alpha$.
+
 _Page 4, line 5: The authors refer to gender but then provide sex options (male/female)._
@@
 We revised to remove such confusion.
+This includes changing the prototype software app to use 95% credible intervals.
 On the choice of credible intervals, see Item 1(5) above.
@@
 This comment is not clear to us.
-The real data analyses are based on real historical strike data from real criminal cases in the federal district court in Connecticut.
+In section 3.2, we base the real data analyses on real historical strike data from real criminal cases in the federal district court in Connecticut.
+On how we collected this data, see the paper's Supplementary Materials.
 If they wish, users of the app can run the model excluding any historical data (as this reviewer suggested) by not selecting the name of a prosecutor or defense attorney.
 Like our model, the app uses the $N(0,2)$ prior for the bias parameter.
-Similarly, the app user can alter the weight the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
+Similarly, the app user can alter the weight of the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter).
 Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2).
 See also Item 1(6).
+
+Similarly, although we did run simulations for when $\alpha = 0$, Figures 2-4 do not depict the bias detection rate for that scenario, simply because that scenario is equivalent to evaluating bias detection with only strike data from the current trial (no historical strike data) -- a topic of prior papers [e.g.
+@Kadane2018a].
 
@@
 In contrast, in our approach, the statistical model has to qualify as a valid model of the strike procedure of the court of interest.
-That strike procedure can and must be known and modeled before any strike data is collected or analyzed.
+That strike procedure must be known and modeled before any strike data is analyzed.
 More generally, our approach thwarts the strategic selection of $\alpha$ values by making the choice of $\alpha$ transparent.
-See Item 1(4) and section 3.2 of the revised Discussion.
+See Item 1(4) and section 4.2 of the revised Discussion.
 
@@
 
-We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_curr$); and (2) bias detection can be low where current bias is high but bias in historical trials is low.
-The tradeoff is that, given compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
+We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_{curr}$); and (2) bias detection can be low where current bias is high but bias in historical trials is low.
+The tradeoff is that, given high compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial.
 The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $\alpha$.
@@
 
-This information has been added to the text (sec. 2.5, p. 18).
-
-<!-- A normal distribution was assumed for the distribution of b.
-The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in.
-No thinning was performed, as the correlation was weak and the convergence occurred rapidly.
-Convergence was checked from both traceplots and the Gelman-Rubin statistic.
-Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (current data only _versus_ current and historical data).-->
+We added this information to the paper (p. 16-17).
 
@@
 
-See Items 1(2) and 1(4) above and section 3.1. of the revised Discussion.
+See Items 1(2) and 1(4) above and section 4.1 of the revised Discussion.
 
@@
 
-We appreciate this suggestion but we are worried that this would unduly increase the paper's length. Because of this concern, we have already moved other parts of the paper to the (online) Appendix.
+We appreciate this suggestion but we are worried that this would unduly increase the paper's length, particularly given the lack of a closed form solution for the posterior distribution even in a toy example.
+Because of this concern, we have already moved other parts of the paper to the Supplementary Materials.
+We intend the reader to gain understanding by following along using our prototype software application.
+We also now refer the reader to a gentle introduction to the power prior in @Viele2014.
 
@@
 Our statistical approach's validity turns neither on how many courts make available the kind of data it requires nor on proving how likely courts will start collecting such data.
-Indeed, as section 3.2 (paragraphs 1 and 2) of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right at attend jury selection in criminal court proceedings.
+Some courts do, which is how section 3.2 illustrates the approach with real data.
+Indeed, as section 4.2 of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right to attend jury selection in criminal court proceedings.
 
@@
 And that model explicitly accounts for the racial composition of the pool of prospective jurors at each strike opportunity.
+In any case, we re-ran our simulation and found that, as expected, changing the level of defense attorney bias does not alter the bias detection rates for prosecutors (see Table 1 here).
 
@@
 
-Thank you for this comment!  The paper now addresses this issue directly in the Discussion, section 3.1, paragraph 4.
+Thank you for this comment!  The paper now addresses this issue directly in the Discussion, section 4.1 (p. 21-22).
 
@@
 No, this is not correct.
-We have revised the paper to avoid such confusion.
+We have revised the paper and the software application to avoid such confusion.
 See Item 1(5) above.
-We agree that the choice of credible interval should be consistent within different stages of the _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
-[NOTE: Simulations use 95% credible interval, while app uses 80% credible intervals.]
+We agree that the choice of credible interval should be consistent across different stages of a _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.
 
@@
 The paper now cites to the model in Gastwirth (2005).
-On "data from an actual case," section 2.5 of the paper (our discussion about the prototype software application) expressly relies on actual strike data from a sample of criminal cases.
+On "data from an actual case," section 3.2 of the paper (our discussion about the prototype software application) expressly relies on actual strike data from a sample of actual criminal cases.
+As mentioned above, we intend the reader to gain understanding by following along using our prototype software application.
 
@@
 
-On the choice of $\alpha$, see Items 1(2) and 1(3) above and section 3.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
+On the choice of $\alpha$, see Items 1(2)-1(4) above and section 4.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).
 
@@
 
-Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical data in Table 8, however "unrealistic", to underscore that, as  _Batson_ law requires, our approach makes no different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.
+Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical strike tally in Table 1 (p.
+19), however "unrealistic", to underscore that, as  _Batson_ law requires, our approach makes no different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.
 

### 2023-06-23 analysis/memo2_TAS_reviewers.Rmd
updated TAS reviewer memo, paper, references

--- 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
@@
 
-We have moved Figure 3 to the supplement.
+We have moved what was previously Figure 3 to the supplement.
 
@@
 
-We have revised the figures accordingly.  To make the labels more readable, we used a lighter color for them.  We also fixed the reference to $b_curr$.
+We have revised the figures accordingly.  To make the labels more readable, we used a lighter color for them.  We also fixed the reference to $b_{curr}$.
 
@@
 
-See below.
+Section 3.2 of the paper and the related Supplementary Material illustrates our approach using our prototype software application and a _real_ data set of historical strike data.
+For details on this dataset, see section 2.1 of the Supplementary Material.
+Given our likelihood function (see Equation (3)), there is no closed-form solution for the posterior distribution of $b$.
+This is why we used our prototype software application to show how the posterior estimate of $b$ changes (see Figures 7 and S8).
 
-## Associate Editor
+We have made the app available to the public as a R package that includes the underlying dataset of historical strike data.
+We hope this will help statistics instructors in the way the TAS Associate Editor has in mind.
+However, because of time constraints, we are simply unable to provide in the paper's Supplement a complete tutorial tailored for classroom instruction by teachers of Bayesian statistics.
+On an introduction to a Bayesian approach to estimating illegal strike bias (albeit without historical strike data), we recommend @Kadane2018a and @Kadane2018b.
+
+# Associate Editor
 
@@
 
-We have now provided links to the app as well as the code used to generate the simulation results.
-Instructions for using the app are now provided not only in section 3.2 of the paper and the Supplemental Materials, but also a vignette in the app itself [TBD].
+The paper now contains links to the app as well as the code we used to generate the simulation results.
+We intend section 3.2 of the paper (and related Supplementary Material) as instructions for using the app.
+Those links are identifying, which is why we omitted them from the last draft of the paper.
 
@@
 
-We have revised the paper in paragraph 2 to expressly distinguish between "bias" as the illegal strike bias (the parameter to be estimated) from "bias" in the statistical sense.
-We hope this reduces any confusion.[EDITME]
+We appreciate this suggestion, but we worry that making this point here in terms of Type I and Type II error might lead to net confusion.
+The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more Type I error (falsely inferring illegal strike bias), because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence.
+Type I errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not.
+And Type I errors are worse still as the judge assigns more relative weight to any inference from historical strike data as compared to other items of evidence.
 
@@
 
-We now introduce the term "cognizable class" later, in section 2.
-By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias.
-For example, if the focus is on strike bias against Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors".
-If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."  Because _Batson_ applies without regard to the "majority" or "marginalized" status of struck jurors (e.g., Black prospective jurors _and_ White prospective jurors), we have avoided those terms.
+We have revised the paper to introduce the term "cognizable class" in section 2 instead of the Introduction.
+By "cognizable class", we mean the social category to which the struck juror belongs that the law accepts as a possible basis of illegal strike bias.
+For example, if the focus is on strike bias against Black prospective jurors because of their race, the cognizable class is "Black" prospective jurors.
+If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the cognizable class is "Black female" prospective jurors.
+And if the focus is on illegal strike bias against White prospective jurors because of their race, the cognizable class is "White" prospective jurors.
+Because _Batson_ law applies without regard to the "majority" or "marginalized" status of prospective jurors, we have avoided such terms.
 
-We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", for two reasons.
+We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors" for two reasons.
 First, at the time any strike occurs, no one is a "juror", because no one has yet been seated onto the jury.
@@
 As we assume in the paper, only a subset of prospective jurors present during jury selection are ever subject to strike at any one time.
-The size of that subset is usually based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.
+Although courts vary in practice, the size of that subset is often based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.
 
@@
 
-_(page 9, lines 3-19) I’m still a little unsure exactly how bhist and bcurr are used to generate the data.
+_(page 9, lines 3-19) I’m still a little unsure exactly how b_hist and b_curr are used to generate the data.
 I’m assuming you’re generating it via the probability model described in Equation (2) given those selected values, is that right?_
@@
 _(page 10, lines 25-29) Does the application allow users to revise the historical data used?
-That is, should there be guidance about choosing between low $\alpha$ with all the historical data or large α with a subset of the historical data?
+That is, should there be guidance about choosing between low $\alpha$ with all the historical data or large $\alpha$ with a subset of the historical data?
 That is, it seems your approach can be compatible with the stratified approach of Gastwirth and Xu (2014) – would that improve the issue with heterogeneous historical data?_
 
-Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$).
-As we discuss in the paper, it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility.
-We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases.
-Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$.
-The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race).
-However, in principle, it could be modified to add that kind of feature.
+Subsetting the historical data strike does not necessarily reduce incompatibility between the historical strike data and the current trial.
+Rather, it depends on what one _assumes_ about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) affects incompatibility.
+We stress "assumes", because the true values of $b_{hist}$ and $b_{curr}$ are not directly observable in actual cases.
+Accordingly, for any _subset_ of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$.
+The current version of the software prototype does not let users subset the historical data by some trial-level characteristic (e.g., defendant race).
+However, in principle, it could be modified to add that feature.
 
-## Reviewer 2
+_(page 14, Figure 5) the figure legend makes it challenging to compare between 1 and 2, though one could anticipate 2 is the dashed line._
+
+We have revised that figure legend to better distinguish between the different line types.
+
+_(page 16, line 53-54) 110000 doesn’t have a comma and 10,000 does._
+
+Thank you for catching this.  We have revised accordingly.
+
+_(page 19, Figure 8) This demonstration is very helpful for seeing the utility of the app.
+This makes me think of two possible improvements: 1.
+See point three of the other reviewer.
+This situation presents the opportunity to discuss any disagreements between current and historical data._
+
+We have revised the Discussion accordingly.  See also our reply to other reviewer below.
+
+_I understand the concern about paper length .
+.
+.
+There is value to the general audience for demonstrating posterior changes as the reviewer discusses.
+While this would make the paper long (and require too many figures), I do believe this would be helpful supplementary material._
+
+See above reply to TAS editor on this point.
+
+# Reviewer 2
 
@@
 
-Thank you for this comment.  We have revised the paper accordingly. [TBD]
+Thank you for suggesting this contrast between _Batson_ law and these State laws.
+We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." (@flowers2019, p.
+2241) We did not use the phrase "significant factor," because it does not appear in _Miller-El v.
+Dretke_, 545 U.S.
+231 (2005).
+At best, the Court there wrote: "Comparing his strike with the treatment of panel members who expressed similar views supports a conclusion that race was significant in determining who was challenged and who was not.
+_Id._ at 252 (footnote omitted).
 
@@
 
-Thank you for this comment.  We have revised the Discussion to discuss this further. [TBD].
+We took this comment to have two intended meanings.
 
-We take this comment to posit that a judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias against the cognizable class (e.g.
-Black prospective jurors), because $b = 0$ falls within the credible interval, even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias, i.e., $b = 0$ falls outside the credible interval.
-This might occur, for example, where an attorney is motivated to strike Black prospective jurors because they are Black _only_ in cases with Black defendants; all that attorney's past trials involved Black defendants; but the current trial has a White defendant.
-In this extreme scenario, we agree that a judge might properly worry about high incompatibility and not find a _Batson_ prima facie case for that reason, even though no one can directly observe or know with complete certainty whether this is a case where $b_curr = 0$ and $b_hist >> 0$.
-Or the judge may still find enough for a prima face case, but downweight the historical data when ultimately deciding whether illegal bias likely occurred.
-In the reverse extreme situation, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but does not detect illegal bias when $\alpha > 0$.
-Here, too, we agree that a judge might properly worry about high incompatibility, except that down-weighting the historical data here increases the support for finding at least enough a _Batson_ prima facie case.
+First, we read this comment to posit that a judge is unlikely to find even a _Batson_ prima facie case where "data in the present case indicated the removal rates of protected and unprotected members of the venire were (essentially) equal." However, in some cases, illegal strike bias can produce equal strike _rates_.
+To illustrate, consider the following hypothetical pattern of strikes by a prosecutor in a trial with a strike procedure like the one we model in the paper.
+
+```{r}
+#| label: hypodata
+#| echo: false
+
+round <- c(1:10)
+num_cog <- c(7,6,5,4,3,2,2,2,2,2)
+total <- c(30,28,26,24,22,20,18,16,14,12)
+cog <- c(1,1,1,1,1,0,0,0,0,0)
+
+df <- data.frame(round, num_cog, total, cog)
+
+knitr::kable(df, caption = "Hypothetical Strikes by Prosecutor")
+
+```
+
+Here, the strike rates for the cognizable class and the non-cognizable class are equal (5 out of 10).
+However, using the paper's approach and assuming no historical strike data (or $\alpha = 0$), we detect illegal strike bias, i.e., the 95% credible interval for the posterior estimate of $b$ excludes zero.
+The reason turns on the relatively few members of the cognizable class that could have struck in each round.
+Thus, in this hypothetical case, the judge would err by relying on equal strike rates to infer no illegal strike bias.
+
+Second, we took this comment to say that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal strike bias (the credible interval includes zero), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal strike bias (the credible interval excludes zero).
+We have revised the Discussion to further address high incompatibility.
+As we observe there, because no one can directly observe $b_{curr}$ and $b_{hist}$, they cannot know for sure whether their case involves high incompatibility.
+Nonetheless, if the posterior estimate of $b$ is highly sensitive to the choice of $alpha$ -- something one can observe -- then we agree that a judge might properly worry about erroneously detecting strike bias due to high incompatibility.
+As for what a judge might do in this situation, a judge might not find a _Batson_ prima facie case for that reason.
+Or the judge may still find enough for a prima face case, but down-weight the historical data (by selecting an $\alpha$ value less than 1) when ultimately deciding whether illegal strike bias likely occurred.
+In the reverse situation of high incompatibility, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but (erroneously) does not detect illegal bias when $\alpha > 0$.
+Here, we agree that a judge should worry about erroneously failing to detect strike bias due to high incompatibility.
+Down-weighting the historical data here increases the support for a finding that illegal strike bias likely occurred.

### 2023-06-23 analysis/paper_draft4.Rmd
updated TAS reviewer memo, paper, references

--- 2023-06-18 analysis/paper_draft4.Rmd
@@
 Since _Batson v.
-Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membership in a cognizable class, i.e., a category that the law prohibits as a basis for striking that juror.
-The party bringing a _Batson_ challenge bears the burden of proving such illegal bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
-Such illegal strike bias is the parameter of interest, not to be confused with "bias" in the estimator of a parameter.
+Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the United States Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, sex, or membership in another social category that the law prohibits as a basis for striking that juror.
+The party bringing a _Batson_ challenge bears the burden of proving such illegal strike bias is more likely than not to be true [@LaFave2022, $\S$ 22.3(d)].
+Such illegal strike bias is the parameter of interest, not to be confused with the bias of an estimator of a parameter (how much the estimator's expected value differs from the parameter's true value).
 
 A _Batson_ challenge typically proceeds in a four-step sequence.
-First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal bias.
-Second, if challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence of strike bias to raise a sufficient initial inference of illegal bias.
+First, the attorney must decide whether to challenge the opposing attorney's use of one or more peremptory strikes as based on illegal strike bias.
+Second, if challenged, the challenger must meet an initial burden (the _prima facie_ case) of producing just enough evidence to raise a sufficient initial inference of illegal strike bias.
 Third, the striking attorney must proffer permissible reasons for their challenged strikes.
 Fourth, the challenger must discredit those proffered reasons and ultimately persuade the trial judge that it is more likely than not that the striking attorney acted with illegal bias.
-For _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." [@flowers2019, p.
-2241]
+For _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." (@flowers2019, p.
+2241).
 
@@
 Court Rule 1:8-3A(d)).
-Peremptory-challenge procedure, and thus the task of proving such illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@Williams2017].
+Peremptory-challenge procedure, and thus the task of proving illegal strike bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@Williams2017].
 
-In any case, evidence of such illegal bias may include data on the use of peremptory challenges in past cases (@flowers2019, p.
+In any case, evidence of such illegal strike bias may include data on the use of peremptory challenges in past cases (@flowers2019, p.
 2243; Wash.
@@
 
-Without historical strike data, there is a higher risk of underpowered and inflated estimates of an attorney's strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction [@nscs-jury2018].
+Without historical strike data, there is a higher risk of underpowered and inflated estimates of an attorney's illegal strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction [@nscs-jury2023].
 Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data.
@@
 
-As a result, for any given estimate of strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power [@Bennett2021].
+As a result, for any given estimate of illegal strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power [@Bennett2021].
 Other methods to increase power may in turn require adjusting for incompatibility implicitly.
@@
 We proceed in three steps.
-First, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution.
+First, we specify a model of the peremptory-strike process in the court of interest that includes a strike-bias parameter to which we assign an initial prior distribution.
 Second, we use an attorney's strike data from past trials, generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter.
@@
 And _Batson_ law  requires the trial judge to weigh all such evidence to decide whether illegal strike bias likely exists (the _Batson_ violation).
-Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
-After testing for sensitivity of bias detection to different fixed $\alpha$ values for the power prior [@Carvalho2021, p.
+Still, _before_ accounting for other relevant evidence of strike bias, our approach can help attorneys detect illegal strike bias, given current and historical strike data, a valid model of the relevant court's strike procedure, and a fixed value for the power prior's $\alpha$ parameter.
+After testing for sensitivity of strike-bias detection to different fixed $\alpha$ values for the power prior [@Carvalho2021, p.
 5252; @Ibrahim2015, p.
@@
 Under such a procedure, the trial judge rules on all challenges for cause before the parties exercise any peremptory strikes.
-Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that corresponds to the number seats on the jury (plus alternates, if any).
+Then, of the potential jurors who remain, a subset of them are subject to peremptory strike, usually a number that accounts for the number seats on the jury (plus alternates, if any) and the number of strikes allotted to each party.
 The parties exercise their strikes on anyone among this subset of potential jurors in an alternating sequence.
@@
 
-By adding one parameter $w$, we can measure _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
+By adding one parameter $w$, we can measure strike _bias_ by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$.
 To avoid making the weight of the non-cognizable class be the reciprocal of the weight of cognizable class, let $b = log(w)$.
@@
 If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck.
-If $b>0$, we infer that the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority).
-Where $b<0$, the party has a preference for a juror within the cognizable class.
+If $b>0$, we infer that the party has strike bias against a prospective juror falling within the cognizable class (e.g., a racial minority).
+Where $b<0$, the party has strike bias against non-members of the cognizable class (e.g., a white prospective juror).
 
@@
 This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.
-Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal bias are less likely than lower degrees of bias.
+Moreover, we take the law to imply that, absent any data, one must assume that higher degrees of illegal strike bias are less likely than lower degrees of such bias.
 For this reason, we use a normal (Gaussian) distribution with $\sigma = 2$.
@@
 
-For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero.
+For our simulation, we define bias detection as the proportion of times we identify strike bias when strike bias is present in the current trial (i.e., $b_{curr} > 0$), based on whether credible intervals of the bias parameter exclude zero.
 Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus on how accurately our model can detect that bias, _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate).
@@
 For each scenario, we generated 1000 data sets and fit the model on those data sets.
-For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1000 iterations.
+For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1,000 iterations.
 For each replicate, we calculated the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.
@@
 Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution.
-As in the simulations, a normal distribution was assumed for the distribution of $b$.
+As in the simulations, we assumed a normal distribution for the distribution of $b$.
 The length of the Markov chain was 110,000 and the first 10,000 iterations were dropped as burn-in.
@@
 Now the credible intervals clearly exclude zero.
-Thus, we have a credible basis to infer bias against racial-minority prospective jurors in how this attorney uses their peremptory challenges in the present case.
+Thus, we have a credible basis to infer strike bias against racial-minority prospective jurors in how this attorney uses their peremptory challenges in the present case.
 In the Supplementary Material, we present additional illustrations wherein we use the app to detect gender bias and where we assign less weight to the historical strike data.
@@
 By incorporating the power prior, this approach makes assumptions about the degree of incompatibility between current and historical trial data an explicit part of the process of identifying likely strike bias.
-Moreover, with a tool like our prototype app, attorneys can use the approach presented here in real time during jury selection to help them decide whether to bring a  _Batson_ challenge at all.
+Moreover, with a tool like our prototype app, attorneys can use the approach presented here during jury selection to help them decide whether to bring a  _Batson_ challenge at all.
 If argued, the attorney can also introduce that posterior distribution as relevant evidence, along with any other admissible evidence, for presenting a prima facie case or for ultimately proving illegal strike bias (Figure \@ref(fig:batsonworkflowfig)).
@@
 Overall, given high compatibility, increasing $\alpha$ will increase bias detection.
-But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current bias ($b_{curr}$) is low; and bias detection can be (erroneously) low where current bias is high but bias in historical trials ($b_{hist}$) is low.
+
+But our simulations also show that, given high incompatibility, bias detection can be (erroneously) high when current strike bias ($b_{curr}$) is low and strike bias in past trials ($b_{hist}$) is high.
+For example, Figure \@ref(fig:figbd9515) shows that when $b_{curr}$ = 0, $b_{hist}$ = 3, and alpha = 1, our simulation bias detection rate is 1 when there are two or three historical trials.
+This might occur where an attorney is motivated to strike Black prospective jurors _only_ in criminal trials with Black defendants; all that attorney's past criminal trials had Black defendants; but the current trial has a White defendant.
+Similarly, strike-bias detection can be (erroneously) low where current strike bias ($b_{curr}$) is high but bias in historical trials ($b_{hist}$) is low.
+For either case of high incompatibility, the difficulty is that $b_{curr}$ and $b_{hist}$ are never directly observable in real cases.
 
 Where lawyers dispute the degree of incompatibility, they can and should interrogate the choice of $\alpha$ by selecting multiple $\alpha$ values and observing how sensitive the posterior distribution of the striking attorney's bias parameter is to the choice of $\alpha$ value.
-Our prototype app permits this kind of testing of sensitivity by providing multiple options for $\alpha$ values.
-If the inference of bias is sensitive to the choice of $\alpha$ values, all else equal, then one should worry more about incompatibility and choose a lower $\alpha$ value accordingly.
+Our prototype app permits this kind of testing of the degree of sensitivity by providing multiple options for $\alpha$ values.
+If the inference of bias is highly sensitive to the choice of $\alpha$ values, all else equal, then one should worry more about incompatibility and choose a lower $\alpha$ value accordingly.
+For example, Figure \@ref(fig:figbd9515) shows that when $b_{curr}$ = 0, $b_{hist}$ = 3, but we lower $\alpha$ to 0.1, the simulation bias detection rate drops to 0.07 and 0.13 for two and three historical trials, respectively.
 
@@
 If none, one could assign equal weight to the historical data ($\alpha$ = 1).
-If one allows for some incompatibility nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the same trial-characteristics as the bias-salient characteristics of current trial.
+If one allows for some incompatibility nonetheless, one could either subset the historical data (sacrificing power) to include only those past trials that share the bias-salient characteristics of current trial.
 Or one could keep all the historical data; choose an $\alpha$ value to match the degree of suspected incompatibility; and check for how sensitive any bias detection is to different $\alpha$ values.
-
-[EDITME]Our simulations demonstrate that when bias is high in either the historical _or_ current data, bias can be detected, especially when $\alpha$ is close to 1.
-For example, when $b_{curr}$ = 0, $b_{hist}$ = 3, and alpha = 1 our simulation bias detection rate is 1, when there are two or three historical trials (Fig X).
-In a scenario where we are using real data, and the true values of $\alpha$ are not known, these represent cases of extreme incompatibility, and should be detectible through sensitivity checking of the value of $\alpha$.
-For instance, when the value of $\alpha$ is set to 0.1, the bias detection rate goes down to 0.02 for both the two and three historical data set simulations [INSERT REF TO FIGURE].
-Such sensitivity to the choice of $\alpha$ value indicates incompatibility between current and historical data.
 
@@
 @Ibrahim2015, p.
-3724-3738] could appropriately apply in this context while also considering incompatibility (incommensurability) in selecting $\alpha$ [@Ollier2020].
+3724-3738] could appropriately apply in this context while also considering incompatibility in selecting $\alpha$ [@Ollier2020].
 Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.
@@
 Second, the choice of model can affect prospective data collection not only by courts but also by courtroom observers.
-The public has a presumptive First Amendment right to attend jury selection in criminal cases.
-Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from jury selection [@press1984; @presley2010].
-As a result, any person can often attend jury selection and collect relevant strike data based on what they observe in real time.
+The public has a presumptive First Amendment right to attend jury selection in criminal cases [@press1984].
+Similarly, a criminal defendant has a Sixth Amendment right to a public trial that presumptively precludes excluding the public from the courtroom during jury selection [@presley2010].
+As a result, any person can often attend jury selection and collect relevant strike data based on what they observe.
 Thus, such courtroom observers can prospectively collect strike data that they would have been less likely to obtain from court records after the fact.
@@
 
-Figures were produced with ggplot2 [@ggplot2] or Tikz [@Tantau2022].
+Figures were produced with ggplot2 [@ggplot2] and Tikz [@Tantau2022].
 

### 2023-06-27 analysis/memo2_TAS_reviewers.Rmd
minor edits to responses to Reviewer 2

@@
 Thank you for suggesting this contrast between _Batson_ law and these State laws.
-We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." (@flowers2019, p.
-2241) We did not use the phrase "significant factor," because it does not appear in _Miller-El v.
+We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." _Flowers v.
+Mississippi_, 139 S.
+Ct.
+2228, 2241 (2019).
+We did not use the phrase "significant factor," because it does not appear in _Miller-El v.
 Dretke_, 545 U.S.
@@
 Second, we took this comment to say that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal strike bias (the credible interval includes zero), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal strike bias (the credible interval excludes zero).
+
 We have revised the Discussion to further address high incompatibility.
 As we observe there, because no one can directly observe $b_{curr}$ and $b_{hist}$, they cannot know for sure whether their case involves high incompatibility.
-Nonetheless, if the posterior estimate of $b$ is highly sensitive to the choice of $alpha$ -- something one can observe -- then we agree that a judge might properly worry about erroneously detecting strike bias due to high incompatibility.
-As for what a judge might do in this situation, a judge might not find a _Batson_ prima facie case for that reason.
-Or the judge may still find enough for a prima face case, but down-weight the historical data (by selecting an $\alpha$ value less than 1) when ultimately deciding whether illegal strike bias likely occurred.
+Nonetheless, if the posterior estimate of $b$ is highly sensitive to the choice of $\alpha$ -- something one can observe -- then we agree that a judge might properly worry about erroneously detecting strike bias due to high incompatibility.
+
+A judge might respond to this situation, in two ways.
+First, a judge might not find a _Batson_ prima facie case for that reason.
+Second, the judge may still find enough for a prima face case, but down-weight the historical data (by selecting an $\alpha$ value less than 1) when ultimately deciding whether illegal strike bias likely occurred.
+
 In the reverse situation of high incompatibility, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but (erroneously) does not detect illegal bias when $\alpha > 0$.
 Here, we agree that a judge should worry about erroneously failing to detect strike bias due to high incompatibility.
-Down-weighting the historical data here increases the support for a finding that illegal strike bias likely occurred.
+Down-weighting the historical data here, however, increases the support for a finding that illegal strike bias likely occurred, all else equal.

### 2023-06-27 analysis/paper_draft4.Rmd
updated citation to Pawel et al. (2023)
modifications to paper and memo

@@
 
-As a result, for any given estimate of illegal strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated type I error and reduced power [@Bennett2021].
+As a result, for any given estimate of illegal strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated bias detection when no bias exists and reduced power [@Bennett2021].
 Other methods to increase power may in turn require adjusting for incompatibility implicitly.
@@
 Increasing the number of historical trials leads to an increase in bias detection especially when current and historical trials are compatible.
-However, when current and historical trials are incompatible, both type I and type II error can be high if $\alpha$ is close to one.
+However, when current and historical trials are incompatible, both false positives (bias detection when $b_{curr} = 0$) and false negatives (no bias detection when when $b_{curr} > 0$) can be observed if $\alpha$ is close to one.
 As we discuss in section \@ref(selectingalpha) below, the choice of $\alpha$ depends on assumptions about the degree of incompatibility of current and historical trials.
@@
 Knowing exactly how to quantify incompatibility and specify an appropriate prior for $\alpha$ is an area of active research [e.g.
-@Han2023; @Pawel2022].
+@Han2023; @Pawel2023].
 

### 2023-06-27 analysis/paper_draft4.Rmd
updated paper to use revised figures

@@
 
-load_path <- here::here("figures","SD_sensitivity.jpg")
+load_path <- here::here("figures","SD_sensitivity_rev.jpg")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("figures","compatibility_plot.jpg")
+load_path <- here::here("figures","compatibility_plot_rev.jpg")
 knitr::include_graphics(load_path)
@@
 
-load_path <- here::here("figures","simultaneous_sims_plot.jpg")
+load_path <- here::here("figures","simultaneous_sims_plot_rev.jpg")
 knitr::include_graphics(load_path)

### 2023-06-28 analysis/paper_draft4.Rmd
edit to sentence formerly referring to inflated Type I error and reduced power

@@
 
-As a result, for any given estimate of illegal strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to inflated bias detection when no bias exists and reduced power [@Bennett2021].
+As a result, for any given estimate of illegal strike bias in a current trial, incompatibility between strike data from that current trial and strike data from past trials can lead to errors in detecting illegal strike bias (false positives and false negatives) [@Bennett2021].
 Other methods to increase power may in turn require adjusting for incompatibility implicitly.
-For example, @Gastwirth2014 [p.
-295-297] stratify strike data by trial and apply the Cochran-Mantel-Haenszel test, albeit after first selecting only historical strike data from past trials that are "similar" in their bias-salient characteristics to the current trial [p.
+For example, @Gastwirth2014 stratify strike data by trial and apply the Cochran-Mantel-Haenszel test [p.
+295-297], albeit after first selecting only historical strike data from past trials that are "similar" in their bias-salient characteristics to the current trial [p.
 289, 297].

### 2023-06-28 analysis/memo2_TAS_reviewers.Rmd
modifications to paper and memo
additional minor edits to TAS memo; removing old draft

@@
 width: .3
-bibliography: references.bib
+#bibliography: references.bib
 ---
 
-This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and an anonymous reviewer.
+This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and an anonymous reviewer (Reviewer 2).
 
@@
 
-We have made the app available to the public as a R package that includes the underlying dataset of historical strike data.
-We hope this will help statistics instructors in the way the TAS Associate Editor has in mind.
-However, because of time constraints, we are simply unable to provide in the paper's Supplement a complete tutorial tailored for classroom instruction by teachers of Bayesian statistics.
-On an introduction to a Bayesian approach to estimating illegal strike bias (albeit without historical strike data), we recommend @Kadane2018a and @Kadane2018b.
+We are ready to make the app available to the public as a R package that includes a de-identified version of the dataset of real historical strike data.
+(The link to the Github repository for the app identifies one of the authors.
+Accordingly, to preserve the anonymity that TAS requires, we have not yet put that link in the paper itself.) We hope this will help statistics instructors in the way the TAS Associate Editor has in mind.
+However, because of time constraints, we are simply unable to provide in the paper's supplement a complete tutorial tailored for classroom instruction by teachers of Bayesian statistics.
+
+<!--
+On an introduction to a Bayesian approach to estimating illegal strike bias (albeit without historical strike data), we recommend @Kadane2018a and @Kadane2018b.-->
 
@@
 
-The paper now contains links to the app as well as the code we used to generate the simulation results.
-We intend section 3.2 of the paper (and related Supplementary Material) as instructions for using the app.
-Those links are identifying, which is why we omitted them from the last draft of the paper.
+We intended section 3.2 of the paper (and related Supplementary Material) as instructions for using the app.
+Only because they are identifying, we have thus far omitted hyperlinks to the Github repositories for the app (ready to be available as an R package) as well as the code we used to generate the simulation results.
+When the TAS editor indicates it is appropriate, we will gladly add those links to the paper manuscript.
 
@@
 We appreciate this suggestion, but we worry that making this point here in terms of Type I and Type II error might lead to net confusion.
-The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more Type I error (falsely inferring illegal strike bias), because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence.
-Type I errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not.
-And Type I errors are worse still as the judge assigns more relative weight to any inference from historical strike data as compared to other items of evidence.
+The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more false positives (false inferences of illegal strike bias, because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence.
+Such errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not.
+And such errors are worse still as the judge assigns more relative weight to the inference from historical strike data as compared to other items of evidence.
 
@@
 We stress "assumes", because the true values of $b_{hist}$ and $b_{curr}$ are not directly observable in actual cases.
-Accordingly, for any _subset_ of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$.
-The current version of the software prototype does not let users subset the historical data by some trial-level characteristic (e.g., defendant race).
+Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$.
+The current version of the software app does not let users subset the historical data by some trial-level characteristic (e.g., defendant race, courthouse location).
 However, in principle, it could be modified to add that feature.
@@
 
-We have revised the Discussion accordingly.  See also our reply to other reviewer below.
+We have revised the Discussion accordingly.  See also our reply to Reviewer 2 below.
 
@@
 
-```{r}
-#| label: hypodata
-#| echo: false
+```{r, echo = FALSE}
 

### 2023-07-10 analysis/memo2_TAS_reviewers.Rmd
minor edit to memo

@@
 We appreciate this suggestion, but we worry that making this point here in terms of Type I and Type II error might lead to net confusion.
-The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more false positives (false inferences of illegal strike bias, because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence.
+The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more false positives (false inferences of illegal strike bias), because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence.
 Such errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not.

### 2023-07-28 app/app.R
Delete app directory

@@ -1,238 +1,0 @@
-library(shiny)
-library(rhandsontable)
-library(dplyr)
-library(mcmc)
-library(Rcpp)
-library(ggplot2)
-library(shinythemes)
-
-Rcpp::sourceCpp("mh_sampler_pp.cpp")
-Rcpp::sourceCpp("mh_sampler.cpp")
-
-# Load dummy strike dataframe
-source("dummy_strike_data.R")
-
-# Load calc_prior function
-
-source("calc_prior.R")
-
-# Define Cognitive Class choices
-
-cog_c_levels <- c("race","gender")
- 
-# Define UI for application
-    
-ui <- fluidPage(
-
-    # Navbar structure for UI
-    
-    navbarPage("Batson App - Proof of Concept", theme = shinytheme("paper"),
-               
-        tabPanel("Prototype", fluid = TRUE,
-            sidebarPanel(
-
-                "Prior Strike History by Attorney",
-                 selectInput("atty_d", "Defense:", 
-                               choices=atty_levels_d,
-                               selected = "None"),
-                
-                 selectInput("atty_p", "Prosecution:", 
-                               choices=atty_levels_p,
-                               selected = "None"),
-                 
-                selectInput("cog_c", "Cognizable Class:", choices=cog_c_levels),
-                
-                radioButtons("weight", "Prior Strike History Weight:", inline=TRUE, selected = 1,
-                                choiceValues = c(1, 0.5, 0.2),
-                                choiceNames = c("Equal","Half","Minimal") 
-                                    ),
-                hr(),
-                
-                "Current Strike Tally",
-                rHandsontableOutput("hot"),
-                
-                ),
-            mainPanel(
-                #actionButton("updateButton", "Update",
-                #             icon = icon("refresh")),
-                plotOutput("plot"))
-                )
-        ))
-
-    
-# Define server logic 
-server <- function(input, output, session) {
-
-        output$hot <- renderRHandsontable({
-            
-            rhandsontable(df0, width = 600, height = 700) %>%
-                hot_col("round", format = "0") %>%
-                hot_col("num_cog", format = "0") %>%
-                hot_col("total", format = "0") %>%
-                hot_col("cog", format = "0") %>%
-                hot_col("party", type = "dropdown", source = party_choices) %>%
-                hot_validate_numeric(cols = 4, min = 0) %>%
-                hot_context_menu(allowRowEdit = TRUE)
-        })
-        
-        # Calculate and Plot the Posterior Distribution
[…]

### 2023-07-28 app/calc_prior.R

@@ -1,94 +1,0 @@
-### assume dat0 is all the historical data we have
-
-library(dplyr)
-library(mcmc)
-library(Rcpp)
-Rcpp::sourceCpp("mh_sampler_pp.cpp")
-
-dat0 <- readRDS("jury_data_cleaned_new.rds")
-
-
-#### extract_atny() function extracts the subset of jury_data_cleaned that corresponds to 
-#### the attorney name inputted in 'atny_name' parameter. 
-#### if attorney name is from PP, set 'pp' parameter as TRUE
-#### if attorney name is from PD, set 'pp' parameter as FALSE
-#### parameter 'pp' is the same for the following functions, used to control whether the case is from PP or PD
-
-extract_atny <- function(atny_name,pp,dat){
-  if(pp){
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-  else{
-    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
-    return(dat[which(idx),])
-  }
-} 
-
-#### function organize_input will return a matrix with four columns:
-#### round, num of cog_class, total number, and whether the strike is from cognizable class
-#### the return matrix is of the form of the input of make_posterior function
-#### input of cog should be either 'gender' or 'race'
-
-organize_input <- function(dat,pp,cog){
-  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
-  if(cog=='gender'){
-    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
-  }else{
-    dat_strikes <- dat_strikes %>% filter(!is.na(race))
-  }
-  
-  dat_strikes <- dat_strikes[order(dat_strikes$strike_seq),]
-  if(cog=='gender'){
-    dat_strikes$cogb <- ifelse(dat_strikes$sex=="F", 1, 0)
-  }else{
-    dat_strikes$cogb <- ifelse(dat_strikes$race !=3, 1, 0)
-  }
-  
-  num_cog <- sum(dat_strikes$cogb)
-  num_t <- nrow(dat_strikes)
-  subs <- data.frame(round = c(1:num_t), 
-                     num_cog = NA, 
-                     total = NA)
-  subs$num_cog[1] <- num_cog
-  subs$total[1] <- num_t
[…]

### 2023-07-28 app/dpp.R

@@ -1,13 +1,0 @@
-# dpp.R declares all the packages needed to run the Shiny app. They will be
-# installed during the building process of the Docker image for the Shiny app
-
-# list all the packages as a string list, e.g., c("shiny", "ggplot2", ...)
-packages <- c("shiny","rhandsontable","dplyr","mcmc","Rcpp","ggplot2","shinythemes")
-
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(
-    setdiff(packages, rownames(installed.packages())),
-    dependencies = T,
-    repos='https://cloud.r-project.org/'
-  )
-}

### 2023-07-28 app/dummy_strike_data.R

@@ -1,32 +1,0 @@
-#### dummy strike data ####
-
-df0 <- data.frame(round = c(1:10), 
-                  num_cog = c(3, 4, 4, 3, 2, 2, 2, 2, 2,2),
-                  total = rep(9, 10), 
-                  cog = c(0, 1, 1, 0, 0, 1, 1, 0, 0,0),
-                  party = rep(c("PP","PD"),5)
-)
-
-party_choices <- c("PP","PD")
-
-
-#### attorney name choices ####
-
-dat0 <- readRDS("jury_data_cleaned_new.rds")
-
-atty_levels_p <- dat0 %>%
-  pull("P_atty_l") %>%
-  unlist() %>%
-  unique() %>%
-  stringr::str_trim() %>%
-  factor()
-
-atty_levels_d <- dat0 %>%
-  pull("D_atty_l") %>%
-  unlist() %>%
-  unique() %>%
-  stringr::str_trim() %>%
-  factor()
-
-atty_levels_p <- c("None",levels(atty_levels_p))
-atty_levels_d <- c("None",levels(atty_levels_d))

### 2023-07-28 app/extract_strike_seq.R

@@ -1,27 +1,0 @@
-# Create Strike Sequence Variable (from strike_num) for each case (ID)
-
-## Premise: PP starts first (one strike), then PD (two strikes), then PP ...
-## until all strikes used or foregone (skip)
-
-library(dplyr)
-
-load_path <- here::here("prototype","jury_data_cleaned.RDS")
-save_path <- here::here("prototype","jury_data_cleaned_new.rds")
-
-dat0 <- readRDS(load_path)
-
-extract_seq <- function(dat){
-  dat$strike_num <- as.numeric(dat$strike_num)
-  dat = dat[order(dat$strike_num),]
-  min0 <- min(dat$strike_num[which(dat$strike_num>0)])
-  legth <- length(dat$strike_num[which(dat$strike_num>0)])
-  dat$strike_seq[which(dat$strike_num>0)] <- seq(from=min0,to = min0+legth-1,by = 1)
-  return(dat)
-}
-
-dat1 <- dat0 %>% group_split(ID)
-test_l <- lapply(1:length(dat1), function(x) extract_seq(dat1[[x]]))
-dat1 = do.call(rbind.data.frame,test_l)
-
-saveRDS(dat1, file = save_path)
-

### 2023-07-28 app/mh_sampler_pp.cpp

@@ -1,137 +1,0 @@
-#include <Rcpp.h>
-#include <iostream>
-using namespace Rcpp;
-
-// This is a simple example of exporting a C++ function to R. You can
-// source this function into an R session using the Rcpp::sourceCpp 
-// function (or via the Source button on the editor toolbar). Learn
-// more about Rcpp at:
-//
-//   http://www.rcpp.org/
-//   http://adv-r.had.co.nz/Rcpp.html
-//   http://gallery.rcpp.org/
-//
-
-// [[Rcpp::export]]
-double probcpp(bool cog, int c, int m, double d){
-  if(cog){
-    return log(c)+d-log(c*exp(d)+m);
-  }else{
-    return log(m)-log(c*exp(d)+m);
-  }
-}
-
-
-// x, col 2 should be number cognizable, 3 be total, 4 be cognizable
-// [[Rcpp::export]]
-double Lk_cpp_p(double d,NumericMatrix x, bool L){
-  double lglkr=0;
-  for(int i=0; i < x.nrow();++i){
-    lglkr = lglkr + probcpp(x(i,3),x(i,1),x(i,2)-x(i,1),d);
-  }
-  if(L){
-    return lglkr;
-  }else{
-    return exp(lglkr);
-  }
-}
-
-// [[Rcpp::export]]
-double Priorcpp_p(double theta, double mu, double std,NumericMatrix x,double a0) {
-  return pow(Lk_cpp_p(theta,x,0),a0)*1/(sqrt(2*M_PI)*std)*exp(-pow((theta-mu),2)/(2*pow(std,2)));
-}
-
-// [[Rcpp::export]]
-List make_posterior_p(NumericMatrix x,NumericMatrix x_p,double a0, int niter, 
-                             double theta_start_val, double theta_proposal_sd, double prior_mean, double prior_sd){
-  NumericVector theta(niter);
-  double current_theta;
-  double new_theta;
-  double r;
-  double thresh;
-  double acpt;
-  int acptn=0;
- 
-  theta[0] = theta_start_val;
-  
-  for(int i = 1; i < niter; ++i){
-    
-    current_theta = theta[i-1];
-    
-    new_theta = rnorm(1,current_theta, theta_proposal_sd)[0];
-    
-    
-    r = log(Priorcpp_p(new_theta,prior_mean,prior_sd,x_p,a0))+Lk_cpp_p(new_theta,x, 1)-log(Priorcpp_p(current_theta,prior_mean,prior_sd,x_p,a0))-
[…]

### 2023-07-28 app/prior_sensitivity.R

@@ -1,63 +1,0 @@
-setwd('/Users/xiaomengli/Documents/consulting/2020spring/bayesian_project/Test0/prototype/')
-library(Rcpp)
-source("calc_prior.R")
-sourceCpp("mh_sampler.cpp")
-sourceCpp("mh_sampler_pp.cpp")
-library(ggplot2)
-
-#### simulate input table ####
-df0 <- data.frame(round = c(1:9), 
-                  num_cog = NA, 
-                  total = NA,cog=NA)
-df0[,'num_cog']=c(3,4,4,3,2,2,2,2,2)
-df0[,'total']=rep(9,9)
-df0[,'cog']=c(0,1,1,1,1,0,0,0,1)
-df0_m <- as.matrix(df0)
-
-### normal prior with calculated posterior mean and sd ###
-
-prior1 = calc_prior('Brian P. Leaming',TRUE,dat0,'race')
-post1 <- make_posterior(x = df0_m, niter = 11000, 
-                        theta_start_val = 0, theta_proposal_sd =.5, 
-                        prior_mean = prior1$prior_mean, prior_sd = prior1$prior_sd)
-
-### use default vague prior ###
-post4 <- make_posterior(x = df0_m, niter = 11000, 
-                        theta_start_val = 0, theta_proposal_sd =.5, 
-                        prior_mean = 0, prior_sd = 2)
-
-
-### power prior ####
-subset <- function(atny_name,pp,dat,cog){
-  dat_sub <- extract_atny(atny_name,pp,dat)
-  sub1 <- dat_sub %>% group_split(ID)
-  sub1_l <- lapply(1:length(sub1), function(x) organize_input(sub1[[x]],pp,cog))
-  df_m = do.call(rbind.data.frame,sub1_l)
-  df_m <- as.matrix(df_m)
-}
-
-### with a0=1 ###
-sub1 <- subset('Brian P. Leaming',TRUE,dat0,'race')
-post2 <- make_posterior_p(x = df0_m,x_p=sub1,a0=1, niter = 11000, 
-                        theta_start_val = 0, theta_proposal_sd =.5, 
-                        prior_mean = 0, prior_sd = 2)
-
-### with a0=0 which should be the same with the vague normal prior ###
-post3 <- make_posterior_p(x = df0_m,x_p=sub1,a0=0, niter = 11000, 
-                          theta_start_val = 0, theta_proposal_sd =.5, 
-                          prior_mean = 0, prior_sd = 2)
-
[…]

### 2023-07-28 R/dat0.R
Delete R directory

@@ -1,26 +1,0 @@
-#' Jury Selection in Connecticut federal court
-#'
-#' A dataset containing the jury selection information for cases completed in the federal district court
-#'  of Connecticut for 2013-2018
-#'
-#' @format A data frame with 979 rows and 30 variables:
-#' \describe{
-#'   \item{Docket.x}{case docket number}
-#'   \item{Case_Name}{case name on caption}
-#'   \item{courthouse}{courthouse location}
-#'   \item{Defendant.x}{first defendant name on case caption}
-#'   \item{Judge}{ID (initials) for district court judge}
-#'   \item{date}{date jury selection began}
-#'   \item{R_No}{venireperson sequence order}
-#'   \item{Pool_Seq}{jury pool ID number}
-#'   \item{race}{numerical code for race from juror questionaire}
-#'   \item{hisp}{1 = "Hispanic", 0 = not}
-#'   \item{sex}{M = "male", F = "female"}
-#'   \item{Disp}{voir dire disposition. "PP" - prosecutor strike; "PD" - defense strike; "J" - seated juror}
-#'   \item{strike_num}{order of strikes - each side}
-#'   \item{P_atty}{prosecutors on docket sheet at time of jury selection}
-#'   \item{D_atty}{defense attorneys on docket sheet at time of jury selection}
-#'   ...
-#' }
-#' @source Jury Administrator, US District Court - Connecticut
-"dat0"

### 2023-07-28 R/render_paper.R

@@ -1,5 +1,0 @@
-# render paper to separate output folder
-
-output_path <- here::here("analysis","_output")
-input_path <- here::here("analysis","paper_draft3.Rmd")
-rmarkdown::render(input = input_path, output_dir = output_path)

### 2023-07-28 tests/testthat.R
Delete tests directory

@@ -1,4 +1,0 @@
-library(testthat)
-library(batson_app)
-
-test_check("batson_app")

### 2023-07-28 tests/testthat/test-sequential.R

@@ -1,7 +1,0 @@
-test_that("multiplication works", {
-  expect_equal(2 * 2, 4)
-})
-
-test_that('no missing values', {
-  expect_identical(a, na.omit(a))
-})

### 2023-07-28 tests/testthat/test-sim_seq.R

@@ -1,6 +1,0 @@
-load_path <- here::here("sim_seq.R")
-source(load_path)
-
-test_that("multiplication works", {
-  expect_equal(2 * 2, 4)
-})

### 2023-07-28 R-code/conditional_statement.R
Delete conditional_statement.R

@@ -1,31 +1,0 @@
-cond.fn <- function(CI){
-  CI.PP <- filter(CI, party=="PP")
-    
-  if(CI.PP$q1 < 0 & CI.PP$q2 < 0){
-    print("Preference in PP") 
-    
-  }else if(CI.PP$q1 < 0 & CI.PP$q2 > 0) { 
-    print("No bias or preference in PP")
-    
-  } else if (CI.PP$q1 > 0 & CI.PP$q2 > 0){
-    print("Bias in PP")
-    
-  }
-  
-  CI.PD <- filter(CI, party=="PD")
-  
-  if(CI.PP$q1 < 0 & CI.PP$q2 < 0){
-    print("Preference in PD") 
-    
-  }else if(CI.PP$q1 < 0 & CI.PP$q2 > 0) { 
-    print("No bias or preference in PD")
-    
-  } else if (CI.PP$q1 > 0 & CI.PP$q2 > 0){
-    print("Bias in PD")
-    
-  }
-  
-      
-}
-
-cond.fn(CI)

### 2023-07-28 R-code/simultaneous_plots.R
Update simultaneous_plots.R

@@ -1,6 +1,6 @@
 
 # Code for generating plots of simultaneous strike data
-# output from the 'simultaneous.R' script. 
+# output from the 'simultaneous_sims.R' script. 
 #-----------------------------------------------------------------#
 
 library(dplyr)
